### 2009-06-29 papers/SenseTileOverview/paper.tex
Initial proto-outline of journal paper and added directory hierarchy where papers are to live.

%
% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass[times,10pt,twocolumn]{article} 
\usepackage{latex8}
\usepackage{times}

\usepackage{ifpdf}
\usepackage{a4wide}
\usepackage{pdfsync}

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

%\usepackage{todonotes}

% \usepackage{html}
% \usepackage{url}
\usepackage{xspace}
%\usepackage{doublespace}
\usepackage{tabularx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage{stmaryrd}
% \ifpdf
% \usepackage[centredisplay]{diagrams}
% \else
% \usepackage[centredisplay,PostScript=dvips]{diagrams}
% \fi
\usepackage{float}

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

%% \input{kt}

%=====================================================================

\begin{document}

\title{The UCD CASL SenseTile System}

\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
}

\maketitle

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}

An abstract.

\end{abstract}

%======================================================================
\Section{Introduction}

An introduction.

%=====================================================================
\Section{Conclusion}

A conclusion.

%======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{latex8}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% eval: 
%%% TeX-master: t
%%% End:

### 2009-06-29 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
an abstract of the abstract

--- 2009-06-29 papers/SenseTileOverview/paper.tex
@@
 
+OBJECTIVE: 
+Specifications used as a blueprint for system development.
+Specifications to generate test suite to validate the developed system.
+Specifications to validate simulator of the developed system.
+Apply to data stream processing: verify the stream is parsed correctly, 
+verify the parsed stream output stream can be simulated.
+
+METHODS: 
+JML model.
+JML tools to generate JUnit test suite.
+JUnit test suite applied to developed system, to verify its correctness.
+Junit test suite applied to simulator, to verify its correctness.
+
+RESULTS: 
+Proof developed system same behaviour as simulator.
+Proof developed system and simulator verify the specifications.
+
+CONCLUSIONS:
+Specifications serves two purposes, and indirectly a third one.
+
 \todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
 \myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
-
-An abstract.
 

### 2009-07-15 papers/SenseTileOverview/paper.tex
Wrote an abstract and reused the introduction from the grant proposal as the introduction of this paper.

@@
 \newcommand{\todo}{\textbf{TODO:}}
+\newcommand{\ST}{\emph{SenseTile}\xspace}
+\newcommand{\STs}{\emph{SenseTiles}\xspace}
+\newcommand{\datastore}{\STs Scientific Datastore\xspace}
+\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
+\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
+\newcommand{\sensorfarm}{The \STs\xspace}
 
@@
 \todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
-\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
+\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE
+  Pervasive Computing}}.}
 
-An abstract.
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
+sensor system installed at the University College Dublin in Dublin,
+Ireland.  \textbf{Our facility provides a capability---unique in the world in
+terms of its scale and flexibility---for performing in-depth
+investigation of both the specific and general issues in large-scale
+sensor networking.}
+
+This system integrates a sensor platform, a datastore, and a compute
+farm.  The sensor platform is a custom-designed, but inexpensive,
+sensor platform (called the \ST) paired with general-purpose
+small-scale compute nodes, including everything from low-power PDAs to
+powerful touchscreen-based portable computers.  The datastore is a
+multi-terabyte scale scientific datastore into which sensor data
+flows, and in which online and offline scientific computation of
+sensor and other scientific data takes place.  The compute farm is a
+large-scale Linux, Solaris, and OS~X-based compute farm used by
+scientists for these online and offline scientific computations.
+
+This system has several novel features: (1) new deployments of \STs
+requires zero configuration; (2) \STs have several unusual built-in
+sensors and emitters, (3) sensor nodes can be extended via a
+general-purpose USB bus, (4) \STs contain an FPGA that is dynamically
+updatable on a per-experiment basis, and (5) configuring, deploying,
+and managing experiments using \STs is performed by scientists via
+either programmatic APIs or via a declarative graphical or textual
+specification language.
 
@@
 
-An introduction.
+Sensor networks are widely seen as an essential technology for
+supporting next-generation scientific and engineering challenges,
+including environmental monitoring, climate change, assisted living,
+national security and intensive agriculture. Addressing these problems
+requires two complementary strands of research:
+\begin{enumerate}
+\item to understand the \emph{specific physical and/or social
+    phenomena of interest} against which to fit data collection and
+  analysis; and
+\item to understand the \emph{general scientific and technological
+    principles governing scalable sensor networking} in order to
+  maximise the leverage gained from on-going developments.
+\end{enumerate}
+
+Amongst the core general challenges are the collection of large
+volumes of multimedia data, its storage, cataloging, retrieval and
+processing in such a way as to minimise physical and intellectual
+costs of accessing the available data. This is especially important in
+pursuit of cross-disciplinary projects involving different analysis
+methodologies and constraints. At the same time, it is impossible to
+study sensor networking completely separate from practical
+applications which provide real-world validation and verification of
+the techniques being developed.
+
+This paper describes a facility, the \ST system, to support
+large-scale experiments with complex multimedia sensing and processing
+at terabyte scales.  The key piece of equipment we have designed and
+fabricated is a rich and reusable sensor platform, called the \ST,
+that is easily deployed into the built environment for
+experimentation.  This particular piece of hardware is representative
+of platforms suitable for wider uses, including on-board data storage
+and processing capacity and associated high-speed interconnect.
+
+The key research challenges this system supports includes:
+\begin{itemize}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+\item dealing with uncertainty in sensor systems,
+\item robust management of large (multi-terabyte) sensing data on
+  heterogeneous platforms,
+\item interaction and activity recognition,
+\item sparse signal analysis, and
+\item coding and securing sensor interactions.
+\end{itemize}
+
+Underpinning each of these key challenges is a fundamental need for an
+integrated, multi-modal sensor testbed that has two unique
+characteristics. Firstly, it must be \emph{reconfigurable and be
+  easily targeted} to collect defined data sets.  Secondly---and
+perhaps most importantly---it must be \emph{of sufficient size and
+  complexity} to ensure that the challenges listed above, and the
+resulting theories and techniques, can be evaluated in a
+scientifically rigorous manner.  It is our contention that \emph{our
+  facility provides a capability---unique in the world in terms of its
+  scale and flexibility---for performing in-depth investigation of
+  both the specific and general issues in large-scale sensor
+  networking.}  Furthermore, a synergistic benefit is the support for
+the development of reference data sets and services for use by other
+researchers internationally.
 

### 2009-07-17 papers/BON_VHDL/paper.tex
Wrote outline for two more papers, one on VHDL<->BON refinement, and the other on SensorML<->JML refinement via BON.

--- 2009-06-29 papers/SenseTileOverview/paper.tex
@@
 %
-% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
-% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
+% "Refining from the BON High-level Specification Language to the
+% VHDL Hardware Description Language" for UNKNOWN09.
+% Joseph R. Kiniry, Dragan Stosic
 % $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
@@
 
-\documentclass[times,10pt,twocolumn]{article} 
-\usepackage{latex8}
+\documentclass{article} 
 \usepackage{times}
@@
             linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
+            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
 \else
@@
 \newcommand{\todo}{\textbf{TODO:}}
+\newcommand{\ST}{\emph{SenseTile}\xspace}
+\newcommand{\STs}{\emph{SenseTiles}\xspace}
+\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
+\newcommand{\datastore}{\STs Scientific Datastore\xspace}
+\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
+\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
+\newcommand{\sensorfarm}{The \STs\xspace}
 
@@
 
-\title{The UCD CASL SenseTile System}
+\title{Refining from the BON High-level Specification Language to the
+  VHDL Hardware Description Language}
 
-\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
+\author{Joseph R. Kiniry and Dragan Stosic\\
 UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
 Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
+kiniry@acm.org and dragan.stosic@gmail.com\\
 }
@@
 
-\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
-\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
+sensor system installed at the University College Dublin in Dublin,
+Ireland.  \textbf{Our facility provides a capability---unique in the world in
+terms of its scale and flexibility---for performing in-depth
+investigation of both the specific and general issues in large-scale
+sensor networking.}
 
-An abstract.
+This system integrates a sensor platform, a datastore, and a compute
+farm.  The sensor platform is a custom-designed, but inexpensive,
+sensor platform (called the \ST) paired with general-purpose
+small-scale compute nodes, including everything from low-power PDAs to
+powerful touchscreen-based portable computers.  The datastore is a
+multi-terabyte scale scientific datastore into which sensor data
+flows, and in which online and offline scientific computation of
+sensor and other scientific data takes place.  The compute farm is a
+large-scale Linux, Solaris, and OS~X-based compute farm used by
+scientists for these online and offline scientific computations.
+
+The \STs have several unusual built-in sensors and emitters, sensor
+nodes can be extended via a general-purpose USB bus, and \STs contain
+an FPGA that is dynamically updated on a per-experiment basis.  \STs
+are custom-designed in VHDL.  Like many hardware-based systems, to
+write new software for or against the \ST architecture one either uses
+the actual hardware or one runs a \STS.  But this simulator emulates a
+piece of hardware, one would like to know if the simulator
+\emph{actually} simulates the hardware.  We have developed an approach
+through which we refine the VHDL specification up to the formal BON
+specification language, and from this specification we generate an
+executable, functional model-based specification in the JML language.
+We have used this specification to implement a fully validated and
+verified \ST simulator.
 
@@
 %======================================================================
-\Section{Introduction}
+\section{Introduction}
 
@@
 %=====================================================================
-\Section{Conclusion}
+\section{Conclusion}
 
@@
 %% \nocite{ex1,ex2}
-\bibliographystyle{latex8}
+\bibliographystyle{plain}
 \bibliography{extra,%

### 2009-07-17 papers/GeneratingSensorModels/paper.tex
Wrote outline for two more papers, one on VHDL<->BON refinement, and the other on SensorML<->JML refinement via BON.

--- 2009-06-29 papers/SenseTileOverview/paper.tex
@@
 %
-% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
+% "Generating Formal Executable Sensor Models from their Descriptions"
+% for UNKNOWN09
 % Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
@@
 
-\title{The UCD CASL SenseTile System}
+\title{Generating Formal Executable Sensor Models\\ from their Descriptions}
 
@@
 
-\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
-\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
-
-An abstract.
+  In this work, rather than writing sensor simulators by hand, we
+  automatically refine a SensorML specification through a series of
+  stages to a formally specified, executable, functional
+  model-based specification written in the Java Modeling Language.
+  This resulting simulator implementation is completely validated with
+  (for the most part, automatically generated) unit tests and is
+  verified with the Extended Static Checker for Java, ESC/Java2.
 
@@
 %======================================================================
-\Section{Introduction}
+\section{Introduction}
 
-An introduction.
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
+sensor system installed at the University College Dublin in Dublin,
+Ireland.  \textbf{Our facility provides a capability---unique in the world in
+terms of its scale and flexibility---for performing in-depth
+investigation of both the specific and general issues in large-scale
+sensor networking.}
+
+This system integrates a sensor platform, a datastore, and a compute
+farm.  The sensor platform is a custom-designed, but inexpensive,
+sensor platform (called the \ST) paired with general-purpose
+small-scale compute nodes, including everything from low-power PDAs to
+powerful touchscreen-based portable computers.  There are over a dozen
+sensors on the \ST itself, and new sensors can be added by adding them
+to the USB bus.
+
+A software-based \STS has been designed and implemented that is meant
+to mimic the \ST hardware.  A part of this simulation is, of course,
+simulating the sensors themselves.  The challenge in such simulation
+is to ensure that a sensor simulator properly mimics the behavior of
+the actual hardware sensor.  While one can specify the behavior of
+such a sensor from a hardware data sheet and then hand-implement the
+sensor simulator, this process is problematic on several fronts.  Not
+only is one concerned about the accuracy of the simulation, but
+writing and maintaining these simulations is labor-intensive.
+
+A popular means by which to describe sensor data is to the use the
+SensorML specification language.  Using SensorML one can describe
+sensors and measurement processes, including an electronic
+specification sheet for a sensor, the lineage of an observation, etc.
+We would hope that this declarative specification is directly related
+to its simulation---in other words, the simulation must be a
+refinement of the SensorML specification.  But at the moment the is no
+way one can define or perform this refinement other than programming
+against a SensorML API by hand.
+
+To overcome this challenge, we: (1) formally define a sensor
+meta-model in the BON specification language; (2) define a formal
+refinement relation between the (XML) meta-model of SensorML and this
+BON meta-model; (3) using this refinement, we automatically generate
+an informal and formal BON model from any SensorML model; (4) using
+the Beetlz tool, from the formal BON model, we generate a formally
+specified, executable, functional model-based specification written in
+the Java Modeling Language, and; (5) we validate and verify, using
+unit testing and extended static checking respectively, this JML
+sensor model.
+
+Consequently, using formal BON, we can define a \emph{sensor system}
+as a composition of sensor models, and thus derive a complete formally
+specified and verified, executable sensor system model.  To
+demonstrate this process we describe an interesting slice of the \ST
+specification and its simulator.
 
 %=====================================================================
-\Section{Conclusion}
+\section{Conclusion}
 

### 2009-07-17 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
Wrote an abstract for this paper.

@@
 
-\title{The UCD CASL SenseTile System}
+\title{Deriving an Executable Model of a Data Stream Protocol from its
+Specification}
 
@@
 
+\begin{verbatim}
 OBJECTIVE: 
@@
 Specifications serves two purposes, and indirectly a third one.
+\end{verbatim}
 
-\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
-\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
+  Sensor packages typically communicate (control sensors and report
+  sensor measurements) with a streaming packet-based network protocol.
+  The \ST device is an example of such a sensor package, as it
+  communicates with an arbitrary computer system (running Linux or
+  Windows, at the moment) over a USB bus using a streaming
+  packet-based protocol.  This sensor platform is a custom-designed,
+  but inexpensive, sensor platform (called the \ST) paired with
+  general-purpose small-scale compute nodes, including everything from
+  low-power PDAs to powerful touchscreen-based portable computers.
+  There are over a dozen sensors on the \ST itself, and new sensors
+  can be added by adding them to the USB bus.  A third-party low-level
+  C driver, provided by the USB chip vendor, creates and parses
+  generic USB communication packets, and a custom medium-level Java
+  driver creates and parses \ST streaming communication packets.  This
+  latter Java driver is formally specified, validated with rigorous
+  unit testing, and verified using extended static checking.  However,
+  because of this process and combination of technologies, using this
+  specification we are able to both communicate with a real \ST
+  \textbf{and} generate a formally specified, executable, functional
+  model-based specification for the protocol itself written in the
+  Java Modeling Language.  This protocol model is itself an executable
+  stream simulator that can be used to develop new \ST user
+  applications without having the \ST hardware at all.  Consequently,
+  from this single specification, we can test that the \ST hardware
+  communicates as expected, test that our \ST simulator communicates
+  as expected, and simulate the \ST protocol stream itself.
 
@@
 
-An introduction.
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
+sensor system installed at the University College Dublin in Dublin,
+Ireland.  \textbf{Our facility provides a capability---unique in the world in
+terms of its scale and flexibility---for performing in-depth
+investigation of both the specific and general issues in large-scale
+sensor networking.}
+
+This system integrates a sensor platform, a datastore, and a compute
+farm.  The sensor platform is a custom-designed, but inexpensive,
+sensor platform (called the \ST) paired with general-purpose
+small-scale compute nodes, including everything from low-power PDAs to
+powerful touchscreen-based portable computers.  There are over a dozen
+sensors on the \ST itself, and new sensors can be added by adding them
+to the USB bus.
 

### 2009-07-25 papers/BON_VHDL/paper.tex
start paper work

@@
 
-The \STs have several unusual built-in sensors and emitters, sensor
-nodes can be extended via a general-purpose USB bus, and \STs contain
-an FPGA that is dynamically updated on a per-experiment basis.  \STs
-are custom-designed in VHDL.  Like many hardware-based systems, to
-write new software for or against the \ST architecture one either uses
+The \STs, as an embedded system, have several unusual built-in sensors 
+and emitters, sensor nodes can be extended via a general-purpose USB bus, 
+and \STs contain an FPGA that is dynamically updated on a per-experiment 
+basis.  \STs are custom-designed in VHDL.  Like many hardware-based systems,
+ to write new software for or against the \ST architecture one either uses
 the actual hardware or one runs a \STS.  But this simulator emulates a
@@
 \section{Introduction}
+The Embedded system is composed of a heterogeneous collection of digital,
+analog and mixed signal hardware components.The development of complex 
+embedded system requires sophisticated tool and methodology in order to 
+be able to complete a correct design in time. Embedded system complexity 
+requires the need to strive for abstract specification and formal 
+descriptions each component separatelly in order to produce formal 
+specification of the whole system. The main goal is fast design approach 
+in order to produce simulator and to minimize the time spent.
 
-An introduction.
+The \STs represents an embedded system designed in VHDL.
+
+The authors will illustrate in this paper how BON can represent a high 
+level analogue mapping and will further illustrate a procedure for 
+transformation from BON to VHDL-ASM.
+%=====================================================================
+
+\section{VHDL-ASM}
+\subsection{Hardware description language}
+
+\section{Choice of BON as the modeling tool}
+%Brief description and motivation.
+The Business Object Notation (BON) is an analysis and design notation for 
+object-oriented systems  which emphasizes \emph{seamlessness}, \emph{reversibility} and software \emph{contracting}.
+
+\begin{itemize}
+\item \emph{Seamlessness} is the use of a continuous process throughout the software lifecycle.
+%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
+\item \emph{Reversibility} is the support for both forward and backward engineering: from analysis to
+design and implementation, and back.
+%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
+\item \emph{Contracting} is the precise definition, for each software element, of the associated semantic properties. 
+BON is the only one among the popular analysis methods to use a full assertion
+mechanism, allowing analysts to specify not only the structure of a system but also its
+semantics ( preconditions, postconditions and invariants of the expected results ).
+\end{itemize}
+
+BON focuses on the fundamental elements of analysis and design, and the method is
+meant to be integrated with and adapted to the various development frameworks
+including embedded system.
+
+Because of its flexibility, it is not difficult to find mappings into other domains
+which would seem to indicate that object-oriented modeling is really equivalent
+to process modeling with signal passing, or refinement of VHDL language.
+
+Therefore as a distillation of best modelling practices, and as the 
+language based on contract design and refinement calculus, BON
+is the obvious choice for the current work.
+
+\subsection{Mapping between BON and VHDL}
+%\todo{dragan 25/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+\subsection{Adapting BON for VHDL modeling }
+
+\section{Example}
 

### 2009-07-27 papers/BON_VHDL/paper.tex
documents section start

--- 2009-07-17 papers/BON_VHDL/paper.tex
@@
 \newcommand{\sensorfarm}{The \STs\xspace}
-
+\newcommand{\Sc}{\emph{system\_chart}\xspace}
+\newcommand{\Clc}{\emph{cluster\_chart}\xspace}
+\newcommand{\Csc}{\emph{class\_chart}\xspace}
+\newcommand{\Ft}{\emph{feature}\xspace}
+\newcommand{\ent}{\emph{entity}\xspace}
+\newcommand{\arch}{\emph{architecture}\xspace}
 %---------------------------------------------------------------------
@@
 \title{Refining from the BON High-level Specification Language to the
-  VHDL Hardware Description Language}
+  VHDL-ASM Hardware Description Language}
 
@@
 
-The \STs have several unusual built-in sensors and emitters, sensor
-nodes can be extended via a general-purpose USB bus, and \STs contain
-an FPGA that is dynamically updated on a per-experiment basis.  \STs
-are custom-designed in VHDL.  Like many hardware-based systems, to
-write new software for or against the \ST architecture one either uses
+The \STs, as an embedded system, have several unusual built-in sensors 
+and emitters, sensor nodes can be extended via a general-purpose USB bus, 
+and \STs contain an FPGA that is dynamically updated on a per-experiment 
+basis.  \STs are custom-designed in VHDL and VHDL-ASM.  Like many hardware-based systems,
+ to write new software for or against the \ST architecture one either uses
 the actual hardware or one runs a \STS.  But this simulator emulates a
@@
 \section{Introduction}
+The \STs, represents heterogeneous collection of digital, analog 
+and mixed signal hardware components. Embedded system complexity 
+requires the need to strive for formal descriptions each component 
+separatelly in order to produce specification of the \STs simulator. 
+The development of complex system requires sophisticated tool and methodology 
+in order to be able to complete a correct design in time.
 
-An introduction.
+The authors will illustrate in this paper how BON can represent a high 
+level analogue mapping and will further illustrate a procedure for 
+transformation from BON to VHDL-ASM. As an example, 
+we will give practical aspect of refinement from BON to VHDL-ASM,
+applied on A/D convertor, in order to produce formal specification 
+and software simulation.
+%=====================================================================
+
+\section{VHDL-ASM}
+\subsection{Hardware description language}
+
+\section{Choice of BON as the modeling tool}
+%Brief description and motivation.
+The Business Object Notation (BON) is an analysis and design notation for 
+object-oriented systems  which emphasizes \emph{seamlessness}, \emph{reversibility} and software \emph{contracting}.
+Let us define these principles precisely and point out why the principles are important in VHDL-ASM simulation.
+
+\begin{itemize}
+\item \emph{Seamlessness} is the use of a continuous process throughout the software lifecycle.
+\item \emph{Reversibility} is the support for both forward and backward development process: from analysis to
+design and implementation, and back.
+In VHDL-AMS, modelling represents an continious top-down-bottom-up process. Top-down approach serves for design whereas bottom-up
+serves for verification.
+The possibility of seamlessly translating changes made during a certain development phase back into earlier phases,
+facilitates a stable and predictable environment and ensures the various models consistent with the code.
+
+\item \emph{Contracting} is the precise definition, for each software element, of the associated semantic properties. 
+BON use a full assertion mechanism, allowing analysts to specify not only the structure of a system but also its
+semantics ( constraints and invariants of the expected results ).VHDL-AMS provides strongly typed interface constructs 
+that ensure clarity in the interface specifications.
+
+\end{itemize}
+
+%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
+%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
+Therefore as an approach of best modelling practices, and as the language based on contract design 
+and refinement calculus, BON is the obvious choice for the our simulator.
+
+\subsection{Mapping between BON and VHDL-ASM}
+%\todo{dragan 25/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+\begin{itemize}
+\item \Sc: can be represented with high abstract level of the topmost entity in a design hierarchy.
+\item \Clc: to be described.
+\item \Csc: can be equated with the primary unit(\ent) and secondary unit (\arch) structure in VHDL-AMS.
+The \Csc's name corresponds to the name of the entity in VHDL-AMS.
+\item \Ft: The feature section of the \Csc, where the mathematical transfer function of the component
+is entered, holds the same information as the \arch section of the VHDL-AMS.
+\end{itemize}
+\subsection{Refinement from BON to VHDL-ASM}
 

### 2009-07-27 papers/BON_VHDL/paper.tex
more text

@@
 \newcommand{\arch}{\emph{architecture}\xspace}
+\newcommand{\archs}{\emph{architectures}\xspace}
 %---------------------------------------------------------------------
@@
 %Brief description and motivation.
-The Business Object Notation (BON) is an analysis and design notation for 
-object-oriented systems  which emphasizes \emph{seamlessness}, \emph{reversibility} and software \emph{contracting}.
-Let us define these principles precisely and point out why the principles are important in VHDL-ASM simulation.
-
+The Business Object Notation (BON) is an analysis and design notation 
+for object-oriented systems  which emphasizes \emph{seamlessness}, 
+\emph{reversibility} and software \emph{contracting}. Let us define 
+these principles precisely and point out why the principles are 
+important in VHDL-ASM simulation.
+%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
 \begin{itemize}
-\item \emph{Seamlessness} is the use of a continuous process throughout the software lifecycle.
-\item \emph{Reversibility} is the support for both forward and backward development process: from analysis to
-design and implementation, and back.
-In VHDL-AMS, modelling represents an continious top-down-bottom-up process. Top-down approach serves for design whereas bottom-up
-serves for verification.
-The possibility of seamlessly translating changes made during a certain development phase back into earlier phases,
-facilitates a stable and predictable environment and ensures the various models consistent with the code.
-
-\item \emph{Contracting} is the precise definition, for each software element, of the associated semantic properties. 
-BON use a full assertion mechanism, allowing analysts to specify not only the structure of a system but also its
-semantics ( constraints and invariants of the expected results ).VHDL-AMS provides strongly typed interface constructs 
-that ensure clarity in the interface specifications.
-
+\item \emph{Seamlessness} is the use of a continuous process 
+throughout the software lifecycle.
+%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
+\item \emph{Reversibility} is the support for both forward and 
+backward development process: from analysis to design 
+and implementation, and back.
+In VHDL-AMS, modelling represents an continious top-down-bottom-up
+process. Top-down approach serves for design whereas bottom-up
+serves for verification. The possibility of seamlessly translating 
+changes made during a certain development phase back into earlier phases,
+facilitates a stable and predictable environment and ensures the various 
+models consistent with the code.
+\item \emph{Contracting} is the precise definition, for each software element, 
+of the associated semantic properties. BON use a full assertion mechanism, 
+allowing analysts to specify not only the structure of a system but also its
+semantics ( constraints and invariants of the expected results ).
 \end{itemize}
 
-%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
-%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
-Therefore as an approach of best modelling practices, and as the language based on contract design 
-and refinement calculus, BON is the obvious choice for the our simulator.
+Therefore as an approach of best modelling practices, and as the notation 
+based on contract design and refinement calculus, BON is the obvious choice 
+for the our simulator.
 
@@
 \begin{itemize}
-\item \Sc: can be represented with high abstract level of the topmost entity in a design hierarchy.
+\item \Sc: can be represented with high abstract level of the topmost 
+entity in a design hierarchy.
 \item \Clc: to be described.
-\item \Csc: can be equated with the primary unit(\ent) and secondary unit (\arch) structure in VHDL-AMS.
+\item \Csc: can be identify with the primary unit(\ent) and secondary 
+unit (\arch) structure in VHDL-AMS.
 The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-\item \Ft: The feature section of the \Csc, where the mathematical transfer function of the component
-is entered, holds the same information as the \arch section of the VHDL-AMS.
+\item \Ft: The feature section of the \Csc, where the mathematical 
+transfer function of the component is entered, holds the same information 
+as the \arch section of the VHDL-AMS.
 \end{itemize}
+
 \subsection{Refinement from BON to VHDL-ASM}
+
+An VHDL-AMS component are described using an \ent-\arch structure. 
+The component model contains an \ent and one or more \archs. 
+The \ent specifies the interface of the model to the outside world. 
+In other words, \ent introduces a name, list of input/output ports 
+and the definition of the generic parameters. The graphical 
+presentation of \ent is shown in Figure.1
+
+\begin{figure}[htp]
+\centering
+\include{adc-entity}
+\caption{An \ent body for a A/D converter.}
+\label{figure:adc-entity}
+\end{figure}
+
+The \arch body represents description of the internal implementation 
+of an \ent including only process statements, which are collections 
+of actions to be executed in sequence, or simultaneous statements, 
+which describe equations of analogue system behaviour. 
+The graphical presentation of \ent is shown in Figure.2
+\begin{figure}[htp]
+\centering
+\include{adc-arch}
+\caption{An \arch body for a A/D converter.}
+\label{figure:adc-entity}
+\end{figure}
 

### 2009-07-27 papers/BON_VHDL/paper.tex
minor changes
minor grammar correction
todo tasks for tomorrow
more text

@@
 \section{VHDL-ASM}
-\subsection{Hardware description language}
-
+VHDL-AMS , as an extension of VHDL, was developed to allow designers 
+to represent both analog and digital functionality of a system in 
+the same model. Existing hardware description languages, such as VHDL 
+and Verilog, support digital systems modeling, while languages such as 
+SPICE and its numerous variants support analog systems modeling. 
+Integrating these two modeling domains is a significant new capability 
+for designers. VHDL-AMS supports the differential and algebraic equations 
+(DAEs) where the solution varies continuously with time and, in addition, 
+supports both structural composition and behavioural descriptions of 
+analogue systems. 
 \section{Choice of BON as the modeling tool}
@@
 In VHDL-AMS, modelling represents an continious top-down-bottom-up
-process. Top-down approach serves for design whereas bottom-up
+process. Top-down approach serves for design while bottom-up
 serves for verification. The possibility of seamlessly translating 
@@
 \subsection{Mapping between BON and VHDL-ASM}
-%\todo{dragan 25/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
 \begin{itemize}
@@
 \include{adc-entity}
-\caption{An \ent body for a A/D converter.}
+\caption{An \ent body for A/D converter.}
 \label{figure:adc-entity}
@@
 which describe equations of analogue system behaviour. 
-The graphical presentation of \ent is shown in Figure.2
+The graphical presentation of \arch is shown in Figure.2
 \begin{figure}[htp]
@@
 \include{adc-arch}
-\caption{An \arch body for a A/D converter.}
+\caption{An \arch body for A/D converter.}
 \label{figure:adc-entity}
 \end{figure} 
+%\todo{dragan 27/07/2009 - Start with:1) BON charts and diagrams.}
 

### 2009-07-27 papers/BON_VHDL/paper.tex
restore

@@
 \section{VHDL-ASM}
-VHDL-AMS , as an extension of VHDL, was developed to allow designers 
-to represent both analog and digital functionality of a system in 
-the same model. Existing hardware description languages, such as VHDL 
-and Verilog, support digital systems modeling, while languages such as 
-SPICE and its numerous variants support analog systems modeling. 
-Integrating these two modeling domains is a significant new capability 
-for designers. VHDL-AMS supports the differential and algebraic equations 
-(DAEs) where the solution varies continuously with time and, in addition, 
-supports both structural composition and behavioural descriptions of 
-analogue systems. 
+
 \section{Choice of BON as the modeling tool}
@@
 The graphical presentation of \arch is shown in Figure.2
+
 \begin{figure}[htp]

### 2009-07-28 papers/BON_VHDL/paper.tex
grafical --> textual
minor changes

@@
 \section{VHDL-ASM}
-
+VHDL-AMS is the IEEE standard modeling language (standard 1076.1) created 
+to provide compatibility and capability in an open language for modern analog, 
+mixed-signal, multi-domain designs.
 \section{Choice of BON as the modeling tool}
@@
 for the our simulator.
+Throughout this paper we will try to make this clear.  
+\section{Refinement from BON to VHDL-ASM}
 
@@
 \end{itemize}
-
-\subsection{Refinement from BON to VHDL-ASM}
-
 An VHDL-AMS component are described using an \ent-\arch structure. 
@@
 In other words, \ent introduces a name, list of input/output ports 
-and the definition of the generic parameters. The graphical 
-presentation of \ent is shown in Figure.1
-
-\begin{figure}[htp]
-\centering
-\include{adc-entity}
-\caption{An \ent body for A/D converter.}
-\label{figure:adc-entity}
-\end{figure}
+and the definition of the generic parameters. The textual 
+presentation of \ent is shown in Figure.1. 
+\input{adc-entity}
 
@@
 which describe equations of analogue system behaviour. 
-The graphical presentation of \arch is shown in Figure.2
+The textual presentation of \arch is shown in Figure.2.
+\input{adc-arch}
 
-\begin{figure}[htp]
-\centering
-\include{adc-arch}
-\caption{An \arch body for A/D converter.}
-\label{figure:adc-entity}
-\end{figure} 
 %\todo{dragan 27/07/2009 - Start with:1) BON charts and diagrams.}

### 2009-07-28 papers/BON_VHDL/paper.tex
yes

@@
 \title{Refining from the BON High-level Specification Language to the
-  VHDL-ASM Hardware Description Language}
+  VHDL-AMS Hardware Description Language}
 
@@
 and \STs contain an FPGA that is dynamically updated on a per-experiment 
-basis.  \STs are custom-designed in VHDL and VHDL-ASM.  Like many hardware-based systems,
+basis.  \STs are custom-designed in VHDL and VHDL-AMS.  Like many hardware-based systems,
  to write new software for or against the \ST architecture one either uses
@@
 level analogue mapping and will further illustrate a procedure for 
-transformation from BON to VHDL-ASM. As an example, 
-we will give practical aspect of refinement from BON to VHDL-ASM,
+transformation from BON to VHDL-AMS. As an example, 
+we will give practical aspect of refinement from BON to VHDL-AMS,
 applied on A/D convertor, in order to produce formal specification 
@@
 
-\section{VHDL-ASM}
+\section{VHDL-AMS}
 VHDL-AMS is the IEEE standard modeling language (standard 1076.1) created 
@@
 these principles precisely and point out why the principles are 
-important in VHDL-ASM simulation.
-%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
+important in VHDL-AMS simulation.
+%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
 \begin{itemize}
@@
 throughout the software lifecycle.
-%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
+%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
 \item \emph{Reversibility} is the support for both forward and 
@@
 Throughout this paper we will try to make this clear.  
-\section{Refinement from BON to VHDL-ASM}
+\section{Refinement from BON to VHDL-AMS}
 
-\subsection{Mapping between BON and VHDL-ASM}
+\subsection{Mapping between BON and VHDL-AMS}
 %\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}

### 2009-07-29 papers/BON_VHDL/paper.tex
paper progress

@@
 \newcommand{\archs}{\emph{architectures}\xspace}
+\newcommand{\duc}{\emph{design units collection}\xspace}
+\newcommand{\ducs}{\emph{design units collections}\xspace}
+\newcommand{\du}{\emph{design unit}\xspace}
 %---------------------------------------------------------------------
@@
 level analogue mapping and will further illustrate a procedure for 
-transformation from BON to VHDL-AMS. As an example, 
+transformation from BON to VHDL-AMS. As an extremly simple example, 
 we will give practical aspect of refinement from BON to VHDL-AMS,
@@
 \section{VHDL-AMS}
-VHDL-AMS is the IEEE standard modeling language (standard 1076.1) created 
-to provide compatibility and capability in an open language for modern analog, 
-mixed-signal, multi-domain designs.
+IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
+high-level description language for electronic and multi-domain 
+systems. VHDL-AMS, as an superset of VHDL, supports modeling and 
+simulation of analog and mixed-signal systems. The main contribution 
+of VHDL-AMS is ability to describe the behavior of complex system 
+trought differential and algebraic equations (DAEs). For this purpose, 
+VHDL-AMS introduces the following capabilities and therms:
+
+\begin{itemize}
+\item 
+\emph{Terminals} represents phisical connection points or circuit nodes 
+of a network connectivity. Terminals are declared to be of various 
+\emph{natures}, which represent different energy domains of a system.
+\item
+The \emph{simultaneous statements} creates one or more characteristic 
+equations that are used by the "analog solver" to model continuously 
+valued behaviour.  
+\item
+The \emph{quantities} denote a waveform or a time series of values.
+The behaviour of the \emph{quantities} is given by the \emph{simultaneous
+ statements}. They take values as a result of solving the set of simultaneous 
+ordinary differential and algebraic equations (DAEs). To simplify, 
+quantities can be understood as analog unknowns of (DAEs) that specify 
+analog behavior. 
+\item
+The \emph{break statement} is an construct which allows to 
+initialize the DAE system at the beginning of a transient 
+simulation and to mark discontinuities during such a transient 
+simulation.
+
+\end{itemize}
+
 \section{Choice of BON as the modeling tool}
@@
 \begin{itemize}
-\item \Sc: can be represented with high abstract level of the topmost 
-entity in a design hierarchy.
-\item \Clc: to be described.
-\item \Csc: can be identify with the primary unit(\ent) and secondary 
+\item \Sc: VHDL-AMS allows hierarchical structure modeling, 
+meaning that subsistems ( \ducs ), can be developed independently.
+As an top-level description (informal/formal), \Sc connect all the \ducs
+together to a complete design. 
+\item \Clc: An  subsistem which represent \duc.
+\item \Csc: describes \du. It can be identify with the primary unit (\ent) and secondary 
 unit (\arch) structure in VHDL-AMS.

### 2009-07-29 sb_driver/sb_driver/trunk/src/ie/ucd/sensetile/sensorboard/driver/InputStreamPacketInputStream.java
Sensor Board Driver Java: added abstract contract test case for PacketInputStream added concrete test cases to prove the contract holds

@@ -351,6 +351,12 @@ public class InputStreamPacketInputStream implements PacketInputStream {
         final Packet[] array, 
         final int offset, 
         final int length) {
+      if (
+          (offset < 0) ||
+          (length < 0) ||
+          (offset + length > array.length)) {
+        throw new IndexOutOfBoundsException();
+      }
       this.array = array;
       this.offset = offset;
       this.length = length;

### 2009-07-29 sb_driver/sb_driver/trunk/src/ie/ucd/sensetile/sensorboard/simulator/PacketBuilder.java

@@ -2,8 +2,18 @@ package ie.ucd.sensetile.sensorboard.simulator;
 
 import ie.ucd.sensetile.sensorboard.Packet;
 
+/**
+ * Packet builder class to be used to build simulators.
+ * 
+ * @author delbianc
+ */
 public interface PacketBuilder {
   
+  /**
+   * Return a packet.
+   * 
+   * @return next packet
+   */
   Packet getPacket();
   
 }

### 2009-07-29 sb_driver/sb_driver/trunk/src/ie/ucd/sensetile/sensorboard/simulator/SimulatorPacketInputStream.java

@@ -41,6 +41,9 @@ public class SimulatorPacketInputStream implements PacketInputStream {
       final Packet[] array, final int offset, final int length)
       throws IOException, SenseTileException {
     checkClose();
+    if (length < 0) {
+      throw new IndexOutOfBoundsException();
+    }
     for (int index = offset; index < offset + length; index++) {
       array[index] = internalRead();
     }

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/PacketInputStreamUnitTestCase.java

@@ -1,0 +1,221 @@
+package ie.ucd.sensetile.sensorboard;
+
+import static org.junit.Assert.assertEquals;
+import static org.junit.Assert.assertNotNull;
+import static org.junit.Assert.assertNull;
+import static org.junit.Assert.assertTrue;
+import static org.junit.Assert.fail;
+
+import java.io.IOException;
+
+import org.junit.After;
+import org.junit.Before;
+import org.junit.Test;
+
+public abstract class PacketInputStreamUnitTestCase {
+  
+  protected abstract PacketInputStream createPacketInputStream(
+      final int availablePackets);
+  
+  protected abstract void disposePacketInputStream(
+      final PacketInputStream packetInputStream2);
+  
+  protected void setPacketInputStream(
+      final PacketInputStream packetInputStream) {
+    this.packetInputStream = packetInputStream;
+  }
+  
+  protected PacketInputStream getPacketInputStream() {
+    return packetInputStream;
+  }
+  
+  private PacketInputStream packetInputStream;
+  
+  @Before
+  public void setUp() throws Exception {
+    int availablePackets = 100;
+    setPacketInputStream(createPacketInputStream(availablePackets));
+  }
+  
+  @After
+  public void tearDown() throws Exception {
+    disposePacketInputStream(getPacketInputStream());
+  }
+  
+  @Test
+  public void testAvailablePackets() throws IOException {
+    assertTrue(getPacketInputStream().availablePackets() > 0);
+  }
+  
+  @Test
+  public void testRead() throws IOException, SenseTileException {
+    assertNotNull(getPacketInputStream().read());
+  }
+  
+  @Test
+  public void testReadArray() throws IOException, SenseTileException {
+    int arrayLength = 10;
+    Packet[] array = new Packet[arrayLength];
+    assertEquals(arrayLength, getPacketInputStream().read(array));
+    for (int index = 0; index < array.length; index++) {
+      assertNotNull(array[index]);
+    }
+  }
+  
+  @Test (expected = NullPointerException.class)
+  public void testReadArrayNull() throws IOException, SenseTileException {
+    getPacketInputStream().read(null);
+  }
+  
+  @Test
+  public void testReadArrayOffset() throws IOException, SenseTileException {
+    int arrayLength = 20;
+    int offset = 5;
+    int readLength = 10;
+    Packet[] array = new Packet[arrayLength];
[…]

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/driver/AllUnitTests.java

@@ -5,9 +5,10 @@ import org.junit.runners.Suite;
 
 @RunWith(Suite.class)
 @Suite.SuiteClasses({
-  InputStreamPacketInputStreamUnitTest.class,
+  ByteArrayPacketUnitTest.class,
   ByteArrayPacketByteArrayFrameTest.class,
-  ByteArrayPacketUnitTest.class})
+  InputStreamPacketInputStreamAsPacketInputStreamUnitTest.class,
+  InputStreamPacketInputStreamUnitTest.class})
 
 public class AllUnitTests {
   

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/driver/InputStreamPacketInputStreamAsPacketInputStreamUnitTest.java

@@ -1,0 +1,26 @@
+package ie.ucd.sensetile.sensorboard.driver;
+
+
+import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import ie.ucd.sensetile.sensorboard.PacketInputStreamUnitTestCase;
+
+import java.io.ByteArrayInputStream;
+import java.io.InputStream;
+
+public class InputStreamPacketInputStreamAsPacketInputStreamUnitTest 
+    extends PacketInputStreamUnitTestCase {
+  
+  @Override
+  protected PacketInputStream createPacketInputStream(final int availablePackets) {
+    byte[] rawPacket = PacketRawByteArrayBuilder.prepare(
+        ByteArrayPacket.LENGTH * availablePackets);
+    InputStream is = new ByteArrayInputStream(rawPacket);
+    PacketInputStream pis = new InputStreamPacketInputStream(is);
+    return pis;
+  }
+  
+  @Override
+  protected void disposePacketInputStream(final PacketInputStream packetInputStream) {
+  }
+
+}

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/driver/InputStreamPacketInputStreamUnitTest.java

@@ -10,7 +10,6 @@ import java.io.ByteArrayInputStream;
 import java.io.EOFException;
 import java.io.IOException;
 import java.io.InputStream;
-import java.util.Arrays;
 import java.util.Properties;
 
 import org.junit.BeforeClass;
@@ -18,31 +17,7 @@ import org.junit.Test;
 
 public class InputStreamPacketInputStreamUnitTest {
   
-  private static byte[] rawPacketArray;
   private static Properties defaultProperties;
-  
-  @BeforeClass 
-  public static void initRawPacketArray() {
-    InputStreamPacketInputStreamUnitTest.rawPacketArray = 
-      new byte[ByteArrayPacket.LENGTH * 200];
-    byte[] array = rawPacketArray;
-    for (
-        int packetIndex = 0; 
-        packetIndex <= array.length / ByteArrayPacket.LENGTH; 
-        packetIndex ++) {
-      for (
-          int patternIndex = 0; 
-          patternIndex < ByteArrayPacket.PATTERN.length; 
-          patternIndex++) {
-        int index = 
-          (patternIndex + ByteArrayPacket.PATTERN_OFFSET) % 
-          ByteArrayPacket.LENGTH + packetIndex * ByteArrayPacket.LENGTH;
-        if (index < array.length) {
-          array[index] = ByteArrayPacket.PATTERN[patternIndex];
-        }
-      }
-    }
-  }
   
   @BeforeClass 
   public static void initDefaultProperties() {
@@ -81,7 +56,9 @@ public class InputStreamPacketInputStreamUnitTest {
   @Test
   public void testNotValidLongThanValidFullRead() throws IOException, SenseTileException{
     byte[] invalid = new byte[ByteArrayPacket.LENGTH * 100];
-    byte[] valid = prepareRawPacketArray(ByteArrayPacket.LENGTH * 100 + 100, 50);
+    byte[] valid = 
+      PacketRawByteArrayBuilder.prepare(
+          ByteArrayPacket.LENGTH * 100 + 100, 50);
     byte[] total = new byte[invalid.length + valid.length];
     System.arraycopy(invalid, 0, total, 0, invalid.length);
     System.arraycopy(valid, 0, total, invalid.length, valid.length);
@@ -107,7 +84,8 @@ public class InputStreamPacketInputStreamUnitTest {
   
   @Test
   public void testAvailable1() throws IOException{
[…]

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/driver/PacketRawByteArrayBuilder.java

@@ -1,0 +1,71 @@
+package ie.ucd.sensetile.sensorboard.driver;
+
+import java.util.Arrays;
+
+public final class PacketRawByteArrayBuilder {
+  
+  private PacketRawByteArrayBuilder() {
+  }
+
+  private static byte[] array = new byte[0];
+  
+  private static byte[] prepareArray(final int length) {
+    final byte[] rawPacketArray = new byte[ByteArrayPacket.LENGTH * length];
+    final byte[] array = rawPacketArray;
+    for (
+        int packetIndex = 0; 
+        packetIndex <= array.length / ByteArrayPacket.LENGTH; 
+        packetIndex ++) {
+      for (
+          int patternIndex = 0; 
+          patternIndex < ByteArrayPacket.PATTERN.length; 
+          patternIndex++) {
+        int index = 
+          (patternIndex + ByteArrayPacket.PATTERN_OFFSET) % 
+          ByteArrayPacket.LENGTH + packetIndex * ByteArrayPacket.LENGTH;
+        if (index < array.length) {
+          array[index] = ByteArrayPacket.PATTERN[patternIndex];
+        }
+      }
+    }
+    return rawPacketArray;
+  }
+  
+  private static byte[] getArray(final int length) {
+    if (length >= array.length / ByteArrayPacket.LENGTH) {
+      setArray(prepareArray(length));
+    }
+    return array;
+  }
+  
+  private static void setArray(final byte[] rawPacketArray) {
+    PacketRawByteArrayBuilder.array = rawPacketArray;
+  }
+  
+  /**
+   * Builds a raw bytes packet array. 
+   * 
+   * @param length bytes length of array
+   * @return byte array of packets
+   */
+  public static byte[] prepare(final int length) {
+    return prepare(length, 0);
+  }
+  
+  /**
+   * Builds a raw bytes packet array with an offset. 
+   * 
+   * @param length bytes length of array
+   * @param offset packet offset in array
+   * @return byte array of packets
+   */
+  public static byte[] prepare(final int length, final int offset) {
+    int copyOffset = (ByteArrayPacket.LENGTH  - offset) % ByteArrayPacket.LENGTH;
+    if (copyOffset < 0) {
+      copyOffset = copyOffset + ByteArrayPacket.LENGTH;
+    }
+    return Arrays.copyOfRange(
+        getArray(length / ByteArrayPacket.LENGTH + 2), 
+        copyOffset, copyOffset + length);
+  }
+}

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/simulator/AllUnitTests.java

@@ -6,6 +6,7 @@ import org.junit.runners.Suite;
 @RunWith(Suite.class)
 @Suite.SuiteClasses({
   ClonePacketBuilderTest.class, 
+  SimulatorPacketInputStreamAsPacketInputStreamUnitTest.class, 
   SimulatorPacketInputStreamTest.class})
 
 public class AllUnitTests {

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/simulator/SimulatorPacketInputStreamAsPacketInputStreamUnitTest.java

@@ -1,0 +1,20 @@
+package ie.ucd.sensetile.sensorboard.simulator;
+
+import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import ie.ucd.sensetile.sensorboard.PacketInputStreamUnitTestCase;
+
+public class SimulatorPacketInputStreamAsPacketInputStreamUnitTest 
+    extends PacketInputStreamUnitTestCase {
+  
+  @Override
+  protected PacketInputStream createPacketInputStream(final int availablePackets) {
+    PacketBuilder builder = new ClonePacketBuilder(new InstancePacket());
+    SimulatorPacketInputStream simulator = new SimulatorPacketInputStream(builder);
+    return simulator;
+  }
+  
+  @Override
+  protected void disposePacketInputStream(final PacketInputStream packetInputStream) {
+  }
+  
+}

### 2009-07-29 sb_driver/sb_driver/trunk/test/src/ie/ucd/sensetile/sensorboard/simulator/SimulatorPacketInputStreamTest.java

@@ -1,10 +1,11 @@
 package ie.ucd.sensetile.sensorboard.simulator;
 
-import static org.junit.Assert.*;
 import static org.junit.Assert.assertNotNull;
-
-import ie.ucd.sensetile.sensorboard.SenseTileException;
+import static org.junit.Assert.assertNull;
+import static org.junit.Assert.assertTrue;
+import static org.junit.Assert.fail;
 import ie.ucd.sensetile.sensorboard.Packet;
+import ie.ucd.sensetile.sensorboard.SenseTileException;
 
 import java.io.IOException;
 

### 2009-07-29 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol structure

@@
 \begin{verbatim}
+
 OBJECTIVE: 
@@
 Specifications serves two purposes, and indirectly a third one.
+
 \end{verbatim}
@@
   Windows, at the moment) over a USB bus using a streaming
-  packet-based protocol.  This sensor platform is a custom-designed,
+  packet-based protocol. 
+  This sensor platform is a custom-designed,
   but inexpensive, sensor platform (called the \ST) paired with
@@
   There are over a dozen sensors on the \ST itself, and new sensors
-  can be added by adding them to the USB bus.  A third-party low-level
-  C driver, provided by the USB chip vendor, creates and parses
-  generic USB communication packets, and a custom medium-level Java
-  driver creates and parses \ST streaming communication packets.  This
+  can be added by adding them to the USB bus.
+  A third-party low-level
+  library, provided by the USB chip vendor, creates and parses
+  generic USB communication packets, while a custom medium-level Java
+  driver creates and parses \ST streaming communication packets. 
+  This
   latter Java driver is formally specified, validated with rigorous
-  unit testing, and verified using extended static checking.  However,
-  because of this process and combination of technologies, using this
+  unit testing, and verified using extended static checking. 
+  Because of this process and combination of technologies, using this
   specification we are able to both communicate with a real \ST
@@
   model-based specification for the protocol itself written in the
-  Java Modeling Language.  This protocol model is itself an executable
+  Java Modeling Language.
+  \todo executable? how? what does this mean?
+  This protocol model is itself an executable
   stream simulator that can be used to develop new \ST user
-  applications without having the \ST hardware at all.  Consequently,
+  applications without having the \ST hardware at all. 
+  Consequently,
   from this single specification, we can test that the \ST hardware
@@
 
+\todo extend
+
+\paragraph*{Outline}
+
+In section \ref{sec:data_stream_and_protocol_model} the data stream model 
+and protocol is presented, we will give both a higl level overview of the 
+protocol, and a formalized, JML based, specification,
+
+In section \ref{sec:test_cases} the test cases structure is presented, for 
+both the hand made and the generated ones. 
+Both the approaches are analyzed, their pro and contro are discussed, as 
+well as the expected advantages to use them in combination.
+
+In section \ref{sec:test_cases} the test cases structure is presented, for 
+both the hand made and the generated ones. 
+Both the approaches are analyzed, their pro and contro are discussed, as 
+well as the expected advantages to use them in combination.
+
+In section \ref{sec:test_cases_on_the_job} the execution of the test suites
+is described, and the results are analyzed and compared, both from qualitative
+and quantitative points of view.
+
+In section \ref{sec:test_cases_and_the_scenarios} we examine the different 
+scenarios on which test cases based on specifications can be applied, and 
+the resulting benefits it is possible to obtain.
+
+In section \ref{sec:related_works} related works are presented and analyzed.
+
+Finally in section \ref{sec:conclusions} we outline some conclusions and we 
+briefly describe our ongoing and future works on similar topics.
+
 %=====================================================================
-\Section{Conclusion}
+\section{Data stream and protocol model}
+\label{sec:data_stream_and_protocol_model}
 
-A conclusion.
+\todo describe package
+
+\todo describe high level representation of package: sensors and data 
+details remains, exact package structure and sensor data location in the
+packege are lost
+
+\todo describe JML specification of the high levelr representation of the 
+package
+
+%=====================================================================
+\section{Test cases}
+\label{sec:test_cases}
+
+\todo describe dual approach: hand made tests + generated from 
+specification tests
+
+\todo describe hybrid agile and TDD approach: high level informal 
+specifications with BON, iterative refinements of code, to distill 
+test cases and formal specification of the interfaces
+
+\todo describe the necessity to add specification generated tests: to ensure
+that specification is sound, to obtain higher coverage
+
+\paragraph*{Hand made tests}
+
+\todo describe hand made tests structure
+
+\todo describe abstract test case and abstract test suite patterns: how to 
+write tests against interfaces and/or abstract classes
+
+\paragraph*{Generated tests}
+
+\todo describe the tool that generates the testing
+
+\todo describe the structure of the generates tests
+
+%=====================================================================
+\section{Test cases on the job}
+\label{sec:test_cases_on_the_job}
+
+\todo report high level analysis of the test suites
+
+\todo describe measures: coverage, number of new bugs found
+
+\todo compare hand made vs generated tests
+
+\todo compare compare coupled approach vs single approaches
+
+%=====================================================================
+\section{Test cases and the scenarios}
+\label{sec:test_cases_and_the_scenarios}
+
+\todo benefit: the approch can be used to test different part of the system
+that share the same specification
+
+\todo testing the stream
+
+\todo testing the simulator
+
+\todo describe general approach: depending at what level (of abstraction) tests 
+and specifications are written, it is possible to verify both the simulator and 
+the real protocol (isn't this obvious?)
+
+%=====================================================================
+\section{Related works}
+\label{sec:related_works}
+
+\todo
+
+%=====================================================================
+\section{Conclusion}
+\label{sec:conclusion}
+
+\todo
 

### 2009-08-02 papers/BON_VHDL/paper.tex
sistematisation

@@
 \usepackage{times}
-
+\usepackage{multirow}
 \usepackage{ifpdf}
@@
 \newcommand{\sensorfarm}{The \STs\xspace}
-\newcommand{\Sc}{\emph{system\_chart}\xspace}
-\newcommand{\Clc}{\emph{cluster\_chart}\xspace}
-\newcommand{\Csc}{\emph{class\_chart}\xspace}
+\newcommand{\Sc}{\emph{SYSTEM\_CHART}\xspace}
+\newcommand{\Clc}{\emph{CLUSTER\_CHART}\xspace}
+\newcommand{\Csc}{\emph{CLASS\_CHART}\xspace}
 \newcommand{\Ft}{\emph{feature}\xspace}
@@
 \newcommand{\du}{\emph{design unit}\xspace}
+\newcommand{\undef}{\emph{undefined}\xspace}
 %---------------------------------------------------------------------
@@
 Throughout this paper we will try to make this clear.  
+\section{Mapping between BON and VHDL-AMS}
+\subsection{Static model mapping}
+%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+\begin{center}
+    \begin{tabular}{|l|l|l|}
+    	\hline
+    	VHDL-AMS & BON & Description \\ \hline
+    	\ducs (packages) & \Sc & 
+	\begin{minipage}[c]{7cm} 
+                 \vskip 3mm
+    		VHDL-AMS allows hierarchical structure modeling, 
+    		meaning that subsistems (\ducs), can be developed independently.
+    		As an top-level description (informal/formal), \Sc connect all the \ducs
+    		together to a complete design. Actually \Sc can be understood 
+                as VHDL-AMS package as a way of grouping a collection of related 
+                declarations that serve a common purpose.
+                \\ 
+    	\end{minipage}\\ \hline
+        \duc & \Clc &
+	\begin{minipage}[c]{7cm} 
+                \vskip 3mm
+		An  subsistem which represent \duc.
+                \\ 
+    	\end{minipage}\\ \hline
+	 \du & \Csc &
+	\begin{minipage}[c]{7cm}
+	      \vskip 3mm	 
+              \Csc  describes \du. It can be identify with the primary unit (\ent) and secondary 
+	      unit (\arch) structure in VHDL-AMS.
+	      The \Csc's name corresponds to the name of the entity in VHDL-AMS.
+              \\
+    	\end{minipage}\\ \hline
+        \arch & \Ft & 
+	\begin{minipage}[c]{7cm} 
+              \vskip 3mm
+	      The feature section of the \Csc, where the mathematical 
+	      transfer function of the component is entered, holds the same information 
+	      as the \arch section of the VHDL-AMS.
+              \\
+    	\end{minipage}\\ \hline
+
+    \end{tabular}
+\end{center}
+\subsection{Operators-expressions mapping}
+%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
+This subsection summarises all relations on expressions - operators between VHDL-AMS and BON.
+\begin{center}
+    \begin{tabular}{|c|c|l|}
+    	\hline
+    	VHDL-AMS & BON & Description \\ \hline
+    	\begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{**}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{\textasciicircum}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		exponentiation.
+                \\ 
+    	\end{minipage}\\ \hline
+
+       \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{+, \textendash, *, /}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{+, \textendash, *, /}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		identity, addition, negation, subtraction, multiplication, division.
+                \\ 
+    	\end{minipage}\\ \hline
+
+       \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{=, /=, \textless, \textless=, \textgreater, \textgreater=, mod}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{=, /=, \textless, \textless=, \textgreater, \textgreater=, $\backslash\backslash$}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		equality, inequality, less than, less than or equal , greater than, greater than or equal, modulo.
+                \\ 
+    	\end{minipage}\\ \hline
+
+       \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{not, and, or, xor}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{not, and, or, xor}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	negation, logical and, logical or, exclusive or.
+                \\ 
+    	\end{minipage}\\ \hline
+
+	\begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{nand}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{not and}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	negated logical and.
+                \\ 
+    	\end{minipage}\\ \hline
+
+        \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{nor}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{not or}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	negated logical or.
+                \\ 
+    	\end{minipage}\\ \hline
+
+	 \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{xnor}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \textbf{not xor}
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	negated exclusive or.
+                \\ 
+    	\end{minipage}\\ \hline
+
+	\begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{sll, srl, sla, sra, rol, ror}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \undef
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	shift-left/right logical,\\
+                        shift-left/right arithmetic,\\
+                        rotate left/right.
+                \\ 
+    	\end{minipage}\\ \hline
+
+       \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{rem}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \undef
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	reminder.
+                \\ 
+    	\end{minipage}\\ \hline
+
+       \begin{minipage}[c]{2cm} 
+         \centering 
+         \textbf{abs}
+        \end{minipage}  
+        &
+       \begin{minipage}[c]{2cm}
+         \centering 
+         \undef
+        \end{minipage}  
+        & 
+	\begin{minipage}[c]{5cm} 
+                 \vskip 3mm
+    		 	absolute value.
+                \\ 
+    	\end{minipage}\\ \hline
+
+    \end{tabular}
+\end{center}
+\subsection{Type mapping}
+
 \section{Refinement from BON to VHDL-AMS}
 
-\subsection{Mapping between BON and VHDL-AMS}
-%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
-\begin{itemize}
-\item \Sc: VHDL-AMS allows hierarchical structure modeling, 
-meaning that subsistems ( \ducs ), can be developed independently.
-As an top-level description (informal/formal), \Sc connect all the \ducs
-together to a complete design. 
-\item \Clc: An  subsistem which represent \duc.
-\item \Csc: describes \du. It can be identify with the primary unit (\ent) and secondary 
-unit (\arch) structure in VHDL-AMS.
-The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-\item \Ft: The feature section of the \Csc, where the mathematical 
-transfer function of the component is entered, holds the same information 
-as the \arch section of the VHDL-AMS.
-\end{itemize}
 An VHDL-AMS component are described using an \ent-\arch structure. 

### 2009-08-02 papers/BON_VHDL/paper.tex
breaking docs

--- 2009-07-29 papers/BON_VHDL/paper.tex
@@
 in order to be able to complete a correct design in time.  
-
+This article describes a refinement method from BON to VHDL-AMS
 The authors will illustrate in this paper how BON can represent a high 
@@
 Throughout this paper we will try to make this clear.  
+\section{Mapping between BON and VHDL-AMS}
+\subsection{Static model mapping}
+%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+\input{static}
+\subsection{Operators-expressions mapping}
+%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
+This subsection summarises all relations on expressions - operators between VHDL-AMS and BON.
+\input{operators}
+
+\subsection{Type mapping}
+
 \section{Refinement from BON to VHDL-AMS}
 
-\subsection{Mapping between BON and VHDL-AMS}
-%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
-\begin{itemize}
-\item \Sc: VHDL-AMS allows hierarchical structure modeling, 
-meaning that subsistems ( \ducs ), can be developed independently.
-As an top-level description (informal/formal), \Sc connect all the \ducs
-together to a complete design. 
-\item \Clc: An  subsistem which represent \duc.
-\item \Csc: describes \du. It can be identify with the primary unit (\ent) and secondary 
-unit (\arch) structure in VHDL-AMS.
-The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-\item \Ft: The feature section of the \Csc, where the mathematical 
-transfer function of the component is entered, holds the same information 
-as the \arch section of the VHDL-AMS.
-\end{itemize}
 An VHDL-AMS component are described using an \ent-\arch structure. 

### 2009-08-02 papers/BON_VHDL/static.tex
breaking docs

\begin{center}
    \begin{tabular}{|l|l|l|}
    	\hline
    	VHDL-AMS & BON & Description \\ \hline
    	\ducs (packages) & \Sc & 
	\begin{minipage}[c]{7cm} 
                 \vskip 3mm
    		VHDL-AMS allows hierarchical structure modeling, 
    		meaning that subsistems (\ducs), can be developed independently.
    		As an top-level description (informal/formal), \Sc connect all the \ducs
    		together to a complete design. Actually \Sc can be understood 
                as VHDL-AMS package as a way of grouping a collection of related 
                declarations that serve a common purpose.
                \\ 
    	\end{minipage}\\ \hline
        \duc & \Clc &
	\begin{minipage}[c]{7cm} 
                \vskip 3mm
		An  subsistem which represent \duc.
                \\ 
    	\end{minipage}\\ \hline
	 \du & \Csc &
	\begin{minipage}[c]{7cm}
	      \vskip 3mm	 
              \Csc  describes \du. It can be identify with the primary unit (\ent) and secondary 
	      unit (\arch) structure in VHDL-AMS.
	      The \Csc's name corresponds to the name of the entity in VHDL-AMS.
              \\
    	\end{minipage}\\ \hline
        \arch & \Ft & 
	\begin{minipage}[c]{7cm} 
              \vskip 3mm
	      The feature section of the \Csc, where the mathematical 
	      transfer function of the component is entered, holds the same information 
	      as the \arch section of the VHDL-AMS.
              \\
    	\end{minipage}\\ \hline

    \end{tabular}
\end{center}

### 2009-08-02 papers/BON_VHDL/paper.tex
Appendix

@@
 in order to be able to complete a correct design in time.  
-This article describes a refinement method from BON to VHDL-AMS
 The authors will illustrate in this paper how BON can represent a high 
 level analogue mapping and will further illustrate a procedure for 
-transformation from BON to VHDL-AMS. As an extremly simple example, 
-we will give practical aspect of refinement from BON to VHDL-AMS,
+transformation from VHDL-AMS to BON. As an extremly simple example, 
+we will give practical aspect of refinement from VHDL-AMS to BON,
 applied on A/D convertor, in order to produce formal specification 
-and software simulation.
+and software simulation. A complete list of the mapping between BON 
+and VHDL-AMS can be found in the appendix.
 %=====================================================================
@@
 \section{Mapping between BON and VHDL-AMS}
-\subsection{Static model mapping}
-%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
-\input{static}
-\subsection{Operators-expressions mapping}
-%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
-This subsection summarises all relations on expressions - operators between VHDL-AMS and BON.
-\input{operators}
-
-\subsection{Type mapping}
 
@@
 
-An VHDL-AMS component are described using an \ent-\arch structure. 
-The component model contains an \ent and one or more \archs. 
-The \ent specifies the interface of the model to the outside world. 
-In other words, \ent introduces a name, list of input/output ports 
-and the definition of the generic parameters. The textual 
-presentation of \ent is shown in Figure.1. 
+An VHDL-AMS design units are described using an \ent-\arch structure. 
+The design unit contains an \ent and one or more \archs. 
+The \ent specifies the interface that an entity provides of itself 
+to the outside. In other words, \ent introduces a name, list of 
+input/output ports and the definition of the generic parameters. 
+The textual presentation of \ent is shown in Figure.1. 
 \input{adc-entity}
@@
 
-%======================================================================
 %% \nocite{ex1,ex2}
@@
 
+%======================================================================
+\appendix
+\section {APPENDIX}
+%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
+\input{static}
+%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
+\input{operators}
+%======================================================================
+
 \end{document}

### 2009-08-02 papers/BON_VHDL/static.tex
Appendix

@@
 \begin{center}
-    \begin{tabular}{|l|l|l|}
+    \begin{tabular}{|c|c|l|}
+        \hline
+        \multicolumn{3}{|c|}{Static mapping}\\ \hline
     	\hline
     	VHDL-AMS & BON & Description \\ \hline
-    	\ducs (packages) & \Sc & 
-	\begin{minipage}[c]{7cm} 
-                 \vskip 3mm
+    	\ducs & \Sc & 
+	\begin{minipage}[c]{0.5\linewidth}
+                 \vskip 1mm
     		VHDL-AMS allows hierarchical structure modeling, 
@@
                 declarations that serve a common purpose.
-                \\ 
+
     	\end{minipage}\\ \hline
         \duc & \Clc &
-	\begin{minipage}[c]{7cm} 
-                \vskip 3mm
+	\begin{minipage}[c]{0.5\linewidth} 
+                \vskip 1mm
 		An  subsistem which represent \duc.
-                \\ 
+
     	\end{minipage}\\ \hline
 	 \du & \Csc &
-	\begin{minipage}[c]{7cm}
-	      \vskip 3mm	 
+	\begin{minipage}[c]{0.5\linewidth}
+	      \vskip 1mm	 
               \Csc  describes \du. It can be identify with the primary unit (\ent) and secondary 
@@
 	      The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-              \\
+
     	\end{minipage}\\ \hline
         \arch & \Ft & 
-	\begin{minipage}[c]{7cm} 
-              \vskip 3mm
+	\begin{minipage}[c]{0.5\linewidth} 
+              \vskip 1mm
 	      The feature section of the \Csc, where the mathematical 
@@
 	      as the \arch section of the VHDL-AMS.    
-              \\
     	\end{minipage}\\ \hline

### 2009-08-02 papers/BON_VHDL/static.tex
types
text correction

@@
 	\begin{minipage}[c]{0.5\linewidth}
-                 \vskip 1mm
+                 \smallskip
     		VHDL-AMS allows hierarchical structure modeling, 
@@
                 declarations that serve a common purpose.
-
+                \smallskip
     	\end{minipage}\\ \hline
@@
 	\begin{minipage}[c]{0.5\linewidth} 
-                \vskip 1mm
+                \smallskip
 		An  subsistem which represent \duc.
-
     	\end{minipage}\\ \hline
@@
 	\begin{minipage}[c]{0.5\linewidth}
-	      \vskip 1mm	 
+	      \smallskip	 
               \Csc  describes \du. It can be identify with the primary unit (\ent) and secondary 
@@
 	      The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-
+              \smallskip
     	\end{minipage}\\ \hline
@@
 	\begin{minipage}[c]{0.5\linewidth} 
-              \vskip 1mm
+              \smallskip
 	      The feature section of the \Csc, where the mathematical 
@@
 	      as the \arch section of the VHDL-AMS.
+              \smallskip 
     	\end{minipage}\\ \hline

### 2009-08-04 papers/BON_VHDL/paper.tex
added Scalar types
added types

@@
 The \arch body represents description of the internal implementation 
-of an \ent including only process statements, which are collections 
+of an \ent including only process statements(encapsulation), which are collections 
 of actions to be executed in sequence, or simultaneous statements, 
@@
 \input{static}
-%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
+%\todo{dragan 27/07/2009 - Figure out :  static types.}
 \input{operators}
+%\todo{dragan 27/07/2009 - Figure out :  types/subtypes.}
+\input{types}
 %======================================================================

### 2009-08-04 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: added packet structure description

@@
 
-%\usepackage{todonotes}
-
-% \usepackage{html}
-% \usepackage{url}
 \usepackage{xspace}
-%\usepackage{doublespace}
 \usepackage{tabularx}
@@
 \usepackage{stmaryrd}
-% \ifpdf
-% \usepackage[centredisplay]{diagrams}
-% \else
-% \usepackage[centredisplay,PostScript=dvips]{diagrams}
-% \fi
 \usepackage{float}
@@
             linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
+            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan
+Stosic"]{hyperref}
 \else
@@
 \newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
-%\newcommand{\myhref}[2]{\emph{#2}}
+\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3 \myhref{#1}{#1}}}
+
+\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
+\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
+\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:} #1}}
+
 \newcommand{\todo}{\textbf{TODO:}}
+
 \newcommand{\ST}{\emph{SenseTile}\xspace}
+\newcommand{\STSB}{\ST \emph{Sensor Board}\xspace}
+\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
+\newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
+\newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
+\newcommand{\STU}{\ST \emph{Unit}\xspace}
+\newcommand{\STUs}{\ST \emph{Unit}\xspace}
+
 \newcommand{\STs}{\emph{SenseTiles}\xspace}
-\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
-\newcommand{\datastore}{\STs Scientific Datastore\xspace}
-\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
-\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
-\newcommand{\sensorfarm}{The \STs\xspace}
+
+\newcommand{\simulator}{\STSB \emph{Simulator}\xspace}
+
+\newcommand{\datastore}{\ST Scientific Datastore\xspace}
+\newcommand{\computefarm}{The \ST Scientific Compute Farm\xspace}
+\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute Server
+Farm\xspace}
+\newcommand{\sensorfarm}{The \STSBs\xspace}
 
@@
 
-  Sensor packages typically communicate (control sensors and report
-  sensor measurements) with a streaming packet-based network protocol.
-  The \ST device is an example of such a sensor package, as it
-  communicates with an arbitrary computer system (running Linux or
-  Windows, at the moment) over a USB bus using a streaming
-  packet-based protocol. 
-  This sensor platform is a custom-designed,
-  but inexpensive, sensor platform (called the \ST) paired with
-  general-purpose small-scale compute nodes, including everything from
-  low-power PDAs to powerful touchscreen-based portable computers.
-  There are over a dozen sensors on the \ST itself, and new sensors
-  can be added by adding them to the USB bus.
-  A third-party low-level
-  library, provided by the USB chip vendor, creates and parses
-  generic USB communication packets, while a custom medium-level Java
-  driver creates and parses \ST streaming communication packets. 
-  This
-  latter Java driver is formally specified, validated with rigorous
+  Sensor equipped devices typically communicate with a streaming packet-based 
+  network protocol.
+  The communication is focused on controlling the sensors and report the 
+  sensors measurements.
+  The \STSB is an example of such a device, it communicates with an arbitrary 
+  computer system, the \STPU (running Linux or Windows, at the moment) over 
+  a USB bus using a streaming packet-based protocol.
+  The \STU is a custom-designed, but inexpensive, sensor platform: it is 
+  composed by a one or more \STSBs paired with a general-purpose small-scale 
+  compute node, the \STPU (including everything from PDAs to powerful 
+  touchscreen-based portable computers).
+  There are over a dozen sensors on the \STSB itself, and new sensors can be 
+  added either using expansion slots on the \STSB, or adding them directly to 
+  the USB bus.
+  A third-party low-level library, provided by the USB chip vendor, the USB 
+  chip that has been used in the \STSB, creates and parses
+  generic USB communication packets.
+
+  A custom medium-level Java driver creates and parses the \STSB streaming 
+  communication packets. 
+  This latter Java driver is formally specified, validated with rigorous
   unit testing, and verified using extended static checking. 
@@
 to the USB bus.
+
+The \ST
 
@@
 
-\todo describe package
+The communication between \STSB and \STPU is realized through an Universal 
+Serial Bus (USB) interface.
+On top of the USB serial channels a proprietary protocol is used, specified by 
+FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI is the 
+manufacturer of the USB controller chip used in the \STSB, the 
+FT2232H\cite{ftdi_ft232h_2009}.
+
+The proprietary protocol enables the control and management of some of the 
+functionalities in FT232H chip (such as internal EEPROM read and write, reset 
+the chip, etc.) as well as the definition of the communication protocol and 
+properties to be used in the asynchronous channels (input and output channels) 
+provided by FT232H.
+These added functionalities are accessed via an Application Program Interface,
+the API is provided by a driver developed directly by FTDI, which is available 
+for various systems (Linux 32 and 64 bits, Mac OS, Windows)
+\cite{ftdi_d2xx_api_2009}.
+
+Finally, on top of the serial channels provided by the API, a custom protocol 
+has been developed to manage the entire \STSB; that is, read sensor data 
+observations, enable or disable sensors, configure onboard components 
+as the Field-Programmable Gate Array (FPGA), and the micro processor; the 
+protocol main function is to read sensor data.
+
+The protocol is an asynchronous fixed length packed based protocol.
+We will only examine the output sensor data packet, since it is the more used 
+and interesting one.
+
+The internal structure of the output sensor data packet reflects the board 
+capabilities and sensors.
+Basically a single packet has to accomodate various types of data:
+\begin{itemize}
+ \item Fat data rate streams: audio data, 48KHz or 96KHz.
+ \item Medium data rate streams: 5KHz.
+ \item Slow data rate streams: temperature, pression, accellerometer, etc.; all 
+ these sensor have a data rate far lower than 5 KHz.
+ \item Metadatas: metadatas on sensors and board state.
+\end{itemize}
+
+Even if we have to deal with data streams that have different data rates, the 
+internal structure of the package (as we will see) is strictly fixed. 
+This choice originates from a tradeoff between two constraints: the efficiency 
+of the trasmission (redundant data to be minimized) and the complexity to build 
+a package (complexity to be minimized).
+The complexity minimization constraint is simple to understand: it is the FPGA 
+on the \STSB that build the packet, but it has also to be used for various 
+other specific tasks. 
+Hence, it is important to maintain the building packet code on the FPGA as 
+simple and short as possible, to leave enough space for the code implementing 
+other tasks.
+The Output sensor data packet structure is showed in Table 
+\ref{tab:data_packet_structure}, the packet total length is 1024 bytes (512 16 
+bit words).
+
+\begin{table}[htbp]
+\caption{Output sensor data packet structure.}
+\label{tab:data_packet_structure}
+\begin{center}
+\begin{tabular}{|c|l|}\hline
+\textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
+0-1 & sentinel\\\hline
+2-7 & slow data\\\hline
+8-11 & metadata: board state\\\hline
+12-13 & reserved\\\hline
+14-15 & metadata: board state\\\hline
+16 & reserved\\\hline
+17-22 & subsequence 1\\\hline
+\ldots & \ldots\\\hline
+503-508 & subsequence 82\\\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+The subsequence accomodate data from the fast and medium data rate streams 
+(there are 4 fast data rate stream and 8 medium data rate stream channels). 
+There are 82 subsequences in a packet.
+The subsequence structure is shown in Table \ref{tab:subsequence_structure}.
+
+\begin{table}[htbp]
+\caption{Output sensor data packet subsequence structure.}
+\label{tab:subsequence_structure}
+\begin{center}
+\begin{tabular}{|c|l|}\hline
+\textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
+0 & metadata: subsequence description\\\hline
+1 & medium data rate data sample\\\hline
+2-5 & fast data rate data samples\\\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+It is clear that, looking at the structure described and at the data rates 
+frequencies, that there must be something subtle, to make all the figures to 
+add up.
+The subleties lies in two forms of ridundancies, for slow and mediumd data 
+rates:
+
+\begin{itemize}
+  \item Slow data rates: slow data can be found at the beginning of each 
+  packet, the data is simply repeated (that is, it is equal to the previous 
+  packet in the stream) when no new data is available.
+  \item Medium data rates: the 8 medium rate channels are multiplexed together 
+  into the medium data rate data sample of each subsequence, which channel the 
+  sample belongs to is a piece of info to be extracted from the subsequence 
+  metadata. Besides, the medium data rate data sample could contain no data at 
+  all, this info again is to be extracted from the metadata.
+\end{itemize}
 

### 2009-08-11 papers/ScenariosAndEventsAsTests/paper.tex
Title, short intro, and whiteboard photo for new paper.

--- 2009-08-02 papers/BON_VHDL/paper.tex
@@
 %
-% "Refining from the BON High-level Specification Language to the
-% VHDL Hardware Description Language" for UNKNOWN09.
-% Joseph R. Kiniry, Dragan Stosic
+% "Formal Refinement between High-level Specifications and Validation" for UNKNOWN09.
+% Joseph R. Kiniry
 % $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
@@
             linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
+            pdfauthor="Joseph R. Kiniry"]{hyperref}
 \else
@@
 \newcommand{\todo}{\textbf{TODO:}}
-\newcommand{\ST}{\emph{SenseTile}\xspace}
-\newcommand{\STs}{\emph{SenseTiles}\xspace}
-\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
-\newcommand{\datastore}{\STs Scientific Datastore\xspace}
-\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
-\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
-\newcommand{\sensorfarm}{The \STs\xspace}
-\newcommand{\Sc}{\emph{SYSTEM\_CHART}\xspace}
-\newcommand{\Clc}{\emph{CLUSTER\_CHART}\xspace}
-\newcommand{\Csc}{\emph{CLASS\_CHART}\xspace}
-\newcommand{\Ft}{\emph{feature}\xspace}
-\newcommand{\ent}{\emph{entity}\xspace}
-\newcommand{\arch}{\emph{architecture}\xspace}
-\newcommand{\archs}{\emph{architectures}\xspace}
-\newcommand{\duc}{\emph{design units collection}\xspace}
-\newcommand{\ducs}{\emph{design units collections}\xspace}
-\newcommand{\du}{\emph{design unit}\xspace}
-\newcommand{\undef}{\emph{undefined}\xspace}
+\newcommand{\name}{\textbf{IB2JML}}
 %---------------------------------------------------------------------
@@
 
-\title{Refining from the BON High-level Specification Language to the
-  VHDL-AMS Hardware Description Language}
+\title{Formal Refinement between\\
+High-level Specifications and Validation}
 
-\author{Joseph R. Kiniry and Dragan Stosic\\
+\author{Joseph R. Kiniry\\
 UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
 Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org and dragan.stosic@gmail.com\\
+kiniry@acm.org\\
 }
@@
 
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
-sensor system installed at the University College Dublin in Dublin,
-Ireland.  \textbf{Our facility provides a capability---unique in the world in
-terms of its scale and flexibility---for performing in-depth
-investigation of both the specific and general issues in large-scale
-sensor networking.}
-
-This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed, but inexpensive,
-sensor platform (called the \ST) paired with general-purpose
-small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers.  The datastore is a
-multi-terabyte scale scientific datastore into which sensor data
-flows, and in which online and offline scientific computation of
-sensor and other scientific data takes place.  The compute farm is a
-large-scale Linux, Solaris, and OS~X-based compute farm used by
-scientists for these online and offline scientific computations.
-
-The \STs, as an embedded system, have several unusual built-in sensors 
-and emitters, sensor nodes can be extended via a general-purpose USB bus, 
-and \STs contain an FPGA that is dynamically updated on a per-experiment 
-basis.  \STs are custom-designed in VHDL and VHDL-AMS.  Like many hardware-based systems,
- to write new software for or against the \ST architecture one either uses
-the actual hardware or one runs a \STS.  But this simulator emulates a
-piece of hardware, one would like to know if the simulator
-\emph{actually} simulates the hardware.  We have developed an approach
-through which we refine the VHDL specification up to the formal BON
-specification language, and from this specification we generate an
-executable, functional model-based specification in the JML language.
-We have used this specification to implement a fully validated and
-verified \ST simulator.
+  An abstract.
 
@@
 \section{Introduction}
-The \STs, represents heterogeneous collection of digital, analog 
-and mixed signal hardware components. Embedded system complexity 
-requires the need to strive for formal descriptions each component 
-separatelly in order to produce specification of the \STs simulator. 
-The development of complex system requires sophisticated tool and methodology 
-in order to be able to complete a correct design in time.  
-The authors will illustrate in this paper how BON can represent a high 
-level analogue mapping and will further illustrate a procedure for 
-transformation from VHDL-AMS to BON. As an extremly simple example, 
-we will give practical aspect of refinement from VHDL-AMS to BON,
-applied on A/D convertor, in order to produce formal specification 
-and software simulation. A complete list of the mapping between BON 
-and VHDL-AMS can be found in the appendix.
-%=====================================================================
 
-\section{VHDL-AMS}
-IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
-high-level description language for electronic and multi-domain 
-systems. VHDL-AMS, as an superset of VHDL, supports modeling and 
-simulation of analog and mixed-signal systems. The main contribution 
-of VHDL-AMS is ability to describe the behavior of complex system 
-trought differential and algebraic equations (DAEs). For this purpose, 
-VHDL-AMS introduces the following capabilities and therms:
+The typical requirements document for a complex software project
+consists of hundreds or thousands of labeled English sentences written
+in a Microsoft Word document.  There is little-to-no relationship
+between this document's contents and the architecture and
+implementation of the system they purport to describe.  Checking that
+a system conforms to the requirements is a labor-intensive, expensive,
+slow, and manual process performed by some poor engineer that drew the
+short straw.  Moreover, like the vast majority of architecture
+descriptions using languages like UML, the system and its description
+are not kept in sync, so as the implementation moves on, the
+requirements are tossed to the wayside.
 
-\begin{itemize}
-\item 
-\emph{Terminals} represents phisical connection points or circuit nodes 
-of a network connectivity. Terminals are declared to be of various 
-\emph{natures}, which represent different energy domains of a system.
-\item
-The \emph{simultaneous statements} creates one or more characteristic 
-equations that are used by the "analog solver" to model continuously 
-valued behaviour.  
-\item
-The \emph{quantities} denote a waveform or a time series of values.
-The behaviour of the \emph{quantities} is given by the \emph{simultaneous
- statements}. They take values as a result of solving the set of simultaneous 
-ordinary differential and algebraic equations (DAEs). To simplify, 
-quantities can be understood as analog unknowns of (DAEs) that specify 
-analog behavior. 
-\item
-The \emph{break statement} is an construct which allows to 
-initialize the DAE system at the beginning of a transient 
-simulation and to mark discontinuities during such a transient 
-simulation.
+In this paper we describe a a formal refinement between high-level
+specifications and a system's implementation via its validation.  The
+high-level specifications that we focus on include scenarios (aka
+use-cases), requirements, features (a la software product lines), and
+events, all of which are expressed in the Extended Business Object
+Notation (EBON) specification language.  Validation includes
+(sub)system, regression, and unit testing and static verification,
+using the Java Modeling Language (JML).  This refinement is expressed
+as a relation on types which is realizable as a adjoint pair of
+computations.  Consequently, one can use this formal refinement to (1)
+check that an implementation and its architecture conform to its
+high-level specification, (2) automatically update an architecture and
+implementation as their high-level specifications change, \emph{and}
+(3) automatically update high-level specifications as their
+architecture or implementation evolve.
 
-\end{itemize}
-
-\section{Choice of BON as the modeling tool}
-%Brief description and motivation.
-The Business Object Notation (BON) is an analysis and design notation 
-for object-oriented systems  which emphasizes \emph{seamlessness}, 
-\emph{reversibility} and software \emph{contracting}. Let us define 
-these principles precisely and point out why the principles are 
-important in VHDL-AMS simulation.
-%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
-\begin{itemize}
-\item \emph{Seamlessness} is the use of a continuous process 
-throughout the software lifecycle.
-%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
-\item \emph{Reversibility} is the support for both forward and 
-backward development process: from analysis to design 
-and implementation, and back.
-In VHDL-AMS, modelling represents an continious top-down-bottom-up
-process. Top-down approach serves for design while bottom-up
-serves for verification. The possibility of seamlessly translating 
-changes made during a certain development phase back into earlier phases,
-facilitates a stable and predictable environment and ensures the various 
-models consistent with the code.
-\item \emph{Contracting} is the precise definition, for each software element, 
-of the associated semantic properties. BON use a full assertion mechanism, 
-allowing analysts to specify not only the structure of a system but also its
-semantics ( constraints and invariants of the expected results ).
-\end{itemize}
-
-Therefore as an approach of best modelling practices, and as the notation 
-based on contract design and refinement calculus, BON is the obvious choice 
-for the our simulator.
-Throughout this paper we will try to make this clear.  
-\section{Mapping between BON and VHDL-AMS}
-
-\section{Refinement from BON to VHDL-AMS}
-
-An VHDL-AMS design units are described using an \ent-\arch structure. 
-The design unit contains an \ent and one or more \archs. 
-The \ent specifies the interface that an entity provides of itself 
-to the outside. In other words, \ent introduces a name, list of 
-input/output ports and the definition of the generic parameters. 
-The textual presentation of \ent is shown in Figure.1. 
-\input{adc-entity}
-
-The \arch body represents description of the internal implementation 
-of an \ent including only process statements, which are collections 
-of actions to be executed in sequence, or simultaneous statements, 
-which describe equations of analogue system behaviour. 
-The textual presentation of \arch is shown in Figure.2.
-\input{adc-arch}
-
-%\todo{dragan 27/07/2009 - Start with:1) BON charts and diagrams.}
+Our formal refinement is mechanically formally specified in
+higher-order logic in the PVS theorem prover and various appropriate
+metatheoretical properties about it are shown to hold.  Tool support
+for checking and updates is provided by the \name tool, an Eclipse
+plugin that realizes this formal refinement between EBON and JML.
 
@@
 
-%======================================================================
-\appendix
-\section {APPENDIX}
-%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
-\input{static}
-%\todo{dragan 27/07/2009 - Figure out :  VHDL-ASM as BON  represents an strongly typed language.}
-\input{operators}
-%======================================================================
-
 \end{document}

### 2009-08-19 papers/BON_VHDL/iarrayattr.tex
VHDL description

\begin{center}
\begin{tabular}{ c  c  c }
\multicolumn{3}{c}{\textbf{ATTRIBUTES OF ARRAY TYPES AND NATURES}}\\ 
\hline \hline
VHDL-AMS & BON & Description \\ 
\hline \hline
\begin{minipage}[c]{2.4cm} 
\smallskip \smallskip
\lstinlinenc|A'left(n)| \\
\lstinlinenc|A'right(n)|\\
\lstinlinenc|A'high(n)| \\
\lstinlinenc|A'low(n)| \\
\lstinlinenc|A'range(n)| \\
\lstinlinenc|A'reverse_range(n)| \\
\lstinlinenc|A'length(n)| \\
\lstinlinenc|A'ascending(n)| \\
\smallskip \smallskip
\end{minipage}  
&
\begin{minipage}[c]{4.6cm}
\centering 
\begin{lstlisting}[language=Bon]
deferred class IARRAY[T]
 feature -> n:INTEGER
  left:T -> n:INTEGER
  right:T-> n:INTEGER
  low:T  -> n:INTEGER
  high:T -> n:INTEGER
  range:INTEGER     
         -> n:INTEGER
  reverse_range:INTEGER 
         -> n:INTEGER
  length:INTEGER 
         -> n:INTEGER
  ascending:BOOLEAN 
         -> n:INTEGER
end
\end{lstlisting}
\smallskip \smallskip
\end{minipage}  
& 
\begin{minipage}[c]{6cm}  
\smallskip \smallskip
This class represents array nature interface 
with following features: leftmost value in index range of dimension n,\xspace 
rightmost value in index range of dimension n,\xspace 
least value in index range of dimension n,\xspace 
greatest value in index range of dimension n,\xspace
index range of dimension n,\xspace
index range of dimension n reversed in direction and bounds,\xspace
length of index range of dimension n,\xspace and
ordering.\xspace
\smallskip \smallskip
\end{minipage}\\ 
\end{tabular}
\end{center}

### 2009-08-19 papers/BON_VHDL/iquantity.tex
VHDL description

\begin{center}
\begin{tabular}{ c  c  c }
\multicolumn{3}{c}{\textbf{ATTRIBUTES OF QUANTITY}}\\ 
\hline \hline
VHDL-AMS & BON & Description \\ 
\hline \hline

\begin{minipage}[l]{2.4cm} 
\smallskip \smallskip 
\lstinlinenc|Q'tolerance|  \\
\lstinlinenc|Q'above(E)|   \\
\lstinlinenc|Q'delayed(t)| \\
\lstinlinenc|Q'dot|        \\
\lstinlinenc|Q'integ|      \\
\lstinlinenc|Q'ztf| \\
\lstinlinenc|(num,den,t| \\
\lstinlinenc|initial_delay)|       \\
\lstinlinenc|Q'ltf| \\
\lstinlinenc|(num,den)| \\
\lstinlinenc|Q'zoh| \\
\lstinlinenc|(t,initial_delay)|       \\
\lstinlinenc|Q'slew|       \\
\lstinlinenc|(max_rising_slope,| \\
\lstinlinenc|max_falling_slope)| \\
\smallskip \smallskip
\end{minipage}  
&
\begin{minipage}[c]{4.6cm}
\centering 
\smallskip \smallskip
\begin{lstlisting}[language=Bon]
deferred class IQUANTITY
 feature 
  tolerance:TOLERANCE
  above:BOOLEAN
   -> e:REAL
  delayed:REAL
   -> t:UNIVERSAL_TIME
  dot:FREE_QUANTITY 
  integ:FREE_QUANTITY
  ztf:FREE_QUANTITY
   -> num:SEQUENCE[REAL]
   -> den:SEQUENCE[REAL]
   -> t:UNIVERSAL_TIME
   -> initial_delay:UNIVERSAL_TIME 
  ltf:FREE_QUANTITY
   -> num:SEQUENCE[REAL]
   -> den:SEQUENCE[REAL] 
  zoh:FREE_QUANTITY
   -> t:UNIVERSAL_TIME
   -> initial_delay:UNIVERSAL_TIME  
  slew:FREE_QUANTITY
   -> max_rising_slope:REAL
   -> max_falling_slope:REAL
end
\end{lstlisting}
\smallskip \smallskip
\end{minipage}  
& 
\begin{minipage}[c]{6cm}  
\smallskip \smallskip
This calss represents quantity interface with following
attributes: \\
Tolerance group string presentation for scalar quantity.
Above attribute is true if the value of Q is grater than 
the value of e, otherwise false. \\
Delayed attribute represents value of quantity delayed by T.NOW-T.\\
Dot attribute represents quantity with the same type 
as quantity Q whose value is the \textbf{derivative} 
of Q with respect to time.\\
Integ attribute represents quantity with the same type 
as quantity Q whose value is the \textbf{integral} 
of Q with respect to time. \\
Ztf quantity represents \emph{z-domain} transfer function of Q.\\
Ltf quantity represents \emph{Laplace-domain} transfer function of Q. \\
Zoh quantity represents \emph{zero order hold} of Q.\\
Slow attribute represents quantity whose value follows that of Q, 
but whose derivative with respect to time is limited by the 
slopes.
\smallskip \smallskip
\end{minipage}\\ 
\end{tabular}
\end{center}

### 2009-08-19 papers/BON_VHDL/paper.tex
VHDL description

@@
 \usepackage{pdfsync}
+\usepackage{amsmath}
+\usepackage{listings}
 
@@
 
+\lstset{
+  language=Java,
+  basicstyle=\footnotesize\sffamily,
+  identifierstyle=\itshape,
+  stringstyle=\footnotesize\ttfamily,
+  commentstyle=\textup,
+  columns=flexible,
+  captionpos=b,
+  morekeywords={},
+  literate=
+  {-}{{$\textendash$}}1
+  {<}{{$\textless$}}1
+  {<=}{{$\textless=$}}1
+  {>}{{$\textgreater$}}2
+  {>=}{{$\textgreater=$}}2
+  {\\}{{$\backslash\backslash$}}1
+  {'}{{$\textprimstress$}}1
+}
+
+\def\lstinlinen{\lstinline[language=Bon,basicstyle=\normalsize\upshape]}
+
+\def\lstinlinenv{\lstinline[language=VHDL,basicstyle=\normalsize\upshape]}
+
+\def\lstinlinenc{\lstinline[language=VHDL,basicstyle=\normalsize\bfseries]}
+
+\def\lstinlinenb{\lstinline[language=Bon,basicstyle=\normalsize\bfseries]}
+
+\lstdefinelanguage{Bon} {
+  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result},
+  morekeywords={class,delta,invariant,inherit,deferred,effective,NONE,make},
+  morecomment=[l]{--}, 
+  morecomment=[s]{/*}{*/}, 
+  morestring=[b]", 
+ literate=,
+  showstringspaces=false
+}
+
+\lstdefinelanguage{VHDL} {
+  morekeywords={not,xnor,and,nand, nor,or,xor,integer,byte,real},
+  morekeywords={character,enum,type,boolean,bit,time,null,std_ulogic},
+  morekeywords={sll, srl, sla, sra, rol, ror},
+  morekeywords={rem,abs,mod},
+  morekeywords={terminal,T'reference, T'contribution,T'tolerance},
+  morekeywords={reference,contribution},
+  morekeywords={nature,subnature,electrical,magnetic,translational,translational_v},
+  morekeywords={rotational,rotational_v,fluidic,thermal,radiant},
+  morekeywords={voltage,angular_velocity,pressure,velocity},
+  morekeywords={current,force,torque,volumetric,flow_rate},
+  morekeywords={T'left,T'right,T'high,T'ascending,T'image(x),T'value(s)},
+  morekeywords={left,right,high,ascending,image,value},
+  morekeywords={T'pos(s),T'val(x),T'succ(x),T'pred(x)},
+  morekeywords={pos,val,succ,pred},
+  morekeywords={A'left(n),A'right(n),A'low(n),A'high(n)},
+  morekeywords={left,right,low,high},
+  morekeywords={A'range(n),A'reverse_range(n),A'length(n),A'ascending(n)},
+  morekeywords={range,reverse_range,length,ascending},
+  morekeywords={T'leftof(x),T'rightof(x),T'base},
+  morekeywords={leftof,rightof,base},
+  morekeywords={Q'tolerance, Q'above,Q'delayed,Q'dot,Q'integ},
+  morekeywords={delayed,dot,integ},
+  morekeywords={Q'slew,Q'ztf,Q'ltf,Q'zoh},
+  morekeywords={N'across, N'through},
+  morekeywords={across,through,zoh,ltf,slew,ztf},
+  morekeywords={above,tolerance,group,quantity},
+  morecomment=[l]{--}, 
+  morecomment=[s]{/*}{*/}, 
+  morestring=[b]", 
+ literate=,
+  showstringspaces=false
+}
+
 \newcommand{\tablesize}{\footnotesize}
@@
 \newcommand{\sensorfarm}{The \STs\xspace}
-\newcommand{\Sc}{\emph{SYSTEM\_CHART}\xspace}
-\newcommand{\Clc}{\emph{CLUSTER\_CHART}\xspace}
-\newcommand{\Csc}{\emph{CLASS\_CHART}\xspace}
-\newcommand{\Ft}{\emph{feature}\xspace}
 \newcommand{\ent}{\emph{entity}\xspace}
@@
 \newcommand{\archs}{\emph{architectures}\xspace}
-\newcommand{\duc}{\emph{design units collection}\xspace}
-\newcommand{\ducs}{\emph{design units collections}\xspace}
-\newcommand{\du}{\emph{design unit}\xspace}
+
 \newcommand{\undef}{\emph{undefined}\xspace}
+\newcommand{\arcoss}{\emph{across}\xspace}
+\newcommand{\through}{\emph{through}\xspace}
 %---------------------------------------------------------------------
@@
 \appendix
+\newpage
 \section {APPENDIX}
-%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
 \input{static}
-%\todo{dragan 27/07/2009 - Figure out :  static types.}
-\input{operators}
-%\todo{dragan 27/07/2009 - Figure out :  types/subtypes.}
+\input{arith}
+\input{shift}
+\input{relation}
+\input{logic}
 \input{types}
+\input{assignment}
+\input{iscaltypeattr}
+\input{tolerance}
+\input{effort}
+\input{flow}
+\input{nature}
+\input{iscalnature}
+\input{iarrayattr}
+\input{scalnat}
+\input{iterminal}
+\input{terminal}
+\input{iquantity}
+\input{quantity}
+\input{quantityport}
+\input{terminalport}
+\input{signal}
+\input{signalattr}
+\input{wait}
+%\input{quantityattr1}
+%\input{quantityattr2}
+%\input{quantityattr3}
 %======================================================================

### 2009-08-19 papers/BON_VHDL/static.tex
added more types
VHDL description

--- 2009-08-02 papers/BON_VHDL/static.tex
@@
+\newcommand{\duc}{\emph{design units collection}\xspace}
+\newcommand{\ducs}{\emph{design units collections}\xspace}
+\newcommand{\du}{\emph{design unit}\xspace}
 \begin{center}
-    \begin{tabular}{|l|l|l|}
-    	\hline
-    	VHDL-AMS & BON & Description \\ \hline
-    	\ducs (packages) & \Sc & 
-	\begin{minipage}[c]{7cm} 
-                 \vskip 3mm
+\begin{tabular}{ c  c  c }
+\multicolumn{3}{c}{\textbf{STATIC TYPES}}\\ 
+\hline \hline
+VHDL-AMS & BON & Description \\ 
+\hline \hline
+% first element
+\begin{minipage}[c]{2.4cm}
+\smallskip \smallskip
+\centering
+\ducs
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{4.6cm}
+\smallskip \smallskip
+\centering
+\lstinlinen|SYSTEM_CHART|  
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{6cm}
+         \smallskip \smallskip
 	VHDL-AMS allows hierarchical structure modeling, 
-    		meaning that subsistems (\ducs), can be developed independently.
-    		As an top-level description (informal/formal), \Sc connect all the \ducs
-    		together to a complete design. Actually \Sc can be understood 
-                as VHDL-AMS package as a way of grouping a collection of related 
-                declarations that serve a common purpose.
-                \\ 
-    	\end{minipage}\\ \hline
-        \duc & \Clc &
-	\begin{minipage}[c]{7cm} 
-                \vskip 3mm
+	meaning that subsistems (\ducs), can be developed 
+        independently. As an top-level description 
+        (informal/formal), \lstinlinen|SYSTEM_CHART|
+        connect all the \ducs together 
+        to a complete design. Actually \lstinlinen|SYSTEM_CHART|
+        can be understood as VHDL-AMS package as a way of grouping 
+        a collection of related declarations that serve a common 
+        purpose.
+        \smallskip \smallskip
+\end{minipage}\\
+
+% next element
+
+\begin{minipage}[c]{2.4cm}
+\smallskip \smallskip
+\centering
+\duc
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{4.6cm}
+\smallskip \smallskip
+\centering
+\lstinlinen|CLUSTER_CHART| 
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{6cm}
+      \smallskip \smallskip	 
       An  subsistem which represent \duc.
-                \\ 
-    	\end{minipage}\\ \hline
-	 \du & \Csc &
-	\begin{minipage}[c]{7cm}
-	      \vskip 3mm	 
-              \Csc  describes \du. It can be identify with the primary unit (\ent) and secondary 
-	      unit (\arch) structure in VHDL-AMS.
-	      The \Csc's name corresponds to the name of the entity in VHDL-AMS.
-              \\
-    	\end{minipage}\\ \hline
-        \arch & \Ft & 
-	\begin{minipage}[c]{7cm} 
-              \vskip 3mm
-	      The feature section of the \Csc, where the mathematical 
-	      transfer function of the component is entered, holds the same information 
-	      as the \arch section of the VHDL-AMS.
-              \\
-    	\end{minipage}\\ \hline
+      \smallskip \smallskip
+\end{minipage}\\
+
+% next element
+
+\begin{minipage}[c]{2.4cm}
+\smallskip \smallskip
+\centering
+\du
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{4.6cm}
+\smallskip \smallskip
+\centering
+\lstinlinen|CLASS_CHART| 
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{6cm}
+      \smallskip \smallskip	 
+      \lstinlinen|CLASS_CHART|  describes \du. 
+      It can be identify with the primary unit (\ent) 
+      and secondary unit (\arch) structure in VHDL-AMS. 
+      The \lstinlinen|CLASS_CHART|'s name corresponds 
+      to the name of the entity in VHDL-AMS.
+      \smallskip \smallskip
+\end{minipage}\\
+
+% next element
+
+\begin{minipage}[c]{2.4cm}
+\smallskip \smallskip
+\centering
+\arch
+\smallskip \smallskip
+\end{minipage}
+& 
+\begin{minipage}[c]{4.6cm}
+\smallskip \smallskip
+\centering
+\lstinlinen|feature|
+\smallskip \smallskip
+\end{minipage}
+&
+\begin{minipage}[c]{6cm} 
+      \smallskip \smallskip
+      The feature section of the \lstinlinen|CLASS_CHART|, 
+      where the mathematical transfer function of the component 
+      is entered, holds the same information as the \arch section 
+      of the VHDL-AMS.
+      \smallskip \smallskip 
+\end{minipage}\\
 

### 2009-08-20 papers/BON_VHDL/paper.tex
paperwork

@@
 The \emph{simultaneous statements} creates one or more characteristic 
-equations that are used by the "analog solver" to model continuously 
-valued behaviour.  
+equations that are solved by the dedicated simulation kernel a.k.a.
+"analog solver", to model continuously valued behaviour.  
 \item
@@
 semantics ( constraints and invariants of the expected results ). 
+In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
+continuous-time waveform. In other words, simulator evaluates quantities such 
+that the constraints specified by the simultaneous statements are satisfied. 
+We admit that the constraint of waveform is nothing else than \emph{invariant}.
+Another example of \emph{invariant} is conservation of energy law in which the summation
+of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
+around any closed path (loop) equals zero and the summation of
+through quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node (point) equals zero.
+Finally, any predefined (implicite) attribute in VHDL has defined \emph{pre(post)condition} trought
+input/output parameters constraints.
+
 \end{itemize}
@@
 Therefore as an approach of best modelling practices, and as the notation 
-based on contract design and refinement calculus, BON is the obvious choice 
+based on contract design, BON is the obvious choice 
 for the our simulator.
@@
 \input{wait}
-%\input{quantityattr1}
-%\input{quantityattr2}
-%\input{quantityattr3}
 %======================================================================

### 2009-08-20 papers/BON_VHDL/paper.tex
more text
text

@@
 through quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node (point) equals zero.
-Finally, any predefined (implicite) attribute in VHDL has defined \emph{pre(post)condition} trought
-input/output parameters constraints.
+Finally, any predefined (implicite) attribute in VHDL has defined \emph{pre}/\emph{(post)condition} trought
+input/output parameters and theirs constraints.
 
@@
 Throughout this paper we will try to make this clear.
-\section{Mapping between BON and VHDL-AMS}
 
-\section{Refinement from BON to VHDL-AMS}
+\section{Refinement from VHDL-AMS to BON }
 
@@
 \input{adc-arch}
+
+\section{Mapping between BON and VHDL-AMS}
 

### 2009-08-21 papers/BON_VHDL/paper.tex
more text

@@
 applied on A/D convertor, in order to produce formal specification 
-and software simulation. A complete list of the mapping between BON 
+and software simulation. A comprehensive list of the mapping between BON 
 and VHDL-AMS can be found in the appendix.
@@
 \emph{Terminals} represents phisical connection points or circuit nodes 
-of a network connectivity. Terminals are declared to be of various 
+of a network connectivity. They are declared to be of various 
 \emph{natures}, which represent different energy domains of a system.
+Terminals hold no values; their only role is to facilitate the formation
+of conservative equation sets that, in turn, constrain the associated 
+\emph{branch quatities}. 
 \item
@@
 quantities can be understood as analog unknowns of (DAEs) that specify 
-analog behavior. 
+analog behavior. There are three kinds of quantity in VHDL-AMS: free, 
+branch and source quantities. A \emph{free quantity} can be used in 
+signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
+systems. A \emph{source quantity} is used for frequency and noise modeling.
 \item
@@
 simulation and to mark discontinuities during such a transient 
-simulation.
-
+simulation. A break statement allows restart of the analog kernel,
+than overrides quantity values and executes concurrently with the analogue model. 
+For any concurrent break statement there is an associated process.
 \end{itemize}
@@
 important in VHDL-AMS simulation.
-%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
-\begin{itemize}
-\item \emph{Seamlessness} is the use of a continuous process 
+
+\subsection{Seamlessness and Reversibility}
+\emph{Seamlessness} is the use of a continuous process 
 throughout the software lifecycle.
-%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
-\item \emph{Reversibility} is the support for both forward and 
+\emph{Reversibility} is the support for both forward and 
 backward development process: from analysis to design 
@@
 facilitates a stable and predictable environment and ensures the various 
-models consistent with the code.
-\item \emph{Contracting} is the precise definition, for each software element, 
+models consistent with the code. Keeping in sync various models, documetation
+and specification from forward progress to backward and redesign, preclude 
+that design and implementation diverge to an inconsistent state.   
+\subsection{Contracting}
+\emph{Contracting} is the precise definition, for each software element, 
 of the associated semantic properties. BON use a full assertion mechanism, 
@@
 We admit that the constraint of waveform is nothing else than \emph{invariant}.
-Another example of \emph{invariant} is conservation of energy law in which the summation
-of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
-around any closed path (loop) equals zero and the summation of
-through quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node (point) equals zero.
-Finally, any predefined (implicite) attribute in VHDL has defined \emph{pre}/\emph{(post)condition} trought
-input/output parameters and theirs constraints.
-
-\end{itemize}
+Another example of \emph{invariant} is conservation of energy law in which 
+the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
+around any closed path (loop) equals zero and the summation of through 
+quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
+(point) equals zero. Finally, any predefined (implicite) attribute in VHDL 
+has defined \emph{pre}/\emph{post condition} trought input/output 
+parameters and theirs constraints.
 
@@
 The \arch body represents description of the internal implementation 
-of an \ent including only process statements(encapsulation), which are collections 
-of actions to be executed in sequence, or simultaneous statements, 
+of an \ent including only process statements(encapsulation), which 
+are collections  of actions to be executed in sequence, or simultaneous statements, 
 which describe equations of analogue system behaviour. 

### 2009-08-22 papers/BON_VHDL/paper.tex
text

@@
 around any closed path (loop) equals zero and the summation of through 
-quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
-(point) equals zero. Finally, any predefined (implicite) attribute in VHDL 
+quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node equals zero. 
+Moreover, any predefined (implicite) attribute in VHDL 
 has defined \emph{pre}/\emph{post condition} trought input/output 
-parameters and theirs constraints.
-
+parameters and theirs constraints. \\
 Therefore as an approach of best modelling practices, and as the notation 
@@
 
-%\todo{dragan 27/07/2009 - Start with:1) BON charts and diagrams.}
-
 %=====================================================================
@@
 A conclusion.
+
+\section {Acknowledgments}
+The research upon which this paper is based was caried out with
+sponsorship from Science Foundation Ireland via the 
+\textbf{UCD CASL SenseTile System} grant.
 

### 2009-08-23 papers/BON_VHDL/paper.tex
change sections
continious work

@@
 \lstdefinelanguage{Bon} {
-  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result},
+  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result,INTEGER},
   morekeywords={class,delta,invariant,inherit,deferred,effective,NONE,make},
@@
 and specification from forward progress to backward and redesign, preclude 
-that design and implementation diverge to an inconsistent state.   
+that design and implementation diverge to an inconsistent state. \emph{Seamlessness}
+also plays important role in relating behaviours at differing levels of abstraction. 
+For example, one may wish to relate the structural views of an A/D converter at the 
+level of individual bits and at the level of busses. Therefore capturing similarities 
+between the differing levels of abstractions far outweigh their inevitable differences, 
+making it possible to obtain a continuous process that ensures a direct mapping between 
+a formal specification and its software solution.
+
 \subsection{Contracting}
@@
 has defined \emph{pre}/\emph{post condition} trought input/output 
-parameters and theirs constraints. \\
+parameters and theirs constraints. Furthermore any \emph{subtype} represents restricted
+set of values from a base \emph{type}. The condition that determines which values are in the
+\emph{subtype} may be \emph{contracted} by specifying characteristic predicate. 
+For example, the \emph{subtype} \lstinlinenc|bit| of \lstinlinen|INTEGRER| may be defined as:\\
 Therefore as an approach of best modelling practices, and as the notation 
@@
 Throughout this paper we will try to make this clear.
+
+\section{Mapping between VHDL-AMS and BON}
 
@@
 \input{adc-arch}
-
-\section{Mapping between BON and VHDL-AMS}
 

### 2009-08-24 papers/BON_VHDL/paper.tex
added more in section 3

@@
 \lstdefinelanguage{Bon} {
-  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result,INTEGER},
-  morekeywords={class,delta,invariant,inherit,deferred,effective,NONE,make},
+  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result, Void},
+  morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
+  morekeywords={explanation,query,command,constraint},
   morecomment=[l]{--}, 
@@
 and specification from forward progress to backward and redesign, preclude 
-that design and implementation diverge to an inconsistent state. \emph{Seamlessness}
-also plays important role in relating behaviours at differing levels of abstraction. 
-For example, one may wish to relate the structural views of an A/D converter at the 
-level of individual bits and at the level of busses. Therefore capturing similarities 
-between the differing levels of abstractions far outweigh their inevitable differences, 
-making it possible to obtain a continuous process that ensures a direct mapping between 
-a formal specification and its software solution.
+that design and implementation diverge to an inconsistent state. 
+\emph{Seamlessness} also plays important role in relating behaviours at 
+differing levels of abstraction. For example, one may wish to relate the 
+structural views of an A/D converter at the level of individual bits and 
+at the level of busses. Therefore capturing similarities between the 
+differing levels of abstractions far outweigh their inevitable differences, 
+making it possible to obtain a continuous process that ensures a direct 
+mapping between a formal specification and its software solution.
 
@@
 In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
-continuous-time waveform. In other words, simulator evaluates quantities such 
+continuous-time waveform. In other words, analog kernel calculate quantities such 
 that the constraints specified by the simultaneous statements are satisfied. 
-We admit that the constraint of waveform is nothing else than \emph{invariant}.
+We will agree that the constraint of waveform is nothing else than \emph{invariant}.
 Another example of \emph{invariant} is conservation of energy law in which 
 the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
-around any closed path (loop) equals zero and the summation of through 
+around any loop equals zero and the summation of through 
 quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node equals zero. 
 Moreover, any predefined (implicite) attribute in VHDL 
-has defined \emph{pre}/\emph{post condition} trought input/output 
-parameters and theirs constraints. Furthermore any \emph{subtype} represents restricted
-set of values from a base \emph{type}. The condition that determines which values are in the
-\emph{subtype} may be \emph{contracted} by specifying characteristic predicate. 
-For example, the \emph{subtype} \lstinlinenc|bit| of \lstinlinen|INTEGRER| may be defined as:\\
-Therefore as an approach of best modelling practices, and as the notation 
-based on contract design, BON is the obvious choice 
-for the our simulator.
-Throughout this paper we will try to make this clear.
+has defined pre/post condition trought input/output 
+parameters and theirs constraints. Furthermore any subtype/subnature 
+represents restricted set of values from a base type/nature. 
+The condition that determines which values are in the subtype/subnature 
+may be contracted by specifying characteristic predicate. 
+For example, in typed $\lambda$ calculi, the subtype \lstinlinenc|bit| of 
+positive integer may be defined as \texttt{bit:=$\lambda$ n:nat.n < 2}. 
+The corresponding informal and formal BON transformation looks like:
+\input{bitbon}
+Therefore as an notition based on contract design, and as  methodology of 
+best modelling practices, BON is the obvious choice for the our simulator.
+Throughout this paper we will try to make this more clearly.
 

### 2009-08-24 papers/BON_VHDL/paper.tex
more text and description

@@
 may be contracted by specifying characteristic predicate. 
-For example, in typed $\lambda$ calculi, the subtype \lstinlinenc|bit| of 
-positive integer may be defined as \texttt{bit:=$\lambda$ n:nat.n < 2}. 
+For example, in typed $\lambda$ calculi, the VHDL-AMS subtype \lstinlinenc|bit| of 
+\lstinlinenc|integer|, may be defined as \texttt{bit:=$\lambda$ n:nat.n < 2}. 
 The corresponding informal and formal BON transformation looks like:

### 2009-08-24 papers/BON_VHDL/paper.tex
nothing specially
minor change

@@
 \subsection{Contracting}
-\emph{Contracting} is the precise definition, for each software element, 
+\emph{Contracting} is the definition, for each software element, 
 of the associated semantic properties. BON use a full assertion mechanism, 
@@
 represents restricted set of values from a base type/nature. 
-The condition that determines which values are in the subtype/subnature 
+The condition that determines which values are in the subtype/subnature,  
 may be contracted by specifying characteristic predicate. 

### 2009-08-24 papers/BON_VHDL/paper.tex
minor again
minor corrections

@@
 Moreover, any predefined (implicite) attribute in VHDL 
-has defined pre/post condition trought input/output 
-parameters and theirs constraints. Furthermore any subtype/subnature 
-represents restricted set of values from a base type/nature. 
-The condition that determines which values are in the subtype/subnature,  
+has defined pre(post) condition trought input(output) 
+parameters and theirs constraints. Furthermore any subtype(subnature) 
+represents restricted set of values from a base type(nature). 
+The condition that determines which values are in the subtype(subnature)  
 may be contracted by specifying characteristic predicate. 
@@
 \lstinlinenc|integer|, may be defined as \texttt{bit:=$\lambda$ n:nat.n < 2}. 
-The corresponding informal and formal BON transformation looks like:
+The corresponding informal and formal BON transformations are shown 
+by the example below:
 \input{bitbon}

### 2009-08-24 papers/BON_VHDL/paper.tex
minor correction
text correction

@@
 for object-oriented systems  which emphasizes \emph{seamlessness}, 
-\emph{reversibility} and software \emph{contracting}. Let us define 
-these principles precisely and point out why the principles are 
-important in VHDL-AMS simulation.
+\emph{reversibility},software \emph{contracting} and \emph{refinement}. 
+Let us define these principles precisely and point out why the principles 
+are important in VHDL-AMS simulation.
 
@@
 
-\subsection{Contracting}
+\subsection{Contracting and Refinement}
 \emph{Contracting} is the definition, for each software element, 
@@
 \input{bitbon}
+The current IEEE standard VHDL-AMS language reference manual 
+tries to define VHDL-AMS as well as possible in a descriptive way, 
+explaining the semantics in English. As we have shown in example above,
+BON has an easily-readable, English-based textual notation.\\
 Therefore as an notition based on contract design, and as  methodology of 

### 2009-08-24 papers/BON_VHDL/paper.tex
text correction

--- 2009-08-24 papers/BON_VHDL/paper.tex
@@
 \input{bitbon}
+Another advantage of BON as mapping tool is an easily-readable, 
+English-based textual notation. The current IEEE standard VHDL-AMS 
+language reference manual tries to define VHDL-AMS as well as possible 
+in a descriptive way, explaining the semantics in English.
 Therefore as an notition based on contract design, and as  methodology of 

### 2009-08-24 papers/BON_VHDL/paper.tex
text correction

--- 2009-08-24 papers/BON_VHDL/paper.tex
@@
 \input{bitbon}
-Therefore as an notition based on contract design, and as  methodology of 
-best modelling practices, BON is the obvious choice for the our simulator.
-Throughout this paper we will try to make this more clearly.
-
+Another advantage of BON is an easily-readable, English-based textual notation. 
+The current IEEE standard VHDL-AMS language reference manual tries to define 
+VHDL-AMS as well as possible in a descriptive way, explaining the 
+semantics in English.\\
+Therefore, on the basis of the above, BON is the obvious choice 
+for the our \ST simulator.
 \section{Mapping between VHDL-AMS and BON}

### 2009-08-24 papers/BON_VHDL/paper.tex
minor correction

@@
 Therefore, on the basis of the above, BON is the obvious choice 
-for the our \ST simulator.
+for the \ST simulator.
 \section{Mapping between VHDL-AMS and BON}

### 2009-08-26 papers/BON_VHDL/paper.tex
minor correction
added more..

@@
 that the constraints specified by the simultaneous statements are satisfied. 
-We will agree that the constraint of waveform is nothing else than \emph{invariant}.
+We will agree that the constraint of waveform is nothing more than \emph{invariant}.
 Another example of \emph{invariant} is conservation of energy law in which 
@@
 \section{Mapping between VHDL-AMS and BON}
-
+The main theme of this paper is the mapping between VHDL-AMS and BON.
 \section{Refinement from VHDL-AMS to BON }

### 2009-08-26 papers/BON_VHDL/paper.tex
refined paper

--- 2009-08-24 papers/BON_VHDL/paper.tex
@@
   morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
-  morekeywords={explanation,query,command,constraint},
+  morekeywords={explanation,query,command,constraint,exists},
   morecomment=[l]{--}, 
@@
 \newcommand{\archs}{\emph{architectures}\xspace}
-
+\newcommand{\inv}{\emph{invariant}\xspace}
 \newcommand{\undef}{\emph{undefined}\xspace}
@@
 and \STs contain an FPGA that is dynamically updated on a per-experiment 
-basis.  \STs are custom-designed in VHDL and VHDL-AMS.  Like many hardware-based systems,
+basis.  \STs are custom-designed in VHDL-AMS.  Like many hardware-based systems,
  to write new software for or against the \ST architecture one either uses
@@
 \section{Introduction}
-The \STs, represents heterogeneous collection of digital, analog 
-and mixed signal hardware components. Embedded system complexity 
+The \ST sensor platform, represents heterogeneous collection of digital, analog 
+and mixed signal hardware circuits. Embedded system complexity 
 requires the need to strive for formal descriptions each component 
-separatelly in order to produce specification of the \STs simulator. 
-The development of complex system requires sophisticated tool and methodology 
-in order to be able to complete a correct design in time.  
-The authors will illustrate in this paper how BON can represent a high 
+separatelly in order to produce specification of the \STS. 
+Our aim is to be able to precisely describe the structures of circuits,
+the behaviour of circuits and the formal link between structures and behaviours.
+In other words, the behavioural specification of the \STS may be 
+specified as a \inv on the signals at the circuit's ports and may be obtained 
+by taking the conjunction of the \emph{invariants} that describe 
+behaviours of the individual components separately. 
+For this purpose we need  sophisticated tool and methodology 
+in order to be able to achieve this goal.  
+The authors will illustrate in this article how BON can represent a high 
 level analogue mapping and will further illustrate a procedure for 
-transformation from VHDL-AMS to BON. As an extremly simple example, 
-we will give practical aspect of refinement from VHDL-AMS to BON,
-applied on A/D convertor, in order to produce formal specification 
-and software simulation. A comprehensive list of the mapping between BON 
-and VHDL-AMS can be found in the appendix.
+transformation from VHDL-AMS to BON.\\ 
+This article provides two main contributions.
+\begin{enumerate}
+\item 
+ It provides a solution how to formally emulate
+ piece of hardware using BON-JML approach; up to date BON and JML
+ has been used mostly in the pure object oriented software construction.    
+\item
+It relates hardware description code and formal calculus. 
+This means that if there  is a code that is "correct by construction", 
+then it will be derivable from informal BON specification, 
+via the refinement calculus to formal BON specification 
+and JML implementation.     
+\end{enumerate}
+As an extremly simple example, 
+we will give practical aspect of refinement, applied on real capacitor component, 
+in order to produce formal specification  and software simulation.\\  
+A comprehensive list of the mapping between BON and VHDL-AMS can be 
+found in the appendix.
 %=====================================================================
@@
 simulation of analog and mixed-signal systems. The main contribution 
-of VHDL-AMS is ability to describe the behavior of complex system 
+of VHDL-AMS is ability to describe the behaviour of complex system 
 trought differential and algebraic equations (DAEs). For this purpose, 
@@
 \item
-The \emph{simultaneous statements} creates one or more characteristic 
-equations that are solved by the dedicated simulation kernel a.k.a.
-"analog solver", to model continuously valued behaviour.  
-\item
 The \emph{quantities} denote a waveform or a time series of values.
@@
 quantities can be understood as analog unknowns of (DAEs) that specify 
-analog behavior. There are three kinds of quantity in VHDL-AMS: free, 
+analog behaviour. There are three kinds of quantity in VHDL-AMS: free, 
 branch and source quantities. A \emph{free quantity} can be used in 
@@
 systems. A \emph{source quantity} is used for frequency and noise modeling.
+\item
+The \emph{simultaneous statements} creates one or more characteristic 
+equations that are solved by the dedicated simulation kernel a.k.a.
+"analog solver", to model continuously valued behaviour.  
+\item 
+The \emph{procedural statements} or \emph{procedural} express analog behaviour 
+using sequential statements. Actually procedural describes analog behaviour 
+by computing new values for quantities from the current value of quantities.
 \item
@@
 The Business Object Notation (BON) is an analysis and design notation 
-for object-oriented systems  which emphasizes \emph{seamlessness}, 
-\emph{reversibility},software \emph{contracting} and \emph{refinement}. 
+for object-oriented systems  which emphasizes \emph{seamlessness},\xspace 
+\emph{reversibility} and software \emph{contracting}. 
 Let us define these principles precisely and point out why the principles 
@@
 that design and implementation diverge to an inconsistent state. 
-\emph{Seamlessness} also plays important role in relating behaviours at 
-differing levels of abstraction. For example, one may wish to relate the 
-structural views of an A/D converter at the level of individual bits and 
-at the level of busses. Therefore capturing similarities between the 
-differing levels of abstractions far outweigh their inevitable differences, 
-making it possible to obtain a continuous process that ensures a direct 
-mapping between a formal specification and its software solution.
+\emph{Seamlessness} also plays an important role in relating behaviours at
+different levels of abstraction. For example, one may wish to analyze
+a "real capacitor" at the level of individual components (a resistor and ideal capacitor)
+and their behaviours, or at the more abstract level through general behaviour (\inv) and
+using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
+Therefore capturing similarities between the differing levels of abstractions 
+far outweigh their inevitable differences, making it possible to obtain a 
+continuous process that ensures a direct mapping between a formal specification 
+and its software solution.
 
-\subsection{Contracting and Refinement}
+\subsection{Contracting}
 \emph{Contracting} is the definition, for each software element, 
@@
 that the constraints specified by the simultaneous statements are satisfied. 
-We will agree that the constraint of waveform is nothing else than \emph{invariant}.
-Another example of \emph{invariant} is conservation of energy law in which 
+We will agree that the constraint of waveform is nothing more than \inv.
+Another example of \inv is conservation of energy law in which 
 the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
 around any loop equals zero and the summation of through 
-quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node equals zero. 
-Moreover, any predefined (implicite) attribute in VHDL 
-has defined pre(post) condition trought input(output) 
-parameters and theirs constraints. Furthermore any subtype(subnature) 
-represents restricted set of values from a base type(nature). 
-The condition that determines which values are in the subtype(subnature)  
-may be contracted by specifying characteristic predicate. 
-For example, in typed $\lambda$ calculi, the VHDL-AMS subtype \lstinlinenc|bit| of 
-\lstinlinenc|integer|, may be defined as \texttt{bit:=$\lambda$ n:nat.n < 2}. 
-The corresponding informal and formal BON transformations are shown 
-by the example below:
-\input{bitbon}
+quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
+(\texttt{terminal}) equals zero. Moreover, any predefined (implicite) attribute in VHDL-AMS  
+has defined pre (post) condition trought input (output) 
+parameters and theirs constraints. Furthermore any subtype (subnature) 
+represents restricted set of values from a base type (nature). 
+The condition that determines which values are in the subtype (subnature)  
+may be contracted by specifying characteristic predicate which represents \inv.\\ 
 Another advantage of BON is an easily-readable, English-based textual notation. 
@@
 Therefore, on the basis of the above, BON is the obvious choice 
-for the \ST simulator.
+for the \STS.
+Throughput this article we will make this more clearly.  
 \section{Mapping between VHDL-AMS and BON}
-
+The main theme of this article is the mapping between VHDL-AMS and BON.
 \section{Refinement from VHDL-AMS to BON }
@@
 \section {Acknowledgments}
-The research upon which this paper is based was caried out with
+The research upon which this article is based was caried out with
 sponsorship from Science Foundation Ireland via the 

### 2009-08-26 papers/BON_VHDL/paper.tex
intro elaborated

@@
 \section{Introduction}
+\label{sec:introduction}
 The \ST sensor platform, represents heterogeneous collection of digital, analog 
@@
 This means that if there  is a code that is "correct by construction", 
-then it will be derivable from informal BON specification, 
-via the refinement calculus to formal BON specification 
-and JML implementation.     
+then it will be derivable from BON (informal/formal) specification, 
+via the refinement calculus to JML implementation.     
 \end{enumerate}
-As an extremly simple example, 
-we will give practical aspect of refinement, applied on real capacitor component, 
-in order to produce formal specification  and software simulation.\\  
+The rest of this article is organized as follows.
+In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
+\autoref{sec:choiceobon} defines  BON principles and explains the 
+reasons why we have decided to use BON method.
+The central part of the our article is \autoref{sec:mappingVB}, which 
+shows mapping between VHDL-AMS and BON. 
+In \autoref{sec:refVB} we will give practical aspect of refinement, 
+applied on real capacitor component, in order to produce formal specification  
+and software simulation.\\  
 A comprehensive list of the mapping between BON and VHDL-AMS can be 
@@
 \section{VHDL-AMS}
+\label{sec:vhdlams}
 IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
@@
 \section{Choice of BON as the modeling tool}
+\label{sec:choiceobon}
 %Brief description and motivation.
@@
 \subsection{Seamlessness and Reversibility}
+\label{subsec:seamrever}
 \emph{Seamlessness} is the use of a continuous process 
@@
 \subsection{Contracting}
+\label{subsec:contracting}
 \emph{Contracting} is the definition, for each software element, 
@@
 \section{Mapping between VHDL-AMS and BON}
+\label{sec:mappingVB}
 The main theme of this article is the mapping between VHDL-AMS and BON.
 \section{Refinement from VHDL-AMS to BON }
-
+\label{sec:refVB} 
 An VHDL-AMS design units are described using an \ent-\arch structure. 

### 2009-08-26 papers/BON_VHDL/paper.tex
added minor text

@@
 and their behaviours, or at the more abstract level through general behaviour (\inv) and
-using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
+using existential quantification (\texttt{exists}) to encapsulate all internal structure. 
 Therefore capturing similarities between the differing levels of abstractions 

### 2009-08-26 papers/BON_VHDL/paper.tex
correction

@@
 In other words, the behavioural specification of the \STS may be 
-specified as a \inv on the signals at the circuit's ports and may be obtained 
+represented as a \inv on the signals at the circuit's ports and may be obtained 
 by taking the conjunction of the \emph{invariants} that describe 

### 2009-08-27 papers/BON_VHDL/paper.tex
more ...

--- 2009-08-26 papers/BON_VHDL/paper.tex
@@
 requires the need to strive for formal descriptions each component
-separatelly in order to produce specification of the \STS. 
+separately in order to produce specification of the \STS.
 Our aim is to be able to precisely describe the structures of circuits,
-the behaviour of circuits and the formal link between structures and behaviours.
-In other words, the behavioural specification of the \STS may be 
-specified as a \inv on the signals at the circuit's ports and may be obtained 
+the behavior of circuits and the formal link between structures and behaviors.
+In other words, the behavioral specification of the \STS may be
+represented as a \inv on the signals at the circuit's ports and may be obtained
 by taking the conjunction of the \emph{invariants} that describe
-behaviours of the individual components separately. 
+behaviors of the individual components separately.
 For this purpose we need  sophisticated tool and methodology
@@
  piece of hardware using BON-JML approach; up to date BON and JML
- has been used mostly in the pure object oriented software construction.    
+ have been used mostly in the pure object oriented software construction.   
 \item
@@
 
-\subsection{Seamlessness and Reversibility}
-\label{subsec:seamrever}
-\emph{Seamlessness} is the use of a continuous process 
-throughout the software lifecycle.
-\emph{Reversibility} is the support for both forward and 
-backward development process: from analysis to design 
+\textbf{Seamlessness and  Reversibility} \xspace \xspace
+Seamlessness is the use of a continuous process
+throughout the software lifecycle. Reversibility is the support for
+both forward and backward development process: from analysis to design
 and implementation, and back.
@@
 and specification from forward progress to backward and redesign, preclude
-that design and implementation diverge to an inconsistent state. 
+that design and implementation diverge to an inconsistent state.\\
 \emph{Seamlessness} also plays an important role in relating behaviours at
 different levels of abstraction. For example, one may wish to analyze
-a "real capacitor" at the level of individual components (a resistor and ideal capacitor)
-and their behaviours, or at the more abstract level through general behaviour (\inv) and
-using existential quantification (\texttt{exists}) to encapsulate all internal signals. 
-Therefore capturing similarities between the differing levels of abstractions 
-far outweigh their inevitable differences, making it possible to obtain a 
-continuous process that ensures a direct mapping between a formal specification 
-and its software solution.
-
-\subsection{Contracting}
-\label{subsec:contracting}
-\emph{Contracting} is the definition, for each software element, 
-of the associated semantic properties. BON use a full assertion mechanism, 
-allowing analysts to specify not only the structure of a system but also its
-semantics ( constraints and invariants of the expected results ). 
-In VHDL-AMS \emph{simultaneous statement} represents analogue constraint of 
-continuous-time waveform. In other words, analog kernel calculate quantities such 
-that the constraints specified by the simultaneous statements are satisfied. 
-We will agree that the constraint of waveform is nothing more than \inv.
-Another example of \inv is conservation of energy law in which 
-the summation of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) 
-around any loop equals zero and the summation of through 
-quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node 
-(\texttt{terminal}) equals zero. Moreover, any predefined (implicite) attribute in VHDL-AMS  
-has defined pre (post) condition trought input (output) 
-parameters and theirs constraints. Furthermore any subtype (subnature) 
+a "real capacitor" at the level of individual components (a resistor and
+ideal capacitor) and their behaviours, or at the more abstract level through
+general behaviour (\inv) and using existential quantification (\texttt{exists})
+to encapsulate all internal structure. Therefore capturing similarities between
+the differing levels of abstractions far outweigh their inevitable differences,
+making it possible to obtain a continuous process that ensures a direct mapping
+between a formal specification and its software solution.\\
+\textbf{Contracting} \xspace \xspace Contracting is the definition, for each
+software element, of the associated semantic properties. BON use a full
+assertion mechanism, allowing analysts to specify not only the structure of
+a system but also its semantics ( constraints and invariants of the expected results ).
+The common way to specify a program statement is to give a \emph{precondition} and
+\emph{postcondition} on the state that the implementation must satisfy, assuming that
+only certain program variables are changed.
+In VHDL-AMS any predefined (implicite) attribute 
+has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
+parameters and theirs constraints. Furthermore, any subtype (subnature)
 represents restricted set of values from a base type (nature).
@@
 may be contracted by specifying characteristic predicate which represents \inv.\\
+Let us next look how \emph{simultaneous statements} are handled in contract design.
+The \emph{simultaneous statement} represents analogue constraint of
+continuous-time waveform. We will assume that the constraint of waveform
+represents \inv. This premise deserve an informal explanation.\\
+Briefly stated, analog kernel calculate quantities such
+that the constraints specified by the simultaneous statements are satisfied.
+More accurately, \emph{simultaneous statements} can be thought of as
+algebraic equations. It means that the number of simultaneous statements
+should be equal to the number of unknowns, i.e. \emph{free} and \emph{branch}
+\emph{quantities}. When we reasoning about conservation of energy law, for example,
+the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
+in a model.
+In other words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0})
+around any closed path must be equal to zero and the sum of \through
+quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
+(terminal) must be equal to zero. To be more precise, the results of summations 
+above remains zero when a particular transformation is applied to quantities.
+In general, they are \inv. \\
+However if our implementation does not satisfy conservation of energy law, 
+ then the invariant will be broken.
+In practice, one can use \emph{invariant} against the already existing code in
+order to check the correctness.
+It is important because we cannot patch a fabricated chip once company started
+massive production. \\
 Another advantage of BON is an easily-readable, English-based textual notation.

### 2009-08-31 papers/BON_VHDL/paper.tex
added more

@@
 \usepackage{a4wide}
-\usepackage{pdfsync}
+%\usepackage{pdfsync}
 \usepackage{amsmath}
@@
 \fi
+\ifpdf
+  \DeclareGraphicsRule{*}{mps}{*}{}
+\fi
 
@@
 }
+
+\def\sectionautorefname{Section}
 
@@
 transformation from VHDL-AMS to BON.\\
-This article provides two main contributions.
+This article provides three main contributions.
 \begin{enumerate}
@@
 reasons why we have decided to use BON method.
-The central part of the our article is \autoref{sec:mappingVB}, which
-shows mapping between VHDL-AMS and BON.
-In \autoref{sec:refVB} we will give practical aspect of refinement,
-applied on real capacitor component, in order to produce formal specification 
-and software simulation.\\ 
+In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
 A comprehensive list of the mapping between BON and VHDL-AMS can be
 found in the appendix. 
+The central part of the our article is \autoref{sec:refVB}, in which
+we will give practical aspect of refinement, applied on lowpass filter, 
+in order to produce formal specification and software simulation.\\ 
 %=====================================================================
@@
 simulation and to mark discontinuities during such a transient
-simulation. A break statement allows restart of the analog kernel,
+simulation.For any concurrent break statement there is an associated process.
+A break statement allows restart of the analog kernel,
 than overrides quantity values and executes concurrently with the analogue model.
-For any concurrent break statement there is an associated process.
+
 \end{itemize}
@@
 Let us define these principles precisely and point out why the principles
-are important in VHDL-AMS simulation.
-
+are important in VHDL-AMS simulation.\\
 \textbf{Seamlessness and  Reversibility} \xspace \xspace
@@
 \emph{Seamlessness} also plays an important role in relating behaviours at
-different levels of abstraction. For example, one may wish to analyze
+different levels of abstraction. For instance, one may wish to analyze
 a "real capacitor" at the level of individual components (a resistor and
@@
 should be equal to the number of unknowns, i.e. \emph{free} and \emph{branch}
-\emph{quantities}. When we reasoning about conservation of energy law, for example,
+\emph{quantities}. For instance, when we reasoning about conservation of energy law,
 the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
@@
 In general, they are \inv. \\
-However if our implementation does not satisfy conservation of energy law, 
+However, if our implementation does not satisfy conservation of energy law, 
  then the invariant will be broken. 
-In practice, one can use \emph{invariant} against the already existing code in
-order to check the correctness.
-It is important because we cannot patch a fabricated chip once company started
+Another conclusion may be deducted from the example given above.\\
+Intuitively \inv may be useful for testing the function of the hardware components.
+In practice, one can runs simulator against the component in order to check 
+the properties. Testing can then be done by checking whether the \inv hold or not.
+For example, the most common problem of the real components is the glich - an electrical pulse 
+of short duration that is usually the result of a design error.
+The number of glitches that occur can be known by the fact how many times per day, 
+the \inv was broken and reestablished. 
+It is important because we cannot patch a fabricated component once we started
 massive production. \\
@@
 \label{sec:refVB}
-An VHDL-AMS design units are described using an \ent-\arch structure.
-The design unit contains an \ent and one or more \archs.
-The \ent specifies the interface that an entity provides of itself
-to the outside. In other words, \ent introduces a name, list of
-input/output ports and the definition of the generic parameters.
-The textual presentation of \ent is shown in Figure.1.
-\input{adc-entity}
-
-The \arch body represents description of the internal implementation
-of an \ent including only process statements(encapsulation), which
-are collections  of actions to be executed in sequence, or simultaneous statements,
-which describe equations of analogue system behaviour.
-The textual presentation of \arch is shown in Figure.2.
-\input{adc-arch}
+\includegraphics{circ.1}
 

### 2009-09-04 papers/BON_VHDL/paper.tex
reorganized paper

@@
 \usepackage{float}
+\usepackage{subfig}
 
@@
 \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
+
 \fi
@@
 \fi
+\def\figureautorefname{Fig.}
 
@@
   morekeywords={above,tolerance,group,quantity},
+  morekeywords={entity,end,use,port,generic,is,library,all},
+  morekeywords={architecture,of,through,begin,constant},
   morecomment=[l]{--},
@@
 \emph{actually} simulates the hardware.  We have developed an approach
-through which we refine the VHDL specification up to the formal BON
+through which we refine the VHDL-AMS specification up to the formal BON
 specification language, and from this specification we generate an
@@
 separately in order to produce specification of the \STS.
-Our aim is to be able to precisely describe the structures of circuits,
+Our aim is to be able to describe the structures of circuits,
 the behavior of circuits and the formal link between structures and behaviors.
@@
 transformation from VHDL-AMS to BON.\\
-This article provides three main contributions.
+This article provides two main contributions.
 \begin{enumerate}
@@
 \item
-It relates hardware description code and formal calculus.
+It relates hardware description code and refinement.
 This means that if there  is a code that is "correct by construction",
-then it will be derivable from BON (informal/formal) specification,
-via the refinement calculus to JML implementation.
+then it will be derivable from BON (informal/formal) specification to 
+JML implementation.
 \end{enumerate}
@@
 The central part of the our article is \autoref{sec:refVB}, in which
-we will give practical aspect of refinement, applied on lowpass filter, 
-in order to produce formal specification and software simulation.\\ 
+we will give practical aspect of refinement , applied on 
+first order passive RC Low-Pass filter, in order to produce formal 
+specification and software simulation.\\ 
 %=====================================================================
@@
 simulation and to mark discontinuities during such a transient
-simulation.For any concurrent break statement there is an associated process.
+simulation. For any concurrent break statement there is an associated process.
 A break statement allows restart of the analog kernel,
@@
 different levels of abstraction. For instance, one may wish to analyze
-a "real capacitor" at the level of individual components (a resistor and
-ideal capacitor) and their behaviours, or at the more abstract level through
+a "lowpass-filter" at the level of individual components (a resistor and 
+capacitor) and their behaviours, or at the more abstract level through
 general behaviour (\inv) and using existential quantification (\texttt{exists})
@@
 \label{sec:mappingVB}
-The main theme of this article is the mapping between VHDL-AMS and BON.
+
 \section{Refinement from VHDL-AMS to BON }
 \label{sec:refVB}
-\includegraphics{circ.1}
+The main theme of this article is the translation from VHDL-AMS to BON using 
+refinement approach.
+We will refine the VHDL-AMS code up to the formal BON, and from this 
+specification we will generate an executable specification in the JML language.
+For this purpose, we will demonstrate our approach in BON-JML fashion trought 
+low-pass filter example.
+\input{lowpassRC}
+We can implement the last equation (see \autoref{fig:lowpassfig}) using the \texttt{'dot} 
+implicite attribute of quantity. For this purpose we introduce relation 
+\texttt{$s \rightarrow d/dt \rightarrow 'dot$} . The transformation looks as follows: \\
+\begin{equation}
+v_{in} = v_{out} + \tau_psv_{out} 
+   \begin{array}[b]{c}                                                                                                                                    
+      {\scriptscriptstyle s \rightarrow d/dt}  \\=                                                                                                                  
+   \end{array} 
+v_{out} + \tau_p\cfrac{dv_{out}}{dt}
+\begin{array}[b]{c}                                                                                                                                    
+      {\scriptscriptstyle d/dt \rightarrow } {\scriptscriptstyle 'dot} \\=                                                                                                                  
+   \end{array}
+  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
+\end{equation} 
+\input{lowpass}
+We have not yet said what specification and refinement are.
+Let us introduce these therms first.\\ 
+A specification of a program component is a contract that gives some constraints 
+on how the component is to behave but leaves freedom for the implementator to 
+decide how the actual behavior of the component is to be realized.
+Specification involves \emph{choice} between a number of alternatives; this 
+relationship is one-to-many. As we have shown in \autoref{tab:rescap}, corresponding 
+informal specifications represents high level of abstraction derived from VHDL-AMS code.
 
+A refinement represents the concept that can be understud as series of 
+correctness-preserving transformations from general perspective to 
+a precise level of detail. Accurately, if we compare two specifications,
+\texttt{S} and \texttt{S'}, then we can conclude that the \texttt{S} 
+is more abstract than \texttt{S'} if any condition that we can establish 
+with the first specification can also be established by second. 
+It means that "\texttt{S is \emph{refined} by S'}", 
+and we denote this  as \texttt{S $\rightsquigarrow$ S'}.
+Let say that the first informal BON specification is denoted by 
+\texttt{S} (see \autoref{tab:rescap}), and second, which is given 
+in Table2  is denoted by \texttt{S'}.
+In this case, the refinement relation represents an \emph{isomorphism} 
+and maps the following:
 %=====================================================================
@@
 %======================================================================
-\appendix
-\newpage
-\section {APPENDIX}
-\input{static}
-\input{arith}
-\input{shift}
-\input{relation}
-\input{logic}
-\input{types}
-\input{assignment}
-\input{iscaltypeattr}
-\input{tolerance}
-\input{effort}
-\input{flow}
-\input{nature}
-\input{iscalnature}
-\input{iarrayattr}
-\input{scalnat}
-\input{iterminal}
-\input{terminal}
-\input{iquantity}
-\input{quantity}
-\input{quantityport}
-\input{terminalport}
-\input{signal}
-\input{signalattr}
-\input{wait}
+%\appendix
+%\newpage
+%\section {APPENDIX}
+%\input{static}
+%\input{arith}
+%\input{shift}
+%\input{relation}
+%\input{logic}
+%\input{types}
+%\input{assignment}
+%\input{iscaltypeattr}
+%\input{tolerance}
+%\input{effort}
+%\input{flow}
+%\input{nature}
+%\input{iscalnature}
+%\input{iarrayattr}
+%\input{scalnat}
+%\input{iterminal}
+%\input{terminal}
+%\input{iquantity}
+%\input{quantity}
+%\input{quantityport}
+%\input{terminalport}
+%\input{signal}
+%\input{signalattr}
+%\input{wait}
+
 %======================================================================

### 2009-09-13 papers/BON_VHDL/paper.tex
finalisation

@@
 \newcommand{\sensorfarm}{The \STs\xspace}
-\newcommand{\ent}{\emph{entity}\xspace}
-\newcommand{\arch}{\emph{architecture}\xspace}
-\newcommand{\archs}{\emph{architectures}\xspace}
+\newcommand{\ent}{entity\xspace}
+\newcommand{\arch}{architecture\xspace}
+\newcommand{\archs}{architectures\xspace}
 \newcommand{\inv}{\emph{invariant}\xspace}
@@
 \newcommand{\through}{\emph{through}\xspace}
+
 %---------------------------------------------------------------------
@@
 
-\title{Refining from the BON High-level Specification Language to the
-  VHDL-AMS Hardware Description Language}
+\title{Refining from the VHDL-AMS Hardware Description Language 
+to the BON High-level Specification Language}
 
@@
 \label{sec:introduction}
-The \ST sensor platform, represents heterogeneous collection of digital, analog
-and mixed signal hardware circuits. Embedded system complexity
-requires the need to strive for formal descriptions each component
-separately in order to produce specification of the \STS.
-Our aim is to be able to describe the structures of circuits,
-the behavior of circuits and the formal link between structures and behaviors.
-In other words, the behavioral specification of the \STS may be
-represented as a \inv on the signals at the circuit's ports and may be obtained
-by taking the conjunction of the \emph{invariants} that describe
-behaviors of the individual components separately.
-For this purpose we need  sophisticated tool and methodology
-in order to be able to achieve this goal. 
+The \ST sensor platform, represents heterogeneous 
+collection of digital, analog and mixed signal hardware circuits. 
+The complexity of our platform requires the need to strive for formal 
+descriptions each component separately and  behavior of their 
+complex interdependencies, in order to produce executable specification.
+To make sure that the sensor platform works properly we need 
+to translate the VHDL-AMS code into the specification, and to refine and 
+verify this specification.
+In other words, our aim is to establish formal link between behavior of a given 
+sensor platform, and derived specification of components  in order to implement 
+fully validated and verified \ST simulator.
+For this purpose, we need notation and methodology 
+which will be able to satisfy the requirements given below.\\ 
+First, both of them, should be in sync with  VHDL-AMS language and reference 
+manual which tries to define VHDL-AMS as well as possible in a descriptive way,
+explaining the semantics in English.\\
+Second, specification derived from the notation should  be easy to attach to 
+the rest of the system architecture\xspace(driver, microprocessor emulator, 
+test-benches e.t.c.).\\
+Third, methodology must aid efforts on creation of target system 
+correctness criteria, such that are independent of implementation.
+It means that verification of simulator model should take place on specification 
+using reasoning techniques provided for algebraic formalism before coding starts.\\
+Fourth,a precise semantics should  provides the basis for
+automated forward-backward engineering.\\
+Managing these requirements, and insuring that the platform definition 
+(constraints) are preserved and verified through all levels of 
+abstraction is where one becomes aware of gaps in already 
+used design methodologies.
+In our opinion, the BON-JML approach is in line with requirements given above.
 The authors will illustrate in this article how BON can represent a high
-level analogue mapping and will further illustrate a procedure for
-transformation from VHDL-AMS to BON.\\
-This article provides two main contributions.
+level analogue mapping notation and will further illustrate a refinement 
+technique from VHDL-AMS to BON and JML.\\
+This article provides three main contributions.
 \begin{enumerate}
@@
 \item
-It relates hardware description code and refinement.
-This means that if there  is a code that is "correct by construction",
-then it will be derivable from BON (informal/formal) specification to 
-JML implementation.
+It relates hardware description code and refinement technique.
+It means that if a satisfactory VHDL-AMS code exists, then it can be simulated  from the 
+BON (informal/formal) specification, and JML implementation by an 
+appropriate sequence of refinements.
+\item
+It introduces \inv as a general concept of modeling hardware 
+components behavior and their interactions.
+To be more precisely, the behavioral specification of the \STS may 
+be represented as a \inv on the signals at the circuit's ports and 
+may be obtained by taking the conjunction of the \emph{invariants} 
+that describe behaviors of the individual components separately.\\ 
 \end{enumerate}
@@
 In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON. 
-A comprehensive list of the mapping between BON and VHDL-AMS can be
-found in the appendix. 
 The central part of the our article is \autoref{sec:refVB}, in which
@@
 first order passive RC Low-Pass filter, in order to produce formal 
-specification and software simulation.\\ 
+specification and software simulation.
+In \autoref{sec:conclusion} the conclusions are given,
+and directions of future studies are outlined. Finally, a comprehensive list 
+of the mapping between BON and VHDL-AMS can be found in the appendix. \\ 
 %=====================================================================
@@
 simulation of analog and mixed-signal systems. The main contribution
-of VHDL-AMS is ability to describe the behaviour of complex system
+of VHDL-AMS is ability to describe the behavior of complex system
 trought differential and algebraic equations (DAEs). For this purpose,
@@
 of conservative equation sets that, in turn, constrain the associated
-\emph{branch quatities}.
+\emph{branch qualities}.
 \item
 The \emph{quantities} denote a waveform or a time series of values.
-The behaviour of the \emph{quantities} is given by the \emph{simultaneous
+The behavior of the \emph{quantities} is given by the \emph{simultaneous
  statements}. They take values as a result of solving the set of simultaneous
@@
 equations that are solved by the dedicated simulation kernel a.k.a.
-"analog solver", to model continuously valued behaviour. 
+"analog solver", to model continuously valued behavior. 
 \item
@@
 The Business Object Notation (BON) is an analysis and design notation
-for object-oriented systems  which emphasizes \emph{seamlessness},\xspace
+for object-oriented systems  which brings \emph{seamlessness},\xspace
 \emph{reversibility} and software \emph{contracting}.
@@
 Seamlessness is the use of a continuous process
-throughout the software lifecycle. Reversibility is the support for
+throughout the software life-cycle. Reversibility is the support for
 both forward and backward development process: from analysis to design
 and implementation, and back.
-In VHDL-AMS, modelling represents an continious top-down-bottom-up
+In VHDL-AMS, modelling represents an continuous top-down-bottom-up
 process. Top-down approach serves for design while bottom-up
@@
 facilitates a stable and predictable environment and ensures the various
-models consistent with the code. Keeping in sync various models, documetation
+models consistent with the code. Keeping in sync various models, documentation
 and specification from forward progress to backward and redesign, preclude
 that design and implementation diverge to an inconsistent state.\\
-\emph{Seamlessness} also plays an important role in relating behaviours at
+\emph{Seamlessness} also plays an important role in relating behaviors at
 different levels of abstraction. For instance, one may wish to analyze
@@
 capacitor) and their behaviours, or at the more abstract level through
-general behaviour (\inv) and using existential quantification (\texttt{exists})
-to encapsulate all internal structure. Therefore capturing similarities between
-the differing levels of abstractions far outweigh their inevitable differences,
+general behaviour (\inv).
+ Therefore, capturing similarities between the differing levels of abstractions,
 making it possible to obtain a continuous process that ensures a direct mapping
 between a formal specification and its software solution.\\
-\textbf{Contracting} \xspace \xspace Contracting is the definition, for each
-software element, of the associated semantic properties. BON use a full
-assertion mechanism, allowing analysts to specify not only the structure of
-a system but also its semantics ( constraints and invariants of the expected results ).
+\textbf{Contracting} \xspace \xspace Contract specification describe 
+system components interaction, based on mutual obligations and benefits, 
+and, for each software element, associated semantic properties. BON use a full
+assertion mechanism, allowing analysts to specify both:\xspace (1) the \emph{structure} of
+a system and (2)\xspace \emph{semantics} ( constraints and invariants).
 The common way to specify a program statement is to give a \emph{precondition} and
@@
 only certain program variables are changed.
-In VHDL-AMS any predefined (implicite) attribute 
+In VHDL-AMS any predefined (implicate) attribute 
 has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
@@
 algebraic equations. It means that the number of simultaneous statements
-should be equal to the number of unknowns, i.e. \emph{free} and \emph{branch}
-\emph{quantities}. For instance, when we reasoning about conservation of energy law,
+should be equal to the number of unknowns (\emph{quantities}). 
+For instance, when we reasoning about the conservation of energy law,
 the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
@@
 In general, they are \inv. \\
-However, if our implementation does not satisfy conservation of energy law, 
+However, if our implementation does not satisfy the conservation of energy law, 
  then the invariant will be broken. 
@@
 The number of glitches that occur can be known by the fact how many times per day, 
-the \inv was broken and reestablished. 
-It is important because we cannot patch a fabricated component once we started
-massive production. \\
-Another advantage of BON is an easily-readable, English-based textual notation.
-The current IEEE standard VHDL-AMS language reference manual tries to define
-VHDL-AMS as well as possible in a descriptive way, explaining the
-semantics in English.\\
+the \inv was broken and reestablished. According to that we can conclude whether the 
+quality of a component is satisfactory. Unlike in the case of software, where 
+correction of an error is relatively inexpensive, an detected hardware error 
+may require product replacement, which is associated with great expenses. 
+In other words, we cannot patch a fabricated component once we started massive production. \\
 Therefore, on the basis of the above, BON is the obvious choice
@@
 \label{sec:mappingVB}
+In this section, we briefly describe the mapping between VHDL-AMS and BON notation 
+which is discussed in the points below.\\
+\textbf{Structure} \xspace \xspace VHDL-AMS allows hierarchical structure 
+modeling, meaning that the model subsystems can be developed independently.
+As an top-level description, \lstinlinen|SYSTEM_CHART| in the BON notation  
+can connect all the model subsystems  together to a model design.
+In BON, model subsystem is denoted as the \lstinlinen|CLUSTER_CHART| whereas 
+module is represented as the \lstinlinen|CLASS_CHART|.
+Each module has a \texttt{port} interface: a set of \emph{inputs} and \emph{outputs} used for 
+connecting modules with environment. The \texttt{port} section in 
+\texttt{\ent}structure (input, output and generic values) are represented as 
+a \emph{features} in the \lstinlinen|CLASS_CHART|.
+The \lstinlinen|CLASS_CHART| \xspace can be equated with \ent-\arch 
+structure in VHDL-AMS and the name of the \lstinlinen|CLASS_CHART|\xspace 
+corresponds to the name of \texttt{\ent}in VHDL-AMS.\\ 
+\textbf{Behavior} \xspace \xspace The module also contains an 
+\texttt{\arch} implementation, which determines the way the module 
+processes \emph{input}/\emph{output} signals and performs computations.
+The \emph{constraint} section in the \lstinlinen|CLASS_CHART| 
+contains the transfer function and holds the same information as 
+the \texttt{\arch}section of the VHDL-AMS.\\ 
+\textbf{Types and operators} \xspace \xspace There are many features of 
+VHDL-AMS which are already mapped in BON by default, such as arithmetic, 
+relational and logical operators and types like \xspace \texttt{integer},
+\xspace \texttt{byte},\xspace\texttt{real},\xspace\texttt{character} and 
+\xspace\texttt{boolean}.
+Vector may be represented as \lstinlinen|SEQUENCE[TYPE]|, 
+whereas record, whose elements are of heterogeneous types, can be mapped 
+as \lstinlinen|SET| or \lstinlinen|TABLE|.
+In BON, the \texttt{time} can be represented as the
+\lstinlinen|UNIVERSAL_TIME|, in order to specify \emph{delays}. Actually 
+time can be thought of as object with integer as a value and measurement 
+units as a \xspace \lstinlinen|SEQUENCE| feature.     
+In order to specify energy domain for nodes of circuits we have developed 
+various standard \texttt{nature} templates for \texttt{terminals}.
+We have introduced \xspace\lstinlinen|FREE|,\xspace \lstinlinen|BRANCH|
+and  \xspace\lstinlinen|SOURCE| \space \lstinlinen|QUANTITY| objects, 
+which are intended to describe signal-flow modeling, the 
+conservation of energy low and frequency-noise related problems. 
+There are also defined specifications for effort and 
+flow domains, tolerance as well as appropriate contracts for 
+scalar types(natures), array types(natures), 
+terminals and quantity attributes.
+Whereas the value assignment, in VHDL-AMS has the same purpose as in the BON,
+the signal assignment, which schedules a new value to be applied to a 
+signal later, is simulated in specification, but  denotational semantic
+is not yet translated in the BON as well as for concurrency and 
+the \texttt{wait} statement.
 
@@
 \input{lowpassRC}
-We can implement the last equation (see \autoref{fig:lowpassfig}) using the \texttt{'dot} 
-implicite attribute of quantity. For this purpose we introduce relation 
-\texttt{$s \rightarrow d/dt \rightarrow 'dot$} . The transformation looks as follows: \\
+According to equation given above (see \autoref{fig:lowpassfig}) 
+we can produce VHDL-AMS \emph{simultaneous statement} using 
+\texttt{$s \rightarrow d/dt \rightarrow 'dot$} relation.
+The transformation looks as follows: \\
 \begin{equation}
@@
 We have not yet said what specification and refinement are.
-Let us introduce these therms first.\\ 
+Let us introduce these terms first.\\ 
 A specification of a program component is a contract that gives some constraints 
@@
 Let say that the first informal BON specification is denoted by 
-\texttt{S} (see \autoref{tab:rescap}), and second, which is given 
-in Table2  is denoted by \texttt{S'}.
+\texttt{S} (see \autoref{tab:rescap}), and second, which is given in Table2  is denoted by \texttt{S'}.
 In this case, the refinement relation represents an \emph{isomorphism} 
@@
 \section{Conclusion}
-
-A conclusion.
+\label{sec:conclusion}
+The BON notation has originally been used for description of 
+software models. However, after certain elaboration it can be used 
+for description of hardware components. Practical utilization of the approach
+in the project on mapping between VHDL-AMS and BON using refinement technique,
+has demonstrated through low-pass filter example and shown that simulation
+of component requires relatively little and inexpensive  efforts. Moreover, our article shown
+that the development of the specification produced by the BON and translated 
+into JML can be applied as refinement procedure: from VHDL-AMS to informal BON, 
+next, from informal to more formal BON and finally, to JML specification. 
+The BON support general software development and should be adapted for concrete
+applications and language as VHDL-AMS. For this purpose we have developed various 
+VHDL-AMS related types, operators and relations and we have defined appropriate 
+specifications for all of them. However, the behavior of
+\texttt{break} statement and  concurrency, for example, have not been translated 
+yet in the BON notation. We have started a new phase of project named \emph{eBon} in which
+the issues mentioned above, and creation of appropriate components and relations 
+are main subject.
 
@@
 %======================================================================
-%\appendix
-%\newpage
-%\section {APPENDIX}
-%\input{static}
-%\input{arith}
-%\input{shift}
-%\input{relation}
-%\input{logic}
-%\input{types}
-%\input{assignment}
-%\input{iscaltypeattr}
-%\input{tolerance}
-%\input{effort}
-%\input{flow}
-%\input{nature}
-%\input{iscalnature}
-%\input{iarrayattr}
-%\input{scalnat}
-%\input{iterminal}
-%\input{terminal}
-%\input{iquantity}
-%\input{quantity}
-%\input{quantityport}
-%\input{terminalport}
-%\input{signal}
-%\input{signalattr}
-%\input{wait}
-
+\appendix
+\newpage
+\section {APPENDIX}
+\input{static}
+\input{arith}
+\input{shift}
+\input{relation}
+\input{logic}
+\input{types}
+\input{assignment}
+\input{iscaltypeattr}
+\input{tolerance}
+\input{effort}
+\input{flow}
+\input{nature}
+\input{iscalnature}
+\input{iarrayattr}
+\input{scalnat}
+\input{iterminal}
+\input{terminal}
+\input{iquantity}
+\input{quantity}
+\input{quantityport}
+\input{terminalport}
+\input{signal}
+\input{signalattr}
+\input{wait}
+\input{bitbon}
 %======================================================================

### 2009-09-13 papers/BON_VHDL/static.tex
finalisation

@@
-\newcommand{\duc}{\emph{design units collection}\xspace}
-\newcommand{\ducs}{\emph{design units collections}\xspace}
-\newcommand{\du}{\emph{design unit}\xspace}
 \begin{center}
@@
 \centering
-\ducs
+\emph{model}
 \smallskip \smallskip
@@
          \smallskip \smallskip
-	VHDL-AMS allows hierarchical structure modeling, 
-	meaning that subsistems (\ducs), can be developed 
-        independently. As an top-level description 
+	As an top-level description 
         (informal/formal), \lstinlinen|SYSTEM_CHART|
-        connect all the \ducs together 
-        to a complete design. Actually \lstinlinen|SYSTEM_CHART|
-        can be understood as VHDL-AMS package as a way of grouping 
-        a collection of related declarations that serve a common 
-        purpose.
+        connect all the model subsystems together 
+        to a model.
         \smallskip \smallskip
@@
 \centering
-\duc
+model subsystem 
 \smallskip \smallskip
@@
       \smallskip \smallskip	 
-      An  subsistem which represent \duc.
+      An  subsistem which represent module colleaction.
+        VHDL-AMS allows hierarchical structure modeling, 
+	meaning that subsystems (modules), can be developed 
+        independently.
       \smallskip \smallskip
@@
 \centering
-\du
+modul
 \smallskip \smallskip
@@
       \smallskip \smallskip	 
-      \lstinlinen|CLASS_CHART|  describes \du. 
+      \lstinlinen|CLASS_CHART|  describes module. 
       It can be identify with the primary unit (\ent) 

### 2009-09-15 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol: modified introduction added parts

@@
 JUnit test suite applied to developed system, to verify its correctness.
-Junit test suite applied to simulator, to verify its correctness.
+JUnit test suite applied to simulator, to verify its correctness.
 
 RESULTS: 
-Proof developed system same behaviour as simulator.
+Proof developed system same behavior as simulator.
 Proof developed system and simulator verify the specifications.
@@
 The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
-sensor system installed at the University College Dublin in Dublin,
-Ireland.  \textbf{Our facility provides a capability---unique in the world in
-terms of its scale and flexibility---for performing in-depth
-investigation of both the specific and general issues in large-scale
-sensor networking.}
+sensor system to be installed at the University College Dublin in Dublin,
+Ireland. 
+Our facility aims to provide a scalable and extensible 
+infrastructure for performing in-depth investigation of both the specific 
+and general issues in large-scale sensor networking.
 
 This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed, but inexpensive,
+farm.  
+The sensor platform is a custom-designed, but inexpensive,
 sensor platform (called the \ST) paired with general-purpose
 small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers.  There are over a dozen
+powerful touchscreen-based portable computers. 
+There are over a dozen
 sensors on the \ST itself, and new sensors can be added by adding them
-to the USB bus.
+to the board.
 
-The \ST
+Because of hard time constraints, and the initial unavailability of the 
+custom board, we needed to progress concurrently with all the development tasks: 
+the specification of the communication protocol with the /SB, the development of 
+the controlling code in the /SB, the development of the driver to communicate 
+with the /SB, and the development of the simulators needed for future tests 
+of the system (the controlling code of the board was developed by the external 
+manufacturer of the board, hence, it is not taken into account here).
 
-\todo extend
+\todo explain and refine the problem
+
+\todo other approaches to the problem
+
+To tackle the complexity of the concurrent tasks the formal model of the protocol 
+was used as a focus point that lead all the other tasks. 
+The formal model of the protocol was used to check at run time the correctness of 
+the /SB, the driver, the simulator, and the hand made test suites of the driver 
+and the simulator.
+The formal model was used to automatically generate a test suite to verify that
+both the driver and the simulator were compliant to the specification, and eventually 
+to verify the specification itself.
+
+The approach proved itself effective, at the point that, when the board was finally 
+released, no defects were found in the driver or the simulator, while the errors and 
+not agreed changes introduced in the board where discovered during the first execution.
 
@@
 In section \ref{sec:data_stream_and_protocol_model} the data stream model 
-and protocol is presented, we will give both a higl level overview of the 
-protocol, and a formalized, JML based, specification,
+and protocol is presented, we will present both a high level overview of the 
+protocol, and a formalized, JML based, specification.
 
@@
 both the hand made and the generated ones. 
-Both the approaches are analyzed, their pro and contro are discussed, as 
-well as the expected advantages to use them in combination.
-
-In section \ref{sec:test_cases} the test cases structure is presented, for 
-both the hand made and the generated ones. 
-Both the approaches are analyzed, their pro and contro are discussed, as 
+Both the approaches are analyzed, their pro and contra are discussed, as 
 well as the expected advantages to use them in combination.
@@
 
-Finally in section \ref{sec:conclusions} we outline some conclusions and we 
-briefly describe our ongoing and future works on similar topics.
+Finally in section \ref{sec:conclusions} conclusions are drawed and 
+ongoing and future works on similar topics is introduced.
 
@@
 has been developed to manage the entire \STSB; that is, read sensor data 
-observations, enable or disable sensors, configure onboard components 
+observations, enable or disable sensors, configure on-board components 
 as the Field-Programmable Gate Array (FPGA), and the micro processor; the 
@@
 capabilities and sensors.
-Basically a single packet has to accomodate various types of data:
+Basically a single packet has to accommodate various types of data:
 \begin{itemize}
- \item Fat data rate streams: audio data, 48KHz or 96KHz.
+ \item Fast data rate streams: audio data, 48KHz or 96KHz.
  \item Medium data rate streams: 5KHz.
- \item Slow data rate streams: temperature, pression, accellerometer, etc.; all 
+ \item Slow data rate streams: temperature, pression, accelerometer, etc.; all 
  these sensor have a data rate far lower than 5 KHz.
@@
 internal structure of the package (as we will see) is strictly fixed. 
-This choice originates from a tradeoff between two constraints: the efficiency 
-of the trasmission (redundant data to be minimized) and the complexity to build 
+This choice originates from a trade-off between two constraints: the efficiency 
+of the transmission (redundant data to be minimized) and the complexity to build 
 a package (complexity to be minimized).
@@
 
-The subsequence accomodate data from the fast and medium data rate streams 
+The subsequence accommodate data from the fast and medium data rate streams 
 (there are 4 fast data rate stream and 8 medium data rate stream channels). 
@@
 
-It is clear that, looking at the structure described and at the data rates 
-frequencies, that there must be something subtle, to make all the figures to 
+Looking at the structure described and at the data rates frequencies, it
+can be seen that there must be something subtle, to make all the figures to 
 add up.
-The subleties lies in two forms of ridundancies, for slow and mediumd data 
+The subtleties lies in two forms of redundancies, for slow and medium data 
 rates:
@@
 details remains, exact package structure and sensor data location in the
-packege are lost
+package are lost
 
-\todo describe JML specification of the high levelr representation of the 
+\todo describe JML specification of the high level representation of the 
 package
@@
 
-\todo benefit: the approch can be used to test different part of the system
+\todo benefit: the approach can be used to test different part of the system
 that share the same specification

### 2009-09-17 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol: specification initial description

@@
   communication packets. 
-  This latter Java driver is formally specified, validated with rigorous
-  unit testing, and verified using extended static checking. 
+  The driver is formally specified, validated and verified with rigorous
+  unit testing. 
   Because of this process and combination of technologies, using this
@@
   Java Modeling Language.
-  \todo executable? how? what does this mean?
+  \notev{executable? how? what does this mean?}
   This protocol model is itself an executable
@@
 custom board, we needed to progress concurrently with all the development tasks: 
-the specification of the communication protocol with the /SB, the development of 
-the controlling code in the /SB, the development of the driver to communicate 
-with the /SB, and the development of the simulators needed for future tests 
+the specification of the communication protocol with the \STSB, the development of 
+the controlling code in the \STSB, the development of the driver to communicate 
+with the \STSB, and the development of the simulators needed for future tests 
 of the system (the controlling code of the board was developed by the external 
@@
 To tackle the complexity of the concurrent tasks the formal model of the protocol 
-was used as a focus point that lead all the other tasks. 
+was used as a focus point that led all the other development tasks. 
 The formal model of the protocol was used to check at run time the correctness of 
@@
 
-In section \ref{sec:data_stream_and_protocol_model} the data stream model 
-and protocol is presented, we will present both a high level overview of the 
-protocol, and a formalized, JML based, specification.
+In section \ref{sec:data_stream_and_protocol_description} the data stream model 
+and protocol is presented.
+
+In section \ref{sec:data_stream_and_protocol_specification} the data stream model 
+and protocol is specified.
 
@@
 
-Finally in section \ref{sec:conclusions} conclusions are drawed and 
+Finally in section \ref{sec:conclusion} conclusions are drawed and 
 ongoing and future works on similar topics is introduced.
@@
 %=====================================================================
-\section{Data stream and protocol model}
-\label{sec:data_stream_and_protocol_model}
+\section{Data stream and protocol description}
+\label{sec:data_stream_and_protocol_description}
 
@@
 The protocol is an asynchronous fixed length packed based protocol.
+The protocol has been divided in multiple layers to ease its formal description.
+The rationale is the following: mixing up different abstraction levels of a 
+protocol description can lead to overly complex and unmanageable descriptions, 
+expecially when dealing with formal ones.
+The layers identified in protocol are the following:
+
+\begin{description}
+ \item[Packet byte structure]: the internal representation of the packet.
+ \item[Packet info structure]: the meaningful fields contained in the packets.
+ \item[Single packet rules]: the content acceptable values, and how they 
+influence each other.
+ \item[Packet sequence rules]: the content acceptable values, based on values 
+of previously received packets.
+ \item[Packet input output asynchronous rules]: the content acceptable values 
+and reaction constraints on output packets, based on previously transmitted 
+input packets.
+\end{description}
+
 We will only examine the output sensor data packet, since it is the more used 
 and interesting one.
+We will examine the first three layers, that is: the packet byte structure, the 
+packet info structure and the single packet rules.
 
-The internal structure of the output sensor data packet reflects the board 
+The packet byte and info structures of the output sensor data packet reflects the board 
 capabilities and sensors.
@@
 a package (complexity to be minimized).
-The complexity minimization constraint is simple to understand: it is the FPGA 
-on the \STSB that build the packet, but it has also to be used for various 
-other specific tasks. 
-Hence, it is important to maintain the building packet code on the FPGA as 
-simple and short as possible, to leave enough space for the code implementing 
-other tasks.
+The complexity minimization constraint rationale is the following: the FPGA 
+on the \STSB builds the packet, but the FPGA is also to be used for various 
+other specific tasks; it is important to maintain the building packet code on the 
+FPGA as simple and short as possible, to leave enough space to implement other 
+tasks.
 The Output sensor data packet structure is showed in Table 
@@
 16 & reserved\\\hline
-17-22 & subsequence 1\\\hline
+17-22 & frame 1\\\hline
 \ldots & \ldots\\\hline
@@
 
-The subsequence accommodate data from the fast and medium data rate streams 
+The frames accommodate data from the fast and medium data rate streams 
 (there are 4 fast data rate stream and 8 medium data rate stream channels). 
-There are 82 subsequences in a packet.
-The subsequence structure is shown in Table \ref{tab:subsequence_structure}.
+There are 82 frames in a packet.
+The frame structure is shown in Table \ref{tab:subsequence_structure}.
 
@@
 can be seen that there must be something subtle, to make all the figures to 
-add up.
+add up, that is, to be capable to accommodate three different rates outputs
+in a fixed structure, fixed length packet.
 The subtleties lies in two forms of redundancies, for slow and medium data 
@@
 
-\todo describe high level representation of package: sensors and data 
-details remains, exact package structure and sensor data location in the
-package are lost
+The single packet rules delimit the boundaries of the values obtainable from 
+the packet: each defined sensor represented in the packet, as well as the 
+metadata describing the \STSB and the packet and frame contents,  have a defined 
+range of acceptable values.
+There are also rules affecting more than one value, basically validity rules (i.e. if 
+a specific sensor is active than the reading value has to be in a defined range), 
+and rules specifing a correct sequence of frames (i.e. there can be only one switch 
+in sampling rate in the frame sequence of a single packet).
 
-\todo describe JML specification of the high level representation of the 
-package
+%=====================================================================
+\section{Data stream and protocol specification}
+\label{sec:data_stream_and_protocol_specification}
+
+The specification of the protocol has been used as a focus point to lead all the 
+other development tasks.
+
+We needed a specification capable of:
+
+\begin{itemize}
+ \item Be executed at run time. If something on the protocol is not respected, 
+the run time execution environment is capable to detect and signal the fault.
+ \item Generate tests. 
+ \item Verify simulators.
+ \item Be used in conjunction of hand written unit tests.
+\end{itemize}
+
+We also needed a specification capable to tackle properly all the protocol layers.
+Since the different layers have very different characteristics, different context
+dependant specification approaches were chosen.
+
+\paragraph{Packet byte structure specification}
+
+The specification is provided by a reference Java implementation capable of 
+recognizing the proper packet structure in a byte data stream. The implementation 
+is tricky, since it checks for the sentinels repeatedly to be sufficiently confidant
+that the byte stream can be properly decomposed into a packet stream.
+
+The implementation can be obviously executed at run time (the specification is the 
+implementation code). 
+Meaningful automatic tests cannot be generated. 
+The implementation can be used to verify a packet byte structure simulator. 
+The implementation were actually driven (Test Driven Development) by the test suite.
+
+\paragraph{Packet info structure specification}
+
+The specification is provided both by the Java API that has been built, and by the 
+annotation with JML language\footnote{
+\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}}. 
+JML has been chosen because it is capable to specify Java code: JML is a behavioral 
+interface specification language. 
+JML specification language combines the design by contract approach (applayed to 
+Java language) and the model-based specification approach of interface specification 
+languages.
+
+\todo insert proper JML references
+
+Many tools are available supporting JML specification language, focusing on different 
+aspects of the language. 
+JML2 suite of tools
+\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}}
+was chosen, since both run time verification support and test generation support are 
+provided directly by the suite.
+
+\todo insert proper JML2 references
+
+The Java API specification part is checked automatically by the Java compiler, while
+the JML specification part can be checked at run time using the provided JML2 run time 
+environment. 
+Meaningful automatic tests can be generated from the JML specification. 
+The JML specification can be used to verify an abstract packet generation simulator, 
+making the simulator an extension of the Java interfaces. 
+The specification can also be used to verify the real driver implementation, based on 
+the packet byte structure implementation presented above. 
+The implementation was driven by the JML specification and by an hand made test suite.
+Finally, hand made unit tests can be checked, during execution, against the 
+specification, using the provided JML2 run time environment.
+
+\paragraph{Single packet rules}
+
+The specification is provided exclusively by annotations with JML language.
+
+The capabilities of a JML specification have been already descibed in the previous 
+paragraph, that is: specification checkable at run time, tests generation, simulator 
+verification, usable with hand written tests.
+
+\subsection{On simulators}
+\label{subsec:on_simulators}
+
+\todo different layers need different simulators
+
+\todo describe simulators structure
+
+\subsection{A JML specification example}
+\label{subsec:a_jml_specification_example}
+
+\todo find interesting JML specification
+
+\todo describe the example
 
@@
 test cases and formal specification of the interfaces
+
+\notev{maybe this part is to be saved for a different paper? BON has 
+not been used properly}
 
@@
 \todo benefit: the approach can be used to test different part of the system
-that share the same specification
+that share the same specifications
 
-\todo testing the stream
+\todo testing the protocols layers
 
-\todo testing the simulator
+\todo testing the simulators
 

### 2009-09-17 papers/BON_VHDL/paper.tex
first draft

@@
   {>=}{{$\textgreater=$}}2
-  {\\}{{$\backslash\backslash$}}1
   {'}{{$\textprimstress$}}1
@@
   morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
-  morekeywords={explanation,query,command,constraint,exists},
+  morekeywords={explanation,query,command,constraint,exists,for_all,such_that,it_holds},
   morecomment=[l]{--},
@@
   morekeywords={sll, srl, sla, sra, rol, ror},
-  morekeywords={rem,abs,mod},
+  morekeywords={rem,abs,mod,to},
   morekeywords={terminal,T'reference, T'contribution,T'tolerance},
@@
 using reasoning techniques provided for algebraic formalism before coding starts.\\
-Fourth,a precise semantics should  provides the basis for
+Fourth, a precise semantics should  provides the basis for
 automated forward-backward engineering.\\
@@
 The central part of the our article is \autoref{sec:refVB}, in which
-we will give practical aspect of refinement , applied on 
-first order passive RC Low-Pass filter, in order to produce formal 
-specification and software simulation.
+we will give practical aspect of refinement, applied on 
+first order passive RC low-pass filter, in order to produce formal 
+specification and software simulation. In this section we 
+explore this approach in detail.
 In \autoref{sec:conclusion} the conclusions are given,
@@
 For this purpose, we will demonstrate our approach in BON-JML fashion trought  
-low-pass filter example.
+low-pass filter example.\\
+\subsection{Low-Pass filter transfer function }
+\label{subsec:lptransfunc} 
 \input{lowpassRC}
-According to equation given above (see \autoref{fig:lowpassfig}) 
+The most simple low-pass filter is passive RC network 
+shown in \autoref{fig:lowpassfig}.
+According to equation given above (see \autoref{fig:transfun}) 
 we can produce VHDL-AMS \emph{simultaneous statement} using 
-\texttt{$s \rightarrow d/dt \rightarrow 'dot$} relation.
+\texttt{$s = d/dt = 'dot$} relation.
 The transformation looks as follows: \\
@@
    \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle s \rightarrow d/dt}  \\=                                                                                                                  
+      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
    \end{array} 
@@
 \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle d/dt \rightarrow } {\scriptscriptstyle 'dot} \\=                                                                                                                  
+      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
    \end{array}
@@
 \end{equation} 
-\input{lowpass}
+ The last equation can be described as follows. 
+The input voltage\xspace($v_{in}$)\xspace equals summation 
+of output voltage\xspace($v_{out}$)\xspace and multiplication 
+of the time constant\xspace($\tau_p$)\xspace of the pole and the
+first derivative of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace 
+over time.\\
+Now we can discretize previous equation. Let us represent $v_{in}$ by a series 
+of samples $v_{in(1..n)}$. At the same points in time, we can represent $v_{out}$ by 
+series of sample $v_{out(1..n)}$. We can assume that samples are taken at points in 
+time incrementing by $\triangle$t. Now we have equivalent equations:
+\begin{equation}
+v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                                                  
+\end{equation}
+
+\begin{equation}
+v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
+ v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                                                                  
+\end{equation}
+
+\begin{equation}
+v_{out(i)} 
+ \begin{array}[b]{c}                                                                                                                                    
+      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                                                           
+ \end{array}
+ \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                                                                  
+\end{equation}
+
+The last equation gives us a way to calculate the output samples in terms of the 
+input samples and the preceding output.
+But how to represent these unknowns in BON ?
+We introduce the  \lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
+with features as fallows:
+\begin{itemize}
+\item
+\emph{across} which represent current:$i_{out}$ 
+\item 
+\emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
+\item 
+\emph{terminal sequences} which describes input/output terminal(s). 
+\end{itemize}
+Now we are in presence of BON notation and we may start
+with the  refinement of the VHDL-AMS code.\\
+\subsection{Refine example }
+\label{subsec:refine}     
 We have not yet said what specification and refinement are.
 Let us introduce these terms first.\\
-A specification of a program component is a contract that gives some constraints 
+A \emph{specification} of a program component is a 
+contract that gives some constraints 
 on how the component is to behave but leaves freedom for the implementator to 
 decide how the actual behavior of the component is to be realized. 
-Specification involves \emph{choice} between a number of alternatives; this 
- relationship is one-to-many. As we have shown in \autoref{tab:rescap}, corresponding 
-informal specifications represents high level of abstraction derived from VHDL-AMS code.
-
-A refinement represents the concept that can be understud as series of 
+As we have shown in \autoref{tab:rescapfirst}, corresponding 
+informal specifications represents high level of abstraction derived 
+from VHDL-AMS code.\\
+A \emph{refinement} represents the concept that can be understood as series of 
 correctness-preserving transformations from general perspective to 
@@
 It means that "\texttt{S is \emph{refined} by S'}", 
-and we denote this  as \texttt{S $\rightsquigarrow$ S'}.
+and we denote this  as \texttt{S $\sqsubseteq$ S'}.
+The \emph{refinement} which was derived in three steps, 
+from informal to more formal BON and finally to JML, 
+is represented using relation \texttt{$\sqsubseteq$}. This relation represent 
+\emph{isomorphism} and has inverse relation \texttt{$\sqsupseteq$}. 
+It means that once we have more formal specification, we can refine 
+and seamlessly  translate specification in previous abstract level,  
+keeping specification validity and documentation consistency. 
+In other  words, we use two important BON principles to achieve 
+as much as precise model consistency: the seamlessness and  
+reversibility(see \autoref{sec:choiceobon}). 
 Let say that the first informal BON specification is denoted by 
-\texttt{S} (see \autoref{tab:rescap}), and second, which is given in Table2  is denoted by \texttt{S'}.
-In this case, the refinement relation represents an \emph{isomorphism} 
-and maps the following:
+\texttt{S} (see \autoref{tab:rescapfirst}), and second and third,
+which are given in \autoref{tab:rescapsecond}  is denoted by \texttt{S',S''}.
+The forward process refinement can be represented as:\xspace
+\texttt{S $\sqsubseteq$ S' $\sqsubseteq$ S''}, while the backward process
+can be represented as \xspace \texttt{S''$\sqsubseteq$ S' $\sqsubseteq$ S}.  
+Using all BON principles, we have developed an specification of RC low-pass
+filter, and we have produced fully validated and verified low-pass filter.
+\input{lowpass1}
+\input{lowpass2}
+
 %=====================================================================
@@
 \input{quantity}
-\input{quantityport}
-\input{terminalport}
-\input{signal}
-\input{signalattr}
-\input{wait}
-\input{bitbon}
+%\input{quantityport}
+%\input{terminalport}
+%\input{signal}
+%\input{signalattr}
+%\input{wait}
+%\input{bitbon}
 %======================================================================

### 2009-09-24 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: added JML specification examples

@@
 % Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
-% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
+% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz 
 %
@@
 \usepackage{float}
+\usepackage{listings}
+\input{jml-listings}
+\lstset{language=[JML]Java,xleftmargin=20pt,xrightmargin=20pt}
 
@@
 \newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3 \myhref{#1}{#1}}}
+
+\newcommand{\lil}[1]{\texttt{\lstinline|#1|}}
 
@@
 
-\todo other approaches to the problem
+\todo search: other approaches to similar problems
 
@@
 
-\todo different layers need different simulators
+Simulators are needed to test in isolation parts of the system; especially the upper 
+layers, yet to be developed. 
+During the development the simulators were essential, since the board was not available.
+Simulators were built concurrently with the specification, the real driver code, and the 
+tests.
 
-\todo describe simulators structure
+Theoretically a separate simulator should be needed for each protocol layer; a
+separate simulator for each layer would enable us to separeately specify, develop and test
+each layer of the protocol.
+In our case we evaluated not worth the effort to maintain such a separation in the 
+simulators, motivated by the fact that the protocol layers identified are too fine grained 
+to be separated into different simulators. 
+Two simulators were built.
 
-\subsection{A JML specification example}
+The two simulators were built reflecting the software abstraction layers defined for the 
+driver.
+The driver API is splitted in two structurual layers: the general high level interfaces, that 
+expose the main functionalities of the board and the main contents of the packets hiding the 
+details, and the lower level implementation that parse the data streams in and out the board, 
+to translate the higher level intructions and data in properly formed packets.
+Hence, we have implemented two simulators: one high level simulator, whicj reimplements the 
+high level interfaces providing corresponding high level mechanisms to control the data and 
+the behaviour of the simulated board, it has nothing to do with the details of the data 
+streams.
+One low level simulator which is capable to rebuild the data stream out of the board based on 
+the data stream into the board, the in and out data streams are built exaclty as the sensor 
+board is expected to be able to parse or generate.
+
+\subsection{JML specification examples}
 \label{subsec:a_jml_specification_example}
 
-\todo find interesting JML specification
+The JML specifications used to annotate Java code are of varing complexities.
+Some of the specifications are simple, focusing on constraints that should hold when 
+calling a method (preconditions) and the constraints on the return value (a very basic form of 
+postconditions).
+In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the 
+method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that 
+it cannot change the state of any object (a postcondition), and the lower and upper bound of 
+its return value are defined (another postcondition).
 
-\todo describe the example
+\begin{lstlisting}[
+  caption={A simple specification with JML annotation: simple postconditions.},
+  label=lst:jml_example_simple,
+  float=htbp]
+/*@ 
+  ensures \result >= -880;
+  ensures \result <= 2047;
+@*/
+/*@ pure @*/ short getTemperature();
+\end{lstlisting}
+
+Some of the specifications are rather complex, usually focusing on properties regarding the 
+behaviour of a whole object.
+In listing \ref{lst:jml_example_invariant} a complex invariant 
+(a property that has to be maintained, that is, that is assumed on entry and guaranteed on 
+exit of each method of an object)
+example is shown.
+
+\sloppy
+
+The invariant is constraining the number of samples for each medium data rate streams: 
+the total number of streams is \lil{Frame.ADC_CHANNELS}; the total number of frames is 
+\lil{FRAMS}; the constant constraining the number of samples is 
+\lil{FRAMES/Frame.ADC_CHANNELS+1}, it means that the samples contained in a frame are
+fairly distributed on the channels. 
+The valid samples are counted parsing all the frames contained in a packet, selecting only 
+the matching valid samples.
+A medium data rate stream sample is considered valid when \lil{isADCActive()} method 
+returns \lil{true}.
+
+\fussy
+
+\begin{lstlisting}[
+  caption={A specification with JML annotation: a complex object invariant which constraints 
+    the number of samples for each medium data rate stream channel.},
+  label=lst:jml_example_invariant,
+  float=htbp]
+/*@ 
+  invariant (
+    \forall int channel;
+    0 <= channel && channel < Frame.ADC_CHANNELS; 
+    (
+      \num_of int i;
+      0<= i && i<(FRAMES-1);
+      (
+        getFrame(i).isADCActive() &&
+        (getFrame(i).getADCChannel() == channel)
+      )
+    ) <= (FRAMES / Frame.ADC_CHANNELS + 1)
+  );
+@*/
+\end{lstlisting}
 
@@
 specification tests
+
+\todo reason about the usual structure oracle-driver-test-stub(s). a simulator is sometimes a 
+stub, sometimes a driver, sometimes an oracle...
+
+\todo insert references to testing embedded systems with simulators and/or with driver/stubs
 

### 2009-09-24 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: added UML model, not yet finished explained part of the test structure

@@
 In section \ref{sec:test_cases_on_the_job} the execution of the test suites
-is described, and the results are analyzed and compared, both from qualitative
-and quantitative points of view.
+is described, the data collected is presented.
 
-In section \ref{sec:test_cases_and_the_scenarios} we examine the different 
-scenarios on which test cases based on specifications can be applied, and 
-the resulting benefits it is possible to obtain.
+In section \ref{sec:test_cases_retrospectives} the data collected is analyzed, 
+both from a qualitative and a quantitative points of view.
+
+In section \ref{sec:test_cases_and_the_scenarios} the different scenarios on 
+which test cases based on specifications can be applied are analyzed, and 
+the resulting benefits it is possible to obtain are detailed.
 
@@
 method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that 
-it cannot change the state of any object (a postcondition), and the lower and upper bound of 
-its return value are defined (another postcondition).
+it cannot change the state of any object (a postcondition), the specification also constraints 
+the return value with a lower and upper bound (another postcondition).
 
@@
 
+\sloppy
+
 Some of the specifications are rather complex, usually focusing on properties regarding the 
@@
 example is shown. 
-
-\sloppy
-
 The invariant is constraining the number of samples for each medium data rate streams: 
-the total number of streams is \lil{Frame.ADC_CHANNELS}; the total number of frames is 
-\lil{FRAMS}; the constant constraining the number of samples is 
-\lil{FRAMES/Frame.ADC_CHANNELS+1}, it means that the samples contained in a frame are
+the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is 
+\lil{FRAMS}, the constant constraining the number of samples is 
+\lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are
 fairly distributed on the channels. 
@@
 \label{sec:test_cases}
+
+The unit test cases that have been built to verify the protocol driver 
+are of two kinds: hand made unit tests and unit tests automatically 
+generated based on JML specifications.
+The tests have been tought to be used together, complementarly. 
+Despite of this premise, the tests effectiveness has been evaluated 
+separately and together, the evaluation is carried out in section 
+\ref{sec:test_cases_retrospectives}.
+The test effectiveness evaluation has been based on three elements:
+code coverage (quantitative), development help and usefulness 
+(qualitative), effective results on piloting the real board 
+(quantitative).
+
+The dual testing approach was thought as a winning option when 
+compared to hand based tests or completely automated tests, in testing 
+a driver for a stream communication protocol with hard constraints 
+on development time.
+\ST stream communication protocol is characterized by a well definite 
+packet structure, and a well definite semantic on the packet contents 
+as well.
+On the other hand there are not strong constraints on synchronization, 
+since there exists two concurrent communication channels and the protocol 
+is asynchronous by construction.
+The lack of hard synchronization constraints reduce the advantage to use 
+a classical state machine based specification.
+
+\todo reference to protocol tests generated from a state machine based 
+specification is needed
+
+Hard development time constraints were also a precise requirement; the time 
+constraints pushed us towards a specification \emph{just good enough} for
+our purposes.
+A \emph{just good enough} specification is simply a specification that is not 
+specifing everything, but only the software parts that are considered most 
+critical, or the ones that are really easy to specify with the chosen 
+specification language.
+In \STSB driver JML was used to specify very simple behaviours (method 
+oriented) on the whole driver implementaion, and more complex behaviour on 
+the driver core interfaces.
+The not covered areas were addressed by hand made unit tests.
+
+Hand made unit tests cover behaviour not covered by the automatically 
+generated tests, that is, they cover the not specified behaviour; but this 
+is not the only purpose they have.
+Hand made unit tests are used to drive the development, using the well known 
+Test Driven Development approach, as well as to drive the most error prone 
+specification as well.
+Tests are used to drive the specification, a hand made unit test that should 
+fail if the specification is not working as expected.
+
+The rationale is the following: a stream communication protocol is 
+not completely focused on a synchronised interaction
+
+\begin{itemize}
+ \item 
+\end{itemize}
 
@@
 \todo compare compare coupled approach vs single approaches
+
+%=====================================================================
+\section{Retrospective on test cases effectiveness}
+\label{sec:test_cases_retrospectives}
+
+The test effectiveness evaluation criteria is of two types.
+
+\todo describe quantitative: coverage
+
+\todo describe qualitative: errors found, help in design, ...
 

### 2009-10-05 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper ModelBasedTestCaseGenerationDataStreamProtocol: error detection results added

@@
 %
-% "The UCD CASL SenseTile System", for IEEE Pervasive Computing
-% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
-% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz 
+% "The UCD CASL SenseTile System", for IEEE Pervasive Computing Joseph
+% R. Kiniry, Vieri del Bianco, Dragan Stosic $Id: paper.tex 2795
+% 2007-09-23 19:35:53Z dmz
 %
 
-\documentclass{article} 
-\usepackage{times}
+\documentclass{article} \usepackage{times}
 
-\usepackage{ifpdf}
-\usepackage{a4wide}
-\usepackage{pdfsync}
+\usepackage{ifpdf} \usepackage{a4wide} \usepackage{pdfsync}
 
-\ifpdf
-\usepackage[pdftex]{graphicx}
-\else
-\usepackage{graphicx}
-\fi
+\ifpdf \usepackage[pdftex]{graphicx} \else \usepackage{graphicx} \fi
 
-\usepackage{xspace}
-\usepackage{tabularx}
-\usepackage{epsfig}
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{eucal}
-\usepackage{stmaryrd}
-\usepackage{float}
-\usepackage{listings}
-\input{jml-listings}
+\usepackage{xspace} \usepackage{tabularx} \usepackage{epsfig}
+\usepackage{amsmath} \usepackage{amsfonts} \usepackage{amssymb}
+\usepackage{eucal} \usepackage{stmaryrd} \usepackage{float}
+\usepackage{listings} \input{jml-listings}
 \lstset{language=[JML]Java,xleftmargin=20pt,xrightmargin=20pt}
@@
 \ifpdf
+\usepackage{epstopdf}
 \usepackage[pdftex,bookmarks=false,a4paper=false,
@@
             linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan
-Stosic"]{hyperref}
+            pdfauthor="Joseph R. Kiniry and Vieri del Bianco"]{hyperref}
 \else
@@
 
-\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
+%\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
+\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:} #1}}
 \newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
@@
 
-\newcommand{\todo}{\textbf{TODO:}}
+\newcommand{\todo}[1]{\texttt{\textbf{TODO:} #1}}
 
 \newcommand{\ST}{\emph{SenseTile}\xspace}
-\newcommand{\STSB}{\ST \emph{Sensor Board}\xspace}
+\newcommand{\SB}{\emph{Sensor Board}\xspace}
+\newcommand{\STSB}{\ST \SB\xspace}
 \newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
@@
 
-\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
+\author{Joseph R. Kiniry, Vieri del Bianco\\
 UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
 Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
+kiniry@acm.org, vieri.delbianco@gmail.com\\
 }
@@
 
-\todo explain and refine the problem
+\todo{explain and refine the problem}
 
-\todo search: other approaches to similar problems
+\todo{references: search other approaches to similar problems}
 
@@
 
-Finally in section \ref{sec:conclusion} conclusions are drawed and 
+Finally in section \ref{sec:conclusion} conclusions are drawn and 
 ongoing and future works on similar topics is introduced.
@@
 protocol description can lead to overly complex and unmanageable descriptions, 
-expecially when dealing with formal ones.
+especially when dealing with formal ones.
 The layers identified in protocol are the following:
@@
 a specific sensor is active than the reading value has to be in a defined range), 
-and rules specifing a correct sequence of frames (i.e. there can be only one switch 
+and rules specifying a correct sequence of frames (i.e. there can be only one switch 
 in sampling rate in the frame sequence of a single packet).
@@
 Since the different layers have very different characteristics, different context
-dependant specification approaches were chosen.
+dependent specification approaches were chosen.
 
@@
 The specification is provided both by the Java API that has been built, and by the 
-annotation with JML language\footnote{
+annotation with JML language\cite{JML2}\footnote{
 \myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}}. 
@@
 interface specification language. 
-JML specification language combines the design by contract approach (applayed to 
+JML specification language combines the design by contract approach (applied to 
 Java language) and the model-based specification approach of interface specification 
@@
 
-\todo insert proper JML references
-
 Many tools are available supporting JML specification language, focusing on different 
 aspects of the language. 
-JML2 suite of tools
+JML2 suite of tools\cite{BurdyEtal05-STTT}
 \footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}}
@@
 provided directly by the suite.
-
-\todo insert proper JML2 references
 
@@
 
-The capabilities of a JML specification have been already descibed in the previous 
+The capabilities of a JML specification have been already described in the previous 
 paragraph, that is: specification checkable at run time, tests generation, simulator 
@@
 Theoretically a separate simulator should be needed for each protocol layer; a
-separate simulator for each layer would enable us to separeately specify, develop and test
+separate simulator for each layer would enable us to separately specify, develop and test
 each layer of the protocol.
@@
 driver.
-The driver API is splitted in two structurual layers: the general high level interfaces, that 
+The driver API is splitted in two structural layers: the general high level interfaces, that 
 expose the main functionalities of the board and the main contents of the packets hiding the 
 details, and the lower level implementation that parse the data streams in and out the board, 
-to translate the higher level intructions and data in properly formed packets.
-Hence, we have implemented two simulators: one high level simulator, whicj reimplements the 
+to translate the higher level instructions and data in properly formed packets.
+Hence, we have implemented two simulators: one high level simulator, which re-implements the 
 high level interfaces providing corresponding high level mechanisms to control the data and 
-the behaviour of the simulated board, it has nothing to do with the details of the data 
+the behavior of the simulated board, it has nothing to do with the details of the data 
 streams.
 One low level simulator which is capable to rebuild the data stream out of the board based on 
-the data stream into the board, the in and out data streams are built exaclty as the sensor 
+the data stream into the board, the in and out data streams are built exactly as the sensor 
 board is expected to be able to parse or generate.
@@
 
-The JML specifications used to annotate Java code are of varing complexities.
+The JML specifications used to annotate Java code are of varying complexities.
 Some of the specifications are simple, focusing on constraints that should hold when 
@@
 Some of the specifications are rather complex, usually focusing on properties regarding the 
-behaviour of a whole object.
+behavior of a whole object.
 In listing \ref{lst:jml_example_invariant} a complex invariant 
@@
 the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is 
-\lil{FRAMS}, the constant constraining the number of samples is 
+\lil{FRAMES}, the constant constraining the number of samples is 
 \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are
@@
 are of two kinds: hand made unit tests and unit tests automatically
-generated based on JML specifications.
-The tests have been tought to be used together, complementarly. 
-Despite of this premise, the tests effectiveness has been evaluated 
-separately and together, the evaluation is carried out in section 
-\ref{sec:test_cases_retrospectives}.
-The test effectiveness evaluation has been based on three elements:
-code coverage (quantitative), development help and usefulness 
-(qualitative), effective results on piloting the real board 
-(quantitative).
+generated based on JML specifications.  The tests have been thought to
+be used together, they are complementary.  The tests effectiveness has
+been evaluated, for each kind of test, the evaluation is carried out
+in section \ref{sec:test_cases_retrospectives}.  The test
+effectiveness evaluation has been based on three elements: effective
+results on piloting the real board (quantitative), code coverage
+(quantitative), development help and usefulness (qualitative).
 
-The dual testing approach was thought as a winning option when 
-compared to hand based tests or completely automated tests, in testing 
-a driver for a stream communication protocol with hard constraints 
+The dual testing approach was thought as a winning option, when 
+compared to hand based tests or completely automated tests used alone, 
+testing a driver for a stream communication protocol with hard constraints 
 on development time.
@@
 as well.
-On the other hand there are not strong constraints on synchronization, 
-since there exists two concurrent communication channels and the protocol 
-is asynchronous by construction.
+On the other hand in the protocol there are not strong constraints on 
+synchronization, since the communication channel is splitted in two (in and 
+out) separate concurrent and asynchronous channels.
 The lack of hard synchronization constraints reduce the advantage to use 
-a classical state machine based specification.
-
-\todo reference to protocol tests generated from a state machine based 
-specification is needed
+a classical state machine based specification \todo{references: 
+protocol tests generated from a state machine based specification}.
 
 Hard development time constraints were also a precise requirement; the time 
-constraints pushed us towards a specification \emph{just good enough} for
-our purposes.
-A \emph{just good enough} specification is simply a specification that is not 
-specifing everything, but only the software parts that are considered most 
-critical, or the ones that are really easy to specify with the chosen 
-specification language.
-In \STSB driver JML was used to specify very simple behaviours (method 
-oriented) on the whole driver implementaion, and more complex behaviour on 
+constraints pushed us towards a incomplete (\emph{just good enough}) 
+specification.
+In \STSB driver JML was used to specify very simple behaviors (method 
+oriented) on the whole driver implementation, and more complex behavior on 
 the driver core interfaces.
-The not covered areas were addressed by hand made unit tests.
+The software system parts not specified by the JML2 specification were 
+specified using hand made unit tests.
 
-Hand made unit tests cover behaviour not covered by the automatically 
-generated tests, that is, they cover the not specified behaviour; but this 
+Hand made unit tests cover behavior not covered by the automatically 
+generated tests, that is, they cover the not specified behavior; but this 
 is not the only purpose they have.
-Hand made unit tests are used to drive the development, using the well known 
-Test Driven Development approach, as well as to drive the most error prone 
-specification as well.
-Tests are used to drive the specification, a hand made unit test that should 
-fail if the specification is not working as expected.
+Hand made unit tests are used to drive the development, since the Test Driven 
+Development approach\cite{beck2003test} has been followed during development.
+Hand made tests are also used to drive the development of the most error 
+prone or complex JML2 specifications; a test developed to drive the 
+specification is a test that should fail if the specification is not working 
+as expected.
 
-The rationale is the following: a stream communication protocol is 
-not completely focused on a synchronised interaction
-
-\begin{itemize}
- \item 
-\end{itemize}
-
-\todo describe dual approach: hand made tests + generated from 
-specification tests
-
-\todo reason about the usual structure oracle-driver-test-stub(s). a simulator is sometimes a 
-stub, sometimes a driver, sometimes an oracle...
-
-\todo insert references to testing embedded systems with simulators and/or with driver/stubs
-
-\todo describe hybrid agile and TDD approach: high level informal 
-specifications with BON, iterative refinements of code, to distill 
-test cases and formal specification of the interfaces
-
-\notev{maybe this part is to be saved for a different paper? BON has 
-not been used properly}
-
-\todo describe the necessity to add specification generated tests: to ensure
-that specification is sound, to obtain higher coverage
+\todo{describe the necessity to add specification generated tests: to ensure
+that specification is sound, to obtain higher coverage}
 
@@
 
-\todo describe hand made tests structure
+The package structure of the driver developed is shown in Figure 
+\ref{fig:class_diagram_main}: two independent packages has been defined: 
+\lil{Stream} and \lil{Driver}.
+The packages are abstract, that is, they mainly contain abstractions; in Java 
+language this is translated into a package which contains mainly abstract classes 
+or interfaces.
+The two abstract packages has various implementations, the dependency between
+abstract packages is not direct, but only through an implementation 
+(\lil{StreamDriver}).
+This is needed to maintain a high decoupling of the packages, and is the result of 
+applying the \emph{classical} dependency inversion 
+principle\cite{martin1996dependency} (the dependency structure of the 
+packages has a saw shape, a common shape on systems where the dependency 
+inversion principle has been followed).
 
-\todo describe abstract test case and abstract test suite patterns: how to 
-write tests against interfaces and/or abstract classes
+\begin{figure}[htb!]
+ \centering
+ \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__Main}
+ \label{fig:class_diagram_main}
+ \caption{Main packages class diagram.}
+\end{figure}
+
+The test package structure, which is shown in Figure 
+\ref{fig:class_diagram_test} reflects and mimics the code structure; the
+tests that test an abstract package are abstract, to be implemented by the 
+package that tests a corresponding system implementation; this is a well 
+known test pattern, the Abstract Test Pattern\cite{thomas2004java}, it is 
+used to test that the contracts are respected in the implementations.
+For instance, package \lil{DriverT} contains abstract tests for the 
+abstractions of package \lil{Driver}, package \lil{StreamDriverT} 
+inherits the abstractions contained in \lil{DriverT} and makes them concrete, 
+to test the corresponding implementation \lil{StreamDriver}; package 
+\lil{StreamDriverT} also contains stand alone tests written specifically for
+the implementation contained in \lil{StreamDriver}.
+
+\begin{figure}[htb!]
+ \centering
+ \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__Test}
+ \label{fig:class_diagram_test}
+ \caption{Test packages class diagram.}
+\end{figure}
 
@@
 
-\todo describe the tool that generates the testing
+The JML specification has been used to generate the tests. 
+A specific tool of the JML2 suite can be used to automatically generate 
+them\cite{Cheon-Leavens02}.
 
-\todo describe the structure of the generates tests
+The test suite generated is a collection of fine grained unit tests.
+For each public method a set of tests is generated.
+The specified behavior of the method is checked using the JML2 run time 
+assertion checker, hence, an oracle is not needed. 
+The generated tests send messages to the method under test, they catch 
+assertion violation exceptions from test cases that pass an initial 
+precondition check; that is, if a precondition is not met, the test is 
+not run, if the precondition is met the test is run, and if the postcondition 
+is not me an exception is fired and the test fails.
+The test data that is used to build the tests is the only part that is 
+required to be provided by the user.
+
+The test package structure is shown in Figure 
+\ref{fig:class_diagram_generatedtest}.
+The packages dependencies are defined by the specifications; for instance, 
+\lil{SimulatorDriverST} are generated tests, they test (hence use) the 
+implementation \lil{SimulatorDriver}, they have been generated using the 
+specification corresponding specification \lil{SimulatorDriverS}.
+The generated tests do not maintain the inheritance structure of the code 
+and the specification.
+
+\begin{figure}[htb!]
+ \centering
+ \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
+ \label{fig:class_diagram_generatedtest}
+ \caption{Generated test packages class diagram.}
+\end{figure}
+
+\paragraph*{On simulators, again}
+
+To have proper unit test suite, for each method under test a proper stub, 
+driver and oracle should be identified, these are the usual components found 
+in the harness that enable test execution\cite{binder1999testing}.
+In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework used 
+to manage the implementation and execution of hand made and generated unit 
+tests, there is no clear distinction between drivers, stubs, and oracle of a 
+test; noneless, even if a proper distinction is not imposed by the framework, 
+the conceptual distinction still holds.
+
+Hence, to properly unit test a module in isolation and automatically we need:
+
+\begin{description}
+ \item[Stub or Skeleton]: a piece of code that simulates the activity of 
+the modules that are not under tests.
+ \item[Driver]: a piece of code that calls in a proper way the module under 
+test.
+ \item[Oracle]: a mechanism used for determining whether a test has passed 
+or failed, comparing the output of the module under test to the output that 
+the oracle determines that the module should have.
+\end{description}
+
+Hand made unit tests and JML generated tests differs in how these components 
+are provided. 
+The distinctions and similarities are summarized in Table 
+\ref{tab:test_harness}.
+
+\begin{table}[htbp]
+\caption{Hand made and generated tests harness.}
+\label{tab:test_harness}
+\begin{center}
+\begin{tabular}{|l|l|l|}\hline
+& \textbf{\textit{hand made}} & \textbf{\textit{JML}} \\\hline
+\textbf{\textit{Stub}} & to be provided & to be provided\\\hline
+\textbf{\textit{Driver}} & to be provided & automatic: 1 call\\\hline
+\textbf{\textit{Oracle}} & to be provided & 
+automatic: runtime assertion checker\\\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+A module simulator can be used to ease the tasks of developing each component 
+of a test harness; that is, it can be used as a stub, to simulate a module 
+needed by the module under test, it can be used as a driver, to stimulate the 
+module under test, it can be used as an oracle, if it simulates the module 
+under test.
+
+In our system, the simulators has been used mainly as stubs, so that we have,
+during testing, a structure like the one shown in Figure 
+\ref{fig:class_diagram_streamdriver_test}.
+In Figure \ref{fig:class_diagram_streamdriver_test} the test suite 
+\lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate 
+the \lil{Stream} that is used by the system under test, that is 
+\lil{StreamDriver}.
+
+\begin{figure}[htb!]
+ \centering
+ \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__StreamDriverT}
+ \label{fig:class_diagram_streamdriver_test}
+ \caption{StreamDriver test packages class diagram: use of the stream simulator.}
+\end{figure}
+
+In our case, the usage of simulators to enable unit testing, was essential to 
+test the \STSB driver, since the physical board was not available.
+The simulator were used in both the hand made and the generated tests, since
+both of them need a proper set of stubs to be executed.
+
+\todo{insert references to testing embedded systems with simulators and/or 
+with driver/stubs}
+
+\paragraph*{BON}
+
+\todo{describe approach: high level informal specifications with BON, 
+iterative refinements of code, to distill test cases and formal 
+specification of the interfaces} \notev{remove, BON has not been used 
+properly, maybe a follow up paper}
 
@@
 
-\todo report high level analysis of the test suites
+The use of both test suites, hand made and generated, has proven 
+effective on finding protocols errors on the \STSB when the board 
+prototypes have been delivered.
 
-\todo describe measures: coverage, number of new bugs found
+The first \STSB prototype delivered was meant to build empty packets, 
+but the packets were built in a way not conformant to the 
+specification, the sentinel values used to delimit a packet were not 
+correct.
+The board implementation error has been detected by the 
+\lil{StreamDriver} package.
 
-\todo compare hand made vs generated tests
+The second prototype delivered was meant to build packets with values
+taken from rear sensors installed on the board.  The only sensors that
+were not yet working were the sensors devoted to the fast rate data
+streams, that is, the audio sensors.  Only one error needed to be
+corrected in the driver implementation, because of a (rare)
+combinations of conditions that were not covered by the test cases.
+One error have been discovered in the low level driver that the driver
+was using; to eliminate the error an upgrade of the low level drover
+to a new unreleased version was necessary.  No other errors were
+detected on the driver side. All the errors on the \SB side have been
+detected.
 
-\todo compare compare coupled approach vs single approaches
+The formal model and the run time assertion checker enabled the
+detection of errors on the \SB itself.  Errors regarding the expected
+values and values ranges of the filds contained in the packets and the
+frames (4 protocol errors), and errors regarding the dependencies
+between values in field (2 protocol errors) were detected. No other
+errors, other than the ones immediately discovered by the run time
+assertion checker were later on discovered.
+
+The hand made test suite is composed by over 140 tests, while the
+generated one is 100 times bigger, totalling over 14000 tests.
+
+\todo{describe why such a difference: random selection, many tests 
+pruned if pre condition is not satisfied}
+
+\todo{compare hand made, many calls, vs generated, single call, tests}
+
+\todo{describe measures: coverage, number of new bugs found}
+
+\todo{compare compare coupled approach vs single approaches}
 
@@
 
-\todo describe quantitative: coverage
+\todo{describe quantitative: coverage}
 
-\todo describe qualitative: errors found, help in design, ...
+\todo{describe qualitative: errors found, help in design, ...}
 
@@
 
-\todo benefit: the approach can be used to test different part of the system
-that share the same specifications
+\todo{benefit: the approach can be used to test different part of the system
+that share the same specifications}
 
-\todo testing the protocols layers
+\todo{testing the protocols layers}
 
-\todo testing the simulators
+\todo{testing the simulators}
 
-\todo describe general approach: depending at what level (of abstraction) tests 
+\todo{describe general approach: depending at what level (of abstraction) tests 
 and specifications are written, it is possible to verify both the simulator and 
-the real protocol (isn't this obvious?)
+the real protocol (isn't this obvious?)}
 
@@
 
-\todo
-
 %=====================================================================
@@
 \label{sec:conclusion}
-
-\todo
 

### 2009-10-06 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: discussion on hand made and generated tests

@@
 
-  Sensor equipped devices typically communicate with a streaming packet-based 
-  network protocol.
-  The communication is focused on controlling the sensors and report the 
-  sensors measurements.
-  The \STSB is an example of such a device, it communicates with an arbitrary 
-  computer system, the \STPU (running Linux or Windows, at the moment) over 
-  a USB bus using a streaming packet-based protocol.
-  The \STU is a custom-designed, but inexpensive, sensor platform: it is 
-  composed by a one or more \STSBs paired with a general-purpose small-scale 
-  compute node, the \STPU (including everything from PDAs to powerful 
-  touchscreen-based portable computers).
-  There are over a dozen sensors on the \STSB itself, and new sensors can be 
-  added either using expansion slots on the \STSB, or adding them directly to 
-  the USB bus.
-  A third-party low-level library, provided by the USB chip vendor, the USB 
-  chip that has been used in the \STSB, creates and parses
-  generic USB communication packets.
+  Sensor equipped devices typically communicate with a streaming
+  packet-based network protocol.  The communication is focused on
+  controlling the sensors and report the sensors measurements.  The
+  \STSB is an example of such a device, it communicates with an
+  arbitrary computer system, the \STPU (running Linux or Windows, at
+  the moment) over a USB bus using a streaming packet-based protocol.
+  The \STU is a custom-designed, but inexpensive, sensor platform: it
+  is composed by a one or more \STSBs paired with a general-purpose
+  small-scale compute node, the \STPU (including everything from PDAs
+  to powerful touchscreen-based portable computers).  There are over a
+  dozen sensors on the \STSB itself, and new sensors can be added
+  either using expansion slots on the \STSB, or adding them directly
+  to the USB bus.  A third-party low-level library, provided by the
+  USB chip vendor, the USB chip that has been used in the \STSB,
+  creates and parses generic USB communication packets.
 
-  A custom medium-level Java driver creates and parses the \STSB streaming 
-  communication packets. 
-  The driver is formally specified, validated and verified with rigorous
-  unit testing. 
-  Because of this process and combination of technologies, using this
-  specification we are able to both communicate with a real \ST
-  \textbf{and} generate a formally specified, executable, functional
-  model-based specification for the protocol itself written in the
-  Java Modeling Language.
-  \notev{executable? how? what does this mean?}
-  This protocol model is itself an executable
-  stream simulator that can be used to develop new \ST user
-  applications without having the \ST hardware at all. 
-  Consequently,
-  from this single specification, we can test that the \ST hardware
-  communicates as expected, test that our \ST simulator communicates
-  as expected, and simulate the \ST protocol stream itself.
+  A custom medium-level Java driver creates and parses the \STSB
+  streaming communication packets.  The driver is formally specified,
+  validated and verified with rigorous unit testing.  Because of this
+  process and combination of technologies, using this specification we
+  are able to both communicate with a real \ST \textbf{and} generate a
+  formally specified, executable, functional model-based specification
+  for the protocol itself written in the Java Modeling Language.
+  \notev{executable? how? what does this mean?}  This protocol model
+  is itself an executable stream simulator that can be used to develop
+  new \ST user applications without having the \ST hardware at all.
+  Consequently, from this single specification, we can test that the
+  \ST hardware communicates as expected, test that our \ST simulator
+  communicates as expected, and simulate the \ST protocol stream
+  itself.
 
@@
 
-The use of both test suites, hand made and generated, has proven 
-effective on finding protocols errors on the \STSB when the board 
-prototypes have been delivered.
+JML specifications, the simulators,  and both test suites has been used to:
+
+\begin{itemize}
+\item verify hardware implementation;
+\item verify driver implementation, with and without the physical
+  board;
+\item guide driver development.
+\end{itemize}
 
@@
 specification, the sentinel values used to delimit a packet were not
-correct.
-The board implementation error has been detected by the 
+correct.  The board implementation error has been detected by the
 \lil{StreamDriver} package.
@@
 The second prototype delivered was meant to build packets with values
-taken from rear sensors installed on the board.  The only sensors that
+taken from real sensors installed on the board.  The only sensors that
 were not yet working were the sensors devoted to the fast rate data
@@
 combinations of conditions that were not covered by the test cases.
-One error have been discovered in the low level driver that the driver
-was using; to eliminate the error an upgrade of the low level drover
-to a new unreleased version was necessary.  No other errors were
-detected on the driver side. All the errors on the \SB side have been
-detected.
+One error have been discovered in the low level system driver that the
+developed driver was using (the system driver was outside of our
+development scope); to eliminate the error an upgrade of the low level
+driver to a new unreleased version was necessary.  No other errors
+were detected on the driver side. All the errors on the \SB side have
+been detected.
 
@@
 
-\todo{describe why such a difference: random selection, many tests 
-pruned if pre condition is not satisfied}
+The ratio of 1 to 100, that is, the big number of generated tests, can
+be understood looking closely at how the generated tests are created
+by the JML framework. The generated tests are exploring all the
+possible input combinations on public methods, combining the type data
+ranges that have been specified for the test suite (see
+\cite{Cheon-Leavens02} for more details on how the JML framework works
+on generating unit tests). For instance, let's suppose we want to test
+a method \lil{m(int p1, int p2)} for a class \lil{C}; the data ranges
+specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C}
+are the instances \lil{\{c1,c2,c3,c4\}}. The JML framework will
+generate $24 = 3 \times 3 \times 4$ tests, the combination of the data
+ranges involved (the data range of the parameters and the data range
+of the type owning the method under test). But not all the generated
+tests are executed.
 
-\todo{compare hand made, many calls, vs generated, single call, tests}
+A generated test, as it was previously mentioned (see section
+\ref{sec:test_cases}), is not executed, unless the preconditions of
+the called method are met. Thus, depending on the preconditions
+specified for a method, the number of the generated tests that are
+actually executed in the test suite can be significantly reduced.
 
-\todo{describe measures: coverage, number of new bugs found}
+To compare in a quantitative way the effectiveness of the test suites,
+code coverage metrics have been used; this is not the only measure
+that have been considered, since it is almost unanimously accepted
+that code coverage alone cannot asses the quality of a test suite
+\todo{reference to: ``How to misuse code coverage'', ??? ``Coverage
+  Metrics for Formal Verification''}.
 
-\todo{compare compare coupled approach vs single approaches}
+Various coverage metrics exist, we analyzed the test suites with
+statement code coverage (one of the simplest types)\footnote{The tool
+  used to obtain statement code coverage metrics is
+  \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage
+reports whether each executable statement has been encountered. This
+code coverage metric is easy to implement, but lacks the capability to
+explore the control flow graph associated to the code, that is,
+decision branches are not explored. The statement code coverage
+results are reported in the Table \ref{tab:statement_code_coverage}.
+
+\begin{table}[htbp]
+  \caption{Statement code coverage result, categorized by source package.}
+  \label{tab:statement_code_coverage}
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|}\hline
+      & \textbf{\textit{hand}} & \textbf{\textit{generated}} &
+      \textbf{\textit{total}} \\\hline
+      \textbf{\textit{Driver}} & $15.9 \%$ & $6.3 \%$ & $15.9 \%$ \\\hline
+      \textbf{\textit{StreamDriver}} & $76 \%$ & $79.7 \%$ & $88.5 \%$ \\\hline
+      \textbf{\textit{SimulatorDriver}} & $64.9 \%$ & $44.2 \%$ &
+      $71.2 \%$ \\\hline
+      \textbf{\textit{Utility}} & $98.1 \%$ & $74.5 \%$ & $98.1 \%$ \\\hline
+    \end{tabular}
+  \end{center}
+\end{table}
 
@@
 
-The test effectiveness evaluation criteria is of two types.
+The quantitative and qualitative specification and test effectiveness
+measures to verify system under development obtained
+in the previous section need to be analyzed to be fully understood.
 
-\todo{describe quantitative: coverage}
+The coverage measures reported in Table
+\ref{tab:statement_code_coverage} are not high, in absolute
+values. This is because of two different effects: non public methods,
+and interfaces. The effect of non public methods (to be precise:
+package Java visibility) is that these methods are taken into account
+as well as public and protected ones during code coverage calculation;
+but these methods are not part of the interface, they are not meant to
+be called by the system, and they are only used in object
+initialization or during explicit object setup during unit tests. In
+fact there are package and private code blocks that are not reachable
+by construction from public and protected methods, and this is a
+precise characteristic reflecting the code architecture, meant to ease
+class instances setup and configuration for testing purposes. The used
+statement coverage tool cannot be configured to distinguish these code
+blocks, so this effect is unavoidable. The effect is mainly seen in
+low \lil{SimulatorDriver} code coverage figures: the
+\lil{SimulatorDriver} package has plenty of setup and initialization
+non public methods.
 
-\todo{describe qualitative: errors found, help in design, ...}
+Regarding interfaces: an interface contains no statements, so when an
+interface method is called, it is the method of the implementation
+class used that is actually covered. This effect is visible in
+\lil{Driver} coverage figures, \lil{Driver} package contains
+interfaces (that are not counted at all) and some very simple real
+class (exceptions) that are not throughly tested: the net result is
+extremely low statement code coverage figures because the very simple
+classes are the only ones that are actually covered. This effect could
+be easily avoided writing specific hand made tests for the simple
+classes and including them in the classes for which tests are
+generated.
+
+The statement code coverage figures shows that the test suites do a
+good job on the most important package, that is, the
+\lil{StreamDriver}. The generated test suit reaching almost $ 80 \% $,
+while the hand made tests have only a slightly lower coverage ratio: $
+76 \% $. But the combination of test suites is interesting,
+raising the coverage ratio to $ 88.5 \% $. This is a clear indication
+that one test suite is not completely overlapping the other; that is,
+the test suites are not to be considered alternatives, but they have
+to be used together to achive the higher benefits, they are
+complementary. This consideration can be confirmed looking at the
+figures obtained for \lil{SimulatorDriver} package as well.
+
+The qualitative difference of the two test suites can be understood
+analysing how the suites are built. The hand made test suite primarly
+focuses on small features and use cases fragments that the system must
+implement, or for which an exceptional case has been considered. A
+test case is to be considered a rough specification written in
+imperative style. The common structure require an object
+initialization, followed by a hopefully not too long sequence of
+public methods to bring one object into a desired state, than the
+method under test is called and the object state is finally inspected
+to verify it is in the expected state. The generated tests have a far
+more limited scope, since only one method call is performed. Thus it
+is difficult to explore complex behaviour. It is true that with a
+complete specification, and specifing large enough data ranges types,
+the genrated tests could teoretically achive the same expressive power
+as the hand made ones, but the price to pay would be expensive: much
+longer an more complex specifications, requiring more time to be
+designed and or updated, and more tests to be run, augmentig the
+already not so short computation time of the generated test suite.
+
+\todo{compare qualitative: hand made, many calls, vs generated, single call, tests}
+
+\todo{compare compare coupled approach vs single approaches}
+
+\todo{comment quantitative: number of bugs found}
+
+\todo{compare qualitative: spec driven dev, test driven dev, test
+  oriented code is an automatic result}
+
+\todo{spec and hand useful to drive, spec and hand and simulator
+  useful to test, spec and generated useful for statement coverage and
+  easy methods, spec useful to test hardware, hand useful to spec,
+  spec useful to hand... test driven spec}
 

### 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: first almost done draft finished, everything smoothed out, only conclusions are missing

@@
 
-\ifpdf
-\usepackage{epstopdf}
+\ifpdf \usepackage{epstopdf}
 \usepackage[pdftex,bookmarks=false,a4paper=false,
-            plainpages=false,naturalnames=true,
-            colorlinks=true,pdfstartview=FitV,
-            linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Vieri del Bianco"]{hyperref}
-\else
+plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,
+linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Joseph
+R. Kiniry and Vieri del Bianco"]{hyperref} \else
 \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
@@
 
-\newcommand{\tablesize}{\footnotesize}
-\newcommand{\eg}{e.g.,\xspace}
-\newcommand{\ie}{i.e.,\xspace}
-\newcommand{\etc}{etc.\xspace}
+\newcommand{\tablesize}{\footnotesize} \newcommand{\eg}{e.g.,\xspace}
+\newcommand{\ie}{i.e.,\xspace} \newcommand{\etc}{etc.\xspace}
 \newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
-\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3 \myhref{#1}{#1}}}
+\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3
+    \myhref{#1}{#1}}}
 
@@
 
-%\newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:} #1}}
-\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:} #1}}
-\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:} #1}}
-\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:} #1}}
+% \newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:}
+%     #1}}
+\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:}
+    #1}}
+\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}
+    #1}}
+\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:}
+    #1}}
 
@@
 \newcommand{\ST}{\emph{SenseTile}\xspace}
-\newcommand{\SB}{\emph{Sensor Board}\xspace}
-\newcommand{\STSB}{\ST \SB\xspace}
-\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
+\newcommand{\SB}{\emph{Sensor Board}\xspace} \newcommand{\STSB}{\ST
+  \SB\xspace} \newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
 \newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
 \newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
-\newcommand{\STU}{\ST \emph{Unit}\xspace}
-\newcommand{\STUs}{\ST \emph{Unit}\xspace}
+\newcommand{\STU}{\ST \emph{Unit}\xspace} \newcommand{\STUs}{\ST
+  \emph{Unit}\xspace}
 
@@
 \newcommand{\computefarm}{The \ST Scientific Compute Farm\xspace}
-\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute Server
-Farm\xspace}
-\newcommand{\sensorfarm}{The \STSBs\xspace}
-
-%---------------------------------------------------------------------
+\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute
+  Server Farm\xspace} \newcommand{\sensorfarm}{The \STSBs\xspace}
+
+% ---------------------------------------------------------------------
 % New commands, macros, \etc
-%---------------------------------------------------------------------
+% ---------------------------------------------------------------------
 
@@
 
-%=====================================================================
+% =====================================================================
 
@@
 
-\author{Joseph R. Kiniry, Vieri del Bianco\\
+\author{Joseph R. Kiniry, Vieri del Bianco and Dragan Stosic\\
   UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
   Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org, vieri.delbianco@gmail.com\\
+  kiniry@acm.org, vieri.delbianco@ucd.ie and dragan.stosic@gmail.com\\
 }
@@
 
-%======================================================================
-%\thispagestyle{empty}
+% ======================================================================
+% \thispagestyle{empty}
 \begin{abstract}
-
-\begin{verbatim}
-
-OBJECTIVE: 
-Specifications used as a blueprint for system development.
-Specifications to generate test suite to validate the developed system.
-Specifications to validate simulator of the developed system.
-Apply to data stream processing: verify the stream is parsed correctly, 
-verify the parsed stream output stream can be simulated.
-
-METHODS: 
-JML model.
-JML tools to generate JUnit test suite.
-JUnit test suite applied to developed system, to verify its correctness.
-JUnit test suite applied to simulator, to verify its correctness.
-
-RESULTS: 
-Proof developed system same behavior as simulator.
-Proof developed system and simulator verify the specifications.
-
-CONCLUSIONS:
-Specifications serves two purposes, and indirectly a third one.
-
-\end{verbatim}
 
@@
   \STSB is an example of such a device, it communicates with an
-  arbitrary computer system, the \STPU (running Linux or Windows, at
-  the moment) over a USB bus using a streaming packet-based protocol.
-  The \STU is a custom-designed, but inexpensive, sensor platform: it
-  is composed by a one or more \STSBs paired with a general-purpose
-  small-scale compute node, the \STPU (including everything from PDAs
-  to powerful touchscreen-based portable computers).  There are over a
-  dozen sensors on the \STSB itself, and new sensors can be added
-  either using expansion slots on the \STSB, or adding them directly
-  to the USB bus.  A third-party low-level library, provided by the
-  USB chip vendor, the USB chip that has been used in the \STSB,
-  creates and parses generic USB communication packets.
+  arbitrary computer system, the \STPU over a USB bus using a
+  streaming packet-based protocol.  The \STU is a custom-designed, but
+  inexpensive, sensor platform: it is composed by a one or more \STSBs
+  paired with a general-purpose small-scale compute node, the \STPU
+  (including everything from PDAs to powerful touchscreen-based
+  portable computers).  There are over a dozen sensors on the \STSB
+  itself, and new sensors can be added either using expansion slots on
+  the \STSB, or adding them directly to the USB bus.  A third-party
+  low-level library, provided by the USB chip vendor, the USB chip
+  that has been used in the \STSB, creates and parses generic USB
+  communication packets.
 
@@
   streaming communication packets.  The driver is formally specified,
-  validated and verified with rigorous unit testing.  Because of this
-  process and combination of technologies, using this specification we
-  are able to both communicate with a real \ST \textbf{and} generate a
-  formally specified, executable, functional model-based specification
-  for the protocol itself written in the Java Modeling Language.
-  \notev{executable? how? what does this mean?}  This protocol model
-  is itself an executable stream simulator that can be used to develop
-  new \ST user applications without having the \ST hardware at all.
-  Consequently, from this single specification, we can test that the
-  \ST hardware communicates as expected, test that our \ST simulator
-  communicates as expected, and simulate the \ST protocol stream
-  itself.
+  validated and verified with rigorous unit testing, both hand made
+  and automatically generated.  Because of this process and
+  combination of technologies, using this specification we are able to
+  both verify a real \ST and a simulated one.  The system have been
+  developed concurrently with the protocol specification and the real
+  \STSB, the board were not available until the last months of
+  development.  The protocol model is itself an executable stream
+  simulator that can be used to develop new \ST user applications
+  without having the \ST hardware at all.  Consequently, from this
+  single specification, we can test that the \ST hardware communicates
+  as expected, test that our \ST simulator communicates as expected,
+  and simulate the \ST protocol stream itself.
 
@@
 
-%======================================================================
+% ======================================================================
 \section{Introduction}
-
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
-sensor system to be installed at the University College Dublin in Dublin,
-Ireland. 
-Our facility aims to provide a scalable and extensible 
-infrastructure for performing in-depth investigation of both the specific 
-and general issues in large-scale sensor networking.
+\label{sec:introduction}
+
+The \textbf{UCD CASL SenseTile System} is a large-scale,
+general-purpose sensor system to be installed at the University
+College Dublin in Dublin, Ireland.  Our facility aims to provide a
+scalable and extensible infrastructure for performing in-depth
+investigation of both the specific and general issues in large-scale
+sensor networking.
 
 This system integrates a sensor platform, a datastore, and a compute
-farm.  
-The sensor platform is a custom-designed, but inexpensive,
+farm.  The sensor platform is a custom-designed, but inexpensive,
 sensor platform (called the \ST) paired with general-purpose
 small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers. 
-There are over a dozen
+powerful touchscreen-based portable computers.  There are over a dozen
 sensors on the \ST itself, and new sensors can be added by adding them
@@
 
-Because of hard time constraints, and the initial unavailability of the 
-custom board, we needed to progress concurrently with all the development tasks: 
-the specification of the communication protocol with the \STSB, the development of 
-the controlling code in the \STSB, the development of the driver to communicate 
-with the \STSB, and the development of the simulators needed for future tests 
-of the system (the controlling code of the board was developed by the external 
-manufacturer of the board, hence, it is not taken into account here).
-
-\todo{explain and refine the problem}
-
-\todo{references: search other approaches to similar problems}
-
-To tackle the complexity of the concurrent tasks the formal model of the protocol 
-was used as a focus point that led all the other development tasks. 
-The formal model of the protocol was used to check at run time the correctness of 
-the /SB, the driver, the simulator, and the hand made test suites of the driver 
-and the simulator.
-The formal model was used to automatically generate a test suite to verify that
-both the driver and the simulator were compliant to the specification, and eventually 
-to verify the specification itself.
-
-The approach proved itself effective, at the point that, when the board was finally 
-released, no defects were found in the driver or the simulator, while the errors and 
-not agreed changes introduced in the board where discovered during the first execution.
+Because of hard time constraints, and the initial unavailability of
+the custom board, we needed to progress concurrently with all the
+development tasks: the specification of the communication protocol
+with the \STSB, the development of the controlling code in the \STSB,
+the development of the driver to communicate with the \STSB, and the
+development of the \STSB simulators needed for future tests of the
+system (the board itself with the embedded software controlling it was
+developed by an external manufacturer, hence, it is not directly taken
+into account here).
+
+There exist various approaches in literature addressing this kind of
+development constraints (rapid development in hardware software
+codesign), but they are all focused on big scale systems.  The
+hardware software formal codesign methodologies usually have a common
+structure\cite{Slomka2000,Hoffman2001}:
+
+\begin{itemize}
+\item High level specification of the system.
+\item Translation of the high level specification to lower level
+  specifications: hardware and software counterparts.
+\item Generation of software code.
+\item Run hardware specification into a simulator.
+\item Build hardware component.
+\end{itemize}
+
+Even when cosidering the development of the device hardware, embedded
+software and software driver as separate entities that possibly have
+to be developed concurrently, the approaches proposed are similar in
+structure to the one seen in the case of hardware software
+codesign\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
+on very accurate formal specification that drive all the other
+development tasks.
+
+This approach is not feasible or convenient in our case: we want to
+develop all the software components concurrently, we want to enhance
+the protocol specification incrementally; we are not able to provide a
+complete specification from the beginning, we are not interested in
+direct code generation from specification, we want to generate
+simulators to be used to feed the remaining system, even if the \STSB
+is not available, and we want to be able to verify the \STSB behavior
+when it is available.
+
+To tackle the complexity of the concurrent tasks the formal model of
+the protocol was used as a focus point that led all the other
+development tasks (in this part, the approach is similar to the one
+used in rapid development in hardware software codesign, as mentioned
+above).  The protocol to specify has been divided in various (thin)
+layers, each of the layer has been specified with a different
+(lighweight) approach, some of the layers are completely formalized.
+The formal model of the protocol was used to check at run time the
+correctness of the /SB, the driver, the simulator, and the hand made
+test suites of the driver and the simulator.  The formal model was
+used to automatically generate a test suite to verify that both the
+driver and the simulator were compliant to the specification, and
+eventually to verify the specification itself.
+
+The approach proved itself effective, at the point that, when the
+board was finally released, only one small defect were found in the
+driver or the simulator, while the errors and not agreed changes
+introduced in the board where discovered with no effort.  Besides, the
+simulators were able to feed with correct data the rest of the system
 
@@
 
-In section \ref{sec:data_stream_and_protocol_description} the data stream model 
-and protocol is presented.
-
-In section \ref{sec:data_stream_and_protocol_specification} the data stream model 
-and protocol is specified.
-
-In section \ref{sec:test_cases} the test cases structure is presented, for 
-both the hand made and the generated ones. 
-Both the approaches are analyzed, their pro and contra are discussed, as 
-well as the expected advantages to use them in combination.
-
-In section \ref{sec:test_cases_on_the_job} the execution of the test suites
-is described, the data collected is presented.
-
-In section \ref{sec:test_cases_retrospectives} the data collected is analyzed, 
-both from a qualitative and a quantitative points of view.
-
-In section \ref{sec:test_cases_and_the_scenarios} the different scenarios on 
-which test cases based on specifications can be applied are analyzed, and 
-the resulting benefits it is possible to obtain are detailed.
-
-In section \ref{sec:related_works} related works are presented and analyzed.
+In section \ref{sec:data_stream_and_protocol_description} the data
+stream model and protocol is presented.
+
+In section \ref{sec:data_stream_and_protocol_specification} the data
+stream model and protocol is specified.
+
+In section \ref{sec:test_cases} the test cases structure is presented,
+for both the hand made and the generated ones.  Both the approaches
+are analyzed, their pro and contra are discussed, as well as the
+expected advantages to use them in combination.
+
+In section \ref{sec:test_cases_on_the_job} the execution of the test
+suites is described, the results are presented.
+
+In section \ref{sec:test_cases_retrospectives} the data collected and
+the results are analyzed, both from a qualitative and a quantitative
+points of view.
+
+In section \ref{sec:related_works} related works are presented and
+analyzed.
 
@@
 
-%=====================================================================
+% =====================================================================
 \section{Data stream and protocol description}
@@
 
-The communication between \STSB and \STPU is realized through an Universal 
-Serial Bus (USB) interface.
-On top of the USB serial channels a proprietary protocol is used, specified by 
-FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI is the 
-manufacturer of the USB controller chip used in the \STSB, the 
+The communication between \STSB and \STPU is realized through an
+Universal Serial Bus (USB) interface.  On top of the USB serial
+channels a proprietary protocol is used, specified by
+FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
+is the manufacturer of the USB controller chip used in the \STSB, the
 FT2232H\cite{ftdi_ft232h_2009}.
 
-The proprietary protocol enables the control and management of some of the 
-functionalities in FT232H chip (such as internal EEPROM read and write, reset 
-the chip, etc.) as well as the definition of the communication protocol and 
-properties to be used in the asynchronous channels (input and output channels) 
-provided by FT232H.
-These added functionalities are accessed via an Application Program Interface,
-the API is provided by a driver developed directly by FTDI, which is available 
-for various systems (Linux 32 and 64 bits, Mac OS, Windows)
+The proprietary protocol enables the control and management of some of
+the functionalities in FT232H chip (such as internal EEPROM read and
+write, reset the chip, etc.) as well as the definition of the
+communication protocol and properties to be used in the asynchronous
+channels (input and output channels) provided by FT232H.  These added
+functionalities are accessed via an Application Program Interface, the
+API is provided by a driver developed directly by FTDI, which is
+available for various systems (Linux 32 and 64 bits, Mac OS, Windows)
 \cite{ftdi_d2xx_api_2009}.
 
-Finally, on top of the serial channels provided by the API, a custom protocol 
-has been developed to manage the entire \STSB; that is, read sensor data 
-observations, enable or disable sensors, configure on-board components 
-as the Field-Programmable Gate Array (FPGA), and the micro processor; the 
-protocol main function is to read sensor data.
+Finally, on top of the serial channels provided by the API, a custom
+protocol has been developed to manage the entire \STSB; that is, read
+sensor data observations, enable or disable sensors, configure
+on-board components as the Field-Programmable Gate Array (FPGA), and
+the micro processor; the protocol main function is to read sensor
+data.
 
 The protocol is an asynchronous fixed length packed based protocol.
-The protocol has been divided in multiple layers to ease its formal description.
-The rationale is the following: mixing up different abstraction levels of a 
-protocol description can lead to overly complex and unmanageable descriptions, 
-especially when dealing with formal ones.
-The layers identified in protocol are the following:
+The protocol has been divided in multiple layers to ease its formal
+description.  The rationale is the following: mixing up different
+abstraction levels of a protocol description can lead to overly
+complex and unmanageable descriptions, especially when dealing with
+formal ones.  The layers identified in protocol are the following:
 
 \begin{description}
- \item[Packet byte structure]: the internal representation of the packet.
- \item[Packet info structure]: the meaningful fields contained in the packets.
- \item[Single packet rules]: the content acceptable values, and how they 
-influence each other.
- \item[Packet sequence rules]: the content acceptable values, based on values 
-of previously received packets.
- \item[Packet input output asynchronous rules]: the content acceptable values 
-and reaction constraints on output packets, based on previously transmitted 
-input packets.
+\item[Packet byte structure]: the internal representation of the
+  packet.
+\item[Packet info structure]: the meaningful fields contained in the
+  packets.
+\item[Single packet rules]: the content acceptable values, and how
+  they influence each other.
+\item[Packet sequence rules]: the content acceptable values, based on
+  values of previously received packets.
+\item[Packet input output asynchronous rules]: the content acceptable
+  values and reaction constraints on output packets, based on
+  previously transmitted input packets.
 \end{description}
 
-We will only examine the output sensor data packet, since it is the more used 
-and interesting one.
-We will examine the first three layers, that is: the packet byte structure, the 
-packet info structure and the single packet rules.
-
-The packet byte and info structures of the output sensor data packet reflects the board 
-capabilities and sensors.
-Basically a single packet has to accommodate various types of data:
+We will only examine the output sensor data packet, since it is the
+more used and interesting one.  We will examine the first three
+layers, that is: the packet byte structure, the packet info structure
+and the single packet rules.
+
+The packet byte and info structures of the output sensor data packet
+reflects the board capabilities and sensors.  Basically a single
+packet has to accommodate various types of data:
 \begin{itemize}
@@
 \item Medium data rate streams: 5KHz.
- \item Slow data rate streams: temperature, pression, accelerometer, etc.; all 
- these sensor have a data rate far lower than 5 KHz.
+\item Slow data rate streams: temperature, pression, accelerometer,
+  etc.; all these sensor have a data rate far lower than 5 KHz.
 \item Metadatas: metadatas on sensors and board state.
@@
 
-Even if we have to deal with data streams that have different data rates, the 
-internal structure of the package (as we will see) is strictly fixed. 
-This choice originates from a trade-off between two constraints: the efficiency 
-of the transmission (redundant data to be minimized) and the complexity to build 
-a package (complexity to be minimized).
-The complexity minimization constraint rationale is the following: the FPGA 
-on the \STSB builds the packet, but the FPGA is also to be used for various 
-other specific tasks; it is important to maintain the building packet code on the 
-FPGA as simple and short as possible, to leave enough space to implement other 
-tasks.
-The Output sensor data packet structure is showed in Table 
-\ref{tab:data_packet_structure}, the packet total length is 1024 bytes (512 16 
-bit words).
+Even if we have to deal with data streams that have different data
+rates, the internal structure of the package (as we will see) is
+strictly fixed.  This choice originates from a trade-off between two
+constraints: the efficiency of the transmission (redundant data to be
+minimized) and the complexity to build a package (complexity to be
+minimized).  The complexity minimization constraint rationale is the
+following: the FPGA on the \STSB builds the packet, but the FPGA is
+also to be used for various other specific tasks; it is important to
+maintain the building packet code on the FPGA as simple and short as
+possible, to leave enough space to implement other tasks.  The Output
+sensor data packet structure is showed in Table
+\ref{tab:data_packet_structure}, the packet total length is 1024 bytes
+(512 16 bit words).
 
@@
 The frames accommodate data from the fast and medium data rate streams
-(there are 4 fast data rate stream and 8 medium data rate stream channels). 
-There are 82 frames in a packet.
-The frame structure is shown in Table \ref{tab:subsequence_structure}.
+(there are 4 fast data rate stream and 8 medium data rate stream
+channels).  There are 82 frames in a packet.  The frame structure is
+shown in Table \ref{tab:subsequence_structure}.
 
@@
 
-Looking at the structure described and at the data rates frequencies, it
-can be seen that there must be something subtle, to make all the figures to 
-add up, that is, to be capable to accommodate three different rates outputs
-in a fixed structure, fixed length packet.
-The subtleties lies in two forms of redundancies, for slow and medium data 
-rates:
+Looking at the structure described and at the data rates frequencies,
+it can be seen that there must be something subtle, to make all the
+figures to add up, that is, to be capable to accommodate three
+different rates outputs in a fixed structure, fixed length packet.
+The subtleties lies in two forms of redundancies, for slow and medium
+data rates:
 
@@
 \item Slow data rates: slow data can be found at the beginning of each
-  packet, the data is simply repeated (that is, it is equal to the previous 
-  packet in the stream) when no new data is available.
-  \item Medium data rates: the 8 medium rate channels are multiplexed together 
-  into the medium data rate data sample of each subsequence, which channel the 
-  sample belongs to is a piece of info to be extracted from the subsequence 
-  metadata. Besides, the medium data rate data sample could contain no data at 
-  all, this info again is to be extracted from the metadata.
+  packet, the data is simply repeated (that is, it is equal to the
+  previous packet in the stream) when no new data is available.
+\item Medium data rates: the 8 medium rate channels are multiplexed
+  together into the medium data rate data sample of each subsequence,
+  which channel the sample belongs to is a piece of info to be
+  extracted from the subsequence metadata. Besides, the medium data
+  rate data sample could contain no data at all, this info again is to
+  be extracted from the metadata.
 \end{itemize}
 
-The single packet rules delimit the boundaries of the values obtainable from 
-the packet: each defined sensor represented in the packet, as well as the 
-metadata describing the \STSB and the packet and frame contents,  have a defined 
-range of acceptable values.
-There are also rules affecting more than one value, basically validity rules (i.e. if 
-a specific sensor is active than the reading value has to be in a defined range), 
-and rules specifying a correct sequence of frames (i.e. there can be only one switch 
-in sampling rate in the frame sequence of a single packet).
-
-%=====================================================================
+The single packet rules delimit the boundaries of the values
+obtainable from the packet: each defined sensor represented in the
+packet, as well as the metadata describing the \STSB and the packet
+and frame contents, have a defined range of acceptable values.  There
+are also rules affecting more than one value, basically validity rules
+(i.e. if a specific sensor is active than the reading value has to be
+in a defined range), and rules specifying a correct sequence of frames
+(i.e. there can be only one switch in sampling rate in the frame
+sequence of a single packet).
+
+% =====================================================================
 \section{Data stream and protocol specification}
@@
 
-The specification of the protocol has been used as a focus point to lead all the 
-other development tasks.
+The specification of the protocol has been used as a focus point to
+lead all the other development tasks.
 
@@
 \begin{itemize}
- \item Be executed at run time. If something on the protocol is not respected, 
-the run time execution environment is capable to detect and signal the fault.
+\item Be executed at run time. If something on the protocol is not
+  respected, the run time execution environment is capable to detect
+  and signal the fault.
 \item Generate tests.
@@
 
-We also needed a specification capable to tackle properly all the protocol layers.
-Since the different layers have very different characteristics, different context
-dependent specification approaches were chosen.
+We also needed a specification capable to tackle properly all the
+protocol layers.  Since the different layers have very different
+characteristics, different context dependent specification approaches
+were chosen.
 
@@
 
-The specification is provided by a reference Java implementation capable of 
-recognizing the proper packet structure in a byte data stream. The implementation 
-is tricky, since it checks for the sentinels repeatedly to be sufficiently confidant
-that the byte stream can be properly decomposed into a packet stream.
-
-The implementation can be obviously executed at run time (the specification is the 
-implementation code). 
-Meaningful automatic tests cannot be generated. 
-The implementation can be used to verify a packet byte structure simulator. 
-The implementation were actually driven (Test Driven Development) by the test suite.
+The specification is provided by a reference Java implementation
+capable of recognizing the proper packet structure in a byte data
+stream. The implementation is tricky, since it checks for the
+sentinels repeatedly to be sufficiently confidant that the byte stream
+can be properly decomposed into a packet stream.
+
+The implementation can be obviously executed at run time (the
+specification is the implementation code).  Meaningful automatic tests
+cannot be generated.  The implementation can be used to verify a
+packet byte structure simulator.  The implementation were actually
+driven (Test Driven Development) by the test suite.
 
@@
 
-The specification is provided both by the Java API that has been built, and by the 
-annotation with JML language\cite{JML2}\footnote{
-\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}}. 
-JML has been chosen because it is capable to specify Java code: JML is a behavioral 
-interface specification language. 
-JML specification language combines the design by contract approach (applied to 
-Java language) and the model-based specification approach of interface specification 
-languages.
-
-Many tools are available supporting JML specification language, focusing on different 
-aspects of the language. 
-JML2 suite of tools\cite{BurdyEtal05-STTT}
-\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}}
-was chosen, since both run time verification support and test generation support are 
-provided directly by the suite.
-
-The Java API specification part is checked automatically by the Java compiler, while
-the JML specification part can be checked at run time using the provided JML2 run time 
-environment. 
-Meaningful automatic tests can be generated from the JML specification. 
-The JML specification can be used to verify an abstract packet generation simulator, 
-making the simulator an extension of the Java interfaces. 
-The specification can also be used to verify the real driver implementation, based on 
-the packet byte structure implementation presented above. 
-The implementation was driven by the JML specification and by an hand made test suite.
-Finally, hand made unit tests can be checked, during execution, against the 
-specification, using the provided JML2 run time environment.
+The specification is provided both by the Java API that has been
+built, and by the annotation with JML language\cite{JML2}\footnote{
+  \myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling
+    Language (JML)}}.  JML has been chosen because it is capable to
+specify Java code: JML is a behavioral interface specification
+language.  JML specification language combines the design by contract
+approach (applied to Java language) and the model-based specification
+approach of interface specification languages.
+
+Many tools are available supporting JML specification language,
+focusing on different aspects of the language.  JML2 suite of
+tools\cite{BurdyEtal05-STTT}
+\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common
+    JML tools}} was chosen, since both run time verification support
+and test generation support are provided directly by the suite.
+
+The Java API specification part is checked automatically by the Java
+compiler, while the JML specification part can be checked at run time
+using the provided JML2 run time environment.  Meaningful automatic
+tests can be generated from the JML specification.  The JML
+specification can be used to verify an abstract packet generation
+simulator, making the simulator an extension of the Java interfaces.
+The specification can also be used to verify the real driver
+implementation, based on the packet byte structure implementation
+presented above.  The implementation was driven by the JML
+specification and by an hand made test suite.  Finally, hand made unit
+tests can be checked, during execution, against the specification,
+using the provided JML2 run time environment.
 
@@
 
-The specification is provided exclusively by annotations with JML language.
-
-The capabilities of a JML specification have been already described in the previous 
-paragraph, that is: specification checkable at run time, tests generation, simulator 
-verification, usable with hand written tests.
+The specification is provided exclusively by annotations with JML
+language.
+
+The capabilities of a JML specification have been already described in
+the previous paragraph, that is: specification checkable at run time,
+tests generation, simulator verification, usable with hand written
+tests.
 
@@
 
-Simulators are needed to test in isolation parts of the system; especially the upper 
-layers, yet to be developed. 
-During the development the simulators were essential, since the board was not available.
-Simulators were built concurrently with the specification, the real driver code, and the 
-tests.
-
-Theoretically a separate simulator should be needed for each protocol layer; a
-separate simulator for each layer would enable us to separately specify, develop and test
-each layer of the protocol.
-In our case we evaluated not worth the effort to maintain such a separation in the 
-simulators, motivated by the fact that the protocol layers identified are too fine grained 
-to be separated into different simulators. 
-Two simulators were built.
-
-The two simulators were built reflecting the software abstraction layers defined for the 
-driver.
-The driver API is splitted in two structural layers: the general high level interfaces, that 
-expose the main functionalities of the board and the main contents of the packets hiding the 
-details, and the lower level implementation that parse the data streams in and out the board, 
-to translate the higher level instructions and data in properly formed packets.
-Hence, we have implemented two simulators: one high level simulator, which re-implements the 
-high level interfaces providing corresponding high level mechanisms to control the data and 
-the behavior of the simulated board, it has nothing to do with the details of the data 
-streams.
-One low level simulator which is capable to rebuild the data stream out of the board based on 
-the data stream into the board, the in and out data streams are built exactly as the sensor 
-board is expected to be able to parse or generate.
+Simulators are needed to test in isolation parts of the system;
+especially the upper layers, yet to be developed.  During the
+development the simulators were essential, since the board was not
+available.  Simulators were built concurrently with the specification,
+the real driver code, and the tests.
+
+Theoretically a separate simulator should be needed for each protocol
+layer; a separate simulator for each layer would enable us to
+separately specify, develop and test each layer of the protocol.  In
+our case we evaluated not worth the effort to maintain such a
+separation in the simulators, motivated by the fact that the protocol
+layers identified are too fine grained to be separated into different
+simulators.  Two simulators were built.
+
+The two simulators were built reflecting the software abstraction
+layers defined for the driver.  The driver API is splitted in two
+structural layers: the general high level interfaces, that expose the
+main functionalities of the board and the main contents of the packets
+hiding the details, and the lower level implementation that parse the
+data streams in and out the board, to translate the higher level
+instructions and data in properly formed packets.  Hence, we have
+implemented two simulators: one high level simulator, which
+re-implements the high level interfaces providing corresponding high
+level mechanisms to control the data and the behavior of the simulated
+board, it has nothing to do with the details of the data streams.  One
+low level simulator which is capable to rebuild the data stream out of
+the board based on the data stream into the board, the in and out data
+streams are built exactly as the sensor board is expected to be able
+to parse or generate.
 
@@
 
-The JML specifications used to annotate Java code are of varying complexities.
-Some of the specifications are simple, focusing on constraints that should hold when 
-calling a method (preconditions) and the constraints on the return value (a very basic form of 
-postconditions).
-In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the 
-method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that 
-it cannot change the state of any object (a postcondition), the specification also constraints 
-the return value with a lower and upper bound (another postcondition).
+The JML specifications used to annotate Java code are of varying
+complexities.  Some of the specifications are simple, focusing on
+constraints that should hold when calling a method (preconditions) and
+the constraints on the return value (a very basic form of
+postconditions).  In listing \ref{lst:jml_example_simple} a simple JML
+specification example is shown, the method \lil{getTemperature} is
+declared to be \lil{/*@ pure @*/}, which means that it cannot change
+the state of any object (a postcondition), the specification also
+constraints the return value with a lower and upper bound (another
+postcondition).
 
@@
   float=htbp]
-/*@ 
-  ensures \result >= -880;
-  ensures \result <= 2047;
-@*/
-/*@ pure @*/ short getTemperature();
+  /*@ ensures \result >= -880; ensures \result <= 2047; @*/ /*@ pure
+  @*/ short getTemperature();
 \end{lstlisting}
@@
 
-Some of the specifications are rather complex, usually focusing on properties regarding the 
-behavior of a whole object.
-In listing \ref{lst:jml_example_invariant} a complex invariant 
-(a property that has to be maintained, that is, that is assumed on entry and guaranteed on 
-exit of each method of an object)
-example is shown. 
-The invariant is constraining the number of samples for each medium data rate streams: 
-the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is 
-\lil{FRAMES}, the constant constraining the number of samples is 
-\lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are
-fairly distributed on the channels. 
-The valid samples are counted parsing all the frames contained in a packet, selecting only 
-the matching valid samples.
-A medium data rate stream sample is considered valid when \lil{isADCActive()} method 
-returns \lil{true}.
+Some of the specifications are rather complex, usually focusing on
+properties regarding the behavior of a whole object.  In listing
+\ref{lst:jml_example_invariant} a complex invariant (a property that
+has to be maintained, that is, that is assumed on entry and guaranteed
+on exit of each method of an object) example is shown.  The invariant
+is constraining the number of samples for each medium data rate
+streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the
+total number of frames is \lil{FRAMES}, the constant constraining the
+number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that
+the samples contained in a frame are fairly distributed on the
+channels.  The valid samples are counted parsing all the frames
+contained in a packet, selecting only the matching valid samples.  A
+medium data rate stream sample is considered valid when
+\lil{isADCActive()} method returns \lil{true}.
 
@@
   float=htbp]
-/*@ 
-  invariant (
-    \forall int channel;
-    0 <= channel && channel < Frame.ADC_CHANNELS; 
-    (
-      \num_of int i;
-      0<= i && i<(FRAMES-1);
-      (
-        getFrame(i).isADCActive() &&
-        (getFrame(i).getADCChannel() == channel)
-      )
-    ) <= (FRAMES / Frame.ADC_CHANNELS + 1)
-  );
-@*/
+  /*@ invariant ( \forall int channel; 0 <= channel && channel <
+  Frame.ADC_CHANNELS; ( \num_of int i; 0<= i && i<(FRAMES-1); (
+  getFrame(i).isADCActive() && (getFrame(i).getADCChannel() ==
+  channel) ) ) <= (FRAMES / Frame.ADC_CHANNELS + 1) ); @*/
 \end{lstlisting}
 
-%=====================================================================
+% =====================================================================
 \section{Test cases}
@@
 compared to hand based tests or completely automated tests used alone,
-testing a driver for a stream communication protocol with hard constraints 
-on development time.
-\ST stream communication protocol is characterized by a well definite 
-packet structure, and a well definite semantic on the packet contents 
-as well.
-On the other hand in the protocol there are not strong constraints on 
-synchronization, since the communication channel is splitted in two (in and 
-out) separate concurrent and asynchronous channels.
-The lack of hard synchronization constraints reduce the advantage to use 
-a classical state machine based specification \todo{references: 
-protocol tests generated from a state machine based specification}.
-
-Hard development time constraints were also a precise requirement; the time 
-constraints pushed us towards a incomplete (\emph{just good enough}) 
-specification.
-In \STSB driver JML was used to specify very simple behaviors (method 
-oriented) on the whole driver implementation, and more complex behavior on 
-the driver core interfaces.
-The software system parts not specified by the JML2 specification were 
-specified using hand made unit tests.
+testing a driver for a stream communication protocol with hard
+constraints on development time.  \ST stream communication protocol is
+characterized by a well definite packet structure, and a well definite
+semantic on the packet contents as well.  On the other hand in the
+protocol there are not strong constraints on synchronization, since
+the communication channel is splitted in two (in and out) separate
+concurrent and asynchronous channels.  The lack of hard
+synchronization constraints reduce the advantage to use a classical
+state machine based specification (as the one presented in
+\cite{Hubbers2004}).
+
+Hard development time constraints were also a precise requirement; the
+time constraints pushed us towards a incomplete (\emph{just good
+  enough}) specification.  In \STSB driver JML was used to specify
+very simple behaviors (method oriented) on the whole driver
+implementation, and more complex behavior on the driver core
+interfaces.  The software system parts not specified by the JML2
+specification were specified using hand made unit tests.
 
 Hand made unit tests cover behavior not covered by the automatically
-generated tests, that is, they cover the not specified behavior; but this 
-is not the only purpose they have.
-Hand made unit tests are used to drive the development, since the Test Driven 
-Development approach\cite{beck2003test} has been followed during development.
-Hand made tests are also used to drive the development of the most error 
-prone or complex JML2 specifications; a test developed to drive the 
-specification is a test that should fail if the specification is not working 
-as expected.
-
-\todo{describe the necessity to add specification generated tests: to ensure
-that specification is sound, to obtain higher coverage}
+generated tests, that is, they cover the not specified behavior; but
+this is not the only purpose they have.  Hand made unit tests are used
+to drive the development, since the Test Driven Development
+approach\cite{beck2003test} has been followed during development.
+Hand made tests are also used to drive the development of the most
+error prone or complex JML2 specifications; a test developed to drive
+the specification is a test that should fail if the specification is
+not working as expected.
 
@@
 The package structure of the driver developed is shown in Figure
-\ref{fig:class_diagram_main}: two independent packages has been defined: 
-\lil{Stream} and \lil{Driver}.
-The packages are abstract, that is, they mainly contain abstractions; in Java 
-language this is translated into a package which contains mainly abstract classes 
-or interfaces.
-The two abstract packages has various implementations, the dependency between
-abstract packages is not direct, but only through an implementation 
-(\lil{StreamDriver}).
-This is needed to maintain a high decoupling of the packages, and is the result of 
+\ref{fig:class_diagram_main}: two independent packages has been
+defined: \lil{Stream} and \lil{Driver}.  The packages are abstract,
+that is, they mainly contain abstractions; in Java language this is
+translated into a package which contains mainly abstract classes or
+interfaces.  The two abstract packages has various implementations,
+the dependency between abstract packages is not direct, but only
+through an implementation (\lil{StreamDriver}).  This is needed to
+maintain a high decoupling of the packages, and is the result of
 applying the \emph{classical} dependency inversion
 principle\cite{martin1996dependency} (the dependency structure of the
-packages has a saw shape, a common shape on systems where the dependency 
-inversion principle has been followed).
+packages has a saw shape, a common shape on systems where the
+dependency inversion principle has been followed).
 
@@
 The test package structure, which is shown in Figure
-\ref{fig:class_diagram_test} reflects and mimics the code structure; the
-tests that test an abstract package are abstract, to be implemented by the 
-package that tests a corresponding system implementation; this is a well 
-known test pattern, the Abstract Test Pattern\cite{thomas2004java}, it is 
-used to test that the contracts are respected in the implementations.
-For instance, package \lil{DriverT} contains abstract tests for the 
-abstractions of package \lil{Driver}, package \lil{StreamDriverT} 
-inherits the abstractions contained in \lil{DriverT} and makes them concrete, 
-to test the corresponding implementation \lil{StreamDriver}; package 
-\lil{StreamDriverT} also contains stand alone tests written specifically for
-the implementation contained in \lil{StreamDriver}.
+\ref{fig:class_diagram_test} reflects and mimics the code structure;
+the tests that test an abstract package are abstract, to be
+implemented by the package that tests a corresponding system
+implementation; this is a well known test pattern, the Abstract Test
+Pattern\cite{thomas2004java}, it is used to test that the contracts
+are respected in the implementations.  For instance, package
+\lil{DriverT} contains abstract tests for the abstractions of package
+\lil{Driver}, package \lil{StreamDriverT} inherits the abstractions
+contained in \lil{DriverT} and makes them concrete, to test the
+corresponding implementation \lil{StreamDriver}; package
+\lil{StreamDriverT} also contains stand alone tests written
+specifically for the implementation contained in \lil{StreamDriver}.
 
@@
 
-The JML specification has been used to generate the tests. 
-A specific tool of the JML2 suite can be used to automatically generate 
+The JML specification has been used to generate the tests.  A specific
+tool of the JML2 suite can be used to automatically generate
 them\cite{Cheon-Leavens02}.
@@
 The test suite generated is a collection of fine grained unit tests.
-For each public method a set of tests is generated.
-The specified behavior of the method is checked using the JML2 run time 
-assertion checker, hence, an oracle is not needed. 
-The generated tests send messages to the method under test, they catch 
-assertion violation exceptions from test cases that pass an initial 
-precondition check; that is, if a precondition is not met, the test is 
-not run, if the precondition is met the test is run, and if the postcondition 
-is not me an exception is fired and the test fails.
-The test data that is used to build the tests is the only part that is 
-required to be provided by the user.
+For each public method a set of tests is generated.  The specified
+behavior of the method is checked using the JML2 run time assertion
+checker, hence, an oracle is not needed.  The generated tests send
+messages to the method under test, they catch assertion violation
+exceptions from test cases that pass an initial precondition check;
+that is, if a precondition is not met, the test is not run, if the
+precondition is met the test is run, and if the postcondition is not
+me an exception is fired and the test fails.  The test data that is
+used to build the tests is the only part that is required to be
+provided by the user.
 
 The test package structure is shown in Figure
-\ref{fig:class_diagram_generatedtest}.
-The packages dependencies are defined by the specifications; for instance, 
-\lil{SimulatorDriverST} are generated tests, they test (hence use) the 
-implementation \lil{SimulatorDriver}, they have been generated using the 
-specification corresponding specification \lil{SimulatorDriverS}.
-The generated tests do not maintain the inheritance structure of the code 
+\ref{fig:class_diagram_generatedtest}.  The packages dependencies are
+defined by the specifications; for instance, \lil{SimulatorDriverST}
+are generated tests, they test (hence use) the implementation
+\lil{SimulatorDriver}, they have been generated using the
+specification corresponding specification \lil{SimulatorDriverS}.  The
+generated tests do not maintain the inheritance structure of the code
 and the specification.
@@
 
-To have proper unit test suite, for each method under test a proper stub, 
-driver and oracle should be identified, these are the usual components found 
-in the harness that enable test execution\cite{binder1999testing}.
-In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework used 
-to manage the implementation and execution of hand made and generated unit 
-tests, there is no clear distinction between drivers, stubs, and oracle of a 
-test; noneless, even if a proper distinction is not imposed by the framework, 
-the conceptual distinction still holds.
-
-Hence, to properly unit test a module in isolation and automatically we need:
+To have proper unit test suite, for each method under test a proper
+stub, driver and oracle should be identified, these are the usual
+components found in the harness that enable test
+execution\cite{Binder1999}.  In
+JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework
+used to manage the implementation and execution of hand made and
+generated unit tests, there is no clear distinction between drivers,
+stubs, and oracle of a test; none-less, even if a proper distinction
+is not imposed by the framework, the conceptual distinction still
+holds.
+
+Hence, to properly unit test a module in isolation and automatically
+we need:
 
 \begin{description}
- \item[Stub or Skeleton]: a piece of code that simulates the activity of 
-the modules that are not under tests.
- \item[Driver]: a piece of code that calls in a proper way the module under 
-test.
- \item[Oracle]: a mechanism used for determining whether a test has passed 
-or failed, comparing the output of the module under test to the output that 
-the oracle determines that the module should have.
+\item[Stub or Skeleton]: a piece of code that simulates the activity
+  of the modules that are not under tests.
+\item[Driver]: a piece of code that calls in a proper way the module
+  under test.
+\item[Oracle]: a mechanism used for determining whether a test has
+  passed or failed, comparing the output of the module under test to
+  the output that the oracle determines that the module should have.
 \end{description}
 
-Hand made unit tests and JML generated tests differs in how these components 
-are provided. 
-The distinctions and similarities are summarized in Table 
-\ref{tab:test_harness}.
+Hand made unit tests and JML generated tests differs in how these
+components are provided.  The distinctions and similarities are
+summarized in Table \ref{tab:test_harness}.
 
@@
 
-A module simulator can be used to ease the tasks of developing each component 
-of a test harness; that is, it can be used as a stub, to simulate a module 
-needed by the module under test, it can be used as a driver, to stimulate the 
-module under test, it can be used as an oracle, if it simulates the module 
-under test.
-
-In our system, the simulators has been used mainly as stubs, so that we have,
-during testing, a structure like the one shown in Figure 
-\ref{fig:class_diagram_streamdriver_test}.
-In Figure \ref{fig:class_diagram_streamdriver_test} the test suite 
-\lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate 
-the \lil{Stream} that is used by the system under test, that is 
-\lil{StreamDriver}.
+A module simulator can be used to ease the tasks of developing each
+component of a test harness; that is, it can be used as a stub, to
+simulate a module needed by the module under test, it can be used as a
+driver, to stimulate the module under test, it can be used as an
+oracle, if it simulates the module under test.
+
+In our system, the simulators has been used mainly as stubs, so that
+we have, during testing, a structure like the one shown in Figure
+\ref{fig:class_diagram_streamdriver_test}.  In Figure
+\ref{fig:class_diagram_streamdriver_test} the test suite
+\lil{StreamDriverT} is using the \lil{SimulatorStream} to properly
+simulate the \lil{Stream} that is used by the system under test, that
+is \lil{StreamDriver}.
 
@@
   \label{fig:class_diagram_streamdriver_test}
- \caption{StreamDriver test packages class diagram: use of the stream simulator.}
+  \caption{StreamDriver test packages class diagram: use of the stream
+    simulator.}
 \end{figure}
 
-In our case, the usage of simulators to enable unit testing, was essential to 
-test the \STSB driver, since the physical board was not available.
-The simulator were used in both the hand made and the generated tests, since
-both of them need a proper set of stubs to be executed.
-
-\todo{insert references to testing embedded systems with simulators and/or 
-with driver/stubs}
-
-\paragraph*{BON}
-
-\todo{describe approach: high level informal specifications with BON, 
-iterative refinements of code, to distill test cases and formal 
-specification of the interfaces} \notev{remove, BON has not been used 
-properly, maybe a follow up paper}
-
-%=====================================================================
+In our case, the usage of simulators to enable unit testing, was
+essential to test the \STSB driver, since the physical board was not
+available.  The simulator were used in both the hand made and the
+generated tests, since both of them need a proper set of stubs to be
+executed.
+
+% =====================================================================
 \section{Test cases on the job}
@@
 
-JML specifications, the simulators,  and both test suites has been used to:
+JML specifications, the simulators, and both test suites has been used
+to:
 
@@
 The first \STSB prototype delivered was meant to build empty packets,
-but the packets were built in a way not conformant to the
+but the packets were built in a way not consistent with the
 specification, the sentinel values used to delimit a packet were not
@@
 detection of errors on the \SB itself.  Errors regarding the expected
-values and values ranges of the filds contained in the packets and the
-frames (4 protocol errors), and errors regarding the dependencies
+values and values ranges of the fields contained in the packets and
+the frames (4 protocol errors), and errors regarding the dependencies
 between values in field (2 protocol errors) were detected. No other
@@
 The hand made test suite is composed by over 140 tests, while the
-generated one is 100 times bigger, totalling over 14000 tests.
+generated one is 100 times bigger, totaling over 14000 tests.
 
@@
 that have been considered, since it is almost unanimously accepted
-that code coverage alone cannot asses the quality of a test suite
-\todo{reference to: ``How to misuse code coverage'', ??? ``Coverage
-  Metrics for Formal Verification''}.
+that code coverage alone cannot asses the quality of a test
+suite\cite{Marick1999,Chockler2006}.
 
@@
 
-%=====================================================================
+Execution time are very different, the hand made test suite, with the
+run time assertion checking active runs, completely in less the one
+minute on an average PC, while the generated test suite runs in more
+than 20 minutes (the rough ratio is $ 1 : 65 $). Most of the time is
+spent on setting up the suite, only considering the real execution of
+the tests, with no suite setup, the execution time reduces to less
+than 4 minutes (the rough ratio is $ 1 : 10 $).
+
+% =====================================================================
 \section{Retrospective on test cases effectiveness}
@@
 The quantitative and qualitative specification and test effectiveness
-measures to verify system under development obtained
-in the previous section need to be analyzed to be fully understood.
+measures to verify system under development obtained in the previous
+section need to be analyzed to be fully understood.
 
@@
 while the hand made tests have only a slightly lower coverage ratio: $
-76 \% $. But the combination of test suites is interesting,
-raising the coverage ratio to $ 88.5 \% $. This is a clear indication
-that one test suite is not completely overlapping the other; that is,
-the test suites are not to be considered alternatives, but they have
-to be used together to achive the higher benefits, they are
-complementary. This consideration can be confirmed looking at the
-figures obtained for \lil{SimulatorDriver} package as well.
+76 \% $. But the combination of test suites is interesting, raising
+the coverage ratio to $ 88.5 \% $. This is a clear indication that one
+test suite is not completely overlapping the other; that is, the test
+suites are not to be considered alternatives, but they have to be used
+together to achieve the higher benefits, they are complementary. This
+consideration can be confirmed looking at the figures obtained for
+\lil{SimulatorDriver} package as well.
 
 The qualitative difference of the two test suites can be understood
-analysing how the suites are built. The hand made test suite primarly
+analyzing how the suites are built. The hand made test suite primarily
 focuses on small features and use cases fragments that the system must
-implement, or for which an exceptional case has been considered. A
-test case is to be considered a rough specification written in
-imperative style. The common structure require an object
-initialization, followed by a hopefully not too long sequence of
-public methods to bring one object into a desired state, than the
-method under test is called and the object state is finally inspected
-to verify it is in the expected state. The generated tests have a far
-more limited scope, since only one method call is performed. Thus it
-is difficult to explore complex behaviour. It is true that with a
-complete specification, and specifing large enough data ranges types,
-the genrated tests could teoretically achive the same expressive power
+implement, or for which an exceptional case has been specified. A test
+case is to be considered a rough specification written in imperative
+style. The common structure require an object initialization, followed
+by a hopefully not too long sequence of public methods to bring one
+object (or set of objects) into a desired state, than the method under
+test is called and the object state is finally inspected to verify it
+is in the expected state. Mock objects can be used to limit the
+dependencies. The test suite should be composed only by unit tests,
+but usually some tests are more complex to be called proper unit
+tests, they can be seen as simple integration tests.
+
+The generated test suite has a more focused and limited scope, since
+only one method call is performed in each test, hence it is difficult
+to explore complex behaviors. It is true that with a complete
+specification, and specifying large enough data ranges (the
+appropariate data ranges are combined to generate the tests), the
+generated tests could theoretically achieve the same expressive power
 as the hand made ones, but the price to pay would be expensive: much
-longer an more complex specifications, requiring more time to be
-designed and or updated, and more tests to be run, augmentig the
+longer and more complex specifications, requiring more time to be
+designed and or updated, and more tests to be run, augmenting the
 already not so short computation time of the generated test suite.
 
-\todo{compare qualitative: hand made, many calls, vs generated, single call, tests}
-
-\todo{compare compare coupled approach vs single approaches}
-
-\todo{comment quantitative: number of bugs found}
-
-\todo{compare qualitative: spec driven dev, test driven dev, test
-  oriented code is an automatic result}
-
-\todo{spec and hand useful to drive, spec and hand and simulator
-  useful to test, spec and generated useful for statement coverage and
-  easy methods, spec useful to test hardware, hand useful to spec,
-  spec useful to hand... test driven spec}
-
-%=====================================================================
-\section{Test cases and the scenarios}
-\label{sec:test_cases_and_the_scenarios}
-
-\todo{benefit: the approach can be used to test different part of the system
-that share the same specifications}
-
-\todo{testing the protocols layers}
-
-\todo{testing the simulators}
-
-\todo{describe general approach: depending at what level (of abstraction) tests 
-and specifications are written, it is possible to verify both the simulator and 
-the real protocol (isn't this obvious?)}
-
-%=====================================================================
+The test suites gave their best when used in a combined approach: the
+generated test suite, checking most simple behaviors, while the hand
+made checking the more complex ones.  During the development of the
+driver code and the simulators, the generated test suit was helpful to
+test the incrementally growing specification and the finer details of
+each method, while the hand made one was helpful to verify the
+behaviour on complex methods calls on objects intialized to specific
+states. Having an hand made test suite also make the code style more
+test oriented, which usually leads to a better quality and more
+modularized source code\cite{Binder1999}.
+
+The test suites has been used to verify different parts of the
+systems, that is, the real system as well as the simulators.  The
+multiple simulators and multiple test suites approach is not new, it
+is often used in embedded systems software testing\cite{Broekman2002}.
+This is a general approach that can (and should) be used to verify
+both the simulators built and the real systems.  This approach can be
+easily used only if the code has been structured following the already
+mentioned dependency inversion principle\cite{martin1996dependency}
+(see Figure \ref{fig:class_diagram_main}).
+
+The combined test approach, driven by the packet protocol
+specification, prooved succesfull on verifing the protocol itself and
+the software implementing it.  Neverthless, it has to be noted that we
+did not verified the highest levels of the protocol (see
+\ref{sec:data_stream_and_protocol_description}: ``Packet sequence
+rules'' and ``Packet input output asynchronous rules'').  Since on
+these levels the protocol properties regards sequences of packets,
+testing them can be tricky, since the sequences of packets are not
+referred to explicitly in the implementation.  Two possible strategies
+can be used:
+
+\begin{itemize}
+\item Translate the packet sequence properties in local packet
+  properties (when this is possible).  The advantage is to have
+  properties shown as the usual precondition, post conditions and
+  invariants.  But the localized properties can be more complex and
+  less intuitive and comprehensible than the original ones; finally,
+  some of the global properties could proove themselves impossible to
+  be translated into local ones.
+\item Use separate utility classes, meant to verify a sequence of
+  packages.  This approach is not clean, since it requires adding and
+  using new classes, unneeded by the core implementation, to verify
+  the protocol.  It has the advantage of keeping the properties that
+  have to be verified simple.
+\end{itemize}
+
+One problem we noticed regards the executiuon time of the generated
+tests.  The executiuon time is long for two reasons:
+
+\begin{itemize}
+\item Number of the tests generated. This problem is not directly
+  addressable, since it is the heart of the test suite generation
+  creation method itself: generating all the combinations of possible
+  inputs for each method under test.
+\item Runtime assertion checker. Each test is executed while RAC
+  checks all the relavant properties. This problem, again, is not
+  directly addressable. But, we can speculate that if the RAC
+  assertions could be triggered on and off on single properties in a
+  dynamic way, that is, during runtime, a test suite could use take
+  advantage of this possibility, turning on the RAC only on the
+  properties that have to be verified for the classes or methods under
+  tests (and not for all the involved classes and methods); this could
+  save time during the test suite execution.
+\end{itemize}
+
+% =====================================================================
 \section{Related works}
@@
 
-%=====================================================================
+JML has been used previously to verify a communication protocol in
+\cite{Hubbers2004}, but the protocol is a smart card protocol, and it
+is well represented through state machines, which are then refined
+into JML properties.  The approach is in waterfall style, the need to
+develop the protocol, the specification and the implementation
+concurrently is not addressed.  No hand made or generated test suites
+are used, the code is specified with JML assertions, refined by the
+state machine specifications, and then checked statically with
+ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2
+tools RAC\cite{BurdyEtal05-STTT}.
+
+The JML-JUnit automatic test suite generation tool has not been widely
+used or assessed. It is described by its own creators in
+\cite{Cheon2002,Cheon2004,Cheon2005}, it has been used in
+\cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test
+suite, and, more importantly, only against proof of concepts system,
+not real complex systems.  A brief assessment of the JML-JUnit tool
+has been made in \cite{Tan2004}, the analysis is based on injecting
+random generated errors in the source code, and then evaluating the
+ratio of the errors found by the test suite: the results shows that
+the JML-JUnit tool is not effective enough if used as a stand alone
+tool. No attempts on running or evaluating the generated test suite
+alongside with an hand made test suite are made.
+
+% =====================================================================
 \section{Conclusion}
@@
 
-%======================================================================
+\todo{specification and hand tests useful to drive, specification and
+  hand tests and simulator useful to test, specification and generated
+  tests useful for statement coverage and easy methods, specification
+  useful to test hardware, hand useful to spec, spec useful to
+  hand... test driven spec}
+
+\section{Next papers}
+
+\todo{remove and push this on a new paper}
+
+\paragraph*{TDS: Test Driven Specification}
+
+Approach:
+\begin{itemize}
+\item more specification used together: formal and not formal
+\item formal specification is not complete: it is just good
+  ``enough'', static checkers unfortunately cannot be used, since the
+  specification is partial
+\item use runtime asserion checker for formal specifications, on tests
+\item use generated tests for formal specifications
+\end{itemize}
+
+Who drives what:
+\begin{itemize}
+\item specification (jml - formal specification or unit test -
+  examples fragments) driven development
+\item test driven specification
+\item formal specification is not complete: it is just good
+  ``enough'', static checkers unfortunately cannot be used, since the
+  specification is partial
+\item use runtime asserion checker for formal specifications, on tests
+\item use generated tests for formal specifications
+\end{itemize}
+
+Benefits:
+\begin{itemize}
+\item real agile approach
+\item real formal approach
+\item use best of both worlds, do not try to use only on of them
+\end{itemize}
+
+Case example: actual paper, modified to fit the new parts, expecially
+the test driven specification part.
+
+\paragraph*{BON}
+
+\todo{describe approach: high level informal specifications with BON,
+  iterative refinements of code, to distill test cases and formal
+  specification of the interfaces}
+
+% ======================================================================
 %% \nocite{ex1,ex2}
-\bibliographystyle{latex8}
-\bibliography{extra,%
+\bibliographystyle{latex8} \bibliography{extra,%
   abbrev,%
@@
 
-%======================================================================
+% ======================================================================
 % Fin

### 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: syntax errors corrections

@@
 development constraints (rapid development in hardware software
-codesign), but they are all focused on big scale systems.  The
-hardware software formal codesign methodologies usually have a common
+co-design), but they are all focused on big scale systems.  The
+hardware software formal co-design methodologies usually have a common
 structure\cite{Slomka2000,Hoffman2001}:
@@
 
-Even when cosidering the development of the device hardware, embedded
+Even when considering the development of the device hardware, embedded
 software and software driver as separate entities that possibly have
@@
 structure to the one seen in the case of hardware software
-codesign\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
+co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
 on very accurate formal specification that drive all the other
@@
 development tasks (in this part, the approach is similar to the one
-used in rapid development in hardware software codesign, as mentioned
+used in rapid development in hardware software co-design, as mentioned
 above).  The protocol to specify has been divided in various (thin)
 layers, each of the layer has been specified with a different
-(lighweight) approach, some of the layers are completely formalized.
+(lightweight) approach, some of the layers are completely formalized.
 The formal model of the protocol was used to check at run time the
@@
 specification, and specifying large enough data ranges (the
-appropariate data ranges are combined to generate the tests), the
+appropriate data ranges are combined to generate the tests), the
 generated tests could theoretically achieve the same expressive power
@@
 each method, while the hand made one was helpful to verify the
-behaviour on complex methods calls on objects intialized to specific
+behavior on complex methods calls on objects initialized to specific
 states. Having an hand made test suite also make the code style more
@@
 The combined test approach, driven by the packet protocol
-specification, prooved succesfull on verifing the protocol itself and
-the software implementing it.  Neverthless, it has to be noted that we
+specification, proved successful on verifying the protocol itself and
+the software implementing it.  Nevertheless, it has to be noted that we
 did not verified the highest levels of the protocol (see
@@
   less intuitive and comprehensible than the original ones; finally,
-  some of the global properties could proove themselves impossible to
+  some of the global properties could prove themselves impossible to
   be translated into local ones.
@@
 
-One problem we noticed regards the executiuon time of the generated
-tests.  The executiuon time is long for two reasons:
+One problem we noticed regards the execution time of the generated
+tests.  The execution time is long for two reasons:
 
@@
 \item Runtime assertion checker. Each test is executed while RAC
-  checks all the relavant properties. This problem, again, is not
+  checks all the relevant properties. This problem, again, is not
   directly addressable. But, we can speculate that if the RAC
@@
   specification is partial
-\item use runtime asserion checker for formal specifications, on tests
+\item use runtime assertion checker for formal specifications, on tests
 \item use generated tests for formal specifications
@@
 
-Case example: actual paper, modified to fit the new parts, expecially
+Case example: actual paper, modified to fit the new parts, especially
 the test driven specification part.

### 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: syntax errors

@@
   float=htbp]
-  /*@ ensures \result >= -880; ensures \result <= 2047; @*/ /*@ pure
-  @*/ short getTemperature();
+  /*@ 
+    ensures \result >= -880; 
+    ensures \result <= 2047; 
+  @*/ 
+  /*@ pure @*/ short getTemperature();
 \end{lstlisting}
@@
   float=htbp]
-  /*@ invariant ( \forall int channel; 0 <= channel && channel <
-  Frame.ADC_CHANNELS; ( \num_of int i; 0<= i && i<(FRAMES-1); (
-  getFrame(i).isADCActive() && (getFrame(i).getADCChannel() ==
-  channel) ) ) <= (FRAMES / Frame.ADC_CHANNELS + 1) ); @*/
+  /*@ 
+    invariant ( 
+      \forall int channel; 0 <= channel && channel < Frame.ADC_CHANNELS; ( 
+        \num_of int i; 0<= i && i<(FRAMES-1); (
+          getFrame(i).isADCActive() && 
+          (getFrame(i).getADCChannel() == channel) 
+        ) 
+      ) <= (FRAMES / Frame.ADC_CHANNELS + 1) 
+    ); 
+  @*/
 \end{lstlisting}

### 2009-10-14 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
ModelBasedTestCaseGenerationDataStreamProtocol paper: revisione

--- 2009-10-12 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
@@
   packet-based network protocol.  The communication is focused on
-  controlling the sensors and report the sensors measurements.  The
-  \STSB is an example of such a device, it communicates with an
-  arbitrary computer system, the \STPU over a USB bus using a
+  controlling the sensors and reporting the sensors measurements.  The
+  \STSB is an example of such a device: it communicates with an
+  arbitrary computer system, the \STPU, over a USB bus using a
   streaming packet-based protocol.  The \STU is a custom-designed, but
@@
   the \STSB, or adding them directly to the USB bus.  A third-party
-  low-level library, provided by the USB chip vendor, the USB chip
-  that has been used in the \STSB, creates and parses generic USB
-  communication packets.
+  low-level library creates and parses generic USB communication
+  packets; the library has been provided by the vendor of the USB chip
+  vendor, the USB chip that has been used in the \STSB.
 
@@
   and automatically generated.  Because of this process and
-  combination of technologies, using this specification we are able to
-  both verify a real \ST and a simulated one.  The system have been
-  developed concurrently with the protocol specification and the real
-  \STSB, the board were not available until the last months of
-  development.  The protocol model is itself an executable stream
-  simulator that can be used to develop new \ST user applications
-  without having the \ST hardware at all.  Consequently, from this
-  single specification, we can test that the \ST hardware communicates
-  as expected, test that our \ST simulator communicates as expected,
-  and simulate the \ST protocol stream itself.
+  combination of technologies, we are able to verify a real \ST and a
+  simulated one as well.  The system has been developed concurrently
+  with the protocol specification and the real \STSB, the boards have
+  not been available until the last months of development.  The
+  protocol model is itself an executable stream simulator that can be
+  used to develop new \ST user applications without having the \ST
+  hardware at all.  Consequently, from this single specification, we
+  are able to test that the \ST hardware communicates as expected,
+  that our \ST simulator communicates as expected, and simulate the
+  \ST protocol stream itself.
 
@@
 powerful touchscreen-based portable computers.  There are over a dozen
-sensors on the \ST itself, and new sensors can be added by adding them
-to the board.
+sensors packaged on the \ST itself, the board is expandable as well,
+new sensors can be added to it easily.
 
@@
 the custom board, we needed to progress concurrently with all the
-development tasks: the specification of the communication protocol
-with the \STSB, the development of the controlling code in the \STSB,
-the development of the driver to communicate with the \STSB, and the
-development of the \STSB simulators needed for future tests of the
-system (the board itself with the embedded software controlling it was
-developed by an external manufacturer, hence, it is not directly taken
-into account here).
+development tasks:
+
+\begin{itemize}
+\item specification of the communication protocol;
+\item development of the physical board specification and following
+  realization;
+\item development of the board embedded software;
+\item development of the software driver to communicate with the
+  board;
+\item development of the software simulators of the board.
+\end{itemize}
+
+The development of the physical board, along with its embedded
+software, has been done by an external manufacturer, thus it is not
+directly taken into account here.
 
@@
 development constraints (rapid development in hardware software
-co-design), but they are all focused on big scale systems.  The
+co-design), but they are all focused on big-scale systems.  The
 hardware software formal co-design methodologies usually have a common
@@
 Even when considering the development of the device hardware, embedded
-software and software driver as separate entities that possibly have
+software and software driver as separate entities, that possibly have
 to be developed concurrently, the approaches proposed are similar in
@@
 co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
-on very accurate formal specification that drive all the other
-development tasks.
-
-This approach is not feasible or convenient in our case: we want to
-develop all the software components concurrently, we want to enhance
-the protocol specification incrementally; we are not able to provide a
-complete specification from the beginning, we are not interested in
-direct code generation from specification, we want to generate
-simulators to be used to feed the remaining system, even if the \STSB
-is not available, and we want to be able to verify the \STSB behavior
-when it is available.
-
-To tackle the complexity of the concurrent tasks the formal model of
-the protocol was used as a focus point that led all the other
-development tasks (in this part, the approach is similar to the one
-used in rapid development in hardware software co-design, as mentioned
-above).  The protocol to specify has been divided in various (thin)
-layers, each of the layer has been specified with a different
-(lightweight) approach, some of the layers are completely formalized.
-The formal model of the protocol was used to check at run time the
-correctness of the /SB, the driver, the simulator, and the hand made
-test suites of the driver and the simulator.  The formal model was
-used to automatically generate a test suite to verify that both the
-driver and the simulator were compliant to the specification, and
-eventually to verify the specification itself.
+on a specification that aims to be as complete as possible. The
+specification, once it is finalized, is used to define the remaining
+development tasks, through automatic code generation, refinement,
+model checking, etc.
+
+This approach has been considered not feasible nor convenient if
+applied to our case since it does not respect some of the
+requirements: first of all, we want to develop all the software
+components concurrently, we want to enhance the protocol specification
+incrementally and we are not able to provide a complete specification
+from the beginning, hence, a properly defined specification is not
+available in advance. Secondly, we are not interested in direct code
+generation from specification, since we do not want to have any
+dependencies on code generators; in fact the programming language to
+be used to develop the driver has been postponed for a while.
+
+To tackle the complexity of the concurrent development tasks, the
+formal model of the protocol was used as a focus point that led all
+the other development tasks. As it can be seen, the central role of
+the specification is unchanged; what is changed is how and when the
+specification is developed.
+
+The protocol to specify has been divided in various (thin) layers,
+each of the layer has been specified with a different lightweight
+approach: some of the layers have been specified formally, while
+others have been specified in an unprecise way.  The formal
+specifications have been used to check at run time the correctness of
+the \SB, the driver and the simulators.  The formal specifications
+have also been used to automatically generate a test suite to verify
+the driver and the simulators, and eventually to verify the
+specification itself.  An hand made test suite has been used to verify
+the driver and the simulators as well.  The generated and hand made
+test suites have been used together, complementing each other.
 
 The approach proved itself effective, at the point that, when the
-board was finally released, only one small defect were found in the
-driver or the simulator, while the errors and not agreed changes
-introduced in the board where discovered with no effort.  Besides, the
-simulators were able to feed with correct data the rest of the system
+board has been finally released, only one small defect has been found
+in the driver, while the errors and not agreed changes introduced in
+the board and in the protocol it implemented, have been discovered
+immediately, with no additional effort.  Besides, the simulators are
+able to feed with correct data the rest of the system.
 
@@
 In section \ref{sec:data_stream_and_protocol_description} the data
-stream model and protocol is presented.
+stream model and protocol are presented.
 
 In section \ref{sec:data_stream_and_protocol_specification} the data
-stream model and protocol is specified.
+stream model and protocol are specified.
 
 In section \ref{sec:test_cases} the test cases structure is presented,
-for both the hand made and the generated ones.  Both the approaches
-are analyzed, their pro and contra are discussed, as well as the
-expected advantages to use them in combination.
+for both the hand made and the generated test suites.  Both the
+approaches are analyzed, their pro and contra are discussed, as well
+as the expected advantages to use them in combination.
 
 In section \ref{sec:test_cases_on_the_job} the execution of the test
-suites is described, the results are presented.
-
-In section \ref{sec:test_cases_retrospectives} the data collected and
-the results are analyzed, both from a qualitative and a quantitative
-points of view.
-
-In section \ref{sec:related_works} related works are presented and
-analyzed.
-
-Finally in section \ref{sec:conclusion} conclusions are drawn and
-ongoing and future works on similar topics is introduced.
+suites is described; the results and measures obtained by their
+execution are presented.
+
+In section \ref{sec:test_cases_retrospectives} the data collected in
+the previous section are analyzed, both from a qualitative and a
+quantitative points of view. The effectiveness of the approach is
+discussed as well.
+
+In section \ref{sec:related_works} the related works that have not
+been mentioned in the current section are presented and analyzed.
+
+Finally in section \ref{sec:conclusion} conclusions are drawn, and
+ongoing and future works on similar topics are introduced.
 
@@
 
-The communication between \STSB and \STPU is realized through an
-Universal Serial Bus (USB) interface.  On top of the USB serial
-channels a proprietary protocol is used, specified by
+\STSB and \STPU communicates via an Universal Serial Bus (USB)
+interface.  A proprietary protocol is used on top of the USB serial
+channels, the protocol is specified by
 FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
@@
 the functionalities in FT232H chip (such as internal EEPROM read and
-write, reset the chip, etc.) as well as the definition of the
+write, reset of the chip, etc.), as well as the definition of the
 communication protocol and properties to be used in the asynchronous
-channels (input and output channels) provided by FT232H.  These added
+channels (input and output channels) provided by FT232H.  These
 functionalities are accessed via an Application Program Interface, the
@@
 
-Finally, on top of the serial channels provided by the API, a custom
-protocol has been developed to manage the entire \STSB; that is, read
-sensor data observations, enable or disable sensors, configure
-on-board components as the Field-Programmable Gate Array (FPGA), and
-the micro processor; the protocol main function is to read sensor
-data.
-
-The protocol is an asynchronous fixed length packed based protocol.
-The protocol has been divided in multiple layers to ease its formal
-description.  The rationale is the following: mixing up different
-abstraction levels of a protocol description can lead to overly
-complex and unmanageable descriptions, especially when dealing with
-formal ones.  The layers identified in protocol are the following:
+Finally, a custom protocol has been developed to manage the entire
+\STSB, built on top of the serial channels provided by the API.  The
+custom protocol is meant to be used to read sensor data observations,
+enable or disable sensors, configure on-board components (as the
+Field-Programmable Gate Array and the micro processor); the protocol
+main function is to read sensor data.
+
+The protocol is asynchronous and packet based, the packets have a
+fixed length.  The protocol has been divided in multiple layers to
+ease its formal description.  The rationale is the following: mixing
+up different abstraction levels of a protocol description can lead to
+overly complex and unmanageable descriptions, especially when dealing
+with formal ones.  The layers identified in protocol are the
+following:
 
@@
 
-We will only examine the output sensor data packet, since it is the
-more used and interesting one.  We will examine the first three
-layers, that is: the packet byte structure, the packet info structure
-and the single packet rules.
+We will examine the output sensor data packet, since it is the most
+used and interesting one; we won't describe input or other output
+packets.  We will examine the first three layers, that is: the packet
+byte structure, the packet info structure and the single packet rules.
 
 The packet byte and info structures of the output sensor data packet
-reflects the board capabilities and sensors.  Basically a single
-packet has to accommodate various types of data:
-\begin{itemize}
-\item Fast data rate streams: audio data, 48KHz or 96KHz.
+reflect the board capabilities and built in sensors.  Essentially a
+single packet has to accommodate various types of data:
+
+\begin{itemize}
+\item Fast data rate streams: audio data, 48KHz (standard audio) or
+  96KHz (ultra audio).
 \item Medium data rate streams: 5KHz.
 \item Slow data rate streams: temperature, pression, accelerometer,
-  etc.; all these sensor have a data rate far lower than 5 KHz.
+  etc.; all these sensor have a data rate lower than 5 KHz.
 \item Metadatas: metadatas on sensors and board state.
@@
 
-Even if we have to deal with data streams that have different data
-rates, the internal structure of the package (as we will see) is
-strictly fixed.  This choice originates from a trade-off between two
-constraints: the efficiency of the transmission (redundant data to be
-minimized) and the complexity to build a package (complexity to be
-minimized).  The complexity minimization constraint rationale is the
-following: the FPGA on the \STSB builds the packet, but the FPGA is
-also to be used for various other specific tasks; it is important to
-maintain the building packet code on the FPGA as simple and short as
-possible, to leave enough space to implement other tasks.  The Output
-sensor data packet structure is showed in Table
-\ref{tab:data_packet_structure}, the packet total length is 1024 bytes
-(512 16 bit words).
+The internal structure of the package (as we will see) is strictly
+fixed, even if the data streams have different data rates.  This
+choice originates from a trade-off between two constraints: the
+efficiency of the transmission (redundant data to be minimized) and
+the complexity to build a package (complexity to be minimized).  The
+complexity minimization constraint rationale is the following: the
+FPGA on the \STSB builds the packet, but the FPGA is also to be used
+for various other specific tasks; it is important to maintain the
+building packet code on the FPGA as simple and short as possible, to
+leave enough space to implement other tasks.  The Output sensor data
+packet structure is shown in Table \ref{tab:data_packet_structure},
+the packet total length is 1024 bytes (512 16 bit words).
 
@@
       \ldots & \ldots\\\hline
-      503-508 & subsequence 82\\\hline
+      503-508 & frame 82\\\hline
     \end{tabular}
@@
 channels).  There are 82 frames in a packet.  The frame structure is
-shown in Table \ref{tab:subsequence_structure}.
+shown in Table \ref{tab:frame_structure}.
 
 \begin{table}[htbp]
-  \caption{Output sensor data packet subsequence structure.}
-  \label{tab:subsequence_structure}
+  \caption{Output sensor data packet frame structure.}
+  \label{tab:frame_structure}
   \begin{center}
@@
       \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0 & metadata: subsequence description\\\hline
+      0 & metadata: frame description\\\hline
       1 & medium data rate data sample\\\hline
@@
 
-Looking at the structure described and at the data rates frequencies,
-it can be seen that there must be something subtle, to make all the
-figures to add up, that is, to be capable to accommodate three
+Looking at the structure described and the data rates frequencies, it
+can be seen that there must be some subtle rule, to make all the
+figures to add up; that is, to be capable to accommodate three
 different rates outputs in a fixed structure, fixed length packet.
@@
 \item Medium data rates: the 8 medium rate channels are multiplexed
-  together into the medium data rate data sample of each subsequence,
-  which channel the sample belongs to is a piece of info to be
-  extracted from the subsequence metadata. Besides, the medium data
-  rate data sample could contain no data at all, this info again is to
-  be extracted from the metadata.
+  together into the medium data rate data sample of each frame.  The
+  medium data rate sample channel is a piece of info to be extracted
+  from the frame description. Besides, the medium data rate data
+  sample could contain no data at all, this information is to be
+  extracted from the frame description. So, one frame could contain a
+  medium data rate sample from channel 0, the following frame a sample
+  from channel 5, the following frame no sample at all, etc.
 \end{itemize}
@@
 packet, as well as the metadata describing the \STSB and the packet
-and frame contents, have a defined range of acceptable values.  There
-are also rules affecting more than one value, basically validity rules
-(i.e. if a specific sensor is active than the reading value has to be
-in a defined range), and rules specifying a correct sequence of frames
-(i.e. there can be only one switch in sampling rate in the frame
-sequence of a single packet).
+and frame contents, are constrained by a range of acceptable values.
+There are also rules affecting more than one value, these are
+essentially validity rules (i.e. if a specific sensor is active the
+reading value has to be in a defined range), and rules specifying a
+correct sequence of frames (i.e. there can be only one switch in
+sampling rate in the frame sequence of a single packet).
 
@@
 
-The specification of the protocol has been used as a focus point to
+The specifications of the protocol has been used as a central point to
 lead all the other development tasks.
 
-We needed a specification capable of:
-
-\begin{itemize}
-\item Be executed at run time. If something on the protocol is not
+Regarding the formal specification used, we needed a specification
+capable of:
+
+\begin{itemize}
+\item Be executed at run time: if something on the protocol is not
   respected, the run time execution environment is capable to detect
@@
 
-We also needed a specification capable to tackle properly all the
-protocol layers.  Since the different layers have very different
-characteristics, different context dependent specification approaches
-were chosen.
+We needed specifications capable to tackle properly all the protocol
+layers.  Since different layers have different characteristics,
+different specification approaches has been chosen for each layer: the
+specification are context dependent.
 
@@
 
-The specification is provided by a reference Java implementation
-capable of recognizing the proper packet structure in a byte data
-stream. The implementation is tricky, since it checks for the
-sentinels repeatedly to be sufficiently confidant that the byte stream
-can be properly decomposed into a packet stream.
-
-The implementation can be obviously executed at run time (the
-specification is the implementation code).  Meaningful automatic tests
-cannot be generated.  The implementation can be used to verify a
-packet byte structure simulator.  The implementation were actually
-driven (Test Driven Development) by the test suite.
+The specification is provided by a hand made test suite, composed of
+unit and integration tests; the tests specifies the behaviors of the
+Java implementation, which is capable of recognizing the proper packet
+structure in a binary data stream. The implementation is tricky: it
+checks for the sentinels repeatedly to be sufficiently confidant that
+the byte stream can be properly decomposed into a packet stream.
+
+Fulfilled requirements:
+
+\begin{itemize}
+\item The specification cannot be checked at run time.
+\item Meaningful automatic tests cannot be generated.
+\item The specification can be used to verify a packet byte structure
+  simulator.
+\item The implementation were actually driven (Test Driven
+  Development) by the test suite specification.
+\end{itemize}
 
@@
 
-The specification is provided both by the Java API that has been
-built, and by the annotation with JML language\cite{JML2}\footnote{
+The specification is provided by JML annotations\cite{JML2}\footnote{
   \myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling
-    Language (JML)}}.  JML has been chosen because it is capable to
-specify Java code: JML is a behavioral interface specification
-language.  JML specification language combines the design by contract
-approach (applied to Java language) and the model-based specification
-approach of interface specification languages.
-
-Many tools are available supporting JML specification language,
+    Language (JML)}} and by a hand made test suite.
+
+JML has been chosen because it is capable to specify Java code: JML is
+a behavioral interface specification language.  JML specification
+language combines the design by contract approach (applied to Java
+language) and the model-based specification approach of interface
+specification languages.
+
+Many tools able to support JML specification language are available,
 focusing on different aspects of the language.  JML2 suite of
@@
 \footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common
-    JML tools}} was chosen, since both run time verification support
-and test generation support are provided directly by the suite.
-
-The Java API specification part is checked automatically by the Java
-compiler, while the JML specification part can be checked at run time
-using the provided JML2 run time environment.  Meaningful automatic
-tests can be generated from the JML specification.  The JML
-specification can be used to verify an abstract packet generation
-simulator, making the simulator an extension of the Java interfaces.
-The specification can also be used to verify the real driver
-implementation, based on the packet byte structure implementation
-presented above.  The implementation was driven by the JML
-specification and by an hand made test suite.  Finally, hand made unit
-tests can be checked, during execution, against the specification,
+    JML tools}} has been chosen, since the suite provides both run
+time verification support and test generation support.
+
+Fulfilled requirements:
+
+\begin{itemize}
+\item The Java API specification part is checked automatically by the
+  Java compiler, while the JML specification part can be checked at
+  run time using the provided JML2 run time environment.
+\item Meaningful automatic tests can be generated from the JML
+  specification.
+\item The JML specification can be used to verify a simulator, which
+  generates packets as the real board would do .The specification can
+  also be used to verify the real driver implementation, based on the
+  packet byte structure implementation presented above.
+\item The implementation was driven by the formal JML specification
+  and by the hand made test suite.  Finally, hand made unit tests can
+  be checked, during execution, against the formal specification,
   using the provided JML2 run time environment.
+\end{itemize}
 
@@
 
-The specification is provided exclusively by annotations with JML
-language.
-
-The capabilities of a JML specification have been already described in
-the previous paragraph, that is: specification checkable at run time,
-tests generation, simulator verification, usable with hand written
-tests.
+The specification is provided exclusively by the formal specification
+with JML language.
+
+The fulfilled requirements of a JML specification have been already
+described in the previous paragraph, that is:
+
+\begin{itemize}
+\item The JML specification is checkable at run time.
+\item Tests can be automatically generated.
+\item Simulators can be verified.
+\item The specification can be used with hand written tests.
+\end{itemize}
 
@@
 
-Simulators are needed to test in isolation parts of the system;
-especially the upper layers, yet to be developed.  During the
-development the simulators were essential, since the board was not
-available.  Simulators were built concurrently with the specification,
-the real driver code, and the tests.
+Simulators can be used to test parts of the system, to minimizing
+their dependencies: a simulator can be used by upper layers, with no
+need to provide the functionalities of the lower layers.  During the
+development of the \STSB the simulators have been essential, since the
+physical board was not available: we needed at least to remove the
+dependencies over the real board to execute the driver implementation.
+Simulators have been built concurrently with the specification, the
+driver implementation, and the tests.
 
 Theoretically a separate simulator should be needed for each protocol
-layer; a separate simulator for each layer would enable us to
-separately specify, develop and test each layer of the protocol.  In
-our case we evaluated not worth the effort to maintain such a
-separation in the simulators, motivated by the fact that the protocol
-layers identified are too fine grained to be separated into different
-simulators.  Two simulators were built.
-
-The two simulators were built reflecting the software abstraction
+layer: this would enable us to specify, develop and test each layer in
+isolation.  In our case we evaluated not worth the effort to replicate
+the layer structure into simulators; the choice was motivated by the
+fact that such a distinction would have not provided us any
+perceivable benefit.  Two simulators were built.
+
+The simulators have been built reflecting the software abstraction
 layers defined for the driver.  The driver API is splitted in two
-structural layers: the general high level interfaces, that expose the
-main functionalities of the board and the main contents of the packets
-hiding the details, and the lower level implementation that parse the
-data streams in and out the board, to translate the higher level
-instructions and data in properly formed packets.  Hence, we have
-implemented two simulators: one high level simulator, which
-re-implements the high level interfaces providing corresponding high
-level mechanisms to control the data and the behavior of the simulated
-board, it has nothing to do with the details of the data streams.  One
-low level simulator which is capable to rebuild the data stream out of
-the board based on the data stream into the board, the in and out data
-streams are built exactly as the sensor board is expected to be able
-to parse or generate.
+abstraction layers: the general high level interfaces, that expose the
+main functionalities of the board and the main contents of the
+packets, and the lower level implementation that parse the data
+streams in and out the board, meant to translate the higher level
+instructions and data in properly formed packets.
+
+The high level simulator implements the interfaces providing the
+methods to deal with an abstracted \SB: the simulator knows nothing
+about the details of the real format of the binary data streams.  The
+low level simulator is capable to rebuild the \SB data streams: the in
+and out data streams are built exactly as the sensor board is expected
+to be able to parse or generate.
 
@@
 
-The JML specifications used to annotate Java code are of varying
+The JML specifications that have been written are of varying
 complexities.  Some of the specifications are simple, focusing on
-constraints that should hold when calling a method (preconditions) and
-the constraints on the return value (a very basic form of
+constraints that should hold when calling a method (the preconditions)
+and constraints on the return value (the very basic form of
 postconditions).  In listing \ref{lst:jml_example_simple} a simple JML
@@
 properties regarding the behavior of a whole object.  In listing
-\ref{lst:jml_example_invariant} a complex invariant (a property that
-has to be maintained, that is, that is assumed on entry and guaranteed
-on exit of each method of an object) example is shown.  The invariant
-is constraining the number of samples for each medium data rate
-streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the
-total number of frames is \lil{FRAMES}, the constant constraining the
-number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that
-the samples contained in a frame are fairly distributed on the
-channels.  The valid samples are counted parsing all the frames
-contained in a packet, selecting only the matching valid samples.  A
-medium data rate stream sample is considered valid when
-\lil{isADCActive()} method returns \lil{true}.
+\ref{lst:jml_example_invariant} a complex invariant example is shown;
+an invariant is a property that has to be maintained during the life
+cycle of an object, more precisly, an invariant is assumed on entry
+and guaranteed on exit of each method of an object.  The invariant is
+constraining the number of samples for each medium data rate streams:
+the total number of streams is \lil{Frame.ADC_CHANNELS}, the total
+number of frames is \lil{FRAMES}, the constant constraining the number
+of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the
+samples contained in a frame are fairly distributed on the channels.
+The valid samples are counted parsing all the frames contained in a
+packet, selecting only the matching valid samples.  A medium data rate
+stream sample is considered valid when \lil{isADCActive()} method
+returns \lil{true}.
 
@@
 The unit test cases that have been built to verify the protocol driver
-are of two kinds: hand made unit tests and unit tests automatically
-generated based on JML specifications.  The tests have been thought to
-be used together, they are complementary.  The tests effectiveness has
-been evaluated, for each kind of test, the evaluation is carried out
+are of two kinds: hand made unit tests and automatically generated
+unit tests based on JML specifications.  The tests have been thought
+to be used together, they are complementary.  The tests effectiveness
+has been evaluated for each test suite; the evaluation is carried out
 in section \ref{sec:test_cases_retrospectives}.  The test
@@
 
-The dual testing approach was thought as a winning option, when
-compared to hand based tests or completely automated tests used alone,
-testing a driver for a stream communication protocol with hard
-constraints on development time.  \ST stream communication protocol is
-characterized by a well definite packet structure, and a well definite
-semantic on the packet contents as well.  On the other hand in the
-protocol there are not strong constraints on synchronization, since
-the communication channel is splitted in two (in and out) separate
-concurrent and asynchronous channels.  The lack of hard
-synchronization constraints reduce the advantage to use a classical
-state machine based specification (as the one presented in
-\cite{Hubbers2004}).
+The dual testing approach was thought a winning option, when compared
+to hand based tests or completely automated tests used alone.  \ST
+stream communication protocol is characterized by a well definite
+packet structure, and a well definite semantic on the packet contents
+as well.  On the other hand, in the protocol, there are not strong
+constraints on synchronization, since the communication channel is
+splitted in two (in and out) separate concurrent and asynchronous
+channels.  The lack of hard synchronization constraints reduces the
+advantage to use a classical state machine based specification (as the
+one presented in \cite{Hubbers2004}), hence this option has been ruled
+out.
 
 Hard development time constraints were also a precise requirement; the
-time constraints pushed us towards a incomplete (\emph{just good
+time constraints pushed us towards an incomplete (\emph{just good
   enough}) specification.  In \STSB driver JML was used to specify
@@
 
-Hand made unit tests cover behavior not covered by the automatically
-generated tests, that is, they cover the not specified behavior; but
-this is not the only purpose they have.  Hand made unit tests are used
-to drive the development, since the Test Driven Development
-approach\cite{beck2003test} has been followed during development.
-Hand made tests are also used to drive the development of the most
-error prone or complex JML2 specifications; a test developed to drive
-the specification is a test that should fail if the specification is
-not working as expected.
+Hand made unit tests specifies the behavior not specified by formal
+specification; nevertheless this is not the only purpose they have.
+Hand made unit tests have been used to drive the development, since
+the Test Driven Development approach\cite{beck2003test} has been
+followed during development.
 
@@
 \ref{fig:class_diagram_main}: two independent packages has been
-defined: \lil{Stream} and \lil{Driver}.  The packages are abstract,
+defined (\lil{Stream} and \lil{Driver}).  The packages are abstract,
 that is, they mainly contain abstractions; in Java language this is
 translated into a package which contains mainly abstract classes or
-interfaces.  The two abstract packages has various implementations,
-the dependency between abstract packages is not direct, but only
+interfaces.  The two abstract packages have various implementations,
+and the dependency between the two is not direct, since it is realized
 through an implementation (\lil{StreamDriver}).  This is needed to
 maintain a high decoupling of the packages, and is the result of
-applying the \emph{classical} dependency inversion
-principle\cite{martin1996dependency} (the dependency structure of the
+applying the classical dependency inversion
+principle\cite{martin1996dependency}: the dependency structure of the
 packages has a saw shape, a common shape on systems where the
-dependency inversion principle has been followed).
+dependency inversion principle has been followed.
 
@@
 \ref{fig:class_diagram_test} reflects and mimics the code structure;
-the tests that test an abstract package are abstract, to be
-implemented by the package that tests a corresponding system
-implementation; this is a well known test pattern, the Abstract Test
-Pattern\cite{thomas2004java}, it is used to test that the contracts
-are respected in the implementations.  For instance, package
-\lil{DriverT} contains abstract tests for the abstractions of package
-\lil{Driver}, package \lil{StreamDriverT} inherits the abstractions
-contained in \lil{DriverT} and makes them concrete, to test the
-corresponding implementation \lil{StreamDriver}; package
-\lil{StreamDriverT} also contains stand alone tests written
-specifically for the implementation contained in \lil{StreamDriver}.
+the tests for an abstract package are abstract, to be implemented by
+the package that tests a corresponding system implementation; this is
+a well known test pattern (the Abstract Test
+Pattern\cite{thomas2004java}) used to test that the contracts are
+respected in the implementations.  For instance, package \lil{DriverT}
+contains abstract tests for the abstractions of package \lil{Driver},
+package \lil{StreamDriverT} inherits the abstractions contained in
+\lil{DriverT} and makes them concrete, to test the corresponding
+implementation \lil{StreamDriver}; package \lil{StreamDriverT} also
+contains stand alone tests written specifically for the implementation
+contained in \lil{StreamDriver}.
 
@@
 The JML specification has been used to generate the tests.  A specific
-tool of the JML2 suite can be used to automatically generate
+tool of the JML2 suite is able to automatically generate
 them\cite{Cheon-Leavens02}.
@@
 checker, hence, an oracle is not needed.  The generated tests send
-messages to the method under test, they catch assertion violation
-exceptions from test cases that pass an initial precondition check;
-that is, if a precondition is not met, the test is not run, if the
-precondition is met the test is run, and if the postcondition is not
-me an exception is fired and the test fails.  The test data that is
+messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation
+exceptions;
+if a precondition is not met the test is not run, if the
+precondition is met the test is run and if the postcondition is not
+met an exception is fired and the test fails.  The test data
 used to build the tests is the only part that is required to be
@@
 
-The test package structure is shown in Figure
-\ref{fig:class_diagram_generatedtest}.  The packages dependencies are
+The resulting test package structure is shown in Figure
+\ref{fig:class_diagram_generatedtest}.  The package dependencies are
 defined by the specifications; for instance, \lil{SimulatorDriverST}
 are generated tests, they test (hence use) the implementation
-\lil{SimulatorDriver}, they have been generated using the
-specification corresponding specification \lil{SimulatorDriverS}.  The
+\lil{SimulatorDriver}, and have been generated using the
+corresponding specification \lil{SimulatorDriverS}.  The
 generated tests do not maintain the inheritance structure of the code
@@
 
-To have proper unit test suite, for each method under test a proper
-stub, driver and oracle should be identified, these are the usual
-components found in the harness that enable test
+A proper stub, driver and oracle should be identified for each method
+under test to have proper unit test suite; stub, driver and oracle are
+the usual components found in the harness that enable test
 execution\cite{Binder1999}.  In
-JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, the framework
-used to manage the implementation and execution of hand made and
-generated unit tests, there is no clear distinction between drivers,
-stubs, and oracle of a test; none-less, even if a proper distinction
+JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the
+framework used to manage the implementation and execution of hand made
+and generated unit tests, there is no clear distinction between
+drivers, stubs, and oracles; none-less, even if a proper distinction
 is not imposed by the framework, the conceptual distinction still
@@
 
-Hence, to properly unit test a module in isolation and automatically
-we need:
+Hence, to properly unit test a package, we need:
 
@@
   passed or failed, comparing the output of the module under test to
-  the output that the oracle determines that the module should have.
+  the output the oracle determines the module should have.
 \end{description}
 
-Hand made unit tests and JML generated tests differs in how these
+Hand made unit tests and JML generated tests differ in how these
 components are provided.  The distinctions and similarities are
@@
 A module simulator can be used to ease the tasks of developing each
-component of a test harness; that is, it can be used as a stub, to
-simulate a module needed by the module under test, it can be used as a
-driver, to stimulate the module under test, it can be used as an
-oracle, if it simulates the module under test.
-
-In our system, the simulators has been used mainly as stubs, so that
-we have, during testing, a structure like the one shown in Figure
+component of a test harness: it can be used as a stub, to simulate a
+module needed by the module under test, it can be used as a driver, to
+stimulate the module under test, it can be used as an oracle, if it
+simulates the module under test.
+
+In our system, the simulators has been used as stubs, so that we have,
+during testing, a structure like the one shown in Figure
 \ref{fig:class_diagram_streamdriver_test}.  In Figure
@@
 \lil{StreamDriverT} is using the \lil{SimulatorStream} to properly
-simulate the \lil{Stream} that is used by the system under test, that
-is \lil{StreamDriver}.
+simulate the \lil{Stream} used by \lil{StreamDriver}, which is the
+system under test.
 
@@
 
-In our case, the usage of simulators to enable unit testing, was
+In our case, the use of simulators to enable unit testing, has been
 essential to test the \STSB driver, since the physical board was not
-available.  The simulator were used in both the hand made and the
-generated tests, since both of them need a proper set of stubs to be
-executed.
+available.  The simulators were have been used in both the hand made
+and the generated test suite, since both of them need a proper set of
+stubs to be executed.
 
@@
 specification, the sentinel values used to delimit a packet were not
-correct.  The board implementation error has been detected by the
-\lil{StreamDriver} package.
+correct.  The board implementation errors were detected by simply
+using the driver implementation, the \lil{StreamDriver} package.
 
@@
 were not yet working were the sensors devoted to the fast rate data
-streams, that is, the audio sensors.  Only one error needed to be
-corrected in the driver implementation, because of a (rare)
-combinations of conditions that were not covered by the test cases.
-One error have been discovered in the low level system driver that the
-developed driver was using (the system driver was outside of our
-development scope); to eliminate the error an upgrade of the low level
-driver to a new unreleased version was necessary.  No other errors
-were detected on the driver side. All the errors on the \SB side have
-been detected.
-
-The formal model and the run time assertion checker enabled the
-detection of errors on the \SB itself.  Errors regarding the expected
-values and values ranges of the fields contained in the packets and
-the frames (4 protocol errors), and errors regarding the dependencies
-between values in field (2 protocol errors) were detected. No other
-errors, other than the ones immediately discovered by the run time
-assertion checker were later on discovered.
+streams, the audio sensors.  Only one error needed to be corrected in
+the driver implementation, because of a (rare) combinations of
+conditions that were not covered by the test cases, but that
+occasionally showed up during real use.  One error have been
+discovered in the low level system driver that the developed driver
+was using (the system driver was outside of our development scope); to
+eliminate the error an upgrade of the low level driver to a new
+unreleased version was necessary.  No other errors were detected on
+the driver side.
+
+The second prototype respected the packet byte structure
+specification, but was not fully compliant to the packet info
+structure specification and the single packet rules.  The formal model
+and the run time assertion checker enabled the detection of these
+errors.  4 protocol errors regarding the packet info structure
+specification, and 2 protocol errors regarding the single packet rules
+has been found. No other errors, except the mentioned ones, have
+been later on discovered.
 
 The hand made test suite is composed by over 140 tests, while the
-generated one is 100 times bigger, totaling over 14000 tests.
-
-The ratio of 1 to 100, that is, the big number of generated tests, can
-be understood looking closely at how the generated tests are created
-by the JML framework. The generated tests are exploring all the
-possible input combinations on public methods, combining the type data
-ranges that have been specified for the test suite (see
-\cite{Cheon-Leavens02} for more details on how the JML framework works
-on generating unit tests). For instance, let's suppose we want to test
-a method \lil{m(int p1, int p2)} for a class \lil{C}; the data ranges
-specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C}
-are the instances \lil{\{c1,c2,c3,c4\}}. The JML framework will
-generate $24 = 3 \times 3 \times 4$ tests, the combination of the data
+generated one is 100 times bigger, totaling over 14000 tests.  The big
+number of generated tests (1 to 100 ratio), can be understood looking
+closely at how the generated tests are created by the JML
+framework. The generated tests are exploring the possible input
+combinations on public methods, combining the type data ranges that
+have been specified for the test suite (see \cite{Cheon-Leavens02} for
+more details on how the JML framework works on generating unit
+tests). For instance, let's suppose we want to test a method
+\lil{m(int p1, int p2)} for a class \lil{C}; the data ranges specified
+for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the
+instances \lil{\{c1,c2,c3,c4\}}. The JML framework generates $24 = 3
+\times 3 \times 4$ tests, since they are the combination of the data
 ranges involved (the data range of the parameters and the data range
-of the type owning the method under test). But not all the generated
-tests are executed.
+of the type owning the method under test). Fortunately, not all the
+generated tests are executed.
 
@@
 \ref{sec:test_cases}), is not executed, unless the preconditions of
-the called method are met. Thus, depending on the preconditions
+the called method are satisfied. Thus, depending on the preconditions
 specified for a method, the number of the generated tests that are
-actually executed in the test suite can be significantly reduced.
+actually executed in the test suite can be significantly smaller 
+compared to the total number of generated tests.
 
@@
 explore the control flow graph associated to the code, that is,
-decision branches are not explored. The statement code coverage
-results are reported in the Table \ref{tab:statement_code_coverage}.
+decision branches are not explored. The coverage results are reported
+in the Table \ref{tab:statement_code_coverage}.
 
@@
 
-Execution time are very different, the hand made test suite, with the
-run time assertion checking active runs, completely in less the one
+Test suites execution time are very different, the hand made test suite, with the
+run time assertion checking active runs completely in less than one
 minute on an average PC, while the generated test suite runs in more
@@
 \label{sec:test_cases_retrospectives}
-
-The quantitative and qualitative specification and test effectiveness
-measures to verify system under development obtained in the previous
-section need to be analyzed to be fully understood.
 
@@
 be called by the system, and they are only used in object
-initialization or during explicit object setup during unit tests. In
+initialization or in object setups performed during unit tests. In
 fact there are package and private code blocks that are not reachable
@@
 
-Regarding interfaces: an interface contains no statements, so when an
+Regarding interfaces, an interface contains no statements, so when an
 interface method is called, it is the method of the implementation
 class used that is actually covered. This effect is visible in
-\lil{Driver} coverage figures, \lil{Driver} package contains
+\lil{Driver} coverage figures: \lil{Driver} package contains
 interfaces (that are not counted at all) and some very simple real
-class (exceptions) that are not throughly tested: the net result is
+class (exceptions) that are not throughly tested. The net result is
 extremely low statement code coverage figures because the very simple
@@
 
-The statement code coverage figures shows that the test suites do a
+The statement code coverage figures show that the test suites do a
 good job on the most important package, that is, the
-\lil{StreamDriver}. The generated test suit reaching almost $ 80 \% $,
+\lil{StreamDriver}. The generated test suite reaches almost $ 80 \% $,
 while the hand made tests have only a slightly lower coverage ratio: $
@@
 the coverage ratio to $ 88.5 \% $. This is a clear indication that one
-test suite is not completely overlapping the other; that is, the test
-suites are not to be considered alternatives, but they have to be used
+test suite is not completely overlapping the other: the test suites
+are not to be considered alternatives, but they have to be used
 together to achieve the higher benefits, they are complementary. This
@@
 case is to be considered a rough specification written in imperative
-style. The common structure require an object initialization, followed
-by a hopefully not too long sequence of public methods to bring one
-object (or set of objects) into a desired state, than the method under
-test is called and the object state is finally inspected to verify it
-is in the expected state. Mock objects can be used to limit the
-dependencies. The test suite should be composed only by unit tests,
-but usually some tests are more complex to be called proper unit
-tests, they can be seen as simple integration tests.
-
-The generated test suite has a more focused and limited scope, since
+style. The common structure requires an object initialization,
+followed by a hopefully not too long sequence of public methods taking
+an object (or a set of objects) into a desired state; than the method
+under test is called and the object state is finally inspected to
+verify if it is in the expected state. Mock objects can be used to
+limit the dependencies. The test suite should be composed only by unit
+tests, but usually some tests are more complex to be called proper
+unit tests; they can be seen as simple integration tests.
+
+The generated test suite has a more focused and limited scope: since
 only one method call is performed in each test, hence it is difficult
@@
 longer and more complex specifications, requiring more time to be
-designed and or updated, and more tests to be run, augmenting the
+designed and or updated, and more tests to be run, increasing the
 already not so short computation time of the generated test suite.
@@
 The test suites gave their best when used in a combined approach: the
-generated test suite, checking most simple behaviors, while the hand
-made checking the more complex ones.  During the development of the
-driver code and the simulators, the generated test suit was helpful to
-test the incrementally growing specification and the finer details of
-each method, while the hand made one was helpful to verify the
-behavior on complex methods calls on objects initialized to specific
+generated test suite, checking simpler behaviors, and the hand made,
+checking the more complex ones.  During the development of the driver
+code and the simulators, the generated test suite was helpful to test
+the incrementally growing specifications and the finer details of each
+method, while the hand made one was helpful to verify the behavior on
+complex methods calls on objects initialized to specific
 states. Having an hand made test suite also make the code style more
@@
 systems, that is, the real system as well as the simulators.  The
-multiple simulators and multiple test suites approach is not new, it
-is often used in embedded systems software testing\cite{Broekman2002}.
-This is a general approach that can (and should) be used to verify
-both the simulators built and the real systems.  This approach can be
-easily used only if the code has been structured following the already
-mentioned dependency inversion principle\cite{martin1996dependency}
-(see Figure \ref{fig:class_diagram_main}).
+multiple simulators and multiple test suites approach is not new,
+since it has been often used in embedded systems software
+testing\cite{Broekman2002}.  This is a general approach that can (and
+should) be used to verify both the simulators built and the real
+systems.  It can be easily implemented only if the code has been
+structured following the already mentioned dependency inversion
+principle\cite{martin1996dependency} (see Figure
+\ref{fig:class_diagram_main}).
 
 The combined test approach, driven by the packet protocol
-specification, proved successful on verifying the protocol itself and
-the software implementing it.  Nevertheless, it has to be noted that we
-did not verified the highest levels of the protocol (see
-\ref{sec:data_stream_and_protocol_description}: ``Packet sequence
-rules'' and ``Packet input output asynchronous rules'').  Since on
-these levels the protocol properties regards sequences of packets,
+specification, has been proved successful on verifying the protocol
+itself and the software implementing it.  Nevertheless, it has to be
+noted that the highest levels of the protocol are more problematic
+(see \ref{sec:data_stream_and_protocol_description}: ``Packet sequence
+rules'' and ``Packet input output asynchronous rules'').  On these
+levels, the protocol properties regard sequences of packets, and
 testing them can be tricky, since the sequences of packets are not
-referred to explicitly in the implementation.  Two possible strategies
-can be used:
+referred to explicitly, in the implementation.  Two possible
+strategies can be used:
 
@@
   properties (when this is possible).  The advantage is to have
-  properties shown as the usual precondition, post conditions and
-  invariants.  But the localized properties can be more complex and
-  less intuitive and comprehensible than the original ones; finally,
-  some of the global properties could prove themselves impossible to
-  be translated into local ones.
+  properties shown as the usual pre-conditions, post-conditions and
+  invariants.  However, the localized properties can be more complex
+  and less intuitive and comprehensible than the original ones;
+  finally, some of the global properties could prove themselves
+  impossible to be translated into local ones.
 \item Use separate utility classes, meant to verify a sequence of
@@
   the protocol.  It has the advantage of keeping the properties that
-  have to be verified simple.
+  have to be verified mostly unchanged.
 \end{itemize}
@@
 One problem we noticed regards the execution time of the generated
-tests.  The execution time is long for two reasons:
+tests.  It is long for two reasons:
 
@@
   checks all the relevant properties. This problem, again, is not
-  directly addressable. But, we can speculate that if the RAC
-  assertions could be triggered on and off on single properties in a
-  dynamic way, that is, during runtime, a test suite could use take
-  advantage of this possibility, turning on the RAC only on the
-  properties that have to be verified for the classes or methods under
-  tests (and not for all the involved classes and methods); this could
-  save time during the test suite execution.
+  directly addressable. We can speculate that if the RAC assertions
+  could be triggered on and off on single properties in a dynamic way
+  (during runtime), a test suite would take advantage of this
+  possibility, turning on the RAC only on the properties that have to
+  be verified for the classes or methods under tests (and not for all
+  the involved classes and methods); this could save time during the
+  test suite execution.
 \end{itemize}
@@
 is well represented through state machines, which are then refined
-into JML properties.  The approach is in waterfall style, the need to
-develop the protocol, the specification and the implementation
+into JML properties.  The approach is in waterfall style; than the
+need to develop the protocol, the specification and the implementation
 concurrently is not addressed.  No hand made or generated test suites
@@
 used or assessed. It is described by its own creators in
-\cite{Cheon2002,Cheon2004,Cheon2005}, it has been used in
+\cite{Cheon2002,Cheon2004,Cheon2005}; it has been used in
 \cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test
 suite, and, more importantly, only against proof of concepts system,
-not real complex systems.  A brief assessment of the JML-JUnit tool
-has been made in \cite{Tan2004}, the analysis is based on injecting
-random generated errors in the source code, and then evaluating the
-ratio of the errors found by the test suite: the results shows that
-the JML-JUnit tool is not effective enough if used as a stand alone
-tool. No attempts on running or evaluating the generated test suite
-alongside with an hand made test suite are made.
+it has never be proved against real complex systems.  A brief
+assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the
+analysis is based on injecting random generated errors in the source
+code, and then evaluating the ratio of the errors found by the test
+suite: the results shows that the JML-JUnit tool is not effective
+enough if used as a stand alone tool. No attempts on running or
+evaluating the generated test suite alongside with an hand made test
+suite are made.
 
@@
   specification is partial
-\item use runtime assertion checker for formal specifications, on tests
+\item use runtime assertion checker for formal specifications, on
+  tests
 \item use generated tests for formal specifications

### 2009-11-02 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
Initial proposed template for thesis

\documentclass[]{final_report}
\usepackage{graphicx}
\usepackage{hyperref}

%%%%%%%%%%%%%%%%%%%%%%
%%% Input project details
\def\studentname{Ciaran Palmer}
\def\projecttitle{SensorML on SenseTile}
\def\supervisorname{Your Supervisor Name}
\def\moderatorname{Your Moderator Name}

\begin{document}

\maketitle
\tableofcontents\pdfbookmark[0]{Table of Contents}{toc}\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% Your Abstract here

\begin{abstract}

\end{abstract}
\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% Acknowledgments

\chapter*{Acknowledgments}

%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction

\chapter{Introduction}

Problem to solve
Whay is it interesting?
How is the problem going to be solved
Why is it innovate?

\chapter{ Background Research}

How was problem solved before.

\chapter{SensorML}
Basic description.
How the sy

\chapter{Design}
High level design

\chapter{ Detailed Design and Implementation}

implementation technology used
issues that arouse

\chapter{Results}

\chapter{ Conclusions and Future Work}

what has been achieved
the weaknesses of your approach

\chapter{References}

\chapter{Appendices}

\newpage
\begin{thebibliography}{99}
\bibitem{DAWSON:2000} Christian Dawson. \emph{The Essence of Computing Projects -- A Student's Guide}. 192 pages. ISBN: 013021972X. Pearson Education, 2000.
\end{thebibliography}
\label{endpage}

\end{document}

\end{article}

### 2009-11-09 papers/BON_VHDL/paper.tex
Reformatted as proper LaTeX and used fill-region to reformat syntax for easier reading and writing.  Paper has been revised and annotated by-hand by Joe.

@@
 
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
-sensor system installed at the University College Dublin in Dublin,
-Ireland.  \textbf{Our facility provides a capability---unique in the world in
-terms of its scale and flexibility---for performing in-depth
-investigation of both the specific and general issues in large-scale
-sensor networking.}
+The \textbf{UCD CASL SenseTile System} is a large-scale,
+general-purpose sensor system installed at the University College
+Dublin in Dublin, Ireland.  \textbf{Our facility provides a
+capability---unique in the world in terms of its scale and
+flexibility---for performing in-depth investigation of both the
+specific and general issues in large-scale sensor networking.}
 
@@
 The \STs, as an embedded system, have several unusual built-in sensors
-and emitters, sensor nodes can be extended via a general-purpose USB bus,
-and \STs contain an FPGA that is dynamically updated on a per-experiment
-basis.  \STs are custom-designed in VHDL-AMS.  Like many hardware-based systems,
- to write new software for or against the \ST architecture one either uses
-the actual hardware or one runs a \STS.  But this simulator emulates a
-piece of hardware, one would like to know if the simulator
-\emph{actually} simulates the hardware.  We have developed an approach
-through which we refine the VHDL-AMS specification up to the formal BON
-specification language, and from this specification we generate an
-executable, functional model-based specification in the JML language.
-We have used this specification to implement a fully validated and
-verified \ST simulator.
+and emitters, sensor nodes can be extended via a general-purpose USB
+bus, and \STs contain an FPGA that is dynamically updated on a
+per-experiment basis.  \STs are custom-designed in VHDL-AMS.  Like
+many hardware-based systems, to write new software for or against the
+\ST architecture one either uses the actual hardware or one runs a
+\STS.  But this simulator emulates a piece of hardware, one would like
+to know if the simulator \emph{actually} simulates the hardware.  We
+have developed an approach through which we refine the VHDL-AMS
+specification up to the formal BON specification language, and from
+this specification we generate an executable, functional model-based
+specification in the JML language.  We have used this specification to
+implement a fully validated and verified \ST simulator.
 
@@
 \label{sec:introduction}
-The \ST sensor platform, represents heterogeneous 
-collection of digital, analog and mixed signal hardware circuits. 
-The complexity of our platform requires the need to strive for formal 
-descriptions each component separately and  behavior of their 
-complex interdependencies, in order to produce executable specification.
-To make sure that the sensor platform works properly we need 
-to translate the VHDL-AMS code into the specification, and to refine and 
-verify this specification.
-In other words, our aim is to establish formal link between behavior of a given 
-sensor platform, and derived specification of components  in order to implement 
-fully validated and verified \ST simulator.
-For this purpose, we need notation and methodology 
-which will be able to satisfy the requirements given below.\\ 
-First, both of them, should be in sync with  VHDL-AMS language and reference 
-manual which tries to define VHDL-AMS as well as possible in a descriptive way,
-explaining the semantics in English.\\
-Second, specification derived from the notation should  be easy to attach to 
-the rest of the system architecture\xspace(driver, microprocessor emulator, 
-test-benches e.t.c.).\\
+
+The \ST sensor platform, represents heterogeneous collection of
+digital, analog and mixed signal hardware circuits.  The complexity of
+our platform requires the need to strive for formal descriptions each
+component separately and behavior of their complex interdependencies,
+in order to produce executable specification.\footnote{This is not a
+  good reason for performing this research.  Complexity is often the
+  reason to use \emph{less} formality, nor more.  Also, why must the
+  specification be \emph{executable}?}  To make sure that the sensor
+platform works properly we need to translate the VHDL-AMS code into
+the specification, and to refine and verify this
+specification.\footnote{Too vague a statement---what are the
+  \emph{real} reasons we are doing this work?}  In other words, our
+aim is to establish formal link between behavior of a given sensor
+platform, and derived specification of components in order to
+implement fully validated and verified \ST simulator.  For this
+purpose, we need notation and methodology which will be able to
+satisfy the requirements given below.
+
+First, both of them, should be in sync with VHDL-AMS language and
+reference manual which tries to define VHDL-AMS as well as possible in
+a descriptive way, explaining the semantics in English.
+
+Second, specification derived from the notation should be easy to
+attach\footnote{Meaning?} to the rest of the system
+architecture\xspace(driver, microprocessor emulator, test-benches
+e.t.c.).
+
 Third, methodology must aid efforts on creation of target system
-correctness criteria, such that are independent of implementation.
-It means that verification of simulator model should take place on specification 
-using reasoning techniques provided for algebraic formalism before coding starts.\\
-Fourth, a precise semantics should  provides the basis for
-automated forward-backward engineering.\\
+correctness criteria, such that are independent of implementation.  It
+means that verification of simulator model should take place on
+specification using reasoning techniques provided for algebraic
+formalism before coding starts.
+
+Fourth, a precise semantics should provides the basis for automated
+forward-backward engineering.
+
 Managing these requirements, and insuring that the platform definition
 (constraints) are preserved and verified through all levels of
-abstraction is where one becomes aware of gaps in already 
-used design methodologies.
-In our opinion, the BON-JML approach is in line with requirements given above.
-The authors will illustrate in this article how BON can represent a high
-level analogue mapping notation and will further illustrate a refinement 
-technique from VHDL-AMS to BON and JML.\\
+abstraction is where one becomes aware of gaps in already used design
+methodologies.  In our opinion, the BON-JML approach is in line with
+requirements given above.  The authors will illustrate in this article
+how BON can represent a high level analogue mapping notation and will
+further illustrate a refinement technique from VHDL-AMS to BON and
+JML.
+
 This article provides three main contributions.
 \begin{enumerate}
-\item
- It provides a solution how to formally emulate
- piece of hardware using BON-JML approach; up to date BON and JML
- have been used mostly in the pure object oriented software construction.   
-\item
-It relates hardware description code and refinement technique.
-It means that if a satisfactory VHDL-AMS code exists, then it can be simulated  from the 
-BON (informal/formal) specification, and JML implementation by an 
-appropriate sequence of refinements.
-\item
-It introduces \inv as a general concept of modeling hardware 
-components behavior and their interactions.
-To be more precisely, the behavioral specification of the \STS may 
-be represented as a \inv on the signals at the circuit's ports and 
-may be obtained by taking the conjunction of the \emph{invariants} 
-that describe behaviors of the individual components separately.\\ 
-\end{enumerate}
-The rest of this article is organized as follows.
-In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
-\autoref{sec:choiceobon} defines  BON principles and explains the
-reasons why we have decided to use BON method.
-In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON. 
+\item It provides a solution how to formally emulate piece of hardware
+using BON-JML approach; up to date BON and JML have been used mostly
+in the pure object oriented software construction.
+\item It relates hardware description code and refinement technique.
+It means that if a satisfactory VHDL-AMS code exists, then it can be
+simulated from the BON (informal/formal) specification, and JML
+implementation by an appropriate sequence of refinements.
+\item It introduces \inv as a general concept of modeling hardware
+components behavior and their interactions.  To be more precisely, the
+behavioral specification of the \STS may be represented as a \inv on
+the signals at the circuit's ports and may be obtained by taking the
+conjunction of the \emph{invariants} that describe behaviors of the
+individual components separately.
+
+\end{enumerate} The rest of this article is organized as follows.  In
+\autoref{sec:vhdlams} we have presented an overview of VHDL-AMS
+language.  \autoref{sec:choiceobon} defines BON principles and
+explains the reasons why we have decided to use BON method.  In
+\autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
 The central part of the our article is \autoref{sec:refVB}, in which
-we will give practical aspect of refinement, applied on 
-first order passive RC low-pass filter, in order to produce formal 
-specification and software simulation. In this section we 
-explore this approach in detail.
-In \autoref{sec:conclusion} the conclusions are given,
-and directions of future studies are outlined. Finally, a comprehensive list 
-of the mapping between BON and VHDL-AMS can be found in the appendix. \\ 
+we will give practical aspect of refinement, applied on first order
+passive RC low-pass filter, in order to produce formal specification
+and software simulation. In this section we explore this approach in
+detail.  In \autoref{sec:conclusion} the conclusions are given, and
+directions of future studies are outlined. Finally, a comprehensive
+list of the mapping between BON and VHDL-AMS can be found in the
+appendix.
+
 %=====================================================================
@@
 \section{VHDL-AMS}
-\label{sec:vhdlams}
-IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal
-high-level description language for electronic and multi-domain
-systems. VHDL-AMS, as an superset of VHDL, supports modeling and
-simulation of analog and mixed-signal systems. The main contribution
-of VHDL-AMS is ability to describe the behavior of complex system
-trought differential and algebraic equations (DAEs). For this purpose,
-VHDL-AMS introduces the following capabilities and therms:
+\label{sec:vhdlams} IEEE VHDL-AMS 1076.1 is the industry standard
+mixed-signal high-level description language for electronic and
+multi-domain systems. VHDL-AMS, as an superset of VHDL, supports
+modeling and simulation of analog and mixed-signal systems. The main
+contribution of VHDL-AMS is ability to describe the behavior of
+complex system trought differential and algebraic equations
+(DAEs). For this purpose, VHDL-AMS introduces the following
+capabilities and therms:
 
 \begin{itemize}
-\item
-\emph{Terminals} represents phisical connection points or circuit nodes
-of a network connectivity. They are declared to be of various
-\emph{natures}, which represent different energy domains of a system.
-Terminals hold no values; their only role is to facilitate the formation
-of conservative equation sets that, in turn, constrain the associated
-\emph{branch qualities}.
-\item
-The \emph{quantities} denote a waveform or a time series of values.
-The behavior of the \emph{quantities} is given by the \emph{simultaneous
- statements}. They take values as a result of solving the set of simultaneous
-ordinary differential and algebraic equations (DAEs). To simplify,
-quantities can be understood as analog unknowns of (DAEs) that specify
-analog behaviour. There are three kinds of quantity in VHDL-AMS: free,
-branch and source quantities. A \emph{free quantity} can be used in
-signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
-systems. A \emph{source quantity} is used for frequency and noise modeling.
-\item
-The \emph{simultaneous statements} creates one or more characteristic
-equations that are solved by the dedicated simulation kernel a.k.a.
-"analog solver", to model continuously valued behavior. 
-\item
-The \emph{procedural statements} or \emph{procedural} express analog behaviour
-using sequential statements. Actually procedural describes analog behaviour
-by computing new values for quantities from the current value of quantities.
-\item
-The \emph{break statement} is an construct which allows to
-initialize the DAE system at the beginning of a transient
-simulation and to mark discontinuities during such a transient
-simulation. For any concurrent break statement there is an associated process.
-A break statement allows restart of the analog kernel,
-than overrides quantity values and executes concurrently with the analogue model.
+\item \emph{Terminals} represents phisical connection points or
+circuit nodes of a network connectivity. They are declared to be of
+various \emph{natures}, which represent different energy domains of a
+system.  Terminals hold no values; their only role is to facilitate
+the formation of conservative equation sets that, in turn, constrain
+the associated \emph{branch qualities}.
+\item The \emph{quantities} denote a waveform or a time series of
+values.  The behavior of the \emph{quantities} is given by the
+\emph{simultaneous statements}. They take values as a result of
+solving the set of simultaneous ordinary differential and algebraic
+equations (DAEs). To simplify, quantities can be understood as analog
+unknowns of (DAEs) that specify analog behaviour. There are three
+kinds of quantity in VHDL-AMS: free, branch and source quantities. A
+\emph{free quantity} can be used in signal-flow modeling. A
+\emph{branch quantity} is used to model conservative energy systems. A
+\emph{source quantity} is used for frequency and noise modeling.
+\item The \emph{simultaneous statements} creates one or more
+characteristic equations that are solved by the dedicated simulation
+kernel a.k.a.  "analog solver", to model continuously valued behavior.
+\item The \emph{procedural statements} or \emph{procedural} express
+analog behaviour using sequential statements. Actually procedural
+describes analog behaviour by computing new values for quantities from
+the current value of quantities.
+\item The \emph{break statement} is an construct which allows to
+initialize the DAE system at the beginning of a transient simulation
+and to mark discontinuities during such a transient simulation. For
+any concurrent break statement there is an associated process.  A
+break statement allows restart of the analog kernel, than overrides
+quantity values and executes concurrently with the analogue model.
 
@@
 \section{Choice of BON as the modeling tool}
-\label{sec:choiceobon}
-%Brief description and motivation.
-The Business Object Notation (BON) is an analysis and design notation
-for object-oriented systems  which brings \emph{seamlessness},\xspace
-\emph{reversibility} and software \emph{contracting}.
-Let us define these principles precisely and point out why the principles
-are important in VHDL-AMS simulation.\\
-\textbf{Seamlessness and  Reversibility} \xspace \xspace
-Seamlessness is the use of a continuous process
-throughout the software life-cycle. Reversibility is the support for
-both forward and backward development process: from analysis to design
-and implementation, and back.
-In VHDL-AMS, modelling represents an continuous top-down-bottom-up
-process. Top-down approach serves for design while bottom-up
-serves for verification. The possibility of seamlessly translating
-changes made during a certain development phase back into earlier phases,
-facilitates a stable and predictable environment and ensures the various
-models consistent with the code. Keeping in sync various models, documentation
-and specification from forward progress to backward and redesign, preclude
-that design and implementation diverge to an inconsistent state.\\
-\emph{Seamlessness} also plays an important role in relating behaviors at
-different levels of abstraction. For instance, one may wish to analyze
-a "lowpass-filter" at the level of individual components (a resistor and 
-capacitor) and their behaviours, or at the more abstract level through
-general behaviour (\inv).
- Therefore, capturing similarities between the differing levels of abstractions,
-making it possible to obtain a continuous process that ensures a direct mapping
-between a formal specification and its software solution.\\
+\label{sec:choiceobon} %Brief description and motivation.  The
+Business Object Notation (BON) is an analysis and design notation for
+object-oriented systems which brings \emph{seamlessness},\xspace
+\emph{reversibility} and software \emph{contracting}.  Let us define
+these principles precisely and point out why the principles are
+important in VHDL-AMS simulation.
+
+\textbf{Seamlessness and Reversibility} \xspace \xspace Seamlessness
+is the use of a continuous process throughout the software
+life-cycle. Reversibility is the support for both forward and backward
+development process: from analysis to design and implementation, and
+back.  In VHDL-AMS, modelling represents an continuous
+top-down-bottom-up process. Top-down approach serves for design while
+bottom-up serves for verification. The possibility of seamlessly
+translating changes made during a certain development phase back into
+earlier phases, facilitates a stable and predictable environment and
+ensures the various models consistent with the code. Keeping in sync
+various models, documentation and specification from forward progress
+to backward and redesign, preclude that design and implementation
+diverge to an inconsistent state.
+
+\emph{Seamlessness} also plays an important role in relating behaviors
+at different levels of abstraction. For instance, one may wish to
+analyze a "lowpass-filter" at the level of individual components (a
+resistor and capacitor) and their behaviours, or at the more abstract
+level through general behaviour (\inv).  Therefore, capturing
+similarities between the differing levels of abstractions, making it
+possible to obtain a continuous process that ensures a direct mapping
+between a formal specification and its software solution.
+
 \textbf{Contracting} \xspace \xspace Contract specification describe
-system components interaction, based on mutual obligations and benefits, 
-and, for each software element, associated semantic properties. BON use a full
-assertion mechanism, allowing analysts to specify both:\xspace (1) the \emph{structure} of
-a system and (2)\xspace \emph{semantics} ( constraints and invariants).
-The common way to specify a program statement is to give a \emph{precondition} and
-\emph{postcondition} on the state that the implementation must satisfy, assuming that
-only certain program variables are changed.
-In VHDL-AMS any predefined (implicate) attribute 
-has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
-parameters and theirs constraints. Furthermore, any subtype (subnature)
-represents restricted set of values from a base type (nature).
-The condition that determines which values are in the subtype (subnature) 
-may be contracted by specifying characteristic predicate which represents \inv.\\
-Let us next look how \emph{simultaneous statements} are handled in contract design.
-The \emph{simultaneous statement} represents analogue constraint of
-continuous-time waveform. We will assume that the constraint of waveform
-represents \inv. This premise deserve an informal explanation.\\
-Briefly stated, analog kernel calculate quantities such
-that the constraints specified by the simultaneous statements are satisfied.
+system components interaction, based on mutual obligations and
+benefits, and, for each software element, associated semantic
+properties. BON use a full assertion mechanism, allowing analysts to
+specify both:\xspace (1) the \emph{structure} of a system and
+(2)\xspace \emph{semantics} ( constraints and invariants).  The common
+way to specify a program statement is to give a \emph{precondition}
+and \emph{postcondition} on the state that the implementation must
+satisfy, assuming that only certain program variables are changed.  In
+VHDL-AMS any predefined (implicate) attribute has defined \emph{pre}
+(\emph{post}) \emph{condition} trought input (output) parameters and
+theirs constraints. Furthermore, any subtype (subnature) represents
+restricted set of values from a base type (nature).  The condition
+that determines which values are in the subtype (subnature) may be
+contracted by specifying characteristic predicate which represents
+\inv.
+
+Let us next look how \emph{simultaneous statements} are handled in
+contract design.  The \emph{simultaneous statement} represents
+analogue constraint of continuous-time waveform. We will assume that
+the constraint of waveform represents \inv. This premise deserve an
+informal explanation.
+
+Briefly stated, analog kernel calculate quantities such that the
+constraints specified by the simultaneous statements are satisfied.
 More accurately, \emph{simultaneous statements} can be thought of as
-algebraic equations. It means that the number of simultaneous statements
-should be equal to the number of unknowns (\emph{quantities}). 
-For instance, when we reasoning about the conservation of energy law,
-the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
-in a model.
-In other words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0})
-around any closed path must be equal to zero and the sum of \through
-quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
-(terminal) must be equal to zero. To be more precise, the results of summations 
-above remains zero when a particular transformation is applied to quantities.
-In general, they are \inv. \\
-However, if our implementation does not satisfy the conservation of energy law, 
- then the invariant will be broken. 
-Another conclusion may be deducted from the example given above.\\
-Intuitively \inv may be useful for testing the function of the hardware components.
-In practice, one can runs simulator against the component in order to check 
-the properties. Testing can then be done by checking whether the \inv hold or not.
-For example, the most common problem of the real components is the glich - an electrical pulse 
-of short duration that is usually the result of a design error.
-The number of glitches that occur can be known by the fact how many times per day, 
-the \inv was broken and reestablished. According to that we can conclude whether the 
-quality of a component is satisfactory. Unlike in the case of software, where 
-correction of an error is relatively inexpensive, an detected hardware error 
-may require product replacement, which is associated with great expenses. 
-In other words, we cannot patch a fabricated component once we started massive production. \\
-Therefore, on the basis of the above, BON is the obvious choice
-for the \STS.
-Throughput this article we will make this more clearly. 
+algebraic equations. It means that the number of simultaneous
+statements should be equal to the number of unknowns
+(\emph{quantities}).  For instance, when we reasoning about the
+conservation of energy law, the \emph{simultaneous statements}
+constrain the values of \emph{branch quantities} in a model.  In other
+words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across
+quatities =0}) around any closed path must be equal to zero and the
+sum of \through quantities (\texttt{$\sum_{node}$ through quatities
+=0}) at any node (terminal) must be equal to zero. To be more precise,
+the results of summations above remains zero when a particular
+transformation is applied to quantities.  In general, they are \inv.
+
+However, if our implementation does not satisfy the conservation of
+energy law, then the invariant will be broken.  Another conclusion may
+be deducted from the example given above.
+
+Intuitively \inv may be useful for testing the function of the
+hardware components.  In practice, one can runs simulator against the
+component in order to check the properties. Testing can then be done
+by checking whether the \inv hold or not.  For example, the most
+common problem of the real components is the glich - an electrical
+pulse of short duration that is usually the result of a design error.
+The number of glitches that occur can be known by the fact how many
+times per day, the \inv was broken and reestablished. According to
+that we can conclude whether the quality of a component is
+satisfactory. Unlike in the case of software, where correction of an
+error is relatively inexpensive, an detected hardware error may
+require product replacement, which is associated with great expenses.
+In other words, we cannot patch a fabricated component once we started
+massive production.
+
+Therefore, on the basis of the above, BON is the obvious choice for
+the \STS.  Throughput this article we will make this more clearly.
 \section{Mapping between VHDL-AMS and BON}
-\label{sec:mappingVB}
-In this section, we briefly describe the mapping between VHDL-AMS and BON notation 
-which is discussed in the points below.\\
-\textbf{Structure} \xspace \xspace VHDL-AMS allows hierarchical structure 
-modeling, meaning that the model subsystems can be developed independently.
-As an top-level description, \lstinlinen|SYSTEM_CHART| in the BON notation  
-can connect all the model subsystems  together to a model design.
-In BON, model subsystem is denoted as the \lstinlinen|CLUSTER_CHART| whereas 
-module is represented as the \lstinlinen|CLASS_CHART|.
-Each module has a \texttt{port} interface: a set of \emph{inputs} and \emph{outputs} used for 
-connecting modules with environment. The \texttt{port} section in 
-\texttt{\ent}structure (input, output and generic values) are represented as 
-a \emph{features} in the \lstinlinen|CLASS_CHART|.
-The \lstinlinen|CLASS_CHART| \xspace can be equated with \ent-\arch 
-structure in VHDL-AMS and the name of the \lstinlinen|CLASS_CHART|\xspace 
-corresponds to the name of \texttt{\ent}in VHDL-AMS.\\ 
+\label{sec:mappingVB} In this section, we briefly describe the mapping
+between VHDL-AMS and BON notation which is discussed in the points
+below.
+
+\textbf{Structure} \xspace \xspace VHDL-AMS allows hierarchical
+structure modeling, meaning that the model subsystems can be developed
+independently.  As an top-level description, \lstinlinen|SYSTEM_CHART|
+in the BON notation can connect all the model subsystems together to a
+model design.  In BON, model subsystem is denoted as the
+\lstinlinen|CLUSTER_CHART| whereas module is represented as the
+\lstinlinen|CLASS_CHART|.  Each module has a \texttt{port} interface:
+a set of \emph{inputs} and \emph{outputs} used for connecting modules
+with environment. The \texttt{port} section in \texttt{\ent}structure
+(input, output and generic values) are represented as a
+\emph{features} in the \lstinlinen|CLASS_CHART|.  The
+\lstinlinen|CLASS_CHART| \xspace can be equated with \ent-\arch
+structure in VHDL-AMS and the name of the
+\lstinlinen|CLASS_CHART|\xspace corresponds to the name of
+\texttt{\ent}in VHDL-AMS.
+
 \textbf{Behavior} \xspace \xspace The module also contains an
 \texttt{\arch} implementation, which determines the way the module
-processes \emph{input}/\emph{output} signals and performs computations.
-The \emph{constraint} section in the \lstinlinen|CLASS_CHART| 
-contains the transfer function and holds the same information as 
-the \texttt{\arch}section of the VHDL-AMS.\\ 
-\textbf{Types and operators} \xspace \xspace There are many features of 
-VHDL-AMS which are already mapped in BON by default, such as arithmetic, 
-relational and logical operators and types like \xspace \texttt{integer},
-\xspace \texttt{byte},\xspace\texttt{real},\xspace\texttt{character} and 
-\xspace\texttt{boolean}.
-Vector may be represented as \lstinlinen|SEQUENCE[TYPE]|, 
-whereas record, whose elements are of heterogeneous types, can be mapped 
-as \lstinlinen|SET| or \lstinlinen|TABLE|.
-In BON, the \texttt{time} can be represented as the
-\lstinlinen|UNIVERSAL_TIME|, in order to specify \emph{delays}. Actually 
-time can be thought of as object with integer as a value and measurement 
-units as a \xspace \lstinlinen|SEQUENCE| feature.     
-In order to specify energy domain for nodes of circuits we have developed 
-various standard \texttt{nature} templates for \texttt{terminals}.
-We have introduced \xspace\lstinlinen|FREE|,\xspace \lstinlinen|BRANCH|
-and  \xspace\lstinlinen|SOURCE| \space \lstinlinen|QUANTITY| objects, 
-which are intended to describe signal-flow modeling, the 
-conservation of energy low and frequency-noise related problems. 
-There are also defined specifications for effort and 
-flow domains, tolerance as well as appropriate contracts for 
-scalar types(natures), array types(natures), 
-terminals and quantity attributes.
-Whereas the value assignment, in VHDL-AMS has the same purpose as in the BON,
-the signal assignment, which schedules a new value to be applied to a 
-signal later, is simulated in specification, but  denotational semantic
-is not yet translated in the BON as well as for concurrency and 
-the \texttt{wait} statement.
+processes \emph{input}/\emph{output} signals and performs
+computations.  The \emph{constraint} section in the
+\lstinlinen|CLASS_CHART| contains the transfer function and holds the
+same information as the \texttt{\arch}section of the VHDL-AMS.
+
+\textbf{Types and operators} \xspace \xspace There are many features
+of VHDL-AMS which are already mapped in BON by default, such as
+arithmetic, relational and logical operators and types like \xspace
+\texttt{integer}, \xspace
+\texttt{byte},\xspace\texttt{real},\xspace\texttt{character} and
+\xspace\texttt{boolean}.  Vector may be represented as
+\lstinlinen|SEQUENCE[TYPE]|, whereas record, whose elements are of
+heterogeneous types, can be mapped as \lstinlinen|SET| or
+\lstinlinen|TABLE|.  In BON, the \texttt{time} can be represented as
+the \lstinlinen|UNIVERSAL_TIME|, in order to specify
+\emph{delays}. Actually time can be thought of as object with integer
+as a value and measurement units as a \xspace \lstinlinen|SEQUENCE|
+feature.  In order to specify energy domain for nodes of circuits we
+have developed various standard \texttt{nature} templates for
+\texttt{terminals}.  We have introduced
+\xspace\lstinlinen|FREE|,\xspace \lstinlinen|BRANCH| and
+\xspace\lstinlinen|SOURCE| \space \lstinlinen|QUANTITY| objects, which
+are intended to describe signal-flow modeling, the conservation of
+energy low and frequency-noise related problems.  There are also
+defined specifications for effort and flow domains, tolerance as well
+as appropriate contracts for scalar types(natures), array
+types(natures), terminals and quantity attributes.  Whereas the value
+assignment, in VHDL-AMS has the same purpose as in the BON, the signal
+assignment, which schedules a new value to be applied to a signal
+later, is simulated in specification, but denotational semantic is not
+yet translated in the BON as well as for concurrency and the
+\texttt{wait} statement.
 
 \section{Refinement from VHDL-AMS to BON }
-\label{sec:refVB}
-The main theme of this article is the translation from VHDL-AMS to BON using 
-refinement approach.
-We will refine the VHDL-AMS code up to the formal BON, and from this 
-specification we will generate an executable specification in the JML language.
-For this purpose, we will demonstrate our approach in BON-JML fashion trought  
-low-pass filter example.\\
+\label{sec:refVB} The main theme of this article is the translation
+from VHDL-AMS to BON using refinement approach.  We will refine the
+VHDL-AMS code up to the formal BON, and from this specification we
+will generate an executable specification in the JML language.  For
+this purpose, we will demonstrate our approach in BON-JML fashion
+trought low-pass filter example.
+
 \subsection{Low-Pass filter transfer function }
@@
 \input{lowpassRC}
-The most simple low-pass filter is passive RC network 
-shown in \autoref{fig:lowpassfig}.
-According to equation given above (see \autoref{fig:transfun}) 
-we can produce VHDL-AMS \emph{simultaneous statement} using 
-\texttt{$s = d/dt = 'dot$} relation.
-The transformation looks as follows: \\
+
+The most simple low-pass filter is passive RC network shown in
+\autoref{fig:lowpassfig}.  According to equation given above (see
+\autoref{fig:transfun}) we can produce VHDL-AMS \emph{simultaneous
+  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
+transformation looks as follows:
 \begin{equation}
@@
 \end{equation}
- The last equation can be described as follows. 
-The input voltage\xspace($v_{in}$)\xspace equals summation 
-of output voltage\xspace($v_{out}$)\xspace and multiplication 
-of the time constant\xspace($\tau_p$)\xspace of the pole and the
-first derivative of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace 
-over time.\\
-Now we can discretize previous equation. Let us represent $v_{in}$ by a series 
-of samples $v_{in(1..n)}$. At the same points in time, we can represent $v_{out}$ by 
-series of sample $v_{out(1..n)}$. We can assume that samples are taken at points in 
-time incrementing by $\triangle$t. Now we have equivalent equations:
+
+The last equation can be described as follows.  The input
+voltage\xspace($v_{in}$)\xspace equals summation of output
+voltage\xspace($v_{out}$)\xspace and multiplication of the time
+constant\xspace($\tau_p$)\xspace of the pole and the first derivative
+of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace over time.
+
+Now we can discretize previous equation. Let us represent $v_{in}$ by
+a series of samples $v_{in(1..n)}$. At the same points in time, we can
+represent $v_{out}$ by series of sample $v_{out(1..n)}$. We can assume
+that samples are taken at points in time incrementing by
+$\triangle$t.
+
+Now we have equivalent equations:
 \begin{equation}
-v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                                                  
-\end{equation}
+v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                          \end{equation}
 
@@
  \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                                                           
- \end{array}
+      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                              \end{array}
  \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                                                                  
@@
 
-The last equation gives us a way to calculate the output samples in terms of the 
-input samples and the preceding output.
-But how to represent these unknowns in BON ?
-We introduce the  \lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
-with features as fallows:
+The last equation gives us a way to calculate the output samples in
+terms of the input samples and the preceding output.  But how to
+represent these unknowns in BON ?  We introduce the
+\lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
+with features as follows:
 \begin{itemize}
@@
 \end{itemize}
-Now we are in presence of BON notation and we may start
-with the  refinement of the VHDL-AMS code.\\
+Now we are in presence of BON notation and we may start with the
+refinement of the VHDL-AMS code.
+
 \subsection{Refine example }
 \label{subsec:refine}     
-We have not yet said what specification and refinement are.
-Let us introduce these terms first.\\
-A \emph{specification} of a program component is a 
-contract that gives some constraints 
-on how the component is to behave but leaves freedom for the implementator to 
-decide how the actual behavior of the component is to be realized. 
-As we have shown in \autoref{tab:rescapfirst}, corresponding 
-informal specifications represents high level of abstraction derived 
-from VHDL-AMS code.\\
-A \emph{refinement} represents the concept that can be understood as series of 
-correctness-preserving transformations from general perspective to 
-a precise level of detail. Accurately, if we compare two specifications,
-\texttt{S} and \texttt{S'}, then we can conclude that the \texttt{S} 
-is more abstract than \texttt{S'} if any condition that we can establish 
-with the first specification can also be established by second. 
-It means that "\texttt{S is \emph{refined} by S'}", 
-and we denote this  as \texttt{S $\sqsubseteq$ S'}.
-The \emph{refinement} which was derived in three steps, 
-from informal to more formal BON and finally to JML, 
-is represented using relation \texttt{$\sqsubseteq$}. This relation represent 
-\emph{isomorphism} and has inverse relation \texttt{$\sqsupseteq$}. 
-It means that once we have more formal specification, we can refine 
-and seamlessly  translate specification in previous abstract level,  
-keeping specification validity and documentation consistency. 
-In other  words, we use two important BON principles to achieve 
-as much as precise model consistency: the seamlessness and  
-reversibility(see \autoref{sec:choiceobon}). 
-Let say that the first informal BON specification is denoted by 
-\texttt{S} (see \autoref{tab:rescapfirst}), and second and third,
-which are given in \autoref{tab:rescapsecond}  is denoted by \texttt{S',S''}.
-The forward process refinement can be represented as:\xspace
-\texttt{S $\sqsubseteq$ S' $\sqsubseteq$ S''}, while the backward process
-can be represented as \xspace \texttt{S''$\sqsubseteq$ S' $\sqsubseteq$ S}.  
-Using all BON principles, we have developed an specification of RC low-pass
-filter, and we have produced fully validated and verified low-pass filter.
-\input{lowpass1}
-\input{lowpass2}
+We have not yet said what specification and refinement are.  Let us
+introduce these terms first.
+
+A \emph{specification} of a program component is a contract that gives
+some constraints on how the component is to behave but leaves freedom
+for the implementator to decide how the actual behavior of the
+component is to be realized.  As we have shown in
+\autoref{tab:rescapfirst}, corresponding informal specifications
+represents high level of abstraction derived from VHDL-AMS code.
+
+A \emph{refinement} represents the concept that can be understood as
+series of correctness-preserving transformations from general
+perspective to a precise level of detail. Accurately, if we compare
+two specifications, \texttt{S} and \texttt{S'}, then we can conclude
+that the \texttt{S} is more abstract than \texttt{S'} if any condition
+that we can establish with the first specification can also be
+established by second.  It means that "\texttt{S is \emph{refined} by
+S'}", and we denote this as \texttt{S $\sqsubseteq$ S'}.  The
+\emph{refinement} which was derived in three steps, from informal to
+more formal BON and finally to JML, is represented using relation
+\texttt{$\sqsubseteq$}. This relation represent \emph{isomorphism} and
+has inverse relation \texttt{$\sqsupseteq$}.  It means that once we
+have more formal specification, we can refine and seamlessly translate
+specification in previous abstract level, keeping specification
+validity and documentation consistency.  In other words, we use two
+important BON principles to achieve as much as precise model
+consistency: the seamlessness and reversibility(see
+\autoref{sec:choiceobon}).  Let say that the first informal BON
+specification is denoted by \texttt{S} (see
+\autoref{tab:rescapfirst}), and second and third, which are given in
+\autoref{tab:rescapsecond} is denoted by \texttt{S',S''}.  The forward
+process refinement can be represented as:\xspace \texttt{S
+$\sqsubseteq$ S' $\sqsubseteq$ S''}, while the backward process can be
+represented as \xspace \texttt{S''$\sqsubseteq$ S' $\sqsubseteq$ S}.
+Using all BON principles, we have developed an specification of RC
+low-pass filter, and we have produced fully validated and verified
+low-pass filter.  \input{lowpass1} \input{lowpass2}
 
@@
 \section{Conclusion}
-\label{sec:conclusion}
-The BON notation has originally been used for description of 
-software models. However, after certain elaboration it can be used 
-for description of hardware components. Practical utilization of the approach
-in the project on mapping between VHDL-AMS and BON using refinement technique,
-has demonstrated through low-pass filter example and shown that simulation
-of component requires relatively little and inexpensive  efforts. Moreover, our article shown
-that the development of the specification produced by the BON and translated 
-into JML can be applied as refinement procedure: from VHDL-AMS to informal BON, 
-next, from informal to more formal BON and finally, to JML specification. 
-The BON support general software development and should be adapted for concrete
-applications and language as VHDL-AMS. For this purpose we have developed various 
-VHDL-AMS related types, operators and relations and we have defined appropriate 
-specifications for all of them. However, the behavior of
-\texttt{break} statement and  concurrency, for example, have not been translated 
-yet in the BON notation. We have started a new phase of project named \emph{eBon} in which
-the issues mentioned above, and creation of appropriate components and relations 
-are main subject.
+\label{sec:conclusion} The BON notation has originally been used for
+description of software models. However, after certain elaboration it
+can be used for description of hardware components. Practical
+utilization of the approach in the project on mapping between VHDL-AMS
+and BON using refinement technique, has demonstrated through low-pass
+filter example and shown that simulation of component requires
+relatively little and inexpensive efforts. Moreover, our article shown
+that the development of the specification produced by the BON and
+translated into JML can be applied as refinement procedure: from
+VHDL-AMS to informal BON, next, from informal to more formal BON and
+finally, to JML specification.  The BON support general software
+development and should be adapted for concrete applications and
+language as VHDL-AMS. For this purpose we have developed various
+VHDL-AMS related types, operators and relations and we have defined
+appropriate specifications for all of them. However, the behavior of
+\texttt{break} statement and concurrency, for example, have not been
+translated yet in the BON notation. We have started a new phase of
+project named \emph{eBon} in which the issues mentioned above, and
+creation of appropriate components and relations are main subject.
 
-\section {Acknowledgments}
-The research upon which this article is based was caried out with
-sponsorship from Science Foundation Ireland via the
-\textbf{UCD CASL SenseTile System} grant.
+\section {Acknowledgments} The research upon which this article is
+based was caried out with sponsorship from Science Foundation Ireland
+via the \textbf{UCD CASL SenseTile System} grant.
 

### 2009-11-10 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
Some initial updates

@@
 \begin{abstract}
+SenseTileSensor Board
+
+SensorML description.
+
+WebServer to access sensor observations
+
+SensorML Bon Mapping with a tool and method to develop sensorML descriptions
 
 \end{abstract}
+
 \newpage
@@
 
-Problem to solve
-Whay is it interesting?
-How is the problem going to be solved
-Why is it innovate?
+The aim of this thesis project was to prototype a Web Service for accessing sensor observations from the SenseTile Sensor Board.
+SenseTile is a sensor and processing package including motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
+It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
+
+Descriptions of the SenseTile Sensor Board were to be developed using Sensor Model Language (SensorML)\cite{SensoMLref}.
+SensorML is a language that describes the process of sensor measurement by providing standard models and XML encoding for such processes.
+SensorML was be used as it is becoming a standard for describing sensor data processing as part of the SWE.
+
+Another part of the thesis was to define a SensorML to BON Mapping .
+BON\cite{BONref} is a notation and a method for Object Oriented systems analysis and design.
+Based on this mapping it was proposed to develop a tool/method to automatically generate SensorML descriptions.
+Generally SensorML is hand generated with all the shortcomings of such an approach.
+The proposed approach would allow sensor processing to defined formally in BON and then refined to JML/Java.
 
@@
 \chapter{SensorML}
-Basic description.
-How the sy
+\section{SensorML Overview}
+The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) activity, is being executed through the OGC Web Services (OWS) initiatives (under the Interoperability Program).
+The SWE initiatives are establishing the interfaces and protocols that will enable a “Sensor Web” through which applications and services will be able to access sensors of all types over the Web.
+All processes define their inputs, outputs, parameters, and method, as well as provide relevant metadata. SensorML models detectors and sensors as processes that convert real phenomena to data.
+
+sensors modeled as processes that can be connected and participate equally within a process chain or system, and which utilize the same process model frame as any other process.
+Processes are entities that take one or more inputs and through the application of well-defined methods using specific parameters, results in one or more outputs.
+SensorML also supports linking between processes and thus supports the concept of process chains, which are themselves defined as processes.
+The second possible role is to provide processing chains with which SensorML-enabled software could derive new data from existing observations on-demand.
+
+A standard description format for sensors is important for the long-term definition of the sensor model’s fundamental characteristics and assumptions for use in future reprocessing and refining of sensor data.
+
+a System defines a collection of sensors that are related spatially and temporally to one another, as well as related to some geospatial domain.
+Detectors are typically simple devices (real or virtual) that take one input signal and generate one or more output signals.
+
+In SensorML, these simpler atomic processes are defined as “process models”, if they are pure processes with no real connection to a spatial or temporal domain, and as “components” if they are physical processing devices, such as detectors and actuators.
+
+SensorML is mostly used to describe functional models of the sensors rather than the hardware.
+
+The response characteristics of a sensor determine how the sensor will react to a particular stimulus (i.e., phenomenon) and how it will operate under given environmental conditions.
+Within the sensor response characteristics will be specifications for sensitivity (e.g., threshold, dynamic range, capacity, band width, etc.), accuracy and precision, and behavior under certain environmental conditions.
+
+    Component -
+
+    Physical atomic process that transforms information from one form to another.
+For example, a Detector typically transforms a physical observable property or phenomenon to a digital number.
+Example Components include detectors, actuators, and physical filters.
+
+    System -
+
+    Composite physically-based model of a group or array of components, which can include detectors, actuators, or sub-systems.
+A System relates a process to the real world and therefore provides additional definitions regarding relative positions of its components and communication interfaces.
+
+    Process Model -
+
+    Atomic non-physical processing block usually used within a more complex Process Chain.
+It is associated to a Process Method which defines the process interface as well as how to execute the model.
+It also precisely defines its own inputs, outputs and parameters.
+
+    Process Chain -
+
+    Composite non-physical processing block consisting of interconnected sub-processes, which can in turn be Process Models or Process Chains.
+A process chain also includes possible data sources as well as connections that explicitly link input and output signals of sub-processes together.
+It also precisely defines its own inputs, outputs and parameters.
+
+    Process Method -
+
+    Definition of the behavior and interface of a Process Model.
+It can be stored in a library so that it can be reused by different Process Model instances (by using 'xlink' mechanism).
+It essentially describes the process interface and algorithm, and can point the user to existing implementations.
+
+    Detector -
+
+    Atomic component of a composite Measurement System defining sampling and response characteristic of a simple detection device.
+A detector has only one input and one output, both being scalar quantities.
+More complex Sensors such as a frame camera which are composed of multiple detectors can be described as a detector group or array using a System or Sensor.
+In SensorML a detector is a particular type of Process Model.
+
+    Sensor -
+
+    Specific type of System representing a complete Sensor. This could be for example a complete airborne scanner which includes several Detectors (one for each band).
+
+The SensorML description of the SenseTile Sensor Board is shown in the following section.
+
+\section{SensorML Description of SenseTile}
+
+The SenseTile Sensor Board is modelled a SensorML System. Meta data such as position.
+The Sensors are modelled a detectors. They are process that take scalar values
+and can perform transformations on them.
+
+System is providing the mapping of inputs to processes and the outputs.
+The 
+ \begin{figure}
+\includegraphics{SensorML_SenseTile_block_design.png}
+\caption{Block diagram of SenseTile System process}\label{fig:SensorML_SenseTile_design.png}
+\end{figure}
+
+Types
+ - SWE -
+
+Count - BigInteger
+Quantity - Double
+
+ * Carries an array of int primitives.
+ * All data is casted to other types when requested.
+
+public class DataBlockInt extends AbstractDataBlock
+
+mapping packet data to process:packet->sensor process
+could be multiple inputs.
+
+\section{BON to SensorML mapping}
 
@@
 High level design
+\section{SenseTile Web Service}
+
+DataProducer - use PacketInput stream to access SenseTile observation packets.
+
+DataProvider - provides an RMI interface to the DataProducer to update. Provides a Web Service interface  based on Sensor Observation Service. Not fully as is a very complex interface.
+
+\section{SensorML generation tool}
+
+SensorML BON mapping
+
+BON to Java
+
+Instantiate Java classes with data
+
+Walk Java class and generate sensorml.
 
@@
 
-implementation technology used
-issues that arouse
+\section{SenseTile WebService}
+
+DataProvider is based on AXIS 2.0/Spring
+
+DataProducer POJO/VastLib/UCD SenseTile Driver Lib
 
@@
 \begin{thebibliography}{99}
-\bibitem{DAWSON:2000} Christian Dawson. \emph{The Essence of Computing Projects -- A Student's Guide}. 192 pages. ISBN: 013021972X. Pearson Education, 2000.
+\bibitem{SensoMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
+\bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
 \end{thebibliography}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/AbstractProcess.java
Adding dataproducer classes

@@ -1,0 +1,73 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.sensorboard.Packet;
+
+import org.apache.log4j.BasicConfigurator;
+import org.apache.log4j.Logger;
+import org.vast.data.DataValue;
+import org.vast.process.DataProcess;
+import org.vast.process.ProcessException;
+
+enum ProcessType {SENSOR,CONVERTER}
+public abstract class AbstractProcess extends DataProcess {
+
+    private DataValue input;
+ 
+    private DataValue output;
+
+    protected static Logger logger =
+        Logger.getLogger(AbstractProcess.class);
+
+    private ProcessType processType = ProcessType.SENSOR;
+
+    public AbstractProcess() {
+        BasicConfigurator.configure();
+    }
+
+    public ProcessType getProcessType() {
+        return this.processType;
+    }
+
+    public void setProcessType(ProcessType processType) {
+        this.processType = processType;
+    }
+    /**
+     * * Initializes the process.
+     * * Gets handles to input/output components.
+     */
+    public void init() throws ProcessException {
+        try {
+            System.out.println("abstact name ???? "+this.getName());
+            // I/O mappings
+            input =
+                (DataValue) inputData.getComponent(this.getName() + "Input");
+            output =
+                (DataValue) outputData.getComponent(this.getName() + "Output");
+        } catch (ClassCastException e) {
+            throw new ProcessException("Invalid I/O data", e);
+        }
+    }
+
+    protected final double getDoubleInput() {
+        return input.getData().getDoubleValue();
+    }
+
+    protected final int getIntInput() {
+        return input.getData().getIntValue();
+    }
+
+    protected final void setDoubleOutput(final double value) {
+        output.getData().setDoubleValue(value);
+    }
+
+    protected final void setIntOutput(final int value) {
+        output.getData().setIntValue(value);
+    }
+    
+    /**
+    * Executes process algorithm on inputs and set output data.
+    */
+    public abstract void execute() throws ProcessException;
+
+
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/CelciusConverter.java

@@ -1,0 +1,16 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import org.vast.process.ProcessException;
+
+public class CelciusConverter extends AbstractProcess
+{
+    public static double CONVERSIONVALUE = 16.0;
+    
+    public void execute() throws ProcessException {
+        super.logger.info("Thermistor execute");
+
+        double temp = this.getDoubleInput();
+        this.setDoubleOutput(temp /CONVERSIONVALUE);
+    }
+
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/PacketStreamReader.java

@@ -1,0 +1,94 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import java.util.ArrayList;
+import java.util.List;
+
+import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import org.apache.log4j.Logger;
+import org.apache.log4j.BasicConfigurator;
+import org.apache.log4j.NDC;
+/**
+ * Reads parameterised no of packets every second.
+ *
+ * @author Ciaran Palmer
+ * @version 1
+ */
+
+
+public class PacketStreamReader {
+
+       private static final int MAXNOPACKETS = 1000; //arbitrary
+       private static final int MINNOPACKETS = 1; //arbitrary
+       private int numpackets=MINNOPACKETS;
+
+       private PacketInputStream packetInputStream;
+       private boolean readpackets;
+       private List<Sensor> sensorList;
+
+       static Logger logger =
+           Logger.getLogger(PacketStreamReader.class.getName());
+       private SenseTileSystem senseTileSystem;
+
+       public PacketStreamReader(SenseTileSystem senseTileSystem) {
+           BasicConfigurator.configure();
+           this.senseTileSystem = senseTileSystem;
+           sensorList = new ArrayList<Sensor>();
+
+       }
+      /**
+       * set no of packets to read.
+       * @require num_packets >0;
+       * @require num_packets<=MAXNOPACKETS;
+       * @ensure numpackets = num_packets;
+       */
+       public void setNumPacketsToRead(final int numpackets) {
+           this.numpackets = numpackets;
+       }
+
+       /**
+        * register sensors.
+        */
+        public final void registerSensors(List<Sensor> sensorList) {
+            this.sensorList = sensorList;
+        }
+
+        public final List<Sensor> getSensors() {
+            return this.sensorList;
+        }
+
+        /**
+         * register sensor.
+         */
+         public final void registerSensor(Sensor sensor) {
+             this.sensorList.add(sensor);
+         }
+      /** 
+       * get no of packets to read.
+       */
+       public final /*@ pure @*/ int getNumPacketsToRead() {
+           return this.numpackets;
+       }
+
+       public void setPacketInputStream(final /*@ non_null @*/ PacketInputStream packetInputStream) {
+           this.packetInputStream = packetInputStream;
+       }
+       /**
+        * read packets from input stream.
+        */
+       public void readPackets() {
+           readpackets = true;
+           while (readpackets){
+               try{
+                   senseTileSystem.processSensorData(
[…]

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SenseTileSystem.java

@@ -1,0 +1,122 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.dataprovider.SensorObservations;
+import ie.ucd.sensetile.sensorboard.Packet;
+import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import ie.ucd.sensetile.webservice.sos.SensorObservationIF;
+import java.io.InputStream;
+import java.net.MalformedURLException;
+import java.rmi.Naming;
+import java.rmi.NotBoundException;
+import java.rmi.RemoteException;
+import java.util.ArrayList;
+import java.util.Iterator;
+import java.util.List;
+
+import org.apache.log4j.BasicConfigurator;
+import org.apache.log4j.Logger;
+import org.apache.log4j.NDC;
+import org.vast.data.DataBlockInt;
+import org.vast.process.DataProcess;
+
+/**
+ * Process packets and initiates processing of sensor data by the system.
+ *
+ * @author Ciaran Palmer
+ * @version 1
+ */
+public class SenseTileSystem {
+
+    private PacketStreamReader packetInputStream;
+
+    private SensorMLProcessEngine smlProcessEngine;
+    
+    private String dataproviderName = "rmi://localhost/SenseTileService";
+
+    
+    private static Logger logger =
+        Logger.getLogger(SenseTileSystem.class);
+
+    public SenseTileSystem() {
+       BasicConfigurator.configure();
+       init();
+    }
+    /**
+     * Get Clients for Observations.
+     */
+
+    /**
+     * Initialise the SensoMLProcessEngine with the sensorml files.
+     */
+     public final void init() {
+         InputStream inputxml = getSensorBoardInputStream();
+         smlProcessEngine = new SensorMLProcessEngine(inputxml);
+     }
+
+
+    /**
+     * ProcessSensorData - get packets and send
+     * data toSensorMLProcessEngine.
+     * @throws NotBoundException 
+     * @throws RemoteException 
+     * @throws MalformedURLException 
+     *
+     * @requires this.packetInputStream /= null
+     */
+     public final void processSensorData(final Packet packet) {
+         //logger.info("Received Packet to process");
+         DataBlockInt pktemp = new DataBlockInt(1);
+         pktemp.setIntValue(packet.getTemperature());
+         smlProcessEngine.setInput("txPacketTemperature",pktemp);
+         smlProcessEngine.execute();
+         
[…]

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/Sensor.java

@@ -1,0 +1,11 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.sensorboard.Packet;
+
+public interface Sensor {
+
+    String getName();
+
+    void processSensorData(Packet packet);
+
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SensorMLProcessEngine.java

@@ -1,0 +1,82 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import java.io.InputStream;
+import java.util.ArrayList;
+import java.util.List;
+
+import org.vast.cdm.common.DataComponent;
+import org.vast.data.DataBlockInt;
+import org.vast.process.DataProcess;
+import org.vast.process.ProcessException;
+import org.vast.sensorML.SMLException;
+import org.vast.sensorML.SystemReaderV1;
+import org.vast.sensorML.system.SMLSystem;
+import org.vast.xml.DOMHelper;
+import org.vast.xml.DOMHelperException;
+import org.w3c.dom.Element;
+
+public class SensorMLProcessEngine extends SMLSystem {
+
+    private SMLSystem smlsystem;
+    private SystemReaderV1 sysReaderv1;
+    private DOMHelper domHelper;
+
+    SensorMLProcessEngine(final InputStream inputxml) {
+        super();
+        initialise(inputxml);
+    }
+
+    /**
+     * Initialises the SensorMLProcessEngine.
+     * @param inputxml
+     */
+    private void initialise(final InputStream inputxml) {
+
+        try {
+
+            domHelper = new DOMHelper(inputxml, true);
+        } catch (DOMHelperException e1) {
+            e1.printStackTrace();
+        }
+
+        smlsystem = new SMLSystem();
+        sysReaderv1 = new SystemReaderV1();
+        Element processElt = domHelper.getElement("member/*");
+        try {
+            smlsystem = sysReaderv1.readSystem(domHelper, processElt);
+            smlsystem.init();
+         } catch (SMLException e) {
+            e.printStackTrace();
+         } catch (ProcessException e) {
+            e.printStackTrace();
+         }
+    }
+
+    public void execute(){
+        try {
+            smlsystem.execute();
+        } catch (ProcessException e) {
+            e.printStackTrace();
+        }
+    }
+
+    private DataComponent getInput(final String inputname) {
+       return smlsystem.getInputList().getComponent(inputname);
+    }
+
+
+    public void setInput(final String inputname, final DataBlockInt data) {
+        getInput(inputname).setData(data);
+    }
+
+    public double getOutput(final String outputname) {
+       return smlsystem.getOutputList().getComponent(outputname)
+       .getData().getDoubleValue();
+    }
+    
+    public List<DataProcess >getSensors(){
+        return smlsystem.getProcessList();
+    }
+
+
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/Thermistor.java

@@ -1,0 +1,27 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.sensorboard.Packet;
+
+import org.vast.process.ProcessException;
+
+public class Thermistor extends AbstractProcess implements Sensor {
+
+    private int sensordatasum = 0;
+    private int nosensordata = 0;
+
+    public void execute() throws ProcessException {
+        super.logger.info("Thermistor execute");
+        this.setIntOutput(sensordatasum / nosensordata);
+        sensordatasum = 0;
+        nosensordata = 0;
+    }
+
+    @Override
+    public final void processSensorData(final Packet packet) {
+        super.logger.info("Thermistor processSensorData");
+        sensordatasum += (int) packet.getTemperature();
+        nosensordata++;
+
+    }
+
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/sos/SensorObservationIF.java

@@ -1,0 +1,11 @@
+package ie.ucd.sensetile.webservice.sos;
+
+import java.rmi.Remote;
+import java.rmi.RemoteException;
+
+public interface SensorObservationIF extends Remote{
+
+   public void insertObservation(int obs) throws RemoteException;
+
+   public void resgistorSensor(String sensorId) throws RemoteException;
+}

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/org/vast/sensorML/ProcessLoader.java

@@ -1,0 +1,196 @@
+/***************************** BEGIN LICENSE BLOCK ***************************
+
+ The contents of this file are subject to the Mozilla Public License Version
+ 1.1 (the "License"); you may not use this file except in compliance with
+ the License. You may obtain a copy of the License at
+ http://www.mozilla.org/MPL/MPL-1.1.html
+ 
+ Software distributed under the License is distributed on an "AS IS" basis,
+ WITHOUT WARRANTY OF ANY KIND, either express or implied. See the License
+ for the specific language governing rights and limitations under the License.
+ 
+ The Original Code is the "SensorML DataProcessing Engine".
+ 
+ The Initial Developer of the Original Code is the VAST team at the
 University of Alabama in Huntsville (UAH). <http://vast.uah.edu>
 Portions created by the Initial Developer are Copyright (C) 2007
 the Initial Developer. All Rights Reserved.

 Please Contact Mike Botts <mike.botts@uah.edu> for more information.
+ 
+ Contributor(s): 
+    Alexandre Robin <robin@nsstc.uah.edu>
+ 
+******************************* END LICENSE BLOCK ***************************/
+
+package org.vast.sensorML;
+
+import org.w3c.dom.*;
+import org.vast.xml.DOMHelper;
+import org.vast.xml.DOMHelperException;
+import org.vast.process.*;
+
+import java.io.FileInputStream;
+import java.io.FileNotFoundException;
+import java.lang.reflect.*;
+import java.util.Hashtable;
+
+
+/**
+ * <p><b>Title:</b><br/>
+ * Process Loader
+ * </p>
+ *
+ * <p><b>Description:</b><br/>
+ * This class loads the right process based on the method URI
+ * provided in the SensorML description. If the URI is known locally,
+ * the corresponding class is instantiated, otherwise it is resolved 
+ * and the corresponding ProcessMethod is fetched from registry and
+ * further parsed. A new process implementation is eventually
+ * downloaded if available.
+ * </p>
+ *
+ * <p>Copyright (c) 2005</p>
+ * @author Alexandre Robin
+ * @date Feb 15, 2006
+ * @version 1.0
+ */
+public class ProcessLoader
+{
+    public static Hashtable<String,String> processMap = null;
+    
+    
+    public ProcessLoader()
+    {
+        
+    }    
+        
+
+    /**
+     * Loads the DataProcess implementation corresponding to the given URI.
+     * @param uri The URI of the method (i.e. the specific process to instantiate)
[…]

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/org/vast/sensorML/ProcessReaderV1.java

@@ -1,0 +1,499 @@
+/***************************** BEGIN LICENSE BLOCK ***************************
+
+ The contents of this file are subject to the Mozilla Public License Version
+ 1.1 (the "License"); you may not use this file except in compliance with
+ the License. You may obtain a copy of the License at
+ http://www.mozilla.org/MPL/MPL-1.1.html
+ 
+ Software distributed under the License is distributed on an "AS IS" basis,
+ WITHOUT WARRANTY OF ANY KIND, either express or implied. See the License
+ for the specific language governing rights and limitations under the License.
+ 
+ The Original Code is the "SensorML DataProcessing Engine".
+ 
+ The Initial Developer of the Original Code is the VAST team at the
 University of Alabama in Huntsville (UAH). <http://vast.uah.edu>
 Portions created by the Initial Developer are Copyright (C) 2007
 the Initial Developer. All Rights Reserved.

 Please Contact Mike Botts <mike.botts@uah.edu> for more information.
+ 
+ Contributor(s): 
+    Alexandre Robin <robin@nsstc.uah.edu>
+ 
+******************************* END LICENSE BLOCK ***************************/
+
+package org.vast.sensorML;
+
+import java.net.URL;
+
+import org.w3c.dom.*;
+import org.vast.process.*;
+import org.vast.cdm.common.CDMException;
+import org.vast.cdm.common.DataComponent;
+import org.vast.xml.DOMHelper;
+import org.vast.sensorML.metadata.Metadata;
+import org.vast.sweCommon.SWECommonUtils;
+import org.vast.util.*;
+
+
+/**
+ * <p><b>Title:</b><br/>
+ * Process Reader v1.0
+ * </p>
+ *
+ * <p><b>Description:</b><br/>
+ * Reader for Process, ProcessModel, ProcessChain, ProcessList
+ * Uses Reflection to create particular ProcessModel for plug'n play capability
+ * </p>
+ *
+ * <p>Copyright (c) 2005</p>
+ * @author Alexandre Robin
+ * @version 1.0
+ */
+public class ProcessReaderV1 extends AbstractSMLReader implements ProcessReader
+{
+    protected static final String dataSeparator = "/"; 
+    protected MetadataReaderV1 metadataReader;
+    protected ProcessLoader processLoader;
+    protected SWECommonUtils utils;
+    protected boolean readMetadata = false;
+    protected boolean createExecutableProcess = true;
+        
+    
+    /**
[…]

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/main/src/org/vast/xml/DOMHelper.java

@@ -108,7 +108,7 @@ public class DOMHelper
     {
         this.validation = validation;
         //eeicpr
-        System.out.println("DomHelper const");
+        System.out.println("DomHelper constr");
         try {
 			mainFragment = parseStream(inputStream, true);
 	        System.out.println("mainFragment "+mainFragment.getBaseElement().getPrefix());
@@ -1128,7 +1128,7 @@ public class DOMHelper
         //eeicpr
         newDocument = new XMLDocument(inputStream, validation);
 		//newDocument.setUri(new URI("file:///c://masters//tempproj//SensorML//Schemas//Sensors//SenseTileSensorBoardThermistor.xml"));
-		newDocument.setUri((new File(".")).toURI());
+		//newDocument.setUri((new File(".")).toURI());
         System.out.println("current dir "+System.getProperty("user.dir"));
 
         System.out.println("parseStream "+inputStream.toString());
@@ -1251,8 +1251,11 @@ public class DOMHelper
                     System.out.println("link to Uri "+baseUri);
                     if (baseUri==null){
                     	System.out.println("base uri is null ");
+                    	linkUri =this.getClass().getResource("/"+idRef).toURI();
                     }
-                    linkUri = baseUri.resolve(linkUri);
+                    else {
+                       linkUri = baseUri.resolve(linkUri);
+                    }
                 }     
                 xmlFragment = parseURI(linkUri.toString(), false);
             }

### 2009-11-23 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex

--- 2009-11-02 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
@@
 
-Problem to solve
-Whay is it interesting?
-How is the problem going to be solved
-Why is it innovate?
+The aim of this thesis project was to prototype a Web Service for accessing sensor observations from the SenseTile Sensor Board.
+SenseTile is a sensor and processing package including motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
+It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
+
+Descriptions of the SenseTile Sensor Board were to be developed using Sensor Model Language (SensorML)\cite{SensoMLref}.
+SensorML is a language that describes the process of sensor measurement by providing standard models and XML encoding for such processes.
+SensorML was be used as it is becoming a standard for describing sensor data processing as part of the SWE[ref].
+
+A  SensorML to BON Mapping was also to be explored .
+BON\cite{BONref} is a notation and a method for Object Oriented systems analysis and design.
+Based on this mapping it was proposed to develop a tool/method to automatically generate SensorML descriptions.
+Generally SensorML is hand generated with all the shortcomings of such an approach.
+The proposed approach would allow sensor processing to defined formally in BON and then refined to JML/Java.
 
@@
 \chapter{SensorML}
-Basic description.
-How the sy
+\section{SensorML Overview}
+The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standerised “Sensor Web”.
+It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.
+Sensor Modelling Language (SensorML) is one part of the SWE initiative and is used to describe sensor systems and the processing of observations from sensor systems.
+
+The purposes of SensorML are to:
+• Provide descriptions of sensors and sensor systems for inventory management
+• Provide sensor and process information in support of resource and observation discovery
+• Support the processing and analysis of the sensor observations
+• Support the geolocation of observed values (measured data)
+• Provide performance characteristics (e.g., accuracy, threshold, etc.)
+• Provide an explicit description of the process by which an observation was obtained (i.e., it’s lineage)
+• Provide an executable process chain for deriving new data products on demand (i.e., derivable observation)
+• Archive fundamental properties and assumptions regarding sensor systems
+
+SensorML provides a functional model of sensors and an XML encoding to describe sensors and their observations.
+Sensors are modelled as processes that convert real phenomena to data.
+Processes take inputs, process them and result in one or several outputs.
+They can also define relevant metadata.
+The processes can be connected together in chains using SensorML ProcessChains or Systems.
+
+When Modelling a sensor system in SensorML the main entities used are Systems and Components.
+A SensorML System is used to group sensors that are related spatially to one another.
+A Component is an atomic process that generally converts a physical phenomen to a digital number.
+ProcessModels are another modelling that is used to describe a pure process that has no physicality
+
+Quantity, Count, Boolean, Category, and Time provide the basic primitives. Within SensorML, these serve as the basis for specifying all inputs, outputs, and parameters within a Process
+
+DataRecord or DataArray.
+
+Detector model.
+
+The SensorML description of the SenseTile Sensor Board is shown in the following section.
+
+\section{SensorML Description of SenseTile}
+
+SensorML System and Component were used to describe the SenseTile system and it's sensors. The sensor specifications were used to generate the metadata.
+The Sensors are modelled a detectors. They are process that take scalar values
+and can perform transformations on them.
+
+System is providing the mapping of inputs to processes and the outputs.
+
+It is further recognized that most sensor observations consist of at least two processes: a sampling process and a conversion process.
+The 
+ \begin{figure}
+\includegraphics{SensorML_SenseTile_System_design_tuple_pa3.png}
+\caption{Block diagram of SenseTile System process}\label{fig:SensorML_SenseTile_design.png}
+\end{figure}
+
+\subsection{SenseTile System}
+\subsection{Light Sensor}
+\subsection{Thermistor}
+\subsection{Pressure Sensor}
+\subsection{Accelerometer}
+\subsection{Acoustic Sensor}
+
+Types
+ - SWE -
+
+Count - BigInteger
+Quantity - Double
+
+ * Carries an array of int primitives.
+ * All data is casted to other types when requested.
+
+public class DataBlockInt extends AbstractDataBlock
+
+\newpage
+\section{BON to SensorML mapping}
 
@@
 High level design
+\section{SenseTile Web Service}
+
+DataProducer - use PacketInput stream to access SenseTile observation packets.
+
+Steps in the process these sensorml components are roughly.
+
+System reads packets and averages them for 1 second.
+
+Thermistor components reads out the temperature value
+and its output is sent to both the raw data and the celcius
+converter component and on to the tuples.
+
+These tuples are sent to webservice provider as observations.
+
+DataProvider - provides an RMI interface to the DataProducer to update. Provides a Web Service interface  based on Sensor Observation Service. Not fully as is a very complex interface.
+
+\section{SensorML generation tool}
+
+SensorML to BON mapping
+
+BON to Java
+
+Instantiate Java classes with data
+
+Walk Java class and generate sensorml.
 
@@
 
-implementation technology used
-issues that arouse
+\section{SenseTile WebService}
+
+DataProvider is based on AXIS 2.0/Spring
+
+DataProducer POJO/VastLib/UCD SenseTile Driver Lib
 
 \chapter{Results}
+
+Tested against real SenseTile and sensor data retrieved by a testclient.
 

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
add static diagram

@@
 
+\subsection{Sensor Networks}
+\subsection{BON}
+\subsection{Sensorml}
+reference chapter
+\subsection{WebService}
+\subsection{JIBX}
+
 \chapter{SensorML}
@@
 
+Naming rules inputs and outputs - allow more generic handling of implementation classes.
+
+Processed verus raw data.
+
 It is further recognized that most sensor observations consist of at least two processes: a sampling process and a conversion process.
@@
 
+DataArray
+
  * Carries an array of int primitives.
@@
 \section{SenseTile Web Service}
-
+ \begin{figure}
+\scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
+\caption{SenseTile Static Diagram}\label{fig:sensetile_static_diagam.png}
+\end{figure}
 DataProducer - use PacketInput stream to access SenseTile observation packets.
@@
 
-System reads packets and averages them for 1 second.
+System reads packets and averages them for every configurable seconds. Defalut 1.
 
@@
 
+\newpage
 \section{SensorML generation tool}
 
-SensorML to BON mapping
+BON to SensorML mapping.
 
-BON to Java
+Features to be elements or attributes.
 
-Instantiate Java classes with data
+Name rule suffix "Attrib" to make it an attribute.
 
-Walk Java class and generate sensorml.
+List handling - convention for mapping.
+
+Java binding wth JIBX  generate sensorml.
+
+Simple xml file with Sensor metadata.
+
+Naming rules allow connection of inputs and outputs
+in sensorML processes.
 
@@
 
+RMI used between DataProvider and DataProducer.
 \chapter{Results}
@@
 \chapter{References}
-
-\chapter{Appendices}
-
 \newpage
@@
 
+\chapter{Appendices}
+
 \end{document}

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/AbstractProcess.java

@@ -8,7 +8,8 @@ import org.vast.data.DataValue;
 import org.vast.process.DataProcess;
 import org.vast.process.ProcessException;
 
-enum ProcessType {SENSOR,CONVERTER}
+enum ProcessType { SENSOR, CONVERTER }
+
 public abstract class AbstractProcess extends DataProcess {
 
     private DataValue input;
@@ -32,17 +33,20 @@ public abstract class AbstractProcess extends DataProcess {
         this.processType = processType;
     }
     /**
-     * * Initializes the process.
+     * * Initialises the process.
      * * Gets handles to input/output components.
+     * @throws ProcessException - Vast Library processing exception
      */
-    public void init() throws ProcessException {
+    public final void init() throws ProcessException {
         try {
-            System.out.println("abstact name ???? "+this.getName());
+            System.out.println("abstact name ???? " + this.getName());
             // I/O mappings
             input =
                 (DataValue) inputData.getComponent(this.getName() + "Input");
+
             output =
                 (DataValue) outputData.getComponent(this.getName() + "Output");
+
         } catch (ClassCastException e) {
             throw new ProcessException("Invalid I/O data", e);
         }
@@ -63,9 +67,10 @@ public abstract class AbstractProcess extends DataProcess {
     protected final void setIntOutput(final int value) {
         output.getData().setIntValue(value);
     }
-    
+
     /**
     * Executes process algorithm on inputs and set output data.
+    * @throws ProcessException  - Vast Library processing exception
     */
     public abstract void execute() throws ProcessException;
 

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/CelciusConverter.java

@@ -2,15 +2,32 @@ package ie.ucd.sensetile.webservice.dataproducer;
 
 import org.vast.process.ProcessException;
 
-public class CelciusConverter extends AbstractProcess
-{
-    public static double CONVERSIONVALUE = 16.0;
-    
-    public void execute() throws ProcessException {
-        super.logger.info("Thermistor execute");
+/**
+ * converts sensor data to a celcius value.
+ */
+public class CelciusConverter extends AbstractProcess {
+    /**
+     * conversion value need to divide by to get a celcius value.
+     */
+    private static double CONVERSIONVALUE = 16.0;
 
-        double temp = this.getDoubleInput();
-        this.setDoubleOutput(temp /CONVERSIONVALUE);
+    /**
+     *set type at construction.
+     */
+    public CelciusConverter() {
+        super();
+        this.setProcessType(ProcessType.CONVERTER);
+    }
+
+    /**
+     * process sensor data and output.
+     * @throws ProcessException - Vast exception
+     */
+    public final void execute() throws ProcessException {
+        super.logger.info("CelciusConveter execute");
+
+        double temp = this.getIntInput();
+        this.setDoubleOutput(temp / CONVERSIONVALUE);
     }
 
 }

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/PacketStreamReader.java

@@ -1,9 +1,14 @@
 package ie.ucd.sensetile.webservice.dataproducer;
 
+import java.io.IOException;
 import java.util.ArrayList;
+import java.util.Iterator;
 import java.util.List;
 
+import ie.ucd.sensetile.sensorboard.Packet;
 import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import ie.ucd.sensetile.sensorboard.SenseTileException;
+
 import org.apache.log4j.Logger;
 import org.apache.log4j.BasicConfigurator;
 import org.apache.log4j.NDC;
@@ -32,7 +37,7 @@ public class PacketStreamReader {
        public PacketStreamReader(SenseTileSystem senseTileSystem) {
            BasicConfigurator.configure();
            this.senseTileSystem = senseTileSystem;
-           sensorList = new ArrayList<Sensor>();
+           sensorList = new ArrayList < Sensor >();
 
        }
       /**
@@ -48,21 +53,21 @@ public class PacketStreamReader {
        /**
         * register sensors.
         */
-        public final void registerSensors(List<Sensor> sensorList) {
-            this.sensorList = sensorList;
+        public final void registerSensors(final List < Sensor > sensorlist) {
+            this.sensorList = sensorlist;
         }
 
-        public final List<Sensor> getSensors() {
+        public final List < Sensor > getSensors() {
             return this.sensorList;
         }
 
         /**
          * register sensor.
          */
-         public final void registerSensor(Sensor sensor) {
+         public final void registerSensor(final Sensor sensor) {
              this.sensorList.add(sensor);
          }
-      /** 
+      /**
        * get no of packets to read.
        */
        public final /*@ pure @*/ int getNumPacketsToRead() {
@@ -72,23 +77,46 @@ public class PacketStreamReader {
        public void setPacketInputStream(final /*@ non_null @*/ PacketInputStream packetInputStream) {
            this.packetInputStream = packetInputStream;
        }
+
        /**
         * read packets from input stream.
         */
        public void readPackets() {
-           readpackets = true;
-           while (readpackets){
-               try{
-                   senseTileSystem.processSensorData(
-                           this.packetInputStream.read());
-                   
-                   Thread.currentThread().sleep(5000);//sleep for 1000 ms
 
+
+
+           readpackets = true;
+           while (readpackets) {
+               try {
+                   Thread.currentThread().sleep(5000);
+                 } catch (InterruptedException e1) {
+                   // TODO Auto-generated catch block
+                   e1.printStackTrace();
                  }
[…]

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SenseTileSystem.java

@@ -33,7 +33,8 @@ public class SenseTileSystem {
     
     private String dataproviderName = "rmi://localhost/SenseTileService";
 
-    
+    SensorObservationIF sensorObs = null;
+
     private static Logger logger =
         Logger.getLogger(SenseTileSystem.class);
 
@@ -51,15 +52,27 @@ public class SenseTileSystem {
      public final void init() {
          InputStream inputxml = getSensorBoardInputStream();
          smlProcessEngine = new SensorMLProcessEngine(inputxml);
+         try {
+             sensorObs = (SensorObservationIF) Naming.lookup(dataproviderName);
+         } catch (MalformedURLException e1) {
+             // TODO Auto-generated catch block
+             e1.printStackTrace();
+         } catch (RemoteException e1) {
+             // TODO Auto-generated catch block
+             e1.printStackTrace();
+         } catch (NotBoundException e1) {
+             // TODO Auto-generated catch block
+             e1.printStackTrace();
+         }
      }
 
 
     /**
      * ProcessSensorData - get packets and send
      * data toSensorMLProcessEngine.
-     * @throws NotBoundException 
-     * @throws RemoteException 
-     * @throws MalformedURLException 
+     * @throws NotBoundException
+     * @throws RemoteException
+     * @throws MalformedURLException
      *
      * @requires this.packetInputStream /= null
      */
@@ -69,13 +82,13 @@ public class SenseTileSystem {
          pktemp.setIntValue(packet.getTemperature());
          smlProcessEngine.setInput("txPacketTemperature",pktemp);
          smlProcessEngine.execute();
-         
+
          //logger.info("result "+smlProcessEngine.getOutput("temperature"));
-         
-         SensorObservationIF sensorObs =null;
-         
+
+         SensorObservationIF sensorObs = null;
+
         try {
-            sensorObs = (SensorObservationIF)Naming.lookup(dataproviderName);
+            sensorObs = (SensorObservationIF) Naming.lookup(dataproviderName);
         } catch (MalformedURLException e1) {
             // TODO Auto-generated catch block
             e1.printStackTrace();
@@ -86,8 +99,8 @@ public class SenseTileSystem {
             // TODO Auto-generated catch block
             e1.printStackTrace();
         }
-         
-         try {
+
+        try {
              
             sensorObs.insertObservation((int)smlProcessEngine.getOutput("temperature"));
         } catch (RemoteException e) {
@@ -98,17 +111,22 @@ public class SenseTileSystem {
 
     /**
[…]

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SensorMLProcessEngine.java

@@ -73,7 +73,11 @@ public class SensorMLProcessEngine extends SMLSystem {
        return smlsystem.getOutputList().getComponent(outputname)
        .getData().getDoubleValue();
     }
-    
+
+    public double getIntOutput(final String outputname) {
+        return smlsystem.getOutputList().getComponent(outputname)
+        .getData().getIntValue();
+     }
     public List<DataProcess >getSensors(){
         return smlsystem.getProcessList();
     }

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/Thermistor.java

@@ -1,27 +1,0 @@
-package ie.ucd.sensetile.webservice.dataproducer;
-
-import ie.ucd.sensetile.sensorboard.Packet;
-
-import org.vast.process.ProcessException;
-
-public class Thermistor extends AbstractProcess implements Sensor {
-
-    private int sensordatasum = 0;
-    private int nosensordata = 0;
-
-    public void execute() throws ProcessException {
-        super.logger.info("Thermistor execute");
-        this.setIntOutput(sensordatasum / nosensordata);
-        sensordatasum = 0;
-        nosensordata = 0;
-    }
-
-    @Override
-    public final void processSensorData(final Packet packet) {
-        super.logger.info("Thermistor processSensorData");
-        sensordatasum += (int) packet.getTemperature();
-        nosensordata++;
-
-    }
-
-}

### 2009-11-24 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/ThermistorSensor.java

@@ -1,0 +1,51 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.sensorboard.Packet;
+
+import org.vast.process.ProcessException;
+
+/**
+ *sum of sensor data readings received.
+ */
+public class ThermistorSensor extends AbstractProcess implements Sensor {
+
+    /**
+     *sum of sensor data readings received.
+     */
+    private int sensordatasum = 0;
+
+    /**
+     *number of sensor data readings received.
+     */
+    private int nosensordata = 0;
+
+    /**
+     *set type at construction.
+     */
+    public ThermistorSensor() {
+        super();
+        this.setProcessType(ProcessType.SENSOR);
+    }
+
+    /**
+     * process sensor data and output.
+     * @throws ProcessException - Vast exception
+     */
+    public final void execute() throws ProcessException {
+        AbstractProcess.logger.info("Thermistor execute");
+        this.setIntOutput(sensordatasum / nosensordata);
+        sensordatasum = 0;
+        nosensordata = 0;
+    }
+
+    /**
+     *sum of sensor data readings received.
+     *@param packet Packet
+     */
+    @Override
+    public final void processSensorData(final Packet packet) {
+        AbstractProcess.logger.info("Thermistor processSensorData");
+        sensordatasum += (int) packet.getTemperature();
+        nosensordata++;
+    }
+}

### 2009-11-26 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
outline version.

@@
 It is further recognized that most sensor observations consist of at least two processes: a sampling process and a conversion process.
-The 
+
  \begin{figure}
@@
 
-public class DataBlockInt extends AbstractDataBlock
-
 \newpage
 \section{BON to SensorML mapping}
+Features to be elements or attributes.
 
+Name rule suffix "Attrib" to make it an attribute.
+
+List handling - convention for mapping.
+
+SML prefix to handled key word clashes between sensorML and BON.
+
+How to display this in thesis?
 \chapter{Design}
-High level design
+
 \section{SenseTile Web Service}
+\subsection{Overview}
+ \begin{figure}
+\scalebox{0.65}{\includegraphics{SystemOverview.png}}
+\caption{SenseTile System Overview}\label{fig:SystemOverview.png}
+\end{figure}
+
  \begin{figure}
 \scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
-\caption{SenseTile Static Diagram}\label{fig:sensetile_static_diagam.png}
+\caption{DataProducer Static Diagram}\label{fig:sensetile_static_diagam.png}
 \end{figure}
-DataProducer - use PacketInput stream to access SenseTile observation packets.
 
-Steps in the process these sensorml components are roughly.
+\subsection{DataProducer}
+Use PacketInput stream to access SenseTile observation packets.
 
-System reads packets and averages them for every configurable seconds. Defalut 1.
+System reads packets and averages them for every configurable seconds. Default 1.
 
@@
 
-DataProvider - provides an RMI interface to the DataProducer to update. Provides a Web Service interface  based on Sensor Observation Service. Not fully as is a very complex interface.
+Observation as described in SOS.
+
+\subsection {DataProvider}
+
+Keeps track of sensorboards that register
+
+provides an RMI interface to the DataProducer to update with observations.
+
+GetObservaton - marshal Observation to xml string and send as response.
+
+StreamObservation -  registered clients get new observation sent over UDP connection?
+
+Provides a Web Service interface  based on Sensor Observation Service. Not fully as is a very complex interface.
 
 \newpage
-\section{SensorML generation tool}
-
-BON to SensorML mapping.
-
-Features to be elements or attributes.
-
-Name rule suffix "Attrib" to make it an attribute.
-
-List handling - convention for mapping.
+\section{SensorML generation}
 
@@
 in sensorML processes.
+
+ component id +Input
+
+ component id+Output
 
@@
 RMI used between DataProvider and DataProducer.
+
+\section{SensorML Generation}
+
 \chapter{Results}
@@
 Tested against real SenseTile and sensor data retrieved by a testclient.
+
+Generation of SensorML description of a Senstile Sensor
 

### 2009-12-01 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
New brutal/punchy abstract.

@@
 
-  Sensor equipped devices typically communicate with a streaming
-  packet-based network protocol.  The communication is focused on
-  controlling the sensors and reporting the sensors measurements.  The
-  \STSB is an example of such a device: it communicates with an
-  arbitrary computer system, the \STPU, over a USB bus using a
-  streaming packet-based protocol.  The \STU is a custom-designed, but
-  inexpensive, sensor platform: it is composed by a one or more \STSBs
-  paired with a general-purpose small-scale compute node, the \STPU
-  (including everything from PDAs to powerful touchscreen-based
-  portable computers).  There are over a dozen sensors on the \STSB
-  itself, and new sensors can be added either using expansion slots on
-  the \STSB, or adding them directly to the USB bus.  A third-party
-  low-level library creates and parses generic USB communication
-  packets; the library has been provided by the vendor of the USB chip
-  vendor, the USB chip that has been used in the \STSB.
-
-  A custom medium-level Java driver creates and parses the \STSB
-  streaming communication packets.  The driver is formally specified,
-  validated and verified with rigorous unit testing, both hand made
-  and automatically generated.  Because of this process and
-  combination of technologies, we are able to verify a real \ST and a
-  simulated one as well.  The system has been developed concurrently
-  with the protocol specification and the real \STSB, the boards have
-  not been available until the last months of development.  The
-  protocol model is itself an executable stream simulator that can be
-  used to develop new \ST user applications without having the \ST
-  hardware at all.  Consequently, from this single specification, we
-  are able to test that the \ST hardware communicates as expected,
-  that our \ST simulator communicates as expected, and simulate the
-  \ST protocol stream itself.
+%  A punchy abstract that makes the reader want to know more.
+
+  Agile people hate formal people.  Formal people think agile people
+  are flakes.  These two worlds do not talk to or learn from each
+  other.  This is bullshit.  This paper summarizes a complex case
+  study which shows that it is not only possible, but tasty, to
+  combine the garlic of formal methods and the chocolate of agile
+  programming.
+
+%   Sensor equipped devices typically communicate with a streaming
+%   packet-based network protocol.  The communication is focused on
+%   controlling the sensors and reporting the sensors measurements.  The
+%   \STSB is an example of such a device: it communicates with an
+%   arbitrary computer system, the \STPU, over a USB bus using a
+%   streaming packet-based protocol.  The \STU is a custom-designed, but
+%   inexpensive, sensor platform: it is composed by a one or more \STSBs
+%   paired with a general-purpose small-scale compute node, the \STPU
+%   (including everything from PDAs to powerful touchscreen-based
+%   portable computers).  There are over a dozen sensors on the \STSB
+%   itself, and new sensors can be added either using expansion slots on
+%   the \STSB, or adding them directly to the USB bus.  A third-party
+%   low-level library creates and parses generic USB communication
+%   packets; the library has been provided by the vendor of the USB chip
+%   vendor, the USB chip that has been used in the \STSB.
+
+%   A custom medium-level Java driver creates and parses the \STSB
+%   streaming communication packets.  The driver is formally specified,
+%   validated and verified with rigorous unit testing, both hand made
+%   and automatically generated.  Because of this process and
+%   combination of technologies, we are able to verify a real \ST and a
+%   simulated one as well.  The system has been developed concurrently
+%   with the protocol specification and the real \STSB, the boards have
+%   not been available until the last months of development.  The
+%   protocol model is itself an executable stream simulator that can be
+%   used to develop new \ST user applications without having the \ST
+%   hardware at all.  Consequently, from this single specification, we
+%   are able to test that the \ST hardware communicates as expected,
+%   that our \ST simulator communicates as expected, and simulate the
+%   \ST protocol stream itself.
 

### 2010-01-04 papers/BON_VHDL/paper.tex
New abstract written in OmmWriter.  Revised abstract and most of introduction.

@@
 
-The \textbf{UCD CASL SenseTile System} is a large-scale,
-general-purpose sensor system installed at the University College
-Dublin in Dublin, Ireland.  \textbf{Our facility provides a
-capability---unique in the world in terms of its scale and
-flexibility---for performing in-depth investigation of both the
-specific and general issues in large-scale sensor networking.}
-
-This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed, but inexpensive,
-sensor platform (called the \ST) paired with general-purpose
-small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers.  The datastore is a
-multi-terabyte scale scientific datastore into which sensor data
-flows, and in which online and offline scientific computation of
-sensor and other scientific data takes place.  The compute farm is a
-large-scale Linux, Solaris, and OS~X-based compute farm used by
-scientists for these online and offline scientific computations.
-
-The \STs, as an embedded system, have several unusual built-in sensors
-and emitters, sensor nodes can be extended via a general-purpose USB
-bus, and \STs contain an FPGA that is dynamically updated on a
-per-experiment basis.  \STs are custom-designed in VHDL-AMS.  Like
-many hardware-based systems, to write new software for or against the
-\ST architecture one either uses the actual hardware or one runs a
-\STS.  But this simulator emulates a piece of hardware, one would like
-to know if the simulator \emph{actually} simulates the hardware.  We
-have developed an approach through which we refine the VHDL-AMS
-specification up to the formal BON specification language, and from
-this specification we generate an executable, functional model-based
-specification in the JML language.  We have used this specification to
-implement a fully validated and verified \ST simulator.
+Complex software engineering projects often use high-level
+specification languages to describe system abstractions.  These
+abstractions, which typically capture static and dynamic architectural
+aspects, are meant to be more approachable, comprehensible, and useful
+than witnessing the gory details of the systems they describe.  In the
+hardware world, systems are described with hardware description
+languages like VHDL.  These languages exist at a similar level of
+abstraction as traditional software programming languages, like Java.
+But whence are the high-level hardware description languages?
+Moreover, if one is doing hardware/software co-design, what
+architecture description languages are usable and useful to both
+facets of the problem domain?  In this paper, we describe a
+non-trivial co-design effort, called SenseTile, and how requirements
+on its design and verification led us to define a formal refinement of
+VHDL-AMS to the EBON high-level specification language.  This
+refinement enables us to abstract away, in a seamless, reversible, and
+contract-based fashion, from the details of a particular hardware
+system to an neutral architecture description which, in turn, can be
+refined to an executable Java-based software simulator.  This pair of
+refinements permits full round-trip hardware/software co-design with
+verification, as the Java-based system is annotated with formal
+specifications using the Java Modeling Language.
 
@@
 \label{sec:introduction}
+
+The \textbf{UCD CASL SenseTile System} is a large-scale,
+general-purpose sensor system under development at the University
+College Dublin in Dublin, Ireland.  \emph{After its installation, it
+  is our intention that this facility will provide a
+  capability---unique in the world in terms of its scale and
+  flexibility---for performing in-depth investigation of both the
+  specific and general issues in large-scale sensor networking.}  This
+system integrates a \emph{sensor platform}, a \emph{datastore}, and a
+\emph{compute farm}.
+
+The \emph{sensor platform} is a custom-designed, but inexpensive,
+sensor platform called the \ST.  Each \ST is paired with
+general-purpose, small-scale compute node ranging is capability from a
+low-power PDA to a powerful touchscreen-based portable computer.
+
+The \emph{datastore} is a multi-terabyte scale scientific datastore
+located in a server room across campus.  Sensor data flows into this
+datastore during experiments.  It is also used for online and offline
+computation of sensor and other scientific data.
+
+The \emph{compute farm} is a large-scale Linux, Solaris, Windows, and
+OS~X-based compute farm used by scientists for these computations.
+
+\subsection{The \ST}
+\label{subsec:st}
+
+The \STs contain several built-in sensors, ranging from typical
+sensors like temperature and humidity to more unusual ones like
+ultrasonic microphones.  They also contain several emitters, including
+visible and infrared light, audible and ultrasonic sound, and more.
+Sensor nodes are extended via a USB bus, and \STs contain an FPGA that
+is dynamically updated on a per-experiment basis.
+
+\STs are custom-designed by a third party based upon our high-level
+specifications in VHDL-AMS.  Common to many hardware prototype
+development efforts, writing software for the \ST architecture means
+either using the actual hardware, or running a \STS.  But this
+simulator must emulate a piece of hardware \emph{that does not yet
+  exist}.  Worse, as is typical with these kinds of development
+efforts, the hardware is several quarters late, and the show must go
+on.
+
+Consequently, the \STS becomes the primary focus of development
+efforts and, as we do not wish to loose any development time, we would
+like to know if the simulator \emph{actually} simulates the hardware,
+thus ensuring that the software we write for the simulator will run
+immediately on the hardware when it arrives.
+
+\subsection{Software}
+\label{subsec:software}
+
+Complementing these hardware development efforts, the software
+architecture that runs the \STs is under development.  This system
+consists of a low-level USB driver written in C and a packet framer
+and control system written in Java.
+
+It is vitally important that these low-level software systems are
+well-understood, work correctly from day zero, and have adequate
+non-functional characteristics---primarily that they do not use too
+much memory or CPU on the low-power devices for which they are
+destined.  As such, and following the standard development conventions
+of our research group, they are formally specified using the BON and
+JML specification languages.  Moreover, they are validated and
+verified using various forms of testing and static analysis, all of
+which are the focus of several other papers.
+
+\subsection{Co-design}
+\label{subsec:co-design}
+
+Even in the best of hardware-software co-design processes, hardware
+development typically focuses on correctness through rigorous
+specification, static analysis, and simulation, while software
+development uses the ``by the seat of the pants'' approach.  As is
+obvious from the previous section, that does not fly in our research
+group.
+
+Consequently, we have developed an approach through which we abstract
+(via a formal refinement) a VHDL-AMS specification up to a formal BON
+specification.  Using this specification, we then generate (again, via
+a formally defined refinement) an executable, functional, model-based
+specification in the JML language.  We have subsequently used this
+specification to implement a fully validated and verified \ST
+simulator.  Due to the semantic properties of these refinements, as
+seen in~\autoref{diagram}, we are guaranteed that all functional
+properties of the original VHDL-AMS specification are preserved in the
+software simulator, thereby witnessing full hardware-software
+co-design in a formal sense.
+
+XXX
 

### 2010-01-06 papers/BON_VHDL/paper.tex
Rewrote another section or two.

@@
 \newcommand{\through}{\emph{through}\xspace}
+\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}
+    #1}}
 
@@
 specification to implement a fully validated and verified \ST
-simulator.  Due to the semantic properties of these refinements, as
-seen in~\autoref{diagram}, we are guaranteed that all functional
-properties of the original VHDL-AMS specification are preserved in the
-software simulator, thereby witnessing full hardware-software
-co-design in a formal sense.
+simulator.
 
-XXX
+Due to the semantic properties of these refinements, as seen
+in~\autoref{diagram}, we are guaranteed that all functional properties
+of the original VHDL-AMS specification are preserved in the software
+simulator, thereby witnessing full hardware-software co-design in a
+formal sense.
 
-The \ST sensor platform, represents heterogeneous collection of
-digital, analog and mixed signal hardware circuits.  The complexity of
-our platform requires the need to strive for formal descriptions each
-component separately and behavior of their complex interdependencies,
-in order to produce executable specification.\footnote{This is not a
-  good reason for performing this research.  Complexity is often the
-  reason to use \emph{less} formality, nor more.  Also, why must the
-  specification be \emph{executable}?}  To make sure that the sensor
-platform works properly we need to translate the VHDL-AMS code into
-the specification, and to refine and verify this
-specification.\footnote{Too vague a statement---what are the
-  \emph{real} reasons we are doing this work?}  In other words, our
-aim is to establish formal link between behavior of a given sensor
-platform, and derived specification of components in order to
-implement fully validated and verified \ST simulator.  For this
-purpose, we need notation and methodology which will be able to
-satisfy the requirements given below.
+In summary, this article provides three main contributions.
+\begin{enumerate}
+\item We define a relation between a hardware description language,
+  VHDL-AMS, and a high-level specification language (BON).  This
+  relation means that, given a well-formed VHDL-AMS program, it can be
+  described using a BON (informal/formal) specification, and hence to
+  a JML implementation, through a sequence of refinements.
+\item We describe how to formally emulate piece of hardware using a
+  refinement-based approach.  Our approach uses object-oriented
+  languages, including a high-level and a low-level specification
+  language, BON and JML respectively.
+\item We introduce the use of \emph{invariants} as a general concept
+  for modeling hardware components behavior and their interactions.
+  To be more precise, the behavioral specification of a hardware
+  component is represented as an invariant on the signals at a
+  circuit's ports, which in turn is derived from the conjunction of
+  the invariants that describe behaviors of the component's
+  subsystems.
+\end{enumerate}
 
-First, both of them, should be in sync with VHDL-AMS language and
-reference manual which tries to define VHDL-AMS as well as possible in
-a descriptive way, explaining the semantics in English.
-
-Second, specification derived from the notation should be easy to
-attach\footnote{Meaning?} to the rest of the system
-architecture\xspace(driver, microprocessor emulator, test-benches
-e.t.c.).
-
-Third, methodology must aid efforts on creation of target system
-correctness criteria, such that are independent of implementation.  It
-means that verification of simulator model should take place on
-specification using reasoning techniques provided for algebraic
-formalism before coding starts.
-
-Fourth, a precise semantics should provides the basis for automated
-forward-backward engineering.
-
-Managing these requirements, and insuring that the platform definition
-(constraints) are preserved and verified through all levels of
-abstraction is where one becomes aware of gaps in already used design
-methodologies.  In our opinion, the BON-JML approach is in line with
-requirements given above.  The authors will illustrate in this article
-how BON can represent a high level analogue mapping notation and will
-further illustrate a refinement technique from VHDL-AMS to BON and
-JML.
-
-This article provides three main contributions.
-\begin{enumerate}
-\item It provides a solution how to formally emulate piece of hardware
-using BON-JML approach; up to date BON and JML have been used mostly
-in the pure object oriented software construction.
-\item It relates hardware description code and refinement technique.
-It means that if a satisfactory VHDL-AMS code exists, then it can be
-simulated from the BON (informal/formal) specification, and JML
-implementation by an appropriate sequence of refinements.
-\item It introduces \inv as a general concept of modeling hardware
-components behavior and their interactions.  To be more precisely, the
-behavioral specification of the \STS may be represented as a \inv on
-the signals at the circuit's ports and may be obtained by taking the
-conjunction of the \emph{invariants} that describe behaviors of the
-individual components separately.
-
-\end{enumerate} The rest of this article is organized as follows.  In
-\autoref{sec:vhdlams} we have presented an overview of VHDL-AMS
-language.  \autoref{sec:choiceobon} defines BON principles and
-explains the reasons why we have decided to use BON method.  In
-\autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
-The central part of the our article is \autoref{sec:refVB}, in which
-we will give practical aspect of refinement, applied on first order
-passive RC low-pass filter, in order to produce formal specification
-and software simulation. In this section we explore this approach in
-detail.  In \autoref{sec:conclusion} the conclusions are given, and
-directions of future studies are outlined. Finally, a comprehensive
-list of the mapping between BON and VHDL-AMS can be found in the
-appendix.
+The rest of this article is organized as follows.  In
+\autoref{sec:vhdlams} we present an overview of VHDL-AMS language.
+\autoref{sec:choiceobon} defines BON principles and explains the
+reasons why we have used the BON method.  In \autoref{sec:mappingVB}
+the relation between VHDL-AMS and BON is summarized.  A comprehensive
+description of the mapping between BON and VHDL-AMS is found in a
+separate technical report.  This refinement is applied to a small
+example in \autoref{sec:refVB}, in which we abstract from a passive RC
+low-pass filter in order to produce formal specification and software
+simulation.  In \autoref{sec:conclusion} we reflect on this work and
+point to some open problems and research opportunities.
 
 %=====================================================================
+\section{VHDL-AMS}
+\label{sec:vhdlams}
 
-\section{VHDL-AMS}
-\label{sec:vhdlams} IEEE VHDL-AMS 1076.1 is the industry standard
-mixed-signal high-level description language for electronic and
-multi-domain systems. VHDL-AMS, as an superset of VHDL, supports
-modeling and simulation of analog and mixed-signal systems. The main
-contribution of VHDL-AMS is ability to describe the behavior of
-complex system trought differential and algebraic equations
+The VHDL-AMS language, defined by IEEE VHDL-AMS 1076.1, is the
+industry standard mixed-signal high-level description language for
+electronic and multi-domain systems.  VHDL-AMS, as an superset of
+VHDL, supports modeling and simulation of \emph{analog} and
+mixed-signal systems.
+
+\subsection{Key Qualities of VHDL-AM}
+\label{subsec:key-qualities-vhdl}
+
+The main contribution of VHDL-AMS is ability to describe the behavior
+of a complex system through differential and algebraic equations
 (DAEs).  For this purpose, VHDL-AMS introduces the following
-capabilities and therms:
-
+notions:
 \begin{itemize}
-\item \emph{Terminals} represents phisical connection points or
-circuit nodes of a network connectivity. They are declared to be of
-various \emph{natures}, which represent different energy domains of a
-system.  Terminals hold no values; their only role is to facilitate
+\item \emph{Terminals} represents physical connection points or
+  circuit nodes of a network connectivity.  Each terminal has a
+  \emph{nature}, which represents different energy domains of a
+  system.  Terminals hold no values---their only role is to facilitate
   the formation of conservative equation sets that, in turn, constrain
   the associated \emph{branch qualities}.
-\item The \emph{quantities} denote a waveform or a time series of
-values.  The behavior of the \emph{quantities} is given by the
-\emph{simultaneous statements}. They take values as a result of
-solving the set of simultaneous ordinary differential and algebraic
-equations (DAEs). To simplify, quantities can be understood as analog
-unknowns of (DAEs) that specify analog behaviour. There are three
-kinds of quantity in VHDL-AMS: free, branch and source quantities. A
-\emph{free quantity} can be used in signal-flow modeling. A
-\emph{branch quantity} is used to model conservative energy systems. A
+
+\item \emph{Quantities} denote a waveform or a time series of values.
+  The behavior of the \emph{quantities} is given through the use of
+  \emph{simultaneous statements}.  Quantities take values as a result
+  of solving a set of simultaneous ordinary differential and algebraic
+  equations (DAEs).  \notej{We need a concrete example here.  This
+    does not make much sense in isolation.  Perhaps this detail should
+    just be moved or replicated later.  -JRK}
+
+  To simplify, quantities are understood as analog unknowns of (DAEs)
+  that specify analog behaviour.  There are three kinds of quantities
+  in VHDL-AMS: free, branch, and source quantities.  A \emph{free
+    quantity} is used in signal-flow modeling.  A \emph{branch
+    quantity} is used to model conservative energy systems.  A
   \emph{source quantity} is used for frequency and noise modeling.
-\item The \emph{simultaneous statements} creates one or more
-characteristic equations that are solved by the dedicated simulation
-kernel a.k.a.  "analog solver", to model continuously valued behavior.
-\item The \emph{procedural statements} or \emph{procedural} express
-analog behaviour using sequential statements. Actually procedural
-describes analog behaviour by computing new values for quantities from
-the current value of quantities.
-\item The \emph{break statement} is an construct which allows to
-initialize the DAE system at the beginning of a transient simulation
-and to mark discontinuities during such a transient simulation. For
-any concurrent break statement there is an associated process.  A
-break statement allows restart of the analog kernel, than overrides
+
+\item \emph{Simultaneous statements} define one or more characteristic
+  equations that are solved by the dedicated simulation kernel, a.k.a.
+  an "analog solver," to model continuously valued behavior.
+
+\item \emph{Procedural statements} express analog behaviour using
+  sequential statements. Actually, procedural statements describe
+  analog behaviour by computing new values for quantities from the
+  current value of quantities.
+
+\item The \emph{break statement} is an construct used to initialize
+  the DAE system at the beginning of a transient simulation, and to
+  mark discontinuities during such a transient simulation.  For any
+  concurrent break statement there is an associated process.  A break
+  statement allows restart of the analog kernel, than overrides
   quantity values and executes concurrently with the analogue model.
-
+  \notje{I have no idea what this last sentence means. -JRK}
 \end{itemize}
 
-\section{Choice of BON as the modeling tool}
-\label{sec:choiceobon} %Brief description and motivation.  The
-Business Object Notation (BON) is an analysis and design notation for
-object-oriented systems which brings \emph{seamlessness},\xspace
-\emph{reversibility} and software \emph{contracting}.  Let us define
-these principles precisely and point out why the principles are
-important in VHDL-AMS simulation.
+\subsection{A Running Example: A Low-Pass Filter}
+\label{subsec:lptransfunc}
 
-\textbf{Seamlessness and Reversibility} \xspace \xspace Seamlessness
-is the use of a continuous process throughout the software
-life-cycle. Reversibility is the support for both forward and backward
+\input{lowpassRC}
+
+A simple low-pass filter is a passive RC network shown in
+\autoref{fig:lowpassfig}.  According to the equation given above (see
+\autoref{fig:transfun}), we can produce VHDL-AMS \emph{simultaneous
+  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
+transformation looks as follows:
+\begin{equation}
+v_{in} = v_{out} + \tau_psv_{out} 
+   \begin{array}[b]{c}                                                                                                                                    
+      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
+   \end{array} 
+v_{out} + \tau_p\cfrac{dv_{out}}{dt}
+\begin{array}[b]{c}                                                                                                                                    
+      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
+   \end{array}
+  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
+\end{equation}
+
+This last equation is interpreted as follows.  The input voltage
+($v_{in}$) equals the sum of output voltage ($v_{out}$) and
+multiplication of the time constant ($\tau_p$) of the pole and the
+first derivative of the capacitor's voltage ($v_{out}{'dot}$) over
+time.
+
+Now we discretize previous equation.  Let us represent $v_{in}$ by a
+series of samples $v_{in(1..n)}$.  At the same points in time, we
+represent $v_{out}$ by a series of samples $v_{out(1..n)}$.  We assume
+that samples are taken at points in time incrementing by $\triangle$t.
+
+Now we have equivalent equations:
+\begin{equation}
+  v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}}
+  {\triangle t});
+\end{equation}
+
+\begin{equation}
+  v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
+  v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                \end{equation}
+
+\begin{equation}
+  v_{out(i)} 
+ \begin{array}[b]{c}
+   {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}
+   \\=
+ \end{array}
+ \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                           \end{equation}
+
+The last equation gives us a way to calculate the output samples in
+terms of the input samples and the preceding output.
+
+But how to represent these unknowns in BON\notej{This whole last bit
+  should be moved to after we introduce BON and the VHDL-AMS
+  metamodel.}?  We introduce the \lstinlinen|BRANCH_QUANTITY| objects
+like wrappers around unknowns, with features as follows:
+\begin{itemize}
+\item \emph{across} which represent current:$i_{out}$ 
+\item \emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
+\item \emph{terminal sequences} which describes input/output
+  terminal(s). 
+\end{itemize}
+
+%===================================================================== 
+\section{The Business Object Notation}
+\label{sec:choiceobon}
+
+The Business Object Notation (BON) is an analysis and design notation
+for object-oriented systems which supports \emph{seamlessness},
+\emph{reversibility}, and software \emph{contracting}.  Let us define
+these principles precisely and point out why they are important in
+VHDL-AMS simulation.
+
+\paragraph{Seamlessness and Reversibility} \emph{Seamlessness} is the
+use of a continuous process throughout the software life-cycle.
+\emph{Reversibility} is the support for both forward and backward
 development process: from analysis to design and implementation, and
-back.  In VHDL-AMS, modelling represents an continuous
-top-down-bottom-up process. Top-down approach serves for design while
-bottom-up serves for verification. The possibility of seamlessly
-translating changes made during a certain development phase back into
-earlier phases, facilitates a stable and predictable environment and
-ensures the various models consistent with the code. Keeping in sync
-various models, documentation and specification from forward progress
-to backward and redesign, preclude that design and implementation
-diverge to an inconsistent state.
+back.
+
+In VHDL-AMS, modeling represents an continuous top-down/bottom-up
+process.  A top-down approach supports design, while bottom-up
+supports verification.  The possibility of seamlessly translating
+changes made during a certain development phase back into earlier
+phases facilitates a stable and predictable environment and ensures
+the various models consistent with the code.  Keeping in sync various
+models, documentation and specification from forward progress to
+backward and redesign preclude that design and implementation diverge
+to an inconsistent state.
 
@@
 at different levels of abstraction.  For instance, one may wish to
-analyze a "lowpass-filter" at the level of individual components (a
-resistor and capacitor) and their behaviours, or at the more abstract
-level through general behaviour (\inv).  Therefore, capturing
-similarities between the differing levels of abstractions, making it
-possible to obtain a continuous process that ensures a direct mapping
-between a formal specification and its software solution.
+analyze a lowpass-filter at the level of individual components (a
+resistor and capacitor) and their behaviours or, at the more abstract
+level, through general behaviour expressed via an invariant.
+Therefore, capturing similarities between the differing levels of
+abstractions makes it possible to obtain a continuous process that
+ensures a direct mapping between a formal specification and its
+software solution.
 
-\textbf{Contracting} \xspace \xspace Contract specification describe
-system components interaction, based on mutual obligations and
-benefits, and, for each software element, associated semantic
-properties. BON use a full assertion mechanism, allowing analysts to
-specify both:\xspace (1) the \emph{structure} of a system and
-(2)\xspace \emph{semantics} ( constraints and invariants).  The common
-way to specify a program statement is to give a \emph{precondition}
-and \emph{postcondition} on the state that the implementation must
-satisfy, assuming that only certain program variables are changed.  In
-VHDL-AMS any predefined (implicate) attribute has defined \emph{pre}
-(\emph{post}) \emph{condition} trought input (output) parameters and
-theirs constraints. Furthermore, any subtype (subnature) represents
-restricted set of values from a base type (nature).  The condition
-that determines which values are in the subtype (subnature) may be
-contracted by specifying characteristic predicate which represents
-\inv.
+\paragraph{Contracting} Contract specifications describe system
+components and their interactions based on mutual obligations, and,
+for each software element, its associated semantic properties.  BON
+use an assertion mechanism, supporting analysts' need to specify
+both:~(1) the \emph{structure} of a system, and (2)~the
+\emph{semantics} of a system via a set of constraints and invariants.
+
+The common way to specify a program statement is to give a
+\emph{precondition} and \emph{postcondition} on the state that the
+implementation must satisfy, assuming that only certain program
+variables are changed.  In VHDL-AMS, any predefined (implicate)
+attribute has defined \emph{pre} (\emph{post}) \emph{condition}
+through input (output) parameters and their constraints.  Furthermore,
+any subtype (subnature) represents restricted set of values from a
+base type (nature).  The condition that determines which values are in
+the subtype (subnature) may be contracted by specifying a
+characteristic predicate which represents an invariant.
 
 Let us next look how \emph{simultaneous statements} are handled in
-contract design.  The \emph{simultaneous statement} represents
+contract-based design.  The \emph{simultaneous statement} represents
 analogue constraint of continuous-time waveform.  We will assume that
-the constraint of waveform represents \inv. This premise deserve an
-informal explanation.
+the constraint of a waveform is represented by an invariant. This
+premise deserves an informal explanation.
 
-Briefly stated, analog kernel calculate quantities such that the
+Briefly stated, the analog kernel calculates quantities such that the
 constraints specified by the simultaneous statements are satisfied.
@@
 statements should be equal to the number of unknowns
-(\emph{quantities}).  For instance, when we reasoning about the
-conservation of energy law, the \emph{simultaneous statements}
-constrain the values of \emph{branch quantities} in a model.  In other
-words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across
-quatities =0}) around any closed path must be equal to zero and the
-sum of \through quantities (\texttt{$\sum_{node}$ through quatities
-=0}) at any node (terminal) must be equal to zero. To be more precise,
-the results of summations above remains zero when a particular
-transformation is applied to quantities.  In general, they are \inv.
+(\emph{quantities}).
+
+For instance, when we reasoning about the conservation of energy law,
+the \emph{simultaneous statements} constrain the values of
+\emph{branch quantities} in a model.  In other words, the sum of
+\arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0}) around
+any closed path must be equal to zero and the sum of \through
+quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
+(terminal) must be equal to zero.  To be more precise, the results of
+summations above remains zero when a particular transformation is
+applied to quantities.  In general, they are invariant.
 
@@
 
-Therefore, on the basis of the above, BON is the obvious choice for
-the \STS.  Throughput this article we will make this more clearly.
+%=====================================================================
 \section{Mapping between VHDL-AMS and BON}
-\label{sec:mappingVB} In this section, we briefly describe the mapping
-between VHDL-AMS and BON notation which is discussed in the points
-below.
+\label{sec:mappingVB}
+
+In this section, we briefly describe the mapping between VHDL-AMS and
+BON notation which is discussed in the points below.
 
@@
 
-\subsection{Low-Pass filter transfer function }
-\label{subsec:lptransfunc} 
-\input{lowpassRC}
-
-The most simple low-pass filter is passive RC network shown in
-\autoref{fig:lowpassfig}.  According to equation given above (see
-\autoref{fig:transfun}) we can produce VHDL-AMS \emph{simultaneous
-  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
-transformation looks as follows:
-\begin{equation}
-v_{in} = v_{out} + \tau_psv_{out} 
-   \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
-   \end{array} 
-v_{out} + \tau_p\cfrac{dv_{out}}{dt}
-\begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
-   \end{array}
-  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
-\end{equation}
-
-The last equation can be described as follows.  The input
-voltage\xspace($v_{in}$)\xspace equals summation of output
-voltage\xspace($v_{out}$)\xspace and multiplication of the time
-constant\xspace($\tau_p$)\xspace of the pole and the first derivative
-of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace over time.
-
-Now we can discretize previous equation. Let us represent $v_{in}$ by
-a series of samples $v_{in(1..n)}$. At the same points in time, we can
-represent $v_{out}$ by series of sample $v_{out(1..n)}$. We can assume
-that samples are taken at points in time incrementing by
-$\triangle$t.
-
-Now we have equivalent equations:
-\begin{equation}
-v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                          \end{equation}
-
-\begin{equation}
-v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
- v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                                                                  
-\end{equation}
-
-\begin{equation}
-v_{out(i)} 
- \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                              \end{array}
- \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                                                                  
-\end{equation}
-
-The last equation gives us a way to calculate the output samples in
-terms of the input samples and the preceding output.  But how to
-represent these unknowns in BON ?  We introduce the
-\lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
-with features as follows:
-\begin{itemize}
-\item
-\emph{across} which represent current:$i_{out}$ 
-\item 
-\emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
-\item 
-\emph{terminal sequences} which describes input/output terminal(s). 
-\end{itemize}
-Now we are in presence of BON notation and we may start with the
-refinement of the VHDL-AMS code.
-
 \subsection{Refine example }
@@
 \section{Conclusion}
-\label{sec:conclusion} The BON notation has originally been used for
-description of software models. However, after certain elaboration it
-can be used for description of hardware components. Practical
-utilization of the approach in the project on mapping between VHDL-AMS
-and BON using refinement technique, has demonstrated through low-pass
-filter example and shown that simulation of component requires
-relatively little and inexpensive efforts. Moreover, our article shown
-that the development of the specification produced by the BON and
-translated into JML can be applied as refinement procedure: from
-VHDL-AMS to informal BON, next, from informal to more formal BON and
-finally, to JML specification.  The BON support general software
-development and should be adapted for concrete applications and
-language as VHDL-AMS. For this purpose we have developed various
-VHDL-AMS related types, operators and relations and we have defined
-appropriate specifications for all of them. However, the behavior of
+\label{sec:conclusion}
+
+The BON notation has originally been used for description of software
+models. However, after certain elaboration it can be used for
+description of hardware components. Practical utilization of the
+approach in the project on mapping between VHDL-AMS and BON using
+refinement technique, has demonstrated through low-pass filter example
+and shown that simulation of component requires relatively little and
+inexpensive efforts. Moreover, our article shown that the development
+of the specification produced by the BON and translated into JML can
+be applied as refinement procedure: from VHDL-AMS to informal BON,
+next, from informal to more formal BON and finally, to JML
+specification.  The BON support general software development and
+should be adapted for concrete applications and language as
+VHDL-AMS. For this purpose we have developed various VHDL-AMS related
+types, operators and relations and we have defined appropriate
+specifications for all of them. However, the behavior of
 \texttt{break} statement and concurrency, for example, have not been

### 2010-01-17 papers/BON_VHDL/paper.tex
Cleaning up some equations and LaTeX.

@@
 
-%======================================================================
+%=============================================
 %\thispagestyle{empty}
@@
   specification languages to describe system abstractions.  These
-abstractions, which typically capture static and dynamic architectural
-aspects, are meant to be more approachable, comprehensible, and useful
-than witnessing the gory details of the systems they describe.  In the
-hardware world, systems are described with hardware description
-languages like VHDL.  These languages exist at a similar level of
-abstraction as traditional software programming languages, like Java.
-But whence are the high-level hardware description languages?
-Moreover, if one is doing hardware/software co-design, what
-architecture description languages are usable and useful to both
-facets of the problem domain?  In this paper, we describe a
-non-trivial co-design effort, called SenseTile, and how requirements
-on its design and verification led us to define a formal refinement of
-VHDL-AMS to the EBON high-level specification language.  This
-refinement enables us to abstract away, in a seamless, reversible, and
-contract-based fashion, from the details of a particular hardware
-system to an neutral architecture description which, in turn, can be
-refined to an executable Java-based software simulator.  This pair of
-refinements permits full round-trip hardware/software co-design with
-verification, as the Java-based system is annotated with formal
-specifications using the Java Modeling Language.
+  abstractions, which typically capture static and dynamic
+  architectural aspects, are meant to be more approachable,
+  comprehensible, and useful than witnessing the gory details of the
+  systems they describe.  In the hardware world, systems are described
+  with hardware description languages like VHDL.  These languages
+  exist at a similar level of abstraction as traditional software
+  programming languages, like Java.  But whence are the high-level
+  hardware description languages?  Moreover, if one is doing
+  hardware/software co-design, what architecture description languages
+  are usable and useful to both facets of the problem domain?  In this
+  paper, we describe a non-trivial co-design effort, called SenseTile,
+  and how requirements on its design and verification led us to define
+  a formal refinement of VHDL-AMS to the EBON high-level specification
+  language.  This refinement enables us to abstract away, in a
+  seamless, reversible, and contract-based fashion, from the details
+  of a particular hardware system to an neutral architecture
+  description which, in turn, can be refined to an executable
+  Java-based software simulator.  This pair of refinements permits
+  full round-trip hardware/software co-design with verification, as
+  the Java-based system is annotated with formal specifications using
+  the Java Modeling Language.
 
@@
 
-%======================================================================
+%=============================================
 \section{Introduction}
@@
 
-%=====================================================================
+%=============================================
 \section{VHDL-AMS}
@@
 The VHDL-AMS language, defined by IEEE VHDL-AMS 1076.1, is the
-industry standard mixed-signal high-level description language for
+industry-standard mixed-signal high-level description language for
 electronic and multi-domain systems.  VHDL-AMS, as an superset of
 VHDL, supports modeling and simulation of \emph{analog} and
-mixed-signal systems.
+mixed-signal systems.  In short, one uses VHDL-AMS to describe
+hardware components ranging in complexity from filters to CPUs.
 
-\subsection{Key Qualities of VHDL-AM}
+\subsection{Key Qualities of VHDL-AMS}
 \label{subsec:key-qualities-vhdl}
@@
 
+A simple example will highlight some of the key aspects of VHDL-AMS
+language and will be used to motivate and explain our refinement.
+
 \subsection{A Running Example: A Low-Pass Filter}
@@
 \autoref{fig:transfun}), we can produce VHDL-AMS \emph{simultaneous
-  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
-transformation looks as follows:
+  statement} using \texttt{$s = d/dt = 'dot$} relation\notej{I don't
+  know what this means}.
+
+The transformation looks like the following:
 \begin{equation}
   v_{in} = v_{out} + \tau_psv_{out} 
-   \begin{array}[b]{c}                                                                                                                                    
-      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
-   \end{array} 
+  \begin{array}[b]{c}                                                                                                                {\scriptscriptstyle s = d/dt}  \\=                                                                                 \end{array}
   v_{out} + \tau_p\cfrac{dv_{out}}{dt}
@@
   v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}}
-  {\triangle t});
+  {\triangle t})
 \end{equation}
@@
   v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
-  v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                \end{equation}
+  v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t})                                                                \end{equation}
 
@@
  \end{array}
- \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                           \end{equation}
+ \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)}                                                                           \end{equation}
 
@@
 
-%===================================================================== 
+%=============================================
 \section{The Business Object Notation}
@@
 
-%=====================================================================
+%=============================================
 \section{Mapping between VHDL-AMS and BON}
@@
 
+%=============================================  
 \section{Refinement from VHDL-AMS to BON }
@@
 
-%=====================================================================
+%=============================================
 \section{Conclusion}
@@
 
-\section {Acknowledgments} The research upon which this article is
+\paragraph{Acknowledgments} The research upon which this article is
 based was caried out with sponsorship from Science Foundation Ireland

### 2010-01-27 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
minor updates
static diagram for data provider added

@@
 
+This description is executed on the processing unit as part of a webservice to generate data.
+SOS. Implemation added to allow the description to be used to generate observations. Processing of data.
+
 A  SensorML to BON Mapping was also to be explored .
@@
 \section{SensorML Overview}
-The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standerised “Sensor Web”.
+The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised “Sensor Web”.
 It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.
@@
 The processes can be connected together in chains using SensorML ProcessChains or Systems.
+
+including measurement by a sensor system, as well as post-measurement
+processing.
 
@@
  \begin{figure}
-\includegraphics{SensorML_SenseTile_System_design_tuple_pa3.png}
-\caption{Block diagram of SenseTile System process}\label{fig:SensorML_SenseTile_design.png}
+\scalebox{0.65}{\includegraphics{SenseTileSystem.pdf}}
+\caption{SenseTile System Overview}\label{fig:SenseTileSystem.pdf}
+\end{figure}
+
+ \begin{figure}
+\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_2_pa1.pdf}}
+\caption{SenseTile System Components}\label{fig:SensorML_SenseTile_System_2_pa1.pdf}
 \end{figure}
@@
 \subsection{Overview}
+
+Web Service - SOS -standard access to sensor observations
+
+SOS diagram
+
+DataProvider
+
+DataProducer
+
+Deployment Scenarios - diagram
+
+Flexible Division - several dataproducers can register with one DataProvider.
+
+Seperate processes rmi.
  \begin{figure}
 \scalebox{0.65}{\includegraphics{SystemOverview.png}}
-\caption{SenseTile System Overview}\label{fig:SystemOverview.png}
+\caption{Overview}\label{fig:SystemOverview.png}
 \end{figure}
@@
 \subsection {DataProvider}
-
+ \begin{figure}
+\scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
+\caption{DataProvider Static Diagram}\label{fig:bon_static_diagam_provider.png}
+\end{figure}
 Keeps track of sensorboards that register

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
more updates. Intro etc

@@
 
-The aim of this thesis project was to prototype a Web Service for accessing sensor observations from the SenseTile Sensor Board.
+The aim of this thesis project was to model the SenseTile Sensor Board using Sensor Model Language (SensorML)\cite{SensoMLref} .
+A prototype of a Web Service for accessing SenseTile Sensor Board sensor observations was to be developed using this SensorML description of the SenseTile Sensor Board.
+An evaluation of the suitability of SensorML for use as part of the SenseTile System was also to be performed.
+
 SenseTile is a sensor and processing package including motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
 It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
+The SenseTile System is made up of two units a Sensor Board and Processor Unit.
+The sensors are interfaced or hosted on the Sensor Board.
+The Processor Unit interfaces to the Sensor Board allowing data from the Sensor Board to be processed futher as required.
 
-Descriptions of the SenseTile Sensor Board were to be developed using Sensor Model Language (SensorML)\cite{SensoMLref}.
-SensorML is a language that describes the process of sensor measurement by providing standard models and XML encoding for such processes.
-SensorML was be used as it is becoming a standard for describing sensor data processing as part of the SWE[ref].
+SensorML was developed to describe sensor systems and the processing of sensor observations.
+It provides standard models and XML encodings to describe the process of sensor measurement.
+SensorML is one of a suite of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
+The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
 
-This description is executed on the processing unit as part of a webservice to generate data.
-SOS. Implemation added to allow the description to be used to generate observations. Processing of data.
+The SenseTile Web Service allows access to the Sensor Board sensor observations and SensorML description.
+It would run on the SenseTile Processor Unit.
+The Web Service is based on another OGC SWE  specification, the Sensor Observation Service (SOS)[ref].
+This specification defines a standard Web interface for retrieving information about sensor systems and sensor observations.
+Sensor Board measurements that have been post processed and converted to real values or the raw Sensor Board data can be accessed using the Web Service.
+
+SensorML and SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref].
+This specification provides standard models and XML encoding for sensor observations.SWE Common - common data models and schema
 
@@
 
+\section{Sensor Networks and SWE}
 How was problem solved before.
 
-\subsection{Sensor Networks}
-\subsection{BON}
-\subsection{Sensorml}
-reference chapter
-\subsection{WebService}
-\subsection{JIBX}
-
-\chapter{SensorML}
-\section{SensorML Overview}
+\section{SensorML}
 The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised “Sensor Web”.
@@
 The SensorML description of the SenseTile Sensor Board is shown in the following section.
+\section{BON}
+\section{SOS}
+\section{OGC Lib}
 
+\chapter{Design}
 \section{SensorML Description of SenseTile}
@@
 How to display this in thesis?
-\chapter{Design}
-
 \section{SenseTile Web Service}
@@
 Web Service - SOS -standard access to sensor observations
+The Sensor Observation Service (SOS) is an approved Open Geospatial Consortium standard.
+The standard defines a web service interface for the discovery and retrieval of real time or archived data produced by all kinds of sensors like mobile or stationary as well as in-situ or remote sensors.
 
+The sensor data can be either observations or descriptions of the producing sensors containing metadata like calibration information, positions, etc.
+Observations are returned encoded as OM Observations, while information about sensors are returned encoded in SensorML or TML.
+Used in conjunction with these OGC standards and other OGC specifications, the SOS provides a broad range of interoperable capability for discovering, binding to and interrogating individual sensors, sensor platforms, or networked constellations of sensors in real-time, archived or simulated environments.
 SOS diagram
@@
 \bibitem{SensoMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
+\bibitem{SOSref}Open Geospatial Consortium Inc., OpenGIS  Implementation Specification, 2007
+\bibitem{OMref}Open Geospatial Consortium Inc., OpenGIS  Implementation Specification, 2007
+\bibitem{SWEArchref}Open Geospatial Consortium Inc., , 2007
+\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., , 2007
+
 \bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebServiceClient/src/ie/ucd/sensetile/sos/dataprovider/service/SenseTileServiceClient.java
latest

@@ -16,18 +16,21 @@
  * specific language governing permissions and limitations
  * under the License.
  */
-package ie.ucd.sensetile.dataprovider.service;
+package ie.ucd.sensetile.sos.dataprovider.service;
 
 
-import ie.ucd.sensetile.dataprovider.service.SenseTileServiceStub.GetSensorObsResponse;
-
+import ie.ucd.sensetile.sos.dataprovider.service.SenseTileServiceStub.GetSensorObsResponse;
+import ie.ucd.sensetile.sos.dataprovider.service.SenseTileServiceStub.GetCapabilitiesResponse;
 public class SenseTileServiceClient {
  
     public static void main(String[] args) throws Exception {
  
         SenseTileServiceStub stub = new SenseTileServiceStub();
-        
+        GetCapabilitiesResponse capabilities = stub.getCapabilities();
+		
         //Invoke the service
+        System.out.println("Response : " + capabilities.get_return() );
+
         GetSensorObsResponse response = stub.getSensorObs();
  
         System.out.println("Response : " + response.get_return().getObservation());

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebService/src/ie/ucd/sensetile/sos/dataprovider/PreFixMapper.java
latest

@@ -1,0 +1,45 @@
+package ie.ucd.sensetile.sos.dataprovider;
+
+import com.sun.xml.bind.marshaller.NamespacePrefixMapper;
+
+public class PreFixMapper extends NamespacePrefixMapper {
+    @Override
+    public String getPreferredPrefix(String namespace, String suggestion, boolean requirePrefix) {
+        String prefix = "";
+        if (namespace.contains("http://www.opengis.net/gml")){
+            prefix = "gml";
+        }
+        else if (namespace.contains("http://relaxng.org/ns/structure")){
+            prefix = "rng";
+        }       
+        else if (namespace.contains("http://www.opengis.net/sensorML")){
+            prefix = "sml";
+        }
+        else if (namespace.contains("http://www.opengis.net/swe")){
+            prefix = "swe";
+        }
+        else if (namespace.contains("http://www.w3.org/1999/xlink")){
+            prefix = "xlink";
+        }
+        else if (namespace.contains("http://www.opengis.net/sos/1.0")){
+            prefix = "sos";
+        }
+        else if (namespace.contains("http://www.opengis.net/tml")){
+            prefix = "tml";
+        }
+        else if (namespace.contains("http://www.w3.org/2001/XMLSchema-instance")){
+            prefix = "xsi";
+        }
+        else if (namespace.contains("http://www.opengis.net/ows")){
+            prefix = "ows";
+        }
+        else if (namespace.contains("http://www.opengis.net/ogc")){
+            prefix = "ogc";
+        }
+        else if (namespace.contains("http://www.opengis.net/om")){
+            prefix = "om";
+        }
+
+        return prefix;
+    }
+}

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebService/src/ie/ucd/sensetile/sos/dataprovider/SOSCapability.java

@@ -1,0 +1,136 @@
+package ie.ucd.sensetile.sos.dataprovider;
+
+import java.io.ByteArrayOutputStream;
+import java.util.List;
+
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Marshaller;
+import javax.xml.bind.PropertyException;
+import javax.xml.namespace.QName;
+
+import net.opengis.gml.v_3_1_1.CodeType;
+import net.opengis.gml.v_3_1_1.ReferenceType;
+import net.opengis.gml.v_3_1_1.StringOrRefType;
+import net.opengis.ows._1.LanguageStringType;
+import net.opengis.ows._1.ObjectFactory;
+import net.opengis.ows._1.ServiceIdentification;
+import net.opengis.ows._1.ServiceProvider;
+import net.opengis.sensorml._1_0.SystemType;
+import net.opengis.sos._1.Capabilities;
+import net.opengis.sos._1.Contents;
+import net.opengis.sos._1.ObservationOfferingType;
+import net.opengis.swe._1_0.PhenomenonPropertyType;
+import net.opengis.swe._1_0.PhenomenonType;
+
+public class SOSCapability {
+    
+    private Capabilities capabilities= null;
+    private String offeringXML = null;
+    static final String OFFERID ="offering-";
+    static final String OFFERURN ="urn:MyOrg:offering:";
+
+    static long offerid = 0;
+    
+    SOSCapability(){
+        createCapability();
+    }
+    
+    
+    public String getCapabilitiesXml() {
+        return offeringXML;
+    }
+    private void createCapability() {
+        net.opengis.sos._1.ObjectFactory objFactorySOS = new net.opengis.sos._1.ObjectFactory();
+        capabilities = objFactorySOS.createCapabilities();
+        net.opengis.ows._1.ObjectFactory objFactoryOWS = new net.opengis.ows._1.ObjectFactory();
+        ServiceIdentification serviceId = objFactoryOWS.createServiceIdentification();
+        
+        LanguageStringType title = new LanguageStringType();
+        title.setValue("SenseTile SOS");
+        serviceId.getTitle().add(title);
+        
+        capabilities.setServiceIdentification(serviceId);
+        
+        ServiceProvider serviceProvider = objFactoryOWS.createServiceProvider();
+        Contents contents =  objFactorySOS.createContents();
+        capabilities.setContents(contents);
[…]

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebService/src/ie/ucd/sensetile/sos/dataprovider/Sensor.java

@@ -1,41 +1,174 @@
 package ie.ucd.sensetile.sos.dataprovider;
 
+import java.io.InputStreamReader;
+import java.io.StringBufferInputStream;
+import java.util.ArrayList;
+import java.util.List;
+
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Unmarshaller;
+
+import net.opengis.sensorml._1_0.AbstractProcessType;
+import net.opengis.sensorml._1_0.SensorML;
+import net.opengis.sensorml._1_0.SystemType;
+import net.opengis.sensorml._1_0.SensorML.Member;
+import net.opengis.sos._1.ObservationOfferingType;
+
+
+
 public class Sensor implements SensorIF{
 
-    @Override
-    public void addObservation(Observation obs) {
-        // TODO Auto-generated method stub
+    private String sensorURN = null;
+    private String xmldescription = null;
+    private String version = null;
+    private String observationTemplate = null;
+    private String service = null;
+    static final String JAXB_SENSORML_PACKAGE = "net.opengis.sensorml._1_0";
+    private ObservationOfferingType observationOffering = null;
+    private SensorML sensorml = null;
+
+
+    List <Observation> observations = new ArrayList <Observation>();
+    private SensorML description;
+        
+    public Sensor(String sensorURN, String description, String version,
+            String observationTemplate, String service) {
+        super();
+        this.sensorURN = sensorURN;
+        this.xmldescription = description;
+        this.version = version;
+        this.observationTemplate = observationTemplate;
+        this.service = service;
+        this.sensorml = readDescription(xmldescription);
         
     }
 
+
+    Sensor(String sensorURN) {
+        this.sensorURN = sensorURN;
+    }
+    
+
+    public void setxmlDescription(String description) {
+        this.xmldescription = description;
+    }
+
+    public void setVersion(String version) {
+        this.version = version;
+    }
+
+    public void setObservationTemplate(String observationTemplate) {
+        this.observationTemplate = observationTemplate;
+    }
+
+    public void setService(String service) {
+        this.service = service;
+    }
+    
+    public String getSensorURN() {
+        return sensorURN;
+    }
+
+    public void setSensorURN(String sensorURN) {
+        this.sensorURN = sensorURN;
+    }
+
+    @Override
+    public void addObservation(Observation obs) {
+        this.observations.add(obs);        
+    }
+    
     @Override
     public Observation getObservation() {
[…]

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebService/src/ie/ucd/sensetile/sos/dataprovider/SensorIF.java

@@ -16,11 +16,11 @@ public interface SensorIF {
     /* get version */
     String getVersion();
 
-    /* get SensorDescription */
-    String getDescription();
 
     /* get ObservationTemplate */
-    String ObservationTemplate();
+    String getObservationTemplate();
+
+    String getXMLDescription();
 
 
 }

### 2010-02-17 thesis/SensorMLOnSenseTile/trunk/WebService/src/ie/ucd/sensetile/sos/dataprovider/SensorObservations.java

@@ -23,17 +23,30 @@ import java.rmi.RemoteException;
 import java.rmi.registry.LocateRegistry;
 import java.rmi.registry.Registry;
 import java.rmi.server.UnicastRemoteObject;
+import java.util.ArrayList;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Map;
 
 import ie.ucd.sensetile.webservice.sos.SensorObservationIF;
+
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Marshaller;
+import java.io.ByteArrayOutputStream;
+
+import net.opengis.sos._1.Contents;
+import net.opengis.sos._1.ObjectFactory;
 
 public class SensorObservations
     implements SensorObservationIF {
     private int observation;
     private String sensorIdUrnRoot ="urn:ie.ucd.sensetile:sensor:";
-    private String sensor1 = null;
     private Long sensorId = 0l;
     private String serverObjectName ="rmi://localhost/SenseTileService";
-    private int counter =0;
+    private Map <String, SensorIF> sensors= new HashMap <String, SensorIF>();
+    
+    private SOSCapability sosCapability = null;
     
     public SensorObservations() throws RemoteException{
         //super();
@@ -77,9 +90,17 @@ public class SensorObservations
     public String RegisterSensor(String service, String version,
             String SensorDescription, String ObservationTemplate)
             throws RemoteException {
-        counter++;
-        System.out.println("Observations bean service = " + counter);       
-        return this.sensorIdUrnRoot;
+        String sensorurn = generateSensorURN();
+        Sensor sensor = new Sensor(sensorurn, SensorDescription, version,
+                ObservationTemplate,service);
+        this.sensors.put(sensorurn,sensor);
+        System.out.println("Observations bean service = " + sensorurn);
+        
+        if (sosCapability == null) {
+            sosCapability = new SOSCapability();
+            sosCapability.addOffering(sensor);
+        }
+        return sensorurn;
 
     }
     
@@ -87,5 +108,28 @@ public class SensorObservations
         String id = sensorIdUrnRoot+sensorId;
         sensorId++;
         return id;
+    }
+
+    public String getCapabilites() {
+
+       /* try {
+            JAXBContext jaxbContext = JAXBContext.newInstance("net.opengis.sos._1");
+            Marshaller marshaller = jaxbContext.createMarshaller();
[…]

### 2010-02-20 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates

@@
 
-SensorML was developed to describe sensor systems and the processing of sensor observations.
+SensorML is one of a number of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
+The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
+SensorML is used to describe sensor systems and the processing of sensor observations.
 It provides standard models and XML encodings to describe the process of sensor measurement.
-SensorML is one of a suite of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
-The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
+SensorML is described further in chapter x.
 
-The SenseTile Web Service allows access to the Sensor Board sensor observations and SensorML description.
-It would run on the SenseTile Processor Unit.
+The SenseTile Web Service prototype allows access to the Sensor Board sensor observations and it's SensorML description.
+It runs on the SenseTile Processor Unit.
 The Web Service is based on another OGC SWE  specification, the Sensor Observation Service (SOS)[ref].
@@
 
-SensorML and SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref].
-This specification provides standard models and XML encoding for sensor observations.SWE Common - common data models and schema
+There are othe  SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref]. This specification provides standard models and XML encoding for sensor observations.
 
@@
 
-\section{Sensor Networks and SWE}
-How was problem solved before.
+\section{Sensor Web Enablement}
+The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised “Sensor Web”.
+It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.They provide a framework of open standards for use with Web-connected sensors and sensor systems.
+There are seven main specifications currently:
+
+\begin{itemize}
+\item  SensorModel Language(SensorML) - models and schema for sensor systems and processes surrounding measurements
+\item  Observations \& Measurements (O\&M) - models and schema for packaging observation values
+\item  Transducer Markup Language (TML) - models and schema for multiplexed data from sensor systems
+\item  Sensor Observation Service (SOS) - standard web interface for accessing observations
+\item  Sensor Planning Service (SPS) - standard web interface for tasking sensor systems and model and requesting acquisitions
+\item   Sensor Alert Service (SAS) - standard web interface for publishing and subscribing to sensor alerts
+\item   Web Notification Service (WNS) - standard web interface for asynchronous notification
+\end{itemize}
+There is also the SWE Common that provides common data models and schemas for the above.
+
+Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand.
+The SWE framework seems to cover all needs for modeling sensors and observations.
+Support the use of Web Services is the main rational and fits the needs of a SenseTile Sensor Network.
+In this case the SOS and SensorML are core.
+
+Scable - can be  large systems. Discovery. Process data without having to define an implemenation as this
+can be provodied as part of SensorML description.
+
+Other ways EEML? Small not suitable for Web explotiation. Anything gained by it's simpler approach.
 
 \section{SensorML}
-The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised “Sensor Web”.
-It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.
-Sensor Modelling Language (SensorML) is one part of the SWE initiative and is used to describe sensor systems and the processing of observations from sensor systems.
+Sensor Modelling Language (SensorML) is used to describe sensor systems and the processing of observations from sensor systems.
 
 The purposes of SensorML are to:
-• Provide descriptions of sensors and sensor systems for inventory management
-• Provide sensor and process information in support of resource and observation discovery
-• Support the processing and analysis of the sensor observations
-• Support the geolocation of observed values (measured data)
-• Provide performance characteristics (e.g., accuracy, threshold, etc.)
-• Provide an explicit description of the process by which an observation was obtained (i.e., it’s lineage)
-• Provide an executable process chain for deriving new data products on demand (i.e., derivable observation)
-• Archive fundamental properties and assumptions regarding sensor systems
+\begin{itemize}
+\item Provide sensor and process information in support of resource and observation discovery
+\item Support the processing and analysis of the sensor observations
+\item Support the geolocation of observed values (measured data)
+\item Provide performance characteristics (e.g., accuracy, threshold, etc.)
+\item Provide an explicit description of the process by which an observation was obtained (i.e., it’s lineage)
+\item Provide an executable process chain for deriving new data products on demand (i.e., derivable observation)
+\item Archive fundamental properties and assumptions regarding sensor systems
+\end{itemize}
 
 SensorML provides a functional model of sensors and an XML encoding to describe sensors and their observations.
-Sensors are modelled as processes that convert real phenomena to data.
-Processes take inputs, process them and result in one or several outputs.
-They can also define relevant metadata.
+Sensors are modelled as processes that convert real phenomena to data. Processes take inputs, process them and result in one or several outputs.
+
+Sensor metadata is also supported by SensorML.
+
 The processes can be connected together in chains using SensorML ProcessChains or Systems.
@@
 When Modelling a sensor system in SensorML the main entities used are Systems and Components.
-A SensorML System is used to group sensors that are related spatially to one another.
+
+A SensorML System is used to group sensors that are related spatially to one another.  As it is a processchaing it is hierarchical and so Systems can contain Systems.
+
 A Component is an atomic process that generally converts a physical phenomen to a digital number.
-ProcessModels are another modelling that is used to describe a pure process that has no physicality
 
-Quantity, Count, Boolean, Category, and Time provide the basic primitives. Within SensorML, these serve as the basis for specifying all inputs, outputs, and parameters within a Process
+ProcessModels are another modelling that is used to describe a pure process that has no physicality. They point to an implementation or a description of one.
 
-DataRecord or DataArray.
+The data types used for inputs and out puts come another OGC sepc the SWE common.
+Examles are Quantity which is a decimal number, Count, Boolean, Category, and Time provide the basic primitives.
+Within SensorML,
 
-Detector model.
+Aggregation of primative data types are also provided DataRecord or DataArray.
 
-The SensorML description of the SenseTile Sensor Board is shown in the following section.
+A sensor systems would likely to be based mainly on Systems and
+Detector model. Reference tutorial and sensorml specification.
+
+The SensorML description of the SenseTile Sensor Board is desctibed in chapter .
+
+\section{SOS}
+A Sensor Observation Service provides an API for managing deployed sensors and
+retrieving sensor data and specifically “observation” data.
+
+SOS has three mandatory “core” operations: GetObservation, DescribeSensor, and
+GetCapabilities. The GetObservation operation provides access to sensor observations
+and measurement data via a spatio-temporal query that can be filtered by phenomena.
+The DescribeSensor operation retrieves detailed information about the sensors and
+processes generating those measurements. The GetCapabilities operation provides the
+means to access SOS service metadata. Several optional, non-mandatory operations have
+also been defined. There are two operations to support transactions, RegisterSensor and
+InsertObservation, and six enhanced operations, including GetResult,
+GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
+DescribeObservationType, and DescribeResultModel.
+
+The assumption is that sensor data producers are software entities running on resource
+constrained platforms that are attached to sensors – either physically or through a
+wireless connection of some kind. The producers have just enough processing power to
+generate simple XML insert documents that can be published to the transactional SOS.
+The transactional SOS is expected to be running on a capable server in a data center
+which can easily handle a large number of simultaneous connections from data
+producers.
+Data consumers would request data from the transactional SOS using operations from the
+core profile and would not need to communicate with the ultimate data producers. Figure
+12 shows sensor data producers in the operational context described above. The catalog
+still appears in the diagram but there is only one and it will often be co-located with the
+SOS service. The catalog would be used for discovery of the SOS instance if the data
+producers were capable enough to support a catalog client application. In some cases the
+sensor data publishers will need to be configured with the location of the SOS when they
+are deployed. The diagram is meant to represent a facility that is under the control of a
+single organization. Scalability would be achieved through aggregating a number of
+sensor data producers on the sensor side and by aggregating a number of SOS instances
+on the data center side for providing data to consumers.
+\section{OGC Lib}
+DOM based. Reference on google code. Refernce of how to build. Quicker to build prototype. SensorML aware.
+
 \section{BON}
-\section{SOS}
-\section{OGC Lib}
 
@@
 
-Processed verus raw data.
+Offerings Processed verus raw data.
 

### 2010-02-21 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates.

@@
 
-The SensorML description of the SenseTile Sensor Board is desctibed in chapter .
+The SensorML description of the SenseTile Sensor Board is described in chapter .
 
-\section{SOS}
-A Sensor Observation Service provides an API for managing deployed sensors and
-retrieving sensor data and specifically “observation” data.
+\section{Sensor Observation Service}
+The SenseTile Web Service is based on the concepts described in the Sensor Observation Service (SOS) specification.
+The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.
+Observations are encoded as OGC O\&M Observations.
+Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML with SensorML being used by the SenseTile Web Service.
 
-SOS has three mandatory “core” operations: GetObservation, DescribeSensor, and
-GetCapabilities. The GetObservation operation provides access to sensor observations
-and measurement data via a spatio-temporal query that can be filtered by phenomena.
-The DescribeSensor operation retrieves detailed information about the sensors and
-processes generating those measurements. The GetCapabilities operation provides the
-means to access SOS service metadata. Several optional, non-mandatory operations have
-also been defined. There are two operations to support transactions, RegisterSensor and
-InsertObservation, and six enhanced operations, including GetResult,
+There are there core operations that are provided by an SOS and used by SOS Clients to request sensor information:
+ \begin{itemize}
+\item GetObservation operation is used to obtain sensor observations and measurement data.
+\item GetCapabilities operation is used to retrieve SOS service metadata.
+\item DescribeSensor operation retrieves detailed information about the sensors .
+\end{itemize}
+
+Two other operations non RegisterSensor and InsertObservation are of interest to this thesis
+with the obvious functionality.
+These are part of the SOS transactional interface.
+They are used with SOS data producers which are software entities connected to the sensor and have enough processing power to generate XML descriptions of the sensor data.
+
+Six other enhanced operations, GetResult,
 GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
-DescribeObservationType, and DescribeResultModel.
+DescribeObservationType, and DescribeResultModel are not considered are not included in the SenseTileWeb Service.
 
-The assumption is that sensor data producers are software entities running on resource
-constrained platforms that are attached to sensors – either physically or through a
-wireless connection of some kind. The producers have just enough processing power to
-generate simple XML insert documents that can be published to the transactional SOS.
-The transactional SOS is expected to be running on a capable server in a data center
+These operations are encoded in XML and used in http post and get operations.
+The XML message needs to be parsed to identify the operations, parameters and results.
+There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP.
+I could not find this as part of current OGC SWE framework.
+
+para: The SOS is expected to be running on a  server that  handle large numbers of connections
+from sensor systems.
+
 which can easily handle a large number of simultaneous connections from data
 producers.
-Data consumers would request data from the transactional SOS using operations from the
-core profile and would not need to communicate with the ultimate data producers. Figure
-12 shows sensor data producers in the operational context described above. The catalog
-still appears in the diagram but there is only one and it will often be co-located with the
-SOS service. The catalog would be used for discovery of the SOS instance if the data
-producers were capable enough to support a catalog client application. In some cases the
-sensor data publishers will need to be configured with the location of the SOS when they
-are deployed. The diagram is meant to represent a facility that is under the control of a
-single organization. Scalability would be achieved through aggregating a number of
+
+ The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis.
+
+para: Scalability would be achieved through aggregating a number of
 sensor data producers on the sensor side and by aggregating a number of SOS instances
 on the data center side for providing data to consumers.
+
 \section{OGC Lib}
 DOM based. Reference on google code. Refernce of how to build. Quicker to build prototype. SensorML aware.
+This open source project aims at building a JAVA API and implementation of the SWE (Sensor Web Enablement) Common Data Model, that can be easily used to access and produce SWE Common Data, and convert in to/from existing formats such as NetCDF, HDF, GeoTIFF, JPEG2000, etc...
+It can also be used to quickly develop SWE services such as SOS, SPS and SAS which all rely on the SWE Common Data Model and thus to build a Sensor Network.
+The project was initiated by the VAST Team at the University of Alabama in Huntsville (UAH) and is now also maintained by Sensia Software, Spot Image, and Botts Innovative Research, Inc.. 
+http://code.google.com/p/swe-common-data-framework/
 
 \section{BON}
+\section{Trchnologies}
+JAXB
+JAVA
+AXIS 2.0
+SPRING
 

### 2010-03-03 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates

@@
 \usepackage{hyperref}
+\usepackage{listings}
 
@@
 
-Scable - can be  large systems. Discovery. Process data without having to define an implemenation as this
-can be provodied as part of SensorML description.
+Scalable - can be  large systems. Discovery. Process data without having to define an implementation as this
+can be provided as part of SensorML description.
 
-Other ways EEML? Small not suitable for Web explotiation. Anything gained by it's simpler approach.
+Other ways EEML? DAML?
 
-\section{SensorML}
+\section{Sensor Modeling Language}
 Sensor Modelling Language (SensorML) is used to describe sensor systems and the processing of observations from sensor systems.
@@
 
-When Modelling a sensor system in SensorML the main entities used are Systems and Components.
+When Modeling a sensor system in SensorML the main entities used are Systems and Components.
 
-A SensorML System is used to group sensors that are related spatially to one another.  As it is a processchaing it is hierarchical and so Systems can contain Systems.
+A SensorML System is used to group sensors that are related spatially to one another.  As it is a process chain it is hierarchical and so Systems can contain Systems.
 
-A Component is an atomic process that generally converts a physical phenomen to a digital number.
+A Component is an atomic process that generally converts a physical phenomena to a digital number.
 
-ProcessModels are another modelling that is used to describe a pure process that has no physicality. They point to an implementation or a description of one.
+ProcessModels are another modeling that is used to describe a pure process that has no physicality. They point to an implementation or a description of one.
 
@@
 
-Aggregation of primative data types are also provided DataRecord or DataArray.
+Aggregation of primitive data types are also provided DataRecord or DataArray.
 
@@
 
-Six other enhanced operations, GetResult,
-GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
-DescribeObservationType, and DescribeResultModel are not considered are not included in the SenseTileWeb Service.
-
 These operations are encoded in XML and used in http post and get operations.
@@
 
- The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis.
-
 para: Scalability would be achieved through aggregating a number of
@@
 
+The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis. Also six enhanced operations, GetResult,
+GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
+DescribeObservationType, and DescribeResultModel are not included in the SenseTileWeb Service.
+
 \section{OGC Lib}
-DOM based. Reference on google code. Refernce of how to build. Quicker to build prototype. SensorML aware.
-This open source project aims at building a JAVA API and implementation of the SWE (Sensor Web Enablement) Common Data Model, that can be easily used to access and produce SWE Common Data, and convert in to/from existing formats such as NetCDF, HDF, GeoTIFF, JPEG2000, etc...
-It can also be used to quickly develop SWE services such as SOS, SPS and SAS which all rely on the SWE Common Data Model and thus to build a Sensor Network.
-The project was initiated by the VAST Team at the University of Alabama in Huntsville (UAH) and is now also maintained by Sensia Software, Spot Image, and Botts Innovative Research, Inc.. 
-http://code.google.com/p/swe-common-data-framework/
-
+The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA framework call swe-common-data-framework (http://code.google.com/p/swe-common-data-framework/) to allow the quick development of SWE services such as an SOS.
+In this project the library is used to as the basis for the DataProducer as described in the SOS section.
+\section{UCD Sensor Board library}
 \section{BON}
-\section{Trchnologies}
+\section{Technologies}
 JAXB
@@
 
-SensorML System and Component were used to describe the SenseTile system and it's sensors. The sensor specifications were used to generate the metadata.
-The Sensors are modelled a detectors. They are process that take scalar values
-and can perform transformations on them.
+A SensorML model of SenseTile was developed as part of this thesis.
+The model contains metadata about the SenseTile system and supports the needs of the SenseTile Web Service.
+Two main requirements of the SenseTile Web Service were to be supported by the SensorML model.
+Firstly the raw data from the senors was to be retievable to allow external system process the data as desired.
+Secondly there was a requirement to do post processing on the Processor Unit of the sensor data to convert it to meaningful values.
 
-System is providing the mapping of inputs to processes and the outputs.
-
-Naming rules inputs and outputs - allow more generic handling of implementation classes.
-
-Offerings Processed verus raw data.
-
-It is further recognized that most sensor observations consist of at least two processes: a sampling process and a conversion process.
-
- \begin{figure}
-\scalebox{0.65}{\includegraphics{SenseTileSystem.pdf}}
-\caption{SenseTile System Overview}\label{fig:SenseTileSystem.pdf}
+A block diagram of the Sensetile SensorML model is shown in figure \ref{fig:SenseTileSystem}.
+\begin{figure}[h]
+\scalebox{0.70}{\includegraphics{SenseTileSystem.png}}
+\caption{SenseTile System Block Diagram}\label{fig:SenseTileSystem}
 \end{figure}
 
- \begin{figure}
-\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_2_pa1.pdf}}
-\caption{SenseTile System Components}\label{fig:SensorML_SenseTile_System_2_pa1.pdf}
+Three SensorML Systems are used to model the SenseTile system.
+One SensorML System models the overall SenseTile System and connects the other two SensorML Systems used, the Sensor Board System and the Conversion System.
+The physical phenomena measured by the sensors are the input to the  SenseTile System.
+These values will be read in from packets received from the Sensor Board.
+The Digital Number output is the raw sensor data which is a digital number without a physical meaning.
+The Converted Measurement output contains the values from post processing the sensor data which have some physical meaning such as temperature in Celcius.
+
+The division of the systems is based on the concept that sensor observations are based on at least a sampling process and a conversion process{ref tutorial}.
+The Sensor Board System contains SensorML Components that model the sensors on the board that perform the sampling while the SensorML Components contained in the Conversion System model process that produce meaningful physical measurement values such as temperature and lumens.
+
+A block diagram of the Sensetile SensorML model with Componets is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
+Only the Thermistor Component and the Celcius Component are developed as part of this thesis.
+They are described in model detail in later sections.
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}
+\caption{SenseTile System Components}\label{fig:SensorML_SenseTile_System_comp}
 \end{figure}
 
-\subsection{SenseTile System}
-\subsection{Light Sensor}
-\subsection{Thermistor}
-\subsection{Pressure Sensor}
-\subsection{Accelerometer}
-\subsection{Acoustic Sensor}
+Naming rules for inputs and outputs - allow more generic handling of implementation classes.
 
-Types
- - SWE -
+\subsection{Thermistor Component}
+The SensoML Thermistor component contains the metadata for the Sensor Board Thermistor x.
+It's temperature input is modeled using a quantity value without any units as is a measured phenomena.
+Its output is a digital number from the sensor.
+\subsection{Celcius Converter Component}
 
-Count - BigInteger
-Quantity - Double
-
-DataArray
-
- * Carries an array of int primitives.
- * All data is casted to other types when requested.
+The SensoML Celcius Converter component performs a conversion of the Thermistors digital number output to a real quantity celcius value.
 
@@
 
-How to display this in thesis?
 \section{SenseTile Web Service}
@@
 
-Web Service - SOS -standard access to sensor observations
-The Sensor Observation Service (SOS) is an approved Open Geospatial Consortium standard.
-The standard defines a web service interface for the discovery and retrieval of real time or archived data produced by all kinds of sensors like mobile or stationary as well as in-situ or remote sensors.
+The SenseTile web Service is based on the Sensor Observation Service (SOS) as described previously.
+The concept of a DataProvider as described by the SOS and an SOS are core to the architecture of the prototype.
+The overall system structure is shown in figure \ref {fig:SystemOverview} .
 
-The sensor data can be either observations or descriptions of the producing sensors containing metadata like calibration information, positions, etc.
-Observations are returned encoded as OM Observations, while information about sensors are returned encoded in SensorML or TML.
-Used in conjunction with these OGC standards and other OGC specifications, the SOS provides a broad range of interoperable capability for discovering, binding to and interrogating individual sensors, sensor platforms, or networked constellations of sensors in real-time, archived or simulated environments.
-SOS diagram
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{SystemOverview.png}}
+\caption{Overview}\label{fig:SystemOverview}
+\end{figure}
+In this prototype there is a divergance from how the SOS is envisoned to be implemented.
+In the SenseTile Web Service the SOS runs on the Processor Unit to allow access to the sensor data directly from the SenseTile rather than on a external large server.
 
-DataProvider
+It is envisioned that the SOS does not need to be run on every Processor Unit. Several SenseTile could excute just the DataProducer software and be configured with the address of the SOS to use.
 
-DataProducer
+\begin{table}[!th]
+\begin{tabular}{|l|c|r|}
+\hline
 
-Deployment Scenarios - diagram
+Scenarios\\
+\hline
+Generate Observations from Sensor Data \\
+\hline
+Get Sensor Description from SOS\\
+\hline
+Get  Sensor Observation from SOS\\
+\hline
+\end{tabular}
+\caption{Scenario Chart for SenseTile Web Service}
+\label{ex:table}
+\end{table}
 
-Flexible Division - several dataproducers can register with one DataProvider.
+\subsection{DataProducer}
 
-Seperate processes rmi.
- \begin{figure}
-\scalebox{0.65}{\includegraphics{SystemOverview.png}}
-\caption{Overview}\label{fig:SystemOverview.png}
-\end{figure}
+DataProducer generates observations in Observation xml standard.
 
- \begin{figure}
+Interface to the SensorBoard Library.
+
+Uses observer pattern to update sensors with packets from the board.
+
+reads sensors from SensorBoard System xml
+
+Sensors iimplement interface to facilitate this.
+
+Vast lib connects all the processes
+
+implementation of sensor and convertor is named using URN read from the sensorMl descption.
+
+The DataProducer was developed as a SensorML based software.
+It would be parse the SenseTile SensorML model to instantiate the code to excute the processes described.
+This required two aspects to be developed, a framework to run the processes and classes that implemented the components.
+
+\begin{figure}[h]
 \scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
@@
 
-\subsection{DataProducer}
+Interfaces to the Sensor Board
 Use PacketInput stream to access SenseTile observation packets.
 
-System reads packets and averages them for every configurable seconds. Default 1.
+\subsubsection{Sensors}
 
@@
 and its output is sent to both the raw data and the celcius
-converter component and on to the tuples.
-
+converter component and on to the tuples. System reads packets
+and averages them for every configurable seconds. Default 1.
 These tuples are sent to webservice provider as observations.
@@
 
-\subsection {DataProvider}
- \begin{figure}
+\subsection {SOS}
+ \begin{figure}[h]
 \scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
@@
 
-provides an RMI interface to the DataProducer to update with observations.
+provides an RMI interface to the DataProducer to update with observations. This has lower overhead
+than a web service. Though it would not be hard to change to use a Web Service Interface.
 
@@
 
-\section{SenseTile WebService}
+\section{DataProducer}
 
-DataProvider is based on AXIS 2.0/Spring
+DataProducer VastLib UCD SenseTile Driver Lib
 
-DataProducer POJO/VastLib/UCD SenseTile Driver Lib
+\section{SOS}
+ is based on AXIS 2.0/Spring
 
-RMI used between DataProvider and DataProducer.
+JAXB for SensorML XML to java binding.
+
+RMI used between DataProvider and SOS.
 
@@
 \chapter{Results}
+
+running prototype tested with UCD SensorBoard Simulator code
 
@@
 what has been achieved
+
 the weaknesses of your approach
+SOS on the Processor Board. Is this right.
+
+difficult schemas to program against.
+JIBX did not generate a default binding.
+Jaxb did with some help but did not generate the process type correctly.
+
+Schemas cover all needs for sensor description and data.
+
+SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
 
@@
 
+\lstset{language=XML,basicstyle=\small}
+\lstinputlisting{Thermistor.xml}
+
 \end{document}

### 2010-03-03 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates

@@
 Sensor Board measurements that have been post processed and converted to real values or the raw Sensor Board data can be accessed using the Web Service.
+SOS Lite.
 
@@
 \section{Sensor Web Enablement}
-The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised “Sensor Web”.
-It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.They provide a framework of open standards for use with Web-connected sensors and sensor systems.
+The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised Sensor Web.
+It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
 There are seven main specifications currently:
@@
 
-Other ways EEML? DAML?
+ DAML?
 
@@
 
-The purposes of SensorML are to:
+The purposes of SensorML are:
 \begin{itemize}
@@
 Observations are encoded as OGC O\&M Observations.
-Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML with SensorML being used by the SenseTile Web Service.
-
+Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
+para. An SOS organizes collections of related sensor system observations into Observation
+Offerings.
+These operations are encoded in XML and used in http post and get operations.
+The XML message needs to be parsed to identify the operations, parameters and results.
+There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP.
+I could not find this as part of current OGC SWE framework.
+para: Scalability would be achieved through aggregating a number of
+sensor data producers on the sensor side and by aggregating a number of SOS instances
+on the data center side for providing data to consumers. 
+\subsection{SOS Data Consumer Operations}
 There are there core operations that are provided by an SOS and used by SOS Clients to request sensor information:
@@
 
-Two other operations non RegisterSensor and InsertObservation are of interest to this thesis
-with the obvious functionality.
+\subsection{SOS Data Provider Operations}
+Two other operations RegisterSensor and InsertObservation are of interest to this thesis .
 These are part of the SOS transactional interface.
@@
 
-These operations are encoded in XML and used in http post and get operations.
-The XML message needs to be parsed to identify the operations, parameters and results.
-There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP.
-I could not find this as part of current OGC SWE framework.
-
-para: The SOS is expected to be running on a  server that  handle large numbers of connections
-from sensor systems.
-
-which can easily handle a large number of simultaneous connections from data
-producers.
-
-para: Scalability would be achieved through aggregating a number of
-sensor data producers on the sensor side and by aggregating a number of SOS instances
-on the data center side for providing data to consumers.
-
-The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis. Also six enhanced operations, GetResult,
+\subsection{SOS Catalogs}
+The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis. 
+\subsection{Enhanced Operations}
+Also six enhanced operations, GetResult,
 GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
@@
 The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA framework call swe-common-data-framework (http://code.google.com/p/swe-common-data-framework/) to allow the quick development of SWE services such as an SOS.
-In this project the library is used to as the basis for the DataProducer as described in the SOS section.
+In this project the library is used for it's SensorML parsing and SensorML process execution functionality.
 \section{UCD Sensor Board library}
+UCD Research group providing a Java lib to access the data 
 \section{BON}
-\section{Technologies}
-JAXB
-JAVA
-AXIS 2.0
-SPRING
+\section{JAXB}
+\section{AXIS 2.0}
 
@@
 
-A block diagram of the Sensetile SensorML model with Componets is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
+A block diagram of the Sensetile SensorML model with Components is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
 Only the Thermistor Component and the Celcius Component are developed as part of this thesis.
@@
 
+System and components are in several files. Inline in one file if desired.
+
 \subsection{Thermistor Component}
@@
 Its output is a digital number from the sensor.
+
+contraints.
+
+fragment
 \subsection{Celcius Converter Component}
@@
 The SensoML Celcius Converter component performs a conversion of the Thermistors digital number output to a real quantity celcius value.
+
+make portable
+formula
 
@@
 
-The SenseTile web Service is based on the Sensor Observation Service (SOS) as described previously.
-The concept of a DataProvider as described by the SOS and an SOS are core to the architecture of the prototype.
-The overall system structure is shown in figure \ref {fig:SystemOverview} .
+The SenseTile Web Service is based on the Sensor Observation Service (SOS) as described previously.
+The two main entities in the SOS, the Sensor Data Provider and the SOS,  are used to structure the SenseTile Web Service.
+The two entities are implemented as seperate software component that together provide the service.
+The Sensor Data Provider component will be referred to as the DataProvider in this design.
+
+The two components are run as seperate processes on the SenseTile processor unit to allow a more flexible network architecture.
+This network architecture is described in more detail in section.
+The basic system structure is shown in figure \ref {fig:Deployment_sensetile} .
 
 \begin{figure}[h]
-\scalebox{0.65}{\includegraphics{SystemOverview.png}}
-\caption{Overview}\label{fig:SystemOverview}
+\scalebox{0.65}{\includegraphics{Deployment_sensetile.png}}
+\caption{SenseTile Web Service}\label{fig:Deployment_sensetile}
 \end{figure}
-In this prototype there is a divergance from how the SOS is envisoned to be implemented.
-In the SenseTile Web Service the SOS runs on the Processor Unit to allow access to the sensor data directly from the SenseTile rather than on a external large server.
 
-It is envisioned that the SOS does not need to be run on every Processor Unit. Several SenseTile could excute just the DataProducer software and be configured with the address of the SOS to use.
+The DataProvider will be configured with a SensorML description of the SenseTile System, read sensor data form the Sensor Board and provide these as O\&M Observations to the SOS.
+The SenseTile SOS will implement the SOS core operations to allow a client to get the observations.
+The following sections describe the design of the DataProvider and SOS components as well as the scenarios they support.
 
+\subsection{Supported Scenarios}
 \begin{table}[!th]
@@
 \hline
-
 Scenarios\\
@@
 \end{table}
+\subsection{SenseTile WebService Network Architecture}
+In this prototype there is a divergance from how the SOS is envisoned to be implemented.
+In the SenseTile Web Service the SOS runs on the Processor Unit to allow access to the sensor data directly from the SenseTile rather than on a external large server.
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{Deployment_sensetile_network.png}}
+\caption{SenseTile Web Service}\label{fig:Deployment_network}
+\end{figure}
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
+\caption{SenseTile Web Service}\label{fig:Deployment_ext_sos}
+\end{figure}
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{Deployment_sensetile_sos_aggr.png}}
+\caption{SenseTile Web Service}\label{fig:Deployment_sos_aggr}
+\end{figure}
+It is envisioned that the SOS does not need to be run on every Processor Unit. Several SenseTile could excute just the DataProducer software and be configured with the address of the SOS to use.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
@@
 \subsection {SOS}
- \begin{figure}[h]
+ \begin{figure}[h!]
 \scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
-\caption{DataProvider Static Diagram}\label{fig:bon_static_diagam_provider.png}
+\caption{SOS Static Diagram}\label{fig:bon_static_diagam_provider.png}
 \end{figure}
@@
 Jaxb did with some help but did not generate the process type correctly.
+Why?
 
@@
 SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
+flexible.
+Streaming the data.
 

### 2010-03-07 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates

@@
 \chapter{Design}
-\section{SensorML Description of SenseTile}
+\section{SensorML Model of SenseTile}
 
@@
 Secondly there was a requirement to do post processing on the Processor Unit of the sensor data to convert it to meaningful values.
+A block diagram of the Sensetile SensorML model is shown in figure \ref{fig:SenseTileSystem}.
 
-A block diagram of the Sensetile SensorML model is shown in figure \ref{fig:SenseTileSystem}.
 \begin{figure}[h]
-\scalebox{0.70}{\includegraphics{SenseTileSystem.png}}
+\framebox{\scalebox{0.70}{\includegraphics{SenseTileSystem.png}}}
 \caption{SenseTile System Block Diagram}\label{fig:SenseTileSystem}
@@
 
-Three SensorML Systems are used to model the SenseTile system.
-One SensorML System models the overall SenseTile System and connects the other two SensorML Systems used, the Sensor Board System and the Conversion System.
+Three SensorML Systems are used to model the SenseTile as shown in the block diagram.
+One SensorML System models the overall SenseTile which contains the Sensor Board and Conversion Systems.
+The below XML fragment shows how SensorML describes System containment using a component list and how the other SensorML Systems descriptions are referenced using xlink hrefs.
+
+\lstset{language=XML,basicstyle=\scriptsize,frame=single}
+\begin{lstlisting}
+<sml:components>
+        <sml:ComponentList>
+               <sml:component name="sensorBoardSystem" 
+                                          xlink:href="SensorBoardSystem.xml">
+               </sml:component>
+               <sml:component name="convertorSystem"  
+                                          xlink:href="ConverterSystem.xml">
+               </sml:component>
+        </sml:ComponentList>			   
+ </sml:components>
+\end{lstlisting}
+
 The physical phenomena measured by the sensors are the input to the  SenseTile System.
@@
 The division of the systems is based on the concept that sensor observations are based on at least a sampling process and a conversion process{ref tutorial}.
-The Sensor Board System contains SensorML Components that model the sensors on the board that perform the sampling while the SensorML Components contained in the Conversion System model process that produce meaningful physical measurement values such as temperature and lumens.
+The Sensor Board System contains SensorML Components that model the sensors on the board that perform the sampling.
+The SensorML Components contained in the Conversion System model processes that produce meaningful physical measurement values such as temperature and lumens from the sampled data.
 
-A block diagram of the Sensetile SensorML model with Components is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
+The SenseTile System connects up the inputs and outputs of the SensorBoard and Convertor Systems using SensorML connections.
+The below SensorML fragment shows how the temperature output from the SensorBoard System is connected to the input of the Conversion System is peformed in SensorML.
+\newpage
+\begin{lstlisting}
+<sml:connection name="convertToTemperature">
+    <sml:Link>
+        <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
+        <sml:destination ref="convertorSystem/inputs/celciusConvInput"/>
+     </sml:Link>
+ </sml:connection>
+\end{lstlisting}
+
+A more detailed block diagram of the Sensetile SensorML model with Components is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
 Only the Thermistor Component and the Celcius Component are developed as part of this thesis.
@@
 \begin{figure}[h]
-\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}
+\framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
 \caption{SenseTile System Components}\label{fig:SensorML_SenseTile_System_comp}
@@
 
-System and components are in several files. Inline in one file if desired.
+System and components are in several files. Inline in one file if desired.Choice to seperate files.
 
 \subsection{Thermistor Component}
-The SensoML Thermistor component contains the metadata for the Sensor Board Thermistor x.
-It's temperature input is modeled using a quantity value without any units as is a measured phenomena.
+The Thermistor on the SenseTile Sensor board is the Texas Instruments TMP175 Digital Temperature Sensor and is modeled with a SensorML Component.
+This is an indivisible process and can have a different location information than the Sensor Board if required.
+The SensorML Component allows the capabilities of the Thermsistor to be described.
+For example the TMP175 is  specified for operation over a temperature range of −40°C to +125°C.
+This is captured in the following SensorML:
+
+\begin{lstlisting}
+<swe:field name="TemperatureRange" 
+           xlink:arcrole="urn:ogc:def:property:dynamicRange">
+          <swe:QuantityRange definition="urn:ogc:def:property:temperature">
+                   <swe:uom code="cel" /> 
+                          <swe:value>-40 125/swe:value> 
+           </swe:QuantityRange>
+</swe:field>
+\end{lstlisting}
+
+The SensorML Thermistor temperature input is modeled using a Quantity value without any units as is a measured physical phenomena.
 Its output is a digital number from the sensor.
+The range is constrained to the allowed value range from the Thermistor.
+The following fragment shows how the input and outputs for the Thermistor look in SensorML:
 
-contraints.
+\begin{lstlisting}
+<sml:inputs>
+    <sml:InputList>
+         <sml:input name="thermistorInput">
+             <swe:Quantity definition="urn:ogc:def:phenomenon:temperature">
+         </sml:input>
+    </sml:InputList>
+</sml:inputs>
 
-fragment
+<sml:outputs>
+    <sml:OutputList>
+          <sml:output name="thermistorOutput">
+               <swe:Count>
+	        <swe:constraint>
+                         <swe:AllowedValue id="outputRange">
+                              <swe:interval>-880 2032</swe:interval>
+                         </swe:AllowedValue>
+                     </swe:constraint>
+                </swe:Count>
+           </sml:output>
+     </sml:OutputList>
+ </sml:outputs>
+\end{lstlisting}
+
 \subsection{Celcius Converter Component}
 
-The SensoML Celcius Converter component performs a conversion of the Thermistors digital number output to a real quantity celcius value.
+The SensoML Celcius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celcius value.
+It similar to the Thermistor SensorML description but its inputs and outputs are different.
+It process method performs a simple calcution to perform the celcius conversion.
 
-make portable
-formula
+A Sensor process method was also developed.
+The process method describes the coversion algorithm and points to an implementation class.
+Process Methods are not implemented by the Vast Lib so it not used in the running system.
+Instead the method tag provides a URN that is used to lookup the implementation class.
 
-\newpage
 \section{BON to SensorML mapping}
@@
 The SenseTile Web Service is based on the Sensor Observation Service (SOS) as described previously.
-The two main entities in the SOS, the Sensor Data Provider and the SOS,  are used to structure the SenseTile Web Service.
-The two entities are implemented as seperate software component that together provide the service.
+The two main entities in the SOS Specification, the Sensor Data Provider and the SOS,  are used to structure the SenseTile Web Service.
+These entities are implemented as seperate software components that together provide the service.
 The Sensor Data Provider component will be referred to as the DataProvider in this design.
@@
 The two components are run as seperate processes on the SenseTile processor unit to allow a more flexible network architecture.
-This network architecture is described in more detail in section.
+This network architecture is described in more detail in section \ref{subsec:SenseTile WebService Network Architecture}.
 The basic system structure is shown in figure \ref {fig:Deployment_sensetile} .
@@
 
-The DataProvider will be configured with a SensorML description of the SenseTile System, read sensor data form the Sensor Board and provide these as O\&M Observations to the SOS.
-The SenseTile SOS will implement the SOS core operations to allow a client to get the observations.
+The DataProvider parses a SensorML description of the SenseTile System, reads sensor data form the Sensor Board and provides the sensor data as O\&M Observations to the SOS.
+The SenseTile SOS will implement the SOS core operations to allow a client to get these observations.
 The following sections describe the design of the DataProvider and SOS components as well as the scenarios they support.
@@
 Generate Observations from Sensor Data \\
+The DataProducer accesses data from the sensor board \\
 \hline
@@
 \subsection{SenseTile WebService Network Architecture}
-In this prototype there is a divergance from how the SOS is envisoned to be implemented.
-In the SenseTile Web Service the SOS runs on the Processor Unit to allow access to the sensor data directly from the SenseTile rather than on a external large server.
+\label{subsec:SenseTile WebService Network Architecture}
+In the SenseTile Web Service the SOS acts as a gateway to the SenseTile Sensor Network.
+This prototype was designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
+A client would access multiple SOS instances to get observations from the sensor network.
+An SOS could support several DataProducers and hence would not to be run on every SenseTile.
+
 \begin{figure}[h]
 \scalebox{0.65}{\includegraphics{Deployment_sensetile_network.png}}
-\caption{SenseTile Web Service}\label{fig:Deployment_network}
+\caption{SenseTile Sensor Network Architecture}\label{fig:Deployment_network}
 \end{figure}
+The SOS Specification envisioned that the SOS would be run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}.
+The DataProviders would update this external server with observations and clients would just access this large SOS.
 \begin{figure}[h]
 \scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
-\caption{SenseTile Web Service}\label{fig:Deployment_ext_sos}
+\caption{External SOS}\label{fig:Deployment_ext_sos}
 \end{figure}
+The SOS Specification describes the use of a Proxy SOS as shown in figure \ref{fig:Deployment_sos_aggr}.
+This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.
 \begin{figure}[h]
 \scalebox{0.65}{\includegraphics{Deployment_sensetile_sos_aggr.png}}
-\caption{SenseTile Web Service}\label{fig:Deployment_sos_aggr}
+\caption{Proxy SOS}\label{fig:Deployment_sos_aggr}
 \end{figure}
-It is envisioned that the SOS does not need to be run on every Processor Unit. Several SenseTile could excute just the DataProducer software and be configured with the address of the SOS to use.
 
+\newpage
 \subsection{DataProducer}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
 \scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
@@
 
+\newpage
 \subsection {SOS}
@@
 
-\lstset{language=XML,basicstyle=\small}
 \lstinputlisting{Thermistor.xml}

### 2010-03-11 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
updates

@@
 \subsection{DataProducer}
+The DataProducer's role, as described previously, is to read sensor data from the Sensor Board and generate O\&M observations.
+These observations are then sent to the SOS instance.
+The processing of the sensor data to generate the observations is described by the SensorML models.
+The DataProducer uses the OGC Vast Library to instantiate objects that implement the required data processing chains based on the SensorML models.
+The BON static diagram \ref{fig:bon_static_diagam_producer.png}
+ shows the main clusters and classes of the DataProvider.
 
+The SMLEngine cluster contains classes that provide access to
+the Vast Lib that is providing the SensorML Engine.  The Vast Lib implemenation
+of the SensorML Engine is hidden from the rest of the code thus.
+
+The Sensors and Converters clusters contain classes that provide
+implementation of process methods for SensorML components in the SensorBoard and
+Converter SensorML Models respectively. These are referred to as sensor objects and
+converter objects henceforth.
+
+The DataProducer operates by repeating the below steps continously.
+ \begin{enumerate}
+\item Read Sensor Board Data
+\item Process Sensor Data
+\item Create O\&M Observations
+\end{enumerate}
+The steps above are detailed the in following sections with reference
+to the BON static diagram \ref{fig:bon_static_diagam_producer.png}.
+
+\subsubsection {1. Read Sensor Board Data}
+
+The UCD SensorBoard Library is used by DataProducer to access
+the data generated by the Sensor Board.
+The library provides high level interface to the
+the data from the sensor board. It reads data from the Sensor board and generates packets
+containing the sensor data. These packets provide functionality
+to access a particular sensors data such as the temperature
+meaurement generated by the thermistor. 
+This library is shown as the SensorBoard cluster in \ref{fig:bon_static_diagam_producer.png}
+An instance of the class PacketInputStream to is used access the packets from the Sensor Board.
+
+The DataProducer uses the Observer pattern to provide
+packets to the SensorMLEngine for processing.
+At start up the SenseTileSystem reads the list of sensor objects from the SMLEngine that
+implement the Sensors interface. These objects provide an entry point
+to the SensorML processing chain and can read data from the UCD LIB packets.
+These sensor objects implement reading strategies for the type of data they read. 
+For example the ThermistorSensor reads out the temperature value
+but will average it before it sent to the SensorML Engine for processing.
+
+\subsubsection {2. Process Sensor Data}
+The PacketInputStream will request the SenseTileSystem to exexcute
+the SensorML engine to process the sensor data. The SensoML Enigine
+will execute the process chains described in the SenseTile SensorML model.
+The sensor objects output will be passed to the coverter objects. The
+outputs of both object types are then available to generate observations.
+
+\subsubsection {3. Create O\&M Observations}
+
+When the SensorML Engine is finished
+and there are obsevations generated these
+need to be sent to the SOS. An O\&M Onservation
+xml string containing the data is sent to
+the SOS instance. This is timestamped at
+this stage.
+ containing the output from
+the SensorML processes and send the observations
+to the SOS.
+-register offerings with the SOS. In this design
+each system is an offering.
+Need to generate an observation model?
+
+Questions.
+If have light data but no temp. As temp is only
+every 1000 packets?
+
+How to know when to generate observation
+for the two offerings?
 DataProducer generates observations in Observation xml standard.
@@
 \begin{figure}[h]
-\scalebox{0.65}{\includegraphics{sensetile_static_diagam.png}}
-\caption{DataProducer Static Diagram}\label{fig:sensetile_static_diagam.png}
+\scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
+\caption{DataProducer Static Diagram}\label{fig:bon_static_diagam_producer.png}
 \end{figure}
@@
 Interfaces to the Sensor Board
-Use PacketInput stream to access SenseTile observation packets.
-
-\subsubsection{Sensors}
-
-Thermistor components reads out the temperature value
-and its output is sent to both the raw data and the celcius
-converter component and on to the tuples. System reads packets
-and averages them for every configurable seconds. Default 1.
-These tuples are sent to webservice provider as observations.
-
-Observation as described in SOS.
 
@@
 
+A lot of schemas needed.
+
 SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
 flexible.
+
+Vast Lib use as a references only. Other ways such a XQUERY? XML techs?
+As there are a large number of types etc. A quick way for a prototype.
+Had to update library code to get it to work with the SenseTile Schemas. Schemas
+passed the validation test tool from the site.
+
+Portabiltiy, Vast Types POJOs.
+
 Streaming the data.

### 2010-03-21 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 
+Vieri del Bianco , Dragan Stosic and Joseph R. Kiniry.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
-The aim of this thesis project was to model the SenseTile Sensor Board using Sensor Model Language (SensorML)\cite{SensoMLref} .
+The aim of this thesis project was to model the SenseTile Sensor Board using Sensor Model Language (SensorML)\cite{SMLref} .
 A prototype of a Web Service for accessing SenseTile Sensor Board sensor observations was to be developed using this SensorML description of the SenseTile Sensor Board.
@@
 
-SenseTile is a sensor and processing package including motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
+SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
 It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
@@
 \section{Sensor Web Enablement}
-The Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) [ref]  is establishing interfaces and protocols that will enable a standardised Sensor Web.
-It is hoped that an easier integration and development of sensor applications across sensor networks will be facilitated by this standarising activity.The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
+SensorML is developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative (http://www.opengeospatial.org/projects/groups/sensorweb).
+This initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
+It is hoped that this allows for easier integration and development of sensor applications across sensor networks.
+The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
 There are seven main specifications currently:
@@
 \item  SensorModel Language(SensorML) - models and schema for sensor systems and processes surrounding measurements
+\item  Sensor Observation Service (SOS) - standard web interface for accessing observations
 \item  Observations \& Measurements (O\&M) - models and schema for packaging observation values
 \item  Transducer Markup Language (TML) - models and schema for multiplexed data from sensor systems
-\item  Sensor Observation Service (SOS) - standard web interface for accessing observations
 \item  Sensor Planning Service (SPS) - standard web interface for tasking sensor systems and model and requesting acquisitions
@@
 \end{itemize}
-There is also the SWE Common that provides common data models and schemas for the above.
+
+SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to con-trol observations.
+For a good overview of SWE and it's use in Sensor Web Network please see \cite{SWEArchref}
 
 Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand.
-The SWE framework seems to cover all needs for modeling sensors and observations.
-Support the use of Web Services is the main rational and fits the needs of a SenseTile Sensor Network.
-In this case the SOS and SensorML are core.
-
-Scalable - can be  large systems. Discovery. Process data without having to define an implementation as this
-can be provided as part of SensorML description.
-
- DAML?
+The SWE framework seems to cover all that is needed for modeling sensors and observations.
+Support for easy exchange of sensor meta-data and sensor observations is the main rational of SWE and fits the needs of the SenseTile Sensor Network.
+SensorML, SOS and O\&M specifications are the focus of this thesis and are described in more detail in the following sections.
 
 \section{Sensor Modeling Language}
-Sensor Modelling Language (SensorML) is used to describe sensor systems and the processing of observations from sensor systems.
-
-The purposes of SensorML are:
+Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems.
+The SensorML specification\cite{SMLref} provides a functional model of sensors and an XML encoding to describe sensors and their observations.
+Some of the main uses of SensorML described in the specification are:
 \begin{itemize}
-\item Provide sensor and process information in support of resource and observation discovery
-\item Support the processing and analysis of the sensor observations
-\item Support the geolocation of observed values (measured data)
-\item Provide performance characteristics (e.g., accuracy, threshold, etc.)
-\item Provide an explicit description of the process by which an observation was obtained (i.e., it’s lineage)
-\item Provide an executable process chain for deriving new data products on demand (i.e., derivable observation)
-\item Archive fundamental properties and assumptions regarding sensor systems
+\item support sensor and sensor observation discovery
+\item support processing of the sensor observations
+\item support geolocation information about sensors
+\item provide accuracy information of sensors measurements
+\item Provide an explicit description of the process by which an observation was obtained
+\item Provide an executable process chain for deriving new data products on demand
 \end{itemize}
 
-SensorML provides a functional model of sensors and an XML encoding to describe sensors and their observations.
-Sensors are modelled as processes that convert real phenomena to data. Processes take inputs, process them and result in one or several outputs.
+In SensorML sensors are modeled as processes that convert real phenomena to data.
+Processes take inputs, process them and result in one or several outputs.
+For example, a process can take a digital number generated by a thermistor and transform it to a celcius value.
+Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially.
+These processes and process chains are modeled in SensorML as shown in the  below SensorML Conceptual Model \ref{fig:SMLConceptualModel}.
 
-Sensor metadata is also supported by SensorML.
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{SMLConceptualModel.png}}}
+\caption{SensorML Conceptual Model}\label{fig:SMLConceptualModel}
+\end{figure}
 
-The processes can be connected together in chains using SensorML ProcessChains or Systems.
+Processes in SensorML derive from AbstractProcess with it's inputs, output, parameters and metaDataGroup properties.
+The inputs and outputs represent data connections to other processes.
+Parameters represent inputs other than sensor measurements or outputs of other processes.
+They are used to configure the process or provide other inputs that are not sensor measurements.
+For example a parameter might be the latency time of a measurement needed for a calculation.
+Finally each process provides sensor metadata as modeled by the MetaDataGroup.
 
-including measurement by a sensor system, as well as post-measurement
-processing.
+SensorML divides processes into two types, physical and non-physical.
+The non-physical process is one where  location, position etc.
+is not relevant to the process.
+Physical processes provide physical information, such as location, which can be used in processing if required.
+The ProcessModel is a non-physical process while the Component is a physical process and contains extra properties for defining the needed spatial or temporal information.
+Both process types contain the processMethod property which points to the methodology for transforming inputs to outputs.
 
-When Modeling a sensor system in SensorML the main entities used are Systems and Components.
+There is a similar division of the process chains into the non-physical ProcessChain and physical System.
+Both process chains can contain a mixture of ProcessModel and Component process if desired.
+Process chains have inputs and outputs which define the beginning and end of the processing chain.
+They are points of connection to other processes and process chains if needed.
+The connection property contained in the ProcessChain and System is a list of Link type.
+The Link type contains source and destination propertes that reference inputs and outputs of the processes in the process chain thus connecting the processes.
 
-A SensorML System is used to group sensors that are related spatially to one another.  As it is a process chain it is hierarchical and so Systems can contain Systems.
+The SensorML specification currently contains the Sensor Web Enablement (SWE) Common specification.
+This specification defines basic types and data encodings used by the SWE specifications.
+It describes both primitive data types as well as aggregate types such as arrays and records.
+There are a large number of types defined in this specification, too many to show here.
+Below is a Conceptual Model of some of the simple data types.
 
-A Component is an atomic process that generally converts a physical phenomena to a digital number.
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{SWESimpleConceptualModel.png}}}
+\caption{SWE Common Simple Data Types.}\label{fig:SWESimpleConceptualModel}
+\end{figure}
 
-ProcessModels are another modeling that is used to describe a pure process that has no physicality. They point to an implementation or a description of one.
+The data types in \ref{fig:SWESimpleConceptualModel} are scalar primitive values.
+They are the basis for the inputs, outputs and parameters in SensorML processes.
+Quantity models a floating point number.
+Count models a counting value or integer.
+Boolean models a property that can be true or false.
+These types contain a number of properties that further define the type.
+The uom property provides a unit of measurement that indicate how the value should be interpreted.
+A constraint property allows a value range or a enumerated list of values to be defined for the type.
+The quality property allows for a measure of its quality, for example the confidence level of a value being correct expressed as a percentage.
+Aggregation of these primitive data types is also provided by several types including DataRecord and DataArray.
 
-The data types used for inputs and out puts come another OGC sepc the SWE common.
-Examles are Quantity which is a decimal number, Count, Boolean, Category, and Time provide the basic primitives.
-Within SensorML,
+It is stated in the SensorML specification that the SWE Common will be placed in it's own document in future releases.
 
-Aggregation of primitive data types are also provided DataRecord or DataArray.
-
-A sensor systems would likely to be based mainly on Systems and
-Detector model. Reference tutorial and sensorml specification.
-
-The SensorML description of the SenseTile Sensor Board is described in chapter .
+The XML encoding of the SensorML model is not described here. In section where a detailed description of a SensorML model of SenseTile is described.
 
 \section{Sensor Observation Service}
-The SenseTile Web Service is based on the concepts described in the Sensor Observation Service (SOS) specification.
+SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SWEArchref}.
 The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.
-Observations are encoded as OGC O\&M Observations.
-Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
-para. An SOS organizes collections of related sensor system observations into Observation
+The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
+
+The SOS specification uses four main entities to describe the SOS's operation.
+The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself.
+Data Consumers are clients of the SOS that request sensor observations and sensor metadata.
+Data Providers are entities have to access to the sensor and have enough processing power to generate an XML encoding of the sensor's observations.
+These observations are sent over a network to the SOS.
+The OGC Catalog is used store metadata about the sensor systems and is used as part of the discovery process.
+
+The diagram \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{SOSoperationContext.png}}}
+\caption{SOS Operational Context}\label{fig:SOSoperationContext}
+\end{figure}
+
+Collections of related sensor system observations are organized into Observation
 Offerings.
+
+Observations are encoded as OGC O\&M Observations. Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
+para.
 These operations are encoded in XML and used in http post and get operations.
@@
 \item GetCapabilities operation is used to retrieve SOS service metadata.
-\item DescribeSensor operation retrieves detailed information about the sensors .
+\item DescribeSensor operation retrieves detailed information about the sensors.
 \end{itemize}
@@
 \subsection{SOS Data Provider Operations}
-Two other operations RegisterSensor and InsertObservation are of interest to this thesis .
+Two other operations RegisterSensor and InsertObservation are of interest to this thesis.
 These are part of the SOS transactional interface.
@@
 
-\section{OGC Lib}
-The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA framework call swe-common-data-framework (http://code.google.com/p/swe-common-data-framework/) to allow the quick development of SWE services such as an SOS.
-In this project the library is used for it's SensorML parsing and SensorML process execution functionality.
-\section{UCD Sensor Board library}
-UCD Research group providing a Java lib to access the data 
+\section{SenseTile}
+A case study of the use of SensorML in sensor networks is based on the SenseTile System. SenseTile is a general purpose sensor system developed at UCD CASL (http://kind.ucd.ie/
+products/opensource/SenseTile).
+The SenseTile System is used to investigate issues arising with large scale sensor networking.
+SenseTile is made up of a sensor platform, a datastore, and a compute farm.
+The sensor platform is called the SenseTile Unit and is composed of one or more SenseTile Sensor Boards paired with a SenseTile Processor Unit.
+The Processor could be a PDA or small PC or any processing device that has a USB connection.
+The Sensor Board has a USB connection and this is used for communication with the Processor Unit.
+There are over a dozen sensors on the SenseTile Sensor Board including a thermistor and light sensor.
+Sensors are mounted on the Sensor Board and more sensors can be added using the USB connection.
+
+This SenseTile unit is where thesis will focus on the use of SensorML though SensorML can be used in other the parts of the SenseTile system.
+As the processor unit is  a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible.
+The part SensorML plays in this Web Service is explored in the context of the SenseTile system.
+
+\section{UCD Sensor Board Driver}
+
+\section{VAST SensorML Engine}
+The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA based SensorML Processing Engine.
+This software parses SensorML and instantiates objects that implement SensorML processes and process  chains.
+The code is found at http://code.google.com/p/sensorml-data-processing/.
+It dependent on an implementation of the SWE Common data types described in the chapter on SensorML.
+This code can be found at call swe-common-data-framework http://code.google.com/p/swe-common-data-framework/.
+This software is referred to as the VastSMLEngine in rest of the thesis.
+
 \section{BON}
-\section{JAXB}
-\section{AXIS 2.0}
 
@@
 
-A SensorML model of SenseTile was developed as part of this thesis.
+A SensorML model of SenseTile is developed as part of this thesis.
 The model contains metadata about the SenseTile system and supports the needs of the SenseTile Web Service.
 Two main requirements of the SenseTile Web Service were to be supported by the SensorML model.
-Firstly the raw data from the senors was to be retievable to allow external system process the data as desired.
-Secondly there was a requirement to do post processing on the Processor Unit of the sensor data to convert it to meaningful values.
+Firstly the raw data from the senors is to be retrievable to allow external system process the data as desired.
+Secondly there is a requirement to do post processing on the Processor Unit of the sensor data to convert it to meaningful values.
 A block diagram of the Sensetile SensorML model is shown in figure \ref{fig:SenseTileSystem}.
 
+\ref{fig:SensorML_SenseTile_System_comp}. Only the Thermistor Component and the Celcius Component are developed as part of this thesis. They are described in model detail in later sections.
 \begin{figure}[h]
-\framebox{\scalebox{0.70}{\includegraphics{SenseTileSystem.png}}}
-\caption{SenseTile System Block Diagram}\label{fig:SenseTileSystem}
+\framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
+\caption{SenseTile System SensorML Block Diagram}\label{fig:SensorML_SenseTile_System_comp}
 \end{figure}
@@
 \end{lstlisting}
-
-A more detailed block diagram of the Sensetile SensorML model with Components is shown in figure \ref{fig:SensorML_SenseTile_System_comp}.
-Only the Thermistor Component and the Celcius Component are developed as part of this thesis.
-They are described in model detail in later sections.
-\begin{figure}[h]
-\framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
-\caption{SenseTile System Components}\label{fig:SensorML_SenseTile_System_comp}
-\end{figure}
 
@@
 
-A Sensor process method was also developed.
+A Sensor process method is also developed.
 The process method describes the coversion algorithm and points to an implementation class.
@@
 \end{table}
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
+\caption{SensorML Conceptual Model}\label{fig:InsertObs}
+\end{figure}
+
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
+\caption{SensorML Conceptual Model}\label{fig:GetObs}
+\end{figure}
+\newpage
 \subsection{SenseTile WebService Network Architecture}
@@
 
-\newpage
-\section{SensorML generation}
-
-Java binding wth JIBX  generate sensorml.
-
-Simple xml file with Sensor metadata.
-
-Naming rules allow connection of inputs and outputs
-in sensorML processes.
-
- component id +Input
-
- component id+Output
-
-\chapter{ Detailed Design and Implementation}
+\chapter{Implementation}
 
@@
 \begin{thebibliography}{99}
-\bibitem{SensoMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
-\bibitem{SOSref}Open Geospatial Consortium Inc., OpenGIS  Implementation Specification, 2007
-\bibitem{OMref}Open Geospatial Consortium Inc., OpenGIS  Implementation Specification, 2007
-\bibitem{SWEArchref}Open Geospatial Consortium Inc., , 2007
-\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., , 2007
+\bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
+\bibitem{SOSref}Open Geospatial Consortium Inc.,  Sensor Observation Service, 2007
+\bibitem{OMref}Open Geospatial Consortium Inc., Observations and Measurements – Part 1 - Observation schema, 2007
+\bibitem{SWEArchref}Open Geospatial Consortium Inc., OGC Sensor Web Enablement Architecture, 2008
+\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., Using SensorML to describe a
+Complete Weather Station , 2006
+\bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,
 

### 2010-03-22 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 
-\section{Sensor Modeling Language}
+\section{Sensor Modeling Language}\label{SMLsection}
 Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems.
@@
 Processes take inputs, process them and result in one or several outputs.
-For example, a process can take a digital number generated by a thermistor and transform it to a celcius value.
+For example, a process can take a measurement generated by a thermistor which might be a resistance value and transform it to a celcius value.
 Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially.
@@
 
-It is stated in the SensorML specification that the SWE Common will be placed in it's own document in future releases.
+SensorML makes extensive use of XLink components to support hypertext referencing in XML.
+This allows one to reuse Objects that are either internal or external to the instance document, and allows these objects to perhaps have different associations or roles within one or more documents.
+This is supported by extensive use of the id attribute (taking an xs:ID as its value) within Objects, and utilizing the GML attribute group, gml:AssociationGroup, within property elements.
+SensorML is designed with the concept that there exist dictionaries or registries for definitive specification of various terms.
+these definitions should be specified in SensorML documents as resolvable URNs, while XLink references will be either URLs or URNs.
 
@@
 \section{Sensor Observation Service}
-SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SWEArchref}.
+SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref}.
 The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.
@@
 The OGC Catalog is used store metadata about the sensor systems and is used as part of the discovery process.
-
-The diagram \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context
+The SOS is an access point to a Sensor Web Network and provides the Web Service interface.
+The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context.
 \begin{figure}[h]
@@
 
-Collections of related sensor system observations are organized into Observation
-Offerings.
-
-Observations are encoded as OGC O\&M Observations. Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
-para.
-These operations are encoded in XML and used in http post and get operations.
-The XML message needs to be parsed to identify the operations, parameters and results.
-There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP.
-I could not find this as part of current OGC SWE framework.
-para: Scalability would be achieved through aggregating a number of
-sensor data producers on the sensor side and by aggregating a number of SOS instances
-on the data center side for providing data to consumers. 
-\subsection{SOS Data Consumer Operations}
-There are there core operations that are provided by an SOS and used by SOS Clients to request sensor information:
+The diagram above shows the core mandatory operations and transactional operations provided by an SOS.
+The core operations are used by Data Consumers to request sensor information.
+The operations perfom the following function:
  \begin{itemize}
@@
 
-\subsection{SOS Data Provider Operations}
-Two other operations RegisterSensor and InsertObservation are of interest to this thesis.
-These are part of the SOS transactional interface.
-They are used with SOS data producers which are software entities connected to the sensor and have enough processing power to generate XML descriptions of the sensor data.
+The transactional operations supported by the SOS toward the DataProvider are:
+ \begin{itemize}
+\item RegisterSensor operation registers the DataProvider with the SOS.
+\item InsertObservation operation updates SOS service metadata.
+\end{itemize}
 
-\subsection{SOS Catalogs}
-The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis. 
-\subsection{Enhanced Operations}
-Also six enhanced operations, GetResult,
+Collections of related sensor system observations are organized by the SOS into observation offerings. Observations are encoded as OGC O\&M Observations.
+
+These operations are encoded in XML and used in http post and get operations. The XML message needs to be parsed to identify the operations, parameters and results.
+
+Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
+para.
+
+The SOS Operation Context diargam also shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
+The Data Consumer uses a CS-W catalog to find SOS instances.
+The GetCapabilites are then used toward each of the SOS instances returned to find sensor systems that are of interest to the Data Consumer.
+The CS-W catalog is covered seperate specification \cite{OGCcatref} and is not covered in any real detail in the SOS specification.
+The OGC Catalog Service is not explored in this thesis.
+
+There are also six other operations the SOS refers to as enhanced operations. They are GetResult,
 GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
-DescribeObservationType, and DescribeResultModel are not included in the SenseTileWeb Service.
+DescribeObservationType, and DescribeResultModel. These operations are not covered in this thesis.
 
@@
 
-This SenseTile unit is where thesis will focus on the use of SensorML though SensorML can be used in other the parts of the SenseTile system.
+This SenseTile unit is used here as case study on the use of SensorML, though SensorML can be used in other the parts of the SenseTile system.
 As the processor unit is  a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible.
 The part SensorML plays in this Web Service is explored in the context of the SenseTile system.
+
+Thermistor?
 
@@
 \section{VAST SensorML Engine}
-The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA based SensorML Processing Engine.
-This software parses SensorML and instantiates objects that implement SensorML processes and process  chains.
-The code is found at http://code.google.com/p/sensorml-data-processing/.
-It dependent on an implementation of the SWE Common data types described in the chapter on SensorML.
-This code can be found at call swe-common-data-framework http://code.google.com/p/swe-common-data-framework/.
-This software is referred to as the VastSMLEngine in rest of the thesis.
+The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source Java based SensorML Processing Engine.
+This software provides types that implement SensorML processes and process  chains.
+The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing.
+The code is found at http://code.google.com/p/sensorml-data-processing and is referred to as the VastSMLEngine in rest of the thesis.
+This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}.
+This code is found at http://code.google.com/p/swe-common-data-framework.
 
@@
 
+Scalability - info here.
+
 \newpage
@@
 \chapter{ Conclusions and Future Work}
-
+There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
 what has been achieved
@@
 \bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,
-
+\bibitem{OGCcatref}Open Geospatial Consortium Inc., OpenGIS® Catalogue Services Specification, 2007
 \bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995

### 2010-03-24 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
Latest

@@
 The connection property contained in the ProcessChain and System is a list of Link type.
-The Link type contains source and destination propertes that reference inputs and outputs of the processes in the process chain thus connecting the processes.
+The Link type contains source and destination properties that reference inputs and outputs of the processes in the process chain thus connecting the processes.
 
@@
 
-SensorML makes extensive use of XLink components to support hypertext referencing in XML.
-This allows one to reuse Objects that are either internal or external to the instance document, and allows these objects to perhaps have different associations or roles within one or more documents.
-This is supported by extensive use of the id attribute (taking an xs:ID as its value) within Objects, and utilizing the GML attribute group, gml:AssociationGroup, within property elements.
-SensorML is designed with the concept that there exist dictionaries or registries for definitive specification of various terms.
-these definitions should be specified in SensorML documents as resolvable URNs, while XLink references will be either URLs or URNs.
+SensorML uses XML Linking Language (XLink) (http://www.w3.org/TR/xlink) to support hypertext referencing in SensorML XML documents using URNs and URIs.
+This allows SensorML models to be distributed over several documents.
+XLink attributes are also used to reference on-line unit definitions used to define a sensor measurement.
 
-The XML encoding of the SensorML model is not described here. In section where a detailed description of a SensorML model of SenseTile is described.
+The XML encoding of the SensorML model is not described here.
+The OGC SWE specification uses XML Schema to define the allowable SensorML XML structure.It is noted here that sml is the XML namespace for SensorML and swe is the XML namespace for SWE Common.
+In the later SenseTile SensorML model section examples of the XML encoding are used to describe the SenseTile Model.
 
@@
 The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.
-The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
+It organizes related sensor observations into a collection called an observation offering.
 
@@
 These observations are sent over a network to the SOS.
-The OGC Catalog is used store metadata about the sensor systems and is used as part of the discovery process.
-The SOS is an access point to a Sensor Web Network and provides the Web Service interface.
+The OGC Catalog is used to find SOS instances.
+SOS instances act as access points to a Sensor Web Network and provide the Web Service for retrieval of sensor observations.
 The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context.
@@
 
-The diagram above shows the core mandatory operations and transactional operations provided by an SOS.
-The core operations are used by Data Consumers to request sensor information.
-The operations perfom the following function:
+The diagram above shows the core and transactional operations provided by an SOS.
+The core operations are GetCapabilities, DescribeSensor and GetObservations.
+These are mandatory operations used by Data Consumers to request sensor information.
+The core operations perform the following function:
  \begin{itemize}
-\item GetObservation operation is used to obtain sensor observations and measurement data.
-\item GetCapabilities operation is used to retrieve SOS service metadata.
-\item DescribeSensor operation retrieves detailed information about the sensors.
+\item GetCapabilities  -   operation to retrieve SOS observation offering information.
+\item DescribeSensor - operation to retrieve detailed information about the sensors.
+\item GetObservation -   operation to obtain sensor observations.
 \end{itemize}
 
-The transactional operations supported by the SOS toward the DataProvider are:
+The transactional operations are RegisterSensor and InsertObservation used by the Data Producer to provided sensor information and observations.
+The transactional operations supported are described in the following:
  \begin{itemize}
-\item RegisterSensor operation registers the DataProvider with the SOS.
-\item InsertObservation operation updates SOS service metadata.
+\item RegisterSensor - operation to register the Data Provider with the SOS.
+\item InsertObservation - operation to update SOS with a sensor observation.
 \end{itemize}
 
-Collections of related sensor system observations are organized by the SOS into observation offerings. Observations are encoded as OGC O\&M Observations.
-
-These operations are encoded in XML and used in http post and get operations. The XML message needs to be parsed to identify the operations, parameters and results.
-
-Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
-para.
+These operations are encoded in XML and are intended to be carried http  operations.
+The sensor observations contained in  GetObservation and InsertObservation operations are encoded according to the OGC O\&M Observations \cite{OMref}.
+Sensor systems metadata contained in DescribeSensor and RegisterSensor is defined in SensorML or TML.
 
@@
 The GetCapabilites are then used toward each of the SOS instances returned to find sensor systems that are of interest to the Data Consumer.
-The CS-W catalog is covered seperate specification \cite{OGCcatref} and is not covered in any real detail in the SOS specification.
+The CS-W catalog is covered in a separate specification \cite{OGCcatref}.
+This not covered in any real detail in the SOS specification.
 The OGC Catalog Service is not explored in this thesis.
@@
 DescribeObservationType, and DescribeResultModel. These operations are not covered in this thesis.
+
+The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
+The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
 
@@
 The Processor could be a PDA or small PC or any processing device that has a USB connection.
-The Sensor Board has a USB connection and this is used for communication with the Processor Unit.
+The Sensor Board has a USB connection used for communication with the Processor Unit.
 There are over a dozen sensors on the SenseTile Sensor Board including a thermistor and light sensor.
@@
 
-\section{UCD Sensor Board Driver}
+\section{UCD Sensor Board Driver}\label{SensorBoardDriverSec}
 
-\section{VAST SensorML Engine}
+\section{VAST SensorML Engine}\label{VastSensorMLEngineSec}
 The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source Java based SensorML Processing Engine.
@@
 \chapter{Design}
-\section{SensorML Model of SenseTile}
+\section{SensorML Model of SenseTile}\label{SenseTileModelSec}
 
-A SensorML model of SenseTile is developed as part of this thesis.
-The model contains metadata about the SenseTile system and supports the needs of the SenseTile Web Service.
-Two main requirements of the SenseTile Web Service were to be supported by the SensorML model.
-Firstly the raw data from the senors is to be retrievable to allow external system process the data as desired.
-Secondly there is a requirement to do post processing on the Processor Unit of the sensor data to convert it to meaningful values.
-A block diagram of the Sensetile SensorML model is shown in figure \ref{fig:SenseTileSystem}.
+A SensorML model of SenseTile Sensor Board and its sensors is developed as part of this thesis.
+The model of the Sensor Board is placed in a context to examine SensorML process chaining where processing of Sensor Board data is performed in a separate SensorML process.
+The SensorML Systems and Components that are used to model the system are shown in a  block diagram in figure \ref{fig:SensorML_SenseTile_System_comp}.
 
-\ref{fig:SensorML_SenseTile_System_comp}. Only the Thermistor Component and the Celcius Component are developed as part of this thesis. They are described in model detail in later sections.
 \begin{figure}[h]
@@
 
-Three SensorML Systems are used to model the SenseTile as shown in the block diagram.
-One SensorML System models the overall SenseTile which contains the Sensor Board and Conversion Systems.
-The below XML fragment shows how SensorML describes System containment using a component list and how the other SensorML Systems descriptions are referenced using xlink hrefs.
+In the block diagram there are three SensorML Systems, the SenseTile System, the Sensor Board System and the Conversion System.
+The SenseTile System is used as container for the other two systems linking them in a process chain.
+The Sensor Board system contains metadata about the SenseTile Sensor Board and is a container for SensorML Components that model the Sensor Board's sensors.
+The block diagram shows the physical phenomena measured by the sensors being feed as inputs to the Components in the Sensor Board System.
+The SensorBoard System outputs digital numbers based on the sensors measurement that have no physical meaning yet.
+The SenseTile System outputs the digital numbers directly and also connects to the input of the Conversion System.
+The digital numbers from the sensors are fed into the relevant conversion process that produce a meaningful value such as Celsius of Lumen.
+This allows the SenseTile Web Service provide the raw data from the sensors or the data converted to meaningful values to clients.
 
+Only the Thermistor Component and the Celsius Component are developed as part of this thesis. However the same approach can applied to the other senors.
+
+Why use this structure? eeicpr
+
+A description of the Systems and Components used in the SenseTile are provided in the following sections.
+
+\subsection{SenseTile System}
+The SenseTile System is a SensorML System as described in section \ref{SMLsection} and is a container for the Sensor Board and Conversion Systems connecting their inputs and outputs as appropriate.
+The below XML fragment shows how SensorML encodes the containment of these Systems.
 \lstset{language=XML,basicstyle=\scriptsize,frame=single}
@@
 \end{lstlisting}
+The components and ComponentList elements contain a list of component tags.
+These component elements name the component which may be another SensorML System  Component.
+The component tag also has an XLink href used to reference the location of the SensorML definition of the component.
+In this case the SensorML descriptions of the SensorBoard System and the Conversion System are in separate external files rather than in-line in the SenseTile System description.
+The use of XLink in SensorML is described in the SensorML section \ref{SMLsection}.
 
-The physical phenomena measured by the sensors are the input to the  SenseTile System.
-These values will be read in from packets received from the Sensor Board.
-The Digital Number output is the raw sensor data which is a digital number without a physical meaning.
-The Converted Measurement output contains the values from post processing the sensor data which have some physical meaning such as temperature in Celcius.
+The SenseTile System connects up the inputs and outputs of the SensorBoard and Conversion Systems using SensorML connections.
+The below SensorML fragment shows how the temperature output from the SensorBoard System is connected to the input of the Conversion System.
 
-The division of the systems is based on the concept that sensor observations are based on at least a sampling process and a conversion process{ref tutorial}.
-The Sensor Board System contains SensorML Components that model the sensors on the board that perform the sampling.
-The SensorML Components contained in the Conversion System model processes that produce meaningful physical measurement values such as temperature and lumens from the sampled data.
-
-The SenseTile System connects up the inputs and outputs of the SensorBoard and Convertor Systems using SensorML connections.
-The below SensorML fragment shows how the temperature output from the SensorBoard System is connected to the input of the Conversion System is peformed in SensorML.
-\newpage
 \begin{lstlisting}
@@
 
-Naming rules for inputs and outputs - allow more generic handling of implementation classes.
+The connection name attribute is used here to give a meaningful name to the connection.
+The Link type, as described in \ref{SMLsection}, is encoded with the Link tag.
+The contained source and destination elements contain a ref that indicates the components output and input to use.
+Only one connection is shown here but it is contained in a elements that allow a sequence of connection elements.
 
-System and components are in several files. Inline in one file if desired.Choice to seperate files.
+\subsection{SensorBoard System}
 
@@
 This is an indivisible process and can have a different location information than the Sensor Board if required.
-The SensorML Component allows the capabilities of the Thermsistor to be described.
+The SensorML Component allows the capabilities of the Thermistor to be described.
 For example the TMP175 is  specified for operation over a temperature range of −40°C to +125°C.
@@
 
-\subsection{Celcius Converter Component}
+\subsection{Converter System}
+\subsection{Celsius Converter Component}
 
-The SensoML Celcius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celcius value.
+The SensoML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celsius value.
 It similar to the Thermistor SensorML description but its inputs and outputs are different.
-It process method performs a simple calcution to perform the celcius conversion.
+It process method performs a simple calculation to perform the Celsius conversion.
 
 A Sensor process method is also developed.
-The process method describes the coversion algorithm and points to an implementation class.
+The process method describes the conversion algorithm and points to an implementation class.
 Process Methods are not implemented by the Vast Lib so it not used in the running system.
@@
 \framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
-\caption{SensorML Conceptual Model}\label{fig:InsertObs}
+\caption{Generate Observation}\label{fig:InsertObs}
+\end{figure}
+
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
+\caption{Find SOS sensor}\label{fig:GetMeta}
 \end{figure}
@@
 \framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
-\caption{SensorML Conceptual Model}\label{fig:GetObs}
+\caption{Get Observation from Sensor}\label{fig:GetObs}
 \end{figure}
@@
 This prototype was designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
-A client would access multiple SOS instances to get observations from the sensor network.
-An SOS could support several DataProducers and hence would not to be run on every SenseTile.
+A client accesses multiple SOS instances to get observations from the sensor network.
+An SOS can support several DataProducers and hence would not to be run on every SenseTile Unit.
 
@@
 \end{figure}
-The SOS Specification envisioned that the SOS would be run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}.
-The DataProviders would update this external server with observations and clients would just access this large SOS.
+The SOS Specification envisioned that the SOS is run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}.
+The DataProducers update this external server with observations and SOS Data Consumer access this large SOS.
 \begin{figure}[h]
@@
 
-Scalability - info here.
-
 \newpage
-\subsection{DataProducer}
+\subsection{DataProducer}\label{DataProducerHigh}
 The DataProducer's role, as described previously, is to read sensor data from the Sensor Board and generate O\&M observations.
@@
 The SMLEngine cluster contains classes that provide access to
-the Vast Lib that is providing the SensorML Engine.  The Vast Lib implemenation
+the Vast Lib that is providing the SensorML Engine.  The Vast Lib implementation
 of the SensorML Engine is hidden from the rest of the code thus.
@@
 
-Sensors iimplement interface to facilitate this.
+Sensors implement interface to facilitate this.
 
@@
 \subsection {SOS}
+
+The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the observations from the DataProducer.
+The below is a static diagram showing the important clusters and classes in the SOS design.
+
  \begin{figure}[h!]
@@
 \end{figure}
-Keeps track of sensorboards that register
+The SOS is the main cluster and the Service cluster contains the classes that provide the Web Service interface.
 
-provides an RMI interface to the DataProducer to update with observations. This has lower overhead
-than a web service. Though it would not be hard to change to use a Web Service Interface.
+The DataProvider maintains a list of observation offerings provided by the SOS using the ObservationOffering class.
+When a new sensor type is registered a new ObservationOffering is created.
+If a DataProvider registers and provides an offering that already exists then just a new Sensor instance is added to the existing ObservationOffering.
 
-GetObservaton - marshal Observation to xml string and send as response.
+The Sensor class contains a list of Observations that hold the observations from the DataProvider. The last 50 observations are maintained.
 
-StreamObservation -  registered clients get new observation sent over UDP connection?
+The ObservationService class provides the interface for the DataProducer to update with observations.
 
-Provides a Web Service interface  based on Sensor Observation Service. Not fully as is a very complex interface.
+The SenseTile Service provides Web Service interface to be used by clients to access the sensor observations.
+
+The SOS specification provides a very rich filtering feature that is too complex for this prototype.
+The operations are developed in a way to allow the sensor metadata and observations to be retrieved for a specific sensor board.
+The latest observation or all the stored observations can be retrieved.
 
 \chapter{Implementation}
+\section{Overview}
+The SenseTile Web Service is written in Java.
+As described in the high level design there are two components, the DataProducer and the SOS, that are developed to provide this service.
+The two components are run in separate processes.
+The communication between the DataProducer and the SOS uses Java RMI.
+RMI is used due to the SOS and DataProducer
+processes are running in the SenseTile Network and communicate over a LAN.
+This communication protocol has lower overhead than a Web Service so efficient communication between the DataProducer and SOS is achieved.
+Though it would not be hard to change to use a Web Service as a Java Interface is used to hide the communication mechanism from the main application code.
+Clients, SOS DataConsumers, access the SOS using a SOAP based Web Service interface.
 
+The SOS specification describes an implementation using HTTP carry complete operation information encoded in XML including the operations name and parameters.
+The SenseTile WebService takes a different approach.
+The operation is defined in the RMI interface and the Web Service Interface while the parameters are encoded in XML thus need parsing.
+
+The following sections describe the DataProducer and SOS implementation.
 \section{DataProducer}
+The DataProducer, described in section \ref{DataProducerHigh}, accesses the sensor board data and inserts observations into the SOS.
+Two libraries are used to achieve this: the SensorBoardDriver and the VastSensorMLEngine.
 
-DataProducer VastLib UCD SenseTile Driver Lib
+The SensorBoardDriver is described previously in section \ref{SensorBoardDriverSec}.
+There is also a SensorBoard Simulator library developed by UCS CASL researchers.
+It produces sensor board packets that can be feed into the DataProvider.
+This was used during the development of the DataProvider to facilitate testing.
+
+The VastSensorMLEngine, described in section \ref{VastSensorMLEngineSec}, is used in conjunction with the SenseTile SensorML Model (section \ref{SenseTileModelSec}).
+The SensorML is parsed by the VastSensorMLEngine into a DOM tree.
+To perform the processing it propagates an execute method to the instantiated processes.
+A ProcessMap.xml file maps SensorML ProcessModel and Component method  URNs  (see section \ref{SMLsection}) to Java classes that provide the implementation of the processing to be performed.
+
+There is no documentation that I can find for the VastSensorMLEngine.
+There are some reference implementations of SWE services pointed to from the OGC SWE website (http://www.ogcnetwork.net/SWE) that provide some information how to use it.
+The SenseTile SensorML model was not parsed initially by the VastSensorMLEngine.
+As it opensource the code is available and after a few small changes the model parsed.
+The VastSensorMLEngine processes the model fine and was very stable during testing.
 
 \section{SOS}
- is based on AXIS 2.0/Spring
+The SOS is the provider of the Web Service interface and is built on Apache Axis2 (http://ws.apache.org/axis2/). 
+Axis2 is described as a "Web Services / SOAP / WSDL engine"  with the Apache Axis2/Java implementation used in this thesis.
+The Spring framework (http://www.springsource.org/) was also used.
+Using these libraries a light Web Service is built to run on the reasonably powerful SenseTile Processor Unit.
+Axis2 is run standalone here though it can be on with any Servlet engine if a more powerful Web Service is needed.
 
-JAXB for SensorML XML to java binding.
+Developing a Web Service is straight forward with Axis2 and it supports several approaches.
+The supported POJOs (Plain Old Java Objects) approach was used for the SenseTile Web Service as straight forward and the code could be reused on a different framework if AXIS2 did not work well.
+The Spring framework, which Axis2 supports, was also used for it's dependency injection functionality.
+A SprinitInit class is needed for Axis to load up the framework.
 
-RMI used between DataProvider and SOS.
+Axis2 a tool that automatically generates WSDL from the running Web Service. Another tool allows client side code to be generated for the WSDL easing the development of the Web Service clients.
 
-\section{SensorML Generation}
+JAXB (http://java.sun.com/developer/technicalArticles/WebServices/jaxb) was used for SensorML XML to Java binding in the Web Service.
+JAXB is able to generate a default binding from the complex SWE schema.
+Some name collisions were corrected with a JAXB customization file found on the SensorML newsgroups.
 
@@
 
-running prototype tested with UCD SensorBoard Simulator code
+A  SensorML Model description of a SenseTile Sensor Board and Thermistor was developed. It was
+run on a SensorML engine. Appendix shows the model developed.
 
-Tested against real SenseTile and sensor data retrieved by a testclient.
+Developed lighweight Web Service to explore the use of SensorML in Sensor Networks.
+running prototype tested with UCD SensorBoard Simulator code.
 
-Generation of SensorML description of a Senstile Sensor
+Tested on a low power Atom PC that is used as SenseTile Processing Unit.
+
+BON mapping developed for main SensorML types. A BON description of SenseTile SensorML Model
+was developed based on the mapping.
 
@@
 Why?
+
+Difficult to get connections correct. Excute the model to find problems.
+Need static checking.
+Definitely need a tool to generate SensorML. BON would
+help even if metadata not covered. Could be add by
+style sheet.
 
@@
 
-Streaming the data.
+pointed to implenentation would need to be developed for a specific
+SensorML engine.
 
-\chapter{References}
+further work.
+Complete BON and parser.
+Streaming the data. See Arch doc.
+
 \newpage

### 2010-03-25 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 The sensors are interfaced or hosted on the Sensor Board.
-The Processor Unit interfaces to the Sensor Board allowing data from the Sensor Board to be processed futher as required.
+The Processor Unit interfaces to the Sensor Board allowing data from the Sensor Board to be processed further as required.
 
@@
 
-There are othe  SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref]. This specification provides standard models and XML encoding for sensor observations.
+There are other SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref]. This specification provides standard models and XML encoding for sensor observations.
 
@@
 \section{BON to SensorML mapping}
-Features to be elements or attributes.
-
-Name rule suffix "Attrib" to make it an attribute.
-
-List handling - convention for mapping.
-
-SML prefix to handled key word clashes between sensorML and BON.
 
@@
 In the SenseTile Web Service the SOS acts as a gateway to the SenseTile Sensor Network.
-This prototype was designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
+This prototype is designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
 A client accesses multiple SOS instances to get observations from the sensor network.
@@
 \end{figure}
+
 The SOS Specification describes the use of a Proxy SOS as shown in figure \ref{fig:Deployment_sos_aggr}.
@@
 
-Questions.
-If have light data but no temp. As temp is only
-every 1000 packets?
-
-How to know when to generate observation
-for the two offerings?
-DataProducer generates observations in Observation xml standard.
-
-Interface to the SensorBoard Library.
-
-Uses observer pattern to update sensors with packets from the board.
-
-reads sensors from SensorBoard System xml
-
-Sensors implement interface to facilitate this.
-
-Vast lib connects all the processes
-
-implementation of sensor and convertor is named using URN read from the sensorMl descption.
-
 The DataProducer was developed as a SensorML based software.
@@
 \end{figure}
-
-Interfaces to the Sensor Board
 
@@
 \chapter{ Conclusions and Future Work}
-There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
+This thesis analyses the role SensorML plays in Sensor Web Networks.
+The UCD CASL SenseTile System is used a case study with the development of the SenseTile Web Service to access the SenseTile System sensor data.
+SensorML alone is not sufficient to model all aspects of Sensor Web Networks but is focused on providing a standard way to exchange sensor metadata, describing sensor input/output data and the processing of sensor data.
+This leads to examining other OGC SWE specifications, of which SensorML is one.
+The SWE SOS specification describes the operations for accessing sensor observations and it's concepts were used to develop the SenseTile Web Service.
+It uses SensorML provide sensor metadata to clients and the O\&M specification to encode observations.
+The OGC SWE initiative is trying to standardize a Web Service approach to accessing and processing sensor data.
+
+SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.
+
+The SenseTile model uses a hierarachy to model the SenseTile.
+It seems repetitive to have keep defining the same inputs for each contained component.
+Maybe this is a flaw in the SenseTile model developed.
+
+Working in XML is error prone.
+SensorML connections type was a good example of where a graphical tool or another language, like BON,  that can be parsed should be used to check the connections.
+Running the model through the VastSensorMLEngine without a crash was only way other than by eye to see that the connections were set up correctly.
+
+SensorML is defined by a heavy XML Schema.
+This XML Schema along with the other XML Schema for the SOS and dependent specifications are heavy indeed.
+Achieving a Java binding was difficult.
+JIBX would not generate a default XML binding.
+JAXB could with some help.
+There are many layers of abstraction in these XML Schema.
+This reflects the that the specifications cover a very wide range of needs of a senor system.
+
+The idea of automatically obtaining a SensoML model and executing it depends on the SensorML Engine being used.
+Only one Java implementation can be pointed.
+In the OGC examples the Java implementations of the
+
+Vast Lib use as a references only.
+Had to update library code to get it to work with the SenseTile schema even though the schema
+passed the validation test tool from the site. It uses DOM but XML binding is probably a better approach.
+
+the weaknesses of your approach
+-----------------------------------------
+SOS on the Processor Board. Is this right? RMI only useful if SenseTile Network is a lan.
+
+SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
+flexible. There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
 what has been achieved
 
-the weaknesses of your approach
-SOS on the Processor Board. Is this right.
-
-difficult schemas to program against.
-JIBX did not generate a default binding.
-Jaxb did with some help but did not generate the process type correctly.
-Why?
-
-Difficult to get connections correct. Excute the model to find problems.
-Need static checking.
-Definitely need a tool to generate SensorML. BON would
-help even if metadata not covered. Could be add by
-style sheet.
-
-Schemas cover all needs for sensor description and data.
-
-A lot of schemas needed.
-
-SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
-flexible.
-
-Vast Lib use as a references only. Other ways such a XQUERY? XML techs?
-As there are a large number of types etc. A quick way for a prototype.
-Had to update library code to get it to work with the SenseTile Schemas. Schemas
-passed the validation test tool from the site.
-
-Portabiltiy, Vast Types POJOs.
-
-pointed to implenentation would need to be developed for a specific
-SensorML engine.
-
 further work.
-Complete BON and parser.
-Streaming the data. See Arch doc.
+---------------
+Complete BON mapping and parser.
+Streaming the data. See SWE Architecture dpcument for a description of one approach.
+Other technology approaches such as . Other ways such a XQUERY? XML techs?
 

### 2010-03-25 papers/SenseTileOverview/paper.tex
updates

--- 2009-06-29 papers/SenseTileOverview/paper.tex
@@
             colorlinks=true,pdfstartview=FitV,
-            linkcolor=blue,citecolor=blue,urlcolor=blue,
-            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
+            linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
 \else
@@
 
-\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
-UCD CASL: Complex and Adaptive Systems Laboratory and\\
-School of Computer Science and Informatics,\\
-University College Dublin,\\
-Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
-}
+\author{\ldots}
 
@@
 
-\todo{\emph{We probably have only 8 pages, in the IEEE format, if this is going to
-\myhref{http://ieeexplore.ieee.org/xpl/RecentIssue.jsp?punumber=7756}{IEEE Pervasive Computing}}.}
+The \textbf{\ST System} is a large-scale, general-purpose
+sensor system installed at the University College Dublin in Dublin,
+Ireland. 
+Our facility provides services for performing in-depth
+investigation of aspects and issues related to large-scale sensor 
+networking.
 
-An abstract.
+This system integrates heterogenous hardware and software systems. 
+The hardware systems are composed by a sensors layer (1000+ sensors), a gateway 
+layer (100+ minipc), an High-performance computing cluster (50+ cores) and a datastore (100+ TB). 
+The software systems integrated in the platform comprehend: (1) a set of
+drivers to communicate with the severeal sensors systems deployed, (2)
+a distributed sensor data stream management to enable data streaming between 
+the processor nodes and to the datastore, (3) a data provenance and transformation
+distributed service to enable the search and inspection of the meta data associated to the sensor streams, 
+(4) a processor nodes monitoring and management system to ease deployment and mass 
+installation of software and configuration on processor nodes, (5) a scientific platform
+to process and inspect the data collected, batch or interactive computation are possible, using stand alone 
+servers or using the HPC dedicated cluster.
+
+The \ST Sensor Board is part of the sensors layer, and is a custom-designed, but inexpensive,
+sensor board, paired directly with a processor node through an USB channel.
+
+The \ST system exhibits the following novel aspects in sensor management: 
+(1) multi layer architecture, to keep separate sensors, gateways, HPC processing cluster and the datastore;
+(2) integrated management of a large number of gateways collecting data from several heterogenous sources;
+(3) the integration of several heterogeneous software systems, to support sensor, data streams, gateways and 
+datastore management, and scientifical experiments execution.
 
@@
 
-%======================================================================
 \Section{Introduction}
 
-An introduction.
+Sensor networks are widely seen as an essential technology for supporting next-generation scientific and engineering
+challenges, including environmental monitoring, climate change, assisted living, national security and intensive
+agriculture. 
+Addressing these problems requires two complementary strands of research:
+(1) understand the specific physical and social phenomena of interest against which to fit data collection and analysis;
+(2) understand the general scientific and technological principles governing scalable sensor networking in order to 
+maximise the leverage gained from on-going developments.
+
+There are several technological challanges to be faced to solve these problems: (1) the collection, storage and management 
+of large volumes of multimedia data and the meta data associated; (2) the interactive and offline setup and monitoring of
+the sensors; (3) the interactive and offline setup of live data collection experiments (data streaming, and soft real time data 
+processing); (4) the post processing analysis and visualization of the data collected.
+
+A vertical approach in sensor data collection and analysis, which is the nowadays dominant approach, is not 
+able to grasp the general problem.
+Each research community has its own requirements and constraints on data collection and analysis, and considering the 
+problem only from a single perspective would limit the usefulness of the tools developed, the data collected and the data analysis.
+When the problems addressed are too specific, too focused on a single discipline, the solutions found are hardly 
+reusable in different areas.
+Thus, the sensor management platforms and the data repositories designed and developed, and, most importantly, the data 
+collected are ineffective and inappropriate, if not completely unusable and useless, when they are 
+tried to be exported and reused in even slightly dissimilar domains from the original one.
+
+\TODO{provide examples, provide references}
+
+Merging requirements from heterogeneous disciplines is a challanging task; not only because different domains generates 
+different languages creating a de facto barrier in communication and reciprocal understanding \TODO{provide references}, but 
+also because big tradeoffs are needed to be made to accomodate most of the requirements. 
+The tradeoff, to be sensible, needs to be chosen directly by the involved research communities and domain experts, and 
+this is can be solved only involving them directly in cross-disciplinary projects.
+
+Sensor data collection and analysis, to be fully understood and exploited, and to make the solutions envisioned general 
+enough to be reusable, need to be studied and addressed in cross-disciplinary projects involving different analysis 
+methodologies and constraints. 
+Cross-disciplinary projects are irremissible in the development of general solutions and tools to sensor data acquisition, 
+storage and analysis.
+
+heterogeneous domains involved ->
+some domains are far away from computer science ->
+ease of use of tools, techniques, data ->
+increased probability of tools, techniques, data to be succesfully used
+
+\begin{itemize}
+  \setlength{\itemsep}{1pt}
+  \setlength{\parskip}{0pt}
+  \setlength{\parsep}{0pt}
+\item dealing with uncertainty in sensor systems,
+\item robust management of large (multi-terabyte) sensing data on
+  heterogeneous platforms,
+\item interaction and activity recognition,
+\item sparse signal analysis, and
+\item coding and securing sensor interactions.
+\end{itemize}
 

### 2010-03-28 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \chapter{Introduction}
+  - introduce problems: 
+   - sensor interoperability
+   - reliable sensor processing
+  - SensorML proposed standard for sensor description and interoperability
+  - SOS proposed standard for sensor data retrieval
+  - no formal definition of sensor and sensor data processing
+  - introduce SenseTile, a real case study
+  - objective
+   - describe SenseTile with SensorML, access data through SOS
+    - assess SensorML and SOS
+   - formalize sensor description and process
+    - SensorML <-> BON <-> Java + JML
 
-The aim of this thesis project was to model the SenseTile Sensor Board using Sensor Model Language (SensorML)\cite{SMLref} .
+There is currently very little interoperability between different sensor systems making it difficult to access sensor observations.
+The ability to easily discover, access, and process sensors existing within different communities is very rare.Interoperability The  challenges for Sensor Networks include sensor interoperability and reliable sensor processing.
+These issues of sensor system interopetability and senor data processing are being address through the development of standards.
+
+Sensor Model Language (SensorML)\cite{SMLref} is one such proposed standard for sensor description and interoperability.
+
+SensorML enables this processing without the need for sensor-specific software.
+
+SensorML is used to describe sensor systems and the processing of sensor observations.
+SensorML provides a common way to describe sensor systems.
+SensorML is one of a number of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
+The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
+
+SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
+It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
+
+ .
 A prototype of a Web Service for accessing SenseTile Sensor Board sensor observations was to be developed using this SensorML description of the SenseTile Sensor Board.
@@
 
-SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
-It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
-The SenseTile System is made up of two units a Sensor Board and Processor Unit.
-The sensors are interfaced or hosted on the Sensor Board.
-The Processor Unit interfaces to the Sensor Board allowing data from the Sensor Board to be processed further as required.
-
-SensorML is one of a number of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
-The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
-SensorML is used to describe sensor systems and the processing of sensor observations.
-It provides standard models and XML encodings to describe the process of sensor measurement.
-SensorML is described further in chapter x.
-
-The SenseTile Web Service prototype allows access to the Sensor Board sensor observations and it's SensorML description.
-It runs on the SenseTile Processor Unit.
  The Web Service is based on another OGC SWE  specification, the Sensor Observation Service (SOS)[ref].
 This specification defines a standard Web interface for retrieving information about sensor systems and sensor observations.
-Sensor Board measurements that have been post processed and converted to real values or the raw Sensor Board data can be accessed using the Web Service.
-SOS Lite.
 
@@
 \section{Sensor Web Enablement}
-SensorML is developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative (http://www.opengeospatial.org/projects/groups/sensorweb).
+SensorML is developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
 This initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
@@
 
-SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to con-trol observations.
-For a good overview of SWE and it's use in Sensor Web Network please see \cite{SWEArchref}
-
-Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand.
-The SWE framework seems to cover all that is needed for modeling sensors and observations.
-Support for easy exchange of sensor meta-data and sensor observations is the main rational of SWE and fits the needs of the SenseTile Sensor Network.
+SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations.
+For a good overview of SWE and it's use in Sensor Web Network see \cite{SWEArchref}.
 SensorML, SOS and O\&M specifications are the focus of this thesis and are described in more detail in the following sections.
@@
 \begin{itemize}
-\item support sensor and sensor observation discovery
-\item support processing of the sensor observations
-\item support geolocation information about sensors
-\item provide accuracy information of sensors measurements
-\item Provide an explicit description of the process by which an observation was obtained
-\item Provide an executable process chain for deriving new data products on demand
+\item support sensor and sensor observation discovery.
+\item support processing of the sensor observations.
+\item support geolocation information about sensors.
+\item provide accuracy information of sensors measurements.
+\item provide an explicit description of the process by which an observation was obtained.
+\item provide an executable process chain for deriving new data products on demand.
 \end{itemize}
@@
 
-SensorML uses XML Linking Language (XLink) (http://www.w3.org/TR/xlink) to support hypertext referencing in SensorML XML documents using URNs and URIs.
+SensorML uses XML Linking Language (XLink) \footnote{http://www.w3.org/TR/xlink} to support hypertext referencing in SensorML XML documents using URNs and URIs.
 This allows SensorML models to be distributed over several documents.
@@
 \section{SenseTile}
-A case study of the use of SensorML in sensor networks is based on the SenseTile System. SenseTile is a general purpose sensor system developed at UCD CASL (http://kind.ucd.ie/
-products/opensource/SenseTile).
+A case study of the use of SensorML in sensor networks is based on the SenseTile System. SenseTile is a general purpose sensor system developed at UCD CASL \footnote{http://kind.ucd.ie/
+products/opensource/SenseTile}.
 The SenseTile System is used to investigate issues arising with large scale sensor networking.
 SenseTile is made up of a sensor platform, a datastore, and a compute farm.
+The below figure shows a outline of this system.
+\begin{figure}[h]
+\framebox{\scalebox{0.65}{\includegraphics{SenseTileDescription.png}}}
+\caption{SenseTile System}\label{fig:SenseTileDescription}
+\end{figure}
  The sensor platform is called the SenseTile Unit and is composed of one or more SenseTile Sensor Boards paired with a SenseTile Processor Unit.
@@
 The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing.
-The code is found at http://code.google.com/p/sensorml-data-processing and is referred to as the VastSMLEngine in rest of the thesis.
+The code is found at \footnote{http://code.google.com/p/sensorml-data-processing} and is referred to as the VastSMLEngine in rest of the thesis.
 This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}.
-This code is found at http://code.google.com/p/swe-common-data-framework.
+This code is found at \footnote{http://code.google.com/p/swe-common-data-framework}.
 
 \section{BON}
-
+Business Object Notation (BON) is a method for analysis and design of
+object-oriented systems.
+seamlessness, reversibility and software contracting.
+same same semantic and conceptual base  analysis, design, and implementation by using the
+ for the notation on all three levels. It supports source code changes back into design and analysis.
+The idea is to treat each subprogram as a subcontractor
+who undertakes to deliver some service (specified by a postcondition), but only if
+certain prerequisites are fulfilled (the precondition).
+same concepts and semantics throughout the development of a
+software product allows for a smooth transition from the initial analysis and design models
+down to executable code.
+\cite{BONref}
 \chapter{Design}
+\section{Overview}
+Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand.
+The SWE framework seems to cover all that is needed for modeling sensors and observations.
+Support for easy exchange of sensor meta-data and sensor observations is the main rational of SWE and fits the needs of the SenseTile Sensor Network.
 \section{SensorML Model of SenseTile}\label{SenseTileModelSec}
@@
 There is no documentation that I can find for the VastSensorMLEngine.
-There are some reference implementations of SWE services pointed to from the OGC SWE website (http://www.ogcnetwork.net/SWE) that provide some information how to use it.
+There are some reference implementations of SWE services pointed to from the OGC SWE website \footnote{http://www.ogcnetwork.net/SWE} that provide some information how to use it.
 The SenseTile SensorML model was not parsed initially by the VastSensorMLEngine.
@@
 \section{SOS}
-The SOS is the provider of the Web Service interface and is built on Apache Axis2 (http://ws.apache.org/axis2/). 
+The SOS is the provider of the Web Service interface and is built on Apache Axis2 \footnote{http://ws.apache.org/axis2}. 
 Axis2 is described as a "Web Services / SOAP / WSDL engine"  with the Apache Axis2/Java implementation used in this thesis.
-The Spring framework (http://www.springsource.org/) was also used.
+The Spring framework\footnote{http://www.springsource.org/} was also used.
 Using these libraries a light Web Service is built to run on the reasonably powerful SenseTile Processor Unit.
@@
 
-Axis2 a tool that automatically generates WSDL from the running Web Service. Another tool allows client side code to be generated for the WSDL easing the development of the Web Service clients.
+Axis2 a tool that automatically generates WSDL from the running Web Service. Another tool allows client side code to be generated from the WSDL easing the development of the Web Service clients.
 
-JAXB (http://java.sun.com/developer/technicalArticles/WebServices/jaxb) was used for SensorML XML to Java binding in the Web Service.
+JAXB \footnote{http://java.sun.com/developer/technicalArticles/WebServices/jaxb} was used for SensorML XML to Java binding in the Web Service.
 JAXB is able to generate a default binding from the complex SWE schema.
@@
 \begin{thebibliography}{99}
+\bibitem{OSWAref} Xingchen Chu, Tom Kobialka,  Durnota, and  Buyya, Open Sensor Web Architecture: Core Services
 \bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007

### 2010-04-05 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \chapter{Introduction}
-  - introduce problems: 
-   - sensor interoperability
-   - reliable sensor processing
-  - SensorML proposed standard for sensor description and interoperability
-  - SOS proposed standard for sensor data retrieval
-  - no formal definition of sensor and sensor data processing
-  - introduce SenseTile, a real case study
-  - objective
-   - describe SenseTile with SensorML, access data through SOS
-    - assess SensorML and SOS
-   - formalize sensor description and process
-    - SensorML <-> BON <-> Java + JML
 
-There is currently very little interoperability between different sensor systems making it difficult to access sensor observations.
-The ability to easily discover, access, and process sensors existing within different communities is very rare.Interoperability The  challenges for Sensor Networks include sensor interoperability and reliable sensor processing.
-These issues of sensor system interopetability and senor data processing are being address through the development of standards.
+problems.
 
-Sensor Model Language (SensorML)\cite{SMLref} is one such proposed standard for sensor description and interoperability.
+Sensor Networks are increasingly deployed to provide sensing information for a myriad of applications.
+These Sensor Networks are developed by communities working in different domains and contain many different sensor types.
+The heterogeneity of these systems means that interoperability is difficult and expensive.
+Reliably processing sensor observations from these heterogeneous systems is also issue.
+Need to have a sensor description, data description, observation, retrieval and information on accuracy etc for reliable processing.
+Standards are being developed to address these problems.
+Use internet.
 
-SensorML enables this processing without the need for sensor-specific software.
+Sensor Model Language (SensorML)\cite{SMLref} is one proposed standard for sensor description and processing of sensor observations.
+SensorML specifies models and XML Schema for describing sensors, the process of sensor measurement and the post processing of sensor observations.
 
-SensorML is used to describe sensor systems and the processing of sensor observations.
-SensorML provides a common way to describe sensor systems.
-SensorML is one of a number of specifications provided by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement(SWE) initiative[ref].
-The members of the OGC have the goal of developing open standards to exploit Web connected sensors of all types.
+The Sensor Observation Service (SOS)\cite{SOSref}  specification is a proposed standard for retrieving information about sensor and retrieving sensor observations.
 
-SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensor, audio sensors, pressure sensors, light level sensors video sensors among others.
+The SensorML and SOS specifications are developed by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
+The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as SensorML.
+This thesis investigates the use of these standards in developing sensor systems.
+
+Currently there is no formal definition of sensors or sensor data processing.
+There are now formal verification based software development processes [kiniry] with the goal of developing very robust systems.
+Having formal sensor descriptions allows sensor system development with such a verification based process.
+
+The thesis uses a real sensor system called SenseTile as a case study.
+SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensors, audio sensors, pressure sensors, light level sensors video sensors among others.
 It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
 
- .
-A prototype of a Web Service for accessing SenseTile Sensor Board sensor observations was to be developed using this SensorML description of the SenseTile Sensor Board.
-An evaluation of the suitability of SensorML for use as part of the SenseTile System was also to be performed.
+This thesis describes SenseTile with SensorML.
+Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the sensor SensorML descriptions and sensor observations to be retrieved.
+The Web Service will be based on the SOS specification.
+An assessment of SensorML and SOS is provided based on this work.
 
- The Web Service is based on another OGC SWE  specification, the Sensor Observation Service (SOS)[ref].
-This specification defines a standard Web interface for retrieving information about sensor systems and sensor observations.
-
-There are other SOS specification depend on another SWE specification, Observation and Measurements(O\&M)[ref]. This specification provides standard models and XML encoding for sensor observations.
-
-A  SensorML to BON Mapping was also to be explored .
-BON\cite{BONref} is a notation and a method for Object Oriented systems analysis and design.
-Based on this mapping it was proposed to develop a tool/method to automatically generate SensorML descriptions.
-Generally SensorML is hand generated with all the shortcomings of such an approach.
+To formalize sensor description and process BON\cite{BONref} is a notation and a method for Object Oriented systems analysis and design.
+A BON to SensorML mapping is provided.
 The proposed approach would allow sensor processing to defined formally in BON and then refined to JML/Java.
@@
 \section{BON to SensorML mapping}
+\lstset{frame=none}
+\begin{table}[!th]
+\begin{tabular}{|l|r|}
+\hline
+BON & SensorML\\
+\hline
+   deferred class AbstractProcess  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & not mapped\\
+\hline
+     parameters:Parameters & not mapped\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup & not mapped\\
+ \hline
+
+\end{tabular}
+\caption{Scenario Chart for SenseTile Web Service}
+\label{ex:table}
+\end{table}
 
@@
 \begin{thebibliography}{99}
-\bibitem{OSWAref} Xingchen Chu, Tom Kobialka,  Durnota, and  Buyya, Open Sensor Web Architecture: Core Services
+\bibitem{OSWAref} Chu, Kobialka,  Durnota, and  Buyya, Open Sensor Web Architecture: Core Services
 \bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
@@
 \chapter{Appendices}
+\section{SenseTile SensorML Model}
 
 \lstinputlisting{Thermistor.xml}
+
+\section{BON to SensorML Mapping}
+\lstset{basicstyle=\scriptsize,showspaces=false,showstringspaces=false}
+\lstinputlisting{SensorML_formal.bon}
 

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/ACurveElement.java
updates and JAXB binding

@@ -1,0 +1,28 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.namespace.QName;
+
+public class ACurveElement
+    extends JAXBElement<CurveType>
+{
+
+    protected final static QName NAME = new QName("http://www.opengis.net/gml", "Curve");
+
+    public ACurveElement(CurveType value) {
+        super(NAME, ((Class) CurveType.class), null, value);
+    }
+
+    public ACurveElement() {
+        super(NAME, ((Class) CurveType.class), null, null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/AbstractTimeComplexType.java

@@ -1,0 +1,49 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlSeeAlso;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.swe._1_0.TimeGeometricComplexType;
+import net.opengis.swe._1_0.TimeGridType;
+
+
+/**
+ * The abstract supertype for temporal complexes.
+ * 
+ * <p>Java class for AbstractTimeComplexType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="AbstractTimeComplexType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractTimeObjectType">
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "AbstractTimeComplexType")
+@XmlSeeAlso({
+    TimeTopologyComplexType.class,
+    TimeGridType.class,
+    TimeGeometricComplexType.class
+})
+public abstract class AbstractTimeComplexType
+    extends AbstractTimeObjectType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/AbstractTimeObjectType.java

@@ -1,0 +1,48 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlSeeAlso;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.swe._1_0.TimeAggregateType;
+
+
+/**
+ * The abstract supertype for temporal objects.
+ * 
+ * <p>Java class for AbstractTimeObjectType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="AbstractTimeObjectType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractGMLType">
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "AbstractTimeObjectType")
+@XmlSeeAlso({
+    AbstractTimePrimitiveType.class,
+    AbstractTimeComplexType.class,
+    TimeAggregateType.class
+})
+public abstract class AbstractTimeObjectType
+    extends AbstractGMLType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/CodeListType.java

@@ -1,0 +1,120 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.XmlValue;
+
+
+/**
+ * List of values on a uniform nominal scale.  List of text tokens.   
+ *       In a list context a token should not include any spaces, so xsd:Name is used instead of xsd:string.   
+ *       If a codeSpace attribute is present, then its value is a reference to 
+ *       a Reference System for the value, a dictionary or code list.
+ * 
+ * <p>Java class for CodeListType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="CodeListType">
+ *   &lt;simpleContent>
+ *     &lt;extension base="&lt;http://www.opengis.net/gml>NameList">
+ *       &lt;attribute name="codeSpace" type="{http://www.w3.org/2001/XMLSchema}anyURI" />
+ *     &lt;/extension>
+ *   &lt;/simpleContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "CodeListType", propOrder = {
+    "value"
+})
+public class CodeListType {
+
+    @XmlValue
+    protected List<String> value;
+    @XmlAttribute
+    @XmlSchemaType(name = "anyURI")
+    protected String codeSpace;
+
+    /**
+     * XML List based on XML Schema Name type.  An element of this type contains a space-separated list of Name values Gets the value of the value property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/CompositeCurvePropertyType.java

@@ -1,0 +1,296 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for CompositeCurvePropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="CompositeCurvePropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element ref="{http://www.opengis.net/gml}CompositeCurve"/>
+ *       &lt;/sequence>
+ *       &lt;attGroup ref="{http://www.opengis.net/gml}AssociationAttributeGroup"/>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "CompositeCurvePropertyType", propOrder = {
+    "compositeCurve"
+})
+public class CompositeCurvePropertyType {
+
+    @XmlElement(name = "CompositeCurve")
+    protected CompositeCurveType compositeCurve;
+    @XmlAttribute(namespace = "http://www.opengis.net/gml")
+    @XmlSchemaType(name = "anyURI")
+    protected String remoteSchema;
+    /**
+     * 
+     * 
+     */
+    @XmlAttribute(namespace = "http://www.w3.org/1999/xlink")
+    public final static String TYPE = "simple";
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/CovarianceElementType.java

@@ -1,0 +1,134 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.math.BigInteger;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * An element of a covariance matrix.
+ * 
+ * <p>Java class for CovarianceElementType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="CovarianceElementType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/gml}rowIndex"/>
+ *         &lt;element ref="{http://www.opengis.net/gml}columnIndex"/>
+ *         &lt;element ref="{http://www.opengis.net/gml}covariance"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "CovarianceElementType", propOrder = {
+    "rowIndex",
+    "columnIndex",
+    "covariance"
+})
+public class CovarianceElementType {
+
+    @XmlElement(required = true)
+    @XmlSchemaType(name = "positiveInteger")
+    protected BigInteger rowIndex;
+    @XmlElement(required = true)
+    @XmlSchemaType(name = "positiveInteger")
+    protected BigInteger columnIndex;
+    protected double covariance;
+
+    /**
+     * Gets the value of the rowIndex property.
+     * 
+     * @return
+     *     possible object is
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/CylindricalCSType.java

@@ -1,0 +1,41 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * A three-dimensional coordinate system consisting of a polar coordinate system extended by a straight coordinate axis perpendicular to the plane spanned by the polar coordinate system. A CylindricalCS shall have three usesAxis associations. 
+ * 
+ * <p>Java class for CylindricalCSType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="CylindricalCSType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractCoordinateSystemType">
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "CylindricalCSType")
+public class CylindricalCSType
+    extends AbstractCoordinateSystemType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/DerivedCRSRefType.java

@@ -1,0 +1,298 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Association to a non-projected derived coordinate reference system, either referencing or containing the definition of that reference system. 
+ * 
+ * <p>Java class for DerivedCRSRefType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="DerivedCRSRefType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element ref="{http://www.opengis.net/gml}DerivedCRS"/>
+ *       &lt;/sequence>
+ *       &lt;attGroup ref="{http://www.opengis.net/gml}AssociationAttributeGroup"/>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "DerivedCRSRefType", propOrder = {
+    "derivedCRS"
+})
+public class DerivedCRSRefType {
+
+    @XmlElement(name = "DerivedCRS")
+    protected DerivedCRSType derivedCRS;
+    @XmlAttribute(namespace = "http://www.opengis.net/gml")
+    @XmlSchemaType(name = "anyURI")
+    protected String remoteSchema;
+    /**
+     * 
+     * 
+     */
+    @XmlAttribute(namespace = "http://www.w3.org/1999/xlink")
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/DrawingTypeType.java

@@ -1,0 +1,45 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlEnum;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for DrawingTypeType.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * <p>
+ * <pre>
+ * &lt;simpleType name="DrawingTypeType">
+ *   &lt;restriction base="{http://www.w3.org/2001/XMLSchema}string">
+ *     &lt;enumeration value="POLYLINE"/>
+ *     &lt;enumeration value="ORTHOGONAL"/>
+ *   &lt;/restriction>
+ * &lt;/simpleType>
+ * </pre>
+ * 
+ */
+@XmlType(name = "DrawingTypeType")
+@XmlEnum
+public enum DrawingTypeType {
+
+    POLYLINE,
+    ORTHOGONAL;
+
+    public String value() {
+        return name();
+    }
+
+    public static DrawingTypeType fromValue(String v) {
+        return valueOf(v);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/EllipsoidBaseType.java

@@ -1,0 +1,50 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlSeeAlso;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Basic encoding for ellipsoid objects, simplifying and restricting the DefinitionType as needed. 
+ * 
+ * <p>Java class for EllipsoidBaseType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="EllipsoidBaseType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.opengis.net/gml}DefinitionType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/gml}metaDataProperty" maxOccurs="unbounded" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/gml}ellipsoidName"/>
+ *       &lt;/sequence>
+ *       &lt;attribute ref="{http://www.opengis.net/gml}id use="required""/>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "EllipsoidBaseType")
+@XmlSeeAlso({
+    EllipsoidType.class
+})
+public abstract class EllipsoidBaseType
+    extends DefinitionType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/EngineeringCRSRefType.java

@@ -1,0 +1,298 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Association to an engineering coordinate reference system, either referencing or containing the definition of that reference system. 
+ * 
+ * <p>Java class for EngineeringCRSRefType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="EngineeringCRSRefType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element ref="{http://www.opengis.net/gml}EngineeringCRS"/>
+ *       &lt;/sequence>
+ *       &lt;attGroup ref="{http://www.opengis.net/gml}AssociationAttributeGroup"/>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "EngineeringCRSRefType", propOrder = {
+    "engineeringCRS"
+})
+public class EngineeringCRSRefType {
+
+    @XmlElement(name = "EngineeringCRS")
+    protected EngineeringCRSType engineeringCRS;
+    @XmlAttribute(namespace = "http://www.opengis.net/gml")
+    @XmlSchemaType(name = "anyURI")
+    protected String remoteSchema;
+    /**
+     * 
+     * 
+     */
+    @XmlAttribute(namespace = "http://www.w3.org/1999/xlink")
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/FormulaType.java

@@ -1,0 +1,150 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Paremeters of a simple formula by which a value using this unit of measure can be converted to the corresponding value using the preferred unit of measure. The formula element contains elements a, b, c and d, whose values use the XML Schema type "double". These values are used in the formula y = (a + bx) / (c + dx), where x is a value using this unit, and y is the corresponding value using the preferred unit. The elements a and d are optional, and if values are not provided, those parameters are considered to be zero. If values are not provided for both a and d, the formula is equivalent to a fraction with numerator and denominator parameters.
+ * 
+ * <p>Java class for FormulaType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="FormulaType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="a" type="{http://www.w3.org/2001/XMLSchema}double" minOccurs="0"/>
+ *         &lt;element name="b" type="{http://www.w3.org/2001/XMLSchema}double"/>
+ *         &lt;element name="c" type="{http://www.w3.org/2001/XMLSchema}double"/>
+ *         &lt;element name="d" type="{http://www.w3.org/2001/XMLSchema}double" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "FormulaType", propOrder = {
+    "a",
+    "b",
+    "c",
+    "d"
+})
+public class FormulaType {
+
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/GenericMetaDataType.java

@@ -1,0 +1,42 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for GenericMetaDataType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="GenericMetaDataType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractMetaDataType">
+ *       &lt;sequence>
+ *         &lt;any processContents='lax' maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "GenericMetaDataType")
+public class GenericMetaDataType
+    extends AbstractMetaDataType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/GridLimitsType.java

@@ -1,0 +1,73 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for GridLimitsType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="GridLimitsType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="GridEnvelope" type="{http://www.opengis.net/gml}GridEnvelopeType"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "GridLimitsType", propOrder = {
+    "gridEnvelope"
+})
+public class GridLimitsType {
+
+    @XmlElement(name = "GridEnvelope", required = true)
+    protected GridEnvelopeType gridEnvelope;
+
+    /**
+     * Gets the value of the gridEnvelope property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link GridEnvelopeType }
+     *     
+     */
+    public GridEnvelopeType getGridEnvelope() {
+        return gridEnvelope;
+    }
+
+    /**
+     * Sets the value of the gridEnvelope property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link GridEnvelopeType }
+     *     
+     */
+    public void setGridEnvelope(GridEnvelopeType value) {
+        this.gridEnvelope = value;
+    }
+
+    public boolean isSetGridEnvelope() {
+        return (this.gridEnvelope!= null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/GridType.java

@@ -1,0 +1,159 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.math.BigInteger;
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlSeeAlso;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * An unrectified grid, which is a network composed of two or more sets of equally spaced parallel lines in which the members of each set intersect the members of the other sets at right angles.
+ * 
+ * <p>Java class for GridType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="GridType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractGeometryType">
+ *       &lt;sequence>
+ *         &lt;element name="limits" type="{http://www.opengis.net/gml}GridLimitsType"/>
+ *         &lt;element name="axisName" type="{http://www.w3.org/2001/XMLSchema}string" maxOccurs="unbounded"/>
+ *       &lt;/sequence>
+ *       &lt;attribute name="dimension" use="required" type="{http://www.w3.org/2001/XMLSchema}positiveInteger" />
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "GridType", propOrder = {
+    "limits",
+    "axisName"
+})
+@XmlSeeAlso({
+    RectifiedGridType.class
+})
+public class GridType
+    extends AbstractGeometryType
+{
+
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/MovingObjectStatusType.java

@@ -1,0 +1,237 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * This type encapsulates various dynamic properties of moving objects        
+ *              (points, lines, regions). It is useful for dealing with features whose        
+ *              geometry or topology changes over time.
+ * 
+ * <p>Java class for MovingObjectStatusType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="MovingObjectStatusType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractTimeSliceType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/gml}location"/>
+ *         &lt;element name="speed" type="{http://www.opengis.net/gml}MeasureType" minOccurs="0"/>
+ *         &lt;element name="bearing" type="{http://www.opengis.net/gml}DirectionPropertyType" minOccurs="0"/>
+ *         &lt;element name="acceleration" type="{http://www.opengis.net/gml}MeasureType" minOccurs="0"/>
+ *         &lt;element name="elevation" type="{http://www.opengis.net/gml}MeasureType" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/gml}status" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "MovingObjectStatusType", propOrder = {
+    "location",
+    "speed",
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/PolygonPatchArrayPropertyType.java

@@ -1,0 +1,45 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * This type defines a container for an array of 
+ *    polygon patches.
+ * 
+ * <p>Java class for PolygonPatchArrayPropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="PolygonPatchArrayPropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.opengis.net/gml}SurfacePatchArrayPropertyType">
+ *       &lt;sequence maxOccurs="unbounded" minOccurs="0">
+ *         &lt;element ref="{http://www.opengis.net/gml}PolygonPatch"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "PolygonPatchArrayPropertyType")
+public class PolygonPatchArrayPropertyType
+    extends SurfacePatchArrayPropertyType
+{
+
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/PolygonPropertyType.java

@@ -1,0 +1,299 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * This type is deprecated with GML 3 and shall not be used. It is included for backwards compatibility with GML 2. Use SurfacePropertyType instead.
+ * A property that has a polygon as its value domain can either be an appropriate geometry element encapsulated in an element of this type or an XLink reference to a remote geometry element (where remote includes geometry elements located elsewhere in the same document). Either the reference or the contained element must be given, but neither both nor none.
+ * 
+ * <p>Java class for PolygonPropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="PolygonPropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element ref="{http://www.opengis.net/gml}Polygon"/>
+ *       &lt;/sequence>
+ *       &lt;attGroup ref="{http://www.opengis.net/gml}AssociationAttributeGroup"/>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "PolygonPropertyType", propOrder = {
+    "polygon"
+})
+public class PolygonPropertyType {
+
+    @XmlElement(name = "Polygon")
+    protected PolygonType polygon;
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/SequenceRuleType.java

@@ -1,0 +1,102 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.XmlValue;
+
+
+/**
+ * <p>Java class for SequenceRuleType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="SequenceRuleType">
+ *   &lt;simpleContent>
+ *     &lt;extension base="&lt;http://www.opengis.net/gml>SequenceRuleNames">
+ *       &lt;attribute name="order" type="{http://www.opengis.net/gml}IncrementOrder" />
+ *     &lt;/extension>
+ *   &lt;/simpleContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "SequenceRuleType", propOrder = {
+    "value"
+})
+public class SequenceRuleType {
+
+    @XmlValue
+    protected SequenceRuleNames value;
+    @XmlAttribute
+    protected IncrementOrder order;
+
+    /**
+     * List of codes (adopted from ISO 19123 Annex C) that identifies the rule for traversing a grid to correspond with the sequence of members of the rangeSet.
+     * 
+     * @return
+     *     possible object is
+     *     {@link SequenceRuleNames }
+     *     
+     */
+    public SequenceRuleNames getValue() {
+        return value;
+    }
+
+    /**
+     * Sets the value of the value property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link SequenceRuleNames }
+     *     
+     */
+    public void setValue(SequenceRuleNames value) {
+        this.value = value;
+    }
+
+    public boolean isSetValue() {
+        return (this.value!= null);
+    }
+
+    /**
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/SolidType.java

@@ -1,0 +1,117 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * A solid is the basis for 3-dimensional geometry. The extent of a solid is defined by the boundary surfaces (shells). A shell is represented by a composite surface, where every  shell is used to represent a single connected component of the boundary of a solid. It consists of a composite surface (a list of orientable surfaces) connected in a topological cycle (an object whose boundary is empty). Unlike a Ring, a Shell's elements have no natural sort order. Like Rings, Shells are simple.
+ * 
+ * <p>Java class for SolidType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="SolidType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractSolidType">
+ *       &lt;sequence>
+ *         &lt;element name="exterior" type="{http://www.opengis.net/gml}SurfacePropertyType" minOccurs="0"/>
+ *         &lt;element name="interior" type="{http://www.opengis.net/gml}SurfacePropertyType" maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "SolidType", propOrder = {
+    "exterior",
+    "interior"
+})
+public class SolidType
+    extends AbstractSolidType
+{
+
+    protected SurfacePropertyType exterior;
+    protected List<SurfacePropertyType> interior;
+
+    /**
+     * Gets the value of the exterior property.
+     * 
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/SymbolTypeEnumeration.java

@@ -1,0 +1,61 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlEnum;
+import javax.xml.bind.annotation.XmlEnumValue;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for SymbolTypeEnumeration.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * <p>
+ * <pre>
+ * &lt;simpleType name="SymbolTypeEnumeration">
+ *   &lt;restriction base="{http://www.w3.org/2001/XMLSchema}string">
+ *     &lt;enumeration value="svg"/>
+ *     &lt;enumeration value="xpath"/>
+ *     &lt;enumeration value="other"/>
+ *   &lt;/restriction>
+ * &lt;/simpleType>
+ * </pre>
+ * 
+ */
+@XmlType(name = "SymbolTypeEnumeration")
+@XmlEnum
+public enum SymbolTypeEnumeration {
+
+    @XmlEnumValue("svg")
+    SVG("svg"),
+    @XmlEnumValue("xpath")
+    XPATH("xpath"),
+    @XmlEnumValue("other")
+    OTHER("other");
+    private final String value;
+
+    SymbolTypeEnumeration(String v) {
+        value = v;
+    }
+
+    public String value() {
+        return value;
+    }
+
+    public static SymbolTypeEnumeration fromValue(String v) {
+        for (SymbolTypeEnumeration c: SymbolTypeEnumeration.values()) {
+            if (c.value.equals(v)) {
+                return c;
+            }
+        }
+        throw new IllegalArgumentException(v);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/TimeClockType.java

@@ -1,0 +1,189 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.datatype.XMLGregorianCalendar;
+
+
+/**
+ * A clock provides a basis for defining temporal position within a day. 
+ *       A clock must be used with a calendar in order to provide a complete description of a temporal position 
+ *       within a specific day.
+ * 
+ * <p>Java class for TimeClockType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="TimeClockType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractTimeReferenceSystemType">
+ *       &lt;sequence>
+ *         &lt;element name="referenceEvent" type="{http://www.opengis.net/gml}StringOrRefType"/>
+ *         &lt;element name="referenceTime" type="{http://www.w3.org/2001/XMLSchema}time"/>
+ *         &lt;element name="utcReference" type="{http://www.w3.org/2001/XMLSchema}time"/>
+ *         &lt;element name="dateBasis" type="{http://www.opengis.net/gml}TimeCalendarPropertyType" maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "TimeClockType", propOrder = {
+    "referenceEvent",
+    "referenceTime",
+    "utcReference",
+    "dateBasis"
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/TopoSolidType.java

@@ -1,0 +1,88 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * The topological boundary of a TopoSolid consists of a set of directed faces. Note that all faces associated with the TopoSolid, including dangling faces, appear in the boundary. The coboundary of a TopoSolid is empty and hence requires no representation.
+ * 
+ * <p>Java class for TopoSolidType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="TopoSolidType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/gml}AbstractTopoPrimitiveType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/gml}directedFace" maxOccurs="unbounded"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "TopoSolidType", propOrder = {
+    "directedFace"
+})
+public class TopoSolidType
+    extends AbstractTopoPrimitiveType
+{
+
+    @XmlElement(required = true)
+    protected List<DirectedFacePropertyType> directedFace;
+
+    /**
+     * Gets the value of the directedFace property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the directedFace property.
+     * 
+     * <p>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/TopoVolumePropertyType.java

@@ -1,0 +1,73 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.gml.v_3_1_1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for TopoVolumePropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="TopoVolumePropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/gml}TopoVolume"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "TopoVolumePropertyType", propOrder = {
+    "topoVolume"
+})
+public class TopoVolumePropertyType {
+
+    @XmlElement(name = "TopoVolume", required = true)
+    protected TopoVolumeType topoVolume;
+
+    /**
+     * Gets the value of the topoVolume property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link TopoVolumeType }
+     *     
+     */
+    public TopoVolumeType getTopoVolume() {
+        return topoVolume;
+    }
+
+    /**
+     * Sets the value of the topoVolume property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link TopoVolumeType }
+     *     
+     */
+    public void setTopoVolume(TopoVolumeType value) {
+        this.topoVolume = value;
+    }
+
+    public boolean isSetTopoVolume() {
+        return (this.topoVolume!= null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/gml/v_3_1_1/package-info.java

@@ -1,0 +1,9 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+@javax.xml.bind.annotation.XmlSchema(namespace = "http://www.opengis.net/gml", elementFormDefault = javax.xml.bind.annotation.XmlNsForm.QUALIFIED)
+package net.opengis.gml.v_3_1_1;

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/BBOXType.java

@@ -1,0 +1,113 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.gml.v_3_1_1.EnvelopeType;
+import net.opengis.gml.v_3_1_1.EnvelopeWithTimePeriodType;
+
+
+/**
+ * <p>Java class for BBOXType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="BBOXType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/ogc}SpatialOpsType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/ogc}PropertyName"/>
+ *         &lt;element ref="{http://www.opengis.net/gml}Envelope"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "BBOXType", propOrder = {
+    "propertyName",
+    "envelope"
+})
+public class BBOXType
+    extends SpatialOpsType
+{
+
+    @XmlElement(name = "PropertyName", required = true)
+    protected PropertyNameType propertyName;
+    @XmlElementRef(name = "Envelope", namespace = "http://www.opengis.net/gml", type = JAXBElement.class)
+    protected JAXBElement<? extends EnvelopeType> envelope;
+
+    /**
+     * Gets the value of the propertyName property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link PropertyNameType }
+     *     
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/DistanceBufferType.java

@@ -1,0 +1,240 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.gml.v_3_1_1.ACurveElement;
+import net.opengis.gml.v_3_1_1.ARingElement;
+import net.opengis.gml.v_3_1_1.ASolidElement;
+import net.opengis.gml.v_3_1_1.ASurfaceElement;
+import net.opengis.gml.v_3_1_1.AbstractCurveType;
+import net.opengis.gml.v_3_1_1.AbstractGeometricAggregateType;
+import net.opengis.gml.v_3_1_1.AbstractGeometricPrimitiveType;
+import net.opengis.gml.v_3_1_1.AbstractGeometryType;
+import net.opengis.gml.v_3_1_1.AbstractRingType;
+import net.opengis.gml.v_3_1_1.AbstractSolidType;
+import net.opengis.gml.v_3_1_1.AbstractSurfaceType;
+import net.opengis.gml.v_3_1_1.CompositeCurveType;
+import net.opengis.gml.v_3_1_1.CompositeSolidType;
+import net.opengis.gml.v_3_1_1.CompositeSurfaceType;
+import net.opengis.gml.v_3_1_1.GeometricComplexType;
+import net.opengis.gml.v_3_1_1.GridType;
+import net.opengis.gml.v_3_1_1.LineStringType;
+import net.opengis.gml.v_3_1_1.LinearRingType;
+import net.opengis.gml.v_3_1_1.MultiCurveType;
+import net.opengis.gml.v_3_1_1.MultiGeometryType;
+import net.opengis.gml.v_3_1_1.MultiLineStringType;
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/FilterType.java

@@ -1,0 +1,229 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for FilterType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="FilterType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;choice>
+ *         &lt;element ref="{http://www.opengis.net/ogc}spatialOps"/>
+ *         &lt;element ref="{http://www.opengis.net/ogc}comparisonOps"/>
+ *         &lt;element ref="{http://www.opengis.net/ogc}logicOps"/>
+ *         &lt;element ref="{http://www.opengis.net/ogc}_Id" maxOccurs="unbounded"/>
+ *       &lt;/choice>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "FilterType", propOrder = {
+    "spatialOps",
+    "comparisonOps",
+    "logicOps",
+    "id"
+})
+public class FilterType {
+
+    @XmlElementRef(name = "spatialOps", namespace = "http://www.opengis.net/ogc", type = JAXBElement.class)
+    protected JAXBElement<? extends SpatialOpsType> spatialOps;
+    @XmlElementRef(name = "comparisonOps", namespace = "http://www.opengis.net/ogc", type = JAXBElement.class)
+    protected JAXBElement<? extends ComparisonOpsType> comparisonOps;
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/FunctionNamesType.java

@@ -1,0 +1,84 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for FunctionNamesType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="FunctionNamesType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence maxOccurs="unbounded">
+ *         &lt;element name="FunctionName" type="{http://www.opengis.net/ogc}FunctionNameType"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "FunctionNamesType", propOrder = {
+    "functionName"
+})
+public class FunctionNamesType {
+
+    @XmlElement(name = "FunctionName", required = true)
+    protected List<FunctionNameType> functionName;
+
+    /**
+     * Gets the value of the functionName property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the functionName property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getFunctionName().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
+     * Objects of the following type(s) are allowed in the list
+     * {@link FunctionNameType }
+     * 
+     * 
+     */
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/GeometryOperandsType.java

@@ -1,0 +1,85 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.namespace.QName;
+
+
+/**
+ * <p>Java class for GeometryOperandsType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="GeometryOperandsType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="GeometryOperand" type="{http://www.opengis.net/ogc}GeometryOperandType" maxOccurs="unbounded"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "GeometryOperandsType", propOrder = {
+    "geometryOperand"
+})
+public class GeometryOperandsType {
+
+    @XmlElement(name = "GeometryOperand", required = true)
+    protected List<QName> geometryOperand;
+
+    /**
+     * Gets the value of the geometryOperand property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the geometryOperand property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getGeometryOperand().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/SpatialOperatorType.java

@@ -1,0 +1,105 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for SpatialOperatorType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="SpatialOperatorType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="GeometryOperands" type="{http://www.opengis.net/ogc}GeometryOperandsType" minOccurs="0"/>
+ *       &lt;/sequence>
+ *       &lt;attribute name="name" type="{http://www.opengis.net/ogc}SpatialOperatorNameType" />
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "SpatialOperatorType", propOrder = {
+    "geometryOperands"
+})
+public class SpatialOperatorType {
+
+    @XmlElement(name = "GeometryOperands")
+    protected GeometryOperandsType geometryOperands;
+    @XmlAttribute
+    protected SpatialOperatorNameType name;
+
+    /**
+     * Gets the value of the geometryOperands property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link GeometryOperandsType }
+     *     
+     */
+    public GeometryOperandsType getGeometryOperands() {
+        return geometryOperands;
+    }
+
+    /**
+     * Sets the value of the geometryOperands property.
+     * 
+     * @param value
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ogc/SpatialOperatorsType.java

@@ -1,0 +1,84 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ogc;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for SpatialOperatorsType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="SpatialOperatorsType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="SpatialOperator" type="{http://www.opengis.net/ogc}SpatialOperatorType" maxOccurs="unbounded"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "SpatialOperatorsType", propOrder = {
+    "spatialOperator"
+})
+public class SpatialOperatorsType {
+
+    @XmlElement(name = "SpatialOperator", required = true)
+    protected List<SpatialOperatorType> spatialOperator;
+
+    /**
+     * Gets the value of the spatialOperator property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the spatialOperator property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getSpatialOperator().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ows/_1/DCP.java

@@ -1,0 +1,75 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ows._1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;choice>
+ *         &lt;element ref="{http://www.opengis.net/ows/1.1}HTTP"/>
+ *       &lt;/choice>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "", propOrder = {
+    "http"
+})
+@XmlRootElement(name = "DCP")
+public class DCP {
+
+    @XmlElement(name = "HTTP")
+    protected HTTP http;
+
+    /**
+     * Gets the value of the http property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link HTTP }
+     *     
+     */
+    public HTTP getHTTP() {
+        return http;
+    }
+
+    /**
+     * Sets the value of the http property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link HTTP }
+     *     
+     */
+    public void setHTTP(HTTP value) {
+        this.http = value;
+    }
+
+    public boolean isSetHTTP() {
+        return (this.http!= null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ows/_1/LanguageStringType.java

@@ -1,0 +1,109 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ows._1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.XmlValue;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * Text string with the language of the string identified as recommended in the XML 1.0 W3C Recommendation, section 2.12. 
+ * 
+ * <p>Java class for LanguageStringType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="LanguageStringType">
+ *   &lt;simpleContent>
+ *     &lt;extension base="&lt;http://www.w3.org/2001/XMLSchema>string">
+ *       &lt;attribute ref="{http://www.w3.org/XML/1998/namespace}langx"/>
+ *     &lt;/extension>
+ *   &lt;/simpleContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "LanguageStringType", propOrder = {
+    "value"
+})
+public class LanguageStringType {
+
+    @XmlValue
+    protected String value;
+    @XmlAttribute(namespace = "http://www.w3.org/XML/1998/namespace")
+    @XmlJavaTypeAdapter(CollapsedStringAdapter.class)
+    @XmlSchemaType(name = "language")
+    protected String langx;
+
+    /**
+     * Gets the value of the value property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getValue() {
+        return value;
+    }
+
+    /**
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ows/_1/ManifestType.java

@@ -1,0 +1,88 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ows._1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Unordered list of one or more groups of references to remote and/or local resources. 
+ * 
+ * <p>Java class for ManifestType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="ManifestType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/ows/1.1}BasicIdentificationType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/ows/1.1}ReferenceGroup" maxOccurs="unbounded"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "ManifestType", propOrder = {
+    "referenceGroup"
+})
+public class ManifestType
+    extends BasicIdentificationType
+{
+
+    @XmlElement(name = "ReferenceGroup", required = true)
+    protected List<ReferenceGroupType> referenceGroup;
+
+    /**
+     * Gets the value of the referenceGroup property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the referenceGroup property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getReferenceGroup().add(newItem);
+     * </pre>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ows/_1/ServiceIdentification.java

@@ -1,0 +1,236 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ows._1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/ows/1.1}DescriptionType">
+ *       &lt;sequence>
+ *         &lt;element name="ServiceType" type="{http://www.opengis.net/ows/1.1}CodeType"/>
+ *         &lt;element name="ServiceTypeVersion" type="{http://www.opengis.net/ows/1.1}VersionType" maxOccurs="unbounded"/>
+ *         &lt;element name="Profile" type="{http://www.w3.org/2001/XMLSchema}anyURI" maxOccurs="unbounded" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/ows/1.1}Fees" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/ows/1.1}AccessConstraints" maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "", propOrder = {
+    "serviceType",
+    "serviceTypeVersion",
+    "profile",
+    "fees",
+    "accessConstraints"
+})
+@XmlRootElement(name = "ServiceIdentification")
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/ows/_1/TelephoneType.java

@@ -1,0 +1,127 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.ows._1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * Telephone numbers for contacting the responsible individual or organization. 
+ * 
+ * <p>Java class for TelephoneType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="TelephoneType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="Voice" type="{http://www.w3.org/2001/XMLSchema}string" maxOccurs="unbounded" minOccurs="0"/>
+ *         &lt;element name="Facsimile" type="{http://www.w3.org/2001/XMLSchema}string" maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "TelephoneType", propOrder = {
+    "voice",
+    "facsimile"
+})
+public class TelephoneType {
+
+    @XmlElement(name = "Voice")
+    protected List<String> voice;
+    @XmlElement(name = "Facsimile")
+    protected List<String> facsimile;
+
+    /**
+     * Gets the value of the voice property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sensorml/_1_0/Classification.java

@@ -1,0 +1,530 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sensorml._1_0;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlID;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element name="ClassifierList">
+ *           &lt;complexType>
+ *             &lt;complexContent>
+ *               &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *                 &lt;sequence>
+ *                   &lt;element name="classifier" maxOccurs="unbounded">
+ *                     &lt;complexType>
+ *                       &lt;complexContent>
+ *                         &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *                           &lt;sequence>
+ *                             &lt;element ref="{http://www.opengis.net/sensorML/1.0.1}Term"/>
+ *                           &lt;/sequence>
+ *                           &lt;attribute name="name" type="{http://www.w3.org/2001/XMLSchema}token" />
+ *                         &lt;/restriction>
+ *                       &lt;/complexContent>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sensorml/_1_0/IoComponentPropertyType.java

@@ -1,0 +1,723 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sensorml._1_0;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+import net.opengis.swe._1_0.AbstractDataArrayType;
+import net.opengis.swe._1_0.AbstractDataRecordType;
+import net.opengis.swe._1_0.Boolean;
+import net.opengis.swe._1_0.Category;
+import net.opengis.swe._1_0.ConditionalDataType;
+import net.opengis.swe._1_0.ConditionalValueType;
+import net.opengis.swe._1_0.Count;
+import net.opengis.swe._1_0.CountRange;
+import net.opengis.swe._1_0.CurveType;
+import net.opengis.swe._1_0.DataArrayType;
+import net.opengis.swe._1_0.DataRecordType;
+import net.opengis.swe._1_0.EnvelopeType;
+import net.opengis.swe._1_0.GeoLocationArea;
+import net.opengis.swe._1_0.NormalizedCurveType;
+import net.opengis.swe._1_0.ObservableProperty;
+import net.opengis.swe._1_0.PositionType;
+import net.opengis.swe._1_0.Quantity;
+import net.opengis.swe._1_0.QuantityRange;
+import net.opengis.swe._1_0.SimpleDataRecordType;
+import net.opengis.swe._1_0.SquareMatrixType;
+import net.opengis.swe._1_0.Text;
+import net.opengis.swe._1_0.Time;
+import net.opengis.swe._1_0.TimeRange;
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sensorml/_1_0/Link.java

@@ -1,0 +1,271 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sensorml._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="source">
+ *           &lt;complexType>
+ *             &lt;complexContent>
+ *               &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *                 &lt;attribute name="ref" use="required" type="{http://www.opengis.net/sensorML/1.0.1}linkRef" />
+ *               &lt;/restriction>
+ *             &lt;/complexContent>
+ *           &lt;/complexType>
+ *         &lt;/element>
+ *         &lt;element name="destination">
+ *           &lt;complexType>
+ *             &lt;complexContent>
+ *               &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *                 &lt;attribute name="ref" use="required" type="{http://www.opengis.net/sensorML/1.0.1}linkRef" />
+ *               &lt;/restriction>
+ *             &lt;/complexContent>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sensorml/_1_0/SystemType.java

@@ -1,0 +1,43 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sensorml._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for SystemType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="SystemType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/sensorML/1.0.1}AbstractComponentType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/sensorML/1.0.1}components" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/sensorML/1.0.1}positions" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/sensorML/1.0.1}connections" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "SystemType")
+public class SystemType
+    extends AbstractDerivableProcessType
+{
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sensorml/_1_0/package-info.java

@@ -1,0 +1,9 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+@javax.xml.bind.annotation.XmlSchema(namespace = "http://www.opengis.net/sensorML/1.0.1", elementFormDefault = javax.xml.bind.annotation.XmlNsForm.QUALIFIED)
+package net.opengis.sensorml._1_0;

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sos/_1/Capabilities.java

@@ -1,0 +1,110 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sos._1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.ows._1.CapabilitiesBaseType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/ows/1.1}CapabilitiesBaseType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/sos/1.0}Filter_Capabilities" minOccurs="0"/>
+ *         &lt;element ref="{http://www.opengis.net/sos/1.0}Contents" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "", propOrder = {
+    "filterCapabilities",
+    "contents"
+})
+@XmlRootElement(name = "Capabilities")
+public class Capabilities
+    extends CapabilitiesBaseType
+{
+
+    @XmlElement(name = "Filter_Capabilities")
+    protected FilterCapabilities filterCapabilities;
+    @XmlElement(name = "Contents")
+    protected Contents contents;
+
+    /**
+     * Gets the value of the filterCapabilities property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link FilterCapabilities }
+     *     
+     */
+    public FilterCapabilities getFilterCapabilities() {
+        return filterCapabilities;
+    }
+
+    /**
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sos/_1/GetObservation.java

@@ -1,0 +1,720 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sos._1;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.namespace.QName;
+import net.opengis.ogc.BBOXType;
+import net.opengis.ogc.BinaryComparisonOpType;
+import net.opengis.ogc.BinarySpatialOpType;
+import net.opengis.ogc.BinaryTemporalOpType;
+import net.opengis.ogc.ComparisonOpsType;
+import net.opengis.ogc.DistanceBufferType;
+import net.opengis.ogc.PropertyIsBetweenType;
+import net.opengis.ogc.PropertyIsLikeType;
+import net.opengis.ogc.PropertyIsNullType;
+import net.opengis.ogc.SpatialOpsType;
+import net.opengis.ogc.TemporalOpsType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/sos/1.0}RequestBaseType">
+ *       &lt;sequence minOccurs="0">
+ *         &lt;element name="offering" type="{http://www.w3.org/2001/XMLSchema}anyURI"/>
+ *         &lt;element name="eventTime" maxOccurs="unbounded" minOccurs="0">
+ *           &lt;complexType>
+ *             &lt;complexContent>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/sos/_1/InsertObservation.java

@@ -1,0 +1,112 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.sos._1;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import net.opengis.om._1.ObservationType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/sos/1.0}RequestBaseType">
+ *       &lt;sequence>
+ *         &lt;element name="AssignedSensorId" type="{http://www.w3.org/2001/XMLSchema}anyURI"/>
+ *         &lt;element ref="{http://www.opengis.net/om/1.0}Observation"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "", propOrder = {
+    "assignedSensorId",
+    "observation"
+})
+@XmlRootElement(name = "InsertObservation")
+public class InsertObservation
+    extends RequestBaseType
+{
+
+    @XmlElement(name = "AssignedSensorId", required = true)
+    @XmlSchemaType(name = "anyURI")
+    protected String assignedSensorId;
+    @XmlElement(name = "Observation", namespace = "http://www.opengis.net/om/1.0", required = true)
+    protected ObservationType observation;
+
+    /**
+     * Gets the value of the assignedSensorId property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/AbstractMatrixType.java

@@ -1,0 +1,109 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlSeeAlso;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for AbstractMatrixType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="AbstractMatrixType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/swe/1.0.1}AbstractDataArrayType">
+ *       &lt;attribute name="referenceFrame" type="{http://www.w3.org/2001/XMLSchema}anyURI" />
+ *       &lt;attribute name="localFrame" type="{http://www.w3.org/2001/XMLSchema}anyURI" />
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "AbstractMatrixType")
+@XmlSeeAlso({
+    SquareMatrixType.class
+})
+public class AbstractMatrixType
+    extends AbstractDataArrayType
+{
+
+    @XmlAttribute
+    @XmlSchemaType(name = "anyURI")
+    protected String referenceFrame;
+    @XmlAttribute
+    @XmlSchemaType(name = "anyURI")
+    protected String localFrame;
+
+    /**
+     * Gets the value of the referenceFrame property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getReferenceFrame() {
+        return referenceFrame;
+    }
+
+    /**
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/AllowedTimes.java

@@ -1,0 +1,219 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlElementRefs;
+import javax.xml.bind.annotation.XmlID;
+import javax.xml.bind.annotation.XmlList;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;choice>
+ *         &lt;choice>
+ *           &lt;element name="min" type="{http://www.opengis.net/swe/1.0.1}timePositionType"/>
+ *           &lt;element name="max" type="{http://www.opengis.net/swe/1.0.1}timePositionType"/>
+ *         &lt;/choice>
+ *         &lt;choice maxOccurs="unbounded">
+ *           &lt;element name="interval" type="{http://www.opengis.net/swe/1.0.1}timePair"/>
+ *           &lt;element name="valueList" type="{http://www.opengis.net/swe/1.0.1}timeList"/>
+ *         &lt;/choice>
+ *       &lt;/choice>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/AllowedValues.java

@@ -1,0 +1,198 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import java.util.ArrayList;
+import java.util.List;
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlElementRefs;
+import javax.xml.bind.annotation.XmlID;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;choice>
+ *         &lt;choice>
+ *           &lt;element name="min" type="{http://www.w3.org/2001/XMLSchema}double"/>
+ *           &lt;element name="max" type="{http://www.w3.org/2001/XMLSchema}double"/>
+ *         &lt;/choice>
+ *         &lt;choice maxOccurs="unbounded">
+ *           &lt;element name="interval" type="{http://www.opengis.net/swe/1.0.1}decimalPair"/>
+ *           &lt;element name="valueList" type="{http://www.opengis.net/swe/1.0.1}decimalList"/>
+ *         &lt;/choice>
+ *       &lt;/choice>
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/ArrayPropertyType.java

@@ -1,0 +1,73 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for ArrayPropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="ArrayPropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element ref="{http://www.opengis.net/swe/1.0.1}Array"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "ArrayPropertyType", propOrder = {
+    "array"
+})
+public class ArrayPropertyType {
+
+    @XmlElement(name = "Array", required = true)
+    protected ArrayType array;
+
+    /**
+     * Gets the value of the array property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link ArrayType }
+     *     
+     */
+    public ArrayType getArray() {
+        return array;
+    }
+
+    /**
+     * Sets the value of the array property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link ArrayType }
+     *     
+     */
+    public void setArray(ArrayType value) {
+        this.array = value;
+    }
+
+    public boolean isSetArray() {
+        return (this.array!= null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/DataComponentPropertyType.java

@@ -1,0 +1,667 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlElementRef;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+import javax.xml.bind.annotation.adapters.CollapsedStringAdapter;
+import javax.xml.bind.annotation.adapters.XmlJavaTypeAdapter;
+
+
+/**
+ * Complex Type for all properties taking the AnyData Group with a "name" key
+ * 
+ * <p>Java class for DataComponentPropertyType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="DataComponentPropertyType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;group ref="{http://www.opengis.net/swe/1.0.1}AnyData" minOccurs="0"/>
+ *       &lt;attGroup ref="{http://www.opengis.net/gml}AssociationAttributeGroup"/>
+ *       &lt;attribute name="name" use="required" type="{http://www.w3.org/2001/XMLSchema}token" />
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "DataComponentPropertyType", propOrder = {
+    "count",
+    "quantity",
+    "time",
+    "_boolean",
+    "category",
+    "text",
+    "quantityRange",
+    "countRange",
+    "timeRange",
+    "abstractDataRecord",
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/IntervalType.java

@@ -1,0 +1,107 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * A generic interval. The type of the two limits will normally be the same.
+ * 
+ * <p>Java class for IntervalType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="IntervalType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="lowerBound" type="{http://www.w3.org/2001/XMLSchema}anyType"/>
+ *         &lt;element name="upperBound" type="{http://www.w3.org/2001/XMLSchema}anyType"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "IntervalType", propOrder = {
+    "lowerBound",
+    "upperBound"
+})
+public class IntervalType {
+
+    @XmlElement(required = true)
+    protected Object lowerBound;
+    @XmlElement(required = true)
+    protected Object upperBound;
+
+    /**
+     * Gets the value of the lowerBound property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link Object }
+     *     
+     */
+    public Object getLowerBound() {
+        return lowerBound;
+    }
+
+    /**
+     * Sets the value of the lowerBound property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link Object }
+     *     
+     */
+    public void setLowerBound(Object value) {
+        this.lowerBound = value;
+    }
+
+    public boolean isSetLowerBound() {
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/MultiplexedStreamFormatType.java

@@ -1,0 +1,104 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlAttribute;
+import javax.xml.bind.annotation.XmlSchemaType;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for MultiplexedStreamFormatType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="MultiplexedStreamFormatType">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/swe/1.0.1}AbstractEncodingType">
+ *       &lt;attribute name="type" use="required" type="{http://www.w3.org/2001/XMLSchema}anyURI" />
+ *       &lt;attribute name="version" use="required" type="{http://www.w3.org/2001/XMLSchema}string" />
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "MultiplexedStreamFormatType")
+public class MultiplexedStreamFormatType
+    extends AbstractEncodingType
+{
+
+    @XmlAttribute(required = true)
+    @XmlSchemaType(name = "anyURI")
+    protected String type;
+    @XmlAttribute(required = true)
+    protected String version;
+
+    /**
+     * Gets the value of the type property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getType() {
+        return type;
+    }
+
+    /**
+     * Sets the value of the type property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link String }
+     *     
+     */
+    public void setType(String value) {
+        this.type = value;
+    }
+
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/net/opengis/swe/_1_0/Text.java

@@ -1,0 +1,75 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package net.opengis.swe._1_0;
+
+import javax.xml.bind.annotation.XmlAccessType;
+import javax.xml.bind.annotation.XmlAccessorType;
+import javax.xml.bind.annotation.XmlRootElement;
+import javax.xml.bind.annotation.XmlType;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.opengis.net/swe/1.0.1}AbstractDataComponentType">
+ *       &lt;sequence>
+ *         &lt;element name="value" type="{http://www.w3.org/2001/XMLSchema}string" minOccurs="0"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+@XmlAccessorType(XmlAccessType.FIELD)
+@XmlType(name = "", propOrder = {
+    "value"
+})
+@XmlRootElement(name = "Text")
+public class Text
+    extends AbstractDataComponentType
+{
+
+    protected String value;
+
+    /**
+     * Gets the value of the value property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getValue() {
+        return value;
+    }
+
+    /**
+     * Sets the value of the value property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link String }
+     *     
+     */
+    public void setValue(String value) {
+        this.value = value;
+    }
+
+    public boolean isSetValue() {
+        return (this.value!= null);
+    }
+
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/WebService/src/org/w3/_2001/smil20/ObjectFactory.java

@@ -1,0 +1,118 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vJAXB 2.1.10 in JDK 6 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2010.02.01 at 05:02:44 PM GMT 
+//
+
+
+package org.w3._2001.smil20;
+
+import javax.xml.bind.JAXBElement;
+import javax.xml.bind.annotation.XmlElementDecl;
+import javax.xml.bind.annotation.XmlRegistry;
+import javax.xml.namespace.QName;
+import org.w3._2001.smil20.language.AnimateColorType;
+import org.w3._2001.smil20.language.AnimateMotionType;
+import org.w3._2001.smil20.language.AnimateType;
+import org.w3._2001.smil20.language.SetType;
+
+
+/**
+ * This object contains factory methods for each 
+ * Java content interface and Java element interface 
+ * generated in the org.w3._2001.smil20 package. 
+ * <p>An ObjectFactory allows you to programatically 
+ * construct new instances of the Java representation 
+ * for XML content. The Java representation of XML 
+ * content can consist of schema derived interfaces 
+ * and classes representing the binding of schema 
+ * type definitions, element declarations and model 
+ * groups.  Factory methods for each of these are 
+ * provided in this class.
+ * 
+ */
+@XmlRegistry
+public class ObjectFactory {
+
+    private final static QName _AnimateMotion_QNAME = new QName("http://www.w3.org/2001/SMIL20/", "animateMotion");
+    private final static QName _Animate_QNAME = new QName("http://www.w3.org/2001/SMIL20/", "animate");
+    private final static QName _AnimateColor_QNAME = new QName("http://www.w3.org/2001/SMIL20/", "animateColor");
+    private final static QName _Set_QNAME = new QName("http://www.w3.org/2001/SMIL20/", "set");
+
+    /**
+     * Create a new ObjectFactory that can be used to create new instances of schema derived classes for package: org.w3._2001.smil20
+     * 
+     */
+    public ObjectFactory() {
+    }
+
+    /**
+     * Create an instance of {@link AnimateMotionPrototype }
+     * 
+     */
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/DataProviderFactory.java
latest

@@ -1,15 +1,0 @@
-package ie.ucd.sensetile.webservice.dataproducer;
-
-import ie.ucd.sensetile.webservice.dataproducer.DataProviderProxyRMI;
-
-public class DataProviderFactory {
-
-    private static DataProviderProxyRMI proxy = null;
-
-    public static DataProviderProxyIf getProxy() {
-        if (proxy == null) {
-            proxy = new DataProviderProxyRMI();
-        }
-        return proxy;
-    }
-}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/PacketStreamReader.java

@@ -102,12 +102,13 @@ public class PacketStreamReader {
        }
 
        /**
-        * get no of packets to read.
+        * get no of list of sensors and provide them
+        * with the next packet from the packInputSteam.
         */
        private void updateSensors() {
          try {
             Packet packet = this.packetInputStream.read();
-            Iterator<Sensor> iter = sensorList.iterator();
+            Iterator < Sensor > iter = sensorList.iterator();
             while (iter.hasNext()) {
                 iter.next().processSensorData(packet);
             }

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SOSProxyFactory.java

@@ -1,0 +1,15 @@
+package ie.ucd.sensetile.webservice.dataproducer;
+
+import ie.ucd.sensetile.webservice.dataproducer.SOSProxyRMI;
+
+public class SOSProxyFactory {
+
+    private static SOSProxyRMI proxy = null;
+
+    public static SOSProxyIf getProxy() {
+        if (proxy == null) {
+            proxy = new SOSProxyRMI();
+        }
+        return proxy;
+    }
+}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SOSProxyIf.java

@@ -1,8 +1,8 @@
 package ie.ucd.sensetile.webservice.dataproducer;
 
-public interface DataProviderProxyIf {
+public interface SOSProxyIf {
 
-    void insertObservation(int observation);
+    void insertObservation(String sosId, String observation);
     
     public String registerSensor( String service,
              String version,

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SOSProxyRMI.java

@@ -1,6 +1,6 @@
 package ie.ucd.sensetile.webservice.dataproducer;
 
-import ie.ucd.sensetile.webservice.sos.SensorObservationIF;
+import ie.ucd.sensetile.dataprovider.rmi.SensorObservationIF;
 
 import java.net.MalformedURLException;
 import java.rmi.Naming;
@@ -9,14 +9,14 @@ import java.rmi.RemoteException;
 import java.rmi.registry.LocateRegistry;
 import java.rmi.registry.Registry;
 
-public class DataProviderProxyRMI implements DataProviderProxyIf{
+public class SOSProxyRMI implements SOSProxyIf{
 
-    private String dataproviderName = "rmi://localhost/SenseTileService";
+    private String sosName = "rmi://localhost/SenseTileService";
 
     private SensorObservationIF sensorObs = null;
 
 
-    public DataProviderProxyRMI() {
+    public SOSProxyRMI() {
         getDataProviderRef();
     }
 
@@ -24,8 +24,7 @@ public class DataProviderProxyRMI implements DataProviderProxyIf{
        try {
            //currently just local host
            Registry registry = LocateRegistry.getRegistry();
-           sensorObs = (SensorObservationIF) registry.lookup(dataproviderName);
-           //sensorObs = (SensorObservationIF) Naming.lookup(dataproviderName);
+           sensorObs = (SensorObservationIF) registry.lookup(sosName);
        } catch (RemoteException e1) {
            e1.printStackTrace();
        } catch (NotBoundException e1) {
@@ -33,9 +32,9 @@ public class DataProviderProxyRMI implements DataProviderProxyIf{
        }
     }
 
-    public final void insertObservation(final int observation) {
+    public final void insertObservation(final String sosId, final String observation) {
         try {
-            sensorObs.insertObservation(observation);
+            sensorObs.insertObservation(sosId, observation);
         } catch (RemoteException e) {
             // TODO Auto-generated catch block
             e.printStackTrace();

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SenseTileSystem.java

@@ -3,6 +3,9 @@ package ie.ucd.sensetile.webservice.dataproducer;
 import ie.ucd.sensetile.dataprovider.SensorObservations;
 import ie.ucd.sensetile.sensorboard.Packet;
 import ie.ucd.sensetile.sensorboard.PacketInputStream;
+import ie.ucd.sensetile.webservice.dataproducer.smlengine.AbstractProcess;
+import ie.ucd.sensetile.webservice.dataproducer.smlengine.SensorMLSystem;
+
 import java.io.InputStream;
 import java.net.MalformedURLException;
 import java.rmi.Naming;
@@ -30,17 +33,10 @@ public class SenseTileSystem {
 
     private SensorMLSystem smlSenseTileSystem;
     private SystemOutput sensorBoardOutput;
+    private SystemOutput convertorOutput;
+
     private SensorMLSystem smlConvetorSystem;
 
-    private DataProviderProxyIf dataProviderProxy = null;
-
-    public DataProviderProxyIf getDataProviderProxy() {
-        return dataProviderProxy;
-    }
-
-    public void setDataProviderProxy(DataProviderProxyIf dataProviderProxy) {
-        this.dataProviderProxy = dataProviderProxy;
-    }
 
     private static Logger logger =
         Logger.getLogger(SenseTileSystem.class);
@@ -75,7 +71,7 @@ public class SenseTileSystem {
      * @throws MalformedURLException
      * @requires this.packetInputStream /= null
      */
-     public final void processSensorData(final Packet packet) {
+     /*public final void processSensorData(final Packet packet) {
          //logger.info("Received Packet to process");
          DataBlockInt pktemp = new DataBlockInt(1);
          pktemp.setIntValue(packet.getTemperature());
@@ -84,7 +80,7 @@ public class SenseTileSystem {
          dataProviderProxy.insertObservation((int) smlSenseTileSystem.
                                getOutput("temperature"));
 
-     }
+     }*/
 
     /**
      * Set the packet reader.
@@ -97,30 +93,41 @@ public class SenseTileSystem {
 
     /**
      * Get an Input stream for the sensetilesystem xml.
+     * @param name
      * @return InputSteam
      */
-     private InputStream getSensorBoardInputStream(String name) {
-         return this.getClass().getResourceAsStream("/"+name);
+     private InputStream getSensorBoardInputStream(final String name) {
+         return this.getClass().getResourceAsStream("/" + name);
 
      }
 
      /**
       * Register Sensors for to receive packets from a PacketStreamReader.
       */
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/SystemOutput.java

@@ -1,13 +1,27 @@
 package ie.ucd.sensetile.webservice.dataproducer;
 
+import java.text.DateFormat;
+import java.text.DateFormatSymbols;
+import java.text.SimpleDateFormat;
+import java.util.Date;
+
+
 public class SystemOutput {
-    
-    DataProviderProxyIf dataProvider = null;
+
+    SOSProxyIf sosProxy = null;
     String xmlDescription = null;
     ObservationIf observation = null;
     String sosId = null;
     String observationTemplate="initial";
-    
+
+    SystemOutput(){
+        init();
+    }
+
+    private void init(){
+        sosProxy = SOSProxyFactory.getProxy();
+    }
+
     public String getSosId() {
         return sosId;
     }
@@ -18,13 +32,13 @@ public class SystemOutput {
     }
 
 
-    public DataProviderProxyIf getDataProvider() {
-        return dataProvider;
+    public SOSProxyIf getDataProvider() {
+        return sosProxy;
     }
 
 
-    public void setDataProvider(DataProviderProxyIf dataProvider) {
-        this.dataProvider = dataProvider;
+    public void setDataProvider(SOSProxyIf dataProvider) {
+        this.sosProxy = dataProvider;
     }
 
 
@@ -48,18 +62,96 @@ public class SystemOutput {
     }
 
 
-    public SystemOutput(DataProviderProxyIf dataProvider) {
-        this.dataProvider = dataProvider;
+    public SystemOutput(SOSProxyIf dataProvider) {
+        this.sosProxy = dataProvider;
     }
-    
+
 
     /*Register with SOS */
-    public void registerSOS() {
-        this.sosId = dataProvider.registerSensor("SOS", "1.0",
+    public final void registerSOS() {
+        this.sosId = sosProxy.registerSensor("SOS", "1.0",
                                     this.xmlDescription,
                                     this.observationTemplate);
+        //System.out.println(this.xmlDescription);
+
+    }
+
+    /*Update SOS with system outputs given sosId*/
+    public final void insertObservation(int observation) {
+        sosProxy.insertObservation(this.sosId, createObservationQuantity(
+                Integer.toString(observation),sosId));
+    }
+
+    /*Update SOS with system outputs given sosId*/
+    public final void insertObservation(double observation) {
+        sosProxy.insertObservation(this.sosId,createObservationQuantity(
+                Double.toString(observation),sosId));
+    }
+
[…]

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/converters/CelciusConverter.java

@@ -1,4 +1,6 @@
-package ie.ucd.sensetile.webservice.dataproducer;
+package ie.ucd.sensetile.webservice.dataproducer.converters;
+
+import ie.ucd.sensetile.webservice.dataproducer.smlengine.AbstractProcess;
 
 import org.vast.process.ProcessException;
 

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/sensors/ThermistorSensor.java

@@ -1,6 +1,8 @@
-package ie.ucd.sensetile.webservice.dataproducer;
+package ie.ucd.sensetile.webservice.dataproducer.sensors;
 
 import ie.ucd.sensetile.sensorboard.Packet;
+import ie.ucd.sensetile.webservice.dataproducer.Sensor;
+import ie.ucd.sensetile.webservice.dataproducer.smlengine.AbstractProcess;
 
 import org.vast.process.ProcessException;
 

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/smlengine/AbstractProcess.java

@@ -1,4 +1,4 @@
-package ie.ucd.sensetile.webservice.dataproducer;
+package ie.ucd.sensetile.webservice.dataproducer.smlengine;
 
 import ie.ucd.sensetile.sensorboard.Packet;
 
@@ -10,10 +10,10 @@ import org.vast.process.ProcessException;
 
 enum ProcessType { SENSOR, CONVERTER }
 
-public abstract class AbstractProcess extends DataProcess {
+public class AbstractProcess extends DataProcess {
 
     private DataValue input;
- 
+   
     private DataValue output;
 
     protected static Logger logger =
@@ -36,10 +36,12 @@ public abstract class AbstractProcess extends DataProcess {
             System.out.println("abstact name ???? " + this.getName());
             // I/O mappings
             input =
-                (DataValue) inputData.getComponent(this.getName() + "Input");
+                (DataValue) this.inputData.getComponent(
+                                           this.getName() + "Input");
 
             output =
-                (DataValue) outputData.getComponent(this.getName() + "Output");
+                (DataValue) this.outputData.getComponent(
+                                       this.getName() + "Output");
 
         } catch (ClassCastException e) {
             throw new ProcessException("Invalid I/O data", e);
@@ -66,7 +68,7 @@ public abstract class AbstractProcess extends DataProcess {
     * Executes process algorithm on inputs and set output data.
     * @throws ProcessException  - Vast Library processing exception
     */
-    public abstract void execute() throws ProcessException;
+    public void execute() throws ProcessException {};
 
 
 }

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/main/src/ie/ucd/sensetile/webservice/dataproducer/smlengine/SensorMLSystem.java

@@ -1,4 +1,6 @@
-package ie.ucd.sensetile.webservice.dataproducer;
+package ie.ucd.sensetile.webservice.dataproducer.smlengine;
+
+import ie.ucd.sensetile.webservice.dataproducer.Sensor;
 
 import java.io.ByteArrayOutputStream;
 import java.io.IOException;
@@ -26,7 +28,7 @@ public class SensorMLSystem {
     private SystemReaderV1 sysReaderv1;
     private DOMHelper domHelper;
 
-    SensorMLSystem(final InputStream inputxml) {
+    public SensorMLSystem(final InputStream inputxml) {
         initialise(inputxml);
     }
 
@@ -78,13 +80,13 @@ public class SensorMLSystem {
        return smlsystem.getOutputList().getComponent(outputname)
        .getData().getDoubleValue();
     }
-
+    
     public double getIntOutput(final String outputname) {
         return smlsystem.getOutputList().getComponent(outputname)
         .getData().getIntValue();
      }
 
-    public final List < DataProcess > getSensors() {
+    public final List < Sensor > getSensors() {
         List  < DataProcess > processlist = smlsystem.getProcessList();
         SMLSystem dp = null;
         List < DataProcess > sensorList = null;
@@ -95,7 +97,22 @@ public class SensorMLSystem {
                 sensorList = dp.getProcessList();
             }
         }
-        return sensorList;
+        return convertSensorList(sensorList);
+    }
+
+    /**
+     * Produces a list of Sensor type given a list of DataProcess
+     * @param dplist List <DataProcess>
+     */
+    private List < Sensor > convertSensorList(List <DataProcess> dplist){
+        List <Sensor> aplist = new ArrayList <Sensor>();
+        DataProcess dp = null;
+        Iterator < DataProcess > iter = dplist.iterator();
+        while (iter.hasNext()) {
+            dp = iter.next();
+            aplist.add((Sensor)dp);       
+        }
+        return aplist;
     }
 
     public final SMLSystem getSMLSystem(String systemName) {

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \def\supervisorname{Your Supervisor Name}
-\def\moderatorname{Your Moderator Name}
+\def\moderatorname{-}
 
@@
 \begin{abstract}
-SenseTileSensor Board
-
-SensorML description.
-
-WebServer to access sensor observations
-
-SensorML Bon Mapping with a tool and method to develop sensorML descriptions
+This thesis evaluates the use of SensorML and SOS in the development of Sensor Systems.
+The SenseTile system is used as a case study for this evaluation.
+A SenseTile web service is developed to access SenseTile sensor data based on a SensorML model of SenseTile.
+Formalising senor descriptions is also looked using BON to SensorML mapping.
 
@@
 
-problems.
+Sensor systems are increasingly deployed to provide sensing information for a myriad of applications.
+These sensor systems contain many different sensor types and are developed by communities working in different domains.
+The heterogeneity of these systems means that interoperability is difficult and expensive.
+Reliably processing sensor observations retrieved from these heterogeneous systems is also problematic.
+A standard way to retrieve sensor data and a standard way to describe sensors are needed to address the current problem of sensor system heterogeneity.
 
-Sensor Networks are increasingly deployed to provide sensing information for a myriad of applications.
-These Sensor Networks are developed by communities working in different domains and contain many different sensor types.
-The heterogeneity of these systems means that interoperability is difficult and expensive.
-Reliably processing sensor observations from these heterogeneous systems is also issue.
-Need to have a sensor description, data description, observation, retrieval and information on accuracy etc for reliable processing.
-Standards are being developed to address these problems.
-Use internet.
-
-Sensor Model Language (SensorML)\cite{SMLref} is one proposed standard for sensor description and processing of sensor observations.
+Standards are being developed to address the above issues.
+Two such standards are the Sensor Model Language (SensorML)\cite{SMLref} specification and the Sensor Observation Service (SOS)\cite{SOSref} specification.
+SensorML is a proposed standard for sensor description and processing of sensor observations.
 SensorML specifies models and XML Schema for describing sensors, the process of sensor measurement and the post processing of sensor observations.
-
-The Sensor Observation Service (SOS)\cite{SOSref}  specification is a proposed standard for retrieving information about sensor and retrieving sensor observations.
+The SOS specification is a proposed standard for retrieving sensor information and sensor observations using Web Services.
 
 The SensorML and SOS specifications are developed by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
-The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as SensorML.
-This thesis investigates the use of these standards in developing sensor systems.
+The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as the SensorML and SOS specifications.
+This thesis investigates the use of these SWE standards in developing sensor systems with emphasis on the use of SensorML and SOS specifications.
 
-Currently there is no formal definition of sensors or sensor data processing.
-There are now formal verification based software development processes [kiniry] with the goal of developing very robust systems.
-Having formal sensor descriptions allows sensor system development with such a verification based process.
+Developing robust sensor systems is also a natural concern.
+There are now formal verification based software development processes \cite{Kiniryref} with the goal of developing very robust systems.
+Currently there is no widely available formal definition of sensors or sensor data processing.
+Formal sensor descriptions allows sensor system development with such a verification based process and their development is examined also in this thesis.
 
-The thesis uses a real sensor system called SenseTile as a case study.
-SenseTile is a sensor and processing package that can contain motion sensors, RFID sensors, temperature sensors, audio sensors, pressure sensors, light level sensors video sensors among others.
+A real sensor system called SenseTile[ref] is used as a case study in this thesis.
+SenseTile is a sensor and processing package contains motion sensors, temperature sensors, audio sensors, pressure sensors and light level sensors among others.
 It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
 
-This thesis describes SenseTile with SensorML.
-Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the sensor SensorML descriptions and sensor observations to be retrieved.
-The Web Service will be based on the SOS specification.
-An assessment of SensorML and SOS is provided based on this work.
+This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board .
+Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor SensorML descriptions and sensor observations.
+The Web Service is based on the SOS specification.
+An assessment of SensorML and SOS specifications is provided based on this work.
 
-To formalize sensor description and process BON\cite{BONref} is a notation and a method for Object Oriented systems analysis and design.
-A BON to SensorML mapping is provided.
-The proposed approach would allow sensor processing to defined formally in BON and then refined to JML/Java.
+To formalize sensor description BON\cite{BONref} is used.
+BON is a notation and a method for Object Oriented systems analysis and design.
+A BON to SensorML mapping is described in thesis which allows sensor data and processing to defined formally in BON and refined to JML/Java as described in \cite{Kiniryref}.
 
@@
 
-\section{Sensor Web Enablement}
-SensorML is developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
+\section{Sensor Web Enablement}\label{SWESec}
+As described in the introduction the SensorML and SOS specifications are developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
 This initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
-It is hoped that this allows for easier integration and development of sensor applications across sensor networks.
+It is hoped that this allows for easier integration of sensor applications across sensor networks.
 The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
@@
 SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations.
-For a good overview of SWE and it's use in Sensor Web Network see \cite{SWEArchref}.
-SensorML, SOS and O\&M specifications are the focus of this thesis and are described in more detail in the following sections.
+For a good overview of SWE and it's use in Sensor Web Networks see \cite{SWEArchref}.
+Along with with the SensorML and SOS specification, the O\&M specification is also used in developing the SenseTile Web Service as it defines the encoding of sensor observations.
 
-\section{Sensor Modeling Language}\label{SMLsection}
+\section{Sensor Model Language}\label{SMLsection}
 Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems.
@@
 \begin{itemize}
-\item support sensor and sensor observation discovery.
-\item support processing of the sensor observations.
-\item support geolocation information about sensors.
-\item provide accuracy information of sensors measurements.
-\item provide an explicit description of the process by which an observation was obtained.
-\item provide an executable process chain for deriving new data products on demand.
+\item Support sensor and sensor observation discovery.
+\item Support processing of the sensor observations.
+\item Support geolocation information about sensors.
+\item Provide accuracy information of sensors measurements.
+\item Provide an explicit description of the process by which an observation was obtained.
+\item Provide an executable process chain for deriving new data on demand.
 \end{itemize}
@@
 The OGC SWE specification uses XML Schema to define the allowable SensorML XML structure.It is noted here that sml is the XML namespace for SensorML and swe is the XML namespace for SWE Common.
-In the later SenseTile SensorML model section examples of the XML encoding are used to describe the SenseTile Model.
 
-\section{Sensor Observation Service}
+\section{Sensor Observation Service}\label{SOSSec}
 SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref}.
@@
 
-The SOS Operation Context diargam also shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
+The SOS Operation Context diagram also shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
 The Data Consumer uses a CS-W catalog to find SOS instances.
@@
 \section{Overview}
+This chapter describes the design of the SenseTile Web Service.
 Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand.
-The SWE framework seems to cover all that is needed for modeling sensors and observations.
-Support for easy exchange of sensor meta-data and sensor observations is the main rational of SWE and fits the needs of the SenseTile Sensor Network.
+The OGC SWE framework as described in section\ref{SWESec} seems to cover all that is needed for modeling sensors and observations.
+Central to the design is a SensorML model of the SenseTile System.
+The section SensorML Model of SenseTile covers the details of this model.
+This model acts as both metadata about the SenseTile system and provides a description of how to process the sensor data for the SenseTile Web Service.
+The Web Service design is based on the SOS specification and allows access to the SensorML model of SenseTile sensors as well as SenseTile sensor observations.
+Two SOS specification concepts as described in section \ref{SOSSec}, the Data Producer and the SOS itself, structure the design and are described in the SenseTile Web Service section.
+
 \section{SensorML Model of SenseTile}\label{SenseTileModelSec}
 
-A SensorML model of SenseTile Sensor Board and its sensors is developed as part of this thesis.
+A SensorML model of SenseTile is developed as part of this thesis.
+The SenseTile Sensor Board and its sensors is modeled in the most detailed.
 The model of the Sensor Board is placed in a context to examine SensorML process chaining where processing of Sensor Board data is performed in a separate SensorML process.
@@
 \begin{lstlisting}
+<sml:connections>
+   <sml:ConnectionList>
+       <sml:connection name="ambientTemperature">
+          <sml:Link>
+              <sml:source ref="this/inputs/ambientTemperature"/>
+              <sml:destination ref="sensorBoardSystem/inputs/ambientTemperatureInput"/>
+          </sml:Link>
+   </sml:connection>
+   <sml:connection name="temperatureDN">
+          <sml:Link>
+              <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
+              <sml:destination ref="this/outputs/temperatureDNOutput"/>
+          </sml:Link>
+   </sml:connection>
    <sml:connection name="convertToTemperature">
@@
    </sml:connection>
+   <sml:connection name="outputTemperature">
+          <sml:Link>
+              <sml:source ref="convertorSystem/outputs/celciusConvOutput"/>
+              <sml:destination ref="this/outputs/temperatureOutput"/>
+          </sml:Link>
+   </sml:connection>
+</sml:ConnectionList>
 \end{lstlisting}
@@
 \subsection{SensorBoard System}
+
+This System models the Sensor Board and contains meta data about the Sensor Board. It connects up the inputs and outputs the sensors.
 
@@
 Instead the method tag provides a URN that is used to lookup the implementation class.
-
-\section{BON to SensorML mapping}
-\lstset{frame=none}
-\begin{table}[!th]
-\begin{tabular}{|l|r|}
-\hline
-BON & SensorML\\
-\hline
-   deferred class AbstractProcess  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & not mapped\\
-\hline
-     parameters:Parameters & not mapped\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup & not mapped\\
- \hline
-
-\end{tabular}
-\caption{Scenario Chart for SenseTile Web Service}
-\label{ex:table}
-\end{table}
 
@@
 Some name collisions were corrected with a JAXB customization file found on the SensorML newsgroups.
+\section{BON to SensorML mapping}
+\lstset{frame=none}
+\begin{table}[!th]
+\begin{tabular}{|l|r|}
+\hline
+BON & SensorML\\
+\hline
+   deferred class AbstractProcess  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & not mapped\\
+\hline
+     parameters:Parameters & not mapped\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup & not mapped\\
+ \hline
+
+\end{tabular}
+\caption{Scenario Chart for SenseTile Web Service}
+\label{ex:table}
+\end{table}
 
@@
 \bibitem{SOSref}Open Geospatial Consortium Inc.,  Sensor Observation Service, 2007
+\bibitem{Kiniryref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
 \bibitem{OMref}Open Geospatial Consortium Inc., Observations and Measurements – Part 1 - Observation schema, 2007
@@
 \end{thebibliography}
-\label{endpage}
 
@@
 
+\label{endpage}
 \end{document}

### 2010-04-09 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 A SenseTile web service is developed to access SenseTile sensor data based on a SensorML model of SenseTile.
-Formalising senor descriptions is also looked using BON to SensorML mapping.
+Formalizing senor descriptions is also looked using BON to SensorML mapping.
 
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{SMLConceptualModel.png}}}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{SWESimpleConceptualModel.png}}}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{SenseTileDescription.png}}}
@@
 
-\section{BON}
-Business Object Notation (BON) is a method for analysis and design of
-object-oriented systems.
-seamlessness, reversibility and software contracting.
-same same semantic and conceptual base  analysis, design, and implementation by using the
- for the notation on all three levels. It supports source code changes back into design and analysis.
-The idea is to treat each subprogram as a subcontractor
-who undertakes to deliver some service (specified by a postcondition), but only if
-certain prerequisites are fulfilled (the precondition).
-same concepts and semantics throughout the development of a
-software product allows for a smooth transition from the initial analysis and design models
-down to executable code.
-\cite{BONref}
+\section{BON} \label{BONsec}
+Business Object Notation (BON) is a notation and a method for analysis and design of Object-Oriented (OO) systems.
+It supports a seamless transition from the design to the source code as it based on the same Object Oriented (OO) concepts supported by most OO languages.
+It is reversible as changes in the source code can be easily mapped back into the design and analysis models.
+It supports software contracting where each function will guarantee a postcondition specified service if a specified precondition is fulfilled by the caller of the function.
+BOB support for software contracting allows for a formal description of a system to be developed.
+For a detailed description of BON see \cite{BONref}.
 \chapter{Design}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
@@
 \begin{figure}[h]
+\centering
 \scalebox{0.65}{\includegraphics{Deployment_sensetile.png}}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
@@
 \begin{figure}[h]
+\centering
 \framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
@@
 \begin{figure}[h]
+\centering
 \scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
@@
 
-The SOS Specification describes the use of a Proxy SOS as shown in figure \ref{fig:Deployment_sos_aggr}.
-This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.
-\begin{figure}[h]
-\scalebox{0.65}{\includegraphics{Deployment_sensetile_sos_aggr.png}}
-\caption{Proxy SOS}\label{fig:Deployment_sos_aggr}
-\end{figure}
+The SOS Specification describes the use of a Proxy SOS eeicpr fix. This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.
 
@@
 
-The DataProducer operates by repeating the below steps continously.
+The DataProducer operates by repeating the below steps continuously.
  \begin{enumerate}
@@
 to access a particular sensors data such as the temperature
-meaurement generated by the thermistor. 
+measurement generated by the thermistor. 
 This library is shown as the SensorBoard cluster in \ref{fig:bon_static_diagam_producer.png}
@@
 will execute the process chains described in the SenseTile SensorML model.
-The sensor objects output will be passed to the coverter objects. The
+The sensor objects output will be passed to the converter objects. The
 outputs of both object types are then available to generate observations.
@@
 need to be sent to the SOS. An O\&M Onservation
-xml string containing the data is sent to
+XML string containing the data is sent to
 the SOS instance. This is timestamped at
@@
 The DataProducer was developed as a SensorML based software.
-It would be parse the SenseTile SensorML model to instantiate the code to excute the processes described.
+It would be parse the SenseTile SensorML model to instantiate the code to execute the processes described.
 This required two aspects to be developed, a framework to run the processes and classes that implemented the components.
@@
 \begin{figure}[h]
+\centering
 \scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
@@
  \begin{figure}[h!]
+\centering
 \scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
@@
 Developing a Web Service is straight forward with Axis2 and it supports several approaches.
-The supported POJOs (Plain Old Java Objects) approach was used for the SenseTile Web Service as straight forward and the code could be reused on a different framework if AXIS2 did not work well.
-The Spring framework, which Axis2 supports, was also used for it's dependency injection functionality.
-A SprinitInit class is needed for Axis to load up the framework.
+The supported POJOs (Plain Old Java Objects) approach was used for the SenseTile Web Service allowing the code to be reused on a different framework if AXIS2 did not work well.
+Axis2 supports the The Spring framework which is used in the SenseTile WebService for it's dependency injection functionality.
+A SprinitInit class is needed for Axis2 to load up the Spring framework.
 
-Axis2 a tool that automatically generates WSDL from the running Web Service. Another tool allows client side code to be generated from the WSDL easing the development of the Web Service clients.
+Axis2 provides a tool that automatically generates WSDL from the running Web Service.
+Another tool allows client side code to be generated from the WSDL easing the development of the Web Service clients.
 
@@
 Some name collisions were corrected with a JAXB customization file found on the SensorML newsgroups.
-\section{BON to SensorML mapping}
+However the default binding was not fully successful for the System type.
+It input, output and parameter properties were missing and the inheritance hierarchy did not provide them.
+A small change was needed to add these properties and then everything worked fine with marshaling and unmarshaling SensorML Systems.
+
+\chapter{BON to SensorML mapping}
+\section{Overview}
+In this chapter a BON(see section\ref{BONsec}) to SensorML( see section \ref{SMLsection}) mapping is described.
+One motivation for this mapping is take advantage of formal software development processes described in \cite{Kiniryref}.
+This is a verification based process that refines BON to JML and then to Java.
+Robust software can thus be developed.
+Another motivation is that it XML is verbose and a more compact description of a SensorML model as facilitated by BON is naturally preferred.
+Lastly SensorML is complex and it is difficult to develop correct SenorML models based only on using the SensorML XML Schema definition.
+How a BON specification can help here is described later in the chapter.
+
+\section{SensorML BON Specification}
+This sections shows an example of one BON specification to explain the approach used to develop BON specifications for SensorML.
+The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.
+The BON specification developed in this thesis does not address the issue of the SensorML metadata.
+This information  is for descriptive purposes only and is not used for processing sensor data.
+BON adds no value here and a different way to add this property  to SensorML sensor descriptions is suggested to be used.
+One potential way to add this metadata is through the use of XSLT\footnote{http://www.w3.org/TR/xslt20/} but this is not examined further in this thesis.
+
+The listing \ref{BONListing} shows the BON specification for the SensorML AbstractProcess as described in section \ref{SMLsection}.
+The AbstractProcess's inputs, outputs and parameters properties are specified as BON features of type SEQUENCE of the appropriate types.
+These AbstractProcess properties are optional according to the SensorML specification and so there is no postcondition to enforce that a value must be returned.
+Setter features are provided and also a constructor feature, make, with the most likely used features as the constructor parameters.
+
+The AbstractProcess's metaDataGroup property is not optional in AbstractProcess.
+In this case the metaDataGroup feature is used a placeholder for the metadata and returns a BON STRING rather than the expected MetaDataGroup.
+
+The complete BON SensorML specification listing can be found in appendix \ref{appenBSMLBONSpec}.
+\newpage
+
+ \begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON AbstractProcess,label=BONListing]
+   deferred class AbstractProcess
+     indexing
+     about: "base substitution group for all processes.";
+     title:        "AbstractProcess.";
+     author:       "Ciaran Palmer.";
+     copyright:    "none.";
+     organisation: "School of Computer Science and Informatics, UCD.";
+     date:         "2009/10/29.";
+     version:      "Revision: 1.00.";
+
+     feature
+
+     make
+       ->inputsIn:SEQUENCE[Input]
+       ->outputsIn:SEQUENCE[Output]
+       require
+         inputsIn/=Void;
+         inputsIn/=Void;
+       ensure
+         delta{inputs,outputs};
+         inputs = inputsIn;
+         outputs = outputsIn;
+       end
+
+     --optional       
+     inputs:SEQUENCE[Input]
+
+      --optional        
+     outputs:SEQUENCE[Output]
+
+      --optional       
+     parameters:Parameters
+
+     metaDataGroup:STRING
+        ensure
+          Result = "metaDataGroup";
+       end
+
+     setInputs
+       ->inputsIn:SEQUENCE[Input]
+       require
+         inputsIn/=Void;
+       ensure
+         delta{inputs};
+         inputs = inputsIn;
+       end
+
+      setOutputs
+       ->outputsIn:SEQUENCE[Output]
+       require
+         outputsIn/=Void;
+       ensure
+         delta{outputs};
+         outputs = outputsIn;
+       end
+
+      setParameters
+       ->ParametersIn:SEQUENCE[Parameter]
+       require
+         ParametersIn/=Void;
+       ensure
+         delta{Parameters};
+         Parameters = ParametersIn;
+       end               
+   end --AbstractProcess
+ \end{lstlisting}
+
+\section{BON to SensorML Mapping}
+Tables of mapping, like table\ref{tab:bonsmlexample} below, are developed to show the mapping of BON features to SensorML XML tags.
+The mapping is shown from a BON feature to the appropriate SensorML XML tags.
+The table below shows a mapping for AbstractProcess.
+The name of the class is not mapped as it is abstract with the sub-classes Component and ProcessModel providing the name of the enclosing tag names.
+The  BON SEQUENCE types such as SEQUENCE[Input] map to a sequence of Input mapped SensorML XML tags found in the Input mapping table.
+The full set of tables are in appendix \ref{appenBSMLBONTables}.
+
 \lstset{frame=none}
+
 \begin{table}[!th]
-\begin{tabular}{|l|r|}
+
+\centering
+\begin{tabular}{|l|l|}
 \hline
-BON & SensorML\\
+BON features & SensorML XML tags\\
 \hline
@@
 \hline 
-     outputs:SEQUENCE[Output] & not mapped\\
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
 \hline
-     parameters:Parameters & not mapped\\
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
 
 \hline                 
-     metaDataGroup:MetaDataGroup & not mapped\\
+     metaDataGroup:STRING &  \begin{lstlisting}
+<\sml:metaDataGroup>metaDataGroup<\sml:metaDataGroup>\end{lstlisting}\\
  \hline
@@
 \end{tabular}
-\caption{Scenario Chart for SenseTile Web Service}
-\label{ex:table}
+
+\caption{BON to SensorML AbstractProcess Mapping table}
+\label{tab:bonsmlexample}
 \end{table}
+
+The MetaDataGroup is not described in BON as explained in the overview section. Only a place holder value mapped from this feature.
+
+Connections as described in section x are connected up using strings that have to be parsed to find out the inputs and outputs to connect up.
+Invariants written in BON can enforce some level of corrects in that inputs in connections must be inputs in the components contained in that chain.
+This is shown in the BON fragment below.
+
+BON fragment invariant.
+
+SWE and BON Clusters.
+\section{BON Thermistor}
+Thermistor can be generated by the following BON class.
+\section{method}
+BON to SensorML
+BON to Java\\JML.
 
@@
 
-Developed lighweight Web Service to explore the use of SensorML in Sensor Networks.
+Developed lightweight Web Service to explore the use of SensorML in Sensor Networks.
 running prototype tested with UCD SensorBoard Simulator code.
@@
 
-\chapter{Appendices}
-\section{SenseTile SensorML Model}
+\appendix
+\chapter{Appendix: SenseTile SensorML Model}
+\section{SenseTile SensorML System}
+\section{SenseTile SensorML SensorBoardSystem}
+\section{SenseTile SensorML ConvetorSystem}
+\section{SenseTile SensorML Thermistor}
+\lstinputlisting{Thermistor.xml}
+\section{SenseTile SensorML CelciusConvertor}
 
-\lstinputlisting{Thermistor.xml}
+\chapter{Appendix: BON to SensorML Mapping}\label{appenB}
+\section {SensorML BON Specifcation}\label{appenBSMLBONSpec}
 
-\section{BON to SensorML Mapping}
 \lstset{basicstyle=\scriptsize,showspaces=false,showstringspaces=false}
 \lstinputlisting{SensorML_formal.bon}
+\section {BON to SensorML mapping tables} \label{appenBSMLBONTables}
+\lstset{frame=none}
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class AbstractProcess  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
 
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
 \label{endpage}

### 2010-04-10 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 A real sensor system called SenseTile[ref] is used as a case study in this thesis.
-SenseTile is a sensor and processing package contains motion sensors, temperature sensors, audio sensors, pressure sensors and light level sensors among others.
-It is used as a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
+SenseTile is a sensor and processing package that contains motion sensors, temperature sensors, audio sensors, pressure sensors and light level sensors among others.
+It is a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.
 
 This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board .
-Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor SensorML descriptions and sensor observations.
+Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor descriptions and sensor observations.
 The Web Service is based on the SOS specification.
@@
 
-To formalize sensor description BON\cite{BONref} is used.
+Buisness Object Notatation(BON)\cite{BONref} isi used to formalize SensorML sensor descriptions in this thesis.
 BON is a notation and a method for Object Oriented systems analysis and design.
-A BON to SensorML mapping is described in thesis which allows sensor data and processing to defined formally in BON and refined to JML/Java as described in \cite{Kiniryref}.
+The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to Java as described in \cite{Kiniryref} as part of developing a robust Sensor System.
 
@@
 The DataProducer was developed as a SensorML based software.
-It would be parse the SenseTile SensorML model to instantiate the code to execute the processes described.
+It parses the SenseTile SensorML model to instantiate the code to execute the processes described.
 This required two aspects to be developed, a framework to run the processes and classes that implemented the components.
@@
 The VastSensorMLEngine, described in section \ref{VastSensorMLEngineSec}, is used in conjunction with the SenseTile SensorML Model (section \ref{SenseTileModelSec}).
-The SensorML is parsed by the VastSensorMLEngine into a DOM tree.
+The SensorML is parsed by the VastSensorMLEngine into a DOM\footnote{http://www.w3.org/DOM/} tree.
 To perform the processing it propagates an execute method to the instantiated processes.
@@
 Lastly SensorML is complex and it is difficult to develop correct SenorML models based only on using the SensorML XML Schema definition.
-How a BON specification can help here is described later in the chapter.
+How a BON specification of SensorML can help here is described later in the chapter.
 
 \section{SensorML BON Specification}
-This sections shows an example of one BON specification to explain the approach used to develop BON specifications for SensorML.
-The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.
-The BON specification developed in this thesis does not address the issue of the SensorML metadata.
+The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.  The BON SensorML model is shown in figure \ref{fig:bonSensoMLModel}. 
+\begin{figure}[h]
+\centering
+\scalebox{0.65}{\includegraphics{bonSensoMLModel.png}}
+\caption{BON SensorML Model}\label{fig:bonSensoMLModel}
+\end{figure}
+The inheritance heirarhcy is changed slightly with System inheriting from ProcessChain and Component inheriting from ProcessModel.
+A new abstraction called Port is added.
+Port contains all the properties of SensorML inputs, outputs and parameters with classes Input, Output and Parameter subtyping to make mapping to SensorML easier.
+A BON specification SWE Common data types Count and Quantity are also described along with needed dependent classes.
+The abstract class DataComponent is the base type for these.
+There are a many other types but only Count and Quantity are developed as they were the ones used in the SenseTile SensorML model.
+The BON specification developed in this thesis does not address the issue of the SensorML metadata and the MetaDataGroup is not specified.
 This information is for descriptive purposes only and is not used for processing sensor data.
-BON adds no value here and a different way to add this property  to SensorML sensor descriptions is suggested to be used.
-One potential way to add this metadata is through the use of XSLT\footnote{http://www.w3.org/TR/xslt20/} but this is not examined further in this thesis.
-
-The listing \ref{BONListing} shows the BON specification for the SensorML AbstractProcess as described in section \ref{SMLsection}.
-The AbstractProcess's inputs, outputs and parameters properties are specified as BON features of type SEQUENCE of the appropriate types.
-These AbstractProcess properties are optional according to the SensorML specification and so there is no postcondition to enforce that a value must be returned.
-Setter features are provided and also a constructor feature, make, with the most likely used features as the constructor parameters.
-
-The AbstractProcess's metaDataGroup property is not optional in AbstractProcess.
-In this case the metaDataGroup feature is used a placeholder for the metadata and returns a BON STRING rather than the expected MetaDataGroup.
+BON adds no value here.
+A different way to add this information to SensorML sensor descriptions is suggested to be used.
 
 The complete BON SensorML specification listing can be found in appendix \ref{appenBSMLBONSpec}.
-\newpage
+
+\section{BON to SensorML XML Mapping}
+Tables of mapping, like table\ref{tab:bonsmlexample} below, are developed to show the mapping of BON features to SensorML XML tags.
+The mapping is shown from a BON feature to the appropriate SensorML XML tags.
+The table below shows a mapping for AbstractProcess.
+The name of the class is not mapped as it is abstract with the sub-classes Component and ProcessModel providing the name of the enclosing tag names.
+The  BON SEQUENCE types such as SEQUENCE[Input] map to a sequence of Input mapped SensorML XML tags found in the Input mapping table.
+The full set of tables are in appendix \ref{appenBSMLBONTables}.
+
+\lstset{frame=none}
+
+\begin{table}[!th]
+
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class AbstractProcess  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:STRING &  \begin{lstlisting}
+<\sml:metaDataGroup>metaDataGroup<\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+
+\caption{BON to SensorML AbstractProcess Mapping table}
+\label{tab:bonsmlexample}
+\end{table}
+
+The MetaDataGroup is not described in BON as explained previously and only a place holder value is mapped from this feature.
+
+\section{ProcessChain connections problem}
+One of the major features of SensorML is the ablity to describe chains of processes.
+The SensorML ProcessChain and System types have a connections property that allows this chaining as decribed in section \ref{SMLref} .However strings are used to describe the connection.
+These connection strings are parsed to find out the inputs and outputs to connect up as described in SenseTile System section \ref.
+This is very error prone and the ability to check that connections are correct to a least some degree is desirable.
+Below are two fragents form SenseTile System and SenseTile Thermistor that descibe a connection form th
+
+\begin{lstlisting}
+   <sml:connection name="convertToTemperature">
+          <sml:Link>
+              <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
+              <sml:destination ref="convertorSystem/inputs/celciusConvInput"/>
+          </sml:Link>
+   </sml:connection>
+\begin{lstlisting}
+
+ BON specification can enforce some level of correctness in that inputs in connections must be inputs in the components contained in that chain. This is shown in the BON fragment below.
 
@@
 
-\section{BON to SensorML Mapping}
-Tables of mapping, like table\ref{tab:bonsmlexample} below, are developed to show the mapping of BON features to SensorML XML tags.
-The mapping is shown from a BON feature to the appropriate SensorML XML tags.
-The table below shows a mapping for AbstractProcess.
-The name of the class is not mapped as it is abstract with the sub-classes Component and ProcessModel providing the name of the enclosing tag names.
-The  BON SEQUENCE types such as SEQUENCE[Input] map to a sequence of Input mapped SensorML XML tags found in the Input mapping table.
-The full set of tables are in appendix \ref{appenBSMLBONTables}.
-
-\lstset{frame=none}
-
-\begin{table}[!th]
-
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class AbstractProcess  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:STRING &  \begin{lstlisting}
-<\sml:metaDataGroup>metaDataGroup<\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-
-\caption{BON to SensorML AbstractProcess Mapping table}
-\label{tab:bonsmlexample}
-\end{table}
-
-The MetaDataGroup is not described in BON as explained in the overview section. Only a place holder value mapped from this feature.
-
-Connections as described in section x are connected up using strings that have to be parsed to find out the inputs and outputs to connect up.
-Invariants written in BON can enforce some level of corrects in that inputs in connections must be inputs in the components contained in that chain.
-This is shown in the BON fragment below.
-
-BON fragment invariant.
-
 SWE and BON Clusters.
-\section{BON Thermistor}
+\section{SenseTile Thermistor BON SensorML Specification }
 Thermistor can be generated by the following BON class.
-\section{method}
-BON to SensorML
-BON to Java\\JML.
 
@@
 \chapter{ Conclusions and Future Work}
-This thesis analyses the role SensorML plays in Sensor Web Networks.
+This thesis analyses the role SensorML and plays in Sensor Web Networks.
 The UCD CASL SenseTile System is used a case study with the development of the SenseTile Web Service to access the SenseTile System sensor data.
@@
 \end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class System  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class C  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class Link  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class ProcessChain  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class ProcessModel  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class Port  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class Quantity  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
+\begin{table}[!th]
+\centering
+\begin{tabular}{|l|l|}
+\hline
+BON features & SensorML XML tags\\
+\hline
+   deferred class Count  & not mapped\\
+\hline     
+     inputs:SEQUENCE[Input] & \begin{lstlisting}
+<sml:inputs><sml:inputlist>
+SEQUENCE[Input]
+<\sml:InputList><\sml:inputs>\end{lstlisting}\\
+
+\hline 
+     outputs:SEQUENCE[Output] & \begin{lstlisting}
+<sml:outputs><sml:outputlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+\hline
+     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
+<sml:parameters><sml:parametertlist>
+SEQUENCE[Input]
+<\sml:outputList><\sml:outputs>\end{lstlisting}\\
+
+\hline                 
+     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
+<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
+ \hline
+
+\end{tabular}
+\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
+\label{ex:table}
+\end{table}
+
 \label{endpage}

### 2010-04-11 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \def\projecttitle{SensorML on SenseTile}
-\def\supervisorname{Your Supervisor Name}
+\def\supervisorname{Dr. Joseph Kiniry}
 \def\moderatorname{-}
@@
  \begin{itemize}
-\item GetCapabilities  -   operation to retrieve SOS observation offering information.
-\item DescribeSensor - operation to retrieve detailed information about the sensors.
-\item GetObservation -   operation to obtain sensor observations.
+\item GetCapabilities  -    retrieve SOS observation offering information.
+\item DescribeSensor -    retrieve detailed information about the sensors.
+\item GetObservation -   obtain sensor observations.
 \end{itemize}
@@
 
-The SOS Operation Context diagram also shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
+The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
+The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
+
+The OGC Catalog Service is not explored in this thesis.
+The SOS Operation Context diagram shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
 The Data Consumer uses a CS-W catalog to find SOS instances.
@@
 This not covered in any real detail in the SOS specification.
-The OGC Catalog Service is not explored in this thesis.
-
-There are also six other operations the SOS refers to as enhanced operations. They are GetResult,
-GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
-DescribeObservationType, and DescribeResultModel. These operations are not covered in this thesis.
-
-The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
-The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
+
+There several optional SOS operations not covered in this thesis. The SOS refers to these as enhanced operations. They are GetResult,
+GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest, DescribeObservationType, and DescribeResultModel.
 
@@
 SenseTile is made up of a sensor platform, a datastore, and a compute farm.
-The below figure shows a outline of this system.
+Figure \ref{SenseTileDescription} shows a outline of this system.
 \begin{figure}[h]
@@
 It is reversible as changes in the source code can be easily mapped back into the design and analysis models.
-It supports software contracting where each function will guarantee a postcondition specified service if a specified precondition is fulfilled by the caller of the function.
-BOB support for software contracting allows for a formal description of a system to be developed.
+It supports design by contract\ref{meyerref} where a function will guarantee a postcondition if a specified precondition is fulfilled by the caller of the function.
+BON support for design by contract allows for a formal description of a system to be developed.
 For a detailed description of BON see \cite{BONref}.
@@
 The Sensor Board system contains metadata about the SenseTile Sensor Board and is a container for SensorML Components that model the Sensor Board's sensors.
+
 The block diagram shows the physical phenomena measured by the sensors being feed as inputs to the Components in the Sensor Board System.
 The SensorBoard System outputs digital numbers based on the sensors measurement that have no physical meaning yet.
-The SenseTile System outputs the digital numbers directly and also connects to the input of the Conversion System.
-The digital numbers from the sensors are fed into the relevant conversion process that produce a meaningful value such as Celsius of Lumen.
+The SenseTile System outputs the digital numbers directly and also connects this output to the input of the Conversion System.
+The digital numbers from the sensors are fed into the relevant conversion process that produce a meaningful value such as Celsius or Lumen.
 This allows the SenseTile Web Service provide the raw data from the sensors or the data converted to meaningful values to clients.
 
-Only the Thermistor Component and the Celsius Component are developed as part of this thesis. However the same approach can applied to the other senors.
-
-Why use this structure? eeicpr
-
-A description of the Systems and Components used in the SenseTile are provided in the following sections.
+Only the Thermistor is modeled in this thesis.
+The  CelsiusConv Component converts the digital number form the Thermistor Component to a Celcius value.
+The same approach can applied to the other senors though for some components no conversion may be necessary.
+
+A description of the SenseTile Systems and Components used in the SenseTile model is provided in the following sections.
 
@@
 \subsection{Converter System}
+The Converter System is a container for the processes that convert the sensor data to a meaningful value. It this case it connects up the inputs and outputs the CelsiusConv Component. 
 \subsection{Celsius Converter Component}
 
-The SensoML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celsius value.
+The SensorML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celsius value.
 It similar to the Thermistor SensorML description but its inputs and outputs are different.
-It process method performs a simple calculation to perform the Celsius conversion.
+Its process method performs a simple calculation to perform the Celsius conversion.
 
@@
 
+\newpage
 \subsection{Supported Scenarios}
-\begin{table}[!th]
-\begin{tabular}{|l|c|r|}
-\hline
-Scenarios\\
-\hline
-Generate Observations from Sensor Data \\
-The DataProducer accesses data from the sensor board \\
-\hline
-Get Sensor Description from SOS\\
-\hline
-Get  Sensor Observation from SOS\\
-\hline
-\end{tabular}
-\caption{Scenario Chart for SenseTile Web Service}
-\label{ex:table}
-\end{table}
+This section provides UML sequence diagrams to show the how the DataProducer and SOS interact to provide the sensor to the DataConsumer client.
+
+\subsubsection{Generate Observations from Sensor Data}
+The DataProducer accesses data from the sensor board 
 \begin{figure}[h]
@@
 \end{figure}
-
+ \begin{enumerate}
+\item -
+\item -
+\item -
+\end{enumerate}
+
+\subsubsection{Get Sensor Description from SOS}
 \begin{figure}[h]
@@
 \end{figure}
-
+ \begin{enumerate}
+\item -
+\item -
+\end{enumerate}
+\subsubsection{Get  Sensor Observation from SOS}
 \begin{figure}[h]
@@
 \end{figure}
-\newpage
+ \begin{enumerate}
+\item -
+\end{enumerate}
+
 \subsection{SenseTile WebService Network Architecture}
@@
 The PacketInputStream will request the SenseTileSystem to exexcute
-the SensorML engine to process the sensor data. The SensoML Enigine
+the SensorML engine to process the sensor data. The SensorML Enigine
 will execute the process chains described in the SenseTile SensorML model.
@@
 A new abstraction called Port is added.
-Port contains all the properties of SensorML inputs, outputs and parameters with classes Input, Output and Parameter subtyping to make mapping to SensorML easier.
+Port contains all the properties of SensorML input, output and parameter as they all have the same properties.
+A PortType class that ensures legal port types of Input, Output and Parameter is also added to the SenorML BON model.
+
 A BON specification SWE Common data types Count and Quantity are also described along with needed dependent classes.
@@
 There are a many other types but only Count and Quantity are developed as they were the ones used in the SenseTile SensorML model.
+
 The BON specification developed in this thesis does not address the issue of the SensorML metadata and the MetaDataGroup is not specified.
@@
 A different way to add this information to SensorML sensor descriptions is suggested to be used.
-
+AssociationAttributeGroup
+SensorML
+Member
 The complete BON SensorML specification listing can be found in appendix \ref{appenBSMLBONSpec}.
 
-\section{BON to SensorML XML Mapping}
-Tables of mapping, like table\ref{tab:bonsmlexample} below, are developed to show the mapping of BON features to SensorML XML tags.
-The mapping is shown from a BON feature to the appropriate SensorML XML tags.
-The table below shows a mapping for AbstractProcess.
-The name of the class is not mapped as it is abstract with the sub-classes Component and ProcessModel providing the name of the enclosing tag names.
-The  BON SEQUENCE types such as SEQUENCE[Input] map to a sequence of Input mapped SensorML XML tags found in the Input mapping table.
-The full set of tables are in appendix \ref{appenBSMLBONTables}.
-
-\lstset{frame=none}
-
-\begin{table}[!th]
-
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class AbstractProcess  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:STRING &  \begin{lstlisting}
-<\sml:metaDataGroup>metaDataGroup<\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-
-\caption{BON to SensorML AbstractProcess Mapping table}
-\label{tab:bonsmlexample}
-\end{table}
-
-The MetaDataGroup is not described in BON as explained previously and only a place holder value is mapped from this feature.
+\section{SensorML BON to XML Mapping}
+SensorML is essential XML defined by XML Schema with its main purpose being interopeability.
+Therefore a mapping SensorML BON specification to the SensorML XML is needed.
+The BON types map closely to SensorML complex types and so a relatively simple scheme is needed.
+Effective BON classes map to XML elements.
+Features in the class map either to attributes of the XML element or a contained XML element.
+Whether a feature is an attribute is indicated in the class as a comment using @SensorML to make this comment as a mapping instruction.
+Otherwise an XML element is assumed.
+The type names of the BON classes and features are the XML element or attribute names.
+Values.
+Case of names.
+
+BON SEQUENCE types need a special rule for mapping to SensorML.
+
+Connections rule.
 
 \section{ProcessChain connections problem}
-One of the major features of SensorML is the ablity to describe chains of processes.
-The SensorML ProcessChain and System types have a connections property that allows this chaining as decribed in section \ref{SMLref} .However strings are used to describe the connection.
-These connection strings are parsed to find out the inputs and outputs to connect up as described in SenseTile System section \ref.
-This is very error prone and the ability to check that connections are correct to a least some degree is desirable.
-Below are two fragents form SenseTile System and SenseTile Thermistor that descibe a connection form th
-
-\begin{lstlisting}
+One of the major features of SensorML is the ablity to describe processing chains.
+The SensorML ProcessChain and System types have a connections property that allows process chaining as decribed in section \ref{SMLref}.
+Strings are used to describe the connection.
+An example of this is shown in listing \ref{connectionListing} shows one of the connections from the SenseTile System as described in section \ref{SenseTileModelSec}
+\begin{lstlisting}[label=connectionListing,caption=SenseTile System connectionconvertToTemperature]
    <sml:connection name="convertToTemperature">
@@
    </sml:connection>
-\begin{lstlisting}
-
- BON specification can enforce some level of correctness in that inputs in connections must be inputs in the components contained in that chain. This is shown in the BON fragment below.
-
-\begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON AbstractProcess,label=BONListing]
-   deferred class AbstractProcess
+
+\end{lstlisting}
+
+The source and destination reference strings need to be parsed to find out the inputs and outputs to connect.
+This is very error prone and the ability to check that connections are correct to a least some degree is desirable.
+A BON specification can enforce some level of correctness in setting up the connections using preconditions (ensure).
+The source reference must be either an input of the process chain itself or an output of one of the contained components.
+In the BON ProcessChain class a feature AddLink checks that Links are legal before adding the Link to the connections feature.
+This is shown in the BON listing below.
+
+\begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON ProcessChaing]  
+effective class ProcessChain
      indexing
-     about: "base substitution group for all processes.";
-     title:        "AbstractProcess.";
+     about:        "Process formed by chaining sub-processes.";
+     title:        "ProcessChain.";
      author:       "Ciaran Palmer.";
@@
      organisation: "School of Computer Science and Informatics, UCD.";
-     date:         "2009/10/29.";
+     date:         "2010/04/01.";
      version:      "Revision: 1.00.";
 
+     inherit AbstractProcess
      feature
@@
      make
-       ->inputsIn:SEQUENCE[Input]
-       ->outputsIn:SEQUENCE[Output]
+       ->componentsIn:AbstractProcess
        require
-         inputsIn/=Void;
-         inputsIn/=Void;
+         componentsIn/=Void;         
        ensure
-         delta{inputs,outputs};
-         inputs = inputsIn;
-         outputs = outputsIn;
+         delta{components};
+         components = componentsIn;
        end
 
-     --optional       
-     inputs:SEQUENCE[Input]
-
-      --optional        
-     outputs:SEQUENCE[Output]
-
-      --optional       
-     parameters:Parameters
-
-     metaDataGroup:STRING
+     -- Collection of subprocesses that can be chained using connections      
+     components:SEQUENCE[AbstractProcess]
+
+     -- Links between processes
+     connections:SEQUENCE[Link]
+
+     -- Add a link with check that Link is legal
+     AddLink
+        ->linkIn:Link
+      require
+          linkIn /= Void;
+
+           -- check Link source is legal
+          (exists i:INTEGER such_that linkIn.source = Current.inputs[i]) or\
+          \(exists i:INTEGER and j:INTEGER such_that linkIn.source =\
+          \components[j].outputs[i]);
+
+          -- check Link destination is legal
+          (exists i:INTEGER such_that linkIn.destination = Current.outputs[i]) or\
+          \(exists i:INTEGER and j:INTEGER such_that linkIn.destination =\
+          \components[j].inputs[i]);
+
         ensure
-          Result = "metaDataGroup";
-       end
-
-     setInputs
-       ->inputsIn:SEQUENCE[Input]
-       require
-         inputsIn/=Void;
-       ensure
-         delta{inputs};
-         inputs = inputsIn;
-       end
-
-      setOutputs
-       ->outputsIn:SEQUENCE[Output]
-       require
-         outputsIn/=Void;
-       ensure
-         delta{outputs};
-         outputs = outputsIn;
-       end
-
-      setParameters
-       ->ParametersIn:SEQUENCE[Parameter]
-       require
-         ParametersIn/=Void;
-       ensure
-         delta{Parameters};
-         Parameters = ParametersIn;
-       end               
-   end --AbstractProcess
+           delta{connections};
+          (exists i:INTEGER such_that connections[i] linkIn);
+
+      invariant
+        components /= Void;
+
+  end --ProcessChain
  \end{lstlisting}
 
-SWE and BON Clusters.
 \section{SenseTile Thermistor BON SensorML Specification }
@@
 \chapter{ Conclusions and Future Work}
-This thesis analyses the role SensorML and plays in Sensor Web Networks.
-The UCD CASL SenseTile System is used a case study with the development of the SenseTile Web Service to access the SenseTile System sensor data.
-SensorML alone is not sufficient to model all aspects of Sensor Web Networks but is focused on providing a standard way to exchange sensor metadata, describing sensor input/output data and the processing of sensor data.
-This leads to examining other OGC SWE specifications, of which SensorML is one.
-The SWE SOS specification describes the operations for accessing sensor observations and it's concepts were used to develop the SenseTile Web Service.
-It uses SensorML provide sensor metadata to clients and the O\&M specification to encode observations.
-The OGC SWE initiative is trying to standardize a Web Service approach to accessing and processing sensor data.
-
+This thesis analyses the role of SensorML and SOS specifications in Sensor Systems.
+The UCD CASL SenseTile System is used as a case study for the use of these specifications.
+A SensorML model of SenseTile and a SenseTile Web Service prototype.
+
+SensorML address the issues of interoperabilty of Sensor Systems
+rich description of a sensor systems. Hard to think of any system that could no be described by SensorML. Weather systems to Geloloaction services. 
+Accurate elements. Processing data. Metadata.
+The supporting SWE Common data types are many data types.
+SensorML Frameworks.
+Precisely define its data.
+It also is bases for developement for Sensor Ontology work.\footnote{http://www.calebgoodwin.com/SEMANTIC.html}
 SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.
@@
 Maybe this is a flaw in the SenseTile model developed.
-
-Working in XML is error prone.
-SensorML connections type was a good example of where a graphical tool or another language, like BON,  that can be parsed should be used to check the connections.
-Running the model through the VastSensorMLEngine without a crash was only way other than by eye to see that the connections were set up correctly.
 
@@
 
-The idea of automatically obtaining a SensoML model and executing it depends on the SensorML Engine being used.
+The idea of automatically obtaining a SensorML model and executing it depends on the SensorML Engine being used.
 Only one Java implementation can be pointed.
@@
 
-the weaknesses of your approach
------------------------------------------
-SOS on the Processor Board. Is this right? RMI only useful if SenseTile Network is a lan.
+The SWE SOS specification describes the operations for accessing sensor observations and it's concepts were used to develop the SenseTile Web Service.
+It uses SensorML provide sensor metadata to clients and the O\&M specification to encode observations.
+Implement all functionality would be difficult.
+Maybe too heavy for SenseTile System.
+Simple querires.
+Or better a push system to send the data to the client as it is generated could be a better approach here.
 
@@
 flexible. There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
-what has been achieved
+what has been achieve
+
+SOS on the Processor Board. Is this right? RMI only useful if SenseTile Network is a lan.
+
+BON
+
+Working in XML is error prone.
+SensorML connections type was a good example of where a graphical tool or another language, like BON,  that can be parsed should be used to check the connections.
+Running the model through the VastSensorMLEngine without a crash was only way other than by eye to see that the connections were set up correctly.
 
@@
 ---------------
+SensorML sensor accuracy.
 Complete BON mapping and parser.
@@
 \bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
+\bibitem{Meyerref}Bertrand Meyer., Object-Oriented Software Construction, 1997
+
 \end{thebibliography}

### 2010-04-13 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 Reliably processing sensor observations retrieved from these heterogeneous systems is also problematic.
-A standard way to retrieve sensor data and a standard way to describe sensors are needed to address the current problem of sensor system heterogeneity.
+Standard ways to describe sensors and to retrieve sensor data and are needed to address the current problem of sensor system heterogeneity.
 
 Standards are being developed to address the above issues.
-Two such standards are the Sensor Model Language (SensorML)\cite{SMLref} specification and the Sensor Observation Service (SOS)\cite{SOSref} specification.
+Two such standards are the Sensor Model Language (SensorML) specification\cite{SMLref} and the Sensor Observation Service (SOS) specification\cite{SOSref}.
 SensorML is a proposed standard for sensor description and processing of sensor observations.
@@
 This thesis investigates the use of these SWE standards in developing sensor systems with emphasis on the use of SensorML and SOS specifications.
-
-Developing robust sensor systems is also a natural concern.
-There are now formal verification based software development processes \cite{Kiniryref} with the goal of developing very robust systems.
-Currently there is no widely available formal definition of sensors or sensor data processing.
-Formal sensor descriptions allows sensor system development with such a verification based process and their development is examined also in this thesis.
 
@@
 
-Buisness Object Notatation(BON)\cite{BONref} isi used to formalize SensorML sensor descriptions in this thesis.
-BON is a notation and a method for Object Oriented systems analysis and design.
+Currently there is no widely available formal definition of sensors or sensor data processing.
+There are now formal verification based software development processes \cite{Kiniryref} with the goal of developing very robust systems.
+Formal sensor descriptions allows sensor system development with such a verification based process.
+Business Object Notatation(BON)\cite{BONref} is used in \cite{Kiniryref} to formally specify the system being developed.
+BON is a notation and a method for analysis and design of Object Oriented systems.
+
+This thesis looks at how BON can be used to formally specify SensorML.
 The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to Java as described in \cite{Kiniryref} as part of developing a robust Sensor System.
@@
 \section{Sensor Web Enablement}\label{SWESec}
-As described in the introduction the SensorML and SOS specifications are developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
-This initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
-It is hoped that this allows for easier integration of sensor applications across sensor networks.
-The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
+As described in the introduction the SensorML and SOS specifications are developed as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
+This Open Geospatial Consortium (OGC) initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
+This allows for easier integration of sensor applications across sensor systems.
+SWE provides a framework of open standards for use with Web-connected sensors and sensor systems.
 There are seven main specifications currently:
@@
 
-SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations.
+SWE services based on the above specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations.
 For a good overview of SWE and it's use in Sensor Web Networks see \cite{SWEArchref}.
@@
 
+\subsection{SensorML Core Model}
 In SensorML sensors are modeled as processes that convert real phenomena to data.
@@
 The inputs and outputs represent data connections to other processes.
-Parameters represent inputs other than sensor measurements or outputs of other processes.
-They are used to configure the process or provide other inputs that are not sensor measurements.
+Parameters are used to configure the process or provide other inputs that are not sensor measurements.
 For example a parameter might be the latency time of a measurement needed for a calculation.
+
 Finally each process provides sensor metadata as modeled by the MetaDataGroup.
@@
 
+\subsection{Common Data}
 The SensorML specification currently contains the Sensor Web Enablement (SWE) Common specification.
@@
 Boolean models a property that can be true or false.
+
 These types contain a number of properties that further define the type.
 The uom property provides a unit of measurement that indicate how the value should be interpreted.
-A constraint property allows a value range or a enumerated list of values to be defined for the type.
+A constraint property allows a value range or a enumerated list of allowed values to be defined for the type.
 The quality property allows for a measure of its quality, for example the confidence level of a value being correct expressed as a percentage.
-Aggregation of these primitive data types is also provided by several types including DataRecord and DataArray.
+Aggregation of these data types is also provided by several types including DataRecord and DataArray.
 
@@
 
+\subsection{VAST SensorML Engine}\label{VastSensorMLEngineSec}
+The VAST Team at the University of Alabama in Huntsville (UAH) has developed an open source Java based SensorML Processing Engine.
+This software provides types that implement SensorML processes and process  chains.
+The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing.
+The code is found at \footnote{http://code.google.com/p/sensorml-data-processing} and is referred to as the VastSMLEngine in rest of the thesis.
+This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}.
+This code is found at \footnote{http://code.google.com/p/swe-common-data-framework}.
+
 \section{Sensor Observation Service}\label{SOSSec}
-SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref}.
-The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.
+The SOS specifications defines a web service interface for the retrieval of sensor descriptions and the sensor observations from sensor systems.
 It organizes related sensor observations into a collection called an observation offering.
+SensorML is used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref} for sensor descriptions.
 
@@
 SOS instances act as access points to a Sensor Web Network and provide the Web Service for retrieval of sensor observations.
-The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context.
+The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operational context.
 \begin{figure}[h]
@@
 The sensor observations contained in  GetObservation and InsertObservation operations are encoded according to the OGC O\&M Observations \cite{OMref}.
-Sensor systems metadata contained in DescribeSensor and RegisterSensor is defined in SensorML or TML.
+Sensor systems metadata contained in DescribeSensor and RegisterSensor is defined in SensorML.
 
@@
 
-The OGC Catalog Service is not explored in this thesis.
-The SOS Operation Context diagram shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog.
+The OGC Catalog Service shown in the SOS Operation Context diagram \ref{fig:SOSoperationContext} is not explored in this thesis.
 The Data Consumer uses a CS-W catalog to find SOS instances.
-The GetCapabilites are then used toward each of the SOS instances returned to find sensor systems that are of interest to the Data Consumer.
-The CS-W catalog is covered in a separate specification \cite{OGCcatref}.
-This not covered in any real detail in the SOS specification.
-
-There several optional SOS operations not covered in this thesis. The SOS refers to these as enhanced operations. They are GetResult,
-GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest, DescribeObservationType, and DescribeResultModel.
+The GetCapabilites operation is then used toward each of the SOS instances found to find sensor offerings that are of interest to the Data Consumer.
+The CS-W catalog is covered in a separate specification \cite{OGCcatref} and is not covered in any real detail in the SOS specification.
+
+There several optional SOS operations not covered in this thesis.
+The SOS refers to these as enhanced operations.
+They are GetResult, GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest, DescribeObservationType, and DescribeResultModel.
 
@@
 
-Thermistor?
-
-\section{UCD Sensor Board Driver}\label{SensorBoardDriverSec}
-
-\section{VAST SensorML Engine}\label{VastSensorMLEngineSec}
-The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source Java based SensorML Processing Engine.
-This software provides types that implement SensorML processes and process  chains.
-The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing.
-The code is found at \footnote{http://code.google.com/p/sensorml-data-processing} and is referred to as the VastSMLEngine in rest of the thesis.
-This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}.
-This code is found at \footnote{http://code.google.com/p/swe-common-data-framework}.
+The TMP175 and TMP75 are digital temperature sensors
+that are optimal for thermal management and thermal
+protection applications. The TMP175 and TMP75 are
+Two-Wire and SMBus interface-compatible, and are
+specified over a temperature range of −40°C to +125°C.
+
+\subsection{UCD Sensor Board Driver}\label{SensorBoardDriverSec}
 
@@
 The two main entities in the SOS Specification, the Sensor Data Provider and the SOS,  are used to structure the SenseTile Web Service.
-These entities are implemented as seperate software components that together provide the service.
+These entities are implemented as separate software components that together provide the service.
 The Sensor Data Provider component will be referred to as the DataProvider in this design.
 
-The two components are run as seperate processes on the SenseTile processor unit to allow a more flexible network architecture.
+The two components are run as separate processes on the SenseTile processor unit to allow a more flexible network architecture.
 This network architecture is described in more detail in section \ref{subsec:SenseTile WebService Network Architecture}.
@@
 The following sections describe the design of the DataProvider and SOS components as well as the scenarios they support.
-
-\newpage
-\subsection{Supported Scenarios}
-This section provides UML sequence diagrams to show the how the DataProducer and SOS interact to provide the sensor to the DataConsumer client.
-
-\subsubsection{Generate Observations from Sensor Data}
-The DataProducer accesses data from the sensor board 
-\begin{figure}[h]
-\centering
-\framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
-\caption{Generate Observation}\label{fig:InsertObs}
-\end{figure}
- \begin{enumerate}
-\item -
-\item -
-\item -
-\end{enumerate}
-
-\subsubsection{Get Sensor Description from SOS}
-\begin{figure}[h]
-\centering
-\framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
-\caption{Find SOS sensor}\label{fig:GetMeta}
-\end{figure}
- \begin{enumerate}
-\item -
-\item -
-\end{enumerate}
-\subsubsection{Get  Sensor Observation from SOS}
-\begin{figure}[h]
-\centering
-\framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
-\caption{Get Observation from Sensor}\label{fig:GetObs}
-\end{figure}
- \begin{enumerate}
-\item -
-\end{enumerate}
 
@@
 The SOS Specification describes the use of a Proxy SOS eeicpr fix. This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.
+
+\newpage
+\subsection{SenseTile Web Service Scenarios}
+This section contains UML sequence diagrams to show the how the DataProducer and SOS interact to provide the sensor to the DataConsumer client.
+There are three main scenarios supported by the SenseTile Web Service
+
+\subsubsection{Generate Observations from Sensor Data}
+The DataProducer accesses data from the sensor board and updates the SOS with sensor observations.
+\begin{figure}[h]
+\centering
+\framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
+\caption{Generate Observation}\label{fig:InsertObs}
+\end{figure}
+ \begin{enumerate}
+\item At start up the DataProducer registers the SensorBoard System and the Converter System with the SOS in separate RegistorSensor operations.
+The SOS creates sensor offerings for each of the Systems and stores the SensorML descriptions of the Systems provided in the RegistorSensor operation.
+It then returns a unique id (URN) in the RegistorSensor response for use by the DataProducer for subsequent communication with the SOS.
+\item The DataProducer reads Sensor Board data packets from the SensorBoardDriver.
+\item The DataProducer updates the SOS with sensor observations with the InsertObservation operation.
+\end{enumerate}
+
+\subsubsection{Get Sensor Description from SOS}
+The SOS allows DataConsumers to obtain SensorML descriptions of the SesneTile Systems.
+\begin{figure}[h]
+\centering
+\framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
+\caption{Get Sensor Descriptions}\label{fig:GetMeta}
+\end{figure}
+ \begin{enumerate}
+\item The DataConsumer requests information on sensor offerings provided by the SOS using the GetCapabilities operation.
+\item Using the information in the GetCapabilites response the DataConsumer requests a description of a sensor system that is providing an offering using the DescribeSensor operation.
+The SOS returns the SensorML description of the sensor system to the DataConsumer.
+\end{enumerate}
+
+\subsubsection{Get  Sensor Observation from SOS}
+\begin{figure}[h]
+\centering
+\framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
+\caption{Get Observation from Sensor}\label{fig:GetObs}
+\end{figure}
+ \begin{enumerate}
+\item Using a sensor identity received in a GetCapability response the DataConsumer requests a sensor observation. The SOS returns the latest O\&M observation to the DataConsumer.
+\end{enumerate}
 
@@
 \chapter{ Conclusions and Future Work}
-This thesis analyses the role of SensorML and SOS specifications in Sensor Systems.
+This thesis analyses the role of the SensorML and the SOS specifications in Sensor Systems.
 The UCD CASL SenseTile System is used as a case study for the use of these specifications.
-A SensorML model of SenseTile and a SenseTile Web Service prototype.
-
+A SensorML model of the SenseTile system is developed to evaluate SensorML.
+A SenseTile Web Service prototype is used to evaluate the SOS specification which uses SensorML to provide sensor descriptions to clients.
+
+SensorML provides a good model of sensor systems in terms of a an intuitive process concept.
+to describe of a sensor systems.
+It is hard to think of any system that could no be described by SensorML due to number of data types that can.
+Weather systems to Geloloaction services.
 SensorML address the issues of interoperabilty of Sensor Systems
-rich description of a sensor systems. Hard to think of any system that could no be described by SensorML. Weather systems to Geloloaction services. 
-Accurate elements. Processing data. Metadata.
-The supporting SWE Common data types are many data types.
-SensorML Frameworks.
-Precisely define its data.
-It also is bases for developement for Sensor Ontology work.\footnote{http://www.calebgoodwin.com/SEMANTIC.html}
+
+For descriptions there good Metadata.
+
+SensorML provides good support for the reliable processing of sensor data.
+The data types have several properties that can reference definitions to precisely describe their meaning.
+The accuracy of the sensor data can be modelled and provided used by systems to process the data with these variables include to get a better result.
+
+Processing data. 
+SensorML Frameworks. DataProducer.
+
+It also is bases for development for Sensor Ontology work.\footnote{http://www.calebgoodwin.com/SEMANTIC.html}
 SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.
@@
 Or better a push system to send the data to the client as it is generated could be a better approach here.
+
+SOS Scalable hierarcy of instances.
 
@@
 \bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
-\bibitem{Meyerref}Bertrand Meyer., Object-Oriented Software Construction, 1997
+\bibitem{Meyerref}Bertrand Meyer., "Object-Oriented Software Construction", 1997
 

### 2010-04-13 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \begin{abstract}
-This thesis evaluates the use of SensorML and SOS in the development of Sensor Systems.
-The SenseTile system is used as a case study for this evaluation.
-A SenseTile web service is developed to access SenseTile sensor data based on a SensorML model of SenseTile.
-Formalizing senor descriptions is also looked using BON to SensorML mapping.
+This thesis evaluates the Sensor Model Language(SensorML) and Sensor Observation Service(SOS) specifications.
+These specifications are intended for use in the development of open sensor systems.
+The UCD CASL SenseTile sensor system is used as a case study for this evaluation.
+A SensorML model of SenseTile is developed.
+A SenseTile Web Service is also developed based on the SenseTile SensorML model and the SOS specification.
+This Web Service provides access to SenseTile sensor metadata and sensor observations.
+The thesis finally examines formalizing SensorML using a formal specification language called Business Object Notation.
 
@@
 Reliably processing sensor observations retrieved from these heterogeneous systems is also problematic.
-Standard ways to describe sensors and to retrieve sensor data and are needed to address the current problem of sensor system heterogeneity.
+Standard ways to describe sensors and to retrieve sensor data are needed to address the problem of sensor system heterogeneity.
 
@@
 Two such standards are the Sensor Model Language (SensorML) specification\cite{SMLref} and the Sensor Observation Service (SOS) specification\cite{SOSref}.
-SensorML is a proposed standard for sensor description and processing of sensor observations.
-SensorML specifies models and XML Schema for describing sensors, the process of sensor measurement and the post processing of sensor observations.
-The SOS specification is a proposed standard for retrieving sensor information and sensor observations using Web Services.
+SensorML is a proposed standard for both describing sensors and describing the processing of sensor observations.
+The SOS specification is a proposed standard for retrieving sensor descriptions and sensor observations using Web Services.
 
@@
 The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as the SensorML and SOS specifications.
-This thesis investigates the use of these SWE standards in developing sensor systems with emphasis on the use of SensorML and SOS specifications.
-
-A real sensor system called SenseTile[ref] is used as a case study in this thesis.
+This thesis investigates the use of these SWE SensorML and SOS standards in developing sensor systems.
+
+A real sensor system called SenseTile[ref] is used as a case study for the evaluation of SensorML and SOS specifications.
 SenseTile is a sensor and processing package that contains motion sensors, temperature sensors, audio sensors, pressure sensors and light level sensors among others.
@@
 
-This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board .
+This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board.
 Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor descriptions and sensor observations.
@@
 Currently there is no widely available formal definition of sensors or sensor data processing.
-There are now formal verification based software development processes \cite{Kiniryref} with the goal of developing very robust systems.
-Formal sensor descriptions allows sensor system development with such a verification based process.
-Business Object Notatation(BON)\cite{BONref} is used in \cite{Kiniryref} to formally specify the system being developed.
+Formal sensor descriptions allows sensor system development to use a formal verification based process.
+A verification based software development processes is described in \cite{Kiniryref} which has the goal of developing very robust software systems.
+Business Object Notatation(BON)\cite{BONref} is used in this process to formally specify the system being developed.
 BON is a notation and a method for analysis and design of Object Oriented systems.
-
 This thesis looks at how BON can be used to formally specify SensorML.
-The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to Java as described in \cite{Kiniryref} as part of developing a robust Sensor System.
+The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to JML\footnote{http://www.eecs.ucf.edu/~leavens/JML/} and then to Java\footnote{http://www.java.com/}, as described in \cite{Kiniryref}.
 
@@
 \section{Sensor Web Enablement}\label{SWESec}
-As described in the introduction the SensorML and SOS specifications are developed as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
-This Open Geospatial Consortium (OGC) initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
+As described in the introduction the SensorML and SOS specifications are developed by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}.
+The initiative is establishing interfaces and protocols to enable a standardized Sensor Web.
 This allows for easier integration of sensor applications across sensor systems.
@@
 \begin{itemize}
-\item  SensorModel Language(SensorML) - models and schema for sensor systems and processes surrounding measurements
-\item  Sensor Observation Service (SOS) - standard web interface for accessing observations
-\item  Observations \& Measurements (O\&M) - models and schema for packaging observation values
+\item  SensorModel Language(SensorML) - models and schema for sensor system description and sensor measurement processing
+\item  Sensor Observation Service (SOS) - standard web interface for accessing sensor observations
+\item  Observations \& Measurements (O\&M) - models and schema for packaging sensor observation values
 \item  Transducer Markup Language (TML) - models and schema for multiplexed data from sensor systems
-\item  Sensor Planning Service (SPS) - standard web interface for tasking sensor systems and model and requesting acquisitions
+\item  Sensor Planning Service (SPS) - standard web interface for tasking sensor systems
 \item  Sensor Alert Service (SAS) - standard web interface for publishing and subscribing to sensor alerts
@@
 
-SWE services based on the above specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations.
+SWE services based on the above specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control sensor meaurement.
 For a good overview of SWE and it's use in Sensor Web Networks see \cite{SWEArchref}.
-Along with with the SensorML and SOS specification, the O\&M specification is also used in developing the SenseTile Web Service as it defines the encoding of sensor observations.
+The SensorML and SOS specifications, as mentioned, are used to develop the SenseTile Web Service.
+The O\&M specification is also used in developing the SenseTile Web Service.
+This specification defines the encoding of sensor observations and is used by the SOS specification.
 
@@
 \item Provide accuracy information of sensors measurements.
-\item Provide an explicit description of the process by which an observation was obtained.
 \item Provide an executable process chain for deriving new data on demand.
@@
 
-\subsection{SensorML Core Model}
-In SensorML sensors are modeled as processes that convert real phenomena to data.
+XML Schema is used to specify the XML encoding of SensorML in the SensorML specification.
+This XML schema is based on two conceptual UML\footnote{http://www.uml.org/} models, the SensorML Conceptual Model and the SWE Common Conceptual Model, that describe SensorML.
+These models are explained in the following sections.
+
+\subsection{SensorML Conceptual Model}
+In SensorML sensors are modeled as processes that convert physical phenomena to data.
 Processes take inputs, process them and result in one or several outputs.
-For example, a process can take a measurement generated by a thermistor which might be a resistance value and transform it to a celcius value.
+For example, a process can take a measurement generated by a thermistor which might be an electrical resistance value and transform it to a Celsius value.
 Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially.
@@
 \centering
-\framebox{\scalebox{0.65}{\includegraphics{SMLConceptualModel.png}}}
+\framebox{\scalebox{0.62}{\includegraphics{SMLConceptualModel.png}}}
 \caption{SensorML Conceptual Model}\label{fig:SMLConceptualModel}
 \end{figure}
-
-Processes in SensorML derive from AbstractProcess with it's inputs, output, parameters and metaDataGroup properties.
-The inputs and outputs represent data connections to other processes.
-Parameters are used to configure the process or provide other inputs that are not sensor measurements.
-For example a parameter might be the latency time of a measurement needed for a calculation.
-
+\newpage
+
+\subsubsection{Processes}
+The AbstractProcess models processes in SensorML and contains input, output, parameter and metaDataGroup properties.
+The input and output properties represent sensor data measurements or connections to other processes.
+The parameter property is used to configure the process or provide other inputs that are not sensor measurements.
+For example a parameter might be the latency time of a measurement needed for a calculation performed by the process.
 Finally each process provides sensor metadata as modeled by the MetaDataGroup.
+The MetaDataGroup contains properties such as capabilities of the sensor, contact information and documentation references.
+It is generally used to assist humans in understanding the sensor system described and not used when executing a process.
+AbstractProcess derives from AbstractFeature which contains a name and description properties.
 
@@
 The non-physical process is one where  location, position etc.
-is not relevant to the process.
-Physical processes provide physical information, such as location, which can be used in processing if required.
-The ProcessModel is a non-physical process while the Component is a physical process and contains extra properties for defining the needed spatial or temporal information.
-Both process types contain the processMethod property which points to the methodology for transforming inputs to outputs.
-
+is not relevant to the process and is modeled by the ProcessModel type.
+Physical processes provide physical information, such as location, which can be used in processing if required and is modeled by the Component type.
+Both process types contain the method property.
+The method property is a ProcessMethod type describes the methodology for transforming the process inputs to outputs.
+
+\subsubsection{Process Chains}
 There is a similar division of the process chains into the non-physical ProcessChain and physical System.
-Both process chains can contain a mixture of ProcessModel and Component process if desired.
+Both process chains can contain a mixture of ProcessModel and Component processes if desired.
 Process chains have inputs and outputs which define the beginning and end of the processing chain.
-They are points of connection to other processes and process chains if needed.
-The connection property contained in the ProcessChain and System is a list of Link type.
+They are points of connection to other processes and process chains.
+The connection property contained in the CompositePropertGroup referenced by the  ProcessChain and System types is a sequence of type Link.
 The Link type contains source and destination properties that reference inputs and outputs of the processes in the process chain thus connecting the processes.
 
-\subsection{Common Data}
+\subsubsection{Linkable Properties}
+SensorML uses XML Linking Language (XLink) \footnote{http://www.w3.org/TR/xlink} to support hypertext referencing in SensorML XML documents using URNs\footnote{http://tools.ietf.org/html/rfc2141} and URIs\footnote{http://tools.ietf.org/html/rfc3986}.
+This allows SensorML models to be distributed over several XML documents.
+XLink attributes are also used to reference on-line unit definitions used to define a sensor measurement.
+
+\subsection{SWE Common Conceptual Model}\label{SMLsectionSWE}
 The SensorML specification currently contains the Sensor Web Enablement (SWE) Common specification.
 This specification defines basic types and data encodings used by the SWE specifications.
-It describes both primitive data types as well as aggregate types such as arrays and records.
+It describes both simple data types as well as aggregate types such as arrays and records.
 There are a large number of types defined in this specification, too many to show here.
@@
 
-The data types in \ref{fig:SWESimpleConceptualModel} are scalar primitive values.
-They are the basis for the inputs, outputs and parameters in SensorML processes.
-Quantity models a floating point number.
-Count models a counting value or integer.
-Boolean models a property that can be true or false.
-
-These types contain a number of properties that further define the type.
-The uom property provides a unit of measurement that indicate how the value should be interpreted.
+The data types in \ref{fig:SWESimpleConceptualModel} model scalar primitive values.
+They are used to defined the type of day used by the inputs, outputs and parameters in SensorML processes.
+Quantity, for example, models a floating point number and can be used to represent Celsius temperature values.
+The data types in SWE Common derive from AbstractDataComponent which contains naming and description properties.
+
+To allow a more reliable processing of sensor data the types contain a number of properties to provide metadata about the value carried in the simple type.
+The uom property provides a unit of measurement reference that indicates how the value should be interpreted.
 A constraint property allows a value range or a enumerated list of allowed values to be defined for the type.
-The quality property allows for a measure of its quality, for example the confidence level of a value being correct expressed as a percentage.
-Aggregation of these data types is also provided by several types including DataRecord and DataArray.
-
-SensorML uses XML Linking Language (XLink) \footnote{http://www.w3.org/TR/xlink} to support hypertext referencing in SensorML XML documents using URNs and URIs.
-This allows SensorML models to be distributed over several documents.
-XLink attributes are also used to reference on-line unit definitions used to define a sensor measurement.
-
-The XML encoding of the SensorML model is not described here.
-The OGC SWE specification uses XML Schema to define the allowable SensorML XML structure.It is noted here that sml is the XML namespace for SensorML and swe is the XML namespace for SWE Common.
+The quality property provides for a measure of the quality of the value.
+For example the confidence level of a value being correct can be expressed as a percentage.
+
+Aggregation of these data types is also provided by several types including the DataRecord and DataArray types.
 
 \subsection{VAST SensorML Engine}\label{VastSensorMLEngineSec}
-The VAST Team at the University of Alabama in Huntsville (UAH) has developed an open source Java based SensorML Processing Engine.
+The VAST Team at the University of Alabama in Huntsville (UAH) has developed an open source Java based SensorML processing engine.
 This software provides types that implement SensorML processes and process chains.
 The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing.
-The code is found at \footnote{http://code.google.com/p/sensorml-data-processing} and is referred to as the VastSMLEngine in rest of the thesis.
-This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}.
+The code for the SensorML processing engine is found at \footnote{http://code.google.com/p/sensorml-data-processing}.
+This engine is used in the SenseTile Web Service and is referred to as the VastSMLEngine in rest of the thesis.
+
+The SensorML processing engine is dependent on an implementation of the SWE Common data types described in the section \ref{SMLsectionSWE}.
 This code is found at \footnote{http://code.google.com/p/swe-common-data-framework}.
@@
 It organizes related sensor observations into a collection called an observation offering.
-SensorML is used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref} for sensor descriptions.
-
-The SOS specification uses four main entities to describe the SOS's operation.
-The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself.
-Data Consumers are clients of the SOS that request sensor observations and sensor metadata.
-Data Providers are entities have to access to the sensor and have enough processing power to generate an XML encoding of the sensor's observations.
+The offering information is requested by clients to get the identity of senors they are interested in getting observations from.
+
+The SOS specification uses four main entities to describe the SOS's operation. The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself. 
+\begin{itemize}
+\item Data Consumers are clients of the SOS that request sensor observations and sensor metadata. 
+\item Data Providers are entities read the observations from the sensors and have enough processing power to generate an XML encoding of the sensor's observations.
 These observations are sent over a network to the SOS.
-The OGC Catalog is used to find SOS instances.
-SOS instances act as access points to a Sensor Web Network and provide the Web Service for retrieval of sensor observations.
+\item OGC Catalogs are used to locate SOS instances. 
+\item SOS instances act as access points to a Sensor Web Network and provide the Web Service for retrieval of sensor descriptions and sensor observations.
+\end{itemize}
 The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operational context.
@@
 
-The diagram above shows the core and transactional operations provided by an SOS.
+The diagram above shows the core and transactional operations provided by an SOS used by DataConsumers and DataProducers.
 The core operations are GetCapabilities, DescribeSensor and GetObservations.
 These are mandatory operations used by Data Consumers to request sensor information.
-The core operations perform the following function:
+The core operations provide the following service:
  \begin{itemize}
@@
  \begin{itemize}
-\item RegisterSensor - operation to register the Data Provider with the SOS.
-\item InsertObservation - operation to update SOS with a sensor observation.
+\item RegisterSensor -  register the Data Provider with the SOS.
+\item InsertObservation - update SOS with a sensor observation.
 \end{itemize}
 
-These operations are encoded in XML and are intended to be carried http  operations.
+These operations are encoded in XML and are intended to be carried in http\footnote{http://www.w3.org/Protocols/}  operations.
 The sensor observations contained in GetObservation and InsertObservation operations are encoded according to the OGC O\&M Observations \cite{OMref}.
-Sensor systems metadata contained in DescribeSensor and RegisterSensor is defined in SensorML.
+Sensor systems metadata used in DescribeSensor and RegisterSensor operations is provided with SensorML.
 

### 2010-04-13 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 It organizes related sensor observations into a collection called an observation offering.
-The offering information is requested by clients to get the identity of senors they are interested in getting observations from.
+The observation offering information is requested by clients to obtain the identity of senors they are interested in getting observations from.
 
@@
 \section{SenseTile}
-A case study of the use of SensorML in sensor networks is based on the SenseTile System. SenseTile is a general purpose sensor system developed at UCD CASL \footnote{http://kind.ucd.ie/
-products/opensource/SenseTile}.
+SenseTile \footnote{http://kind.ucd.ie/
+products/opensource/SenseTile} is used as case study of the evaluation of SensorML and SOS.
+SenseTile is a general purpose sensor system developed at UCD CASL\footnote{http://casl.ucd.ie/}.
 The SenseTile System is used to investigate issues arising with large scale sensor networking.
@@
 \centering
-\framebox{\scalebox{0.65}{\includegraphics{SenseTileDescription.png}}}
+\framebox{\scalebox{0.62}{\includegraphics{SenseTileDescription.png}}}
 \caption{SenseTile System}\label{fig:SenseTileDescription}
 \end{figure}
+\newpage
+
+The datastore is a multi-terabyte scale datastore . The sensor data is retrieved and stored here where it is used in further processing of the sensor data.
+
+The compute farm is a large-scale compute farm that supports Linux, Solaris, and Mac OS and is where further processing of sensor data is performed.
+
 The sensor platform is called the SenseTile Unit and is composed of one or more SenseTile Sensor Boards paired with a SenseTile Processor Unit.
@@
 The Sensor Board has a USB connection used for communication with the Processor Unit.
+
 There are over a dozen sensors on the SenseTile Sensor Board including a thermistor and light sensor.
 Sensors are mounted on the Sensor Board and more sensors can be added using the USB connection.
-
-This SenseTile unit is used here as case study on the use of SensorML, though SensorML can be used in other the parts of the SenseTile system.
-As the processor unit is a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible.
-The part SensorML plays in this Web Service is explored in the context of the SenseTile system.
-
-The TMP175 and TMP75 are digital temperature sensors
-that are optimal for thermal management and thermal
-protection applications. The TMP175 and TMP75 are
-Two-Wire and SMBus interface-compatible, and are
-specified over a temperature range of −40°C to +125°C.
+The SensorBoard and its sensors are the main focus of the SensorML model in this thesis and as the processor unit is a relatively powerful computer, providing a lightweight Web Service on the SenseTile Processor Unit itself is feasible.
 
 \subsection{UCD Sensor Board Driver}\label{SensorBoardDriverSec}
+
+The researchers at UCD CASL have developed a custom asynchronous packet based protocol to manage the SenseTile Sensor Board and to read sensor data from the Sensor Board.
+A Java driver is developed also and it creates and parses the SenseTile Sensor Board streaming communication packets.
+The SenseTile Web Service uses this Java driver API to access the Sensor Board sensor data.
+This driver is referred to as the SensorBoardDriver in this thesis.
+
+There is a Sensor Board simulator also developed by the UCD CASL group and it is used to develop the SenseTile Web Service.
 
@@
 It is reversible as changes in the source code can be easily mapped back into the design and analysis models.
-It supports design by contract\ref{meyerref} where a function will guarantee a postcondition if a specified precondition is fulfilled by the caller of the function.
+It supports design by contract\cite{Meyerref} where a function will guarantee a postcondition if a specified precondition is fulfilled by the caller of the function.
 BON support for design by contract allows for a formal description of a system to be developed.
@@
 Central to the design is a SensorML model of the SenseTile System.
-The section SensorML Model of SenseTile covers the details of this model.
-This model acts as both metadata about the SenseTile system and provides a description of how to process the sensor data for the SenseTile Web Service.
-The Web Service design is based on the SOS specification and allows access to the SensorML model of SenseTile sensors as well as SenseTile sensor observations.
+
+The section SensorML Model of SenseTile covers the details of the SenseTile SensorML model.
+This model acts as both metadata about the SenseTile system and provides a description of how to process the sensor data.
+
+The Web Service design is based on the SOS specification and allows access to the SensorML description of SenseTile sensors as well as SenseTile sensor observations.
 Two SOS specification concepts as described in section \ref{SOSSec}, the Data Producer and the SOS itself, structure the design and are described in the SenseTile Web Service section.
@@
 The SenseTile Sensor Board and its sensors is modeled in the most detailed.
-The model of the Sensor Board is placed in a context to examine SensorML process chaining where processing of Sensor Board data is performed in a separate SensorML process.
-The SensorML Systems and Components that are used to model the system are shown in a  block diagram in figure \ref{fig:SensorML_SenseTile_System_comp}.
+The model of the Sensor Board is placed in a context to examine SensorML process chaining where further processing of Sensor Board data is performed in a separate SensorML process.
+The SensorML Systems and Components that are used to model the system are shown in a block diagram \ref{fig:SensorML_SenseTile_System_comp}.
 
@@
 \centering
-\framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
+\framebox{\scalebox{0.62}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
 \caption{SenseTile System SensorML Block Diagram}\label{fig:SensorML_SenseTile_System_comp}
@@
 The Sensor Board system contains metadata about the SenseTile Sensor Board and is a container for SensorML Components that model the Sensor Board's sensors.
-
+\newpage
 The block diagram shows the physical phenomena measured by the sensors being feed as inputs to the Components in the Sensor Board System.
@@
 The digital numbers from the sensors are fed into the relevant conversion process that produce a meaningful value such as Celsius or Lumen.
-This allows the SenseTile Web Service provide the raw data from the sensors or the data converted to meaningful values to clients.
-
-Only the Thermistor is modeled in this thesis.
-The  CelsiusConv Component converts the digital number form the Thermistor Component to a Celcius value.
-The same approach can applied to the other senors though for some components no conversion may be necessary.
-
-A description of the SenseTile Systems and Components used in the SenseTile model is provided in the following sections.
+This allows the SenseTile Web Service to provide the raw data from the sensors or provide the data converted to meaningful values to clients.
+
+Of the SenseTile sensors only the Thermistor is modeled along with the  CelsiusConv Component that converts the digital number received from the Thermistor Component to a Celcius value.
+However the same approach can applied to the other senors though for some components no conversion process may be needed.
+
+A description of the  import aspects of the SenseTile Systems and Components used in the SenseTile model is provided in the following sections. The complete model is provided in appendix \ref{appenA}.
 
 \subsection{SenseTile System}
-The SenseTile System is a SensorML System as described in section \ref{SMLsection} and is a container for the Sensor Board and Conversion Systems connecting their inputs and outputs as appropriate.
-The below XML fragment shows how SensorML encodes the containment of these Systems.
+The SenseTile System is a SensorML System as described in section \ref{SMLsection} and is a container for the Sensor Board and Conversion Systems connecting their inputs and outputs.
+The below listing shows the XML encoding of the Systems component property.
+\newline 
 \lstset{language=XML,basicstyle=\scriptsize,frame=single}
-\begin{lstlisting}
+\begin{lstlisting}[label=componentListing,caption=SenseTile System Components List]
 <sml:components>
@@
 \end{lstlisting}
-The components and ComponentList elements contain a list of component tags.
-These component elements name the component which may be another SensorML System  Component.
-The component tag also has an XLink href used to reference the location of the SensorML definition of the component.
+
+The components and ComponentList elements contain a list of component elements.
+These component elements have a name attribute to reference the component within the SenseTile System.
+For example in this case the Sensor Board System is called sensorBoardSystem.
+The component can be a SensorML process or a processes chain.
+
+The component element also has an XLink href attribute that is used to reference the location of the SensorML definition of the component.
 In this case the SensorML descriptions of the SensorBoard System and the Conversion System are in separate external files rather than in-line in the SenseTile System description.
 The use of XLink in SensorML is described in the SensorML section \ref{SMLsection}.
-
-The SenseTile System connects up the inputs and outputs of the SensorBoard and Conversion Systems using SensorML connections.
-The below SensorML fragment shows how the temperature output from the SensorBoard System is connected to the input of the Conversion System.
+As no path or URI is used in the hrefs the files must be in the current directory.
+
+The SenseTile System connects up the inputs and outputs of the SensorBoard and Conversion Systems using connections.
+The below SensorML fragment shows how the SenseTile Systems connections.
+Here the temperature output from the SensorBoard System is connected both to the input of the Conversion System and the output of the SenseTile System itself.
+Also the output of the Conversion System is connected to a different output of the SenseTile System.
 
@@
 The connection name attribute is used here to give a meaningful name to the connection.
-The Link type, as described in \ref{SMLsection}, is encoded with the Link tag.
-The contained source and destination elements contain a ref that indicates the components output and input to use.
-Only one connection is shown here but it is contained in a elements that allow a sequence of connection elements.
+The Link type, as described in \ref{SMLsection}, is encoded with the Link element.
+The Link source and destination elements contain a ref attribute that indicates the components output and input that are connected based on the component name and the input or output names in the process referenced by component name.
+The special "this" string refers to the SenseTile System itself and is used to reference it's inputs and outputs.
 
@@
 
-This System models the Sensor Board and contains meta data about the Sensor Board. It connects up the inputs and outputs the sensors.
+This System models the Sensor Board and contains meta data about the Sensor Board.
+It connects up the inputs and outputs of the SenseTile sensors components.
+In this case it names the Thermistor Component "thermistor" and references the Thermistor SensorML file, Thermistor.xml.
+It connects up the Components inputs and outputs as shown in the below XML fragment.
+
+\begin{lstlisting}[label=SensorBoardConnectionListing,caption=Sensor Board System Components and Connections List]
+         <sml:components>
+            <sml:ComponentList>
+               <sml:component name="thermistor"  xlink:href="Thermistor.xml"/>
+            </sml:ComponentList>			   
+         </sml:components>
+
+         <!-- Connections -->
+         <sml:connections>
+            <sml:ConnectionList>
+	   <sml:connection name="ambientTemperatureInput">
+                  <sml:Link>
+                     <sml:source ref="this/inputs/ambientTemperatureInput"/>
+                     <sml:destination ref="thermistor/inputs/thermistorInput"/>
+                  </sml:Link>
+               </sml:connection>
+	   <sml:connection name="temperatureDN">
+                  <sml:Link>
+                     <sml:source ref="thermistor/outputs/thermistorOutput"/>
+                     <sml:destination ref="this/outputs/temperatureDNOutput"/>
+                  </sml:Link>
+               </sml:connection>
+            </sml:ConnectionList>
+         </sml:connections>
+\end{lstlisting}
 
@@
 This is an indivisible process and can have a different location information than the Sensor Board if required.
-The SensorML Component allows the capabilities of the Thermistor to be described.
+The SensorML Component allows the capabilities of the Thermistor to be described as metadata.
 For example the TMP175 is  specified for operation over a temperature range of −40°C to +125°C.
@@
 
-\begin{lstlisting}
+\begin{lstlisting} [label=ThernmistorMetadataListing,caption=Thermistor metadata example]
 <swe:field name="TemperatureRange" 
@@
 Its output is a digital number from the sensor.
-The range is constrained to the allowed value range from the Thermistor.
-The following fragment shows how the input and outputs for the Thermistor look in SensorML:
-
-\begin{lstlisting}
+This is modeled as a Count type with the range constrained to the allowed value range from the Thermistor.
+The following XML fragment shows how the input and outputs for the Thermistor look in SensorML:
+
+\begin{lstlisting} [label=ThermistorListing,caption=Thermistor inputs and outputs]
 <sml:inputs>
@@
 \subsection{Converter System}
-The Converter System is a container for the processes that convert the sensor data to a meaningful value. It this case it connects up the inputs and outputs the CelsiusConv Component. 
+The Converter System is a container for the processes that convert sensor data to a meaningful value. It this case it connects up the inputs and outputs of the CelsiusConv Component. 
 \subsection{Celsius Converter Component}
 
-The SensorML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celsius value.
-It similar to the Thermistor SensorML description but its inputs and outputs are different.
+The SensorML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real Celsius value.
+It similar to the Thermistor SensorML but its inputs and outputs are have different types.
 Its process method performs a simple calculation to perform the Celsius conversion.
@@
 
-A  SensorML Model description of a SenseTile Sensor Board and Thermistor was developed. It was
-run on a SensorML engine. Appendix shows the model developed.
-
-Developed lightweight Web Service to explore the use of SensorML in Sensor Networks.
-running prototype tested with UCD SensorBoard Simulator code.
-
-Tested on a low power Atom PC that is used as SenseTile Processing Unit.
-
-BON mapping developed for main SensorML types. A BON description of SenseTile SensorML Model
-was developed based on the mapping.
+A  SensorML Model of the UCD CASL SenseTile sensor system was developed to evaluate SensorML.
+The SenseTile Sensor Board and Thermistor sensor were the focus of the model.
+Appendix \ref{appenA} shows the full SensorML SenseTile model developed.
+
+A lightweight Web Service was developed to explore the use of SensorML and SOS specifications in sensor systems.
+The Web Service is implemented in Java and is built on AXIS2.0, the Spring framework, the VASTSensorMLEngine and the UCD CASL SensorBoardDriver.
+It was tested against SenseTile SensorBoard Simulator classes using a basic test client.
+
+A BON specification was developed for the core SensorML types and the SWE Common Data types used in the SenseTile SensorML model.
+The BON specification is included in appendix \ref{appenB} A BON to SensorML XML encoding mapping is also developed and is included in appendix \ref{appenC}.
 
 \chapter{ Conclusions and Future Work}
-This thesis analyses the role of the SensorML and the SOS specifications in Sensor Systems.
-The UCD CASL SenseTile System is used as a case study for the use of these specifications.
-A SensorML model of the SenseTile system is developed to evaluate SensorML.
-A SenseTile Web Service prototype is used to evaluate the SOS specification which uses SensorML to provide sensor descriptions to clients.
-
-SensorML provides a good model of sensor systems in terms of a an intuitive process concept.
-to describe of a sensor systems.
-It is hard to think of any system that could no be described by SensorML due to number of data types that can.
-Weather systems to Geloloaction services.
-SensorML address the issues of interoperabilty of Sensor Systems
-
-For descriptions there good Metadata.
+\subsection{Conclusions}
+This project analyses the role of the SensorML and SOS specifications in sensor systems.
+The UCD CASL SenseTile System is used as a case study for the use of these specifications with a SensorML model of the SenseTile system developed to evaluate SensorML.
+A SenseTile Web Service prototype is developed to evaluate the SOS specification which uses SensorML to provide sensor descriptions to clients.
+
+SensorML provides a simple functional model of sensors in terms of an intuitive process concept.
+A process transforms sensor inputs to outputs as modeled by the SensorML AbstractProcess type.
+SensorML provides a process chaining capability to break up complex processing into smaller more manageable steps.
+There is also a rich set of data types to describe the sensor data.
+Due to the simple yet flexible model and the rich set of data types is is hard to see any sensor or sensor system that could not be described by SensorML.
+This is part of its goal to be a standard of interoperability of sensor system.
+The Sensor XML encoding provided furthers this goal.
+SensorML also provides a good level of metadata types to assist with a human understanding of the capabilities of the sensor system.
 
 SensorML provides good support for the reliable processing of sensor data.
-The data types have several properties that can reference definitions to precisely describe their meaning.
-The accuracy of the sensor data can be modelled and provided used by systems to process the data with these variables include to get a better result.
-
-Processing data. 
-SensorML Frameworks. DataProducer.
-
-It also is bases for development for Sensor Ontology work.\footnote{http://www.calebgoodwin.com/SEMANTIC.html}
+The data types have properties that can reference definitions to precisely describe their meaning or give an indication of the quality of the data.
+The accuracy of the sensor can be modeled and can be provided to processes using parameters to get a more accurate output from the process.
+
+It could model the SenseTile system well at this functional level. 
 SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.
 
-The SenseTile model uses a hierarachy to model the SenseTile.
-It seems repetitive to have keep defining the same inputs for each contained component.
-Maybe this is a flaw in the SenseTile model developed.
-
-SensorML is defined by a heavy XML Schema.
-This XML Schema along with the other XML Schema for the SOS and dependent specifications are heavy indeed.
-Achieving a Java binding was difficult.
-JIBX would not generate a default XML binding.
+SensorML is not intended to provide the framework for encoding the actual
+
+SensorML is defined by a complex set of XML Schema.
+Achieving a Java to XML binding was difficult based on these schema.
+JIBX\footnote{http://jibx.sourceforge.net/} would not generate a default XML binding.
 JAXB could with some help.
@@
 
-The idea of automatically obtaining a SensorML model and executing it depends on the SensorML Engine being used.
-Only one Java implementation can be pointed.
-In the OGC examples the Java implementations of the
-
-Vast Lib use as a references only.
-Had to update library code to get it to work with the SenseTile schema even though the schema
-passed the validation test tool from the site. It uses DOM but XML binding is probably a better approach.
+The Vast Lib seems to be the reference implementation of a SensorML processing framework.
+There was little or no documentation that I could find on this and examples of it's use needed some research.
+It would be better if a working example could be provided with the framework and also some basic documentation.
+The Vast code needed to be slightly modified to parse the SenseTile model files though they passed a validation check\footnote{http://vast.uah.edu/SensorMLforms/validate.jsp} provided by the Vast group.
+After these updates allowed the parsing to finish the code was indeed very stable and never crashed except when there were mistakes in the SenseTile SensorML XML files.
 
@@
 
+The SOS specification is complex.
+If there no need for interoperability or a rich filtering ability of sensor data provided by the SOS operations then it is better to go with some other simpler approach.
+This approach could be use standard Web Services and use SensorML for sensor descriptions.
+
 SOS Scalable hierarcy of instances.
@@
 
-further work.
----------------
+\subsection{Future Work}
 SensorML sensor accuracy.
@@
 \bibitem{OGCcatref}Open Geospatial Consortium Inc., OpenGIS® Catalogue Services Specification, 2007
-\bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
 \bibitem{Meyerref}Bertrand Meyer., "Object-Oriented Software Construction", 1997
@@
 \appendix
-\chapter{Appendix: SenseTile SensorML Model}
+\chapter{Appendix: SenseTile SensorML Model}\label{appenA}
 \section{SenseTile SensorML System}

### 2010-04-14 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 \begin{abstract}
-This thesis evaluates the Sensor Model Language(SensorML) and Sensor Observation Service(SOS) specifications.
+This project evaluates the Sensor Model Language(SensorML) and Sensor Observation Service(SOS) specifications.
 These specifications are intended for use in the development of open sensor systems.
@@
 This Web Service provides access to SenseTile sensor metadata and sensor observations.
-The thesis finally examines formalizing SensorML using a formal specification language called Business Object Notation.
+The project finally examines formalizing SensorML using a formal specification language called Business Object Notation.
 
@@
 The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as the SensorML and SOS specifications.
-This thesis investigates the use of these SWE SensorML and SOS standards in developing sensor systems.
+This project investigates the use of these SWE SensorML and SOS standards in developing sensor systems.
 
@@
 
-This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board.
+This project describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board.
 Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor descriptions and sensor observations.
@@
 BON is a notation and a method for analysis and design of Object Oriented systems.
-This thesis looks at how BON can be used to formally specify SensorML.
-The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to JML\footnote{http://www.eecs.ucf.edu/~leavens/JML/} and then to Java\footnote{http://www.java.com/}, as described in \cite{Kiniryref}.
+This project looks at how BON can be used to formally specify SensorML.
+The SensorML BON specification developed in this project allows SensorML sensor descriptions to then be refined to JML\footnote{http://www.eecs.ucf.edu/~leavens/JML/} and then to Java\footnote{http://www.java.com/}, as described in \cite{Kiniryref}.
 
@@
 
-The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.
+The SenseTile Web Service developed in this project is based on the concepts described in the Sensor Observation Service (SOS) specification.
 The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
 
-The OGC Catalog Service shown in the SOS Operation Context diagram \ref{fig:SOSoperationContext} is not explored in this thesis.
+The OGC Catalog Service shown in the SOS Operation Context diagram \ref{fig:SOSoperationContext} is not explored in this project.
 The Data Consumer uses a CS-W catalog to find SOS instances.
@@
 
-There several optional SOS operations not covered in this thesis.
+There several optional SOS operations not covered in this project.
 The SOS refers to these as enhanced operations.
@@
 Sensors are mounted on the Sensor Board and more sensors can be added using the USB connection.
-The SensorBoard and its sensors are the main focus of the SensorML model in this thesis and as the processor unit is a relatively powerful computer, providing a lightweight Web Service on the SenseTile Processor Unit itself is feasible.
+The SensorBoard and its sensors are the main focus of the SensorML model in this project and as the processor unit is a relatively powerful computer, providing a lightweight Web Service on the SenseTile Processor Unit itself is feasible.
 
@@
 
-A SensorML model of SenseTile is developed as part of this thesis.
+A SensorML model of SenseTile is developed as part of this project.
 The SenseTile Sensor Board and its sensors is modeled in the most detailed.
@@
 The DataProvider parses a SensorML description of the SenseTile System, reads sensor data form the Sensor Board and provides the sensor data as O\&M Observations to the SOS.
-The SenseTile SOS will implement the SOS core operations to allow a client to get these observations.
-The following sections describe the design of the DataProvider and SOS components as well as the scenarios they support.
+The SenseTile SOS implements the SOS core operations to allow a client to get sensor observations.
+The following sections describe the SenseTile network architecture, the supported scenarios of the system and the design of the DataProvider and SOS components.
 
@@
 \label{subsec:SenseTile WebService Network Architecture}
-In the SenseTile Web Service the SOS acts as a gateway to the SenseTile Sensor Network.
-This prototype is designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
+The SOS Specification envisioned that the SOS is run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}. 
+\begin{figure}[h]
+\centering
+\scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
+\caption{SOS Specification Architecture}\label{fig:Deployment_ext_sos}
+\end{figure}
+\newline
+The DataProducers update this external server with observations and SOS Data Consumer access this large SOS. Several SOS servers can be used in a hierarchy to allow the system to scale.
+
+The SenseTile Web Service is designed with the network architecture as shown in fig \ref{fig:Deployment_network}.
+\begin{figure}[h]
+\scalebox{0.65}{\includegraphics{Deployment_sensetile_network.png}}
+\caption{SenseTile Web Service Network Architecture}\label{fig:Deployment_network}
+\end{figure}
+\newline
+This is a different approach than outlined in the SOS specification.
 A client accesses multiple SOS instances to get observations from the sensor network.
-An SOS can support several DataProducers and hence would not to be run on every SenseTile Unit.
-
-\begin{figure}[h]
-\scalebox{0.65}{\includegraphics{Deployment_sensetile_network.png}}
-\caption{SenseTile Sensor Network Architecture}\label{fig:Deployment_network}
-\end{figure}
-The SOS Specification envisioned that the SOS is run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}.
-The DataProducers update this external server with observations and SOS Data Consumer access this large SOS.
-\begin{figure}[h]
-\centering
-\scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
-\caption{External SOS}\label{fig:Deployment_ext_sos}
-\end{figure}
-
-The SOS Specification describes the use of a Proxy SOS eeicpr fix. This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.
-
+An SOS can support several DataProducers and hence is not needed on every SenseTile Unit.
+This approach is used to take advantage of the relatively powerful processing capability of the SenseTile Processor Unit eliminating the need for the large SOS server.
+
+\subsection{SenseTile Web Service Scenarios}
+This section contains UML sequence diagrams to show how the DataProducer and SOS interact to provide the sensor obsevatons and sensor descriptions to the DataConsumer client.
+There are three main scenarios supported by the SenseTile Web Service and they are described in the following sections.
+
+\subsubsection{Generate Observations from Sensor Data}
+\begin{figure}[h]
+\centering
+\framebox{\scalebox{0.61}{\includegraphics{InsertObs.png}}}
+\caption{Generate Observation}\label{fig:InsertObs}
+\end{figure}
 \newpage
-\subsection{SenseTile Web Service Scenarios}
-This section contains UML sequence diagrams to show the how the DataProducer and SOS interact to provide the sensor to the DataConsumer client.
-There are three main scenarios supported by the SenseTile Web Service
-
-\subsubsection{Generate Observations from Sensor Data}
-The DataProducer accesses data from the sensor board and updates the SOS with sensor observations.
-\begin{figure}[h]
-\centering
-\framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
-\caption{Generate Observation}\label{fig:InsertObs}
-\end{figure}
  \begin{enumerate}
 \item At start up the DataProducer registers the SensorBoard System and the Converter System with the SOS in separate RegistorSensor operations.
-The SOS creates sensor offerings for each of the Systems and stores the SensorML descriptions of the Systems provided in the RegistorSensor operation.
-It then returns a unique id (URN) in the RegistorSensor response for use by the DataProducer for subsequent communication with the SOS.
-\item The DataProducer reads Sensor Board data packets from the SensorBoardDriver.
-\item The DataProducer updates the SOS with sensor observations with the InsertObservation operation.
+The SOS creates sensor offerings for each of the Systems and stores the SensorML descriptions of the Systems.
+It then returns a unique identity URN in the RegistorSensor response for use by the DataProducer for subsequent communication with the SOS.
+\item The DataProducer reads Sensor Board data packets from the SensorBoardDriver. It waits one second before reading packets.
+\item The DataProducer updates the SOS with sensor observations in an O\&M XML format with the InsertObservation operation.
 \end{enumerate}
@@
 \subsubsection{Get Sensor Description from SOS}
-The SOS allows DataConsumers to obtain SensorML descriptions of the SesneTile Systems.
-\begin{figure}[h]
-\centering
-\framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
+\begin{figure}[h]
+\centering
+\framebox{\scalebox{0.61}{\includegraphics{GetMeta.png}}}
 \caption{Get Sensor Descriptions}\label{fig:GetMeta}
@@
  \begin{enumerate}
-\item The DataConsumer requests information on sensor offerings provided by the SOS using the GetCapabilities operation.
-\item Using the information in the GetCapabilites response the DataConsumer requests a description of a sensor system that is providing an offering using the DescribeSensor operation.
+\item The DataConsumer requests information about SenseTile sensor offerings provided by the SOS using the GetCapabilities operation.
+\item Using the information in the GetCapabilites response, the DataConsumer requests a description of a sensor system that is providing an offering using the DescribeSensor operation.
 The SOS returns the SensorML description of the sensor system to the DataConsumer.
@@
 \centering
-\framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
+\framebox{\scalebox{0.61}{\includegraphics{GetObs.png}}}
 \caption{Get Observation from Sensor}\label{fig:GetObs}
@@
  \begin{enumerate}
-\item Using a sensor identity received in a GetCapability response the DataConsumer requests a sensor observation. The SOS returns the latest O\&M observation to the DataConsumer.
+\item Using the sensor identity received in a GetCapability response the DataConsumer requests a sensor observation. The SOS returns the latest O\&M format sensor observation to the DataConsumer.
 \end{enumerate}
@@
 These observations are then sent to the SOS instance.
+The BON static diagram \ref{fig:bon_static_diagam_producer.png}
+ shows the main clusters and classes of the DataProvider. BON notatation. Not all the classes are described.
+\begin{figure}[h]
+\centering
+\scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
+\caption{DataProducer BON Static Diagram}\label{fig:bon_static_diagam_producer.png}
+\end{figure}
+
 The processing of the sensor data to generate the observations is described by the SensorML models.
-The DataProducer uses the OGC Vast Library to instantiate objects that implement the required data processing chains based on the SensorML models.
-The BON static diagram \ref{fig:bon_static_diagam_producer.png}
- shows the main clusters and classes of the DataProvider.
-
-The SMLEngine cluster contains classes that provide access to
-the Vast Lib that is providing the SensorML Engine.  The Vast Lib implementation
-of the SensorML Engine is hidden from the rest of the code thus.
+The DataProducer uses the VastSMLEngine to instantiate objects that implement the required data processing chains based on the SensorML models.
+The SMLEngine cluster contains classes that provide access to the VastSMLEngine objects to access the data in the SenseTile SensorML model and to execute the processing as described in the SenseTile SensorML model.
 
@@
 \end{enumerate}
-The steps above are detailed the in following sections with reference
-to the BON static diagram \ref{fig:bon_static_diagam_producer.png}.
 
@@
 
-The UCD SensorBoard Library is used by DataProducer to access
-the data generated by the Sensor Board.
-The library provides high level interface to the
+The UCD CASL SensorBoardDriver is used by DataProducer to access
+the data generated by the Sensor Board. The driver provides a high level interface to the
 the data from the sensor board. It reads data from the Sensor board and generates packets
@@
 measurement generated by the thermistor. 
-This library is shown as the SensorBoard cluster in \ref{fig:bon_static_diagam_producer.png}
+This driver is shown as the SensorBoard cluster in \ref{fig:bon_static_diagam_producer.png}
 An instance of the class PacketInputStream to is used access the packets from the Sensor Board.
@@
 The DataProducer uses the Observer pattern to provide
-packets to the SensorMLEngine for processing.
-At start up the SenseTileSystem reads the list of sensor objects from the SMLEngine that
+packets to SensorML sensor processes that are executed by the VastSMLEngine.
+At start up the SenseTileSystem reads the list of sensor objects from the VastSMLEngine that
 implement the Sensors interface. These objects provide an entry point
-to the SensorML processing chain and can read data from the UCD LIB packets.
-These sensor objects implement reading strategies for the type of data they read. 
-For example the ThermistorSensor reads out the temperature value
-but will average it before it sent to the SensorML Engine for processing.
+to the SensorML processing chain and can read data from the SensorBoardDriver packets.
+
+These sensor objects implement data reading strategies for the type of data they read. 
+For example the ThermistorSensor reads out temperature value, stores it
+and then averages the values before sending a single value to the SensorML Engine for processing.
 
 \subsubsection {2. Process Sensor Data}
-The PacketInputStream will request the SenseTileSystem to exexcute
-the SensorML engine to process the sensor data. The SensorML Enigine
-will execute the process chains described in the SenseTile SensorML model.
-The sensor objects output will be passed to the converter objects. The
-outputs of both object types are then available to generate observations.
+The PacketInputStream will request the SenseTileSystem to execute
+the VastSMLEngine engine to process the sensor data. The VastSMLEngine
+executes the process chains described in the SenseTile SensorML model.
+
+The sensor processes output value is passed to the converter processes for further processing. The
+outputs of both processes are then available to generate observations.
 
@@
 
-When the SensorML Engine is finished
-and there are obsevations generated these
+When the VastSMLEngine Engine is finished processing
+the generated obsevations generated these
 need to be sent to the SOS. An O\&M Onservation
-XML string containing the data is sent to
-the SOS instance. This is timestamped at
-this stage.
- containing the output from
-the SensorML processes and send the observations
-to the SOS.
--register offerings with the SOS. In this design
-each system is an offering.
-Need to generate an observation model?
-
-The DataProducer was developed as a SensorML based software.
-It parses the SenseTile SensorML model to instantiate the code to execute the processes described.
-This required two aspects to be developed, a framework to run the processes and classes that implemented the components.
-
-\begin{figure}[h]
-\centering
-\scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
-\caption{DataProducer Static Diagram}\label{fig:bon_static_diagam_producer.png}
-\end{figure}
+XML string containing the sensor data is sent to
+the SOS instance.
 
@@
 
-The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the observations from the DataProducer.
-The below is a static diagram showing the important clusters and classes in the SOS design.
+The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the senosr observations from the DataProducer.
+The below is a BON static diagram showing the important clusters and classes in the SOS design.
 
@@
 \scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
-\caption{SOS Static Diagram}\label{fig:bon_static_diagam_provider.png}
+\caption{SOS BON Static Diagram}\label{fig:bon_static_diagam_provider.png}
 \end{figure}
@@
 The SOS is the provider of the Web Service interface and is built on Apache Axis2 \footnote{http://ws.apache.org/axis2}. 
-Axis2 is described as a "Web Services / SOAP / WSDL engine"  with the Apache Axis2/Java implementation used in this thesis.
+Axis2 is described as a "Web Services / SOAP / WSDL engine"  with the Apache Axis2/Java implementation used in this project.
 The Spring framework\footnote{http://www.springsource.org/} was also used.
@@
 Port contains all the properties of SensorML input, output and parameter as they all have the same properties.
-A PortType class that ensures legal port types of Input, Output and Parameter is also added to the SenorML BON model.
+A PortType class that ensures legal port types of Input, Output and Parameter is also added to the SensorML BON model.
 
@@
 
-The BON specification developed in this thesis does not address the issue of the SensorML metadata and the MetaDataGroup is not specified.
+The BON specification developed in this project does not address the issue of the SensorML metadata and the MetaDataGroup is not specified.
 This information is for descriptive purposes only and is not used for processing sensor data.
@@
 Member
-The complete BON SensorML specification listing can be found in appendix \ref{appenBSMLBONSpec}.
+The complete BON SensorML specification listing can be found in appendix \ref{appenB}.
 
@@
 A SenseTile Web Service prototype is developed to evaluate the SOS specification which uses SensorML to provide sensor descriptions to clients.
-
+The SensorML and SOS specifications are proposed open standards for sensor description and sensor data access respectively.
+This is part of the OGC SWE initiative goal to improve interoperability of sensor system.
+SensorML and SOS XML encodings are provided by the specifications to support a Web Services approach to sensor system interoperability, the overall aim of the these and other OGC SWE specifications.
+
+\subsubsection{SensorML}
 SensorML provides a simple functional model of sensors in terms of an intuitive process concept.
@@
 SensorML provides a process chaining capability to break up complex processing into smaller more manageable steps.
-There is also a rich set of data types to describe the sensor data.
-Due to the simple yet flexible model and the rich set of data types is is hard to see any sensor or sensor system that could not be described by SensorML.
-This is part of its goal to be a standard of interoperability of sensor system.
-The Sensor XML encoding provided furthers this goal.
+There is also a rich set of data types to describe sensor data.
+Due to the simple model and the rich set of data types is is hard to imagine any sensor or sensor system that could not be described by SensorML.
+It is not intended to describle details of the sensor hardware.
+It is also not intended that SensorML provides the framework for encoding the actual sensor observations.
+This is provided by the SWE O\&M specification.
+
 SensorML also provides a good level of metadata types to assist with a human understanding of the capabilities of the sensor system.
-
-SensorML provides good support for the reliable processing of sensor data.
-The data types have properties that can reference definitions to precisely describe their meaning or give an indication of the quality of the data.
-The accuracy of the sensor can be modeled and can be provided to processes using parameters to get a more accurate output from the process.
-
-It could model the SenseTile system well at this functional level. 
-SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.
-
-SensorML is not intended to provide the framework for encoding the actual
+All SensorML process can provide metadata such as capabilities, properties, contacts information and documentation sources.
+
+Good support for the reliable processing of sensor data is provided by SensorML.
+The data types have properties that can qualify the data.
+For example the data type can reference external definition libraries to precisely describe their meaning and give an indication of the quality of the data.
+The accuracy of the sensor can be modeled using accuracy curves which can be provided to processes using parameters to get a more accurate output from the process.
+
+SensorML can model a simple sensor or complex sensor systems.
+The SensorML specification defines a simple Detector model that limits the process to one input and one output.
+The Detector process outputs a real value such as a Celsius temperature value.
+eeicpr - tutorial ref.
+The SenseTile model uses a different approach as the unconverted data can also be retrieved from the SenseTile Web Service.
+There are good examples eecipr\footnote{} and tutorials available to help develop SensorML models.
+
+The SenseTile model developed in this project is relatively simple and did not use the full richness of SensorML.
+It does not, for example, use the SensorML support for accurate and reliable processing data.
+It did examine a simple process chain which highlighted some of the difficulties of process chaining described in chapter x.
 
 SensorML is defined by a complex set of XML Schema.
-Achieving a Java to XML binding was difficult based on these schema.
-JIBX\footnote{http://jibx.sourceforge.net/} would not generate a default XML binding.
-JAXB could with some help.
-There are many layers of abstraction in these XML Schema.
-This reflects the that the specifications cover a very wide range of needs of a senor system.
-
-The Vast Lib seems to be the reference implementation of a SensorML processing framework.
+Achieving a Java to XML binding is difficult based on these schema.
+JIBX\footnote{http://jibx.sourceforge.net/} did not generate a default XML binding.
+JAXB did generate a default binding with some help but updates are still needed on the generated code to access some elements.
+There are many layers of abstraction in these XML Schema obviously making the binding generation difficult.
+This reflects the fact that the specification tries to cover all possible needs of a sensor system and is thus complex.
+
+The VastSensorMLEngine seems to be the reference implementation of a SensorML processing framework.
 There was little or no documentation that I could find on this and examples of it's use needed some research.
 It would be better if a working example could be provided with the framework and also some basic documentation.
-The Vast code needed to be slightly modified to parse the SenseTile model files though they passed a validation check\footnote{http://vast.uah.edu/SensorMLforms/validate.jsp} provided by the Vast group.
-After these updates allowed the parsing to finish the code was indeed very stable and never crashed except when there were mistakes in the SenseTile SensorML XML files.
-
-The SWE SOS specification describes the operations for accessing sensor observations and it's concepts were used to develop the SenseTile Web Service.
-It uses SensorML provide sensor metadata to clients and the O\&M specification to encode observations.
-Implement all functionality would be difficult.
-Maybe too heavy for SenseTile System.
-Simple querires.
-Or better a push system to send the data to the client as it is generated could be a better approach here.
-
+The VastSensorMLEngine code needed to be slightly modified to parse the SenseTile model files though they passed a validation check\footnote{http://vast.uah.edu/SensorMLforms/validate.jsp} provided by the Vast group.
+After these updates allowed the parsing to complete the code was indeed very stable and never crashed except when there were mistakes in the SenseTile SensorML XML files.
+
+To use SensorML effectively a graphical tool or some other method rather using XML directly is needed.
+Working in SensorML XML is very error prone.
+Running the SenseTile model through the VastSensorMLEngine without a crash was the general procedure to check the correctness of the SensorML descriptions.
+
+\subsubsection{SensorML BON Specification}
+The BON SensorML developed in this project just covers the core of the SenorML types.
+However it has shown how some of the problems of SensorML can be addressed with a formal language such as BON.
+A way to check that connections in SensorML processchains are correct is described using BON assertions.
+The type safety and assertions provided by a BON SensorML specification allows a better way to develop SensoML based sensor sysstems than just using the SensorML XML encoding.
+
+A more complete BON SensorML description is needed than provided in this project and there are still many issues that need to be addressed.
+For example how to represent XLINKs in the SensorML BON specification or is there a need to.
+Based on the SensorML BON specification developed in this project an inline approach would be necessary rather than reference external XML files general used in the examples.
+For a complete SensorML description the SensorML metadata needs also to be included in an SensorML description.
+How to do this in not covered in this project.
+
+\subsubsection{SOS}
+The SenseTile Web Service developed in this project is based on the SOS specification.
 The SOS specification is complex.
-If there no need for interoperability or a rich filtering ability of sensor data provided by the SOS operations then it is better to go with some other simpler approach.
-This approach could be use standard Web Services and use SensorML for sensor descriptions.
-
-SOS Scalable hierarcy of instances.
+Only a small amount of the SOS functionality is implemented in the SenseTile Web Service but this allows sensor descriptions and sensor observations to be retrieved.
+The SenseTile Web Service is very much a prototype and would need more work to make it a real Web Service.
+
+Implementing all SOS functionality is a large task.
+If there no need for interoperability or complex querying of sensor data provided by the SOS operations then it is better to go with some other simpler approach.
+One approach could be to use industry standard Web Services and use SensorML for sensor descriptions.
+
+The approach used in the SenseTile Web Service of having both the SOS and the DataProducer on the SenseTile Processor unit may be too complex.
+It could be better to move the SOS to a separate server as described in the SOS specification.
+This would simplify the system design and management.
+The use of RMI may not be a good choice for the communication between the DataProducer and SOS as it limit this communication to
 
 SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
-flexible. There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
-what has been achieve
-
-SOS on the Processor Board. Is this right? RMI only useful if SenseTile Network is a lan.
-
-BON
-
-Working in XML is error prone.
-SensorML connections type was a good example of where a graphical tool or another language, like BON,  that can be parsed should be used to check the connections.
-Running the model through the VastSensorMLEngine without a crash was only way other than by eye to see that the connections were set up correctly.
+flexible.There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
 
 \subsection{Future Work}
-SensorML sensor accuracy.
-Complete BON mapping and parser.
-Streaming the data. See SWE Architecture dpcument for a description of one approach.
-Other technology approaches such as . Other ways such a XQUERY? XML techs?
+It would be interesting to investigate how accurate the sensor data can be made using the SensorML support for reliable data processing.
+Calibration and accuracy curves could be developed based on the SenseTile sensor hardware specifications.
+A simple Detector model could be developed to see how this works in practice.
+
+A complete SensorML BON specification is still to be done and developing a BON to SensorML tool would be interesting.
+This tool could also incorporate a way to add the sensor metadata.
+
+Streaming the data. See SWE Architecture document for a description of one approach.
+
+Comparing a Restful approach to retrieving sensor observations rather than the SOS approach could be investigated. It may prove to be a simpler way to retrieve the sensor observations.
 
@@
 \bibitem{Meyerref}Bertrand Meyer., "Object-Oriented Software Construction", 1997
+\bibitem{Sensorsref}Bertrand Meyer., ", 1997
 
@@
 
-\chapter{Appendix: BON to SensorML Mapping}\label{appenB}
-\section {SensorML BON Specifcation}\label{appenBSMLBONSpec}
+\chapter{Appendix: SensorML BON Specification}\label{appenB}
 
@@
 \lstinputlisting{SensorML_formal.bon}
-\section {BON to SensorML mapping tables} \label{appenBSMLBONTables}
-\lstset{frame=none}
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class AbstractProcess  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class System  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class C  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class Link  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class ProcessChain  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class ProcessModel  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class Port  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class Quantity  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
-
-\begin{table}[!th]
-\centering
-\begin{tabular}{|l|l|}
-\hline
-BON features & SensorML XML tags\\
-\hline
-   deferred class Count  & not mapped\\
-\hline     
-     inputs:SEQUENCE[Input] & \begin{lstlisting}
-<sml:inputs><sml:inputlist>
-SEQUENCE[Input]
-<\sml:InputList><\sml:inputs>\end{lstlisting}\\
-
-\hline 
-     outputs:SEQUENCE[Output] & \begin{lstlisting}
-<sml:outputs><sml:outputlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-\hline
-     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
-<sml:parameters><sml:parametertlist>
-SEQUENCE[Input]
-<\sml:outputList><\sml:outputs>\end{lstlisting}\\
-
-\hline                 
-     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
-<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
- \hline
-
-\end{tabular}
-\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
-\label{ex:table}
-\end{table}
 

### 2010-04-15 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

@@
 These specifications are intended for use in the development of open sensor systems.
-The UCD CASL SenseTile sensor system is used as a case study for this evaluation.
+The UCD CASL SenseTile Sensor System is used as a case study for this evaluation.
 A SensorML model of SenseTile is developed.
@@
 This Web Service provides access to SenseTile sensor metadata and sensor observations.
-The project finally examines formalizing SensorML using a formal specification language called Business Object Notation.
+The project finally examines formalizing SensorML types using Business Object Notation.
 
@@
 The SenseTile Web Service developed in this project is based on the concepts described in the Sensor Observation Service (SOS) specification.
-The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
+The SOS Data Consumer and SOS Data Provider entities are referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.
 
@@
 It is reversible as changes in the source code can be easily mapped back into the design and analysis models.
-It supports design by contract\cite{Meyerref} where a function will guarantee a postcondition if a specified precondition is fulfilled by the caller of the function.
+It also supports design by contract\cite{Meyerref} where a function guarantees a postcondition if a specified precondition is fulfilled by the caller of the function.
 BON support for design by contract allows for a formal description of a system to be developed.
@@
 
-Of the SenseTile sensors only the Thermistor is modeled along with the  CelsiusConv Component that converts the digital number received from the Thermistor Component to a Celcius value.
+Of the SenseTile sensors only the Thermistor is modeled along with the  CelsiusConv Component that converts the digital number received from the Thermistor Component to a Celsius value.
 However the same approach can applied to the other senors though for some components no conversion process may be needed.
@@
 For example in this case the Sensor Board System is called sensorBoardSystem.
-The component can be a SensorML process or a processes chain.
+The component can be a SensorML process or a process chain.
 
@@
 These entities are implemented as separate software components that together provide the service.
-The Sensor Data Provider component will be referred to as the DataProvider in this design.
+The Sensor Data Provider component is referred to as the DataProvider in this design.
 
@@
 \subsection{SenseTile Web Service Scenarios}
-This section contains UML sequence diagrams to show how the DataProducer and SOS interact to provide the sensor obsevatons and sensor descriptions to the DataConsumer client.
+This section contains UML sequence diagrams to show how the DataProducer and SOS interact to provide the sensor observations and sensor descriptions to the DataConsumer client.
 There are three main scenarios supported by the SenseTile Web Service and they are described in the following sections.
@@
 \scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
-\caption{DataProducer BON Static Diagram}\label{fig:bon_static_diagam_producer.png}
+\caption{DataProducer BON Class Diagram}\label{fig:bon_static_diagam_producer.png}
 \end{figure}
@@
 \subsubsection {2. Process Sensor Data}
-The PacketInputStream will request the SenseTileSystem to execute
+The PacketInputStream requests the SenseTileSystem to execute
 the VastSMLEngine engine to process the sensor data. The VastSMLEngine
@@
 When the VastSMLEngine Engine is finished processing
-the generated obsevations generated these
-need to be sent to the SOS. An O\&M Onservation
+the generated observations generated these
+need to be sent to the SOS. An O\&M Observation
 XML string containing the sensor data is sent to
@@
 
-The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the senosr observations from the DataProducer.
+The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the sensor observations from the DataProducer.
 The below is a BON static diagram showing the important clusters and classes in the SOS design.
@@
 \scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
-\caption{SOS BON Static Diagram}\label{fig:bon_static_diagam_provider.png}
+\caption{SOS BON Class Diagram}\label{fig:bon_static_diagam_provider.png}
 \end{figure}
-The SOS is the main cluster and the Service cluster contains the classes that provide the Web Service interface.
 
-The DataProvider maintains a list of observation offerings provided by the SOS using the ObservationOffering class.
+The DataProvider class maintains a list of observation offerings provided by the SOS using the ObservationOffering class.
 When a new sensor type is registered a new ObservationOffering is created.
-If a DataProvider registers and provides an offering that already exists then just a new Sensor instance is added to the existing ObservationOffering.
+If a DataProvider registers and provides an offering that already exists then just a new Sensor instance is created and its identity is added to the appropriate ObservationOffering.
 
-The Sensor class contains a list of Observations that hold the observations from the DataProvider. The last 50 observations are maintained.
+The DataProducerService class provides the interface for the DataProducer to register senors and update the SOS with observations.
+These sensor observations are stored in  the Sensor class in a list of Observations classes.
+The SenseTileService provides the Web Service interface used by clients to access the sensor observations.
 
-The ObservationService class provides the interface for the DataProducer to update with observations.
-
-The SenseTile Service provides Web Service interface to be used by clients to access the sensor observations.
-
-The SOS specification provides a very rich filtering feature that is too complex for this prototype.
-The operations are developed in a way to allow the sensor metadata and observations to be retrieved for a specific sensor board.
-The latest observation or all the stored observations can be retrieved.
+The SOS specification provides a very rich filtering feature in the GetObservation operation.
+However SenseTile Web Service only implements the functionality to get a sensor observation using a sensors identity.
 
@@
 The SenseTile Web Service is written in Java.
-As described in the high level design there are two components, the DataProducer and the SOS, that are developed to provide this service.
+As described in the design chapter there are two components, the DataProducer and the SOS, that are developed to provide this service.
 The two components are run in separate processes.
-The communication between the DataProducer and the SOS uses Java RMI.
+The communication between the DataProducer and the SOS uses Java RMI\footnote{http://java.sun.com/javase/technologies/core/basic/rmi/index.jsp}.
 RMI is used due to the SOS and DataProducer
 processes are running in the SenseTile Network and communicate over a LAN.
-This communication protocol has lower overhead than a Web Service so efficient communication between the DataProducer and SOS is achieved.
-Though it would not be hard to change to use a Web Service as a Java Interface is used to hide the communication mechanism from the main application code.
-Clients, SOS DataConsumers, access the SOS using a SOAP based Web Service interface.
+This communication has lower overhead than using HTTP so efficient communication between the DataProducer and SOS is achieved.
+It is not be hard to change to use a a different transport layer as a Java Interface is used to hide the communication mechanism from the main application code.
+Clients, SOS DataConsumers, access the SOS using a SOAP\footnote{http://www.w3.org/TR/soap/} based Web Service interface.
 
@@
 The Spring framework\footnote{http://www.springsource.org/} was also used.
-Using these libraries a light Web Service is built to run on the reasonably powerful SenseTile Processor Unit.
-Axis2 is run standalone here though it can be on with any Servlet engine if a more powerful Web Service is needed.
+Using these technologies a light Web Service is built to run on the reasonably powerful SenseTile Processor Unit.
+Axis2 is run standalone here though it can run on any Servlet engine if a more powerful Web Service is needed.
 
-Developing a Web Service is straight forward with Axis2 and it supports several approaches.
+Developing a Web Service is straight forward with Axis2 and it supports several approaches to developing Web applications.
 The supported POJOs (Plain Old Java Objects) approach was used for the SenseTile Web Service allowing the code to be reused on a different framework if AXIS2 did not work well.
-Axis2 supports the The Spring framework which is used in the SenseTile WebService for it's dependency injection functionality.
+Axis2 supports the Spring framework which is used in the SenseTile WebService for it's dependency injection functionality.
 A SprinitInit class is needed for Axis2 to load up the Spring framework.
 
-Axis2 provides a tool that automatically generates WSDL from the running Web Service.
-Another tool allows client side code to be generated from the WSDL easing the development of the Web Service clients.
+Axis2 provides a tool that automatically generates WSDL\footnote{http://www.w3.org/TR/wsdl} from the running Web Service.
+Another tool allows client side code to be generated from the WSDL allowing easier development of the Web Service client.
 
@@
 
-\chapter{BON to SensorML mapping}
+\chapter{SensorML BON Specifications}
 \section{Overview}
-In this chapter a BON(see section\ref{BONsec}) to SensorML( see section \ref{SMLsection}) mapping is described.
-One motivation for this mapping is take advantage of formal software development processes described in \cite{Kiniryref}.
+In this chapter a BON(see section\ref{BONsec}) SensorML( see section \ref{SMLsection}) specification is described.
+One motivation for this is take advantage of formal software development processes described in \cite{Kiniryref}.
 This is a verification based process that refines BON to JML and then to Java.
 Robust software can thus be developed.
-Another motivation is that it XML is verbose and a more compact description of a SensorML model as facilitated by BON is naturally preferred.
-Lastly SensorML is complex and it is difficult to develop correct SenorML models based only on using the SensorML XML Schema definition.
-How a BON specification of SensorML can help here is described later in the chapter.
+Another motivation is that it XML is verbose and a more compact description of a SensorML model as facilitated by BON is preferred.
+Lastly SensorML is complex and it is difficult to develop correct SensorML models based only on using the SensorML XML Schema definition.
+A BON specification of SensorML can help here and an example of this is described later in the chapter.
 
 \section{SensorML BON Specification}
-The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.  The BON SensorML model is shown in figure \ref{fig:bonSensoMLModel}. 
+The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.
+The SensorML BON Class Diagram is shown in figure \ref{fig:bonSensoMLModel}.
 \begin{figure}[h]
@@
 \scalebox{0.65}{\includegraphics{bonSensoMLModel.png}}
-\caption{BON SensorML Model}\label{fig:bonSensoMLModel}
+\caption{SensorML BON Class Diagram}\label{fig:bonSensoMLModel}
 \end{figure}
-The inheritance heirarhcy is changed slightly with System inheriting from ProcessChain and Component inheriting from ProcessModel.
+\newline
+The inheritance hierarchy is changed with System inheriting from ProcessChain and Component inheriting from ProcessModel.
 A new abstraction called Port is added.
-Port contains all the properties of SensorML input, output and parameter as they all have the same properties.
+Port contains all the properties to describe SensorML process's input, output and parameter.
 A PortType class that ensures legal port types of Input, Output and Parameter is also added to the SensorML BON model.
@@
 The abstract class DataComponent is the base type for these.
-There are a many other types but only Count and Quantity are developed as they were the ones used in the SenseTile SensorML model.
+There are a many other types but only Count and Quantity are developed as they are the ones used in the SenseTile SensorML model.
 
@@
 A different way to add this information to SensorML sensor descriptions is suggested to be used.
+
+eeicpr
+
 AssociationAttributeGroup
@@
 Member
-The complete BON SensorML specification listing can be found in appendix \ref{appenB}.
 
-\section{SensorML BON to XML Mapping}
-SensorML is essential XML defined by XML Schema with its main purpose being interopeability.
-Therefore a mapping SensorML BON specification to the SensorML XML is needed.
-The BON types map closely to SensorML complex types and so a relatively simple scheme is needed.
-Effective BON classes map to XML elements.
-Features in the class map either to attributes of the XML element or a contained XML element.
-Whether a feature is an attribute is indicated in the class as a comment using @SensorML to make this comment as a mapping instruction.
-Otherwise an XML element is assumed.
-The type names of the BON classes and features are the XML element or attribute names.
-Values.
-Case of names.
-
-BON SEQUENCE types need a special rule for mapping to SensorML.
-
-Connections rule.
+The SensorML BON specifications developed in this project are be found in appendix \ref{appenB}.
 
@@
 This is very error prone and the ability to check that connections are correct to a least some degree is desirable.
-A BON specification can enforce some level of correctness in setting up the connections using preconditions (ensure).
+A BON specification can enforce some level of correctness in setting up the connections using preconditions.
 The source reference must be either an input of the process chain itself or an output of one of the contained components.
 In the BON ProcessChain class a feature AddLink checks that Links are legal before adding the Link to the connections feature.
-This is shown in the BON listing below.
+This is shown in the BON ProcessChain listing below.
 
-\begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON ProcessChaing]  
+\begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON ProcessChain]  
 effective class ProcessChain
@@
 
-\section{SenseTile Thermistor BON SensorML Specification }
-Thermistor can be generated by the following BON class.
-
 \chapter{Results}
@@
 
-A BON specification was developed for the core SensorML types and the SWE Common Data types used in the SenseTile SensorML model.
-The BON specification is included in appendix \ref{appenB} A BON to SensorML XML encoding mapping is also developed and is included in appendix \ref{appenC}.
+A SensorML BON specification was developed for the core SensorML types and the SWE Common Data types used in the SenseTile SensorML model. This specification is included in appendix \ref{appenB}.
 
@@
 Due to the simple model and the rich set of data types is is hard to imagine any sensor or sensor system that could not be described by SensorML.
-It is not intended to describle details of the sensor hardware.
+It is not intended to describe details of the sensor hardware.
 It is also not intended that SensorML provides the framework for encoding the actual sensor observations.
@@
 The Detector process outputs a real value such as a Celsius temperature value.
-eeicpr - tutorial ref.
 The SenseTile model uses a different approach as the unconverted data can also be retrieved from the SenseTile Web Service.
-There are good examples eecipr\footnote{} and tutorials available to help develop SensorML models.
+There are good examples\footnote{http://www.ogcnetwork.net/SWE\_Projects} and tutorials \cite{SMLTutorialref} available to help develop SensorML models.
 
@@
 It does not, for example, use the SensorML support for accurate and reliable processing data.
-It did examine a simple process chain which highlighted some of the difficulties of process chaining described in chapter x.
+It did examine a simple process chain which highlighted some of the difficulties of process chaining described in chapter \ref{BONsec}..
 
@@
 There was little or no documentation that I could find on this and examples of it's use needed some research.
-It would be better if a working example could be provided with the framework and also some basic documentation.
+It would be good if a working example could be provided with the framework and also some basic documentation.
 The VastSensorMLEngine code needed to be slightly modified to parse the SenseTile model files though they passed a validation check\footnote{http://vast.uah.edu/SensorMLforms/validate.jsp} provided by the Vast group.
@@
 \subsubsection{SensorML BON Specification}
-The BON SensorML developed in this project just covers the core of the SenorML types.
+The BON SensorML developed in this project just covers the core of the SensorML types.
 However it has shown how some of the problems of SensorML can be addressed with a formal language such as BON.
 A way to check that connections in SensorML processchains are correct is described using BON assertions.
-The type safety and assertions provided by a BON SensorML specification allows a better way to develop SensoML based sensor sysstems than just using the SensorML XML encoding.
+The type safety and assertions provided by a BON SensorML specification allows a better way to develop SensorML based sensor sysstems than just using the SensorML XML encoding.
 
 A more complete BON SensorML description is needed than provided in this project and there are still many issues that need to be addressed.
-For example how to represent XLINKs in the SensorML BON specification or is there a need to.
+For example how to represent XLinks in the SensorML BON specification or is there a need to.
 Based on the SensorML BON specification developed in this project an inline approach would be necessary rather than reference external XML files general used in the examples.
@@
 
-Streaming the data. See SWE Architecture document for a description of one approach.
-
-Comparing a Restful approach to retrieving sensor observations rather than the SOS approach could be investigated. It may prove to be a simpler way to retrieve the sensor observations.
+Comparing a RESTful\cite{Restref} approach to retrieving sensor observations rather than the SOS approach could be investigated. It may prove to be a simpler way to retrieve the sensor observations.
 
@@
 Complete Weather Station , 2006
-\bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,
 \bibitem{OGCcatref}Open Geospatial Consortium Inc., OpenGIS® Catalogue Services Specification, 2007
 \bibitem{Meyerref}Bertrand Meyer., "Object-Oriented Software Construction", 1997
-\bibitem{Sensorsref}Bertrand Meyer., ", 1997
+\bibitem{Restref}Fielding, Roy Thomas (2000), Architectural Styles and the Design of Network-based Software Architectures, Doctoral dissertation, University of California, Irvine
 

### 2010-05-25 papers/BON_VHDL/paper.tex
Author update and two writing tweaks.

@@
 
-\author{Joseph R. Kiniry and Dragan Stosic\\
+\author{Joseph R. Kiniry\\
+IT University of Copenhagen\\
+Copenhagen S., 2300, Denmark\\
+kiniry@itu.dk\\
+\\
+Dragan Stosic\\
 UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
 University College Dublin,\\
-Belfield, Dublin 4, Ireland,\\
-kiniry@acm.org and dragan.stosic@gmail.com\\
+Belfield, Dublin 4, Ireland\\
+dragan.stosic@gmail.com
 }
@@
 these principles precisely and point out why they are important in
-VHDL-AMS simulation.
+modeling VHDL-AMS.
 
-\paragraph{Seamlessness and Reversibility} \emph{Seamlessness} is the
-use of a continuous process throughout the software life-cycle.
-\emph{Reversibility} is the support for both forward and backward
-development process: from analysis to design and implementation, and
-back.
+\paragraph{Seamlessness} \emph{Seamlessness} is the use of a
+continuous process throughout the software life-cycle.
+
+\paragraph{Reversibility} \emph{Reversibility} is the support for both
+forward and backward development process: from analysis to design and
+implementation, and back.
 

### 2010-06-21 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
new proposed title
mole abstract

--- 2009-10-14 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
@@
 
-\title{Deriving an Executable Model of a Data Stream Protocol from its
-  Specification}
-
-\author{Joseph R. Kiniry, Vieri del Bianco and Dragan Stosic\\
+\title{Agile Formality: A Mole of Software Processes}
+
+\author{Vieri del Bianco, Joseph R. Kiniry and Dragan Stosic\\
   UCD CASL: Complex and Adaptive Systems Laboratory and\\
@@
   Belfield, Dublin 4, Ireland,\\
-  kiniry@acm.org, vieri.delbianco@ucd.ie and dragan.stosic@gmail.com\\
+  vieri.delbianco@ucd.ie, kiniry@acm.org and dragan.stosic@gmail.com\\
 }
@@
 
-  Sensor equipped devices typically communicate with a streaming
-  packet-based network protocol.  The communication is focused on
-  controlling the sensors and reporting the sensors measurements.  The
-  \STSB is an example of such a device: it communicates with an
-  arbitrary computer system, the \STPU, over a USB bus using a
-  streaming packet-based protocol.  The \STU is a custom-designed, but
-  inexpensive, sensor platform: it is composed by a one or more \STSBs
-  paired with a general-purpose small-scale compute node, the \STPU
-  (including everything from PDAs to powerful touchscreen-based
-  portable computers).  There are over a dozen sensors on the \STSB
-  itself, and new sensors can be added either using expansion slots on
-  the \STSB, or adding them directly to the USB bus.  A third-party
-  low-level library creates and parses generic USB communication
-  packets; the library has been provided by the vendor of the USB chip
-  vendor, the USB chip that has been used in the \STSB.
-
-  A custom medium-level Java driver creates and parses the \STSB
-  streaming communication packets.  The driver is formally specified,
-  validated and verified with rigorous unit testing, both hand made
-  and automatically generated.  Because of this process and
-  combination of technologies, we are able to verify a real \ST and a
-  simulated one as well.  The system has been developed concurrently
-  with the protocol specification and the real \STSB, the boards have
-  not been available until the last months of development.  The
-  protocol model is itself an executable stream simulator that can be
-  used to develop new \ST user applications without having the \ST
-  hardware at all.  Consequently, from this single specification, we
-  are able to test that the \ST hardware communicates as expected,
-  that our \ST simulator communicates as expected, and simulate the
-  \ST protocol stream itself.
+%  A punchy abstract that makes the reader want to know more.
+
+  Agile and formal people do not amalgamate nicely.
+  These two communities do not talk to or learn from each other. 
+  We recognize that the problem solving approches and mind attitudes 
+differs widely, but we strongly believe that this clash of view points 
+can be exploited to grow new development process that actually blends, 
+and not simply compose, practices from the two worlds.
+  This paper summarizes a complex case study which shows that it is not 
+only possible, but tasty, to combine the ``garlic'' of formal methods and the 
+``chocolate'' of agile programming... that is, a ``Mole'' of Software Development 
+Processes.
+
+  The case study focuses on the development of a driver to pilot an 
+embedded custom board equipped with more than a dozen sensors: the communication 
+protocol with the board is asynchronous and packet based.
+ The communication is focused on controlling the sensors and reporting 
+the sensors measurements.
+
+  The software driver have been developed fulfilling contrasting requirements: 
+(1) the driver is formally specified, verified and validated;
+(2) a board simulator is built to match the specifications and the board behaviour;
+(3) if the board components is changed during development, the formal 
+specifications and the simulator reflect the changes accordingly;
+(4) strict software delivery deadlines.
+
+  An innovative development process has been drafted to cope with the contrasting 
+requirements of the case study, blending agile and formal practices; the 
+development process has been succesfully applied to the case study yelding promising 
+results.
 

### 2010-06-21 papers/ModelBasedTestCaseGenerationDataStreamProtocol/talk.tex
mole abstract

--- 2010-06-21 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
@@
 
-\title{Agile Formality: A Mole of Software Processes}
-
-\author{Vieri del Bianco, Joseph R. Kiniry and Dragan Stosic\\
-  UCD CASL: Complex and Adaptive Systems Laboratory and\\
-  School of Computer Science and Informatics,\\
-  University College Dublin,\\
-  Belfield, Dublin 4, Ireland,\\
-  vieri.delbianco@ucd.ie, kiniry@acm.org and dragan.stosic@gmail.com\\
-}
+\title{Deriving an Executable Model of a Data Stream Protocol from its
+  Specification}
 
@@
 % \thispagestyle{empty}
-\begin{abstract}
-
-%  A punchy abstract that makes the reader want to know more.
-
-  Agile and formal people do not amalgamate nicely.
-  These two communities do not talk to or learn from each other. 
-  We recognize that the problem solving approches and mind attitudes 
-differs widely, but we strongly believe that this clash of view points 
-can be exploited to grow new development process that actually blends, 
-and not simply compose, practices from the two worlds.
-  This paper summarizes a complex case study which shows that it is not 
-only possible, but tasty, to combine the ``garlic'' of formal methods and the 
-``chocolate'' of agile programming... that is, a ``Mole'' of Software Development 
-Processes.
-
-  The case study focuses on the development of a driver to pilot an 
-embedded custom board equipped with more than a dozen sensors: the communication 
-protocol with the board is asynchronous and packet based.
- The communication is focused on controlling the sensors and reporting 
-the sensors measurements.
-
-  The software driver have been developed fulfilling contrasting requirements: 
-(1) the driver is formally specified, verified and validated;
-(2) a board simulator is built to match the specifications and the board behaviour;
-(3) if the board components is changed during development, the formal 
-specifications and the simulator reflect the changes accordingly;
-(4) strict software delivery deadlines.
-
-  An innovative development process has been drafted to cope with the contrasting 
-requirements of the case study, blending agile and formal practices; the 
-development process has been succesfully applied to the case study yelding promising 
-results.
-
-\end{abstract}
 
@@
 sensor networking.
-
-This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed, but inexpensive,
-sensor platform (called the \ST) paired with general-purpose
-small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers.  There are over a dozen
-sensors packaged on the \ST itself, the board is expandable as well,
-new sensors can be added to it easily.
 
@@
 development constraints (rapid development in hardware software
-co-design), but they are all focused on big-scale systems.  The
-hardware software formal co-design methodologies usually have a common
-structure\cite{Slomka2000,Hoffman2001}:
-
-\begin{itemize}
-\item High level specification of the system.
-\item Translation of the high level specification to lower level
-  specifications: hardware and software counterparts.
-\item Generation of software code.
-\item Run hardware specification into a simulator.
-\item Build hardware component.
-\end{itemize}
+co-design), but they are all focused on big-scale systems
+\cite{Slomka2000,Hoffman2001}.
 
@@
 co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
-on a specification that aims to be as complete as possible. The
-specification, once it is finalized, is used to define the remaining
-development tasks, through automatic code generation, refinement,
-model checking, etc.
+on a specification that aims to be as complete as possible.
 
@@
 
-\paragraph*{Outline}
-
-In section \ref{sec:data_stream_and_protocol_description} the data
-stream model and protocol are presented.
-
-In section \ref{sec:data_stream_and_protocol_specification} the data
-stream model and protocol are specified.
-
-In section \ref{sec:test_cases} the test cases structure is presented,
-for both the hand made and the generated test suites.  Both the
-approaches are analyzed, their pro and contra are discussed, as well
-as the expected advantages to use them in combination.
-
-In section \ref{sec:test_cases_on_the_job} the execution of the test
-suites is described; the results and measures obtained by their
-execution are presented.
-
-In section \ref{sec:test_cases_retrospectives} the data collected in
-the previous section are analyzed, both from a qualitative and a
-quantitative points of view. The effectiveness of the approach is
-discussed as well.
-
-In section \ref{sec:related_works} the related works that have not
-been mentioned in the current section are presented and analyzed.
-
-Finally in section \ref{sec:conclusion} conclusions are drawn, and
-ongoing and future works on similar topics are introduced.
-
 % =====================================================================
@@
 
-\STSB and \STPU communicates via an Universal Serial Bus (USB)
-interface.  A proprietary protocol is used on top of the USB serial
-channels, the protocol is specified by
-FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
-is the manufacturer of the USB controller chip used in the \STSB, the
-FT2232H\cite{ftdi_ft232h_2009}.
+A custom protocol has been developed to manage the entire \STSB, built
+on top of a USB serial channel; the protocol main function is to read
+sensor data and control the \SB.  The protocol is asynchronous and
+packet based, the packets have a fixed length.  The protocol has been
+divided in multiple layers to ease its formal description.
 
-The proprietary protocol enables the control and management of some of
-the functionalities in FT232H chip (such as internal EEPROM read and
-write, reset of the chip, etc.), as well as the definition of the
-communication protocol and properties to be used in the asynchronous
-channels (input and output channels) provided by FT232H.  These
-functionalities are accessed via an Application Program Interface, the
-API is provided by a driver developed directly by FTDI, which is
-available for various systems (Linux 32 and 64 bits, Mac OS, Windows)
-\cite{ftdi_d2xx_api_2009}.
-
-Finally, a custom protocol has been developed to manage the entire
-\STSB, built on top of the serial channels provided by the API.  The
-custom protocol is meant to be used to read sensor data observations,
-enable or disable sensors, configure on-board components (as the
-Field-Programmable Gate Array and the micro processor); the protocol
-main function is to read sensor data.
-
-The protocol is asynchronous and packet based, the packets have a
-fixed length.  The protocol has been divided in multiple layers to
-ease its formal description.  The rationale is the following: mixing
-up different abstraction levels of a protocol description can lead to
-overly complex and unmanageable descriptions, especially when dealing
-with formal ones.  The layers identified in protocol are the
-following:
-
-\begin{description}
-\item[Packet byte structure]: the internal representation of the
-  packet.
-\item[Packet info structure]: the meaningful fields contained in the
-  packets.
-\item[Single packet rules]: the content acceptable values, and how
-  they influence each other.
-\item[Packet sequence rules]: the content acceptable values, based on
-  values of previously received packets.
-\item[Packet input output asynchronous rules]: the content acceptable
-  values and reaction constraints on output packets, based on
-  previously transmitted input packets.
-\end{description}
-
-We will examine the output sensor data packet, since it is the most
-used and interesting one; we won't describe input or other output
-packets.  We will examine the first three layers, that is: the packet
-byte structure, the packet info structure and the single packet rules.
-
-The packet byte and info structures of the output sensor data packet
-reflect the board capabilities and built in sensors.  Essentially a
-single packet has to accommodate various types of data:
-
-\begin{itemize}
-\item Fast data rate streams: audio data, 48KHz (standard audio) or
-  96KHz (ultra audio).
-\item Medium data rate streams: 5KHz.
-\item Slow data rate streams: temperature, pression, accelerometer,
-  etc.; all these sensor have a data rate lower than 5 KHz.
-\item Metadatas: metadatas on sensors and board state.
-\end{itemize}
-
-The internal structure of the package (as we will see) is strictly
-fixed, even if the data streams have different data rates.  This
-choice originates from a trade-off between two constraints: the
-efficiency of the transmission (redundant data to be minimized) and
-the complexity to build a package (complexity to be minimized).  The
-complexity minimization constraint rationale is the following: the
-FPGA on the \STSB builds the packet, but the FPGA is also to be used
-for various other specific tasks; it is important to maintain the
-building packet code on the FPGA as simple and short as possible, to
-leave enough space to implement other tasks.  The Output sensor data
-packet structure is shown in Table \ref{tab:data_packet_structure},
-the packet total length is 1024 bytes (512 16 bit words).
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet structure.}
-  \label{tab:data_packet_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0-1 & sentinel\\\hline
-      2-7 & slow data\\\hline
-      8-11 & metadata: board state\\\hline
-      12-13 & reserved\\\hline
-      14-15 & metadata: board state\\\hline
-      16 & reserved\\\hline
-      17-22 & frame 1\\\hline
-      \ldots & \ldots\\\hline
-      503-508 & frame 82\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-The frames accommodate data from the fast and medium data rate streams
-(there are 4 fast data rate stream and 8 medium data rate stream
-channels).  There are 82 frames in a packet.  The frame structure is
-shown in Table \ref{tab:frame_structure}.
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet frame structure.}
-  \label{tab:frame_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0 & metadata: frame description\\\hline
-      1 & medium data rate data sample\\\hline
-      2-5 & fast data rate data samples\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-Looking at the structure described and the data rates frequencies, it
-can be seen that there must be some subtle rule, to make all the
-figures to add up; that is, to be capable to accommodate three
-different rates outputs in a fixed structure, fixed length packet.
-The subtleties lies in two forms of redundancies, for slow and medium
-data rates:
-
-\begin{itemize}
-\item Slow data rates: slow data can be found at the beginning of each
-  packet, the data is simply repeated (that is, it is equal to the
-  previous packet in the stream) when no new data is available.
-\item Medium data rates: the 8 medium rate channels are multiplexed
-  together into the medium data rate data sample of each frame.  The
-  medium data rate sample channel is a piece of info to be extracted
-  from the frame description. Besides, the medium data rate data
-  sample could contain no data at all, this information is to be
-  extracted from the frame description. So, one frame could contain a
-  medium data rate sample from channel 0, the following frame a sample
-  from channel 5, the following frame no sample at all, etc.
-\end{itemize}
-
-The single packet rules delimit the boundaries of the values
-obtainable from the packet: each defined sensor represented in the
-packet, as well as the metadata describing the \STSB and the packet
-and frame contents, are constrained by a range of acceptable values.
-There are also rules affecting more than one value, these are
-essentially validity rules (i.e. if a specific sensor is active the
-reading value has to be in a defined range), and rules specifying a
-correct sequence of frames (i.e. there can be only one switch in
-sampling rate in the frame sequence of a single packet).
+We will examine the output sensor data packet.  The packet byte and
+info structures of the output sensor data packet reflect the board
+capabilities and built in sensors.  Essentially a single packet has to
+accommodate various types of data streams chunks.
 
@@
 \end{itemize}
-
-We needed specifications capable to tackle properly all the protocol
-layers.  Since different layers have different characteristics,
-different specification approaches has been chosen for each layer: the
-specification are context dependent.
 
@@
 
-JML has been chosen because it is capable to specify Java code: JML is
-a behavioral interface specification language.  JML specification
-language combines the design by contract approach (applied to Java
-language) and the model-based specification approach of interface
-specification languages.
-
-Many tools able to support JML specification language are available,
-focusing on different aspects of the language.  JML2 suite of
-tools\cite{BurdyEtal05-STTT}
+Many tools are able to support JML specification language are
+available, focusing on different aspects of the language.  JML2 suite
+of tools\cite{BurdyEtal05-STTT}
 \footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common
@@
 
-The dual testing approach was thought a winning option, when compared
-to hand based tests or completely automated tests used alone.  \ST
-stream communication protocol is characterized by a well definite
-packet structure, and a well definite semantic on the packet contents
-as well.  On the other hand, in the protocol, there are not strong
-constraints on synchronization, since the communication channel is
-splitted in two (in and out) separate concurrent and asynchronous
-channels.  The lack of hard synchronization constraints reduces the
-advantage to use a classical state machine based specification (as the
-one presented in \cite{Hubbers2004}), hence this option has been ruled
-out.
-
-Hard development time constraints were also a precise requirement; the
-time constraints pushed us towards an incomplete (\emph{just good
-  enough}) specification.  In \STSB driver JML was used to specify
-very simple behaviors (method oriented) on the whole driver
-implementation, and more complex behavior on the driver core
-interfaces.  The software system parts not specified by the JML2
-specification were specified using hand made unit tests.
-
 Hand made unit tests specifies the behavior not specified by formal
@@
 
-The test suite generated is a collection of fine grained unit tests.
-For each public method a set of tests is generated.  The specified
-behavior of the method is checked using the JML2 run time assertion
-checker, hence, an oracle is not needed.  The generated tests send
-messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation
-exceptions;
-if a precondition is not met the test is not run, if the
-precondition is met the test is run and if the postcondition is not
-met an exception is fired and the test fails.  The test data
-used to build the tests is the only part that is required to be
-provided by the user.
-
 The resulting test package structure is shown in Figure
@@
 are generated tests, they test (hence use) the implementation
-\lil{SimulatorDriver}, and have been generated using the
-corresponding specification \lil{SimulatorDriverS}.  The
-generated tests do not maintain the inheritance structure of the code
-and the specification.
+\lil{SimulatorDriver}, and have been generated using the corresponding
+specification \lil{SimulatorDriverS}.  The generated tests do not
+maintain the inheritance structure of the code and the specification.
 
@@
 
-JML specifications, the simulators, and both test suites has been used
-to:
-
-\begin{itemize}
-\item verify hardware implementation;
-\item verify driver implementation, with and without the physical
-  board;
-\item guide driver development.
-\end{itemize}
-
-The first \STSB prototype delivered was meant to build empty packets,
-but the packets were built in a way not consistent with the
-specification, the sentinel values used to delimit a packet were not
-correct.  The board implementation errors were detected by simply
-using the driver implementation, the \lil{StreamDriver} package.
-
-The second prototype delivered was meant to build packets with values
+The first prototype delivered was meant to build packets with values
 taken from real sensors installed on the board.  The only sensors that
@@
 
-The second prototype respected the packet byte structure
+The prototype respected the packet byte structure
 specification, but was not fully compliant to the packet info
@@
 specification, and 2 protocol errors regarding the single packet rules
-has been found. No other errors, except the mentioned ones, have
-been later on discovered.
+has been found. No other errors, except the mentioned ones, have been
+later on discovered.
 
 The hand made test suite is composed by over 140 tests, while the
-generated one is 100 times bigger, totaling over 14000 tests.  The big
-number of generated tests (1 to 100 ratio), can be understood looking
-closely at how the generated tests are created by the JML
-framework. The generated tests are exploring the possible input
-combinations on public methods, combining the type data ranges that
-have been specified for the test suite (see \cite{Cheon-Leavens02} for
-more details on how the JML framework works on generating unit
-tests). For instance, let's suppose we want to test a method
-\lil{m(int p1, int p2)} for a class \lil{C}; the data ranges specified
-for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the
-instances \lil{\{c1,c2,c3,c4\}}. The JML framework generates $24 = 3
-\times 3 \times 4$ tests, since they are the combination of the data
-ranges involved (the data range of the parameters and the data range
-of the type owning the method under test). Fortunately, not all the
-generated tests are executed.
+generated one is 100 times bigger, totaling over 14000 tests.
 
-A generated test, as it was previously mentioned (see section
-\ref{sec:test_cases}), is not executed, unless the preconditions of
-the called method are satisfied. Thus, depending on the preconditions
-specified for a method, the number of the generated tests that are
-actually executed in the test suite can be significantly smaller 
-compared to the total number of generated tests.
-
-To compare in a quantitative way the effectiveness of the test suites,
-code coverage metrics have been used; this is not the only measure
-that have been considered, since it is almost unanimously accepted
-that code coverage alone cannot asses the quality of a test
-suite\cite{Marick1999,Chockler2006}.
-
-Various coverage metrics exist, we analyzed the test suites with
-statement code coverage (one of the simplest types)\footnote{The tool
+We analyzed the test suites with
+statement code coverage \footnote{The tool
   used to obtain statement code coverage metrics is
   \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage
-reports whether each executable statement has been encountered. This
-code coverage metric is easy to implement, but lacks the capability to
-explore the control flow graph associated to the code, that is,
-decision branches are not explored. The coverage results are reported
+reports whether each executable statement has been encountered. 
+The coverage results are reported
 in the Table \ref{tab:statement_code_coverage}.
@@
 
-Test suites execution time are very different, the hand made test suite, with the
-run time assertion checking active runs completely in less than one
-minute on an average PC, while the generated test suite runs in more
-than 20 minutes (the rough ratio is $ 1 : 65 $). Most of the time is
-spent on setting up the suite, only considering the real execution of
-the tests, with no suite setup, the execution time reduces to less
-than 4 minutes (the rough ratio is $ 1 : 10 $).
-
 % =====================================================================
@@
 values. This is because of two different effects: non public methods,
-and interfaces. The effect of non public methods (to be precise:
-package Java visibility) is that these methods are taken into account
-as well as public and protected ones during code coverage calculation;
-but these methods are not part of the interface, they are not meant to
-be called by the system, and they are only used in object
-initialization or in object setups performed during unit tests. In
-fact there are package and private code blocks that are not reachable
-by construction from public and protected methods, and this is a
-precise characteristic reflecting the code architecture, meant to ease
-class instances setup and configuration for testing purposes. The used
-statement coverage tool cannot be configured to distinguish these code
-blocks, so this effect is unavoidable. The effect is mainly seen in
-low \lil{SimulatorDriver} code coverage figures: the
-\lil{SimulatorDriver} package has plenty of setup and initialization
-non public methods.
+and interfaces.
 
-Regarding interfaces, an interface contains no statements, so when an
-interface method is called, it is the method of the implementation
-class used that is actually covered. This effect is visible in
-\lil{Driver} coverage figures: \lil{Driver} package contains
-interfaces (that are not counted at all) and some very simple real
-class (exceptions) that are not throughly tested. The net result is
-extremely low statement code coverage figures because the very simple
-classes are the only ones that are actually covered. This effect could
-be easily avoided writing specific hand made tests for the simple
-classes and including them in the classes for which tests are
-generated.
-
-The statement code coverage figures show that the test suites do a
-good job on the most important package, that is, the
-\lil{StreamDriver}. The generated test suite reaches almost $ 80 \% $,
-while the hand made tests have only a slightly lower coverage ratio: $
-76 \% $. But the combination of test suites is interesting, raising
-the coverage ratio to $ 88.5 \% $. This is a clear indication that one
+But the combination of test suites is interesting, indicating that one
 test suite is not completely overlapping the other: the test suites
 are not to be considered alternatives, but they have to be used
-together to achieve the higher benefits, they are complementary. This
-consideration can be confirmed looking at the figures obtained for
-\lil{SimulatorDriver} package as well.
-
-The qualitative difference of the two test suites can be understood
-analyzing how the suites are built. The hand made test suite primarily
-focuses on small features and use cases fragments that the system must
-implement, or for which an exceptional case has been specified. A test
-case is to be considered a rough specification written in imperative
-style. The common structure requires an object initialization,
-followed by a hopefully not too long sequence of public methods taking
-an object (or a set of objects) into a desired state; than the method
-under test is called and the object state is finally inspected to
-verify if it is in the expected state. Mock objects can be used to
-limit the dependencies. The test suite should be composed only by unit
-tests, but usually some tests are more complex to be called proper
-unit tests; they can be seen as simple integration tests.
-
-The generated test suite has a more focused and limited scope: since
-only one method call is performed in each test, hence it is difficult
-to explore complex behaviors. It is true that with a complete
-specification, and specifying large enough data ranges (the
-appropriate data ranges are combined to generate the tests), the
-generated tests could theoretically achieve the same expressive power
-as the hand made ones, but the price to pay would be expensive: much
-longer and more complex specifications, requiring more time to be
-designed and or updated, and more tests to be run, increasing the
-already not so short computation time of the generated test suite.
+together to achieve the higher benefits, they are complementary.
 
@@
 generated test suite, checking simpler behaviors, and the hand made,
-checking the more complex ones.  During the development of the driver
-code and the simulators, the generated test suite was helpful to test
-the incrementally growing specifications and the finer details of each
-method, while the hand made one was helpful to verify the behavior on
-complex methods calls on objects initialized to specific
-states. Having an hand made test suite also make the code style more
-test oriented, which usually leads to a better quality and more
-modularized source code\cite{Binder1999}.
-
-The test suites has been used to verify different parts of the
-systems, that is, the real system as well as the simulators.  The
-multiple simulators and multiple test suites approach is not new,
-since it has been often used in embedded systems software
-testing\cite{Broekman2002}.  This is a general approach that can (and
-should) be used to verify both the simulators built and the real
-systems.  It can be easily implemented only if the code has been
-structured following the already mentioned dependency inversion
-principle\cite{martin1996dependency} (see Figure
-\ref{fig:class_diagram_main}).
-
-The combined test approach, driven by the packet protocol
-specification, has been proved successful on verifying the protocol
-itself and the software implementing it.  Nevertheless, it has to be
-noted that the highest levels of the protocol are more problematic
-(see \ref{sec:data_stream_and_protocol_description}: ``Packet sequence
-rules'' and ``Packet input output asynchronous rules'').  On these
-levels, the protocol properties regard sequences of packets, and
-testing them can be tricky, since the sequences of packets are not
-referred to explicitly, in the implementation.  Two possible
-strategies can be used:
-
-\begin{itemize}
-\item Translate the packet sequence properties in local packet
-  properties (when this is possible).  The advantage is to have
-  properties shown as the usual pre-conditions, post-conditions and
-  invariants.  However, the localized properties can be more complex
-  and less intuitive and comprehensible than the original ones;
-  finally, some of the global properties could prove themselves
-  impossible to be translated into local ones.
-\item Use separate utility classes, meant to verify a sequence of
-  packages.  This approach is not clean, since it requires adding and
-  using new classes, unneeded by the core implementation, to verify
-  the protocol.  It has the advantage of keeping the properties that
-  have to be verified mostly unchanged.
-\end{itemize}
-
-One problem we noticed regards the execution time of the generated
-tests.  It is long for two reasons:
-
-\begin{itemize}
-\item Number of the tests generated. This problem is not directly
-  addressable, since it is the heart of the test suite generation
-  creation method itself: generating all the combinations of possible
-  inputs for each method under test.
-\item Runtime assertion checker. Each test is executed while RAC
-  checks all the relevant properties. This problem, again, is not
-  directly addressable. We can speculate that if the RAC assertions
-  could be triggered on and off on single properties in a dynamic way
-  (during runtime), a test suite would take advantage of this
-  possibility, turning on the RAC only on the properties that have to
-  be verified for the classes or methods under tests (and not for all
-  the involved classes and methods); this could save time during the
-  test suite execution.
-\end{itemize}
-
-% =====================================================================
-\section{Related works}
-\label{sec:related_works}
-
-JML has been used previously to verify a communication protocol in
-\cite{Hubbers2004}, but the protocol is a smart card protocol, and it
-is well represented through state machines, which are then refined
-into JML properties.  The approach is in waterfall style; than the
-need to develop the protocol, the specification and the implementation
-concurrently is not addressed.  No hand made or generated test suites
-are used, the code is specified with JML assertions, refined by the
-state machine specifications, and then checked statically with
-ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2
-tools RAC\cite{BurdyEtal05-STTT}.
-
-The JML-JUnit automatic test suite generation tool has not been widely
-used or assessed. It is described by its own creators in
-\cite{Cheon2002,Cheon2004,Cheon2005}; it has been used in
-\cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test
-suite, and, more importantly, only against proof of concepts system,
-it has never be proved against real complex systems.  A brief
-assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the
-analysis is based on injecting random generated errors in the source
-code, and then evaluating the ratio of the errors found by the test
-suite: the results shows that the JML-JUnit tool is not effective
-enough if used as a stand alone tool. No attempts on running or
-evaluating the generated test suite alongside with an hand made test
-suite are made.
-
-% =====================================================================
-\section{Conclusion}
-\label{sec:conclusion}
-
-\todo{specification and hand tests useful to drive, specification and
-  hand tests and simulator useful to test, specification and generated
-  tests useful for statement coverage and easy methods, specification
-  useful to test hardware, hand useful to spec, spec useful to
-  hand... test driven spec}
-
-\section{Next papers}
-
-\todo{remove and push this on a new paper}
-
-\paragraph*{TDS: Test Driven Specification}
-
-Approach:
-\begin{itemize}
-\item more specification used together: formal and not formal
-\item formal specification is not complete: it is just good
-  ``enough'', static checkers unfortunately cannot be used, since the
-  specification is partial
-\item use runtime assertion checker for formal specifications, on
-  tests
-\item use generated tests for formal specifications
-\end{itemize}
-
-Who drives what:
-\begin{itemize}
-\item specification (jml - formal specification or unit test -
-  examples fragments) driven development
-\item test driven specification
-\item formal specification is not complete: it is just good
-  ``enough'', static checkers unfortunately cannot be used, since the
-  specification is partial
-\item use runtime assertion checker for formal specifications, on
-  tests
-\item use generated tests for formal specifications
-\end{itemize}
-
-Benefits:
-\begin{itemize}
-\item real agile approach
-\item real formal approach
-\item use best of both worlds, do not try to use only on of them
-\end{itemize}
-
-Case example: actual paper, modified to fit the new parts, especially
-the test driven specification part.
-
-\paragraph*{BON}
-
-\todo{describe approach: high level informal specifications with BON,
-  iterative refinements of code, to distill test cases and formal
-  specification of the interfaces}
+checking the more complex ones.
 

### 2010-06-29 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
some progress... but still a lot work to do

@@
 
-\title{Agile Formality: A Mole of Software Processes}
+\title{Agile Formality: A Mole of Software Verification Practices}
 
@@
 
-% ======================================================================
-% \thispagestyle{empty}
 \begin{abstract}
-
-%  A punchy abstract that makes the reader want to know more.
 
@@
 These two communities do not talk to or learn from each other.
-  We recognize that the problem solving approches and mind attitudes 
-differs widely, but we strongly believe that this clash of view points 
-can be exploited to grow new development process that actually blends, 
-and not simply compose, practices from the two worlds.
-  This paper summarizes a complex case study which shows that it is not 
-only possible, but tasty, to combine the ``garlic'' of formal methods and the 
-``chocolate'' of agile programming... that is, a ``Mole'' of Software Development 
-Processes.
-
-  The case study focuses on the development of a driver to pilot an 
-embedded custom board equipped with more than a dozen sensors: the communication 
-protocol with the board is asynchronous and packet based.
- The communication is focused on controlling the sensors and reporting 
-the sensors measurements.
-
-  The software driver have been developed fulfilling contrasting requirements: 
-(1) the driver is formally specified, verified and validated;
-(2) a board simulator is built to match the specifications and the board behaviour;
-(3) if the board components is changed during development, the formal 
-specifications and the simulator reflect the changes accordingly;
-(4) strict software delivery deadlines.
-
-  An innovative development process has been drafted to cope with the contrasting 
-requirements of the case study, blending agile and formal practices; the 
-development process has been succesfully applied to the case study yelding promising 
-results.
+We recognize that the problem solving approches and mind attitudes differs widely, but we strongly believe that this clash of view points can be exploited to grow new development process that actually blends, rather than fight or simply compose, practices from the two worlds.
+This paper summarizes a complex case study which shows that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming...
+that is, a ``Mole'' of Software Verification Practices.
+
+A complex case study was used as a test bench of the development process.
+The case study focuses on the development of a driver to pilot an embedded custom board equipped with more than a dozen sensors: the communication protocol with the board is asynchronous and packet based.
+The communication is focused on controlling the sensors and reporting the sensors measurements.
+
+The software driver have been developed fulfilling contrasting requirements: (1) the driver is formally specified, verified and validated; (2) a board simulator is built to match the specifications and the board behaviour; (3) if the board components is changed during development, the formal specifications and the simulator reflect the changes accordingly; (4) strict software delivery deadlines.
+The requirements build up an environment that could benefit from both an agile and a formal approach to software development and verification.
+
+An innovative development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal verification practices; the development process has been succesfully applied to the case study yelding promising results.
 
@@
 
-The \textbf{UCD CASL SenseTile System} is a large-scale,
-general-purpose sensor system to be installed at the University
-College Dublin in Dublin, Ireland.  Our facility aims to provide a
-scalable and extensible infrastructure for performing in-depth
-investigation of both the specific and general issues in large-scale
-sensor networking.
-
-This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed, but inexpensive,
-sensor platform (called the \ST) paired with general-purpose
-small-scale compute nodes, including everything from low-power PDAs to
-powerful touchscreen-based portable computers.  There are over a dozen
-sensors packaged on the \ST itself, the board is expandable as well,
-new sensors can be added to it easily.
-
-Because of hard time constraints, and the initial unavailability of
-the custom board, we needed to progress concurrently with all the
-development tasks:
-
-\begin{itemize}
-\item specification of the communication protocol;
-\item development of the physical board specification and following
-  realization;
-\item development of the board embedded software;
-\item development of the software driver to communicate with the
-  board;
-\item development of the software simulators of the board.
-\end{itemize}
-
-The development of the physical board, along with its embedded
-software, has been done by an external manufacturer, thus it is not
-directly taken into account here.
-
-There exist various approaches in literature addressing this kind of
-development constraints (rapid development in hardware software
-co-design), but they are all focused on big-scale systems.  The
-hardware software formal co-design methodologies usually have a common
-structure\cite{Slomka2000,Hoffman2001}:
-
-\begin{itemize}
-\item High level specification of the system.
-\item Translation of the high level specification to lower level
-  specifications: hardware and software counterparts.
-\item Generation of software code.
-\item Run hardware specification into a simulator.
-\item Build hardware component.
-\end{itemize}
-
-Even when considering the development of the device hardware, embedded
-software and software driver as separate entities, that possibly have
-to be developed concurrently, the approaches proposed are similar in
-structure to the one seen in the case of hardware software
-co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus
-on a specification that aims to be as complete as possible. The
-specification, once it is finalized, is used to define the remaining
-development tasks, through automatic code generation, refinement,
-model checking, etc.
-
-This approach has been considered not feasible nor convenient if
-applied to our case since it does not respect some of the
-requirements: first of all, we want to develop all the software
-components concurrently, we want to enhance the protocol specification
-incrementally and we are not able to provide a complete specification
-from the beginning, hence, a properly defined specification is not
-available in advance. Secondly, we are not interested in direct code
-generation from specification, since we do not want to have any
-dependencies on code generators; in fact the programming language to
-be used to develop the driver has been postponed for a while.
-
-To tackle the complexity of the concurrent development tasks, the
-formal model of the protocol was used as a focus point that led all
-the other development tasks. As it can be seen, the central role of
-the specification is unchanged; what is changed is how and when the
-specification is developed.
-
-The protocol to specify has been divided in various (thin) layers,
-each of the layer has been specified with a different lightweight
-approach: some of the layers have been specified formally, while
-others have been specified in an unprecise way.  The formal
-specifications have been used to check at run time the correctness of
-the \SB, the driver and the simulators.  The formal specifications
-have also been used to automatically generate a test suite to verify
-the driver and the simulators, and eventually to verify the
-specification itself.  An hand made test suite has been used to verify
-the driver and the simulators as well.  The generated and hand made
-test suites have been used together, complementing each other.
-
-The approach proved itself effective, at the point that, when the
-board has been finally released, only one small defect has been found
-in the driver, while the errors and not agreed changes introduced in
-the board and in the protocol it implemented, have been discovered
-immediately, with no additional effort.  Besides, the simulators are
-able to feed with correct data the rest of the system.
-
-\paragraph*{Outline}
-
-In section \ref{sec:data_stream_and_protocol_description} the data
-stream model and protocol are presented.
-
-In section \ref{sec:data_stream_and_protocol_specification} the data
-stream model and protocol are specified.
-
-In section \ref{sec:test_cases} the test cases structure is presented,
-for both the hand made and the generated test suites.  Both the
-approaches are analyzed, their pro and contra are discussed, as well
-as the expected advantages to use them in combination.
-
-In section \ref{sec:test_cases_on_the_job} the execution of the test
-suites is described; the results and measures obtained by their
-execution are presented.
-
-In section \ref{sec:test_cases_retrospectives} the data collected in
-the previous section are analyzed, both from a qualitative and a
-quantitative points of view. The effectiveness of the approach is
-discussed as well.
-
-In section \ref{sec:related_works} the related works that have not
-been mentioned in the current section are presented and analyzed.
-
-Finally in section \ref{sec:conclusion} conclusions are drawn, and
-ongoing and future works on similar topics are introduced.
+% agile and formal difficult to blend: mind attitudes
+Agile and formal development methodologies usually do not blend together easily.
+This is because of several reasons, the most important being a difference of mind attitudes.
+Formal methodologies favor an in-depth think first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall style of development.
+Once a formalization has been developed, several verification steps can be fully automated.
+Agile methodologies favor an higly incremental and iterative approach to cope with changing requirements and precise deadlines; in agile methodologies the problem is only partially understood, focusing on the aspects that need to be implemented in the current iteration, the solution initally developed is usually not optimal at all, but than it gets refined through continuous code refactorings.
+Test suites are used for several purposes, the most important being verification, documentation, and requirement specification, and they are striclty hand made.
+
+% problems would benefit form agile and formal mix, characterize the problem
+The two worlds seem irreconcilable, nevertheless, there exist problems that would greatly benefit from both.
+The problem needs to be at least instable in the requirements definition, constrained by deadlines that cannot be postponed, but also characterized by artifacts that needs to be formally specified and verified.
+The complex problem addressed in the case study fits these needs.
+
+\todo{describe, ``Formal Versus Agile: Survival Of the Fittes'', blending advantages and potentials}
+
+% what is needed, high level
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies needs to be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), need to be adapted to a higly iterative one.
+
+% what needs to be considered: TDD
+The most common software verification practices in agile methodologies need to be considered as well when tring to integrate formal and agile methodologies, since they often are a fundamental part of the development process.
+The most significant verification practice is Test Driven Development (TDD), TDD is a software development technique, originally defined in Extreme Programming (XP) methodology, characterized by writing unit tests before actually writing the code.
+Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
+The test driven apprach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is consider a good software development technique to be used on its own, regardless of the surrounfing development process.
+It is to be noted that, depending on the abstraction level, the test driven approach can introduce activities cycles that can be as long as a complete iteration (functional tests), or as short as a couple of minutes (small increment on a unit test).
+
+% previous very related work
+Some tentative attempts to reconcile formal and agile methodologies have been developed.
+The recurrent approach of the attempts consists in removing the hand made test suites, substituing them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
+
+\todo{describe and reference ``An Agile Formal Development Methodology'', perfect match}
+
+\todo{describe and reference ``Formal Extreme (and Extremely Formal) Programming'', perfect match}
+
+\todo{describe and reference ``XFM: An Incremental Methodology for Developing Formal Models'', perfect match}
+
+% previous work: a common approach
+Unfortunately, the agile and formal reconciliation attempts that have been defined so far are not entirely satisfing.
+As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
+The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract precondition, postconditions and invariants) \todo{reference Meyer} and test suites automatic generation \todo{references, JML, Eiffel, \ldots}.
+At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development \todo{references, ESC/2, \ldots}.
+
+% previous work: why is not satisfying
+The methodologies are sound, but they share some common problems: the loops between development artifacts and related activities (code, tests, design, refactoring), especially the closed and tight ones, typical of TDD, are completely lost.
+This is not a minor issue because it is believed that these loops are the inner engine of many agile methodologies, imposing to the development process an iterative and incremental pace \todo{find meaningful references, I know there are some}.
+Beside possibly disrupting some of the agile methodologies inner mechanincs, only the verification purpose of the tests is mantained, that is, the documentation and requirement specification purposes are lost and need to be moved on top of different artifacts, usually the formal specification.
+Finally a complete system formal specification is needed to apply these methodologies, but this requirement is hard to fulfill since the effort required to write a complete formal specification in most cases is much greater than writing a suite of tests.
+
+% what we want to achieve
+Thus, the objective set is different: we want to maintain an higly iterative and incremental development process, we want to connect development artifacts and activities (code, requirements, tests, formal specifications, refactorings, run-time verification, automatically generated tests, \ldots) through activities cycles similar to what we can find in agile methodologies (especially in XP), we want to be able not to formally specify the complete system in order to cope with constrained resources and
+
+% ======================================================================
+\section{Agile Methodologies}
+\label{sec:agile_methodologies}
+
+\todo{quick introduction}
+
+\todo{define what we are looking at: verification, cycles and dependencies between artifacts}
+
+\todo{the cycles in SCRUM and XP}
+
+\todo{TDD}
+
+% ======================================================================
+\section{Formal Specifications With JML}
+\label{sec:formal_specifications_using_jml}
+
+\todo{quick introduction}
+
+\todo{examine language, really short introduction}
+
+\todo{examine tools, included, JML2, and escluded, ESC/2, ones}
+
+% ======================================================================
+\section{Blending Formal And Agile Development}
+\label{sec:blending_formal_and_agile_development}
+
+\todo{describe loop importance}
+
+\todo{describe partial specification importance, and consequences}
+
+\todo{consequences: no complete formal specification, no theorem provers, no model checking, yes run time assertions, yes generated tests}
+
+\todo{we need both worlds, together}
+
+\todo{FDD: formal specification driven development (formal -> hand made tests -> code, design, refactorings, run-time verification)}
+
+\todo{FDD: formal specification driven development (formal -> generated tests -> code, design, refactorings, run-time verification)}
+
+\todo{TDD: test driven development, classical TDD (hand made tests -> code, design, refactorings)}
+
+\todo{increase code with formal specifications, TDF: test driven formal specification (hand made tests -> formal, design, refactorings, partial run-time verification)}
+
+\todo{result: formal: on some parts of the system}
+
+\todo{result: tests: hand made tests, generated tests}
+
+\todo{result: hand made test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
+
+% =====================================================================
+\section{A Real And Complex Case Study: Rapid Development In Small Scale Hardware Software Co-design}
+\label{sec:a_real_and_complex_case_study}
+
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system to be installed at the University College Dublin in Dublin, Ireland.  
+Our facility aims to provide a scalable and extensible infrastructure for performing in-depth investigation of both the specific and general issues in large-scale sensor networking.
+This system integrates a sensor platform, a datastore, and a compute farm.  
+The sensor platform is a custom-designed, but inexpensive, sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.
+There are over a dozen sensors packaged on the \ST itself, the board is expandable as well, new sensors can be added to it easily.
+
+The case study is focused in building the sensor board and its software driver side by side.
+Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board; (4) board embedded software; (5) communication protocol software driver; (6) development of software simulators of the board.
+
+The development of the physical board, along with its embedded software (development tasks 2, 3, 4), is done by an external manufacturer, thus it is not directly taken into account here.
+The specification of the communication protocol (development task 1) is a joint effort, while the remainang tasks (5, 6) can be developed in isolation.
+
+\todo{illustrate dependencies: 3<>2<>4<>1, 1>5, 1>6}
+
+\todo{explain 3<>2<>4<>1: difficulties encountered}
+
+\todo{explain 1>5, 1>6: clean separation, but 1 is unstable}
+
+There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software
+co-design), but they are all focused on big-scale systems.
+The hardware software formal co-design methodologies usually have several common development artifacts\cite{Slomka2000,Hoffman2001}: an high level specification of the system, a translation (refinement) of the high level specification to low level ones (hardware and software counterparts), the possibility to generate software code from the low level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
+
+When considering the development of the device and of its software driver as separate entities, that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
+
+As it has been already pointed out, a specification as complete as possible is not feasible nor convenient: we want to enhance the protocol specification incrementally, and we are not able to provide a complete specification.
+
+The protocol specification is divided in various (thin) layers, and each of the layer has been specified with a different lightweight approach: some of the layers have been specified formally, while others have been specified in a less formal way.
+The formal specifications are used to check at run time the correctness of the \SB, the driver and the simulators.  
+The formal specifications are used to automatically generate a test suite to verify the driver and the simulators, and eventually to verify the
+specification itself.  
+Hand made test suites are used to verify the driver and the simulators as well.  
+The generated and hand made test suites are used together, complementing each other.
+
+The approach proved itself effective, at the point that, when the board has been finally released and delivered, only one small defect has been found
+in the driver.
+The errors and the not agreed changes introduced in the board and in the protocol by the hardware supplier have been discovered immediately, with no additional effort, while the simulators were able to feed with correct data the rest of the system.
+
+\subsection{Data stream and protocol description}
+\label{sec:data_stream_and_protocol_description}
+
+The \SB protocol is asynchronous and packet based, the packets have a fixed length.  
+The protocol is separated in multiple layers to ease its formal description, since mixing up different abstraction levels can lead to overly complex and unmanageable descriptions, especially when dealing with formal ones.
+The layers identified in protocol are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable
+  values and reaction constraints on output packets, based on previously transmitted input packets.
+
+We focus our attention to the first three layers, considering only the output sensor data packet.
+The packet byte and info structures of the output sensor data packet reflect the board capabilities and built in sensors. 
+Essentially a single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metada describing the sensors and the board state.
+
+The internal structure of the package is strictly fixed, even if the data streams have different data rates. 
+This choice originates from a trade-off between two constraints: the efficiency of the transmission (redundant data to be minimized) and the complexity to build a package (complexity to be minimized, since the FPGA capabilities on the \STSB are limited).
+A packet is internally composed by frames. 
+The frames (82 frames in each packet) accommodate data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
+
+Fast, medium and slow data rates have not a common multiple; a packet, to be capable to accommodate three different rates outputs in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
+
+\begin{itemize}
+\item Slow data rates: slow data can be found at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
+\item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame.
+The medium data rate sample channel is a piece of info to be extracted from the frame description (frame metadata), including the possibility to have no data at all.
+E.g. a frame contain a medium data rate sample from channel 0, the following frame a sample from channel 5, the following frame no sample at all, etc.
+\end{itemize}
+
+The single packet rules delimit the boundaries of the values obtainable from the packet: each defined sensor represented in the packet, as well as the metadata describing the \STSB and the packet and frame contents, are constrained by a range of acceptable values.
+There are also rules affecting more than one value (i.e.
+if a specific sensor is active the reading value has to be in a defined range), and rules specifying a correct sequence of frames (i.e.
+there can be only one switch in sampling rate in the frame sequence of a single packet).
+
+\subsection{\ldots}
+
+\todo{adapt old paper}
+
+% =====================================================================
+\section{Related works}
+\label{sec:related_works}
+
+\todo{Out of scope: agile & formal used together, but not mixed}
+
+\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
+
+\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
+
+\todo{Mixin with test substitution}
+
+% =====================================================================
+\section{Conclusions}
+\label{sec:conclusions}
+
+% old paper
+\chapter{OLD, to be reused}
 
@@
 \section{Related works}
-\label{sec:related_works}
 
@@
 
-% =====================================================================
-\section{Conclusion}
-\label{sec:conclusion}
-
-\todo{specification and hand tests useful to drive, specification and
-  hand tests and simulator useful to test, specification and generated
-  tests useful for statement coverage and easy methods, specification
-  useful to test hardware, hand useful to spec, spec useful to
-  hand... test driven spec}
-
 \section{Next papers}
-
-\todo{remove and push this on a new paper}
 

### 2010-06-29 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
JML subsection written.

@@
 
-\todo{quick introduction}
-
-\todo{examine language, really short introduction}
-
-\todo{examine tools, included, JML2, and escluded, ESC/2, ones}
+The Java Modeling Language (JML) is the de facto standard specification language for describing the behavioral properties of Java programs.
+JML is a rich behavioral interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioral subtyping~\cite{X}.
+JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants.
+It also includes a rich model-based specification layer that includes mathematical models, data refinements~\cite{X}, datagroups~\cite{X}, and many other advanced object-oriented specification features~\cite{advancedJMLpaper}.
+Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{STTTpaper}.
+
+JML is sometimes used in a Design-by-Contract style, where a specification is written from scratch, reusing existing APIs that have specifications of their own, and then an implementation is written conforming to that specification~\cite{DBCjml,DBC}.
+The implementation is checked by whatever means appropriate using some subset of the aforementioned tools.
+At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process we call Contracting-the-Design.
+
+The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{RAC,ESC}.
+The former compiles executable assertions into runtime checks, thereby generating inline test code executed at runtime.
+Using RAC is familiar for developers as unit testing is en vogue at the moment.
+The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
+As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{soundness,completeness}.
 

### 2010-06-30 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
problems with citations

@@
 % problems would benefit form agile and formal mix, characterize the problem
-The two worlds seem irreconcilable, nevertheless, there exist problems that would greatly benefit from both.
-The problem needs to be at least instable in the requirements definition, constrained by deadlines that cannot be postponed, but also characterized by artifacts that needs to be formally specified and verified.
-The complex problem addressed in the case study fits these needs.
-
-\todo{describe, ``Formal Versus Agile: Survival Of the Fittes'', blending advantages and potentials}
+The two worlds seem irreconcilable, nevertheless, there exist problems that would greatly benefit from both~\cite{Black2009}.
+The problems' class we are interested in need to have unstable requirements, to be constrained by deadlines that cannot be postponed, but also to be characterized by artifacts that need to be formally specified and verified.
+
+\subsection{Formal and agile integration}
+\label{subsec:formal_and_agile_integration}
 
@@
 % what needs to be considered: TDD
-The most common software verification practices in agile methodologies need to be considered as well when tring to integrate formal and agile methodologies, since they often are a fundamental part of the development process.
-The most significant verification practice is Test Driven Development (TDD), TDD is a software development technique, originally defined in Extreme Programming (XP) methodology, characterized by writing unit tests before actually writing the code.
+The most common software verification practices in agile methodologies need to be considered, when tring to integrate formal and agile methodologies, since they are a fundamental part of the development process.
+The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)]\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
 Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven apprach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is consider a good software development technique to be used on its own, regardless of the surrounfing development process.
+The test driven apprach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is considered a good software development technique to be used on its own, regardless of the surrounfing development process.
 It is to be noted that, depending on the abstraction level, the test driven approach can introduce activities cycles that can be as long as a complete iteration (functional tests), or as short as a couple of minutes (small increment on a unit test).
 
-% previous very related work
+% previous related work
 Some tentative attempts to reconcile formal and agile methodologies have been developed.
-The recurrent approach of the attempts consists in removing the hand made test suites, substituing them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
-
-\todo{describe and reference ``An Agile Formal Development Methodology'', perfect match}
-
-\todo{describe and reference ``Formal Extreme (and Extremely Formal) Programming'', perfect match}
-
-\todo{describe and reference ``XFM: An Incremental Methodology for Developing Formal Models'', perfect match}
+The recurrent approach of the attempts consists in removing the hand made test suites, substituing them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
+
+% a quick analysis of the related works
+Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an iterative and incremental process to generate X-Machines models, an extension of Finite State Machines. 
+The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines formal methodology. 
+The development process proposed in~\cite{Herranz2003b} by Herranz and Moreno-Navarro is quite similar; they propose an iterative process to model a system using SLAM-SL, an object-oriented specification language, and its tools.
+Some XP practices are fully adopted or considered compatible, as pair-programming, iterative and incremental development, system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
+A different formal methodology, but same approach, can be found also in~\cite{Suhaib2005} where XFM methodology is explained.
 
 % previous work: a common approach
-Unfortunately, the agile and formal reconciliation attempts that have been defined so far are not entirely satisfing.
 As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract precondition, postconditions and invariants) \todo{reference Meyer} and test suites automatic generation \todo{references, JML, Eiffel, \ldots}.
-At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development \todo{references, ESC/2, \ldots}.
+The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} precondition, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
+At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~/cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
 % previous work: why is not satisfying
-The methodologies are sound, but they share some common problems: the loops between development artifacts and related activities (code, tests, design, refactoring), especially the closed and tight ones, typical of TDD, are completely lost.
-This is not a minor issue because it is believed that these loops are the inner engine of many agile methodologies, imposing to the development process an iterative and incremental pace \todo{find meaningful references, I know there are some}.
-Beside possibly disrupting some of the agile methodologies inner mechanincs, only the verification purpose of the tests is mantained, that is, the documentation and requirement specification purposes are lost and need to be moved on top of different artifacts, usually the formal specification.
-Finally a complete system formal specification is needed to apply these methodologies, but this requirement is hard to fulfill since the effort required to write a complete formal specification in most cases is much greater than writing a suite of tests.
+The approach is simple: the verification, documentation and requirement specification purposes of the tests are moved on to the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
+The problem is that this requirement is hard to fulfill, since the effort required to write a complete formal specification of the system, in most real world complex cases, is much greater than writing a suite of tests.
+If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified.
+The parts not formally specified can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are not specified or completely lost.
+The loops and relations between the different development artifacts are the inner engine of many agile methodologies, they impose an iterative and incremental pace to the development process.
 
 % what we want to achieve
-Thus, the objective set is different: we want to maintain an higly iterative and incremental development process, we want to connect development artifacts and activities (code, requirements, tests, formal specifications, refactorings, run-time verification, automatically generated tests, \ldots) through activities cycles similar to what we can find in agile methodologies (especially in XP), we want to be able not to formally specify the complete system in order to cope with constrained resources and
+The objectives we want to achive are the following ones: to be able to formally specify only parts os the system, in order to cope with constrained resources and unstable requirements; to draft an higly iterative and incremental development process that blends formal and agile practices.
+
+% how we want to achive it
+The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem.
+We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
+That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices.
+
+Mole practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and verified.
+The case study fits these needs.
+
+\subsection{A Real And Complex Case Study: Rapid Development In Small Scale Hardware Software Co-design}
+\label{subsec:a_real_and_complex_case_study}
+
+% SenseTile introduction
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system installed at the University College Dublin in Dublin, Ireland.
+Our facility provide a scalable and extensible infrastructure for performing in-depth investigation of both the specific and general issues in large-scale sensor networking.
+This system integrates a sensor platform, a datastore, and a compute farm.  
+The sensor platform is a custom-designed, but inexpensive, sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.
+There are over a dozen sensors packaged on the \ST itself, the board is expandable as well, new sensors can be added to it easily.
+
+% case study: the sensor board development tasks
+The case study is focused in building the sensor board and its software driver side by side.
+Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board; (4) board embedded software; (5) communication protocol software driver; (6) development of software simulators of the board.
+
+% what we have to do, what we outsourced, and the dependencies between tasks
+The development of the physical board, along with its embedded software (development tasks 2, 3, 4), is carried out by an external manufacturer, thus it is not directly taken into account here. 
+The specification of the communication protocol (development task 1) is a joint effort, while the remainang tasks (5, 6) can be developed in isolation.
+The dependencies are straighforward: the specification of the communication protocol (1) is the most stable element, but it still depends on the sensor board (2, 3, 4): big changes in the latter could affect the former.
+The driver and the simulators depends directly on the communication protocol.
+
+% how the problem is usually managed
+There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software
+co-design), but they are all focused on big-scale systems.
+The hardware software formal co-design methodologies usually have several common development artifacts\cite{Slomka2000,Hoffman2001}: an high level specification of the system, a translation (refinement) of the high level specification to low level ones (hardware and software counterparts), the possibility to generate software code from the low level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
+When considering the development of the device and of its software driver as separate entities, that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
+
+% complete specification is not feasible
+A specification as complete as possible is not feasible nor convenient in this case: the protocol specification have to be enhanced incrementally, and its complete specification cannot be provided.
+This makes the problem an ideal candidate to test the Mole practices.
+
+% formal methodology constraints
+An appriate formal methodology needs to be chosen; it must be able to support a specification that is built incrementally and accompanied by tools that allow the verification of the system properties at run time.
+All the formal methods that require a complete specification to be used effectivelt must be avoided.
+The formal language should also be able to automatically generate tests.
+
+% the chosen formal methodology
+Being Java the language chosen for the implementation, the Java Modeling Language (JML)~\cite{advancedJMLpaper}, accompanied by the JML Runtime Assertion Checker suite of tools~\cite{RAC}\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}} has been selected.
+JML and RAC tools provide a run time checker, a test generator, and do not require a complete specification to be used effectively.
+
+\subsection{\todo{paper structure title}}
+
+\todo{explain paper structure}
 
@@
 
-\todo{quick introduction}
-
-\todo{define what we are looking at: verification, cycles and dependencies between artifacts}
-
-\todo{the cycles in SCRUM and XP}
-
-\todo{TDD}
+Agile methodologies are all based on a highly iterative and incremental process.
+They share a common approach on team management, customer relation, simplification and removal of not needed artifacts and activities and importance of working solutions and customers satisfaction.
+The agile manifesto~\cite{Beck2001} summarizes the philosphy and principles shared by all agile methodologies.
+
+In the present context, focused on how to merge agile and formal practices, we are mainly interested in the verification practices used in agile methodologies.
+The agile manifesto does not introduce any specific techniques to be used in the development process, it describes agile methodologies from a very high abstraction level.
+Nevertheless most of the agile methodologies share a similar approach to artifacts verification.
+Functional test suites, to be used as precise requirements or user stories definition, are used in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}, while unit tests and TDD are mentioned, whether not completely accepted, in most of the agile methodologies defined so far.
+
+High abstraction level test suites, functional tests, describe requirements, and actually bind together requirements documentation and system behaviour, the development cycle is one iteration long. 
+Low abstraction level test suites, unit tests, describe a module behaviour, and actually bind together code documentation, code and the contract imposed on the module, the development cycle is usually less than one day.
+
+Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.
+Continuous refactoring~\cite{Fowler1999} enables an interative and incremental development, since the system architecture and design, whether explicitly specified or not, needs to be cleaned up, after implementation, and modified, to accomodate new requirements, services and modules.
+The automated test suites guarantees that the functionalities and behaviour already implemented in the code are not compromised by the system refactorings.
 
@@
 
-The Java Modeling Language (JML) is the de facto standard specification language for describing the behavioral properties of Java programs.
+The Java Modeling Language (JML)\footnote{\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}} is the de facto standard specification language for describing the behavioral properties of Java programs.
 JML is a rich behavioral interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioral subtyping~\cite{X}.
@@
 The implementation is checked by whatever means appropriate using some subset of the aforementioned tools.  
-At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process we call Contracting-the-Design.
-
-The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{RAC,ESC}.
-The former compiles executable assertions into runtime checks, thereby generating inline test code executed at runtime.
-Using RAC is familiar for developers as unit testing is en vogue at the moment.
+At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process called Contracting-the-Design.
+
+The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{RAC,BurdyEtal05-STTT,ESC}.
+The former compiles executable assertions into runtime checks.  
 The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
@@
 % =====================================================================
-\section{A Real And Complex Case Study: Rapid Development In Small Scale Hardware Software Co-design}
-\label{sec:a_real_and_complex_case_study}
-
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system to be installed at the University College Dublin in Dublin, Ireland.  
-Our facility aims to provide a scalable and extensible infrastructure for performing in-depth investigation of both the specific and general issues in large-scale sensor networking.
-This system integrates a sensor platform, a datastore, and a compute farm.  
-The sensor platform is a custom-designed, but inexpensive, sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.
-There are over a dozen sensors packaged on the \ST itself, the board is expandable as well, new sensors can be added to it easily.
-
-The case study is focused in building the sensor board and its software driver side by side.
-Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board; (4) board embedded software; (5) communication protocol software driver; (6) development of software simulators of the board.
-
-The development of the physical board, along with its embedded software (development tasks 2, 3, 4), is done by an external manufacturer, thus it is not directly taken into account here.
-The specification of the communication protocol (development task 1) is a joint effort, while the remainang tasks (5, 6) can be developed in isolation.
-
-\todo{illustrate dependencies: 3<>2<>4<>1, 1>5, 1>6}
-
-\todo{explain 3<>2<>4<>1: difficulties encountered}
-
-\todo{explain 1>5, 1>6: clean separation, but 1 is unstable}
-
-There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software
-co-design), but they are all focused on big-scale systems.
-The hardware software formal co-design methodologies usually have several common development artifacts\cite{Slomka2000,Hoffman2001}: an high level specification of the system, a translation (refinement) of the high level specification to low level ones (hardware and software counterparts), the possibility to generate software code from the low level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
-
-When considering the development of the device and of its software driver as separate entities, that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
-
-As it has been already pointed out, a specification as complete as possible is not feasible nor convenient: we want to enhance the protocol specification incrementally, and we are not able to provide a complete specification.
-
-The protocol specification is divided in various (thin) layers, and each of the layer has been specified with a different lightweight approach: some of the layers have been specified formally, while others have been specified in a less formal way.
-The formal specifications are used to check at run time the correctness of the \SB, the driver and the simulators.  
-The formal specifications are used to automatically generate a test suite to verify the driver and the simulators, and eventually to verify the
-specification itself.  
-Hand made test suites are used to verify the driver and the simulators as well.  
-The generated and hand made test suites are used together, complementing each other.
-
-The approach proved itself effective, at the point that, when the board has been finally released and delivered, only one small defect has been found
-in the driver.
-The errors and the not agreed changes introduced in the board and in the protocol by the hardware supplier have been discovered immediately, with no additional effort, while the simulators were able to feed with correct data the rest of the system.
+\section{\todo{the case study}}
+\label{sec:___}
 
 \subsection{Data stream and protocol description}
-\label{sec:data_stream_and_protocol_description}
-
+\label{subsec:data_stream_and_protocol_description}
+
+% protocol layers
 The \SB protocol is asynchronous and packet based, the packets have a fixed length.  
-The protocol is separated in multiple layers to ease its formal description, since mixing up different abstraction levels can lead to overly complex and unmanageable descriptions, especially when dealing with formal ones.
-The layers identified in protocol are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable
-  values and reaction constraints on output packets, based on previously transmitted input packets.
-
-We focus our attention to the first three layers, considering only the output sensor data packet.
+The protocol is separated in multiple layers to ease its implementation in the driver, since mixing up different abstraction levels can lead to overly complex and unmanageable code.  
+The layers identified in protocol are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable values and reaction constraints on output packets, based on previously transmitted input packets.
+For space reason, the first three layers are analyzed in the case study, considering only the output sensor data packet.
+
+% packet structure
 The packet byte and info structures of the output sensor data packet reflect the board capabilities and built in sensors. 
-Essentially a single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metada describing the sensors and the board state.
-
-The internal structure of the package is strictly fixed, even if the data streams have different data rates. 
-This choice originates from a trade-off between two constraints: the efficiency of the transmission (redundant data to be minimized) and the complexity to build a package (complexity to be minimized, since the FPGA capabilities on the \STSB are limited).
+A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metada describing the sensors and the board state.
+The internal structure of the packet is strictly fixed, even if the data streams have different data rates. 
+This choice originates from a trade-off between two constraints: the efficiency of the transmission (redundant data to be minimized) and the complexity to build a packet (complexity to be minimized, since the processing capabilities in the \STSB are limited).
+
+% packet internal structure: collection of frames
 A packet is internally composed by frames. 
-The frames (82 frames in each packet) accommodate data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
-
-Fast, medium and slow data rates have not a common multiple; a packet, to be capable to accommodate three different rates outputs in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
+A frame (82 frames in each packet) accommodate data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
+Fast, medium and slow data rates have not a common multiple; a packet, to be capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
 
@@
 
-\subsection{\ldots}
-
-\todo{adapt old paper}
-
-% =====================================================================
-\section{Related works}
-\label{sec:related_works}
-
-\todo{Out of scope: agile & formal used together, but not mixed}
-
-\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
-
-\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
-
-\todo{Mixin with test substitution}
-
-% =====================================================================
-\section{Conclusions}
-\label{sec:conclusions}
-
-% old paper
-\chapter{OLD, to be reused}
-
-% =====================================================================
-\section{Data stream and protocol description}
-\label{sec:data_stream_and_protocol_description}
-
-\STSB and \STPU communicates via an Universal Serial Bus (USB)
-interface.  A proprietary protocol is used on top of the USB serial
-channels, the protocol is specified by
-FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
-is the manufacturer of the USB controller chip used in the \STSB, the
-FT2232H\cite{ftdi_ft232h_2009}.
-
-The proprietary protocol enables the control and management of some of
-the functionalities in FT232H chip (such as internal EEPROM read and
-write, reset of the chip, etc.), as well as the definition of the
-communication protocol and properties to be used in the asynchronous
-channels (input and output channels) provided by FT232H.  These
-functionalities are accessed via an Application Program Interface, the
-API is provided by a driver developed directly by FTDI, which is
-available for various systems (Linux 32 and 64 bits, Mac OS, Windows)
-\cite{ftdi_d2xx_api_2009}.
-
-Finally, a custom protocol has been developed to manage the entire
-\STSB, built on top of the serial channels provided by the API.  The
-custom protocol is meant to be used to read sensor data observations,
-enable or disable sensors, configure on-board components (as the
-Field-Programmable Gate Array and the micro processor); the protocol
-main function is to read sensor data.
-
-The protocol is asynchronous and packet based, the packets have a
-fixed length.  The protocol has been divided in multiple layers to
-ease its formal description.  The rationale is the following: mixing
-up different abstraction levels of a protocol description can lead to
-overly complex and unmanageable descriptions, especially when dealing
-with formal ones.  The layers identified in protocol are the
-following:
-
-\begin{description}
-\item[Packet byte structure]: the internal representation of the
-  packet.
-\item[Packet info structure]: the meaningful fields contained in the
-  packets.
-\item[Single packet rules]: the content acceptable values, and how
-  they influence each other.
-\item[Packet sequence rules]: the content acceptable values, based on
-  values of previously received packets.
-\item[Packet input output asynchronous rules]: the content acceptable
-  values and reaction constraints on output packets, based on
-  previously transmitted input packets.
-\end{description}
-
-We will examine the output sensor data packet, since it is the most
-used and interesting one; we won't describe input or other output
-packets.  We will examine the first three layers, that is: the packet
-byte structure, the packet info structure and the single packet rules.
-
-The packet byte and info structures of the output sensor data packet
-reflect the board capabilities and built in sensors.  Essentially a
-single packet has to accommodate various types of data:
-
-\begin{itemize}
-\item Fast data rate streams: audio data, 48KHz (standard audio) or
-  96KHz (ultra audio).
-\item Medium data rate streams: 5KHz.
-\item Slow data rate streams: temperature, pression, accelerometer,
-  etc.; all these sensor have a data rate lower than 5 KHz.
-\item Metadatas: metadatas on sensors and board state.
-\end{itemize}
-
-The internal structure of the package (as we will see) is strictly
-fixed, even if the data streams have different data rates.  This
-choice originates from a trade-off between two constraints: the
-efficiency of the transmission (redundant data to be minimized) and
-the complexity to build a package (complexity to be minimized).  The
-complexity minimization constraint rationale is the following: the
-FPGA on the \STSB builds the packet, but the FPGA is also to be used
-for various other specific tasks; it is important to maintain the
-building packet code on the FPGA as simple and short as possible, to
-leave enough space to implement other tasks.  The Output sensor data
-packet structure is shown in Table \ref{tab:data_packet_structure},
-the packet total length is 1024 bytes (512 16 bit words).
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet structure.}
-  \label{tab:data_packet_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0-1 & sentinel\\\hline
-      2-7 & slow data\\\hline
-      8-11 & metadata: board state\\\hline
-      12-13 & reserved\\\hline
-      14-15 & metadata: board state\\\hline
-      16 & reserved\\\hline
-      17-22 & frame 1\\\hline
-      \ldots & \ldots\\\hline
-      503-508 & frame 82\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-The frames accommodate data from the fast and medium data rate streams
-(there are 4 fast data rate stream and 8 medium data rate stream
-channels).  There are 82 frames in a packet.  The frame structure is
-shown in Table \ref{tab:frame_structure}.
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet frame structure.}
-  \label{tab:frame_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0 & metadata: frame description\\\hline
-      1 & medium data rate data sample\\\hline
-      2-5 & fast data rate data samples\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-Looking at the structure described and the data rates frequencies, it
-can be seen that there must be some subtle rule, to make all the
-figures to add up; that is, to be capable to accommodate three
-different rates outputs in a fixed structure, fixed length packet.
-The subtleties lies in two forms of redundancies, for slow and medium
-data rates:
-
-\begin{itemize}
-\item Slow data rates: slow data can be found at the beginning of each
-  packet, the data is simply repeated (that is, it is equal to the
-  previous packet in the stream) when no new data is available.
-\item Medium data rates: the 8 medium rate channels are multiplexed
-  together into the medium data rate data sample of each frame.  The
-  medium data rate sample channel is a piece of info to be extracted
-  from the frame description. Besides, the medium data rate data
-  sample could contain no data at all, this information is to be
-  extracted from the frame description. So, one frame could contain a
-  medium data rate sample from channel 0, the following frame a sample
-  from channel 5, the following frame no sample at all, etc.
-\end{itemize}
-
-The single packet rules delimit the boundaries of the values
-obtainable from the packet: each defined sensor represented in the
-packet, as well as the metadata describing the \STSB and the packet
-and frame contents, are constrained by a range of acceptable values.
-There are also rules affecting more than one value, these are
-essentially validity rules (i.e. if a specific sensor is active the
-reading value has to be in a defined range), and rules specifying a
-correct sequence of frames (i.e. there can be only one switch in
-sampling rate in the frame sequence of a single packet).
-
-% =====================================================================
-\section{Data stream and protocol specification}
-\label{sec:data_stream_and_protocol_specification}
-
-The specifications of the protocol has been used as a central point to
-lead all the other development tasks.
-
-Regarding the formal specification used, we needed a specification
-capable of:
-
-\begin{itemize}
-\item Be executed at run time: if something on the protocol is not
-  respected, the run time execution environment is capable to detect
-  and signal the fault.
-\item Generate tests.
-\item Verify simulators.
-\item Be used in conjunction of hand written unit tests.
-\end{itemize}
-
-We needed specifications capable to tackle properly all the protocol
-layers.  Since different layers have different characteristics,
-different specification approaches has been chosen for each layer: the
-specification are context dependent.
-
-\paragraph{Packet byte structure specification}
-
-The specification is provided by a hand made test suite, composed of
-unit and integration tests; the tests specifies the behaviors of the
-Java implementation, which is capable of recognizing the proper packet
-structure in a binary data stream. The implementation is tricky: it
-checks for the sentinels repeatedly to be sufficiently confidant that
-the byte stream can be properly decomposed into a packet stream.
-
-Fulfilled requirements:
-
-\begin{itemize}
-\item The specification cannot be checked at run time.
-\item Meaningful automatic tests cannot be generated.
-\item The specification can be used to verify a packet byte structure
-  simulator.
-\item The implementation were actually driven (Test Driven
-  Development) by the test suite specification.
-\end{itemize}
-
-\paragraph{Packet info structure specification}
-
-The specification is provided by JML annotations\cite{JML2}\footnote{
-  \myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling
-    Language (JML)}} and by a hand made test suite.
-
-JML has been chosen because it is capable to specify Java code: JML is
-a behavioral interface specification language.  JML specification
-language combines the design by contract approach (applied to Java
-language) and the model-based specification approach of interface
-specification languages.
-
-Many tools able to support JML specification language are available,
-focusing on different aspects of the language.  JML2 suite of
-tools\cite{BurdyEtal05-STTT}
-\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common
-    JML tools}} has been chosen, since the suite provides both run
-time verification support and test generation support.
-
-Fulfilled requirements:
-
-\begin{itemize}
-\item The Java API specification part is checked automatically by the
-  Java compiler, while the JML specification part can be checked at
-  run time using the provided JML2 run time environment.
-\item Meaningful automatic tests can be generated from the JML
-  specification.
-\item The JML specification can be used to verify a simulator, which
-  generates packets as the real board would do .The specification can
-  also be used to verify the real driver implementation, based on the
-  packet byte structure implementation presented above.
-\item The implementation was driven by the formal JML specification
-  and by the hand made test suite.  Finally, hand made unit tests can
-  be checked, during execution, against the formal specification,
-  using the provided JML2 run time environment.
-\end{itemize}
-
-\paragraph{Single packet rules}
-
-The specification is provided exclusively by the formal specification
-with JML language.
-
-The fulfilled requirements of a JML specification have been already
-described in the previous paragraph, that is:
-
-\begin{itemize}
-\item The JML specification is checkable at run time.
-\item Tests can be automatically generated.
-\item Simulators can be verified.
-\item The specification can be used with hand written tests.
-\end{itemize}
+\subsection{Data stream and protocol specification}
+\label{subsec:data_stream_and_protocol_specification}
+
+% protocol specification
+The specifications of the protocol are distilled and refined incrementally.
+The protocol is divided in various (thin) layers, and each of the layer has been verified with a different approach: some of the layers have been specified formally.
+
+\paragraph*{Packet byte structure specification}
+
+The unformal specification and verification is provided by a hand made test suite, composed of unit and integration tests; the tests specifies the behaviors of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
+The implementation is tricky: it checks for the sentinels repeatedly to be sufficiently confidant that the byte stream can be properly decomposed into a packet stream.
+
+The specification can be used to verify a board simulator, but it cannot be used to check properties at run time.
+Code and tests have been implemented using TDD.
+
+\paragraph*{Packet info structure specification}
+
+The specification is provided by hand made test suites and JML annatations. 
+It has been built starting from hand made tests only, and incrementally adding tests and later on, JML annotations.
+
+\todo{explain: test -> code, test + code -> formal specification, formal specification -> generated tests}
+
+\paragraph*{Single packet rules}
+
+The specification is provided exclusively by the formal specification with JML language.
+It has been built starting from JML annoatations, supported by hand made tests when needed.
+
+\todo{formal specification -> test + generated tests + code}
 
@@
 
-Simulators can be used to test parts of the system, to minimizing
-their dependencies: a simulator can be used by upper layers, with no
-need to provide the functionalities of the lower layers.  During the
-development of the \STSB the simulators have been essential, since the
-physical board was not available: we needed at least to remove the
-dependencies over the real board to execute the driver implementation.
-Simulators have been built concurrently with the specification, the
-driver implementation, and the tests.
-
-Theoretically a separate simulator should be needed for each protocol
-layer: this would enable us to specify, develop and test each layer in
-isolation.  In our case we evaluated not worth the effort to replicate
-the layer structure into simulators; the choice was motivated by the
-fact that such a distinction would have not provided us any
-perceivable benefit.  Two simulators were built.
-
-The simulators have been built reflecting the software abstraction
-layers defined for the driver.  The driver API is splitted in two
-abstraction layers: the general high level interfaces, that expose the
-main functionalities of the board and the main contents of the
-packets, and the lower level implementation that parse the data
-streams in and out the board, meant to translate the higher level
-instructions and data in properly formed packets.
-
-The high level simulator implements the interfaces providing the
-methods to deal with an abstracted \SB: the simulator knows nothing
-about the details of the real format of the binary data streams.  The
-low level simulator is capable to rebuild the \SB data streams: the in
-and out data streams are built exactly as the sensor board is expected
-to be able to parse or generate.
+% why simulators are useful
+Simulators are used to test parts of the system minimizing dependencies: a simulator can be used by upper layers, with no need to provide the functionalities of the lower layers. 
+The simulators proved essential during development, since the physical board was not available.
+
+% how many of them do we need?
+Theoretically a separate simulator is needed for each protocol layer, to specify, develop and test each layer in isolation.  
+But, because of limited resources and affort avilable, the layer structure was not matched by corresponding simulators.
+Two simulators are available.
+
+% simulators development
+Simulators have been built concurrently with the specification, the driver implementation, and the tests.
+
+% simulators description
+The driver API is splitted in two abstraction layers: the general high level interfaces, that expose the main functionalities of the board and the main contents of the packets, and the lower level implementation that parse the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
+The simulators are built according to these two abstraction layers.
+The high level simulator implements the interfaces providing the methods to deal with an abstracted \SB: the simulator knows nothing
+about the details of the real format of the binary data streams.
+The low level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to be able to parse or generate.
 
@@
 
-The JML specifications that have been written are of varying
-complexities.  Some of the specifications are simple, focusing on
-constraints that should hold when calling a method (the preconditions)
-and constraints on the return value (the very basic form of
-postconditions).  In listing \ref{lst:jml_example_simple} a simple JML
-specification example is shown, the method \lil{getTemperature} is
-declared to be \lil{/*@ pure @*/}, which means that it cannot change
-the state of any object (a postcondition), the specification also
-constraints the return value with a lower and upper bound (another
-postcondition).
+% a simple specification example
+The JML specifications are of varying complexities.  
+Some of them are rather simple, focusing on constraints that should hold when calling a method (the preconditions) and constraints on the return value (the very basic form of postconditions). 
+In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition), the specification also constraints the return value with a lower and upper bound (another postcondition).
 
@@
   float=htbp]
-  /*@ ensures \result >= -880; ensures \result <= 2047; @*/ /*@ pure
-  @*/ short getTemperature();
+  /*@ 
+      ensures \result >= -880; 
+      ensures \result <= 2047; 
+  @*/ 
+  /*@ pure @*/ short getTemperature();
 \end{lstlisting}
@@
 
-Some of the specifications are rather complex, usually focusing on
-properties regarding the behavior of a whole object.  In listing
-\ref{lst:jml_example_invariant} a complex invariant example is shown;
-an invariant is a property that has to be maintained during the life
-cycle of an object, more precisly, an invariant is assumed on entry
-and guaranteed on exit of each method of an object.  The invariant is
-constraining the number of samples for each medium data rate streams:
-the total number of streams is \lil{Frame.ADC_CHANNELS}, the total
-number of frames is \lil{FRAMES}, the constant constraining the number
-of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the
-samples contained in a frame are fairly distributed on the channels.
-The valid samples are counted parsing all the frames contained in a
-packet, selecting only the matching valid samples.  A medium data rate
-stream sample is considered valid when \lil{isADCActive()} method
-returns \lil{true}.
+% a complex specification example
+The complex specifications usually focus on properties regarding the behavior of a whole object.
+In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property that has to be maintained during the life cycle of an object, more precisly, an invariant is assumed on entry and guaranteed on exit of each method of an object.
+The invariant is constraining the number of samples for each medium data rate streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is \lil{FRAMES}, the constant constraining the number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are fairly distributed on the channels.
+The valid samples are counted parsing all the frames contained in a packet, selecting only the matching valid samples.  
+A medium data rate stream sample is considered valid when \lil{isADCActive()} method returns \lil{true}.
 
@@
   float=htbp]
-  /*@ invariant ( \forall int channel; 0 <= channel && channel <
-  Frame.ADC_CHANNELS; ( \num_of int i; 0<= i && i<(FRAMES-1); (
-  getFrame(i).isADCActive() && (getFrame(i).getADCChannel() ==
-  channel) ) ) <= (FRAMES / Frame.ADC_CHANNELS + 1) ); @*/
+  /*@ 
+      invariant ( 
+        \forall int channel; 0 <= channel && channel < Frame.ADC_CHANNELS; ( 
+          \num_of int i; 0<= i && i<(FRAMES-1); (
+            (getFrame(i).isADCActive()) && 
+            (getFrame(i).getADCChannel() == channel)
+          ) 
+        ) <= (FRAMES / Frame.ADC_CHANNELS + 1) 
+      ); 
+  @*/
 \end{lstlisting}
 
-% =====================================================================
-\section{Test cases}
-\label{sec:test_cases}
-
-The unit test cases that have been built to verify the protocol driver
-are of two kinds: hand made unit tests and automatically generated
-unit tests based on JML specifications.  The tests have been thought
-to be used together, they are complementary.  The tests effectiveness
-has been evaluated for each test suite; the evaluation is carried out
-in section \ref{sec:test_cases_retrospectives}.  The test
-effectiveness evaluation has been based on three elements: effective
-results on piloting the real board (quantitative), code coverage
+\subsection{Test cases}
+\label{subsec:test_cases}
+
+% unit tests kinds
+The unit test cases that verify the protocol driver are of two kinds: hand made unit tests and automatically generated unit tests based on JML specifications; they are built to be used together, they are complementary.
+
+% unit tests effectiveness
+The tests effectiveness is evaluated for each test suite; the evaluation is carried out in section \ref{subsec:test_cases_retrospectives}.
+The test effectiveness evaluation considers three elements: effective results on piloting the real board (quantitative), code coverage
 (quantitative), development help and usefulness (qualitative).
 
-The dual testing approach was thought a winning option, when compared
-to hand based tests or completely automated tests used alone.  \ST
-stream communication protocol is characterized by a well definite
-packet structure, and a well definite semantic on the packet contents
-as well.  On the other hand, in the protocol, there are not strong
-constraints on synchronization, since the communication channel is
-splitted in two (in and out) separate concurrent and asynchronous
-channels.  The lack of hard synchronization constraints reduces the
-advantage to use a classical state machine based specification (as the
-one presented in \cite{Hubbers2004}), hence this option has been ruled
-out.
-
-Hard development time constraints were also a precise requirement; the
-time constraints pushed us towards an incomplete (\emph{just good
-  enough}) specification.  In \STSB driver JML was used to specify
-very simple behaviors (method oriented) on the whole driver
-implementation, and more complex behavior on the driver core
-interfaces.  The software system parts not specified by the JML2
-specification were specified using hand made unit tests.
-
-Hand made unit tests specifies the behavior not specified by formal
-specification; nevertheless this is not the only purpose they have.
-Hand made unit tests have been used to drive the development, since
-the Test Driven Development approach\cite{beck2003test} has been
-followed during development.
-
 \paragraph*{Hand made tests}
 
-The package structure of the driver developed is shown in Figure
-\ref{fig:class_diagram_main}: two independent packages has been
-defined (\lil{Stream} and \lil{Driver}).  The packages are abstract,
-that is, they mainly contain abstractions; in Java language this is
-translated into a package which contains mainly abstract classes or
-interfaces.  The two abstract packages have various implementations,
-and the dependency between the two is not direct, since it is realized
-through an implementation (\lil{StreamDriver}).  This is needed to
-maintain a high decoupling of the packages, and is the result of
-applying the classical dependency inversion
-principle\cite{martin1996dependency}: the dependency structure of the
-packages has a saw shape, a common shape on systems where the
-dependency inversion principle has been followed.
+% the driver package structure
+The package structure of the driver is shown in Figure \ref{fig:class_diagram_main}: two independent packages are defined (\lil{Stream} and \lil{Driver}).  
+The packages are abstract, that is, they mainly contain abstractions; in Java language this is translated into a package which contains mainly abstract classes or interfaces.  
+The dependency between the two abstract packages is not direct, it is realized through an implementation (\lil{StreamDriver}).
+This is needed to maintain a high decoupling of the packages, and is the result of applying the dependency inversion principle~\cite{Martin1996}.
 
@@
 
-The test package structure, which is shown in Figure
-\ref{fig:class_diagram_test} reflects and mimics the code structure;
-the tests for an abstract package are abstract, to be implemented by
-the package that tests a corresponding system implementation; this is
-a well known test pattern (the Abstract Test
-Pattern\cite{thomas2004java}) used to test that the contracts are
-respected in the implementations.  For instance, package \lil{DriverT}
-contains abstract tests for the abstractions of package \lil{Driver},
-package \lil{StreamDriverT} inherits the abstractions contained in
-\lil{DriverT} and makes them concrete, to test the corresponding
-implementation \lil{StreamDriver}; package \lil{StreamDriverT} also
-contains stand alone tests written specifically for the implementation
-contained in \lil{StreamDriver}.
+% the test package structure
+The test package structure, which is shown in Figure \ref{fig:class_diagram_test} reflects and mimics the code structure; the tests for an abstract package are abstract, and implemented by the package that tests a corresponding system implementation.
+This is a well known test pattern (the Abstract Test Pattern~\cite{Thomas2004}) used to test that the contracts defined in the abstractions are respected in the several implementations.  
+For instance, package \lil{DriverT} contains abstract tests for the abstractions of package \lil{Driver}, package \lil{StreamDriverT} inherits the abstractions contained in \lil{DriverT} and makes them concrete, to test the corresponding implementation \lil{StreamDriver}; package \lil{StreamDriverT} also contains stand alone tests written specifically for the implementation contained in \lil{StreamDriver}.
 
@@
 
-The JML specification has been used to generate the tests.  A specific
-tool of the JML2 suite is able to automatically generate
-them\cite{Cheon-Leavens02}.
-
-The test suite generated is a collection of fine grained unit tests.
-For each public method a set of tests is generated.  The specified
-behavior of the method is checked using the JML2 run time assertion
-checker, hence, an oracle is not needed.  The generated tests send
-messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation
-exceptions;
-if a precondition is not met the test is not run, if the
-precondition is met the test is run and if the postcondition is not
-met an exception is fired and the test fails.  The test data
-used to build the tests is the only part that is required to be
-provided by the user.
-
-The resulting test package structure is shown in Figure
-\ref{fig:class_diagram_generatedtest}.  The package dependencies are
-defined by the specifications; for instance, \lil{SimulatorDriverST}
-are generated tests, they test (hence use) the implementation
-\lil{SimulatorDriver}, and have been generated using the
-corresponding specification \lil{SimulatorDriverS}.  The
-generated tests do not maintain the inheritance structure of the code
-and the specification.
+% the generated test package structure
+The JML specifications are used to generate tests.
+The resulting test package structure is shown in Figure \ref{fig:class_diagram_generatedtest}.  
+The package dependencies are defined by the specifications; for instance, \lil{SimulatorDriverST} are generated tests, they test (hence use) the implementation \lil{SimulatorDriver}, and have been generated using the corresponding specification \lil{SimulatorDriverS}.
+The generated tests do not maintain the inheritance structure of the code and the specification.
 
@@
 
-A proper stub, driver and oracle should be identified for each method
-under test to have proper unit test suite; stub, driver and oracle are
-the usual components found in the harness that enable test
-execution\cite{Binder1999}.  In
-JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the
-framework used to manage the implementation and execution of hand made
-and generated unit tests, there is no clear distinction between
-drivers, stubs, and oracles; none-less, even if a proper distinction
-is not imposed by the framework, the conceptual distinction still
-holds.
-
-Hence, to properly unit test a package, we need:
-
-\begin{description}
-\item[Stub or Skeleton]: a piece of code that simulates the activity
-  of the modules that are not under tests.
-\item[Driver]: a piece of code that calls in a proper way the module
-  under test.
-\item[Oracle]: a mechanism used for determining whether a test has
-  passed or failed, comparing the output of the module under test to
-  the output the oracle determines the module should have.
-\end{description}
-
-Hand made unit tests and JML generated tests differ in how these
-components are provided.  The distinctions and similarities are
-summarized in Table \ref{tab:test_harness}.
-
-\begin{table}[htbp]
-  \caption{Hand made and generated tests harness.}
-  \label{tab:test_harness}
-  \begin{center}
-    \begin{tabular}{|l|l|l|}\hline
-      & \textbf{\textit{hand made}} & \textbf{\textit{JML}} \\\hline
-      \textbf{\textit{Stub}} & to be provided & to be provided\\\hline
-      \textbf{\textit{Driver}} & to be provided & automatic: 1 call\\\hline
-      \textbf{\textit{Oracle}} & to be provided & 
-      automatic: runtime assertion checker\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-A module simulator can be used to ease the tasks of developing each
-component of a test harness: it can be used as a stub, to simulate a
-module needed by the module under test, it can be used as a driver, to
-stimulate the module under test, it can be used as an oracle, if it
-simulates the module under test.
-
-In our system, the simulators has been used as stubs, so that we have,
-during testing, a structure like the one shown in Figure
-\ref{fig:class_diagram_streamdriver_test}.  In Figure
-\ref{fig:class_diagram_streamdriver_test} the test suite
-\lil{StreamDriverT} is using the \lil{SimulatorStream} to properly
-simulate the \lil{Stream} used by \lil{StreamDriver}, which is the
-system under test.
+% simulator roles in tests
+The simulators are used in the hand made test suits as stubs.
+The resulting structure is shown in Figure \ref{fig:class_diagram_streamdriver_test}.  
+For instance, in Figure \ref{fig:class_diagram_streamdriver_test} the test suite \lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate the \lil{Stream} used by \lil{StreamDriver}, which is the system under test.
 
@@
 
-In our case, the use of simulators to enable unit testing, has been
-essential to test the \STSB driver, since the physical board was not
-available.  The simulators were have been used in both the hand made
-and the generated test suite, since both of them need a proper set of
-stubs to be executed.
-
-% =====================================================================
-\section{Test cases on the job}
-\label{sec:test_cases_on_the_job}
-
-JML specifications, the simulators, and both test suites has been used
-to:
-
-\begin{itemize}
-\item verify hardware implementation;
-\item verify driver implementation, with and without the physical
-  board;
-\item guide driver development.
-\end{itemize}
-
-The first \STSB prototype delivered was meant to build empty packets,
-but the packets were built in a way not consistent with the
-specification, the sentinel values used to delimit a packet were not
-correct.  The board implementation errors were detected by simply
-using the driver implementation, the \lil{StreamDriver} package.
-
-The second prototype delivered was meant to build packets with values
-taken from real sensors installed on the board.  The only sensors that
-were not yet working were the sensors devoted to the fast rate data
-streams, the audio sensors.  Only one error needed to be corrected in
-the driver implementation, because of a (rare) combinations of
-conditions that were not covered by the test cases, but that
-occasionally showed up during real use.  One error have been
-discovered in the low level system driver that the developed driver
-was using (the system driver was outside of our development scope); to
-eliminate the error an upgrade of the low level driver to a new
-unreleased version was necessary.  No other errors were detected on
-the driver side.
-
-The second prototype respected the packet byte structure
-specification, but was not fully compliant to the packet info
-structure specification and the single packet rules.  The formal model
-and the run time assertion checker enabled the detection of these
-errors.  4 protocol errors regarding the packet info structure
-specification, and 2 protocol errors regarding the single packet rules
-has been found. No other errors, except the mentioned ones, have
-been later on discovered.
-
-The hand made test suite is composed by over 140 tests, while the
-generated one is 100 times bigger, totaling over 14000 tests.  The big
-number of generated tests (1 to 100 ratio), can be understood looking
-closely at how the generated tests are created by the JML
-framework. The generated tests are exploring the possible input
-combinations on public methods, combining the type data ranges that
-have been specified for the test suite (see \cite{Cheon-Leavens02} for
-more details on how the JML framework works on generating unit
-tests). For instance, let's suppose we want to test a method
-\lil{m(int p1, int p2)} for a class \lil{C}; the data ranges specified
-for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the
-instances \lil{\{c1,c2,c3,c4\}}. The JML framework generates $24 = 3
-\times 3 \times 4$ tests, since they are the combination of the data
-ranges involved (the data range of the parameters and the data range
-of the type owning the method under test). Fortunately, not all the
-generated tests are executed.
-
-A generated test, as it was previously mentioned (see section
-\ref{sec:test_cases}), is not executed, unless the preconditions of
-the called method are satisfied. Thus, depending on the preconditions
-specified for a method, the number of the generated tests that are
-actually executed in the test suite can be significantly smaller 
-compared to the total number of generated tests.
-
-To compare in a quantitative way the effectiveness of the test suites,
-code coverage metrics have been used; this is not the only measure
-that have been considered, since it is almost unanimously accepted
-that code coverage alone cannot asses the quality of a test
-suite\cite{Marick1999,Chockler2006}.
-
-Various coverage metrics exist, we analyzed the test suites with
-statement code coverage (one of the simplest types)\footnote{The tool
-  used to obtain statement code coverage metrics is
-  \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage
-reports whether each executable statement has been encountered. This
-code coverage metric is easy to implement, but lacks the capability to
-explore the control flow graph associated to the code, that is,
-decision branches are not explored. The coverage results are reported
-in the Table \ref{tab:statement_code_coverage}.
+% simulator importance
+The use of simulators to enable unit testing was essential to test the \STSB driver, since the physical board was not available.
+The simulators are used in both the hand made and the generated test suite, since both need a proper set of stubs to be executed.
+
+\subsection{Test cases on the job}
+\label{subsec:test_cases_on_the_job}
+
+% test suites sizes
+The hand made test suite is composed by over 140 tests, while the generated one is 100 times bigger, totaling over 14000 tests.
+The big number of generated tests can be understood looking closely at how the generated tests are created by the JML framework. 
+The generated tests explore the possible input combinations on public methods, combining the type data ranges that are specified for the test suite (see \cite{Cheon-Leavens02} for more details on how the JML framework works on generating unit tests).
+For instance, let's suppose we want to test a method \lil{m(int p1, int p2)} for a class \lil{C}; the data ranges specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the instances \lil{\{c1,c2,c3,c4\}}.
+The JML framework generates $24 = 3 \times 3 \times 4$ tests, since they are the combination of the data ranges involved (the data range of the parameters and the data range of the type owning the method under test).
+Neverteless, not all the generated tests are executed.
+A generated test is not executed, unless the preconditions of the called method are satisfied. 
+Thus, depending on the preconditions specified for a method, the number of the generated tests that are actually executed in the test suite can be significantly smaller compared to the total number of generated tests.
+
+% test execution time, values
+Test suites execution time are different, the hand made test suite, with the run time assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $).
+Most of the time is spent on setting up the suite, only considering the real execution of the tests, with no suite setup, the execution time reduces to less than 4 minutes (the rough ratio is $ 1 : 10 $).
+
+% test effectiveness
+Code coverage metrics is used to compare in a quantitative way the effectiveness of the test suites. 
+This is only an indicator, since it is accepted that code coverage alone cannot asses the quality of a test suite\cite{Marick1999,Chockler2006}.
+Various coverage metrics exist, the test suites are analyzed with statement code coverage (one of the simplest types)\footnote{The tool used to obtain statement code coverage metrics is \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage reports whether each statement has been encountered during execution.
+Statement code coverage is easy to implement, but lacks the capability to explore the control flow graph associated to the code, that is,
+decision branches are not explored. 
+The coverage results are reported in the Table \ref{tab:statement_code_coverage}.
 
@@
 
-Test suites execution time are very different, the hand made test suite, with the
-run time assertion checking active runs completely in less than one
-minute on an average PC, while the generated test suite runs in more
-than 20 minutes (the rough ratio is $ 1 : 65 $). Most of the time is
-spent on setting up the suite, only considering the real execution of
-the tests, with no suite setup, the execution time reduces to less
-than 4 minutes (the rough ratio is $ 1 : 10 $).
+\subsection{Retrospective on test cases effectiveness}
+\label{subsec:test_cases_retrospectives}
+
+% <100 coverage measures, utility methods
+The coverage measures reported in Table \ref{tab:statement_code_coverage} do not reach 100\%. 
+This is because of two different effects: non public utility methods, and abstract methods in interfaces.
+The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, they are not meant to be called by the system, and they are only used in object initialization or in object setups performed during unit tests.
+In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes.
+The used statement coverage tool cannot be configured to distinguish these code blocks, so this effect is unavoidable. 
+The effect is mainly seen in low \lil{SimulatorDriver} code coverage figures: the \lil{SimulatorDriver} package has many utility non public methods.
+
+% <100 coverage measures, interfaces
+Regarding interfaces, an interface contains no statements, so when an interface method is called, it is the method of the implementation class used that is actually covered.
+This effect is visible in \lil{Driver} coverage figures: \lil{Driver} package contains interfaces (that are not counted at all) and some very simple real class (exceptions) that are not throughly tested.
+The net result is extremely low statement code coverage figures because the very simple classes are the only ones that are actually covered.
+
+% code coverage interpretation
+The statement code coverage figures show that the test suites do a good job on the most important package, the \lil{StreamDriver}. 
+The generated test suite reaches almost $ 80 \% $, while the hand made tests have only a slightly lower coverage ratio: $ 76 \% $. 
+The combination of test suites raise the coverage ratio to $ 88.5 \% $. 
+This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not to be considered alternatives, but they need to be used together to achieve the higher benefits, is confirmed.
+A similar indication is provided by \lil{SimulatorDriver} package as well.
+
+% hand made unit test vs generated unit test
+The qualitative difference of the two test suites can be understood observing how the suites are built. 
+An hand made unit test usually requires objects initialization (it can involve several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
+A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviors. 
+With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the hand made ones; but the specification is not complete, and a large enough data ranges would take long to be developed.
+The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviors, and the hand made, checking the more complex ones, whether formalized or not.
 
 % =====================================================================
-\section{Retrospective on test cases effectiveness}
-\label{sec:test_cases_retrospectives}
-
-The coverage measures reported in Table
-\ref{tab:statement_code_coverage} are not high, in absolute
-values. This is because of two different effects: non public methods,
-and interfaces. The effect of non public methods (to be precise:
-package Java visibility) is that these methods are taken into account
-as well as public and protected ones during code coverage calculation;
-but these methods are not part of the interface, they are not meant to
-be called by the system, and they are only used in object
-initialization or in object setups performed during unit tests. In
-fact there are package and private code blocks that are not reachable
-by construction from public and protected methods, and this is a
-precise characteristic reflecting the code architecture, meant to ease
-class instances setup and configuration for testing purposes. The used
-statement coverage tool cannot be configured to distinguish these code
-blocks, so this effect is unavoidable. The effect is mainly seen in
-low \lil{SimulatorDriver} code coverage figures: the
-\lil{SimulatorDriver} package has plenty of setup and initialization
-non public methods.
-
-Regarding interfaces, an interface contains no statements, so when an
-interface method is called, it is the method of the implementation
-class used that is actually covered. This effect is visible in
-\lil{Driver} coverage figures: \lil{Driver} package contains
-interfaces (that are not counted at all) and some very simple real
-class (exceptions) that are not throughly tested. The net result is
-extremely low statement code coverage figures because the very simple
-classes are the only ones that are actually covered. This effect could
-be easily avoided writing specific hand made tests for the simple
-classes and including them in the classes for which tests are
-generated.
-
-The statement code coverage figures show that the test suites do a
-good job on the most important package, that is, the
-\lil{StreamDriver}. The generated test suite reaches almost $ 80 \% $,
-while the hand made tests have only a slightly lower coverage ratio: $
-76 \% $. But the combination of test suites is interesting, raising
-the coverage ratio to $ 88.5 \% $. This is a clear indication that one
-test suite is not completely overlapping the other: the test suites
-are not to be considered alternatives, but they have to be used
-together to achieve the higher benefits, they are complementary. This
-consideration can be confirmed looking at the figures obtained for
-\lil{SimulatorDriver} package as well.
-
-The qualitative difference of the two test suites can be understood
-analyzing how the suites are built. The hand made test suite primarily
-focuses on small features and use cases fragments that the system must
-implement, or for which an exceptional case has been specified. A test
-case is to be considered a rough specification written in imperative
-style. The common structure requires an object initialization,
-followed by a hopefully not too long sequence of public methods taking
-an object (or a set of objects) into a desired state; than the method
-under test is called and the object state is finally inspected to
-verify if it is in the expected state. Mock objects can be used to
-limit the dependencies. The test suite should be composed only by unit
-tests, but usually some tests are more complex to be called proper
-unit tests; they can be seen as simple integration tests.
-
-The generated test suite has a more focused and limited scope: since
-only one method call is performed in each test, hence it is difficult
-to explore complex behaviors. It is true that with a complete
-specification, and specifying large enough data ranges (the
-appropriate data ranges are combined to generate the tests), the
-generated tests could theoretically achieve the same expressive power
-as the hand made ones, but the price to pay would be expensive: much
-longer and more complex specifications, requiring more time to be
-designed and or updated, and more tests to be run, increasing the
-already not so short computation time of the generated test suite.
-
-The test suites gave their best when used in a combined approach: the
-generated test suite, checking simpler behaviors, and the hand made,
-checking the more complex ones.  During the development of the driver
-code and the simulators, the generated test suite was helpful to test
-the incrementally growing specifications and the finer details of each
-method, while the hand made one was helpful to verify the behavior on
-complex methods calls on objects initialized to specific
-states. Having an hand made test suite also make the code style more
-test oriented, which usually leads to a better quality and more
-modularized source code\cite{Binder1999}.
-
-The test suites has been used to verify different parts of the
-systems, that is, the real system as well as the simulators.  The
-multiple simulators and multiple test suites approach is not new,
-since it has been often used in embedded systems software
-testing\cite{Broekman2002}.  This is a general approach that can (and
-should) be used to verify both the simulators built and the real
-systems.  It can be easily implemented only if the code has been
-structured following the already mentioned dependency inversion
-principle\cite{martin1996dependency} (see Figure
+\section{Related works}
+\label{sec:related_works}
+
+\todo{Out of scope: agile & formal used together, but not mixed}
+
+\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
+
+\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
+
+\todo{Mixin with test substitution}
+
+% =====================================================================
+\section{Conclusions}
+\label{sec:conclusions}
+
+\section{Extracts}
+
+\paragraph*{limited resources}
+
+% effect of the limited resources on the formal specifications
+Hard constraints on driver delivery deadlines define the completeness level of the (\emph{just good enough}) specification.  
+In \STSB driver JML specifies very simple behaviors (method oriented) on the whole driver implementation, and more complex behavior on the driver core interfaces.  
+The software system parts not specified by the JML2 specification were specified using hand made unit tests.
+
+\paragraph*{hand made unit tests}
+
+% hand made unit tests purposes
+Hand made unit tests specifies the behavior not specified by formal specification; nevertheless this is not the only purpose they have.
+Hand made unit tests have been used to drive the development, since the Test Driven Development approach\cite{beck2003test} has been
+followed during development.
+
+% hand made unit tests suite structure
+The hand made test suite primarily focuses on small features and use cases fragments that the system must implement, or for which an exceptional case has been specified. 
+A test case is to be considered a rough specification written in imperative style. 
+The common structure requires the objects initialization, followed by a sequence of method calls taking the objects into desired initial states; than the method under test is called and the objects states are finally inspected and verified.
+Mock objects can be used to limit module dependencies. 
+The test suite should be composed only by unit tests, but usually some tests are more complex to be called proper unit tests; they can be seen as simple integration tests.
+
+\paragraph{formal specification of the protocol}
+
+% why a state machine has not been used to formally specify the protocol
+\ST stream communication protocol is characterized by a well definite packet structure, and a well definite semantic on the packet contents.  
+On the other hand, in the protocol, there are no strong constraints on synchronization, since the communication channel is splitted in two (in and out) separate concurrent and asynchronous channels.
+The lack of hard synchronization constraints reduces the advantage to use a classical state machine based specification (as the one presented in \cite{Hubbers2004}), hence this option has been discarded during development.
+
+\paragraph{jml}
+
+% how jml generated tests work
+The test suite generated is a collection of fine grained unit tests.
+For each public method a set of tests is generated.
+The specified behavior of the method is checked using the JML2 run time assertion checker, hence, an oracle is not needed.  
+The generated tests send messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation exceptions; if a precondition is not met the test is not run, if the precondition is met the test is run and if the postcondition is not met an exception is fired and the test fails.
+The test data used to build the tests is the only part that is required to be provided by the user.
+
+\paragraph{unit tests}
+
+% how unit tests should work
+A proper stub, driver and oracle need to be identified for each method under test to obtain a proper unit test suite; stub, driver and oracle are the usual components found in the harness that enable test execution\cite{Binder1999}.
+
+% what is needed to test a package
+To properly unit test a package, the following components need to be identified:
+\begin{description}
+\item[Stub or Skeleton]: a piece of code that simulates the activity of the modules that are not under tests.
+\item[Driver]: a piece of code that calls in a proper way the module under test.
+\item[Oracle]: a mechanism used for determining whether a test has passed or failed, comparing the output of the module under test to the output the oracle determines the module should have.
+\end{description}
+
+% junit
+In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the framework used to manage the implementation and execution of hand made and generated unit tests, there is no clear distinction between drivers, stubs, and oracles; none-less, even if a proper distinction
+is not imposed by the framework, the conceptual distinction still holds.
+
+% who provides what: hand made vs generated
+Hand made unit tests and JML generated tests differ in how test components are provided.
+The distinctions and similarities are summarized in Table \ref{tab:test_harness}.
+\begin{table}[htbp]
+  \caption{Hand made and generated tests harness.}
+  \label{tab:test_harness}
+  \begin{center}
+    \begin{tabular}{|l|l|l|}\hline
+      & \textbf{\textit{hand made}} & \textbf{\textit{JML}} \\\hline
+      \textbf{\textit{Stub}} & to be provided & to be provided\\\hline
+      \textbf{\textit{Driver}} & to be provided & automatic: 1 call\\\hline
+      \textbf{\textit{Oracle}} & to be provided & 
+      automatic: runtime assertion checker\\\hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\todo{expand, test + RAC variations}
+
+% test execution time, reasons
+The execution time of the generated tests is influenced by:
+\begin{itemize}
+\item The number of the generated tests. 
+This problem is not directly addressable, since it is the heart of the test suite generation creation method itself. 
+All the combinations of possible inputs for each method under test must be generated.
+\item The runtime assertion checker overhead. 
+Each test is executed while RAC checks all the relevant properties. 
+We can speculate that if the RAC assertions could be triggered on and off on single properties in a dynamic way (during runtime), a test suite would take advantage of this possibility, turning on the RAC only on the properties that have to be verified for the classes or methods under tests (and not for all the involved classes and methods); this could save time during the test suite execution.
+\end{itemize}
+
+% tests are helful
+The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the hand made one was helpful to verify the behavior of complex methods calls on objects initialized to specific states.
+
+% tests leads to a better design
+An hand made test suite force the code to be test oriented, which usually leads to a better quality and more modularized source code\cite{Binder1999}.
+
+\paragraph*{simulators and tests}
+
+% simulators and tests
+Simulators can be used to ease the tasks of developing the test harness components (Stub, Driver, Oracle).
+A simulator can be used as a stub, to replace a module needed by the module under test.
+A simulator can be used as a driver, stimulating the module under test.
+Finally, a simulator can be used as an oracle, simulating the module under test.
+
+% multiple simulators approach
+The multiple simulators approach is not new, since it has been often used in embedded systems software testing\cite{Broekman2002}.  
+This is a general approach that can (and should) be used to verify both the simulators built and the real systems.  
+It can be easily implemented only if the code has been structured following the dependency inversion principle\cite{Martin1996} (see Figure
 \ref{fig:class_diagram_main}).
 
-The combined test approach, driven by the packet protocol
-specification, has been proved successful on verifying the protocol
-itself and the software implementing it.  Nevertheless, it has to be
-noted that the highest levels of the protocol are more problematic
-(see \ref{sec:data_stream_and_protocol_description}: ``Packet sequence
-rules'' and ``Packet input output asynchronous rules'').  On these
-levels, the protocol properties regard sequences of packets, and
-testing them can be tricky, since the sequences of packets are not
-referred to explicitly, in the implementation.  Two possible
-strategies can be used:
-
+\paragraph*{prototype history}
+
+% tests and the first prototype, board side
+The first \STSB prototype was meant to build empty packets, but the packets were built in a way not consistent with the specification.  
+The board errors have been detected by simply using the driver implementation, the \lil{StreamDriver} package.
+
+% tests and the second prototype, driver side
+The second prototype was meant to build packets with values taken from real sensors installed on the board.  
+The only sensors that were not yet working were the sensors devoted to the fast rate data streams (audio sensors).  
+One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions that were not covered by the test cases, but that occasionally showed up during real use.
+One error have been discovered in the low level system driver that the developed driver is using (the system driver is outside our development scope); to eliminate the error an upgrade of the low level driver to a new unreleased version were necessary.
+No other errors were detected on the driver side.
+
+% tests and the second prototype, board side
+The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
+The formal JML specifications and the run time assertion checker enables the detection of these errors.
+4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found. 
+No other errors, except the mentioned ones, have been discovered.
+
+\paragraph*{future work}
+
+% problem on highest levels of the protocol
+The highest levels of the protocol could be problematic (see \ref{subsec:data_stream_and_protocol_description}: ``Packet sequence rules'' and ``Packet input output asynchronous rules''). 
+On these levels, the protocol properties regard sequences of packets, and testing them can be tricky, since the sequences of packets are not referred to explicitly, in the implementation.
+
+% solutions on highest levels of the protocol
+Two possible strategies could be used:
 \begin{itemize}
-\item Translate the packet sequence properties in local packet
-  properties (when this is possible).  The advantage is to have
-  properties shown as the usual pre-conditions, post-conditions and
-  invariants.  However, the localized properties can be more complex
-  and less intuitive and comprehensible than the original ones;
-  finally, some of the global properties could prove themselves
-  impossible to be translated into local ones.
-\item Use separate utility classes, meant to verify a sequence of
-  packages.  This approach is not clean, since it requires adding and
-  using new classes, unneeded by the core implementation, to verify
-  the protocol.  It has the advantage of keeping the properties that
-  have to be verified mostly unchanged.
+\item Translate the packet sequence properties in local packet properties (when possible).  
+The advantage is to have properties shown as the usual pre-conditions, post-conditions and invariants.  
+However, the localized properties can be more complex and less intuitive and comprehensible than the original ones; finally, some of the global properties could be impossible to translate.
+\item Use separate utility classes, meant to verify a sequence of packages.  
+This approach is not clean, since it requires adding and using new classes, unneeded by the core implementation, to verify the protocol.
+It has the advantage of keeping the properties that have to be verified mostly unchanged.
 \end{itemize}
 
-One problem we noticed regards the execution time of the generated
-tests.  It is long for two reasons:
-
-\begin{itemize}
-\item Number of the tests generated. This problem is not directly
-  addressable, since it is the heart of the test suite generation
-  creation method itself: generating all the combinations of possible
-  inputs for each method under test.
-\item Runtime assertion checker. Each test is executed while RAC
-  checks all the relevant properties. This problem, again, is not
-  directly addressable. We can speculate that if the RAC assertions
-  could be triggered on and off on single properties in a dynamic way
-  (during runtime), a test suite would take advantage of this
-  possibility, turning on the RAC only on the properties that have to
-  be verified for the classes or methods under tests (and not for all
-  the involved classes and methods); this could save time during the
-  test suite execution.
-\end{itemize}
+% old paper
+\section{OLD, to be reused}
+
+% =====================================================================
 
@@
 %%% End:
+
+% =====================================================================
+\section{Data stream and protocol description}
+\label{sec:data_stream_and_protocol_description}
+
+\STSB and \STPU communicates via an Universal Serial Bus (USB)
+interface.  A proprietary protocol is used on top of the USB serial
+channels, the protocol is specified by
+FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
+is the manufacturer of the USB controller chip used in the \STSB, the
+FT2232H\cite{ftdi_ft232h_2009}.
+
+The proprietary protocol enables the control and management of some of
+the functionalities in FT232H chip (such as internal EEPROM read and
+write, reset of the chip, etc.), as well as the definition of the
+communication protocol and properties to be used in the asynchronous
+channels (input and output channels) provided by FT232H.  These
+functionalities are accessed via an Application Program Interface, the
+API is provided by a driver developed directly by FTDI, which is
+available for various systems (Linux 32 and 64 bits, Mac OS, Windows)
+\cite{ftdi_d2xx_api_2009}.
+
+Finally, a custom protocol has been developed to manage the entire
+\STSB, built on top of the serial channels provided by the API.  The
+custom protocol is meant to be used to read sensor data observations,
+enable or disable sensors, configure on-board components (as the
+Field-Programmable Gate Array and the micro processor); the protocol
+main function is to read sensor data.
+
+The protocol is asynchronous and packet based, the packets have a
+fixed length.  The protocol has been divided in multiple layers to
+ease its formal description.  The rationale is the following: mixing
+up different abstraction levels of a protocol description can lead to
+overly complex and unmanageable descriptions, especially when dealing
+with formal ones.  The layers identified in protocol are the
+following:
+
+\begin{description}
+\item[Packet byte structure]: the internal representation of the
+  packet.
+\item[Packet info structure]: the meaningful fields contained in the
+  packets.
+\item[Single packet rules]: the content acceptable values, and how
+  they influence each other.
+\item[Packet sequence rules]: the content acceptable values, based on
+  values of previously received packets.
+\item[Packet input output asynchronous rules]: the content acceptable
+  values and reaction constraints on output packets, based on
+  previously transmitted input packets.
+\end{description}
+
+We will examine the output sensor data packet, since it is the most
+used and interesting one; we won't describe input or other output
+packets.  We will examine the first three layers, that is: the packet
+byte structure, the packet info structure and the single packet rules.
+
+The packet byte and info structures of the output sensor data packet
+reflect the board capabilities and built in sensors.  Essentially a
+single packet has to accommodate various types of data:
+
+\begin{itemize}
+\item Fast data rate streams: audio data, 48KHz (standard audio) or
+  96KHz (ultra audio).
+\item Medium data rate streams: 5KHz.
+\item Slow data rate streams: temperature, pression, accelerometer,
+  etc.; all these sensor have a data rate lower than 5 KHz.
+\item Metadatas: metadatas on sensors and board state.
+\end{itemize}
+
+The internal structure of the package (as we will see) is strictly
+fixed, even if the data streams have different data rates.  This
+choice originates from a trade-off between two constraints: the
+efficiency of the transmission (redundant data to be minimized) and
+the complexity to build a package (complexity to be minimized).  The
+complexity minimization constraint rationale is the following: the
+FPGA on the \STSB builds the packet, but the FPGA is also to be used
+for various other specific tasks; it is important to maintain the
+building packet code on the FPGA as simple and short as possible, to
+leave enough space to implement other tasks.  The Output sensor data
+packet structure is shown in Table \ref{tab:data_packet_structure},
+the packet total length is 1024 bytes (512 16 bit words).
+
+\begin{table}[htbp]
+  \caption{Output sensor data packet structure.}
+  \label{tab:data_packet_structure}
+  \begin{center}
+    \begin{tabular}{|c|l|}\hline
+      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
+      0-1 & sentinel\\\hline
+      2-7 & slow data\\\hline
+      8-11 & metadata: board state\\\hline
+      12-13 & reserved\\\hline
+      14-15 & metadata: board state\\\hline
+      16 & reserved\\\hline
+      17-22 & frame 1\\\hline
+      \ldots & \ldots\\\hline
+      503-508 & frame 82\\\hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+The frames accommodate data from the fast and medium data rate streams
+(there are 4 fast data rate stream and 8 medium data rate stream
+channels).  There are 82 frames in a packet.  The frame structure is
+shown in Table \ref{tab:frame_structure}.
+
+\begin{table}[htbp]
+  \caption{Output sensor data packet frame structure.}
+  \label{tab:frame_structure}
+  \begin{center}
+    \begin{tabular}{|c|l|}\hline
+      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
+      0 & metadata: frame description\\\hline
+      1 & medium data rate data sample\\\hline
+      2-5 & fast data rate data samples\\\hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+Looking at the structure described and the data rates frequencies, it
+can be seen that there must be some subtle rule, to make all the
+figures to add up; that is, to be capable to accommodate three
+different rates outputs in a fixed structure, fixed length packet.
+The subtleties lies in two forms of redundancies, for slow and medium
+data rates:
+
+\begin{itemize}
+\item Slow data rates: slow data can be found at the beginning of each
+  packet, the data is simply repeated (that is, it is equal to the
+  previous packet in the stream) when no new data is available.
+\item Medium data rates: the 8 medium rate channels are multiplexed
+  together into the medium data rate data sample of each frame.  The
+  medium data rate sample channel is a piece of info to be extracted
+  from the frame description. Besides, the medium data rate data
+  sample could contain no data at all, this information is to be
+  extracted from the frame description. So, one frame could contain a
+  medium data rate sample from channel 0, the following frame a sample
+  from channel 5, the following frame no sample at all, etc.
+\end{itemize}
+
+The single packet rules delimit the boundaries of the values
+obtainable from the packet: each defined sensor represented in the
+packet, as well as the metadata describing the \STSB and the packet
+and frame contents, are constrained by a range of acceptable values.
+There are also rules affecting more than one value, these are
+essentially validity rules (i.e. if a specific sensor is active the
+reading value has to be in a defined range), and rules specifying a
+correct sequence of frames (i.e. there can be only one switch in
+sampling rate in the frame sequence of a single packet).

### 2010-07-01 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
almost there, I need a conclusion it has to be reread and spell checked it has to be converted to the format needed by the workshop

@@
 
-\documentclass{article} \usepackage{times}
+\documentclass{article}
 
-\usepackage{ifpdf} \usepackage{a4wide} \usepackage{pdfsync}
+\usepackage{times}
+\usepackage{ifpdf}
+\usepackage{a4wide}
+\usepackage{pdfsync}
 
-\ifpdf \usepackage[pdftex]{graphicx} \else \usepackage{graphicx} \fi
+\ifpdf 
+  \usepackage[pdftex]{graphicx} 
+\else 
+  \usepackage{graphicx} 
+\fi
 
-\usepackage{xspace} \usepackage{tabularx} \usepackage{epsfig}
-\usepackage{amsmath} \usepackage{amsfonts} \usepackage{amssymb}
-\usepackage{eucal} \usepackage{stmaryrd} \usepackage{float}
-\usepackage{listings} \input{jml-listings}
+\usepackage{xspace} 
+\usepackage{tabularx} 
+\usepackage{epsfig}
+\usepackage{amsmath} 
+\usepackage{amsfonts} 
+\usepackage{amssymb}
+\usepackage{eucal} 
+\usepackage{stmaryrd} 
+\usepackage{float}
+\usepackage{listings}
+
+\input{jml-listings}
 \lstset{language=[JML]Java,xleftmargin=20pt,xrightmargin=20pt}
 
-\ifpdf \usepackage{epstopdf}
-\usepackage[pdftex,bookmarks=false,a4paper=false,
-plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,
-linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Joseph
-R. Kiniry and Vieri del Bianco"]{hyperref} \else
+\ifpdf 
+  \usepackage{epstopdf}
+  \usepackage[pdftex,bookmarks=false,a4paper=false,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Joseph R.
+Kiniry and Vieri del Bianco"]{hyperref}
+\else
   \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
@@
 
-\newcommand{\tablesize}{\footnotesize} \newcommand{\eg}{e.g.,\xspace}
-\newcommand{\ie}{i.e.,\xspace} \newcommand{\etc}{etc.\xspace}
+\newcommand{\tablesize}{\footnotesize} 
+\newcommand{\eg}{e.g.,\xspace}
+\newcommand{\ie}{i.e.,\xspace} 
+\newcommand{\etc}{etc.\xspace}
 \newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
-\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3
-    \myhref{#1}{#1}}}
-
+\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3\myhref{#1}{#1}}}
 \newcommand{\lil}[1]{\texttt{\lstinline|#1|}}
 
-% \newcommand{\notev}[1]{\xspace$\textcolor{blue}{\omega^\textsf{vieri}}$\marginpar{\scriptsize\textsf{Vieri:}
-%     #1}}
-\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:}
-    #1}}
-\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}
-    #1}}
-\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:}
-    #1}}
+\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:}#1}}
+\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}#1}}
+\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:}#1}}
 
@@
 \newcommand{\ST}{\emph{SenseTile}\xspace}
-\newcommand{\SB}{\emph{Sensor Board}\xspace} \newcommand{\STSB}{\ST
-  \SB\xspace} \newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
+\newcommand{\STs}{\emph{SenseTiles}\xspace}
+\newcommand{\SB}{\emph{Sensor Board}\xspace} 
+\newcommand{\STSB}{\ST \SB\xspace} 
+\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
 \newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
 \newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
-\newcommand{\STU}{\ST \emph{Unit}\xspace} \newcommand{\STUs}{\ST
-  \emph{Unit}\xspace}
-
-\newcommand{\STs}{\emph{SenseTiles}\xspace}
+\newcommand{\STU}{\ST \emph{Unit}\xspace} 
+\newcommand{\STUs}{\ST \emph{Units}\xspace}
 
 \newcommand{\simulator}{\STSB \emph{Simulator}\xspace}
-
 \newcommand{\datastore}{\ST Scientific Datastore\xspace}
 \newcommand{\computefarm}{The \ST Scientific Compute Farm\xspace}
-\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute
-  Server Farm\xspace} \newcommand{\sensorfarm}{The \STSBs\xspace}
-
-% ---------------------------------------------------------------------
-% New commands, macros, \etc
-% ---------------------------------------------------------------------
+\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute Server Farm\xspace} 
+\newcommand{\sensorfarm}{The \STSBs\xspace}
 
@@
 
+% principles of agile methodologies
 Agile methodologies are all based on a highly iterative and incremental process.
@@
 
-In the present context, focused on how to merge agile and formal practices, we are mainly interested in the verification practices used in agile methodologies.
+% verification practices in agile methodologies
 The agile manifesto does not introduce any specific techniques to be used in the development process, it describes agile methodologies from a very high abstraction level.
 Nevertheless most of the agile methodologies share a similar approach to artifacts verification.
-Functional test suites, to be used as precise requirements or user stories definition, are used in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}, while unit tests and TDD are mentioned, whether not completely accepted, in most of the agile methodologies defined so far.
+Functional test suites, to be used as precise requirements or user stories definition, are used in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
+Unit tests and TDD are mentioned, whether not completely accepted, in most of the agile methodologies defined so far.
 
-High abstraction level test suites, functional tests, describe requirements, and actually bind together requirements documentation and system behaviour, the development cycle is one iteration long. 
-Low abstraction level test suites, unit tests, describe a module behaviour, and actually bind together code documentation, code and the contract imposed on the module, the development cycle is usually less than one day.
+% abstraction in test suites
+High level test suites, composed of functional tests, describe requirements and bind together requirements documentation and system behaviour; the associated development cycle is one iteration long. 
+Low level test suites, composed of unit tests, describe the modules behaviour and bind together code documentation, code and the contract imposed on the module; the associated development cycle is usually less than one day (it can be as short as a few minutes).
 
+% refactoring in iterative development
+Continuous refactoring~\cite{Fowler1999} enables an interative and incremental development, since the system architecture and design, whether explicitly specified or not, need to be cleaned up and modified, to accomodate new requirements, services and modules.
 Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.
-Continuous refactoring~\cite{Fowler1999} enables an interative and incremental development, since the system architecture and design, whether explicitly specified or not, needs to be cleaned up, after implementation, and modified, to accomodate new requirements, services and modules.
 The automated test suites guarantees that the functionalities and behaviour already implemented in the code are not compromised by the system refactorings.
@@
 
+% JML introduction
 The Java Modeling Language (JML)\footnote{\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}} is the de facto standard specification language for describing the behavioral properties of Java programs.
@@
 
+% using JML
 JML is sometimes used in a Design-by-Contract style, where a specification is written from scratch, reusing existing APIs that have specifications of their own, and then an implementation is written conforming to that specification~\cite{DBCjml,DBC}.
@@
 
+% JML tools
 The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{RAC,BurdyEtal05-STTT,ESC}.
@@
 
-\todo{describe loop importance}
+% objectives, repeated
+The objective is to formally specify only parts of the system under development, in an incremental iterative process.
+Formal methodologies artifacts need to coexist with traditional agile development artifacts: requirements, expressed as functional tests or as formal specifications, hand made and generated tests, documentation, source code.
+The development process needs to guarantee consistency all over the involved artifacts.
 
-\todo{describe partial specification importance, and consequences}
+% composition of the two worlds
+Both Test Driven Development (TDD) and classical Formal specification driven development (FDD) guarantee consistency on all the involved artifacts. 
+TDD force an iterative development, whether FDD support it without forcing it.
+The development cycle adopted in TDD is shown in Figure \ref{fig:tdd_diagram} and the one adopted in classic FDD is shown in Figure \ref{fig:fdd_diagram}.
+It can be seen how unit tests in TDD are substituted in FDD by formal specifications supported by the verification environment, in our case the verification environment is the generated unit test suite.
+Formal specifications replace unit tests: they specify, document and verify the system under development.
 
-\todo{consequences: no complete formal specification, no theorem provers, no model checking, yes run time assertions, yes generated tests}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+    \includegraphics[scale=0.5]{diagrams/TDD_classic}
+    \caption{TDD development cycle.}
+    \label{fig:tdd_diagram}
+  \end{minipage}
+  \hspace{0.02\linewidth}
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+    \includegraphics[scale=0.5]{diagrams/FDD_classic}
+    \caption{FDD classic development cycle.}
+    \label{fig:fdd_diagram}
+  \end{minipage}
+\end{figure}
 
-\todo{we need both worlds, together}
+% composition is not enough
+FDD is used for the formal specified parts of the system, whether TDD is used for the remaining ones: a ``composition'' of formal and agile verification practices.
+Yet, it's not enough.
+There are several problem that still needs to be addressed.
+No communication between the two worlds is permitted, it's not defined how to formally specify a module after it was implemented and verified in a non formal way.
+A module only partially specified is a sort of hybrid that needs to be treated accordingly.
+Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is usually not the case.
+Identify the correct formal specification is a difficult task, that needs to be supported and verified by the development process.
 
-\todo{FDD: formal specification driven development (formal -> hand made tests -> code, design, refactorings, run-time verification)}
+% FDD with hand made tests
+A partially specified module cannot be verified properly only through automatically generated tests.
+Hand made tests have to be used together with generated ones.
+Hand made tests verify complex behaviors that are not formally speified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
+Limits in the formal verification tools need to be considered when deciding whether supporting the code with hand made tests, verification tools based on automatically generated tests~\cite{Glass2002} are not an exception.
+The development cycle to be used to blend formal specifications, hand made tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the hand made tests, both hand made tests and the formal specification drive the code development.
 
-\todo{FDD: formal specification driven development (formal -> generated tests -> code, design, refactorings, run-time verification)}
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+    \includegraphics[scale=0.5]{diagrams/FDD_hand_made_tests}
+    \caption{FDD development cycle with hand made tests.}
+    \label{fig:fdd_hand_made_tests_diagram}
+  \end{minipage}
+  \hspace{0.02\linewidth}
+  \begin{minipage}[b]{0.49\linewidth}
+    \centering
+    \includegraphics[scale=0.5]{diagrams/TDF}
+    \caption{TDF development cycle.}
+    \label{fig:tdf_diagram}
+  \end{minipage}
+\end{figure}
 
-\todo{TDD: test driven development, classical TDD (hand made tests -> code, design, refactorings)}
+% test driven formal specification
+The opposite path still needs to be covered, moving incrementally from a tested code to a formally specified code.
+In this case, new hand made tests and existing working code drives the formal specification: Test DRiven Formal specification (TDF).
+The resulting development cycle to be used to incrementally add formal specifications to an existing working code is shown in Figure \ref{fig:tdf_diagram}.
 
-\todo{increase code with formal specifications, TDF: test driven formal specification (hand made tests -> formal, design, refactorings, partial run-time verification)}
+% all together belnds into a mole
+When the Mole verification practices are used together the initial objectives can be achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; complete freedom to decide what needs to be formally specified; the development cycles and the associated verification practice maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
 
-\todo{result: formal: on some parts of the system}
+\todo{if time available: reason about test harness, who is providing what}
 
-\todo{result: tests: hand made tests, generated tests}
+\todo{if time available: reason about RAC, when to use it and when not to use it}
 
-\todo{result: hand made test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
+\todo{if time available: result: hand made test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
+
+\todo{unrevealed problem (hook for a follow up paper): TDF does not work well when testing postconditions, works correctly only when testing preconditions.
+idea1 (complex): a mock+delegate object, that behaves as the original object, except for a specific call on a specific method.
+idea2 (bad, the pre and post needs to be maintained synchronized): use utility class, where the precondition of a method is \~ the postcondition of the tested one}
 
 % =====================================================================
-\section{\todo{the case study}}
-\label{sec:___}
+\section{``Mole'' verification practices applied}
+\label{sec:a_complex_case_study}
 
@@
 Fast, medium and slow data rates have not a common multiple; a packet, to be capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
-
 \begin{itemize}
@@
 
+% single packet rules
 The single packet rules delimit the boundaries of the values obtainable from the packet: each defined sensor represented in the packet, as well as the metadata describing the \STSB and the packet and frame contents, are constrained by a range of acceptable values.
@@
 
-The unformal specification and verification is provided by a hand made test suite, composed of unit and integration tests; the tests specifies the behaviors of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
-The implementation is tricky: it checks for the sentinels repeatedly to be sufficiently confidant that the byte stream can be properly decomposed into a packet stream.
+% verification
+Verification is obtained through a hand made test suite, composed of unit and integration tests; the tests specifies the behaviors of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
+The implementation is tricky: it searchs for repeated patterns into a byte stream to be sufficiently confidant how the stream can be properly decomposed into packets.
+Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
+Therefore, packet byte structure is not formally specified.
 
-The specification can be used to verify a board simulator, but it cannot be used to check properties at run time.
+% development cycles
+The test suite are used to verify board simulators (low level), but it cannot be used to check properties at run time.
 Code and tests have been implemented using TDD.
@@
 
-The specification is provided by hand made test suites and JML annatations. 
-It has been built starting from hand made tests only, and incrementally adding tests and later on, JML annotations.
+% verification
+Verification is provided by hand made test suite, JML annotations and generated unit test suite. 
+It has been built starting from hand made tests only (TDD development cycles).
+Later on JML annotations have been added incrementally, using a mix of mole practices: to move from hand made tests to formal specifications, TDF have been initially used; than all the mole practices have been used according to the artifacts involved.
 
-\todo{explain: test -> code, test + code -> formal specification, formal specification -> generated tests}
+% development cycles
+The test suites are used to verify board simulators (high level), RAC and JML annotations are used to check the formally specified properties at run time.
+Code and tests have been initially implemented using TDD, than using all the Mole verification practices.
 
@@
 
-The specification is provided exclusively by the formal specification with JML language.
+% verification
+The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and hand made test suite to verify the most complex behaviour.
 It has been built starting from JML annoatations, supported by hand made tests when needed.
 
-\todo{formal specification -> test + generated tests + code}
+% development cycles
+The test suites are used to verify board simulators (high level), RAC and JML annotations are used to check most of the rules at run time.
+Code and tests have been implemented using both forms of FDD.
 
@@
 But, because of limited resources and affort avilable, the layer structure was not matched by corresponding simulators.
-Two simulators are available.
-
-% simulators development
-Simulators have been built concurrently with the specification, the driver implementation, and the tests.
+Two simulators are available, they have been built concurrently with the specification, the driver implementation, and the tests.
 
@@
       invariant ( 
-        \forall int channel; 0 <= channel && channel < Frame.ADC_CHANNELS; ( 
-          \num_of int i; 0<= i && i<(FRAMES-1); (
+        \forall int channel; 
+            0 <= channel && 
+            channel < Frame.ADC_CHANNELS; ( 
+          \num_of int i; 
+              0 <= i && 
+              i < (FRAMES-1); (
             (getFrame(i).isADCActive()) && 
@@
   \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__Main}
+  \caption{Main packages class diagram.}
   \label{fig:class_diagram_main}
-  \caption{Main packages class diagram.}
 \end{figure}
@@
   \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__Test}
+  \caption{Test packages class diagram.}
   \label{fig:class_diagram_test}
-  \caption{Test packages class diagram.}
 \end{figure}
@@
   \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
+  \caption{Generated test packages class diagram.}
   \label{fig:class_diagram_generatedtest}
-  \caption{Generated test packages class diagram.}
 \end{figure}
@@
   \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__StreamDriverT}
-  \label{fig:class_diagram_streamdriver_test}
   \caption{StreamDriver test packages class diagram: use of the stream
     simulator.}
+  \label{fig:class_diagram_streamdriver_test}
 \end{figure}
@@
 
-% =====================================================================
-\section{Related works}
-\label{sec:related_works}
-
-\todo{Out of scope: agile & formal used together, but not mixed}
-
-\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
-
-\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
-
-\todo{Mixin with test substitution}
-
-% =====================================================================
 \section{Conclusions}
 \label{sec:conclusions}
+
+\todo{write conclusions}
+
+%% \nocite{ex1,ex2}
+\bibliographystyle{latex8} 
+\bibliography{%
+  extra,%
+  abbrev,%
+  ads,%
+  category,%
+  complexity,%
+  hypertext,%
+  icsr,%
+  knowledge,%
+  languages,%
+  linguistics,%
+  meta,%
+  metrics,%
+  misc,%
+  modeling,%
+  modeltheory,%
+  reuse,%
+  rewriting,%
+  softeng,%
+  specification,%
+  ssr,%
+  technology,%
+  theory,%
+  web,%
+  upcoming,%
+  upcoming_conferences,%
+  conferences,%
+  workshops,%
+  verification,%
+  escjava,%
+  jml,%
+  nijmegen}
+
+\newpage
 
@@
 Hand made unit tests specifies the behavior not specified by formal specification; nevertheless this is not the only purpose they have.
-Hand made unit tests have been used to drive the development, since the Test Driven Development approach\cite{beck2003test} has been
+Hand made unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
 followed during development.
@@
   \begin{center}
-    \begin{tabular}{|l|l|l|}\hline
-      & \textbf{\textit{hand made}} & \textbf{\textit{JML}} \\\hline
-      \textbf{\textit{Stub}} & to be provided & to be provided\\\hline
-      \textbf{\textit{Driver}} & to be provided & automatic: 1 call\\\hline
-      \textbf{\textit{Oracle}} & to be provided & 
-      automatic: runtime assertion checker\\\hline
+    \begin{tabular}{|c|c|c|c|}\hline
+       & \textbf{\textit{hand made}} & \textbf{\textit{hand made + RAC}} & \textbf{\textit{generated}} \\\hline
+      \textbf{\textit{Stub}} & manual & manual & manual\\\hline
+      \textbf{\textit{Driver}} & manual & manual & automatic (1 call generation)\\\hline
+      \textbf{\textit{Oracle}} & manual & manual + automatic (RAC) & automatic (RAC)\\\hline
     \end{tabular}
@@
 \end{table}
-
-\todo{expand, test + RAC variations}
 
@@
 
-% old paper
-\section{OLD, to be reused}
+\paragraph*{out of scope related works}
 
-% =====================================================================
+\todo{Out of scope: agile and formal used together, but not mixed}
 
-% =====================================================================
-\section{Related works}
+\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
 
-JML has been used previously to verify a communication protocol in
-\cite{Hubbers2004}, but the protocol is a smart card protocol, and it
-is well represented through state machines, which are then refined
-into JML properties.  The approach is in waterfall style; than the
-need to develop the protocol, the specification and the implementation
-concurrently is not addressed.  No hand made or generated test suites
-are used, the code is specified with JML assertions, refined by the
-state machine specifications, and then checked statically with
-ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2
-tools RAC\cite{BurdyEtal05-STTT}.
+\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
 
-The JML-JUnit automatic test suite generation tool has not been widely
-used or assessed. It is described by its own creators in
-\cite{Cheon2002,Cheon2004,Cheon2005}; it has been used in
-\cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test
-suite, and, more importantly, only against proof of concepts system,
-it has never be proved against real complex systems.  A brief
-assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the
-analysis is based on injecting random generated errors in the source
-code, and then evaluating the ratio of the errors found by the test
-suite: the results shows that the JML-JUnit tool is not effective
-enough if used as a stand alone tool. No attempts on running or
-evaluating the generated test suite alongside with an hand made test
-suite are made.
+\todo{Mixin with test substitution}
 
-\section{Next papers}
+\paragraph*{case study related works}
 
-\paragraph*{TDS: Test Driven Specification}
+JML has been used previously to verify a communication protocol in \cite{Hubbers2004}, but the protocol is a smart card protocol, and it
+is well represented through state machines, which are then refined into JML properties.  
+The approach is in waterfall style; than the need to develop the protocol, the specification and the implementation concurrently is not addressed.  
+No hand made or generated test suites are used, the code is specified with JML assertions, refined by the state machine specifications, and then checked statically with ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2 tools RAC\cite{BurdyEtal05-STTT}.
 
-Approach:
-\begin{itemize}
-\item more specification used together: formal and not formal
-\item formal specification is not complete: it is just good
-  ``enough'', static checkers unfortunately cannot be used, since the
-  specification is partial
-\item use runtime assertion checker for formal specifications, on
-  tests
-\item use generated tests for formal specifications
-\end{itemize}
-
-Who drives what:
-\begin{itemize}
-\item specification (jml - formal specification or unit test -
-  examples fragments) driven development
-\item test driven specification
-\item formal specification is not complete: it is just good
-  ``enough'', static checkers unfortunately cannot be used, since the
-  specification is partial
-\item use runtime assertion checker for formal specifications, on
-  tests
-\item use generated tests for formal specifications
-\end{itemize}
-
-Benefits:
-\begin{itemize}
-\item real agile approach
-\item real formal approach
-\item use best of both worlds, do not try to use only on of them
-\end{itemize}
-
-Case example: actual paper, modified to fit the new parts, especially
-the test driven specification part.
-
-\paragraph*{BON}
-
-\todo{describe approach: high level informal specifications with BON,
-  iterative refinements of code, to distill test cases and formal
-  specification of the interfaces}
-
-% ======================================================================
-%% \nocite{ex1,ex2}
-\bibliographystyle{latex8} \bibliography{extra,%
-  abbrev,%
-  ads,%
-  category,%
-  complexity,%
-  hypertext,%
-  icsr,%
-  knowledge,%
-  languages,%
-  linguistics,%
-  meta,%
-  metrics,%
-  misc,%
-  modeling,%
-  modeltheory,%
-  reuse,%
-  rewriting,%
-  softeng,%
-  specification,%
-  ssr,%
-  technology,%
-  theory,%
-  web,%
-  upcoming,%
-  upcoming_conferences,%
-  conferences,%
-  workshops,%
-  verification,%
-  escjava,%
-  jml,%
-  nijmegen}
-
-% ======================================================================
-% Fin
+The JML-JUnit automatic test suite generation tool has not been widely used or assessed. 
+It is described by its own creators in \cite{Cheon2002,Cheon2004,Cheon2005}; it has been used in \cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test suite, and, more importantly, only against proof of concepts system, it has never be proved against real complex systems.
+A brief assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the analysis is based on injecting random generated errors in the source code, and then evaluating the ratio of the errors found by the test suite: the results shows that the JML-JUnit tool is not effective enough if used as a stand alone tool.
+No attempts on running or evaluating the generated test suite alongside with an hand made test suite are made.
 
@@
 %%% End:
-
-% =====================================================================
-\section{Data stream and protocol description}
-\label{sec:data_stream_and_protocol_description}
-
-\STSB and \STPU communicates via an Universal Serial Bus (USB)
-interface.  A proprietary protocol is used on top of the USB serial
-channels, the protocol is specified by
-FTDI\footnote{\myhref{http://www.ftdichip.com/}{FTDI web site}}; FTDI
-is the manufacturer of the USB controller chip used in the \STSB, the
-FT2232H\cite{ftdi_ft232h_2009}.
-
-The proprietary protocol enables the control and management of some of
-the functionalities in FT232H chip (such as internal EEPROM read and
-write, reset of the chip, etc.), as well as the definition of the
-communication protocol and properties to be used in the asynchronous
-channels (input and output channels) provided by FT232H.  These
-functionalities are accessed via an Application Program Interface, the
-API is provided by a driver developed directly by FTDI, which is
-available for various systems (Linux 32 and 64 bits, Mac OS, Windows)
-\cite{ftdi_d2xx_api_2009}.
-
-Finally, a custom protocol has been developed to manage the entire
-\STSB, built on top of the serial channels provided by the API.  The
-custom protocol is meant to be used to read sensor data observations,
-enable or disable sensors, configure on-board components (as the
-Field-Programmable Gate Array and the micro processor); the protocol
-main function is to read sensor data.
-
-The protocol is asynchronous and packet based, the packets have a
-fixed length.  The protocol has been divided in multiple layers to
-ease its formal description.  The rationale is the following: mixing
-up different abstraction levels of a protocol description can lead to
-overly complex and unmanageable descriptions, especially when dealing
-with formal ones.  The layers identified in protocol are the
-following:
-
-\begin{description}
-\item[Packet byte structure]: the internal representation of the
-  packet.
-\item[Packet info structure]: the meaningful fields contained in the
-  packets.
-\item[Single packet rules]: the content acceptable values, and how
-  they influence each other.
-\item[Packet sequence rules]: the content acceptable values, based on
-  values of previously received packets.
-\item[Packet input output asynchronous rules]: the content acceptable
-  values and reaction constraints on output packets, based on
-  previously transmitted input packets.
-\end{description}
-
-We will examine the output sensor data packet, since it is the most
-used and interesting one; we won't describe input or other output
-packets.  We will examine the first three layers, that is: the packet
-byte structure, the packet info structure and the single packet rules.
-
-The packet byte and info structures of the output sensor data packet
-reflect the board capabilities and built in sensors.  Essentially a
-single packet has to accommodate various types of data:
-
-\begin{itemize}
-\item Fast data rate streams: audio data, 48KHz (standard audio) or
-  96KHz (ultra audio).
-\item Medium data rate streams: 5KHz.
-\item Slow data rate streams: temperature, pression, accelerometer,
-  etc.; all these sensor have a data rate lower than 5 KHz.
-\item Metadatas: metadatas on sensors and board state.
-\end{itemize}
-
-The internal structure of the package (as we will see) is strictly
-fixed, even if the data streams have different data rates.  This
-choice originates from a trade-off between two constraints: the
-efficiency of the transmission (redundant data to be minimized) and
-the complexity to build a package (complexity to be minimized).  The
-complexity minimization constraint rationale is the following: the
-FPGA on the \STSB builds the packet, but the FPGA is also to be used
-for various other specific tasks; it is important to maintain the
-building packet code on the FPGA as simple and short as possible, to
-leave enough space to implement other tasks.  The Output sensor data
-packet structure is shown in Table \ref{tab:data_packet_structure},
-the packet total length is 1024 bytes (512 16 bit words).
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet structure.}
-  \label{tab:data_packet_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0-1 & sentinel\\\hline
-      2-7 & slow data\\\hline
-      8-11 & metadata: board state\\\hline
-      12-13 & reserved\\\hline
-      14-15 & metadata: board state\\\hline
-      16 & reserved\\\hline
-      17-22 & frame 1\\\hline
-      \ldots & \ldots\\\hline
-      503-508 & frame 82\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-The frames accommodate data from the fast and medium data rate streams
-(there are 4 fast data rate stream and 8 medium data rate stream
-channels).  There are 82 frames in a packet.  The frame structure is
-shown in Table \ref{tab:frame_structure}.
-
-\begin{table}[htbp]
-  \caption{Output sensor data packet frame structure.}
-  \label{tab:frame_structure}
-  \begin{center}
-    \begin{tabular}{|c|l|}\hline
-      \textbf{\textit{index}} & \textbf{\textit{function}}\\\hline
-      0 & metadata: frame description\\\hline
-      1 & medium data rate data sample\\\hline
-      2-5 & fast data rate data samples\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-Looking at the structure described and the data rates frequencies, it
-can be seen that there must be some subtle rule, to make all the
-figures to add up; that is, to be capable to accommodate three
-different rates outputs in a fixed structure, fixed length packet.
-The subtleties lies in two forms of redundancies, for slow and medium
-data rates:
-
-\begin{itemize}
-\item Slow data rates: slow data can be found at the beginning of each
-  packet, the data is simply repeated (that is, it is equal to the
-  previous packet in the stream) when no new data is available.
-\item Medium data rates: the 8 medium rate channels are multiplexed
-  together into the medium data rate data sample of each frame.  The
-  medium data rate sample channel is a piece of info to be extracted
-  from the frame description. Besides, the medium data rate data
-  sample could contain no data at all, this information is to be
-  extracted from the frame description. So, one frame could contain a
-  medium data rate sample from channel 0, the following frame a sample
-  from channel 5, the following frame no sample at all, etc.
-\end{itemize}
-
-The single packet rules delimit the boundaries of the values
-obtainable from the packet: each defined sensor represented in the
-packet, as well as the metadata describing the \STSB and the packet
-and frame contents, are constrained by a range of acceptable values.
-There are also rules affecting more than one value, these are
-essentially validity rules (i.e. if a specific sensor is active the
-reading value has to be in a defined range), and rules specifying a
-correct sequence of frames (i.e. there can be only one switch in
-sampling rate in the frame sequence of a single packet).

### 2010-07-01 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
paper spell checked language: British

@@
 
-\documentclass{article}
+%\documentclass{article}
+\documentclass[english]{lni}
+\usepackage{latin1}
 
 \usepackage{times}
-\usepackage{ifpdf}
+%\usepackage{ifpdf}
 \usepackage{a4wide}
@@
 These two communities do not talk to or learn from each other.
-We recognize that the problem solving approches and mind attitudes differs widely, but we strongly believe that this clash of view points can be exploited to grow new development process that actually blends, rather than fight or simply compose, practices from the two worlds.
+We recognize that the problem solving approaches and mind attitudes differ widely, but we strongly believe that this clash of view points can be exploited to grow new development process that actually blends, rather than fight or simply compose, practices from the two worlds.
 This paper summarizes a complex case study which shows that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming...
@@
 
-An innovative development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal verification practices; the development process has been succesfully applied to the case study yelding promising results.
+An innovative development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal verification practices; the development process has been successfully applied to the case study yielding promising results.
 
@@
 This is because of several reasons, the most important being a difference of mind attitudes.
-Formal methodologies favor an in-depth think first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall style of development.
+Formal methodologies favour an in-depth think first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall style of development.
 Once a formalization has been developed, several verification steps can be fully automated.
-Agile methodologies favor an higly incremental and iterative approach to cope with changing requirements and precise deadlines; in agile methodologies the problem is only partially understood, focusing on the aspects that need to be implemented in the current iteration, the solution initally developed is usually not optimal at all, but than it gets refined through continuous code refactorings.
-Test suites are used for several purposes, the most important being verification, documentation, and requirement specification, and they are striclty hand made.
+Agile methodologies favour an highly incremental and iterative approach to cope with changing requirements and precise deadlines; in agile methodologies the problem is only partially understood, focusing on the aspects that need to be implemented in the current iteration, the solution initially developed is usually not optimal at all, but than it gets refined through continuous code refactorings.
+Test suites are used for several purposes, the most important being verification, documentation, and requirement specification, and they are strictly hand made.
 
@@
 % what is needed, high level
-To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies needs to be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), need to be adapted to a higly iterative one.
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies needs to be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), need to be adapted to a highly iterative one.
 
 % what needs to be considered: TDD
-The most common software verification practices in agile methodologies need to be considered, when tring to integrate formal and agile methodologies, since they are a fundamental part of the development process.
-The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)]\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
+The most common software verification practices in agile methodologies need to be considered, when trying to integrate formal and agile methodologies, since they are a fundamental part of the development process.
+The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
 Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven apprach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is considered a good software development technique to be used on its own, regardless of the surrounfing development process.
+The test driven approach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is considered a good software development technique to be used on its own, regardless of the enclosing development process.
 It is to be noted that, depending on the abstraction level, the test driven approach can introduce activities cycles that can be as long as a complete iteration (functional tests), or as short as a couple of minutes (small increment on a unit test).
@@
 Some tentative attempts to reconcile formal and agile methodologies have been developed.
-The recurrent approach of the attempts consists in removing the hand made test suites, substituing them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
+The recurrent approach of the attempts consists in removing the hand made test suites, substituting them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
 
@@
 As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} precondition, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
+The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
 At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~/cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
@@
 The approach is simple: the verification, documentation and requirement specification purposes of the tests are moved on to the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
-The problem is that this requirement is hard to fulfill, since the effort required to write a complete formal specification of the system, in most real world complex cases, is much greater than writing a suite of tests.
+The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is much greater than writing a suite of tests.
 If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified.
@@
 % what we want to achieve
-The objectives we want to achive are the following ones: to be able to formally specify only parts os the system, in order to cope with constrained resources and unstable requirements; to draft an higly iterative and incremental development process that blends formal and agile practices.
+The objectives we want to achieve are the following ones: to be able to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
 
@@
 The development of the physical board, along with its embedded software (development tasks 2, 3, 4), is carried out by an external manufacturer, thus it is not directly taken into account here. 
-The specification of the communication protocol (development task 1) is a joint effort, while the remainang tasks (5, 6) can be developed in isolation.
-The dependencies are straighforward: the specification of the communication protocol (1) is the most stable element, but it still depends on the sensor board (2, 3, 4): big changes in the latter could affect the former.
+The specification of the communication protocol (development task 1) is a joint effort, while the remaining tasks (5, 6) can be developed in isolation.
+The dependencies are straightforward: the specification of the communication protocol (1) is the most stable element, but it still depends on the sensor board (2, 3, 4): big changes in the latter could affect the former.
 The driver and the simulators depends directly on the communication protocol.
@@
 % formal methodology constraints
-An appriate formal methodology needs to be chosen; it must be able to support a specification that is built incrementally and accompanied by tools that allow the verification of the system properties at run time.
-All the formal methods that require a complete specification to be used effectivelt must be avoided.
+An appropriate formal methodology needs to be chosen; it must be able to support a specification that is built incrementally and accompanied by tools that allow the verification of the system properties at run time.
+All the formal methods that require a complete specification to be used effectively must be avoided.
 The formal language should also be able to automatically generate tests.
@@
 % the chosen formal methodology
-Being Java the language chosen for the implementation, the Java Modeling Language (JML)~\cite{advancedJMLpaper}, accompanied by the JML Runtime Assertion Checker suite of tools~\cite{RAC}\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}} has been selected.
+Being Java the language chosen for the implementation, the Java Modeling Language (JML)~\cite{advancedJMLpaper}, accompanied by the JML Runtime Assertion Checker suite of tools~\cite{Cheon-Leavens02b}\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}} has been selected.
 JML and RAC tools provide a run time checker, a test generator, and do not require a complete specification to be used effectively.
@@
 They share a common approach on team management, customer relation, simplification and removal of not needed artifacts and activities and importance of working solutions and customers satisfaction.
-The agile manifesto~\cite{Beck2001} summarizes the philosphy and principles shared by all agile methodologies.
+The agile manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all agile methodologies.
 
@@
 % refactoring in iterative development
-Continuous refactoring~\cite{Fowler1999} enables an interative and incremental development, since the system architecture and design, whether explicitly specified or not, need to be cleaned up and modified, to accomodate new requirements, services and modules.
+Continuous refactoring~\cite{Fowler1999} enables an iterative and incremental development, since the system architecture and design, whether explicitly specified or not, need to be cleaned up and modified, to accommodate new requirements, services and modules.
 Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.
@@
 % JML introduction
-The Java Modeling Language (JML)\footnote{\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}} is the de facto standard specification language for describing the behavioral properties of Java programs.
-JML is a rich behavioral interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioral subtyping~\cite{X}.
+The Java Modeling Language (JML)~\cite{JML2}\footnote{\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}} is the de facto standard specification language for describing the behavioural properties of Java programs.
+JML is a rich behavioural interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioural subtyping~\cite{X}.
 JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants. 
 It also includes a rich model-based specification layer that includes mathematical models, data refinements~\cite{X}, datagroups~\cite{X}, and many other advanced object-oriented specification features~\cite{advancedJMLpaper}.
-Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{STTTpaper}.
+Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{BurdyEtal05-STTT}.
 
@@
 % JML tools
-The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{RAC,BurdyEtal05-STTT,ESC}.
+The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.
 The former compiles executable assertions into runtime checks.  
 The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
-As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{soundness,completeness}.
+As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{KiniryEtal06}.
 
@@
 Hand made tests have to be used together with generated ones.
-Hand made tests verify complex behaviors that are not formally speified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
+Hand made tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
 Limits in the formal verification tools need to be considered when deciding whether supporting the code with hand made tests, verification tools based on automatically generated tests~\cite{Glass2002} are not an exception.
@@
 The opposite path still needs to be covered, moving incrementally from a tested code to a formally specified code.
-In this case, new hand made tests and existing working code drives the formal specification: Test DRiven Formal specification (TDF).
+In this case, new hand made tests and existing working code drives the formal specification: Test Driven Formal specification (TDF).
 The resulting development cycle to be used to incrementally add formal specifications to an existing working code is shown in Figure \ref{fig:tdf_diagram}.
@@
 The packet byte and info structures of the output sensor data packet reflect the board capabilities and built in sensors. 
-A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metada describing the sensors and the board state.
+A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metadata describing the sensors and the board state.
 The internal structure of the packet is strictly fixed, even if the data streams have different data rates. 
@@
 % verification
-Verification is obtained through a hand made test suite, composed of unit and integration tests; the tests specifies the behaviors of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
-The implementation is tricky: it searchs for repeated patterns into a byte stream to be sufficiently confidant how the stream can be properly decomposed into packets.
+Verification is obtained through a hand made test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
+The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant how the stream can be properly decomposed into packets.
 Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
@@
 The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and hand made test suite to verify the most complex behaviour.
-It has been built starting from JML annoatations, supported by hand made tests when needed.
+It has been built starting from JML annotations, supported by hand made tests when needed.
 
@@
 Theoretically a separate simulator is needed for each protocol layer, to specify, develop and test each layer in isolation.  
-But, because of limited resources and affort avilable, the layer structure was not matched by corresponding simulators.
+But, because of limited resources and effort available, the layer structure was not matched by corresponding simulators.
 Two simulators are available, they have been built concurrently with the specification, the driver implementation, and the tests.
@@
 % a complex specification example
-The complex specifications usually focus on properties regarding the behavior of a whole object.
-In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property that has to be maintained during the life cycle of an object, more precisly, an invariant is assumed on entry and guaranteed on exit of each method of an object.
+The complex specifications usually focus on properties regarding the behaviour of a whole object.
+In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property that has to be maintained during the life cycle of an object, more precisely, an invariant is assumed on entry and guaranteed on exit of each method of an object.
 The invariant is constraining the number of samples for each medium data rate streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is \lil{FRAMES}, the constant constraining the number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are fairly distributed on the channels.
@@
 % test suites sizes
-The hand made test suite is composed by over 140 tests, while the generated one is 100 times bigger, totaling over 14000 tests.
+The hand made test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
 The big number of generated tests can be understood looking closely at how the generated tests are created by the JML framework. 
@@
 The JML framework generates $24 = 3 \times 3 \times 4$ tests, since they are the combination of the data ranges involved (the data range of the parameters and the data range of the type owning the method under test).
-Neverteless, not all the generated tests are executed.
+Nevertheless, not all the generated tests are executed.
 A generated test is not executed, unless the preconditions of the called method are satisfied. 
@@
 Regarding interfaces, an interface contains no statements, so when an interface method is called, it is the method of the implementation class used that is actually covered.
-This effect is visible in \lil{Driver} coverage figures: \lil{Driver} package contains interfaces (that are not counted at all) and some very simple real class (exceptions) that are not throughly tested.
+This effect is visible in \lil{Driver} coverage figures: \lil{Driver} package contains interfaces (that are not counted at all) and some very simple real class (exceptions) that are not thoroughly tested.
 The net result is extremely low statement code coverage figures because the very simple classes are the only ones that are actually covered.
@@
 An hand made unit test usually requires objects initialization (it can involve several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
-A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviors. 
+A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviours. 
 With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the hand made ones; but the specification is not complete, and a large enough data ranges would take long to be developed.
-The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviors, and the hand made, checking the more complex ones, whether formalized or not.
+The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the hand made, checking the more complex ones, whether formalized or not.
 
@@
 
-%% \nocite{ex1,ex2}
-\bibliographystyle{latex8} 
+%\bibliographystyle{latex8} 
+%\bibliographystyle{lniguide} 
 \bibliography{%
@@
 Hard constraints on driver delivery deadlines define the completeness level of the (\emph{just good enough}) specification.  
-In \STSB driver JML specifies very simple behaviors (method oriented) on the whole driver implementation, and more complex behavior on the driver core interfaces.  
+In \STSB driver JML specifies very simple behaviours (method oriented) on the whole driver implementation, and more complex behaviour on the driver core interfaces.  
 The software system parts not specified by the JML2 specification were specified using hand made unit tests.
@@
 % hand made unit tests purposes
-Hand made unit tests specifies the behavior not specified by formal specification; nevertheless this is not the only purpose they have.
+Hand made unit tests specifies the behaviour not specified by formal specification; nevertheless this is not the only purpose they have.
 Hand made unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
@@
 For each public method a set of tests is generated.
-The specified behavior of the method is checked using the JML2 run time assertion checker, hence, an oracle is not needed.  
+The specified behaviour of the method is checked using the JML2 run time assertion checker, hence, an oracle is not needed.  
 The generated tests send messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation exceptions; if a precondition is not met the test is not run, if the precondition is met the test is run and if the postcondition is not met an exception is fired and the test fails.
@@
 % tests are helful
-The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the hand made one was helpful to verify the behavior of complex methods calls on objects initialized to specific states.
+The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the hand made one was helpful to verify the behaviour of complex methods calls on objects initialized to specific states.
 

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
some updates

@@
-%
-% "The UCD CASL SenseTile System", for IEEE Pervasive Computing Joseph
-% R. Kiniry, Vieri del Bianco, Dragan Stosic $Id: paper.tex 2795
-% 2007-09-23 19:35:53Z dmz
-%
-
 %\documentclass{article}
@@
 
-%% \input{kt}
-
-% =====================================================================
-
 \begin{document}
@@
 
-Agile and formal people do not amalgamate nicely.
+Agile and formal people do not blend nicely. 
 These two communities do not talk to or learn from each other. 
-We recognize that the problem solving approaches and mind attitudes differ widely, but we strongly believe that this clash of view points can be exploited to grow new development process that actually blends, rather than fight or simply compose, practices from the two worlds.
-This paper summarizes a complex case study which shows that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming...
-that is, a ``Mole'' of Software Verification Practices.
+The problem-solving approaches and mind attitudes differ widely, but this clash of view points can be exploited into new development processes that actually blend, rather than fight, or simply compose, practices from the two worlds.
+This paper summarizes a complex case study, showing that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming: a ``Mole'' of software verification practices.
 
-A complex case study was used as a test bench of the development process.
+A complex case study is the test bench of the development process. 
 The case study focuses on the development of a driver to pilot an embedded custom board equipped with more than a dozen sensors: the communication protocol with the board is asynchronous and packet based.
-The communication is focused on controlling the sensors and reporting the sensors measurements.
+The protocol is focused on controlling the board and reporting the sensors measurements and the board status.
 
-The software driver have been developed fulfilling contrasting requirements: (1) the driver is formally specified, verified and validated; (2) a board simulator is built to match the specifications and the board behaviour; (3) if the board components is changed during development, the formal specifications and the simulator reflect the changes accordingly; (4) strict software delivery deadlines.
-The requirements build up an environment that could benefit from both an agile and a formal approach to software development and verification.
+The software driver is developed fulfilling contrasting requirements: (1) the protocol implementation needs to be thoroughly verified; (2) the protocol, at least in some of its parts, needs to be formally specified; (3) a board simulator is built to match the specifications and the board behaviour; (4) if the board components changes during development, the formal specifications and the simulator reflect the changes accordingly; (5) strict software delivery deadlines.
+The requirements build up an environment that benefits from both an agile and a formal approach to software development.
 
-An innovative development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal verification practices; the development process has been successfully applied to the case study yielding promising results.
+An innovative development process is drafted to cope with the contrasting requirements of the case study, blending agile and formal practices; the development process has been successfully applied to the case study yielding promising results.
 
@@
 
+% terminology
+In the paper the standard IEEE terminology~/cite{Jay1990} for validation and verification is used: verification is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); validation is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
+While with formal verification it is intended the formal proof on an abstract mathematical model of the system.
+
 \subsection{Formal and agile integration}
@@
 
-% what is needed, high level
+% what is needed, high-level
 To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies needs to be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), need to be adapted to a highly iterative one.
@@
 Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven approach is a cornerstone of XP, it is so popular that have been adopted in other agile methodologies as well; also, it is considered a good software development technique to be used on its own, regardless of the enclosing development process.
-It is to be noted that, depending on the abstraction level, the test driven approach can introduce activities cycles that can be as long as a complete iteration (functional tests), or as short as a couple of minutes (small increment on a unit test).
+The test driven approach is a cornerstone of XP, but it is so popular that have been adopted in many other agile methodologies as well.
+Also, it is considered a good software development technique to be used on its own, regardless of the enclosing development process.
+
+% test driven kinds
+The test driven approach relies on writing the tests before the actual implementation. 
+It can be applied in a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
+It can be applied in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
 
@@
 Some tentative attempts to reconcile formal and agile methodologies have been developed.
-The recurrent approach of the attempts consists in removing the hand made test suites, substituting them with different kinds of automated verification based on the system formal specification; the automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
+The recurrent approach of the attempts consists in removing the hand made test suites, substituting them with different kinds of automated verification based on the system formal specification. 
+The automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
 
@@
 The approach is simple: the verification, documentation and requirement specification purposes of the tests are moved on to the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
-The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is much greater than writing a suite of tests.
-If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified.
-The parts not formally specified can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are not specified or completely lost.
+The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is usually greater than writing a suite of tests~/cite{Glass2002}.
+If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified at all.
+The parts not formally specified can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are not detailed nor enforced.
 The loops and relations between the different development artifacts are the inner engine of many agile methodologies, they impose an iterative and incremental pace to the development process.
@@
 % how we want to achive it
-The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem.
+The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem in an highly iterative development process.
 We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
@@
 
-\subsection{A Real And Complex Case Study: Rapid Development In Small Scale Hardware Software Co-design}
+\subsection{A Real and Complex Case Study: Rapid Development in Small Scale Hardware Software Co-design}
 \label{subsec:a_real_and_complex_case_study}
@@
 % SenseTile introduction
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system installed at the University College Dublin in Dublin, Ireland.
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system developed at the University College Dublin in Dublin, Ireland.
 Our facility provide a scalable and extensible infrastructure for performing in-depth investigation of both the specific and general issues in large-scale sensor networking.
@@
 % case study: the sensor board development tasks
-The case study is focused in building the sensor board and its software driver side by side.
-Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board; (4) board embedded software; (5) communication protocol software driver; (6) development of software simulators of the board.
+The case study is focused in building the sensor board and its software driver concurrently.
+Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board development; (4) development of the board embedded software; (5) communication protocol software driver development; (6) development of software simulators of the board.
 
@@
 
+% communication with external manufacturer
+Communication and synchronization with the external manufacturer is frequent but not optimal.
+The main synchronization artifact is the protocol specification, that remain stable at high-level, but is changed frequently in the low-level details.
+Since domain expertise is very different, a common domain language is hard to find and to agree upon.
+Formal methodologies and programming platforms also differ.
+The informal specification of the communication protocol is the most reliable source of information.
+
 % how the problem is usually managed
@@
 co-design), but they are all focused on big-scale systems.
-The hardware software formal co-design methodologies usually have several common development artifacts\cite{Slomka2000,Hoffman2001}: an high level specification of the system, a translation (refinement) of the high level specification to low level ones (hardware and software counterparts), the possibility to generate software code from the low level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
-When considering the development of the device and of its software driver as separate entities, that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
+The hardware software formal co-design methodologies usually have several common development artifacts~\cite{Slomka2000,Hoffman2001}: an high-level specification of the system, a translation (refinement) of the high-level specification to low-level ones (hardware and software counterparts), the possibility to generate software code from the low-level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
+When considering the development of the device and of its software driver as separate entities that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design~\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
 
@@
 Agile methodologies are all based on a highly iterative and incremental process.
-They share a common approach on team management, customer relation, simplification and removal of not needed artifacts and activities and importance of working solutions and customers satisfaction.
-The agile manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all agile methodologies.
+They share a common approach on team management, customer relation, simplification and removal of unnecessary artifacts and activities, they rate working solutions and custoemrs satisfaction as the most important indicators to be considered during development.
+The Agile Manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all agile methodologies.
 
@@
 % development cycles
-The test suite are used to verify board simulators (low level), but it cannot be used to check properties at run time.
+The test suite are used to verify board simulators (low-level), but it cannot be used to check properties at run time.
 Code and tests have been implemented using TDD.
@@
 % development cycles
-The test suites are used to verify board simulators (high level), RAC and JML annotations are used to check the formally specified properties at run time.
+The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check the formally specified properties at run time.
 Code and tests have been initially implemented using TDD, than using all the Mole verification practices.
@@
 % development cycles
-The test suites are used to verify board simulators (high level), RAC and JML annotations are used to check most of the rules at run time.
+The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check most of the rules at run time.
 Code and tests have been implemented using both forms of FDD.
@@
 % simulators description
-The driver API is splitted in two abstraction layers: the general high level interfaces, that expose the main functionalities of the board and the main contents of the packets, and the lower level implementation that parse the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
+The driver API is splitted in two abstraction layers: the general high-level interfaces, that expose the main functionalities of the board and the main contents of the packets, and the lower level implementation that parse the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
 The simulators are built according to these two abstraction layers.
-The high level simulator implements the interfaces providing the methods to deal with an abstracted \SB: the simulator knows nothing
+The high-level simulator implements the interfaces providing the methods to deal with an abstracted \SB: the simulator knows nothing
 about the details of the real format of the binary data streams.
-The low level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to be able to parse or generate.
+The low-level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to be able to parse or generate.
 
@@
 
+\paragraph*{Effectiveness on board delivery}
+
+% tests and the first prototype, board side
+The first \STSB prototype builds packets with no data, but the packets are built in a way not consistent with the specification.  
+The board errors are detected by simply using the driver implementation, the \lil{StreamDriver} package.
+
+% tests and the second prototype, driver side
+The second prototype builds packets with values taken from real sensors installed on the board.  
+The only sensors that are not working are the sensors devoted to the fast rate data streams (audio sensors).  
+One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions not initially covered by the test cases, but that occasionally showed up during real use.
+One error was discovered in the low-level system driver used by the developed driver (the system driver is outside our development scope): to eliminate the error, an upgrade of the low-level driver to a new beta version was necessary.
+No other errors were detected on the driver side.
+
+% tests and the second prototype, board side
+The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
+The formal JML specifications and the run time assertion checker correctly identified these errors.
+4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, and are expected to be corrected in the following releases. 
+No other errors, except the mentioned ones, have been discovered up to now.
+
 \section{Conclusions}
 \label{sec:conclusions}
+
+% the problem
+In this paper we focus on a specific set of development problems that neither agile methodologies nor formal methodologies are completely confortable to cope with.
+The problems, we are interested in, are characterized by unstable requirements, artifacts that need to be formally specified and verified, and are constrained by deadlines that cannot be postponed.
+We propose a blend of agile and formal verification practices, enclosed by an iterative and incremental development process: the Mole development process.
+
+% Mole
+The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, to be used with formal and non formal artifacts.
+A total of four practices are presented, two of them are conservative, following closely the development practices of agile and formal methodologies, while the other pair are inoovative, blending together elements from both worlds.
+
+% Mole at work
+The practices have been applied succesfully to develop a driver for a custom embedded sensor board, equipped with board simulators and protocol verifiers. 
+The protocol and the board specifications were informal and unstable. 
+Yet a working software product was needed as soon as possible, with incrementally added functionality, and keeping it linked with the last specifications of the protocol and the board.
+Mole verification practices enabled us to keep in sink formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
+We developed hand made test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
+
+% extending Mole
+The practices are very general, and can be applied to different development languages and formal methodologies.
+One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at run time and partially automate the verification process of the formalized properties.
 
@@
 One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions that were not covered by the test cases, but that occasionally showed up during real use.
-One error have been discovered in the low level system driver that the developed driver is using (the system driver is outside our development scope); to eliminate the error an upgrade of the low level driver to a new unreleased version were necessary.
+One error have been discovered in the low-level system driver that the developed driver is using (the system driver is outside our development scope); to eliminate the error an upgrade of the low-level driver to a new unreleased version were necessary.
 No other errors were detected on the driver side.
@@
 \end{document}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% eval: 
-%%% TeX-master: t
-%%% End:

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
new abstract

@@
+%
+% "Agile Formality: A Mole of Software Verification Practices"
+% for Agile Methods+Formal Methods 2010
+% Vieri del Bianco, Joseph Kiniry, Dragan Stosic
+%
+
 %\documentclass{article}
@@
   \usepackage{epstopdf}
-  \usepackage[pdftex,bookmarks=false,a4paper=false,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Joseph R.
-Kiniry and Vieri del Bianco"]{hyperref}
+  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Vieri del Bianco and Joseph R.
+Kiniry and Dragan Stosic"]{hyperref}
 \else
@@
 
-\author{Vieri del Bianco, Joseph R. Kiniry and Dragan Stosic\\
+\author{Vieri del Bianco and Dragan Stosic\\
   UCD CASL: Complex and Adaptive Systems Laboratory and\\
   School of Computer Science and Informatics,\\
-  University College Dublin,\\
-  Belfield, Dublin 4, Ireland,\\
-  vieri.delbianco@ucd.ie, kiniry@acm.org and dragan.stosic@gmail.com\\
+  University College Dublin, Ireland\\
+  \texttt{vieri.delbianco@ucd.ie} and \texttt{dragan.stosic@gmail.com}\\
+  \\
+  Joseph R.~Kiniry \\
+  IT University of Copenhagen, Denmark\\
+  \texttt{kiniry@itu.dk}
 }
@@
 
-Agile and formal people do not blend nicely. 
-These two communities do not talk to or learn from each other. 
-The problem-solving approaches and mind attitudes differ widely, but this clash of view points can be exploited into new development processes that actually blend, rather than fight, or simply compose, practices from the two worlds.
-This paper summarizes a complex case study, showing that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming: a ``Mole'' of software verification practices.
-
-A complex case study is the test bench of the development process. 
-The case study focuses on the development of a driver to pilot an embedded custom board equipped with more than a dozen sensors: the communication protocol with the board is asynchronous and packet based.
-The protocol is focused on controlling the board and reporting the sensors measurements and the board status.
-
-The software driver is developed fulfilling contrasting requirements: (1) the protocol implementation needs to be thoroughly verified; (2) the protocol, at least in some of its parts, needs to be formally specified; (3) a board simulator is built to match the specifications and the board behaviour; (4) if the board components changes during development, the formal specifications and the simulator reflect the changes accordingly; (5) strict software delivery deadlines.
-The requirements build up an environment that benefits from both an agile and a formal approach to software development.
-
-An innovative development process is drafted to cope with the contrasting requirements of the case study, blending agile and formal practices; the development process has been successfully applied to the case study yielding promising results.
+  Members of the agile programming and formal methods communities do
+  not always see eye-to-eye.  These two communities often do not talk
+  to or learn from each other.  Only recently, as highlighted by the
+  September 2009 issue of IEEE Software, the IFIP workshop on
+  balancing agility and formalism in software engineering, and the
+  first edition of the first international workshop for formal methods
+  and agile methods, have ideas from the two communities begun to be
+  synthesized.  While the problem-solving approaches and psychological
+  attitudes of members of the two communities differ widely, we
+  exploit this clash of viewpoints, synthesizing a new development
+  processes that actually blends, rather than mashes together, best
+  practices from the two worlds.  This paper summarizes our process
+  and a supporting complex case study, showing that it is not only
+  possible, but tasty, to combine the ``chili pepper'' of formal
+  methods and the ``chocolate'' of agile programming, thus producing a
+  tasty ``mole'' of software engineering practices.
 
@@
 While with formal verification it is intended the formal proof on an abstract mathematical model of the system.
+
+A complex case study is the test bench of the development process. 
+The case study focuses on the development of a driver to pilot an embedded custom board equipped with more than a dozen sensors: the communication protocol with the board is asynchronous and packet based.
+The protocol is focused on controlling the board and reporting the sensors measurements and the board status.
+
+The software driver is developed fulfilling contrasting requirements: (1) the protocol implementation needs to be thoroughly verified; (2) the protocol, at least in some of its parts, needs to be formally specified; (3) a board simulator is built to match the specifications and the board behaviour; (4) if the board components changes during development, the formal specifications and the simulator reflect the changes accordingly; (5) strict software delivery deadlines.
+The requirements build up an environment that benefits from both an agile and a formal approach to software development.
+
+An innovative development process is drafted to cope with the contrasting requirements of the case study, blending agile and formal practices; the development process has been successfully applied to the case study yielding promising results.
 
@@
 The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
-At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~/cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
@@
 We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
-That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices.
+That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}
 

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
rewrote first two pages

@@
 % agile and formal difficult to blend: mind attitudes
-Agile and formal development methodologies usually do not blend together easily.
-This is because of several reasons, the most important being a difference of mind attitudes.
-Formal methodologies favour an in-depth think first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall style of development.
-Once a formalization has been developed, several verification steps can be fully automated.
-Agile methodologies favour an highly incremental and iterative approach to cope with changing requirements and precise deadlines; in agile methodologies the problem is only partially understood, focusing on the aspects that need to be implemented in the current iteration, the solution initially developed is usually not optimal at all, but than it gets refined through continuous code refactorings.
-Test suites are used for several purposes, the most important being verification, documentation, and requirement specification, and they are strictly hand made.
+
+Agile and formal development methodologies, like oil and water,
+usually do not blend together well.  This is because of several
+reasons, the most important of which is often characterized as a
+radical difference in psychological attitudes about software
+development.
+
+Formal methodologists often favour an in-depth, think-first approach,
+where the problem is understood, formalized, and solved; usually, but
+not always, adopting a waterfall-style of development.  Once a
+formalization is developed, development and verification proceeds, in
+an automated and interactive fashion.  Consequently, projects that use
+formal methodologies (FMs) typically focus on critical systems with
+fixed requirements and somewhat flexible deadlines (``We'll ship it
+when it is right.'')---systems in which if something goes wrong,
+someone dies or, directly or indirectly, looses a very large amount of
+money.
+
+Agile methodologists favour an highly incremental and iterative
+approach, specifically tailored to cope with changing requirements and
+precise deadlines.  In projects that use agile methodologies (AMs), it is
+often the case that the problem is only partially understood.  Thus,
+one focuses on the aspects of a system that must be implemented in the
+current development iteration.  The solution initially developed is
+usually not optimal at all, but as it is refined through continuous
+code refactorings, its quality and validity improve over time.  Test
+suites are used for several purposes, the most important of which are
+system verification, system documentation, and requirement
+specification, and they are strictly handmade.  AMs are typically used
+in development settings where changing requirements and rapid delivery
+of the product are paramount.
 
 % problems would benefit form agile and formal mix, characterize the problem
-The two worlds seem irreconcilable, nevertheless, there exist problems that would greatly benefit from both~\cite{Black2009}.
-The problems' class we are interested in need to have unstable requirements, to be constrained by deadlines that cannot be postponed, but also to be characterized by artifacts that need to be formally specified and verified.
+
+The two worlds seem irreconcilable.  Nevertheless, there exist
+problems that would greatly benefit from both~\cite{Black2009}.  The
+problem class that in which we have particular interest have
+\emph{unstable requirements}, are \emph{constrained by deadlines that
+  cannot be postponed}, but also \emph{have subsystems that must be
+  formally specified and verified}.
 
 % terminology
-In the paper the standard IEEE terminology~/cite{Jay1990} for validation and verification is used: verification is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); validation is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
-While with formal verification it is intended the formal proof on an abstract mathematical model of the system.
 
-A complex case study is the test bench of the development process. 
-The case study focuses on the development of a driver to pilot an embedded custom board equipped with more than a dozen sensors: the communication protocol with the board is asynchronous and packet based.
-The protocol is focused on controlling the board and reporting the sensors measurements and the board status.
+\subsection{Terminology}
+\label{sec:terminology}
 
-The software driver is developed fulfilling contrasting requirements: (1) the protocol implementation needs to be thoroughly verified; (2) the protocol, at least in some of its parts, needs to be formally specified; (3) a board simulator is built to match the specifications and the board behaviour; (4) if the board components changes during development, the formal specifications and the simulator reflect the changes accordingly; (5) strict software delivery deadlines.
-The requirements build up an environment that benefits from both an agile and a formal approach to software development.
+In this paper the standard IEEE terminology for validation and
+verification is used~/cite{Jay1990}.  Paraphrasing the standard
+slightly, \emph{verification} is the process of evaluating software to
+determine whether the products of a given development phase satisfy
+the conditions imposed at the start of that phase, according to the
+requirements and design specifications (``you build the system in the
+right way''); validation is the process of evaluating software during
+or at the end of the development process to determine whether it
+satisfies specified requirements ensuring it meets the user's needs
+(``you build the right system'').
 
-An innovative development process is drafted to cope with the contrasting requirements of the case study, blending agile and formal practices; the development process has been successfully applied to the case study yielding promising results.
+Within the formal methods community, verification means something much
+stronger, and validation something quite different, than in the IEEE
+610 standard.  Loosely, within the FM community, \emph{verification}
+means the use of formal methods to formally, statically (and usually
+modularly) verify that a system conforms to its specification (under
+all possible execution scenarios).  On the other hand,
+\emph{validation} means the use of traditional software engineering
+practices to check, statically or dynamically, that the system
+conforms to its specification (whether in the form of tests,
+requirements, etc.) by hand or via execution under some execution
+scenarios.
 
-\subsection{Formal and agile integration}
+As we come from both the AM and FM communities, we respect and
+acknowledge both perspectives and will attempt to be absolutely clear
+in our use of these terms in the following.
+
+\subsection{Case Study}
+\label{sec:case-study}
+
+Proposing a new process with no supporting evidence is an
+uninteresting proposition.  A complex case study is mandatory to test
+bench a development process.  Our case study focuses on the
+development of a device driver to control and communicate with an
+embedded custom circuit board equipped with more than a dozen sensors.
+The communication protocol with the board is asynchronous and packet
+based and the board is novel and custom-designed.  The protocol is
+focused on controlling the board and reporting sensor measurements and
+board status.
+
+This software driver has been developed using our process while
+statisfying seemingly contradictory requirements: (1) the protocol
+implementation must be thoroughly verified; (2) parts of the protocol
+must be formally specified; (3) a board simulator must be designed,
+implemented, and verified to match the specifications of, and actual
+behavior of, the physical circuit board; (4) if board components
+change during development, the associated formal specifications and
+the simulator whose behavior reflects such changes must be updated
+accordingly; (5) strict software delivery deadlines.  Consequently,
+these process requirements constitute an environment that benefits
+from both an agile and a formal approaches to software development.
+
+Over the course of the past 18 months, an novel development process
+has been refined to cope with the contrasting requirements of the case
+study, blending agile and formal practices.  In the following we
+summarize this development process and report on its successful
+application to the case study, yielding several promising results.
+
+\subsection{Formal and Agile Integration}
 \label{subsec:formal_and_agile_integration}
@@
 Some tentative attempts to reconcile formal and agile methodologies have been developed.
-The recurrent approach of the attempts consists in removing the hand made test suites, substituting them with different kinds of automated verification based on the system formal specification. 
+The recurrent approach of the attempts consists in removing the handmade test suites, substituting them with different kinds of automated verification based on the system formal specification. 
 The automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
@@
 % previous work: a common approach
-As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
+As already mentioned, in all these attempts, the handmade test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
 The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
@@
 The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem in an highly iterative development process.
-We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
+We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how handmade and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
 That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}
@@
 
-\subsection{A Real and Complex Case Study: Rapid Development in Small Scale Hardware Software Co-design}
+\subsection{A Real and Complex Case Study: \\ Rapid Development in Small-scale Hardware Software Co-design}
 \label{subsec:a_real_and_complex_case_study}
@@
 The objective is to formally specify only parts of the system under development, in an incremental iterative process.
-Formal methodologies artifacts need to coexist with traditional agile development artifacts: requirements, expressed as functional tests or as formal specifications, hand made and generated tests, documentation, source code.
+Formal methodologies artifacts need to coexist with traditional agile development artifacts: requirements, expressed as functional tests or as formal specifications, handmade and generated tests, documentation, source code.
 The development process needs to guarantee consistency all over the involved artifacts.
@@
 
-% FDD with hand made tests
+% FDD with handmade tests
 A partially specified module cannot be verified properly only through automatically generated tests.
-Hand made tests have to be used together with generated ones.
-Hand made tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
-Limits in the formal verification tools need to be considered when deciding whether supporting the code with hand made tests, verification tools based on automatically generated tests~\cite{Glass2002} are not an exception.
-The development cycle to be used to blend formal specifications, hand made tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the hand made tests, both hand made tests and the formal specification drive the code development.
+handmade tests have to be used together with generated ones.
+handmade tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
+Limits in the formal verification tools need to be considered when deciding whether supporting the code with handmade tests, verification tools based on automatically generated tests~\cite{Glass2002} are not an exception.
+The development cycle to be used to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests, both handmade tests and the formal specification drive the code development.
 
@@
     \includegraphics[scale=0.5]{diagrams/FDD_hand_made_tests}
-    \caption{FDD development cycle with hand made tests.}
+    \caption{FDD development cycle with handmade tests.}
     \label{fig:fdd_hand_made_tests_diagram}
@@
 The opposite path still needs to be covered, moving incrementally from a tested code to a formally specified code.
-In this case, new hand made tests and existing working code drives the formal specification: Test Driven Formal specification (TDF).
+In this case, new handmade tests and existing working code drives the formal specification: Test Driven Formal specification (TDF).
 The resulting development cycle to be used to incrementally add formal specifications to an existing working code is shown in Figure \ref{fig:tdf_diagram}.
@@
 
-\todo{if time available: result: hand made test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
+\todo{if time available: result: handmade test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
 
@@
 % verification
-Verification is obtained through a hand made test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
+Verification is obtained through a handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
 The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant how the stream can be properly decomposed into packets.
@@
 % verification
-Verification is provided by hand made test suite, JML annotations and generated unit test suite. 
-It has been built starting from hand made tests only (TDD development cycles).
-Later on JML annotations have been added incrementally, using a mix of mole practices: to move from hand made tests to formal specifications, TDF have been initially used; than all the mole practices have been used according to the artifacts involved.
+Verification is provided by handmade test suite, JML annotations and generated unit test suite. 
+It has been built starting from handmade tests only (TDD development cycles).
+Later on JML annotations have been added incrementally, using a mix of mole practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the mole practices have been used according to the artifacts involved.
 
@@
 % verification
-The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and hand made test suite to verify the most complex behaviour.
-It has been built starting from JML annotations, supported by hand made tests when needed.
+The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and handmade test suite to verify the most complex behaviour.
+It has been built starting from JML annotations, supported by handmade tests when needed.
 
@@
 % unit tests kinds
-The unit test cases that verify the protocol driver are of two kinds: hand made unit tests and automatically generated unit tests based on JML specifications; they are built to be used together, they are complementary.
+The unit test cases that verify the protocol driver are of two kinds: handmade unit tests and automatically generated unit tests based on JML specifications; they are built to be used together, they are complementary.
 
@@
 
-\paragraph*{Hand made tests}
+\paragraph*{handmade tests}
 
@@
 % simulator roles in tests
-The simulators are used in the hand made test suits as stubs.
+The simulators are used in the handmade test suits as stubs.
 The resulting structure is shown in Figure \ref{fig:class_diagram_streamdriver_test}.  
@@
 The use of simulators to enable unit testing was essential to test the \STSB driver, since the physical board was not available.
-The simulators are used in both the hand made and the generated test suite, since both need a proper set of stubs to be executed.
+The simulators are used in both the handmade and the generated test suite, since both need a proper set of stubs to be executed.
 
@@
 % test suites sizes
-The hand made test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
+The handmade test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
 The big number of generated tests can be understood looking closely at how the generated tests are created by the JML framework. 
@@
 % test execution time, values
-Test suites execution time are different, the hand made test suite, with the run time assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $).
+Test suites execution time are different, the handmade test suite, with the run time assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $).
 Most of the time is spent on setting up the suite, only considering the real execution of the tests, with no suite setup, the execution time reduces to less than 4 minutes (the rough ratio is $ 1 : 10 $).
@@
 The statement code coverage figures show that the test suites do a good job on the most important package, the \lil{StreamDriver}. 
-The generated test suite reaches almost $ 80 \% $, while the hand made tests have only a slightly lower coverage ratio: $ 76 \% $. 
+The generated test suite reaches almost $ 80 \% $, while the handmade tests have only a slightly lower coverage ratio: $ 76 \% $. 
 The combination of test suites raise the coverage ratio to $ 88.5 \% $. 
@@
 
-% hand made unit test vs generated unit test
+% handmade unit test vs generated unit test
 The qualitative difference of the two test suites can be understood observing how the suites are built. 
-An hand made unit test usually requires objects initialization (it can involve several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
+An handmade unit test usually requires objects initialization (it can involve several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
 A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviours. 
-With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the hand made ones; but the specification is not complete, and a large enough data ranges would take long to be developed.
-The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the hand made, checking the more complex ones, whether formalized or not.
+With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges would take long to be developed.
+The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.
 
@@
 Mole verification practices enabled us to keep in sink formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
-We developed hand made test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
+We developed handmade test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
 
@@
 In \STSB driver JML specifies very simple behaviours (method oriented) on the whole driver implementation, and more complex behaviour on the driver core interfaces.  
-The software system parts not specified by the JML2 specification were specified using hand made unit tests.
+The software system parts not specified by the JML2 specification were specified using handmade unit tests.
 
-\paragraph*{hand made unit tests}
+\paragraph*{handmade unit tests}
 
-% hand made unit tests purposes
-Hand made unit tests specifies the behaviour not specified by formal specification; nevertheless this is not the only purpose they have.
-Hand made unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
+% handmade unit tests purposes
+handmade unit tests specifies the behaviour not specified by formal specification; nevertheless this is not the only purpose they have.
+handmade unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
 followed during development.
 
-% hand made unit tests suite structure
-The hand made test suite primarily focuses on small features and use cases fragments that the system must implement, or for which an exceptional case has been specified. 
+% handmade unit tests suite structure
+The handmade test suite primarily focuses on small features and use cases fragments that the system must implement, or for which an exceptional case has been specified. 
 A test case is to be considered a rough specification written in imperative style. 
@@
 % junit
-In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the framework used to manage the implementation and execution of hand made and generated unit tests, there is no clear distinction between drivers, stubs, and oracles; none-less, even if a proper distinction
+In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the framework used to manage the implementation and execution of handmade and generated unit tests, there is no clear distinction between drivers, stubs, and oracles; none-less, even if a proper distinction
 is not imposed by the framework, the conceptual distinction still holds.
 
-% who provides what: hand made vs generated
-Hand made unit tests and JML generated tests differ in how test components are provided.
+% who provides what: handmade vs generated
+handmade unit tests and JML generated tests differ in how test components are provided.
 The distinctions and similarities are summarized in Table \ref{tab:test_harness}.
 \begin{table}[htbp]
-  \caption{Hand made and generated tests harness.}
+  \caption{handmade and generated tests harness.}
   \label{tab:test_harness}
@@
     \begin{tabular}{|c|c|c|c|}\hline
-       & \textbf{\textit{hand made}} & \textbf{\textit{hand made + RAC}} & \textbf{\textit{generated}} \\\hline
+       & \textbf{\textit{handmade}} & \textbf{\textit{handmade + RAC}} & \textbf{\textit{generated}} \\\hline
       \textbf{\textit{Stub}} & manual & manual & manual\\\hline
@@
 % tests are helful
-The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the hand made one was helpful to verify the behaviour of complex methods calls on objects initialized to specific states.
+The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the handmade one was helpful to verify the behaviour of complex methods calls on objects initialized to specific states.
 
 % tests leads to a better design
-An hand made test suite force the code to be test oriented, which usually leads to a better quality and more modularized source code\cite{Binder1999}.
+An handmade test suite force the code to be test oriented, which usually leads to a better quality and more modularized source code\cite{Binder1999}.
 
@@
 The approach is in waterfall style; than the need to develop the protocol, the specification and the implementation concurrently is not addressed.  
-No hand made or generated test suites are used, the code is specified with JML assertions, refined by the state machine specifications, and then checked statically with ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2 tools RAC\cite{BurdyEtal05-STTT}.
+No handmade or generated test suites are used, the code is specified with JML assertions, refined by the state machine specifications, and then checked statically with ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2 tools RAC\cite{BurdyEtal05-STTT}.
 
@@
 A brief assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the analysis is based on injecting random generated errors in the source code, and then evaluating the ratio of the errors found by the test suite: the results shows that the JML-JUnit tool is not effective enough if used as a stand alone tool.
-No attempts on running or evaluating the generated test suite alongside with an hand made test suite are made.
+No attempts on running or evaluating the generated test suite alongside with an handmade test suite are made.
 

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
merged paper modified figure included many suggestions

@@
-%
-% "Agile Formality: A Mole of Software Verification Practices"
-% for Agile Methods+Formal Methods 2010
-% Vieri del Bianco, Joseph Kiniry, Dragan Stosic
-%
-
 %\documentclass{article}
@@
 % verification practices in agile methodologies
-The agile manifesto does not introduce any specific techniques to be used in the development process, it describes agile methodologies from a very high abstraction level.
-Nevertheless most of the agile methodologies share a similar approach to artifacts verification.
+The Agile Manifesto does not suggest any specific development techniques, it describes agile methodologies from a very high abstraction level.
+Nevertheless, most agile methodologies share a similar approach to artifact verification.
 Functional test suites, to be used as precise requirements or user stories definition, are used in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
@@
 % abstraction in test suites
-High level test suites, composed of functional tests, describe requirements and bind together requirements documentation and system behaviour; the associated development cycle is one iteration long. 
-Low level test suites, composed of unit tests, describe the modules behaviour and bind together code documentation, code and the contract imposed on the module; the associated development cycle is usually less than one day (it can be as short as a few minutes).
+High level test suites, composed of functional tests, describe requirements and bind together requirements documentationand system behaviour: an informal requirement is supported by a set of functional tests.
+The associated development cycle is one iteration long. 
+Low level test suites, composed of unit tests, describe the modules behaviour and bind together code documentation, code and the contract imposed on the module: unit tests represent the contract and the documentation.
+The associated development cycle is less than one day (as short as a few minutes).
 
 % refactoring in iterative development
-Continuous refactoring~\cite{Fowler1999} enables an iterative and incremental development, since the system architecture and design, whether explicitly specified or not, need to be cleaned up and modified, to accommodate new requirements, services and modules.
+Continuous refactoring~\cite{Fowler1999} enables iterative and incremental development, since the system architecture and design, whether explicitly specified or not, must be cleaned up and modified, to accommodate new requirements, services and modules.
 Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.
-The automated test suites guarantees that the functionalities and behaviour already implemented in the code are not compromised by the system refactorings.
+Automated test suites guarantees that the functionalities and behaviour already implemented in the code are not compromised by refactorings.
 
@@
 % objectives, repeated
-The objective is to formally specify only parts of the system under development, in an incremental iterative process.
-Formal methodologies artifacts need to coexist with traditional agile development artifacts: requirements, expressed as functional tests or as formal specifications, handmade and generated tests, documentation, source code.
-The development process needs to guarantee consistency all over the involved artifacts.
+Our objective is to formally specify only parts of the system under development, in an incremental iterative process.
+Formal artifacts must coexist with informal ones: requirements, expressed as functional tests or as formal specifications, handmade and generated tests, documentation, source code.
+The development process needs to guarantee consistency over all the involved artifacts.
 
 % composition of the two worlds
-Both Test Driven Development (TDD) and classical Formal specification driven development (FDD) guarantee consistency on all the involved artifacts. 
-TDD force an iterative development, whether FDD support it without forcing it.
+Both Test Driven Development (TDD) and classical Formal specification Driven Development (FDD) guarantee consistency on all the involved artifacts. 
+Both TDD and FDD support iterative development.
 The development cycle adopted in TDD is shown in Figure \ref{fig:tdd_diagram} and the one adopted in classic FDD is shown in Figure \ref{fig:fdd_diagram}.
-It can be seen how unit tests in TDD are substituted in FDD by formal specifications supported by the verification environment, in our case the verification environment is the generated unit test suite.
+Unit tests, in TDD, are substituted, in FDD, by formal specifications supported by the verification environment, in our case the verification environment is the generated unit test suite.
 Formal specifications replace unit tests: they specify, document and verify the system under development.
@@
 % composition is not enough
+Is this enough to deal with formal and informal artifacts?
 FDD is used for the formal specified parts of the system, whether TDD is used for the remaining ones: a ``composition'' of formal and agile verification practices.
-Yet, it's not enough.
+Yet, it is not enough.
 There are several problem that still needs to be addressed.
-No communication between the two worlds is permitted, it's not defined how to formally specify a module after it was implemented and verified in a non formal way.
+No communication between the two worlds is permitted, it is not defined how to formally specify a module after it was implemented and verified in a non formal way.
 A module only partially specified is a sort of hybrid that needs to be treated accordingly.
-Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is usually not the case.
-Identify the correct formal specification is a difficult task, that needs to be supported and verified by the development process.
+Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is not always straighforward.
+Identify the correct formal specification is a difficult task that needs to be supported and verified by the development process.
 
@@
 A partially specified module cannot be verified properly only through automatically generated tests.
-handmade tests have to be used together with generated ones.
-handmade tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools because of their inherent limits.
-Limits in the formal verification tools need to be considered when deciding whether supporting the code with handmade tests, verification tools based on automatically generated tests~\cite{Glass2002} are not an exception.
-The development cycle to be used to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests, both handmade tests and the formal specification drive the code development.
+Handmade tests have to be used together with generated ones.
+Handmade tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools, because of the limits of the verification tools themselves.
+The constraints imposed by formal verification tools~\cite{Glass2002} must be taken into account when deciding whether supporting the code with handmade tests, and verification tools based on automatically generated tests are not an exception.
+The development cycle able to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests, both handmade tests and the formal specification drive the code development.
 
@@
 The opposite path still needs to be covered, moving incrementally from a tested code to a formally specified code.
-In this case, new handmade tests and existing working code drives the formal specification: Test Driven Formal specification (TDF).
+In this case, new handmade tests and existing working code drives the formal specification, this is what we call Test Driven Formal specification (TDF).
 The resulting development cycle to be used to incrementally add formal specifications to an existing working code is shown in Figure \ref{fig:tdf_diagram}.
@@
 % all together belnds into a mole
-When the Mole verification practices are used together the initial objectives can be achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; complete freedom to decide what needs to be formally specified; the development cycles and the associated verification practice maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
+When the Mole verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
 
@@
 % handmade unit tests purposes
-handmade unit tests specifies the behaviour not specified by formal specification; nevertheless this is not the only purpose they have.
-handmade unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
+Handmade unit tests specifies the behaviour not specified by formal specification; nevertheless, this is not the only purpose they have.
+Handmade unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
 followed during development.

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
rewrote remaining part of the introduction

@@
 
+Additionally, in the following when we informally say ``...development
+of a system,'' we more precisely mean ``...analysis, design,
+development, verification (both testing and formal verification), and
+validation of a system.''
+
 \subsection{Case Study}
@@
 This software driver has been developed using our process while
-statisfying seemingly contradictory requirements: (1) the protocol
-implementation must be thoroughly verified; (2) parts of the protocol
-must be formally specified; (3) a board simulator must be designed,
+statisfying seemingly contradictory requirements: (1)~the protocol
+implementation must be thoroughly verified; (2)~parts of the protocol
+must be formally specified; (3)~a board simulator must be designed,
 implemented, and verified to match the specifications of, and actual
-behavior of, the physical circuit board; (4) if board components
+behavior of, the physical circuit board; (4)~if board components
 change during development, the associated formal specifications and
 the simulator whose behavior reflects such changes must be updated
-accordingly; (5) strict software delivery deadlines.  Consequently,
+accordingly; (5)~strict software delivery deadlines.  Consequently,
 these process requirements constitute an environment that benefits
@@
 
-Mole practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and verified.
-The case study fits these needs.
-
-\subsection{A Real and Complex Case Study: \\ Rapid Development in Small-scale Hardware Software Co-design}
+\section{A Real and Complex Case Study: \\ Rapid Development in Small-scale Hardware Software Co-design}
 \label{subsec:a_real_and_complex_case_study}
 
+Mole practices are applied in the context of a real world case study
+that matches the problem requirements previously detailed: unstable
+requirements and development artifacts that need to be formally
+specified and verified.  The case study fits these needs.
+
+\subsection{The UCD CASL SenseTile System}
+\label{sec:ucd-casl-sensetile}
+
 % SenseTile introduction
-The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system developed at the University College Dublin in Dublin, Ireland.
-Our facility provide a scalable and extensible infrastructure for performing in-depth investigation of both the specific and general issues in large-scale sensor networking.
-This system integrates a sensor platform, a datastore, and a compute farm.  
-The sensor platform is a custom-designed, but inexpensive, sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.
-There are over a dozen sensors packaged on the \ST itself, the board is expandable as well, new sensors can be added to it easily.
+The \textbf{UCD CASL SenseTile System} is a large-scale,
+general-purpose sensor system developed at the University College
+Dublin in Dublin, Ireland.  Our facility provides a scalable and
+extensible infrastructure for performing in-depth investigation into
+both the specific and general issues in large-scale sensor networking.
+This system integrates a sensor platform, a datastore, and a compute
+farm.  The sensor platform is a custom-designed but inexpensive sensor
+platform (called the \ST) paired with general-purpose small-scale
+compute nodes, including everything from low-power PDAs to powerful
+touchscreen-based portable computers.  Besides containing over a dozen
+sensors packaged on the \ST itself, the board is expandable as well,
+as new sensors are added to it easily.
 
 % case study: the sensor board development tasks
-The case study is focused in building the sensor board and its software driver concurrently.
-Because of hard time constraints, and the initial unavailability of the custom board, we need to progress concurrently with all the development tasks, which consist of: (1) specification of the communication protocol; (2) physical board specification; (3) physical board development; (4) development of the board embedded software; (5) communication protocol software driver development; (6) development of software simulators of the board.
+The case study is focused in building the sensor board and its
+software driver concurrently.  Because of hard time constraints and
+the initial unavailability of the custom board, we must progress
+concurrently with all the development tasks, including:
+(1)~specification of the communication protocol; (2)~specification of
+the physical sensor board; (3)~development and fabrication of the
+physical board; (4)~development of the embedded software for the
+board; (5)~development of the communication protocol software driver
+the the board; and (6)~development of a software simulator of the
+board.
 
 % what we have to do, what we outsourced, and the dependencies between tasks
-The development of the physical board, along with its embedded software (development tasks 2, 3, 4), is carried out by an external manufacturer, thus it is not directly taken into account here. 
-The specification of the communication protocol (development task 1) is a joint effort, while the remaining tasks (5, 6) can be developed in isolation.
-The dependencies are straightforward: the specification of the communication protocol (1) is the most stable element, but it still depends on the sensor board (2, 3, 4): big changes in the latter could affect the former.
-The driver and the simulators depends directly on the communication protocol.
+The development of the physical board, along with its embedded
+software (development tasks~2, 3, 4), is carried out by an third party
+under our guidance, thus it is not directly taken into account here.
+The specification of the communication protocol (development task~1)
+is a joint effort between ourselves and the third party, while the
+remaining tasks~(5, 6) are performed in isolation.  Dependencies are
+straightforward: the specification of the communication protocol~(1)
+is the most stable element, but it still depends on the sensor
+board~(2, 3, 4): large changes in the latter affect the former.
+Additionally, the driver and the simulator depends directly on the
+communication protocol.
 
 % communication with external manufacturer
-Communication and synchronization with the external manufacturer is frequent but not optimal.
-The main synchronization artifact is the protocol specification, that remain stable at high-level, but is changed frequently in the low-level details.
-Since domain expertise is very different, a common domain language is hard to find and to agree upon.
-Formal methodologies and programming platforms also differ.
-The informal specification of the communication protocol is the most reliable source of information.
+Communication and synchronization with the external manufacturer is
+frequent but not optimal.  The main synchronization artifact is the
+protocol specification, which remains stable at high-level, but is
+changed frequently in the low-level details.  Since domain expertise
+is very different between the two teams, a common domain language is
+difficult to find and to agree upon.  Formal methodologies and
+programming platforms also differ.  Consequently, more for social and
+psychological reasons than technical ones, the informal specification
+of the communication protocol is the most reliable source of
+information.
+
+\subsubsection{Related Methodologies}
+\label{sec:relat-meth}
 
 % how the problem is usually managed
-There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software
-co-design), but they are all focused on big-scale systems.
-The hardware software formal co-design methodologies usually have several common development artifacts~\cite{Slomka2000,Hoffman2001}: an high-level specification of the system, a translation (refinement) of the high-level specification to low-level ones (hardware and software counterparts), the possibility to generate software code from the low-level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
-When considering the development of the device and of its software driver as separate entities that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware software co-design~\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}: they all focus on a specification that aims to be as complete as possible.
+There exist various approaches in literature addressing this kind of
+development constraints (rapid development in hardware software
+co-design), but all are focused on large-scale systems.
+
+The hardware-software formal co-design methodologies usually have
+several common development artifacts~\cite{Slomka2000, Hoffman2001}
+including a high-level specification of the system, a translation
+(refinement) of the high-level specification to low-level ones
+(hardware and software counterparts), the possibility to generate
+software code from the low-level specifications, a hardware simulator
+capable of simulating an hardware component based on hardware
+specification, the hardware component developed.
+
+When considering the development of the device and of its software
+driver as separate entities that possibly have to be developed
+concurrently, the existing approaches are similar to the ones seen in
+the case of hardware-software co-design~\cite{Valderrama1995,
+  Siegmund2002, Ryzhyk2009}, all of which focus on a specification
+that aims to be as complete.
 
 % complete specification is not feasible
-A specification as complete as possible is not feasible nor convenient in this case: the protocol specification have to be enhanced incrementally, and its complete specification cannot be provided.
-This makes the problem an ideal candidate to test the Mole practices.
+A complete specification is not feasible nor convenient in our case.
+The protocol specification must be enhanced incrementally and its
+complete specification cannot be provided.  This makes the problem an
+ideal candidate to test the our ``mole'' of practices.
 
 % formal methodology constraints
-An appropriate formal methodology needs to be chosen; it must be able to support a specification that is built incrementally and accompanied by tools that allow the verification of the system properties at run time.
-All the formal methods that require a complete specification to be used effectively must be avoided.
-The formal language should also be able to automatically generate tests.
+To perform this case study, an appropriate formal methodology must be
+chosen.  It must be able to support a specification that is built
+incrementally.  Consequently, formal methods that demand a complete
+specification to be used effectively must be avoided.  Additionally,
+to best synthesize with agile methods, the formal language must be
+able to automatically generate tests.  Finally, the formal method must
+be complemented by tools that support the verification of system
+properties both at design-time (via static analysis, for full
+functional verification of critical subsystems) and at run time (for
+dynamic validation for integration with TDD).
 
 % the chosen formal methodology
-Being Java the language chosen for the implementation, the Java Modeling Language (JML)~\cite{advancedJMLpaper}, accompanied by the JML Runtime Assertion Checker suite of tools~\cite{Cheon-Leavens02b}\footnote{\myhref{http://sourceforge.net/projects/jmlspecs/}{Common JML tools}} has been selected.
-JML and RAC tools provide a run time checker, a test generator, and do not require a complete specification to be used effectively.
+Driven by these requirements and our expertise, the Java programming
+language and platform were chosen for the implementation, the Java
+Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} and Business
+Object Notation (BON)~\cite{WaldenNerson95} are used to write formal
+specifications, and the BONc~\cite{BONc}, Beetlz~\cite{Beetlz},
+JML2~\cite{JML2,JML2common}, and Extended Static Checker for Java
+(ESC/Java2)~\cite{ESCJava2,KiniryCok04SmallCite} tool suites are used
+as supporting tools~\cite{BurdyEtal05-STTT}.  The JML2 tool suite
+includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit
+test generator~\cite{Cheon2002}.  The BONc tool supports validity
+checking of high-level requirements and architecture specifications,
+Beetlz supports reasoning about refinement between BON and JML
+specifications, and ESC/Java2 supports automatic, modular, static
+reasoning about the correctness of an implementation against its
+formal specification.  All four tool suites do not require a complete
+specification to be used effectively.
 
@@
 % JML introduction
-The Java Modeling Language (JML)~\cite{JML2}\footnote{\myhref{http://www.cs.ucf.edu/~leavens/JML/}{The Java Modeling Language (JML)}} is the de facto standard specification language for describing the behavioural properties of Java programs.
-JML is a rich behavioural interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioural subtyping~\cite{X}.
-JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants. 
-It also includes a rich model-based specification layer that includes mathematical models, data refinements~\cite{X}, datagroups~\cite{X}, and many other advanced object-oriented specification features~\cite{advancedJMLpaper}.
-Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{BurdyEtal05-STTT}.
+The Java Modeling Language (JML) is the de facto standard
+specification language for describing the behavioural properties of
+Java programs.  JML is a rich behavioural interface specification
+language (BISL) and thus focuses on the modular specification of
+classes and objects in which inheritance is used solely for
+behavioural subtyping~\cite{LeavensBakerRuby99}.  JML includes
+standard specification constructs like assumptions, assertions,
+axioms, and pre- and postconditions, framing clauses, and invariants.
+It also includes a rich model-based specification layer that includes
+mathematical models, data refinements~\cite{}, datagroups~\cite{}, and
+many other advanced object-oriented specification
+features~\cite{chalin06:_beyon_asser}.  Many tools, ranging from
+compilers to static checkers to full-functional verification systems
+support the JML language~\cite{BurdyEtal05-STTT}.
 
 % using JML
-JML is sometimes used in a Design-by-Contract style, where a specification is written from scratch, reusing existing APIs that have specifications of their own, and then an implementation is written conforming to that specification~\cite{DBCjml,DBC}.
-The implementation is checked by whatever means appropriate using some subset of the aforementioned tools.  
-At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process called Contracting-the-Design.
+JML is sometimes used in a Design-by-Contract style, where a
+specification is written from scratch, reusing existing APIs that have
+specifications of their own, and then an implementation is written
+conforming to that specification~\cite{Meyer92b}.  The
+implementation is checked by whatever means appropriate using some
+subset of the aforementioned tools.  At other times an existing piece
+of implemented and tested code is annotated with specifications
+after-the-fact, a process called Contracting-the-Design.
 
 % JML tools
-The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.
-The former compiles executable assertions into runtime checks.  
-The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
-As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{KiniryEtal06}.
+The two tools used most frequently to check the correctness of
+implementations are the JML Runtime Assertion Checker (RAC) and the
+Extended Static Checker for Java, version 2
+(ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.  The
+former compiles executable assertions into runtime checks.  The latter
+tool performs automated modular semantic static analysis to attempt to
+prove that the program under analysis does not misbehave (i.e., crash
+in a variety of ways) and conforms to its specification (lightweight
+functional correctness).  As ESC/Java2 is fully automated, it is not
+sound, nor does it cover the whole Java and JML languages, and is thus
+not complete, but there is support for warning the user about both of
+these situations~\cite{KiniryEtal06}.
 
@@
 %\bibliographystyle{lniguide} 
-\bibliography{%
-  extra,%
-  abbrev,%
-  ads,%
-  category,%
-  complexity,%
-  hypertext,%
-  icsr,%
-  knowledge,%
-  languages,%
-  linguistics,%
-  meta,%
-  metrics,%
-  misc,%
-  modeling,%
-  modeltheory,%
-  reuse,%
-  rewriting,%
-  softeng,%
-  specification,%
-  ssr,%
-  technology,%
-  theory,%
-  web,%
-  upcoming,%
-  upcoming_conferences,%
-  conferences,%
-  workshops,%
-  verification,%
-  escjava,%
-  jml,%
-  nijmegen}
+\bibliography{extra,abbrev,ads,category,complexity,hypertext,icsr,knowledge,languages,linguistics,meta,metrics,misc,modeling,modeltheory,reuse,rewriting,softeng,specification,ssr,technology,theory,web,upcoming,upcoming_conferences,conferences,workshops,verification,escjava,jml,nijmegen}
 

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
added small explanation of mole in abstract

@@
   methods and the ``chocolate'' of agile programming, thus producing a
-  tasty ``mole'' of software engineering practices.
+  tasty ``mole'' (as in the highly-spices Mexican sauce) of software
+  engineering practices.
 
@@
 
-Agile and formal development methodologies, like oil and water,
+Agile and formal development methodologies, like oil and water
 usually do not blend together well.  This is because of several

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
merged! some doubts

@@
 
-\title{Agile Formality: A Mole of Software Verification Practices}
+\title{Agile Formality: A Mole of Software Engineering Practices}
 
@@
 
-Agile and formal development methodologies, like oil and water
+Agile and formal development methodologies
 usually do not blend together well.  This is because of several
@@
 
-In this paper the standard IEEE terminology for validation and
-verification is used~/cite{Jay1990}.  Paraphrasing the standard
+In this paper the IEEE 610~/cite{Jay1990} standard terminology for validation and
+verification is used.  Paraphrasing the standard
 slightly, \emph{verification} is the process of evaluating software to
@@
 % what is needed, high-level
-To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies needs to be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), need to be adapted to a highly iterative one.
-
-% what needs to be considered: TDD
-The most common software verification practices in agile methodologies need to be considered, when trying to integrate formal and agile methodologies, since they are a fundamental part of the development process.
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies must be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), must be adapted to a highly iterative one.
+
+% what must be considered: TDD
+The most common software verification practices in agile methodologies must be considered, when trying to integrate formal and agile methodologies, since they are a fundamental part of the development process.
 The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
@@
 The test driven approach is a cornerstone of XP, but it is so popular that have been adopted in many other agile methodologies as well.
-Also, it is considered a good software development technique to be used on its own, regardless of the enclosing development process.
+It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
 
@@
 The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
-At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, used to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
@@
 % what we want to achieve
-The objectives we want to achieve are the following ones: to be able to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
+The objectives we want to achieve are the following ones: freedom to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
 
@@
 incrementally.  Consequently, formal methods that demand a complete
-specification to be used effectively must be avoided.  Additionally,
+specification must be avoided.  Additionally,
 to best synthesize with agile methods, the formal language must be
@@
 Agile methodologies are all based on a highly iterative and incremental process.
-They share a common approach on team management, customer relation, simplification and removal of unnecessary artifacts and activities, they rate working solutions and custoemrs satisfaction as the most important indicators to be considered during development.
+They share a common approach on team management, customer relation, simplification and removal of unnecessary artifacts and activities, they rate working solutions and customers satisfaction as the most important indicators considered during development.
 The Agile Manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all agile methodologies.
@@
 Nevertheless, most agile methodologies share a similar approach to artifact verification.
-Functional test suites, to be used as precise requirements or user stories definition, are used in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
-Unit tests and TDD are mentioned, whether not completely accepted, in most of the agile methodologies defined so far.
+Functional test suites, used as precise requirements or user stories definition, are applied in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
+Unit tests and TDD are mentioned, and often completely accepted, in most of the agile methodologies defined so far.
 
@@
 No communication between the two worlds is permitted, it is not defined how to formally specify a module after it was implemented and verified in a non formal way.
-A module only partially specified is a sort of hybrid that needs to be treated accordingly.
+A module only partially specified is a sort of hybrid that must be treated accordingly.
 Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is not always straighforward.
-Identify the correct formal specification is a difficult task that needs to be supported and verified by the development process.
+Identify the correct formal specification is a difficult task, it has to be supported and verified by the development process.
 
@@
 In this case, new handmade tests and existing working code drives the formal specification, this is what we call Test Driven Formal specification (TDF).
-The resulting development cycle to be used to incrementally add formal specifications to an existing working code is shown in Figure \ref{fig:tdf_diagram}.
+The resulting development cycle code is shown in Figure \ref{fig:tdf_diagram}, TDF is used to incrementally add formal specifications to an existing working system.
 
@@
 idea1 (complex): a mock+delegate object, that behaves as the original object, except for a specific call on a specific method.
-idea2 (bad, the pre and post needs to be maintained synchronized): use utility class, where the precondition of a method is \~ the postcondition of the tested one}
+idea2 (bad, the pre and post must maintained synchronized): use utility class, where the precondition of a method is \~ the postcondition of the tested one}
 
@@
 A frame (82 frames in each packet) accommodate data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
-Fast, medium and slow data rates have not a common multiple; a packet, to be capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
+Fast, medium and slow data rates have not a common multiple; a packet, capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
 \begin{itemize}
@@
 \item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame.
-The medium data rate sample channel is a piece of info to be extracted from the frame description (frame metadata), including the possibility to have no data at all.
+The medium data rate sample channel is a piece of info extracted from the frame description (frame metadata), including the possibility to have no data at all.
 E.g. a frame contain a medium data rate sample from channel 0, the following frame a sample from channel 5, the following frame no sample at all, etc.
@@
 There are also rules affecting more than one value (i.e.
-if a specific sensor is active the reading value has to be in a defined range), and rules specifying a correct sequence of frames (i.e.
+if a specific sensor is active the reading value must be in a defined range), and rules specifying a correct sequence of frames (i.e.
 there can be only one switch in sampling rate in the frame sequence of a single packet).
@@
 about the details of the real format of the binary data streams.
-The low-level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to be able to parse or generate.
+The low-level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to parse or generate.
 
@@
 Some of them are rather simple, focusing on constraints that should hold when calling a method (the preconditions) and constraints on the return value (the very basic form of postconditions). 
-In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared to be \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition), the specification also constraints the return value with a lower and upper bound (another postcondition).
+In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition), the specification also constraints the return value with a lower and upper bound (another postcondition).
 
@@
 The complex specifications usually focus on properties regarding the behaviour of a whole object.
-In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property that has to be maintained during the life cycle of an object, more precisely, an invariant is assumed on entry and guaranteed on exit of each method of an object.
+In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property maintained during the life cycle of an object, more precisely, an invariant is assumed on entry and guaranteed on exit of each method of an object.
 The invariant is constraining the number of samples for each medium data rate streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is \lil{FRAMES}, the constant constraining the number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are fairly distributed on the channels.
@@
 The use of simulators to enable unit testing was essential to test the \STSB driver, since the physical board was not available.
-The simulators are used in both the handmade and the generated test suite, since both need a proper set of stubs to be executed.
+The simulators are used in both the handmade and the generated test suite; both the suites, to be executed, need a proper set of stubs.
 
@@
 This is because of two different effects: non public utility methods, and abstract methods in interfaces.
-The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, they are not meant to be called by the system, and they are only used in object initialization or in object setups performed during unit tests.
+The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, the system does not depend on them, they are only used in object initialization or in object setups performed during unit tests.
 In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes.
@@
 The combination of test suites raise the coverage ratio to $ 88.5 \% $. 
-This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not to be considered alternatives, but they need to be used together to achieve the higher benefits, is confirmed.
+This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not alternatives, but they should be used together to achieve the higher benefits, is confirmed.
 A similar indication is provided by \lil{SimulatorDriver} package as well.
@@
 A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviours. 
-With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges would take long to be developed.
+With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
 The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.
@@
 The formal JML specifications and the run time assertion checker correctly identified these errors.
-4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, and are expected to be corrected in the following releases. 
+4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, and are expected to be corrected in the following release. 
 No other errors, except the mentioned ones, have been discovered up to now.
@@
 In this paper we focus on a specific set of development problems that neither agile methodologies nor formal methodologies are completely confortable to cope with.
-The problems, we are interested in, are characterized by unstable requirements, artifacts that need to be formally specified and verified, and are constrained by deadlines that cannot be postponed.
-We propose a blend of agile and formal verification practices, enclosed by an iterative and incremental development process: the Mole development process.
+The problems, we are interested in, are characterized by unstable requirements, by artifacts which must be formally specified and verified, and are constrained by deadlines that cannot be postponed.
+We propose a blend of agile and formal engineering practices, enclosed by an iterative and incremental development process: the Mole development process.
 
 % Mole
-The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, to be used with formal and non formal artifacts.
+The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, used with formal and non formal artifacts.
 A total of four practices are presented, two of them are conservative, following closely the development practices of agile and formal methodologies, while the other pair are inoovative, blending together elements from both worlds.
@@
 The handmade test suite primarily focuses on small features and use cases fragments that the system must implement, or for which an exceptional case has been specified. 
-A test case is to be considered a rough specification written in imperative style. 
+A test case is a rough specification written in imperative style. 
 The common structure requires the objects initialization, followed by a sequence of method calls taking the objects into desired initial states; than the method under test is called and the objects states are finally inspected and verified.
@@
 
-\paragraph{formal specification of the protocol}
+\paragraph*{formal specification of the protocol}
 

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
"can" removed

@@
 The test driven approach relies on writing the tests before the actual implementation. 
-It can be applied in a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
-It can be applied in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
+It is appliable in a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
+It ia appliable in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
 
@@
 Some XP practices are fully adopted or considered compatible, as pair-programming, iterative and incremental development, system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
-A different formal methodology, but same approach, can be found also in~\cite{Suhaib2005} where XFM methodology is explained.
+A different formal methodology, but same approach, is found also in~\cite{Suhaib2005} where XFM methodology is explained.
 
@@
 The \SB protocol is asynchronous and packet based, the packets have a fixed length.  
-The protocol is separated in multiple layers to ease its implementation in the driver, since mixing up different abstraction levels can lead to overly complex and unmanageable code.  
+The protocol is separated in multiple layers to ease its implementation in the driver, since mixing up different abstraction levels usually leads to overly complex and unmanageable code.  
 The layers identified in protocol are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable values and reaction constraints on output packets, based on previously transmitted input packets.
@@
 \begin{itemize}
-\item Slow data rates: slow data can be found at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
+\item Slow data rates: slow data is located at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
 \item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame.
@@
 if a specific sensor is active the reading value must be in a defined range), and rules specifying a correct sequence of frames (i.e.
-there can be only one switch in sampling rate in the frame sequence of a single packet).
+only one switch in sampling rate us allowed in the frame sequence of a single packet).
 
@@
 Verification is obtained through a handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
-The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant how the stream can be properly decomposed into packets.
+The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant on how the stream is decomposed into packets.
 Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
@@
 % development cycles
-The test suite are used to verify board simulators (low-level), but it cannot be used to check properties at run time.
+The test suites are used to verify board simulators (low-level), but they cannot be used to check properties at run time.
 Code and tests have been implemented using TDD.
@@
 The handmade test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
-The big number of generated tests can be understood looking closely at how the generated tests are created by the JML framework. 
+The big number of generated tests are explained looking closely at how the generated tests are created by the JML framework. 
 The generated tests explore the possible input combinations on public methods, combining the type data ranges that are specified for the test suite (see \cite{Cheon-Leavens02} for more details on how the JML framework works on generating unit tests).
-For instance, let's suppose we want to test a method \lil{m(int p1, int p2)} for a class \lil{C}; the data ranges specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the instances \lil{\{c1,c2,c3,c4\}}.
+For instance, let's suppose a method \lil{m(int p1, int p2)} must be tested in a class \lil{C}; the data ranges specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the instances \lil{\{c1,c2,c3,c4\}}.
 The JML framework generates $24 = 3 \times 3 \times 4$ tests, since they are the combination of the data ranges involved (the data range of the parameters and the data range of the type owning the method under test).
@@
 A generated test is not executed, unless the preconditions of the called method are satisfied. 
-Thus, depending on the preconditions specified for a method, the number of the generated tests that are actually executed in the test suite can be significantly smaller compared to the total number of generated tests.
+Thus, depending on the preconditions specified for a method, the number of the generated tests that are actually executed in the test suite are significantly smaller compared to the total number of generated tests.
 
@@
 In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes.
-The used statement coverage tool cannot be configured to distinguish these code blocks, so this effect is unavoidable. 
+The used statement coverage tool is not able to distinguish these code blocks, so this effect is unavoidable. 
 The effect is mainly seen in low \lil{SimulatorDriver} code coverage figures: the \lil{SimulatorDriver} package has many utility non public methods.
@@
 % handmade unit test vs generated unit test
-The qualitative difference of the two test suites can be understood observing how the suites are built. 
-An handmade unit test usually requires objects initialization (it can involve several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
+The qualitative difference of the two test suites are understood observing how the suites are built. 
+An handmade unit test usually requires objects initialization (it involvs several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
 A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviours. 
-With a complete specification, and specifying large enough data ranges, the generated tests can achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
+With a complete specification, and specifying large enough data ranges, the generated tests achieves the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
 The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.
@@
 % extending Mole
-The practices are very general, and can be applied to different development languages and formal methodologies.
+The practices are very general, and are appliable to different development languages and formal methodologies.
 One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at run time and partially automate the verification process of the formalized properties.

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
minimal update

@@
 The combination of test suites raise the coverage ratio to $ 88.5 \% $. 
-This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not alternatives, but they should be used together to achieve the higher benefits, is confirmed.
+This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not alternatives, but have to be used together to achieve the higher benefits, is confirmed.
 A similar indication is provided by \lil{SimulatorDriver} package as well.

### 2010-07-05 papers/AgileFormality/paper.tex
Renaming directory to more closely reflect contents of paper.

--- 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
@@
 
-\title{Agile Formality: A Mole of Software Engineering Practices}
+\title{Agile Formality: A ``Mole'' of Software Engineering Practices}
 
@@
 
-  Members of the agile programming and formal methods communities do
-  not always see eye-to-eye.  These two communities often do not talk
-  to or learn from each other.  Only recently, as highlighted by the
-  September 2009 issue of IEEE Software, the IFIP workshop on
-  balancing agility and formalism in software engineering, and the
-  first edition of the first international workshop for formal methods
-  and agile methods, have ideas from the two communities begun to be
-  synthesized.  While the problem-solving approaches and psychological
-  attitudes of members of the two communities differ widely, we
-  exploit this clash of viewpoints, synthesizing a new development
-  processes that actually blends, rather than mashes together, best
-  practices from the two worlds.  This paper summarizes our process
-  and a supporting complex case study, showing that it is not only
-  possible, but tasty, to combine the ``chili pepper'' of formal
-  methods and the ``chocolate'' of agile programming, thus producing a
-  tasty ``mole'' (as in the highly-spices Mexican sauce) of software
-  engineering practices.
+Members of the agile programming and formal methods communities do not always see eye-to-eye.  
+These two communities often do not talk to or learn from each other.  
+Only recently, as highlighted by the September 2009 issue of IEEE Software, the IFIP workshop on balancing agility and formalism in software engineering, and the first edition of the international workshop for formal methods and agile methods, ideas from the two communities begun to be synthesized.
+While the problem-solving approaches and psychological attitudes of members of the two communities differ widely, we exploit this clash of viewpoints, exploiting a new development processes that actually blends, rather than mashes together, best practices from the two worlds.
+This paper summarizes our process and a supporting complex case study, showing that it is not only possible, but tasty, to combine the ``chili pepper'' of formal methods and the ``chocolate'' of agile programming, thus producing a tasty ``Mole'' (as in the highly-spices Mexican sauce) of software engineering practices.
 
@@
 % agile and formal difficult to blend: mind attitudes
+Agile and formal development methodologies usually do not blend together well.  
+This is because of several reasons, the most important of which is often characterized as a radical difference in psychological attitudes about software development.
 
-Agile and formal development methodologies
-usually do not blend together well.  This is because of several
-reasons, the most important of which is often characterized as a
-radical difference in psychological attitudes about software
-development.
+% FM
+Formal methodologists often favour an in-depth, think-first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall-style of development.  
+Once a formalization is developed, development and verification proceed, in an automated and interactive fashion.  
+Consequently, projects that use formal methodologies (FMs) typically focus on critical systems with fixed requirements and somewhat flexible deadlines (``We'll ship it when it is right.'').
 
-Formal methodologists often favour an in-depth, think-first approach,
-where the problem is understood, formalized, and solved; usually, but
-not always, adopting a waterfall-style of development.  Once a
-formalization is developed, development and verification proceeds, in
-an automated and interactive fashion.  Consequently, projects that use
-formal methodologies (FMs) typically focus on critical systems with
-fixed requirements and somewhat flexible deadlines (``We'll ship it
-when it is right.'')---systems in which if something goes wrong,
-someone dies or, directly or indirectly, looses a very large amount of
-money.
-
-Agile methodologists favour an highly incremental and iterative
-approach, specifically tailored to cope with changing requirements and
-precise deadlines.  In projects that use agile methodologies (AMs), it is
-often the case that the problem is only partially understood.  Thus,
-one focuses on the aspects of a system that must be implemented in the
-current development iteration.  The solution initially developed is
-usually not optimal at all, but as it is refined through continuous
-code refactorings, its quality and validity improve over time.  Test
-suites are used for several purposes, the most important of which are
-system verification, system documentation, and requirement
-specification, and they are strictly handmade.  AMs are typically used
-in development settings where changing requirements and rapid delivery
-of the product are paramount.
+% AM
+Agile methodologists favour an highly incremental and iterative approach, specifically tailored to cope with changing requirements and precise deadlines.
+In projects that use agile methodologies (AMs), it is often the case that the problem is only partially understood.
+Thus, one focuses on the aspects of a system that must be implemented in the current development iteration.  
+The solution initially developed is usually not optimal at all, but as it is refined through continuous code refactorings, its quality and validity improve over time.  
+Test suites are used for several purposes, the most important of which are system verification, system documentation, and requirement specification, and they are strictly handmade.  
+AMs are typically used in development settings where changing requirements and rapid delivery of the product are paramount.
 
 % problems would benefit form agile and formal mix, characterize the problem
-
-The two worlds seem irreconcilable.  Nevertheless, there exist
-problems that would greatly benefit from both~\cite{Black2009}.  The
-problem class that in which we have particular interest have
-\emph{unstable requirements}, are \emph{constrained by deadlines that
-  cannot be postponed}, but also \emph{have subsystems that must be
-  formally specified and verified}.
-
-% terminology
+The two worlds seem irreconcilable.  
+Nevertheless, problems exist that would greatly benefit from both of them~\cite{Black2009}.  
+The problem class in which we have particular interest have \emph{unstable requirements}, are \emph{constrained by deadlines that cannot be postponed}, but also have \emph{subsystems that must be formally specified and verified}.
 
@@
 
-In this paper the IEEE 610~/cite{Jay1990} standard terminology for validation and
-verification is used.  Paraphrasing the standard
-slightly, \emph{verification} is the process of evaluating software to
-determine whether the products of a given development phase satisfy
-the conditions imposed at the start of that phase, according to the
-requirements and design specifications (``you build the system in the
-right way''); validation is the process of evaluating software during
-or at the end of the development process to determine whether it
-satisfies specified requirements ensuring it meets the user's needs
-(``you build the right system'').
+% terminology
+In this paper the IEEE 610~/cite{Jay1990} standard terminology for validation and verification is used.  
+Paraphrasing the standard slightly, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the
+right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
 
-Within the formal methods community, verification means something much
-stronger, and validation something quite different, than in the IEEE
-610 standard.  Loosely, within the FM community, \emph{verification}
-means the use of formal methods to formally, statically (and usually
-modularly) verify that a system conforms to its specification (under
-all possible execution scenarios).  On the other hand,
-\emph{validation} means the use of traditional software engineering
-practices to check, statically or dynamically, that the system
-conforms to its specification (whether in the form of tests,
-requirements, etc.) by hand or via execution under some execution
-scenarios.
-
-As we come from both the AM and FM communities, we respect and
-acknowledge both perspectives and will attempt to be absolutely clear
-in our use of these terms in the following.
-
-Additionally, in the following when we informally say ``...development
-of a system,'' we more precisely mean ``...analysis, design,
-development, verification (both testing and formal verification), and
-validation of a system.''
+Within the formal methods community, verification means something much stronger, and validation something quite different, than in the IEEE 610 standard.
+Loosely, within the FM community, \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically (and usually modularly) verify that a system conforms to its specification (under all possible execution scenarios).
+On the other hand, \emph{validation} (in ther paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
+In the present paper both perspectives are respected and acknowledged, with the attempt to be absolutely clear in using these terms unambigously.
+Additionally, ``...development of a system,'' more precisely means ``...analysis, design, development, verification, formal validation of a system''.
 
@@
 
-Proposing a new process with no supporting evidence is an
-uninteresting proposition.  A complex case study is mandatory to test
-bench a development process.  Our case study focuses on the
-development of a device driver to control and communicate with an
-embedded custom circuit board equipped with more than a dozen sensors.
-The communication protocol with the board is asynchronous and packet
-based and the board is novel and custom-designed.  The protocol is
-focused on controlling the board and reporting sensor measurements and
-board status.
+Proposing new practices with no supporting evidence is an uninteresting proposition.  
+A complex case study is mandatory to test bench the proposed practices.  
+Our case study focuses on the development of a device driver to control and communicate with an embedded custom circuit board equipped with more than a dozen sensors.
+The communication protocol with the board is asynchronous and packet-based and the board is novel and custom-designed.  
+The protocol is focused on controlling the board and reporting sensor measurements and board status.
 
-This software driver has been developed using our process while
-statisfying seemingly contradictory requirements: (1)~the protocol
-implementation must be thoroughly verified; (2)~parts of the protocol
-must be formally specified; (3)~a board simulator must be designed,
-implemented, and verified to match the specifications of, and actual
-behavior of, the physical circuit board; (4)~if board components
-change during development, the associated formal specifications and
-the simulator whose behavior reflects such changes must be updated
-accordingly; (5)~strict software delivery deadlines.  Consequently,
-these process requirements constitute an environment that benefits
-from both an agile and a formal approaches to software development.
+This software driver has been developed using our process while statisfying seemingly contradictory requirements: (1)~the protocol implementation must be thoroughly verified; (2)~parts of the protocol must be formally specified; (3)~a board simulator must be designed, implemented, and verified to match the specifications of, and actual behaviour of, the physical circuit board; (4)~if board components change during development, the associated formal specifications and the simulator whose behavior reflects such changes must be updated accordingly; (5)~strict software delivery deadlines.
+Consequently, these process requirements constitute an environment that benefits from both agile and formal approaches to software development.
 
-Over the course of the past 18 months, an novel development process
-has been refined to cope with the contrasting requirements of the case
-study, blending agile and formal practices.  In the following we
-summarize this development process and report on its successful
-application to the case study, yielding several promising results.
+A novel development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal practices.  
+In the following the development process and the practices are summarized, and their successful application to the case study is reported, yielding several promising results.
 
@@
 % what is needed, high-level
-To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies must be maintained, and formal verification  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, verify the implementation against the specification), must be adapted to a highly iterative one.
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies must be maintained, and formal validation  methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, validate the implementation against the specification), must be adapted to a highly iterative one.
 
@@
 Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven approach is a cornerstone of XP, but it is so popular that have been adopted in many other agile methodologies as well.
+The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other agile methodologies as well.
 It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
@@
 It is appliable in a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
-It ia appliable in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
+It is appliable in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
 
@@
 Some tentative attempts to reconcile formal and agile methodologies have been developed.
-The recurrent approach of the attempts consists in removing the handmade test suites, substituting them with different kinds of automated verification based on the system formal specification. 
-The automated verification can be as simple as enabling run time verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers.
+The recurrent approach of the attempts consists in removing the handmade test suites, substituting them with different kinds of automated formal validation based on the system formal specification. 
+The automated validation can be as simple as enabling runtime verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers (formal verification).
 
@@
 The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines formal methodology. 
-The development process proposed in~\cite{Herranz2003b} by Herranz and Moreno-Navarro is quite similar; they propose an iterative process to model a system using SLAM-SL, an object-oriented specification language, and its tools.
+The development process proposed in~\cite{Herranz2003b} by Herranz and Moreno-Navarro is quite similar: they propose an iterative process to model a system using SLAM-SL, an object-oriented specification language, and its tools.
 Some XP practices are fully adopted or considered compatible, as pair-programming, iterative and incremental development, system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
@@
 As already mentioned, in all these attempts, the handmade test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
+The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as runtime verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
 At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, used to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
@@
 The approach is simple: the verification, documentation and requirement specification purposes of the tests are moved on to the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
-The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is usually greater than writing a suite of tests~/cite{Glass2002}.
+The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is usually greater than writing a suite of tests~\cite{Glass2002}.
 If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified at all.
-The parts not formally specified can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are not detailed nor enforced.
+These can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are neither detailed nor enforced.
 The loops and relations between the different development artifacts are the inner engine of many agile methodologies, they impose an iterative and incremental pace to the development process.
@@
 % what we want to achieve
-The objectives we want to achieve are the following ones: freedom to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
+The objectives we want to achieve are the following: to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
 
@@
 We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how handmade and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
-That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}
+That is, we will show how to blend, and not compose, formal validation and agile verification practices: a ``Mole'' of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}.
 
@@
 
-Mole practices are applied in the context of a real world case study
-that matches the problem requirements previously detailed: unstable
-requirements and development artifacts that need to be formally
-specified and verified.  The case study fits these needs.
+``Mole'' practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and validated.
 
@@
 % SenseTile introduction
-The \textbf{UCD CASL SenseTile System} is a large-scale,
-general-purpose sensor system developed at the University College
-Dublin in Dublin, Ireland.  Our facility provides a scalable and
-extensible infrastructure for performing in-depth investigation into
-both the specific and general issues in large-scale sensor networking.
-This system integrates a sensor platform, a datastore, and a compute
-farm.  The sensor platform is a custom-designed but inexpensive sensor
-platform (called the \ST) paired with general-purpose small-scale
-compute nodes, including everything from low-power PDAs to powerful
-touchscreen-based portable computers.  Besides containing over a dozen
-sensors packaged on the \ST itself, the board is expandable as well,
-as new sensors are added to it easily.
+The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system developed at the University College Dublin in Dublin, Ireland.  
+The facility provides a scalable and extensible infrastructure for performing in-depth investigation into both the specific and general issues in large-scale sensor networking. 
+This system integrates a sensor platform, a datastore, and a compute farm.  
+The sensor platform is a custom-designed but inexpensive sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.
+Besides containing over a dozen sensors packaged on the \ST itself, the board is expandable as well, as new sensors are added to it easily.
 
 % case study: the sensor board development tasks
-The case study is focused in building the sensor board and its
-software driver concurrently.  Because of hard time constraints and
-the initial unavailability of the custom board, we must progress
-concurrently with all the development tasks, including:
-(1)~specification of the communication protocol; (2)~specification of
-the physical sensor board; (3)~development and fabrication of the
-physical board; (4)~development of the embedded software for the
-board; (5)~development of the communication protocol software driver
-the the board; and (6)~development of a software simulator of the
-board.
+The case study is focused in building the sensor board and its software driver concurrently.  
+Because of hard time constraints and the initial unavailability of the custom board, we must progress concurrently with all the development tasks, including: (1)~specification of the communication protocol; (2)~specification of the physical sensor board; (3)~development and fabrication of the physical board; (4)~development of the embedded software for the board; (5)~development of the communication protocol software driver of the board; and (6)~development of software simulators of the board.
 
 % what we have to do, what we outsourced, and the dependencies between tasks
-The development of the physical board, along with its embedded
-software (development tasks~2, 3, 4), is carried out by an third party
-under our guidance, thus it is not directly taken into account here.
-The specification of the communication protocol (development task~1)
-is a joint effort between ourselves and the third party, while the
-remaining tasks~(5, 6) are performed in isolation.  Dependencies are
-straightforward: the specification of the communication protocol~(1)
-is the most stable element, but it still depends on the sensor
-board~(2, 3, 4): large changes in the latter affect the former.
-Additionally, the driver and the simulator depends directly on the
-communication protocol.
+The development of the physical board, along with its embedded software (development tasks~2, 3, 4), is carried out by a third party under our guidance, thus it is not directly taken into account here.
+The specification of the communication protocol (development task~1) is a joint effort between ourselves and the third party, while the remaining tasks~(5, 6) are performed in isolation.  
+Dependencies are straightforward: the specification of the communication protocol~(1) is the most stable element, but it still depends on the sensor board~(2, 3, 4): large changes in the latter affect the former.
+Additionally, the driver and the simulator depends directly on the communication protocol.
 
 % communication with external manufacturer
-Communication and synchronization with the external manufacturer is
-frequent but not optimal.  The main synchronization artifact is the
-protocol specification, which remains stable at high-level, but is
-changed frequently in the low-level details.  Since domain expertise
-is very different between the two teams, a common domain language is
-difficult to find and to agree upon.  Formal methodologies and
-programming platforms also differ.  Consequently, more for social and
-psychological reasons than technical ones, the informal specification
-of the communication protocol is the most reliable source of
-information.
+Communication and synchronization with the external manufacturer is frequent but not optimal.  
+The main synchronization artifact is the protocol specification, which remains stable at high-level, but is changed frequently in the low-level details.  
+A common domain language has been found and agreed upon, but the formal methodologies and programming platforms differs widely.  
+Consequently the informal specification of the communication protocol is the most reliable source of information.
 
@@
 % how the problem is usually managed
-There exist various approaches in literature addressing this kind of
-development constraints (rapid development in hardware software
-co-design), but all are focused on large-scale systems.
+There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software co-design), but all are focused on large-scale systems.
 
-The hardware-software formal co-design methodologies usually have
-several common development artifacts~\cite{Slomka2000, Hoffman2001}
-including a high-level specification of the system, a translation
-(refinement) of the high-level specification to low-level ones
-(hardware and software counterparts), the possibility to generate
-software code from the low-level specifications, a hardware simulator
-capable of simulating an hardware component based on hardware
-specification, the hardware component developed.
+% hardware-software
+The hardware-software formal co-design methodologies usually have several common development artifacts~\cite{Slomka2000, Hoffman2001} including a high-level specification of the system, a translation (refinement) of the high-level specification to low-level ones (hardware and software counterparts), the possibility to generate software code from the low-level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.
 
-When considering the development of the device and of its software
-driver as separate entities that possibly have to be developed
-concurrently, the existing approaches are similar to the ones seen in
-the case of hardware-software co-design~\cite{Valderrama1995,
-  Siegmund2002, Ryzhyk2009}, all of which focus on a specification
-that aims to be as complete.
+% device drivers
+When considering the development of the device and of its software driver as separate entities that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware-software co-design~\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}, all of which focus on a specification that aims to be as complete as possible.
 
 % complete specification is not feasible
-A complete specification is not feasible nor convenient in our case.
-The protocol specification must be enhanced incrementally and its
-complete specification cannot be provided.  This makes the problem an
-ideal candidate to test the our ``mole'' of practices.
+A complete specification is neither feasible nor convenient in our case. 
+The protocol specification must be enhanced incrementally and its complete specification cannot be provided.  
+This makes the problem an ideal candidate to test our ``Mole'' of practices.
 
 % formal methodology constraints
-To perform this case study, an appropriate formal methodology must be
-chosen.  It must be able to support a specification that is built
-incrementally.  Consequently, formal methods that demand a complete
-specification must be avoided.  Additionally,
-to best synthesize with agile methods, the formal language must be
-able to automatically generate tests.  Finally, the formal method must
-be complemented by tools that support the verification of system
-properties both at design-time (via static analysis, for full
-functional verification of critical subsystems) and at run time (for
-dynamic validation for integration with TDD).
+To perform this case study, an appropriate formal methodology must be chosen.  
+It must be able to support a specification that is built incrementally.  
+Consequently: formal methods that demand a complete specification must be avoided; additionally, to best synthesize with agile methods, the formal language must be able to automatically generate tests; finally, the formal method must be complemented by tools that support the verification of system properties both at design-time (via static analysis, for full functional verification of critical subsystems) and at runtime (for dynamic validation for integration with TDD).
+
+\subsection{The Chosen Formal Methodology: Formal Specifications with JML}
 
 % the chosen formal methodology
-Driven by these requirements and our expertise, the Java programming
-language and platform were chosen for the implementation, the Java
-Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} and Business
-Object Notation (BON)~\cite{WaldenNerson95} are used to write formal
-specifications, and the BONc~\cite{BONc}, Beetlz~\cite{Beetlz},
-JML2~\cite{JML2,JML2common}, and Extended Static Checker for Java
-(ESC/Java2)~\cite{ESCJava2,KiniryCok04SmallCite} tool suites are used
-as supporting tools~\cite{BurdyEtal05-STTT}.  The JML2 tool suite
-includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit
-test generator~\cite{Cheon2002}.  The BONc tool supports validity
-checking of high-level requirements and architecture specifications,
-Beetlz supports reasoning about refinement between BON and JML
-specifications, and ESC/Java2 supports automatic, modular, static
-reasoning about the correctness of an implementation against its
-formal specification.  All four tool suites do not require a complete
-specification to be used effectively.
+Driven by these requirements and our expertise, the Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2}, and Extended Static Checker for Java (ESC/Java2)~\cite{ESCJava2,KiniryCok04SmallCite} tool suites are used as supporting tools~\cite{BurdyEtal05-STTT}.
+The JML2 tool suite includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit test generator~\cite{Cheon2002}.
+ESC/Java2 supports automatic, modular, static reasoning about the correctness of an implementation against its formal specification.
+These tool suites do not require a complete specification to be used effectively.
 
-\subsection{\todo{paper structure title}}
+% JML introduction
+JML is the de facto standard specification language for describing Java programs behaviour.  
+JML is a rich behavioural interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioural subtyping~\cite{LeavensBakerRuby99}.
+JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants.
+It also includes a rich model-based specification layer that includes mathematical models, data refinements, datagroups, and many other advanced object-oriented specification features~\cite{Chalin06}.
+Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{BurdyEtal05-STTT}.
 
-\todo{explain paper structure}
+% using JML
+JML is sometimes used in a Design-by-Contract style, where a specification is written from scratch, reusing existing APIs that have
+specifications of their own, and then an implementation is written conforming to that specification~\cite{Meyer92b}.  
+The implementation is checked by whatever means appropriate using some subset of the aforementioned tools.  
+At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process called Contracting-the-Design.
+
+% JML tools
+The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the
+Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.  
+The former compiles executable assertions into runtime checks.  
+The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
+As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{KiniryEtal06}.
 
 % ======================================================================
-\section{Agile Methodologies}
-\label{sec:agile_methodologies}
+\section{Blending Formal And Agile Development}
+\label{sec:blending_formal_and_agile_development}
 
@@
 % abstraction in test suites
-High level test suites, composed of functional tests, describe requirements and bind together requirements documentationand system behaviour: an informal requirement is supported by a set of functional tests.
+High level test suites, composed of functional tests, describe requirements and bind together requirement documentation and system behaviour: an informal requirement is supported by a set of functional tests.
 The associated development cycle is one iteration long. 
-Low level test suites, composed of unit tests, describe the modules behaviour and bind together code documentation, code and the contract imposed on the module: unit tests represent the contract and the documentation.
+Low level test suites, composed of unit tests, describe the module behaviour and bind together code documentation, code and the contract imposed on the module: unit tests represent the contract and the documentation.
 The associated development cycle is less than one day (as short as a few minutes).
@@
 
-% ======================================================================
-\section{Formal Specifications With JML}
-\label{sec:formal_specifications_using_jml}
-
-% JML introduction
-The Java Modeling Language (JML) is the de facto standard
-specification language for describing the behavioural properties of
-Java programs.  JML is a rich behavioural interface specification
-language (BISL) and thus focuses on the modular specification of
-classes and objects in which inheritance is used solely for
-behavioural subtyping~\cite{LeavensBakerRuby99}.  JML includes
-standard specification constructs like assumptions, assertions,
-axioms, and pre- and postconditions, framing clauses, and invariants.
-It also includes a rich model-based specification layer that includes
-mathematical models, data refinements~\cite{}, datagroups~\cite{}, and
-many other advanced object-oriented specification
-features~\cite{chalin06:_beyon_asser}.  Many tools, ranging from
-compilers to static checkers to full-functional verification systems
-support the JML language~\cite{BurdyEtal05-STTT}.
-
-% using JML
-JML is sometimes used in a Design-by-Contract style, where a
-specification is written from scratch, reusing existing APIs that have
-specifications of their own, and then an implementation is written
-conforming to that specification~\cite{Meyer92b}.  The
-implementation is checked by whatever means appropriate using some
-subset of the aforementioned tools.  At other times an existing piece
-of implemented and tested code is annotated with specifications
-after-the-fact, a process called Contracting-the-Design.
-
-% JML tools
-The two tools used most frequently to check the correctness of
-implementations are the JML Runtime Assertion Checker (RAC) and the
-Extended Static Checker for Java, version 2
-(ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.  The
-former compiles executable assertions into runtime checks.  The latter
-tool performs automated modular semantic static analysis to attempt to
-prove that the program under analysis does not misbehave (i.e., crash
-in a variety of ways) and conforms to its specification (lightweight
-functional correctness).  As ESC/Java2 is fully automated, it is not
-sound, nor does it cover the whole Java and JML languages, and is thus
-not complete, but there is support for warning the user about both of
-these situations~\cite{KiniryEtal06}.
-
-% ======================================================================
-\section{Blending Formal And Agile Development}
-\label{sec:blending_formal_and_agile_development}
+\subsection{``Mole'' practices}
 
@@
 The development cycle adopted in TDD is shown in Figure \ref{fig:tdd_diagram} and the one adopted in classic FDD is shown in Figure \ref{fig:fdd_diagram}.
-Unit tests, in TDD, are substituted, in FDD, by formal specifications supported by the verification environment, in our case the verification environment is the generated unit test suite.
+Unit tests used in TDD are substituted  in FDD by formal specifications supported by the verification environment; in our case the verification environment is the generated unit test suite.
 Formal specifications replace unit tests: they specify, document and verify the system under development.
@@
 \begin{figure}[ht]
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[b]{0.48\linewidth}
     \centering
@@
   \hspace{0.02\linewidth}
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[b]{0.48\linewidth}
     \centering
@@
 Yet, it is not enough.
-There are several problem that still needs to be addressed.
-No communication between the two worlds is permitted, it is not defined how to formally specify a module after it was implemented and verified in a non formal way.
+There are several problems that still need to be addressed.
+No communication between the two worlds is permitted, it is not defined how to formally specify a module after it has been implemented and verified in a non formal way.
 A module only partially specified is a sort of hybrid that must be treated accordingly.
@@
 The constraints imposed by formal verification tools~\cite{Glass2002} must be taken into account when deciding whether supporting the code with handmade tests, and verification tools based on automatically generated tests are not an exception.
-The development cycle able to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests, both handmade tests and the formal specification drive the code development.
+The development cycle able to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests.
+Both handmade tests and the formal specification drive the code development.
 
 \begin{figure}[ht]
-  \begin{minipage}[b]{0.49\linewidth}
+  \begin{minipage}[b]{0.48\linewidth}
     \centering
@@
 
-% all together belnds into a mole
-When the Mole verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
-
-\todo{if time available: reason about test harness, who is providing what}
-
-\todo{if time available: reason about RAC, when to use it and when not to use it}
-
-\todo{if time available: result: handmade test execution, some should pass with and without run-time verification, some only with run-time verification, some only without run-time verification}
-
-\todo{unrevealed problem (hook for a follow up paper): TDF does not work well when testing postconditions, works correctly only when testing preconditions.
-idea1 (complex): a mock+delegate object, that behaves as the original object, except for a specific call on a specific method.
-idea2 (bad, the pre and post must maintained synchronized): use utility class, where the precondition of a method is \~ the postcondition of the tested one}
+% all together blends into a mole
+When the ``Mole'' verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; the complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
 
@@
 % protocol layers
-The \SB protocol is asynchronous and packet based, the packets have a fixed length.  
+The \SB protocol is asynchronous and packet-based; the packets have a fixed length.  
 The protocol is separated in multiple layers to ease its implementation in the driver, since mixing up different abstraction levels usually leads to overly complex and unmanageable code.  
@@
 % packet internal structure: collection of frames
-A packet is internally composed by frames. 
-A frame (82 frames in each packet) accommodate data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
+A packet is internally composed by 82 frames. 
+A frame accommodates data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
 Fast, medium and slow data rates have not a common multiple; a packet, capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
@@
 if a specific sensor is active the reading value must be in a defined range), and rules specifying a correct sequence of frames (i.e.
-only one switch in sampling rate us allowed in the frame sequence of a single packet).
+only one switch in sampling rate is allowed in the frame sequence of a single packet).
 
@@
 The specifications of the protocol are distilled and refined incrementally.
-The protocol is divided in various (thin) layers, and each of the layer has been verified with a different approach: some of the layers have been specified formally.
+The protocol is divided in various (thin) layers, and each of the layer is verified with a different approach: some of the layers are specified formally.
 
@@
 % verification
-Verification is obtained through a handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
+Verification is obtained through an handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
 The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant on how the stream is decomposed into packets.
@@
 % development cycles
-The test suites are used to verify board simulators (low-level), but they cannot be used to check properties at run time.
-Code and tests have been implemented using TDD.
+The test suites are used to verify board simulators (low-level), but they cannot be used to check properties at runtime.
+Code and tests are implemented using TDD.
 
@@
 It has been built starting from handmade tests only (TDD development cycles).
-Later on JML annotations have been added incrementally, using a mix of mole practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the mole practices have been used according to the artifacts involved.
+Later on JML annotations have been added incrementally, using a mix of ``Mole'' practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the ``Mole'' practices have been used according to the artifacts involved.
 
 % development cycles
-The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check the formally specified properties at run time.
-Code and tests have been initially implemented using TDD, than using all the Mole verification practices.
+The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check the formally specified properties at runtime.
+Code and tests have been initially implemented using TDD, than using all the ``Mole'' verification practices.
 
@@
 The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and handmade test suite to verify the most complex behaviour.
-It has been built starting from JML annotations, supported by handmade tests when needed.
+It is built starting from JML annotations, supported by handmade tests when needed.
 
 % development cycles
-The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check most of the rules at run time.
-Code and tests have been implemented using both forms of FDD.
+The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check most of the rules at runtime.
+Code and tests are implemented using both forms of FDD.
 
@@
 Simulators are used to test parts of the system minimizing dependencies: a simulator can be used by upper layers, with no need to provide the functionalities of the lower layers. 
-The simulators proved essential during development, since the physical board was not available.
+The simulators proved essential during the development, since the physical board was not available.
 
 % how many of them do we need?
-Theoretically a separate simulator is needed for each protocol layer, to specify, develop and test each layer in isolation.  
-But, because of limited resources and effort available, the layer structure was not matched by corresponding simulators.
+Theoretically a separate simulator is needed for each protocol layer to specify, develop and test each layer in isolation.  
+But, because of limited resources and effort available, the layer structure is not matched by corresponding simulators.
 Two simulators are available, they have been built concurrently with the specification, the driver implementation, and the tests.
@@
 % simulators description
-The driver API is splitted in two abstraction layers: the general high-level interfaces, that expose the main functionalities of the board and the main contents of the packets, and the lower level implementation that parse the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
+The driver API is splitted in two abstraction layers: the general high-level interface, that exposes the main functionalities of the board and the main contents of the packets, and the lower level implementation that parses the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
 The simulators are built according to these two abstraction layers.
@@
 Some of them are rather simple, focusing on constraints that should hold when calling a method (the preconditions) and constraints on the return value (the very basic form of postconditions). 
-In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition), the specification also constraints the return value with a lower and upper bound (another postcondition).
+In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition); the specification also constraints the return value with a lower and upper bound (another postcondition).
 
@@
 % unit tests kinds
-The unit test cases that verify the protocol driver are of two kinds: handmade unit tests and automatically generated unit tests based on JML specifications; they are built to be used together, they are complementary.
+The unit test cases that verify the protocol driver are of two kinds: handmade unit tests and automatically generated unit tests based on JML specifications; they are complementary and built to be used together.
 
 % unit tests effectiveness
-The tests effectiveness is evaluated for each test suite; the evaluation is carried out in section \ref{subsec:test_cases_retrospectives}.
+The test effectiveness is evaluated for each test suite; the evaluation is carried out in section \ref{subsec:test_cases_retrospectives}.
 The test effectiveness evaluation considers three elements: effective results on piloting the real board (quantitative), code coverage
@@
 
-\paragraph*{handmade tests}
+\paragraph*{Handmade tests}
 
@@
 % the test package structure
-The test package structure, which is shown in Figure \ref{fig:class_diagram_test} reflects and mimics the code structure; the tests for an abstract package are abstract, and implemented by the package that tests a corresponding system implementation.
-This is a well known test pattern (the Abstract Test Pattern~\cite{Thomas2004}) used to test that the contracts defined in the abstractions are respected in the several implementations.  
-For instance, package \lil{DriverT} contains abstract tests for the abstractions of package \lil{Driver}, package \lil{StreamDriverT} inherits the abstractions contained in \lil{DriverT} and makes them concrete, to test the corresponding implementation \lil{StreamDriver}; package \lil{StreamDriverT} also contains stand alone tests written specifically for the implementation contained in \lil{StreamDriver}.
+The test package structure, shown in Figure \ref{fig:class_diagram_test}, reflects and mimics the code structure; the tests for an abstract package are abstract, and implemented by the package that tests a corresponding system implementation.
+This is a well known test pattern (the Abstract Test Pattern~\cite{Thomas2004}) used to test that the contracts defined in the abstractions are respected in all the implementations.  
+For instance, package \lil{DriverT} contains abstract tests for the abstractions of package \lil{Driver}, package \lil{StreamDriverT} inherits the abstractions of \lil{DriverT} and makes them concrete, to test the corresponding implementation \lil{StreamDriver}; package \lil{StreamDriverT} also contains stand alone tests written specifically for the implementation \lil{StreamDriver}.
 
@@
 The resulting test package structure is shown in Figure \ref{fig:class_diagram_generatedtest}.  
-The package dependencies are defined by the specifications; for instance, \lil{SimulatorDriverST} are generated tests, they test (hence use) the implementation \lil{SimulatorDriver}, and have been generated using the corresponding specification \lil{SimulatorDriverS}.
+The package dependencies are defined by the specifications; for instance, \lil{SimulatorDriverST} are generated tests, they test the implementation \lil{SimulatorDriver}, and are generated using the corresponding specification \lil{SimulatorDriverS}.
 The generated tests do not maintain the inheritance structure of the code and the specification.
@@
 % simulator roles in tests
-The simulators are used in the handmade test suits as stubs.
+The simulators are adopted in the handmade test suits as stubs.
 The resulting structure is shown in Figure \ref{fig:class_diagram_streamdriver_test}.  
-For instance, in Figure \ref{fig:class_diagram_streamdriver_test} the test suite \lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate the \lil{Stream} used by \lil{StreamDriver}, which is the system under test.
+For instance, in Figure \ref{fig:class_diagram_streamdriver_test} the test suite \lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate the \lil{Stream} parsed by \lil{StreamDriver}, which is the system under test.
 
@@
 % simulator importance
-The use of simulators to enable unit testing was essential to test the \STSB driver, since the physical board was not available.
-The simulators are used in both the handmade and the generated test suite; both the suites, to be executed, need a proper set of stubs.
+Simulators were essential to test the \STSB driver, since the physical board was not available.
+The simulators are adopted in both the handmade and the generated test suite; both the suites need a proper set of stubs in order to be executed.
 
@@
 The handmade test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
-The big number of generated tests are explained looking closely at how the generated tests are created by the JML framework. 
+The large number of generated tests are explained looking closely at how the generated tests are created by the JML framework. 
 The generated tests explore the possible input combinations on public methods, combining the type data ranges that are specified for the test suite (see \cite{Cheon-Leavens02} for more details on how the JML framework works on generating unit tests).
@@
 The JML framework generates $24 = 3 \times 3 \times 4$ tests, since they are the combination of the data ranges involved (the data range of the parameters and the data range of the type owning the method under test).
-Nevertheless, not all the generated tests are executed.
-A generated test is not executed, unless the preconditions of the called method are satisfied. 
+Nevertheless, not all the generated tests are executed: a generated test is not executed, unless the preconditions of the called method are satisfied. 
 Thus, depending on the preconditions specified for a method, the number of the generated tests that are actually executed in the test suite are significantly smaller compared to the total number of generated tests.
@@
 % test execution time, values
-Test suites execution time are different, the handmade test suite, with the run time assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $).
-Most of the time is spent on setting up the suite, only considering the real execution of the tests, with no suite setup, the execution time reduces to less than 4 minutes (the rough ratio is $ 1 : 10 $).
+Test suites execution time are different: the handmade test suite, with the runtime assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $).
+Most of the time is spent on setting up the suite; only considering the real execution of the tests, with no suite setup, the execution time reduces to less than 4 minutes (the rough ratio is $ 1 : 10 $).
 
 % test effectiveness
-Code coverage metrics is used to compare in a quantitative way the effectiveness of the test suites. 
+Code coverage metrics are used to compare in a quantitative way the effectiveness of the test suites. 
 This is only an indicator, since it is accepted that code coverage alone cannot asses the quality of a test suite\cite{Marick1999,Chockler2006}.
-Various coverage metrics exist, the test suites are analyzed with statement code coverage (one of the simplest types)\footnote{The tool used to obtain statement code coverage metrics is \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage reports whether each statement has been encountered during execution.
-Statement code coverage is easy to implement, but lacks the capability to explore the control flow graph associated to the code, that is,
-decision branches are not explored. 
+Various coverage metrics exist, the test suites are analyzed with statement code coverage (one of the simplest types)\footnote{The tool used to obtain statement code coverage metrics is \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage reports whether each statement is encountered during execution.
+Statement code coverage is easy to implement, but lacks the capability to explore the control flow graph associated to the code, that is, decision branches are not explored. 
 The coverage results are reported in the Table \ref{tab:statement_code_coverage}.
@@
 The coverage measures reported in Table \ref{tab:statement_code_coverage} do not reach 100\%. 
-This is because of two different effects: non public utility methods, and abstract methods in interfaces.
+This is because of the effects of both non public utility methods, and abstract methods in interfaces.
 The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, the system does not depend on them, they are only used in object initialization or in object setups performed during unit tests.
@@
 The qualitative difference of the two test suites are understood observing how the suites are built. 
-An handmade unit test usually requires objects initialization (it involvs several methods call), the call to the method under test, and the following assertions to verify the expected states change on the objects involved.
-A generated unit test has a more focused and limited scope: only one method call is performed in each test, it is difficult to explore complex behaviours. 
-With a complete specification, and specifying large enough data ranges, the generated tests achieves the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
+An handmade unit test usually requires objects initialization (it involves several method calls), the call to the method under test, and the following assertions to verify the expected state change on the objects involved.
+A generated unit test has a more focused and limited scope: only one method call is performed in each test, therefore it is difficult to explore complex behaviours. 
+With a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
 The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.
@@
 The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
-The formal JML specifications and the run time assertion checker correctly identified these errors.
+The formal JML specifications and the runtime assertion checker correctly identified these errors.
 4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, and are expected to be corrected in the following release. 
@@
 In this paper we focus on a specific set of development problems that neither agile methodologies nor formal methodologies are completely confortable to cope with.
-The problems, we are interested in, are characterized by unstable requirements, by artifacts which must be formally specified and verified, and are constrained by deadlines that cannot be postponed.
-We propose a blend of agile and formal engineering practices, enclosed by an iterative and incremental development process: the Mole development process.
+The problems we are interested in are characterized by unstable requirements and artifacts which must be formally specified and verified, and are constrained by deadlines that cannot be postponed.
+We propose a blend of agile and formal engineering practices, enclosed by an iterative and incremental development process: the ``Mole'' development process.
 
@@
 Yet a working software product was needed as soon as possible, with incrementally added functionality, and keeping it linked with the last specifications of the protocol and the board.
-Mole verification practices enabled us to keep in sink formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
+``Mole'' verification practices enabled us to keep in sink formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
 We developed handmade test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
@@
 The practices are very general, and are appliable to different development languages and formal methodologies.
-One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at run time and partially automate the verification process of the formalized properties.
+One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at runtime and partially automate the verification process of the formalized properties.
 
-\todo{write conclusions}
-
-%\bibliographystyle{latex8} 
-%\bibliographystyle{lniguide} 
 \bibliography{extra,abbrev,ads,category,complexity,hypertext,icsr,knowledge,languages,linguistics,meta,metrics,misc,modeling,modeltheory,reuse,rewriting,softeng,specification,ssr,technology,theory,web,upcoming,upcoming_conferences,conferences,workshops,verification,escjava,jml,nijmegen}
 
-\newpage
-
-\section{Extracts}
-
-\paragraph*{limited resources}
-
-% effect of the limited resources on the formal specifications
-Hard constraints on driver delivery deadlines define the completeness level of the (\emph{just good enough}) specification.  
-In \STSB driver JML specifies very simple behaviours (method oriented) on the whole driver implementation, and more complex behaviour on the driver core interfaces.  
-The software system parts not specified by the JML2 specification were specified using handmade unit tests.
-
-\paragraph*{handmade unit tests}
-
-% handmade unit tests purposes
-Handmade unit tests specifies the behaviour not specified by formal specification; nevertheless, this is not the only purpose they have.
-Handmade unit tests have been used to drive the development, since the Test Driven Development approach\cite{Beck2003} has been
-followed during development.
-
-% handmade unit tests suite structure
-The handmade test suite primarily focuses on small features and use cases fragments that the system must implement, or for which an exceptional case has been specified. 
-A test case is a rough specification written in imperative style. 
-The common structure requires the objects initialization, followed by a sequence of method calls taking the objects into desired initial states; than the method under test is called and the objects states are finally inspected and verified.
-Mock objects can be used to limit module dependencies. 
-The test suite should be composed only by unit tests, but usually some tests are more complex to be called proper unit tests; they can be seen as simple integration tests.
-
-\paragraph*{formal specification of the protocol}
-
-% why a state machine has not been used to formally specify the protocol
-\ST stream communication protocol is characterized by a well definite packet structure, and a well definite semantic on the packet contents.  
-On the other hand, in the protocol, there are no strong constraints on synchronization, since the communication channel is splitted in two (in and out) separate concurrent and asynchronous channels.
-The lack of hard synchronization constraints reduces the advantage to use a classical state machine based specification (as the one presented in \cite{Hubbers2004}), hence this option has been discarded during development.
-
-\paragraph{jml}
-
-% how jml generated tests work
-The test suite generated is a collection of fine grained unit tests.
-For each public method a set of tests is generated.
-The specified behaviour of the method is checked using the JML2 run time assertion checker, hence, an oracle is not needed.  
-The generated tests send messages to the method under test, once the initial precondition check is passed (if it is not passed, the test is discarded), the test catch assertion violation exceptions; if a precondition is not met the test is not run, if the precondition is met the test is run and if the postcondition is not met an exception is fired and the test fails.
-The test data used to build the tests is the only part that is required to be provided by the user.
-
-\paragraph{unit tests}
-
-% how unit tests should work
-A proper stub, driver and oracle need to be identified for each method under test to obtain a proper unit test suite; stub, driver and oracle are the usual components found in the harness that enable test execution\cite{Binder1999}.
-
-% what is needed to test a package
-To properly unit test a package, the following components need to be identified:
-\begin{description}
-\item[Stub or Skeleton]: a piece of code that simulates the activity of the modules that are not under tests.
-\item[Driver]: a piece of code that calls in a proper way the module under test.
-\item[Oracle]: a mechanism used for determining whether a test has passed or failed, comparing the output of the module under test to the output the oracle determines the module should have.
-\end{description}
-
-% junit
-In JUnit\footnote{\myhref{http://www.junit.org/}{JUnit}}, which is the framework used to manage the implementation and execution of handmade and generated unit tests, there is no clear distinction between drivers, stubs, and oracles; none-less, even if a proper distinction
-is not imposed by the framework, the conceptual distinction still holds.
-
-% who provides what: handmade vs generated
-handmade unit tests and JML generated tests differ in how test components are provided.
-The distinctions and similarities are summarized in Table \ref{tab:test_harness}.
-\begin{table}[htbp]
-  \caption{handmade and generated tests harness.}
-  \label{tab:test_harness}
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}\hline
-       & \textbf{\textit{handmade}} & \textbf{\textit{handmade + RAC}} & \textbf{\textit{generated}} \\\hline
-      \textbf{\textit{Stub}} & manual & manual & manual\\\hline
-      \textbf{\textit{Driver}} & manual & manual & automatic (1 call generation)\\\hline
-      \textbf{\textit{Oracle}} & manual & manual + automatic (RAC) & automatic (RAC)\\\hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-% test execution time, reasons
-The execution time of the generated tests is influenced by:
-\begin{itemize}
-\item The number of the generated tests. 
-This problem is not directly addressable, since it is the heart of the test suite generation creation method itself. 
-All the combinations of possible inputs for each method under test must be generated.
-\item The runtime assertion checker overhead. 
-Each test is executed while RAC checks all the relevant properties. 
-We can speculate that if the RAC assertions could be triggered on and off on single properties in a dynamic way (during runtime), a test suite would take advantage of this possibility, turning on the RAC only on the properties that have to be verified for the classes or methods under tests (and not for all the involved classes and methods); this could save time during the test suite execution.
-\end{itemize}
-
-% tests are helful
-The generated test suite has been helpful to test the incrementally growing specifications and the finer details of each method, while the handmade one was helpful to verify the behaviour of complex methods calls on objects initialized to specific states.
-
-% tests leads to a better design
-An handmade test suite force the code to be test oriented, which usually leads to a better quality and more modularized source code\cite{Binder1999}.
-
-\paragraph*{simulators and tests}
-
-% simulators and tests
-Simulators can be used to ease the tasks of developing the test harness components (Stub, Driver, Oracle).
-A simulator can be used as a stub, to replace a module needed by the module under test.
-A simulator can be used as a driver, stimulating the module under test.
-Finally, a simulator can be used as an oracle, simulating the module under test.
-
-% multiple simulators approach
-The multiple simulators approach is not new, since it has been often used in embedded systems software testing\cite{Broekman2002}.  
-This is a general approach that can (and should) be used to verify both the simulators built and the real systems.  
-It can be easily implemented only if the code has been structured following the dependency inversion principle\cite{Martin1996} (see Figure
-\ref{fig:class_diagram_main}).
-
-\paragraph*{prototype history}
-
-% tests and the first prototype, board side
-The first \STSB prototype was meant to build empty packets, but the packets were built in a way not consistent with the specification.  
-The board errors have been detected by simply using the driver implementation, the \lil{StreamDriver} package.
-
-% tests and the second prototype, driver side
-The second prototype was meant to build packets with values taken from real sensors installed on the board.  
-The only sensors that were not yet working were the sensors devoted to the fast rate data streams (audio sensors).  
-One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions that were not covered by the test cases, but that occasionally showed up during real use.
-One error have been discovered in the low-level system driver that the developed driver is using (the system driver is outside our development scope); to eliminate the error an upgrade of the low-level driver to a new unreleased version were necessary.
-No other errors were detected on the driver side.
-
-% tests and the second prototype, board side
-The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
-The formal JML specifications and the run time assertion checker enables the detection of these errors.
-4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found. 
-No other errors, except the mentioned ones, have been discovered.
-
-\paragraph*{future work}
-
-% problem on highest levels of the protocol
-The highest levels of the protocol could be problematic (see \ref{subsec:data_stream_and_protocol_description}: ``Packet sequence rules'' and ``Packet input output asynchronous rules''). 
-On these levels, the protocol properties regard sequences of packets, and testing them can be tricky, since the sequences of packets are not referred to explicitly, in the implementation.
-
-% solutions on highest levels of the protocol
-Two possible strategies could be used:
-\begin{itemize}
-\item Translate the packet sequence properties in local packet properties (when possible).  
-The advantage is to have properties shown as the usual pre-conditions, post-conditions and invariants.  
-However, the localized properties can be more complex and less intuitive and comprehensible than the original ones; finally, some of the global properties could be impossible to translate.
-\item Use separate utility classes, meant to verify a sequence of packages.  
-This approach is not clean, since it requires adding and using new classes, unneeded by the core implementation, to verify the protocol.
-It has the advantage of keeping the properties that have to be verified mostly unchanged.
-\end{itemize}
-
-\paragraph*{out of scope related works}
-
-\todo{Out of scope: agile and formal used together, but not mixed}
-
-\todo{describe and reference ``Agile Formal Method Engineering.'', formal centric}
-
-\todo{describe and reference ``Formal Agility. How much of each?'', formal centric}
-
-\todo{Mixin with test substitution}
-
-\paragraph*{case study related works}
-
-JML has been used previously to verify a communication protocol in \cite{Hubbers2004}, but the protocol is a smart card protocol, and it
-is well represented through state machines, which are then refined into JML properties.  
-The approach is in waterfall style; than the need to develop the protocol, the specification and the implementation concurrently is not addressed.  
-No handmade or generated test suites are used, the code is specified with JML assertions, refined by the state machine specifications, and then checked statically with ESC/Java2 static checker\cite{CokKiniry04} and at runtime with JML2 tools RAC\cite{BurdyEtal05-STTT}.
-
-The JML-JUnit automatic test suite generation tool has not been widely used or assessed. 
-It is described by its own creators in \cite{Cheon2002,Cheon2004,Cheon2005}; it has been used in \cite{Oriat2004} and \cite{Cheon2005}, but only as a stand alone test suite, and, more importantly, only against proof of concepts system, it has never be proved against real complex systems.
-A brief assessment of the JML-JUnit tool has been made in \cite{Tan2004}, the analysis is based on injecting random generated errors in the source code, and then evaluating the ratio of the errors found by the test suite: the results shows that the JML-JUnit tool is not effective enough if used as a stand alone tool.
-No attempts on running or evaluating the generated test suite alongside with an handmade test suite are made.
-
 \end{document}

### 2010-08-19 papers/AgileFormality/paper.tex
changes to reinclude ifpdf after bug solved some packages removed
formatting packages removed
small changes and clarifications introduced. latex style cleaned up. to be done: expand gained experience, clarify prerequisites to apply methodology, reduce to 20 pages

@@
-%\documentclass{article}
-\documentclass[english]{lni}
+%\documentclass[english]{lni}
+\documentclass{article}
 \usepackage{latin1}
 
-\usepackage{times}
-%\usepackage{ifpdf}
-\usepackage{a4wide}
-\usepackage{pdfsync}
+%\usepackage{times}
+\usepackage{ifpdf}
+%\usepackage{a4wide}
+%\usepackage{pdfsync}
 
@@
 \usepackage{xspace} 
-\usepackage{tabularx} 
-\usepackage{epsfig}
-\usepackage{amsmath} 
-\usepackage{amsfonts} 
-\usepackage{amssymb}
-\usepackage{eucal} 
-\usepackage{stmaryrd} 
+%\usepackage{tabularx} 
+%\usepackage{epsfig}
+%\usepackage{amsmath} 
+%\usepackage{amsfonts} 
+%\usepackage{amssymb}
+%\usepackage{eucal} 
+%\usepackage{stmaryrd} 
 \usepackage{float}
@@
 
-Members of the agile programming and formal methods communities do not always see eye-to-eye.  
+Members of the agile programming and formal methods communities do not always see face-to-face.  
 These two communities often do not talk to or learn from each other.  
@@
 While the problem-solving approaches and psychological attitudes of members of the two communities differ widely, we exploit this clash of viewpoints, exploiting a new development processes that actually blends, rather than mashes together, best practices from the two worlds.
-This paper summarizes our process and a supporting complex case study, showing that it is not only possible, but tasty, to combine the ``chili pepper'' of formal methods and the ``chocolate'' of agile programming, thus producing a tasty ``Mole'' (as in the highly-spices Mexican sauce) of software engineering practices.
+This paper summarizes our process and a supporting complex case study, showing that it is not only possible, but tasty, to combine the ``chili pepper'' of formal methods and the ``chocolate'' of agile programming, thus producing a tasty ``Mole'' (as in the highly-spiced Mexican sauce) of software engineering practices.
 
@@
 
-% terminology
+% IEEE terminology
 In this paper the IEEE 610~/cite{Jay1990} standard terminology for validation and verification is used.  
-Paraphrasing the standard slightly, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the
-right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
+Paraphrasing the standard slightly, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
 
+% FM terminology
 Within the formal methods community, verification means something much stronger, and validation something quite different, than in the IEEE 610 standard.
-Loosely, within the FM community, \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically (and usually modularly) verify that a system conforms to its specification (under all possible execution scenarios).
-On the other hand, \emph{validation} (in ther paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
-In the present paper both perspectives are respected and acknowledged, with the attempt to be absolutely clear in using these terms unambigously.
+Loosely, within the FM community, \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically verify that a system conforms to its specification.
+On the other hand, \emph{validation} (in the paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
+In the present paper both perspectives are respected and acknowledged.
+
+% tests terminology
+Tests are usually classified by dichotomies: unit vs. functional, black-box vs. white-box, testing vs. design, tester vs. coder. 
+Yet testing doesn’t always fit in the dichotomies, especially in the case of agile testing approaches~/cite{Beck2007}.
+Black-box vs. white-box dichotomy is particularly problematic, since the box boundaries change the perspective.
+In the paper, functional tests are always black-box tests, coverage tests are always white-box tests, generated tests are always black-box tests (the box is the module specified); other strict rules do not hold.
+
+% development terminology
 Additionally, ``...development of a system,'' more precisely means ``...analysis, design, development, verification, formal validation of a system''.
@@
 The test driven approach relies on writing the tests before the actual implementation. 
-It is appliable in a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
-It is appliable in a large scale as well (requirements immediately translated into functional tests), yeilding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
+It is applicable to a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
+It is applicable to a large scale as well (requirements immediately translated into functional tests), yielding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
 
@@
 % previous work: a common approach
-As already mentioned, in all these attempts, the handmade test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as runtime verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
-At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, used to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+As already mentioned, in all these attempts, the test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
+The requisite to enable these approaches is, at least, a specification of all the relevant aspects of the system under development, to enable simple verification techniques such as runtime verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
+At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, used to statically prove properties about the system~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
@@
 
+% outsider
+Liu in~\cite{Liu2010} takes a different approach where SOFL methodology~\cite{Liu2004} is merged into agile processes. 
+SOFL methodology is based on a three step specification approach: (1) informal specification, (2) semi-formal specification, (3) formal specification. 
+The agile adaptation consists mainly in shorter loops between activities, and into a reduced set of formal specifications.
+The formal specifications are used only to help understanding ambiguous statements in the semi-formal specification, and are not maintened while the system evolves; testing and inspections are based on the semi-formal specification.
+The formal specification are partial, which is interesting, but used only for documentation purposes and than discarded.
+
 % what we want to achieve
-The objectives we want to achieve are the following: to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
+The objectives we want to achieve are the following ones: to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
 
@@
 We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how handmade and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
-That is, we will show how to blend, and not compose, formal validation and agile verification practices: a ``Mole'' of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}.
+That is, we will show how to blend, and not compose, formal validation and agile verification practices: a ``Mole'' of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers (agile) and chocolate (formal), served with meat (working system).}.
 
@@
 Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is not always straighforward.
-Identify the correct formal specification is a difficult task, it has to be supported and verified by the development process.
+Identifying the correct formal specification is a difficult task, it has to be supported and verified by the development process.
 
@@
 Verification is obtained through an handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
-The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confidant on how the stream is decomposed into packets.
+The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confident on how the stream is decomposed into packets.
 Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
@@
 % test effectiveness
-Code coverage metrics are used to compare in a quantitative way the effectiveness of the test suites. 
+Code coverage metrics (white-box testing) are used to compare in a quantitative way the effectiveness of the test suites. 
 This is only an indicator, since it is accepted that code coverage alone cannot asses the quality of a test suite\cite{Marick1999,Chockler2006}.
@@
 The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, used with formal and non formal artifacts.
-A total of four practices are presented, two of them are conservative, following closely the development practices of agile and formal methodologies, while the other pair are inoovative, blending together elements from both worlds.
+A total of four practices are presented, two of them are conservative, following closely the development practices of agile and formal methodologies, while the other pair are innovative, blending together elements from both worlds.
 
@@
 Yet a working software product was needed as soon as possible, with incrementally added functionality, and keeping it linked with the last specifications of the protocol and the board.
-``Mole'' verification practices enabled us to keep in sink formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
+``Mole'' verification practices enabled us to keep in sync formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
 We developed handmade test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
@@
 % extending Mole
-The practices are very general, and are appliable to different development languages and formal methodologies.
+The practices are very general, and are applicable to different development languages and formal methodologies.
 One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at runtime and partially automate the verification process of the formalized properties.
+
+\bibliographystyle{lni}
 

### 2010-08-19 papers/AgileFormality/paper.tex
small changes, refactorings, shortenings

@@
 Agile methodologists favour an highly incremental and iterative approach, specifically tailored to cope with changing requirements and precise deadlines.  
-In projects that use agile methodologies (AMs), it is often the case that the problem is only partially understood.
-Thus, one focuses on the aspects of a system that must be implemented in the current development iteration.  
-The solution initially developed is usually not optimal at all, but as it is refined through continuous code refactorings, its quality and validity improve over time.  
-Test suites are used for several purposes, the most important of which are system verification, system documentation, and requirement specification, and they are strictly handmade.  
+In projects that use agile methodologies (AMs), it is often the case that the problem is only partially understood, focusing only on the aspects that must be implemented in the current development iteration.
+The solution initially developed is usually not optimal, but as it is refined through continuous code refactorings, its quality and validity improve.  
+Test suites are used for several purposes: the most important are system verification, system documentation, and requirement specification, and they are strictly handmade.  
 AMs are typically used in development settings where changing requirements and rapid delivery of the product are paramount.
@@
 % IEEE terminology
-In this paper the IEEE 610~/cite{Jay1990} standard terminology for validation and verification is used.  
-Paraphrasing the standard slightly, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
+In this paper the IEEE 610~\cite{Jay1990} standard terminology for validation and verification is used.  
+Paraphrasing the standard, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
 
@@
 % what is needed, high-level
-To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the agile methodologies must be maintained, and formal validation methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, validate the implementation against the specification), must be adapted to a highly iterative one.
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the AMs must be maintained, and formal validation methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, validate the implementation against the specification), must be adapted to a highly iterative one.
 
 % what must be considered: TDD
-The most common software verification practices in agile methodologies must be considered, when trying to integrate formal and agile methodologies, since they are a fundamental part of the development process.
+The most common software verification practices in AMs must be considered, when trying to integrate FMs and AMs, since they are a fundamental part of the development process. 
 The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
 Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other agile methodologies as well.
+The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other AMs as well.
 It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
@@
 % previous related work
-Some tentative attempts to reconcile formal and agile methodologies have been developed.
+Some tentative attempts to reconcile FMs and AMs have been developed.
 The recurrent approach of the attempts consists in removing the handmade test suites, substituting them with different kinds of automated formal validation based on the system formal specification. 
@@
 Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an iterative and incremental process to generate X-Machines models, an extension of Finite State Machines. 
-The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines formal methodology. 
+The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines FM. 
 The development process proposed in~\cite{Herranz2003b} by Herranz and Moreno-Navarro is quite similar: they propose an iterative process to model a system using SLAM-SL, an object-oriented specification language, and its tools.
 Some XP practices are fully adopted or considered compatible, as pair-programming, iterative and incremental development, system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
-A different formal methodology, but same approach, is found also in~\cite{Suhaib2005} where XFM methodology is explained.
+A different FM, but same approach, is found also in~\cite{Suhaib2005} where XFM methodology is explained.
 
@@
 The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is usually greater than writing a suite of tests~\cite{Glass2002}.
-If a complete formal specification is not built, all the previous methodologies share a common problem: the parts not formally specified are not verified at all.
+If a complete formal specification is not built, all the previous methodologies share a common problem: the parts informally specified are not verified at all.
 These can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are neither detailed nor enforced.
-The loops and relations between the different development artifacts are the inner engine of many agile methodologies, they impose an iterative and incremental pace to the development process.
+The loops and relations between the different development artifacts are the inner engine of many AMs, they impose an iterative and incremental pace to the development process.
 
@@
 % how we want to achive it
-The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem in an highly iterative development process.
+The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in AMs (especially in XP), to solve this problem in an highly iterative development process.
 We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how handmade and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
@@
 The main synchronization artifact is the protocol specification, which remains stable at high-level, but is changed frequently in the low-level details.  
-A common domain language has been found and agreed upon, but the formal methodologies and programming platforms differs widely.  
+A common domain language has been found and agreed upon, but the FMs and programming platforms differs widely.  
 Consequently the informal specification of the communication protocol is the most reliable source of information.
@@
 % formal methodology constraints
-To perform this case study, an appropriate formal methodology must be chosen.  
+To perform this case study, an appropriate FM must be chosen.  
 It must be able to support a specification that is built incrementally.  
@@
 % principles of agile methodologies
-Agile methodologies are all based on a highly iterative and incremental process.
+AMs are all based on a highly iterative and incremental process.
 They share a common approach on team management, customer relation, simplification and removal of unnecessary artifacts and activities, they rate working solutions and customers satisfaction as the most important indicators considered during development.
-The Agile Manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all agile methodologies.
+The Agile Manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all AMs.
 
 % verification practices in agile methodologies
-The Agile Manifesto does not suggest any specific development techniques, it describes agile methodologies from a very high abstraction level.
-Nevertheless, most agile methodologies share a similar approach to artifact verification.
+The Agile Manifesto does not suggest any specific development techniques, it describes AMs from a very high abstraction level.
+Nevertheless, most AMs share a similar approach to artifact verification.
 Functional test suites, used as precise requirements or user stories definition, are applied in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
-Unit tests and TDD are mentioned, and often completely accepted, in most of the agile methodologies defined so far.
+Unit tests and TDD are mentioned, and often completely accepted, in most of the AMs defined so far.
 
@@
 Handmade tests have to be used together with generated ones.
-Handmade tests verify complex behaviours that are not formally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools, because of the limits of the verification tools themselves.
+Handmade tests verify complex behaviours that are informally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools, because of the limits of the verification tools themselves.
 The constraints imposed by formal verification tools~\cite{Glass2002} must be taken into account when deciding whether supporting the code with handmade tests, and verification tools based on automatically generated tests are not an exception.
@@
 Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
-Therefore, packet byte structure is not formally specified.
+Therefore, packet byte structure is informally specified.
 
@@
 % the problem
-In this paper we focus on a specific set of development problems that neither agile methodologies nor formal methodologies are completely confortable to cope with.
+In this paper we focus on a specific set of development problems that neither agile methodologies nor FMs are completely confortable to cope with.
 The problems we are interested in are characterized by unstable requirements and artifacts which must be formally specified and verified, and are constrained by deadlines that cannot be postponed.
@@
 % Mole
-The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, used with formal and non formal artifacts.
-A total of four practices are presented, two of them are conservative, following closely the development practices of agile and formal methodologies, while the other pair are innovative, blending together elements from both worlds.
+The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, used with formal and informal artifacts.
+A total of four practices are presented, two of them are conservative, following closely the development practices of FMs and AMs, while the other pair are innovative, blending together elements from both worlds.
 
@@
 % extending Mole
-The practices are very general, and are applicable to different development languages and formal methodologies.
-One important constraint is on the formal methodology, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at runtime and partially automate the verification process of the formalized properties.
+The practices are very general, and are applicable to different development languages and FMs.
+One important constraint is on the FM, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at runtime and partially automate the verification process of the formalized properties.
 

### 2010-08-23 papers/AgileFormality/paper.tex
some reductions, but not enough, retrospective on mole still to be inserted, 2 pages longer than limit

@@
 In this paper the IEEE 610~\cite{Jay1990} standard terminology for validation and verification is used.  
-Paraphrasing the standard, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed at the start of that phase, according to the requirements and design specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software during or at the end of the development process to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
+Paraphrasing the standard, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed, according to the specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
 
 % FM terminology
-Within the formal methods community, verification means something much stronger, and validation something quite different, than in the IEEE 610 standard.
-Loosely, within the FM community, \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically verify that a system conforms to its specification.
-On the other hand, \emph{validation} (in the paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
+Within the FMs community, verification and validation have different meanings than in the IEEE 610 standard.
+Loosely \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically verify that a system conforms to its specification; \emph{validation} (in the paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
 In the present paper both perspectives are respected and acknowledged.
@@
 Tests are usually classified by dichotomies: unit vs. functional, black-box vs. white-box, testing vs. design, tester vs. coder. 
-Yet testing doesn’t always fit in the dichotomies, especially in the case of agile testing approaches~/cite{Beck2007}.
+Yet testing doesn't always fit in the dichotomies, especially in the case of agile testing approaches~\cite{Beck2007}.
 Black-box vs. white-box dichotomy is particularly problematic, since the box boundaries change the perspective.
@@
 The most common software verification practices in AMs must be considered, when trying to integrate FMs and AMs, since they are a fundamental part of the development process. 
-The most significant verification practice is Test Driven Development (TDD)~\cite{Beck2003}, TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
-Actually, the general technique, consisting in writing the tests before the code, is used at high abstraction levels as well, since user stories (the requirements) are translated into test cases (the functional tests) before implementation.
-The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other AMs as well.
-It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
+The most significant verification related practice is Test Driven Development (TDD)~\cite{Beck2003}: TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
 
@@
 The test driven approach relies on writing the tests before the actual implementation. 
-It is applicable to a very small scale (Test Driven Development) yielding to activity cycles as short as a few minutes.
+It is applicable to a very small scale (TDD) yielding to activity cycles as short as a few minutes.
 It is applicable to a large scale as well (requirements immediately translated into functional tests), yielding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
+The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other AMs as well.
+It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
 
@@
 As already mentioned, in all these attempts, the test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at least, a specification of all the relevant aspects of the system under development, to enable simple verification techniques such as runtime verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
-At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, used to statically prove properties about the system~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+The requisite to enable these approaches is, at minimum, a specification of all the relevant aspects of the system under development, to enable simple verification techniques such as runtime verification (assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
+Whether a complete and sound system specification, supported by an appropriate formal language (to specify model and properties) and a constrained programming language, can be able to statically prove properties about the system~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
@@
 The formal specifications are used only to help understanding ambiguous statements in the semi-formal specification, and are not maintened while the system evolves; testing and inspections are based on the semi-formal specification.
-The formal specification are partial, which is interesting, but used only for documentation purposes and than discarded.
+The formal specification are partial, which is interesting, but used only for documentation purposes and than discarded, they are not used at all in the verification of the system.
 
@@
 
-\subsubsection{Related Methodologies}
+\subsection{Related Methodologies}
 \label{sec:relat-meth}
@@
 
-% device drivers
+% device drivers and formal specification
 When considering the development of the device and of its software driver as separate entities that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware-software co-design~\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}, all of which focus on a specification that aims to be as complete as possible.
-
-% complete specification is not feasible
 A complete specification is neither feasible nor convenient in our case. 
@@
 
+\subsection{The Chosen Formal Methodology: Formal Specifications with JML}
+
 % formal methodology constraints
 To perform this case study, an appropriate FM must be chosen.  
-It must be able to support a specification that is built incrementally.  
-Consequently: formal methods that demand a complete specification must be avoided; additionally, to best synthesize with agile methods, the formal language must be able to automatically generate tests; finally, the formal method must be complemented by tools that support the verification of system properties both at design-time (via static analysis, for full functional verification of critical subsystems) and at runtime (for dynamic validation for integration with TDD).
-
-\subsection{The Chosen Formal Methodology: Formal Specifications with JML}
+It must be able to support a specification that is built incrementally, consequently formal methods that demand a complete specification must be avoided.
+It must be able to automatically generate tests.
+And finally the formal method must be complemented by tools that support the verification of system properties at runtime.
 
 % the chosen formal methodology
-Driven by these requirements and our expertise, the Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2}, and Extended Static Checker for Java (ESC/Java2)~\cite{ESCJava2,KiniryCok04SmallCite} tool suites are used as supporting tools~\cite{BurdyEtal05-STTT}.
+Driven by these requirements and our expertise, the Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2} tool suite is used as supporting tool~\cite{BurdyEtal05-STTT}.
 The JML2 tool suite includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit test generator~\cite{Cheon2002}.
-ESC/Java2 supports automatic, modular, static reasoning about the correctness of an implementation against its formal specification.
-These tool suites do not require a complete specification to be used effectively.
+A complete specification is not required to use JML2 tool suite effectively.
 
@@
 The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
-As ESC/Java2 is fully automated, it is not sound, nor does it cover the whole Java and JML languages, and is thus not complete, but there is support for warning the user about both of these situations~\cite{KiniryEtal06}.
 
@@
 Functional test suites, used as precise requirements or user stories definition, are applied in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
-Unit tests and TDD are mentioned, and often completely accepted, in most of the AMs defined so far.
+Unit tests and TDD are mentioned, and often included, in most of the AMs defined so far.
 
@@
 Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.
-Automated test suites guarantees that the functionalities and behaviour already implemented in the code are not compromised by refactorings.
 
@@
 % composition is not enough
-Is this enough to deal with formal and informal artifacts?
 FDD is used for the formal specified parts of the system, whether TDD is used for the remaining ones: a ``composition'' of formal and agile verification practices.
-Yet, it is not enough.
-There are several problems that still need to be addressed.
+Yet there are several problems that still need to be addressed.
 No communication between the two worlds is permitted, it is not defined how to formally specify a module after it has been implemented and verified in a non formal way.
@@
 The \SB protocol is asynchronous and packet-based; the packets have a fixed length.  
-The protocol is separated in multiple layers to ease its implementation in the driver, since mixing up different abstraction levels usually leads to overly complex and unmanageable code.  
-The layers identified in protocol are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable values and reaction constraints on output packets, based on previously transmitted input packets.
-For space reason, the first three layers are analyzed in the case study, considering only the output sensor data packet.
+The protocol is separated in multiple layers to ease its implementation in the driver.
+The layers identified are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable values and reaction constraints on output packets, based on previously transmitted input packets.
+The first three layers are analyzed in the case study, considering only the output sensor data packet.
 
 % packet structure
-The packet byte and info structures of the output sensor data packet reflect the board capabilities and built in sensors. 
+The packet byte and info structure of the output sensor data packet reflect the board capabilities and built in sensors. 
 A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metadata describing the sensors and the board state.
@@
 The single packet rules delimit the boundaries of the values obtainable from the packet: each defined sensor represented in the packet, as well as the metadata describing the \STSB and the packet and frame contents, are constrained by a range of acceptable values.
-There are also rules affecting more than one value (i.e.
-if a specific sensor is active the reading value must be in a defined range), and rules specifying a correct sequence of frames (i.e.
-only one switch in sampling rate is allowed in the frame sequence of a single packet).
+There are also rules affecting more than one value, and rules specifying a correct sequence of frames.
 
@@
 
-% how many of them do we need?
-Theoretically a separate simulator is needed for each protocol layer to specify, develop and test each layer in isolation.  
-But, because of limited resources and effort available, the layer structure is not matched by corresponding simulators.
-Two simulators are available, they have been built concurrently with the specification, the driver implementation, and the tests.
-
 % simulators description
+The layer structure is not matched by the corresponding simulators.
 The driver API is splitted in two abstraction layers: the general high-level interface, that exposes the main functionalities of the board and the main contents of the packets, and the lower level implementation that parses the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
-The simulators are built according to these two abstraction layers.
-The high-level simulator implements the interfaces providing the methods to deal with an abstracted \SB: the simulator knows nothing
-about the details of the real format of the binary data streams.
+The simulators are built according to these abstraction layers.
+The high-level simulator implements the interfaces providing the methods to deal with an abstracted \SB.
 The low-level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to parse or generate.
@@
 The JML specifications are used to generate tests.
-The resulting test package structure is shown in Figure \ref{fig:class_diagram_generatedtest}.  
-The package dependencies are defined by the specifications; for instance, \lil{SimulatorDriverST} are generated tests, they test the implementation \lil{SimulatorDriver}, and are generated using the corresponding specification \lil{SimulatorDriverS}.
-The generated tests do not maintain the inheritance structure of the code and the specification.
+The resulting test package structure is closely related and reflects the overall package structure; each package have a corresponding generated test package.
 
-\begin{figure}[htb!]
-  \centering
-  \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
-  \caption{Generated test packages class diagram.}
-  \label{fig:class_diagram_generatedtest}
-\end{figure}
+%\begin{figure}[htb!]
+%  \centering
+%  \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
+%  \caption{Generated test packages class diagram.}
+%  \label{fig:class_diagram_generatedtest}
+%\end{figure}
 
@@
 
-% simulator importance
-Simulators were essential to test the \STSB driver, since the physical board was not available.
+% simulator adoption
 The simulators are adopted in both the handmade and the generated test suite; both the suites need a proper set of stubs in order to be executed.
@@
 The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, the system does not depend on them, they are only used in object initialization or in object setups performed during unit tests.
-In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes.
+In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes (white box testing).
 The used statement coverage tool is not able to distinguish these code blocks, so this effect is unavoidable. 
@@
 The qualitative difference of the two test suites are understood observing how the suites are built. 
-An handmade unit test usually requires objects initialization (it involves several method calls), the call to the method under test, and the following assertions to verify the expected state change on the objects involved.
+An handmade unit test usually requires objects initialization, one or more method calls to the modules under test, and the following assertions to verify the expected state change on the modules involved.
 A generated unit test has a more focused and limited scope: only one method call is performed in each test, therefore it is difficult to explore complex behaviours. 
-With a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
+Theoretically with a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
 The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.

### 2010-08-23 papers/AgileFormality/paper.tex
finished and formatted. to bo read and than submitted

@@
-%\documentclass[english]{lni}
-\documentclass{article}
+\documentclass[english]{lni}
 \usepackage{latin1}
@@
 \author{Vieri del Bianco and Dragan Stosic\\
-  UCD CASL: Complex and Adaptive Systems Laboratory and\\
-  School of Computer Science and Informatics,\\
   University College Dublin, Ireland\\
   \texttt{vieri.delbianco@ucd.ie} and \texttt{dragan.stosic@gmail.com}\\
-  \\
   Joseph R.~Kiniry \\
@@
 
-% ======================================================================
 \section{Introduction}
@@
 % formal methodology constraints
-To perform this case study, an appropriate FM must be chosen.  
+An appropriate FM must be chosen.  
 It must be able to support a specification that is built incrementally, consequently formal methods that demand a complete specification must be avoided.
@@
 % the chosen formal methodology
-Driven by these requirements and our expertise, the Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2} tool suite is used as supporting tool~\cite{BurdyEtal05-STTT}.
+The Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2} tool suite is used as supporting tool~\cite{BurdyEtal05-STTT}.
 The JML2 tool suite includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit test generator~\cite{Cheon2002}.
@@
 % JML introduction
-JML is the de facto standard specification language for describing Java programs behaviour.  
-JML is a rich behavioural interface specification language (BISL) and thus focuses on the modular specification of classes and objects in which inheritance is used solely for behavioural subtyping~\cite{LeavensBakerRuby99}.
+JML is a rich behavioural interface specification language (BISL) and focuses on the modular specification of classes and objects~\cite{LeavensBakerRuby99}.  
 JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants.
@@
 specifications of their own, and then an implementation is written conforming to that specification~\cite{Meyer92b}.  
-The implementation is checked by whatever means appropriate using some subset of the aforementioned tools.  
-At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact, a process called Contracting-the-Design.
+At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact.
 
@@
 The former compiles executable assertions into runtime checks.  
-The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave (i.e., crash in a variety of ways) and conforms to its specification (lightweight functional correctness).
+The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave and conforms to its specification (lightweight functional correctness).
 
@@
 
-\subsection{Data stream and protocol description}
+\subsection{Data stream, protocol description and protocol specification}
 \label{subsec:data_stream_and_protocol_description}
@@
 A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metadata describing the sensors and the board state.
-The internal structure of the packet is strictly fixed, even if the data streams have different data rates. 
-This choice originates from a trade-off between two constraints: the efficiency of the transmission (redundant data to be minimized) and the complexity to build a packet (complexity to be minimized, since the processing capabilities in the \STSB are limited).
+The internal structure of the packet is strictly fixed.
 
@@
 A packet is internally composed by 82 frames. 
-A frame accommodates data from the fast and medium data rate streams (there are 4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
-Fast, medium and slow data rates have not a common multiple; a packet, capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
-\begin{itemize}
-\item Slow data rates: slow data is located at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
-\item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame.
+A frame accommodates data from the fast and medium data rate streams (4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
+%Fast, medium and slow data rates have not a common multiple; a packet, capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
+%\begin{itemize}
+%\item Slow data rates: slow data is located at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
+%\item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame.
 The medium data rate sample channel is a piece of info extracted from the frame description (frame metadata), including the possibility to have no data at all.
-E.g. a frame contain a medium data rate sample from channel 0, the following frame a sample from channel 5, the following frame no sample at all, etc.
-\end{itemize}
+%E.g. a frame contain a medium data rate sample from channel 0, the following frame a sample from channel 5, the following frame no sample at all, etc.
+%\end{itemize}
 
@@
 
-\subsection{Data stream and protocol specification}
-\label{subsec:data_stream_and_protocol_specification}
+%\subsection{Data stream and protocol specification}
+%\label{subsec:data_stream_and_protocol_specification}
 
@@
 Verification is obtained through an handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream.
-The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confident on how the stream is decomposed into packets.
-Besides, the packet byte structure is not stable: the internal byte structure changed several times during development.
+%The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confident on how the stream is decomposed into packets.
+The packet byte structure is not stable: the internal byte structure changed several times during development.
 Therefore, packet byte structure is informally specified.
@@
 
-\subsection{On simulators}
-\label{subsec:on_simulators}
+\paragraph*{On simulators}
+%\label{subsec:on_simulators}
 
@@
 Simulators are used to test parts of the system minimizing dependencies: a simulator can be used by upper layers, with no need to provide the functionalities of the lower layers. 
-The simulators proved essential during the development, since the physical board was not available.
+%The simulators proved essential during the development, since the physical board was not available.
 
@@
 
-\paragraph*{On simulators, again}
+\paragraph*{On simulators}
 
 % simulator roles in tests
-The simulators are adopted in the handmade test suits as stubs.
+Simulators are adopted in the handmade test suits as stubs.
 The resulting structure is shown in Figure \ref{fig:class_diagram_streamdriver_test}.  
@@
 Various coverage metrics exist, the test suites are analyzed with statement code coverage (one of the simplest types)\footnote{The tool used to obtain statement code coverage metrics is \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage reports whether each statement is encountered during execution.
-Statement code coverage is easy to implement, but lacks the capability to explore the control flow graph associated to the code, that is, decision branches are not explored. 
 The coverage results are reported in the Table \ref{tab:statement_code_coverage}.
@@
 
-\subsection{Retrospective on test cases effectiveness}
-\label{subsec:test_cases_retrospectives}
+%\subsection{Retrospective on test cases effectiveness}
+%\label{subsec:test_cases_retrospectives}
 
@@
 The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, the system does not depend on them, they are only used in object initialization or in object setups performed during unit tests.
-In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes (white box testing).
-The used statement coverage tool is not able to distinguish these code blocks, so this effect is unavoidable. 
+%In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes (white box testing).
+%The used statement coverage tool is not able to distinguish these code blocks, so this effect is unavoidable. 
 The effect is mainly seen in low \lil{SimulatorDriver} code coverage figures: the \lil{SimulatorDriver} package has many utility non public methods.
@@
 This effect is visible in \lil{Driver} coverage figures: \lil{Driver} package contains interfaces (that are not counted at all) and some very simple real class (exceptions) that are not thoroughly tested.
-The net result is extremely low statement code coverage figures because the very simple classes are the only ones that are actually covered.
+%The net result is extremely low statement code coverage figures because the very simple classes are the only ones that are actually covered.
 
@@
 A generated unit test has a more focused and limited scope: only one method call is performed in each test, therefore it is difficult to explore complex behaviours. 
-Theoretically with a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
+%Theoretically with a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
 The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.
@@
 The first \STSB prototype builds packets with no data, but the packets are built in a way not consistent with the specification.  
-The board errors are detected by simply using the driver implementation, the \lil{StreamDriver} package.
+The board errors are detected by using the driver implementation, the \lil{StreamDriver} package.
 
@@
 One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions not initially covered by the test cases, but that occasionally showed up during real use.
-One error was discovered in the low-level system driver used by the developed driver (the system driver is outside our development scope): to eliminate the error, an upgrade of the low-level driver to a new beta version was necessary.
-No other errors were detected on the driver side.
+%One error was discovered in the low-level system driver used by the developed driver (the system driver is outside our development scope): to eliminate the error, an upgrade of the low-level driver to a new beta version was necessary.
+%No other errors were detected on the driver side.
 
@@
 The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
+It also randomly generated completely invalid packets when overheated.
 The formal JML specifications and the runtime assertion checker correctly identified these errors.
-4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, and are expected to be corrected in the following release. 
-No other errors, except the mentioned ones, have been discovered up to now.
+4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, as well as the invalid packets.
+
+\section{Retrospective on ''Mole`` practices effectiveness}
+\label{subsec:test_cases_retrospectives}
+
+% preconditions on projects
+''Mole`` practices are very easy to be applied and work with.
+Many projects could benefit from the practices, not only the unstable ones that requires a certain level of formality.
+Even projects with no formal specification requirements at all could benefit from the practices, since some system properties are easy to specify and could actually reduce the test suites development effort.
+
+% precondition on formal methodology and associated formal suite
+The cornerstone of the practices is the FM with its supporting tools, and the precondition on the tools are not easy to be fulfilled.
+Methodology that require a complete formalization are impossibile to be applied.
+Beside, it is not clear if a tool suite that does not provide a run time verification environment could be used succesfully, not it is clear whether formal methods not based on the contract paradigma (preconditions and postconditions) could be applied at all.
+These are open questions that will nedd to be addressed in future works to assess the real appliability of the ''Mole`` practices.
+
+% when to use what
+There no strict rules to determine which of the ''Mole`` practices to use to face a specific problem, but only advices.
+TDD if we don't need anu specification, FDD with generated tests if the specification are not too difficult and the scenario is simple enough to be covered by generated tests, FDD with hand made tests if the scenarios is complex, TDF when the specification is complex.
+The experience gathered is yet not enogh to distill idioms or patterns.
+
+% what is working and what is not
+FDD with hand made tests is amazing, it is as natural to work with as TDD. 
+FDD with generated tests can be annoying because of the time needed to run the test suite, often only part of the test suite is executed so that the development loops take not too long.
+TDF is great with complex preconditions, but really difficult with postconditions, especially if the soecification is made on top of already existing code. 
+When developing postconditions after code, the test is not expected to fail, hence the feedback is limited and the specification could not be as strict as it was needed. 
+In this specific case TDF effectiveness is limited.
 
@@
 
-\bibliographystyle{lni}
+%\bibliographystyle{lni}
 

### 2010-08-24 papers/AgileFormality/paper.tex
Revised sections 1, 5, and 6.

@@
   \usepackage{epstopdf}
-  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Vieri del Bianco and Joseph R.
-Kiniry and Dragan Stosic"]{hyperref}
+  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true,
+  colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue,
+  pdfauthor="Vieri del Bianco and Joseph R. Kiniry and Dragan
+  Stosic"]{hyperref} 
 \else
@@
 \newcommand{\STUs}{\ST \emph{Units}\xspace}
+\newcommand{\mole}{``\emph{Mole}''\xspace}
 
@@
 
-\title{Agile Formality: A ``Mole'' of Software Engineering Practices}
+\title{Agile Formality: A \mole of\\Software Engineering Practices}
 
@@
   \texttt{vieri.delbianco@ucd.ie} and \texttt{dragan.stosic@gmail.com}\\
+  \\
   Joseph R.~Kiniry \\
@@
 
-Members of the agile programming and formal methods communities do not always see face-to-face.  
-These two communities often do not talk to or learn from each other.  
-Only recently, as highlighted by the September 2009 issue of IEEE Software, the IFIP workshop on balancing agility and formalism in software engineering, and the first edition of the international workshop for formal methods and agile methods, ideas from the two communities begun to be synthesized.
-While the problem-solving approaches and psychological attitudes of members of the two communities differ widely, we exploit this clash of viewpoints, exploiting a new development processes that actually blends, rather than mashes together, best practices from the two worlds.
-This paper summarizes our process and a supporting complex case study, showing that it is not only possible, but tasty, to combine the ``chili pepper'' of formal methods and the ``chocolate'' of agile programming, thus producing a tasty ``Mole'' (as in the highly-spiced Mexican sauce) of software engineering practices.
+  Members of the agile programming and formal methods communities do
+  not always see eye-to-eye.  These two communities often do not talk
+  to or learn from each other.  Only recently, as highlighted by the
+  September 2009 issue of IEEE Software, the IFIP workshop on
+  balancing agility and formalism in software engineering, and the
+  first edition of the international workshop for formal methods and
+  agile methods, ideas from the two communities begun synthesize.
+  While the problem-solving approaches and psychological attitudes of
+  members of the two communities differ widely, we exploit this clash
+  of viewpoints, creating a new development processes that actually
+  blends, rather than mashes together, best practices from the two
+  worlds.  This paper summarizes our process and a supporting complex
+  case study, showing that it is not only possible, but tasty, to
+  combine the ``chili pepper'' of formal methods and the ``chocolate''
+  of agile programming, thus producing a tasty \mole (as in the
+  highly-spiced Mexican sauce) of software engineering practices.
 
@@
 % agile and formal difficult to blend: mind attitudes
-Agile and formal development methodologies usually do not blend together well.  
-This is because of several reasons, the most important of which is often characterized as a radical difference in psychological attitudes about software development.
+Agile and formal development methodologies usually do not blend
+together well.  This is because of several reasons, the most important
+of which is often characterized as a radical difference in
+psychological attitudes about software development.
 
 % FM
-Formal methodologists often favour an in-depth, think-first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall-style of development.  
-Once a formalization is developed, development and verification proceed, in an automated and interactive fashion.  
-Consequently, projects that use formal methodologies (FMs) typically focus on critical systems with fixed requirements and somewhat flexible deadlines (``We'll ship it when it is right.'').
+Formal methodologists often favour an in-depth, think-first approach,
+where the problem is understood, formalized, and solved; usually, but
+not always, adopting a waterfall-style of development.  Once a
+formalization is developed, development and verification proceed in an
+automated and interactive fashion.  Consequently, projects that use
+formal methodologies (FMs) typically focus on critical systems with
+fixed requirements and somewhat flexible deadlines (``We'll ship it
+when it is right.'').
 
 % AM
-Agile methodologists favour an highly incremental and iterative approach, specifically tailored to cope with changing requirements and precise deadlines.  
-In projects that use agile methodologies (AMs), it is often the case that the problem is only partially understood, focusing only on the aspects that must be implemented in the current development iteration.
-The solution initially developed is usually not optimal, but as it is refined through continuous code refactorings, its quality and validity improve.  
-Test suites are used for several purposes: the most important are system verification, system documentation, and requirement specification, and they are strictly handmade.  
-AMs are typically used in development settings where changing requirements and rapid delivery of the product are paramount.
+Agile methodologists favour an highly incremental and iterative
+approach, specifically tailored to cope with changing requirements and
+precise deadlines.  In projects that use agile methodologies (AMs), it
+is often the case that the problem is only partially understood, and
+there is a focus only on the aspects that must be implemented in the
+current development iteration.  The solution initially developed is
+usually not optimal, but as it is refined through continuous code
+refactorings, its quality and validity improve.  Test suites are used
+for several purposes, the most important of which are system
+verification, system documentation, and requirement specification, and
+they are strictly handmade.  AMs are typically used in development
+settings where changing requirements and rapid delivery of the product
+are paramount.
 
 % problems would benefit form agile and formal mix, characterize the problem
-The two worlds seem irreconcilable.  
-Nevertheless, problems exist that would greatly benefit from both of them~\cite{Black2009}.  
-The problem class in which we have particular interest have \emph{unstable requirements}, are \emph{constrained by deadlines that cannot be postponed}, but also have \emph{subsystems that must be formally specified and verified}.
+The two worlds seem irreconcilable.  Nevertheless, problems exist that
+would greatly benefit from both of them~\cite{Black2009}.  The problem
+class in which we have particular interest have \emph{unstable
+  requirements}, are \emph{constrained by deadlines that cannot be
+  postponed}, but also have \emph{subsystems that must be formally
+  specified and verified}.
 
@@
 % IEEE terminology
-In this paper the IEEE 610~\cite{Jay1990} standard terminology for validation and verification is used.  
-Paraphrasing the standard, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed, according to the specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
+In this paper the IEEE 610~\cite{Jay1990} standard terminology for
+validation and verification is used.  Paraphrasing the standard,
+\emph{verification} is the process of evaluating software to determine
+whether the products of a given development phase satisfy the
+conditions imposed, according to the specifications (``you build the
+system in the right way''); \emph{validation} is the process of
+evaluating software to determine whether it satisfies specified
+requirements ensuring it meets the user's needs (``you build the right
+system'').
 
 % FM terminology
-Within the FMs community, verification and validation have different meanings than in the IEEE 610 standard.
-Loosely \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically verify that a system conforms to its specification; \emph{validation} (in the paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
-In the present paper both perspectives are respected and acknowledged.
+Within the FMs community, verification and validation have different
+meanings than in the IEEE 610 standard.  Loosely \emph{verification}
+(in the paper \emph{formal verification}) means the use of formal
+methods to formally, statically verify that a system conforms to its
+specification; \emph{validation} (in the paper \emph{formal
+  validation}) means the use of traditional software engineering
+practices to check, statically or dynamically, that a system conforms
+to its specification (whether in the form of tests, requirements,
+etc.) by hand or via execution under some execution scenarios.  In the
+present paper both perspectives are respected and acknowledged.
 
 % tests terminology
-Tests are usually classified by dichotomies: unit vs. functional, black-box vs. white-box, testing vs. design, tester vs. coder. 
-Yet testing doesn't always fit in the dichotomies, especially in the case of agile testing approaches~\cite{Beck2007}.
-Black-box vs. white-box dichotomy is particularly problematic, since the box boundaries change the perspective.
-In the paper, functional tests are always black-box tests, coverage tests are always white-box tests, generated tests are always black-box tests (the box is the module specified); other strict rules do not hold.
+Tests are usually classified by dichotomies: unit vs. functional,
+black-box vs. white-box, testing vs. design, tester vs. coder.  Yet
+testing does not always fit into these dichotomies, especially in the
+case of agile testing approaches~\cite{Beck2007}.  The black-box
+vs. white-box dichotomy is particularly problematic, since box
+boundaries change one's perspective.  In this paper, functional tests
+are always black-box tests, coverage tests are always white-box tests,
+generated tests are always black-box tests (the box is the module
+specified); other strict rules do not hold.
 
 % development terminology
-Additionally, ``...development of a system,'' more precisely means ``...analysis, design, development, verification, formal validation of a system''.
+Additionally, ``...development of a system,'' more precisely means
+``...analysis, design, development, verification, formal validation of
+a system.''
 
@@
 
-Proposing new practices with no supporting evidence is an uninteresting proposition.  
-A complex case study is mandatory to test bench the proposed practices.  
-Our case study focuses on the development of a device driver to control and communicate with an embedded custom circuit board equipped with more than a dozen sensors.
-The communication protocol with the board is asynchronous and packet-based and the board is novel and custom-designed.  
-The protocol is focused on controlling the board and reporting sensor measurements and board status.
+Proposing new practices with no supporting evidence is an
+uninteresting proposition.  A complex case study is mandatory to
+test-bench the proposed practices.  Our case study focuses on the
+development of a device driver to control and communicate with an
+embedded custom circuit board equipped with more than a dozen sensors.
+The communication protocol with the board is asynchronous and
+packet-based and the board is novel and custom-designed.  The protocol
+is focused on controlling the board and reporting sensor measurements
+and board status.
 
-This software driver has been developed using our process while statisfying seemingly contradictory requirements: (1)~the protocol implementation must be thoroughly verified; (2)~parts of the protocol must be formally specified; (3)~a board simulator must be designed, implemented, and verified to match the specifications of, and actual behaviour of, the physical circuit board; (4)~if board components change during development, the associated formal specifications and the simulator whose behavior reflects such changes must be updated accordingly; (5)~strict software delivery deadlines.
-Consequently, these process requirements constitute an environment that benefits from both agile and formal approaches to software development.
+This software driver has been developed using our process while
+satisfying seemingly contradictory requirements: (1)~the protocol
+implementation must be thoroughly verified; (2)~parts of the protocol
+must be formally specified; (3)~a board simulator must be designed,
+implemented, and verified to match the specifications of, and actual
+behaviour of, the physical circuit board; (4)~if board components
+change during development, the associated formal specifications and
+the simulator whose behavior reflects such changes must be updated
+accordingly; (5)~there are strict software delivery deadlines.
+Consequently, these process requirements constitute an environment
+that benefits from both agile and formal approaches to software
+development.
 
-A novel development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal practices.  
-In the following the development process and the practices are summarized, and their successful application to the case study is reported, yielding several promising results.
+A novel development process has been drafted to cope with the
+contrasting requirements of the case study, blending agile and formal
+practices.  In the following, the development process and the practices
+are summarized, and their successful application to the case study is
+reported, yielding several promising results.
 
@@
 % what is needed, high-level
-To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the AMs must be maintained, and formal validation methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, validate the implementation against the specification), must be adapted to a highly iterative one.
+To enable the integration of the two worlds, the highly iterative and
+incremental approach found in most of the AMs must be maintained, and
+formal validation methodologies, traditionally used in a waterfall
+development process (formally specify the system, implement the
+system, validate the implementation against the specification) must
+be adapted to a highly iterative one.
 
 % what must be considered: TDD
-The most common software verification practices in AMs must be considered, when trying to integrate FMs and AMs, since they are a fundamental part of the development process. 
-The most significant verification related practice is Test Driven Development (TDD)~\cite{Beck2003}: TDD is a software development technique, originally defined in Extreme Programming (XP)~\cite{Beck2004} methodology, characterized by writing unit tests before writing the code.
+The most common software verification practices in AMs must be
+considered when trying to integrate FMs and AMs since they are a
+fundamental part of the development process.  The most significant
+verification related practice is Test Driven Development
+(TDD)~\cite{Beck2003}.  TDD is a software development technique,
+originally defined in the Extreme Programming (XP)~\cite{Beck2004}
+methodology and is characterized by a process that focuses on writing
+unit tests for a module before writing the code of said module.
 
 % test driven kinds
-The test driven approach relies on writing the tests before the actual implementation. 
-It is applicable to a very small scale (TDD) yielding to activity cycles as short as a few minutes.
-It is applicable to a large scale as well (requirements immediately translated into functional tests), yielding to activity cycles as long as a full delivery iteration (usually no longer than a couple of months).
-The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other AMs as well.
-It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
+The test driven approach relies on writing tests before
+implementations.  It is applicable to a scall and large-scale cycles.
+In the former case, very small scale TDD yields activity cycles as
+short as a few minutes.  In the latter, requirements are immediately
+translated into functional tests resulting in activity cycles as long
+as a full delivery iteration, which is usually no longer than a couple
+of months.  The test driven approach is a cornerstone of XP, but it is
+so popular that has been adopted in many other AMs as well.  It is
+also considered a good software development technique when used on its
+own, as it pre-dates XP and AMs considerably, regardless of the
+enclosing development process.
 
 % previous related work
-Some tentative attempts to reconcile FMs and AMs have been developed.
-The recurrent approach of the attempts consists in removing the handmade test suites, substituting them with different kinds of automated formal validation based on the system formal specification. 
-The automated validation can be as simple as enabling runtime verification of assertions or automatically generating a test suite based on the system properties, or as complex as actually proving the whole system behaviour through static model and code checkers or theorem provers (formal verification).
+Some tentative attempts to reconcile FMs and AMs have been developed
+by others.  The reoccuring theme of these attempts is that idea that
+handmade test suites are no longer used, but instead are replaced by
+different kinds of automated formal validation based on a system-level
+formal specification.  Automated validation can be as simple as
+enabling runtime checking of assertions or automatically generating a
+test suite based on the system properties, or as complex as proving
+the whole system's behaviour through formal verification.
 
 % a quick analysis of the related works
-Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an iterative and incremental process to generate X-Machines models, an extension of Finite State Machines. 
-The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines FM. 
-The development process proposed in~\cite{Herranz2003b} by Herranz and Moreno-Navarro is quite similar: they propose an iterative process to model a system using SLAM-SL, an object-oriented specification language, and its tools.
-Some XP practices are fully adopted or considered compatible, as pair-programming, iterative and incremental development, system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
-A different FM, but same approach, is found also in~\cite{Suhaib2005} where XFM methodology is explained.
+Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an
+iterative and incremental process to generate X-Machines models, an
+extension of Finite State Machines.  The development process impose a
+complete specification of the system, and its verification is done
+exclusively using the tools provided by the X-Machines FM.  The
+development process proposed by Herranz and
+Moreno-Navarro~\cite{Herranz2003b} is quite similar.  They propose an
+iterative process to model a system using SLAM-SL, an object-oriented
+specification language supported by tools.  Some XP practices are
+fully adopted or considered compatible, like pair-programming,
+iterative and incremental development, and system refactoring; while
+tests are completely replaced by the verification tools provided by
+the SLAM suite.  A different FM, but similar approach, is found also
+in the work of Suhaib~\cite{Suhaib2005} where the XFM methodology is
+used.
 
 % previous work: a common approach
-As already mentioned, in all these attempts, the test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
-The requisite to enable these approaches is, at minimum, a specification of all the relevant aspects of the system under development, to enable simple verification techniques such as runtime verification (assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
-Whether a complete and sound system specification, supported by an appropriate formal language (to specify model and properties) and a constrained programming language, can be able to statically prove properties about the system~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+As already mentioned, in all of these attempts, test suites are
+automatically generated using the system formal specification, or are
+completely replaced by formal verification, typically through the use
+of theorem provers and/or static checkers.  The major precondition on
+the use of these approaches is, at the minimum, a specification of all
+the relevant aspects of the system under development.  Such a
+specification enables simple verification techniques such as runtime
+checking (via assertions and design by contract~\cite{Meyer1997} using
+preconditions, postconditions and invariants) and the automatic
+generation of test suites~\cite{Cheon2002,Cheon2004,Cheon2005}.  Via a
+complete and sound system specification, supported by an appropriate
+formal language (to specify the system model and its properties)
+coupled with a (possibly restricted) programming language, one can
+statically prove properties about the system using a variety of tools
+and
+techniques~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
 % previous work: why is not satisfying
-The approach is simple: the verification, documentation and requirement specification purposes of the tests are moved on to the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
-The problem is that this requirement is hard to fulfil, since the effort required to write a complete formal specification of the system, in most real world complex cases, is usually greater than writing a suite of tests~\cite{Glass2002}.
-If a complete formal specification is not built, all the previous methodologies share a common problem: the parts informally specified are not verified at all.
-These can obviously be verified with the traditional methods, but the loops between development artifacts and related activities (code, tests, design, refactoring) are neither detailed nor enforced.
-The loops and relations between the different development artifacts are the inner engine of many AMs, they impose an iterative and incremental pace to the development process.
+The underlying idea of this approach is simple: what was once the
+purpose of handwritten tests (verification, documentation, and
+requirement specification) are now the responsibility of the formal
+specifications, and a complete formal specification of the system
+under development is needed to apply these methodologies.  The problem
+is that this requirement is very difficult to fulfill since the effort
+required to write a complete formal specification of the system in
+most real-world complex cases is usually greater than writing a suite
+of tests~\cite{Glass2002}.
+
+If a complete formal specification is not available, all of the
+previous methodologies share a common problem: the parts informally
+specified are not verified at all.  These parts of the system are
+consequently verified with traditional methods, but the connections
+between development artifacts and related activities (code, tests,
+design, refactoring) are neither detailed nor enforced.  This is
+problematic, as the loops and relations between the different
+development artifacts are the inner engine of many AMs, as they impose
+an iterative and incremental pace to the development process.
 
 % outsider
-Liu in~\cite{Liu2010} takes a different approach where SOFL methodology~\cite{Liu2004} is merged into agile processes. 
-SOFL methodology is based on a three step specification approach: (1) informal specification, (2) semi-formal specification, (3) formal specification. 
-The agile adaptation consists mainly in shorter loops between activities, and into a reduced set of formal specifications.
-The formal specifications are used only to help understanding ambiguous statements in the semi-formal specification, and are not maintened while the system evolves; testing and inspections are based on the semi-formal specification.
-The formal specification are partial, which is interesting, but used only for documentation purposes and than discarded, they are not used at all in the verification of the system.
+Liu takes a different approach wherein the SOFL
+methodology~\cite{Liu2004} is merged into agile
+processes~\cite{Liu2010}.  The SOFL methodology is based on a three
+step specification approach: (1) informal specification, (2)
+semi-formal specification, (3) formal specification.  The agile
+adaptation consists mainly in introducing shorter loops between
+activities and reducing the size of the set of formal specifications.
+The formal specifications are used only to help understand ambiguous
+statements in the semi-formal specification and are not maintained
+while the system evolves.  Testing and inspections are based on the
+semi-formal specification.  The formal specification are partial and
+are only used for documentation purposes and then discarded---they are
+not used at all in the verification of the system.
 
 % what we want to achieve
-The objectives we want to achieve are the following ones: to formally specify only parts of the system, in order to cope with constrained resources and unstable requirements; to draft an highly iterative and incremental development process that blends formal and agile practices.
+In contrast, our objective is to formally specify only those parts of
+the system that must be verified (in order to cope with constrained
+resources and unstable requirements) and to develop and refine a
+highly iterative and incremental development process that blends
+formal and agile practices.
 
 % how we want to achive it
-The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in AMs (especially in XP), to solve this problem in an highly iterative development process.
-We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how handmade and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
-That is, we will show how to blend, and not compose, formal validation and agile verification practices: a ``Mole'' of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers (agile) and chocolate (formal), served with meat (working system).}.
+The main open problem unaddressed in the literature is how to connect
+development artifacts of the two worlds of FMs and AMs together.  To
+answer this challenge we define activity cycles, similar to what is
+found in AMs (especially in XP), to solve this problem in an highly
+iterative development process.  We show how tests drive the formal
+specification, how the formal specifications drive test and code
+development, how handmade and automated tests coexist and support each
+other, how the unspecified parts of the system are incrementally
+specified.  That is, we show how to blend, and not just tack together,
+formal validation and agile verification practices: a \mole of
+verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly
+  spiced Mexican sauce made chiefly from chili peppers (agile) and
+  chocolate (formal), served with meat (working system).}.
 
@@
 
-``Mole'' practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and validated.
+\mole practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and validated.
 
@@
 The protocol specification must be enhanced incrementally and its complete specification cannot be provided.  
-This makes the problem an ideal candidate to test our ``Mole'' of practices.
+This makes the problem an ideal candidate to test our \mole of practices.
 
@@
 
-\subsection{``Mole'' practices}
+\subsection{\mole Practices}
 
@@
 % all together blends into a mole
-When the ``Mole'' verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; the complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
+When the \mole verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; the complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.
 
 % =====================================================================
-\section{``Mole'' verification practices applied}
+\section{\mole Verification Practices Applied}
 \label{sec:a_complex_case_study}
 
-\subsection{Data stream, protocol description and protocol specification}
+\subsection{Data Stream, Protocol Description and Protocol Specification}
 \label{subsec:data_stream_and_protocol_description}
@@
 It has been built starting from handmade tests only (TDD development cycles).
-Later on JML annotations have been added incrementally, using a mix of ``Mole'' practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the ``Mole'' practices have been used according to the artifacts involved.
+Later on JML annotations have been added incrementally, using a mix of \mole practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the \mole practices have been used according to the artifacts involved.
 
@@
 The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check the formally specified properties at runtime.
-Code and tests have been initially implemented using TDD, than using all the ``Mole'' verification practices.
+Code and tests have been initially implemented using TDD, than using all the \mole verification practices.
 
@@
 
-\subsection{JML specification examples}
+\subsection{JML Specification Examples}
 \label{subsec:a_jml_specification_example}
@@
 
-\subsection{Test cases}
+\subsection{Test Cases}
 \label{subsec:test_cases}
@@
 
-\subsection{Test cases on the job}
+\subsection{Test Cases on the Job}
 \label{subsec:test_cases_on_the_job}
@@
 
-\section{Retrospective on ''Mole`` practices effectiveness}
+\section{Retrospective on \mole Practices Effectiveness}
 \label{subsec:test_cases_retrospectives}
@@
 % preconditions on projects
-''Mole`` practices are very easy to be applied and work with.
-Many projects could benefit from the practices, not only the unstable ones that requires a certain level of formality.
-Even projects with no formal specification requirements at all could benefit from the practices, since some system properties are easy to specify and could actually reduce the test suites development effort.
+We believe that \mole practices are relatively easy to introduce and
+use.  Many kinds of projects could benefit from these practices, not
+only those that have unstable requirements but also demands for a
+certain level of formality.  Even projects with no requirements for
+formal specification at all benefit from these practices, since some
+simple or obvious system properties are easy to specify and, though
+their introduction, there is a corresponding reduction in the test
+suite development effort.
 
 % precondition on formal methodology and associated formal suite
-The cornerstone of the practices is the FM with its supporting tools, and the precondition on the tools are not easy to be fulfilled.
-Methodology that require a complete formalization are impossibile to be applied.
-Beside, it is not clear if a tool suite that does not provide a run time verification environment could be used succesfully, not it is clear whether formal methods not based on the contract paradigma (preconditions and postconditions) could be applied at all.
-These are open questions that will nedd to be addressed in future works to assess the real appliability of the ''Mole`` practices.
+The cornerstone of the practices is the formal methodology and its
+supporting tools, and the precondition on the tools' utility is not
+always easily fulfilled.  Methodologies that require a complete
+formalization are nearly impossibile to apply.  In addition, it is not
+clear if a tool suite that does not provide a runtime verification,
+either via runtime checking or simulation, can be broadly used
+succesfully, nor it is clear whether formal methods that are not based
+on the contract paradigm are useful in this context.  These are open
+questions that will need to be addressed in the future to determine
+the broad, concrete applicability of the \mole practices.
 
 % when to use what
-There no strict rules to determine which of the ''Mole`` practices to use to face a specific problem, but only advices.
-TDD if we don't need anu specification, FDD with generated tests if the specification are not too difficult and the scenario is simple enough to be covered by generated tests, FDD with hand made tests if the scenarios is complex, TDF when the specification is complex.
-The experience gathered is yet not enogh to distill idioms or patterns.
+There are no strict rules to determine which of the \mole practices to
+use when facing a specific problem.  We have found that experience
+trumps cookbooks, so we can only give some advice.  We suggesting
+using TDD there is no need for any specifications at all, FDD with
+generated tests if the specification are not too difficult and system
+scenarios is simple enough to be covered by generated tests, FDD with
+hand made tests if the scenarios are complex, and TDF when the
+specification is complex.  We feel that the experience we have
+gathered is not yet enough to distill richer idioms or patterns---such
+is future work.
 
 % what is working and what is not
-FDD with hand made tests is amazing, it is as natural to work with as TDD. 
-FDD with generated tests can be annoying because of the time needed to run the test suite, often only part of the test suite is executed so that the development loops take not too long.
-TDF is great with complex preconditions, but really difficult with postconditions, especially if the soecification is made on top of already existing code. 
-When developing postconditions after code, the test is not expected to fail, hence the feedback is limited and the specification could not be as strict as it was needed. 
-In this specific case TDF effectiveness is limited.
+Upon reflect, we find that the end result obtained via combining FDD
+with hand made tests is remarkable, as it is as natural to work with
+as TDD.  FDD with generated tests is sometimes annoying because of the
+time needed to run the full test suite, thus often only part of the
+test suite is executed so that the development loops do not take too
+long.  TDF is great for complex preconditions, but really difficult
+with postconditions, especially if the specification is written for
+existing code.  When writing after code, the corresponding test is not
+expected to fail, hence the feedback is limited and the contract is
+often not as strong as needed.  In this specific case, TDF
+effectiveness is limited.
 
@@
 % the problem
-In this paper we focus on a specific set of development problems that neither agile methodologies nor FMs are completely confortable to cope with.
-The problems we are interested in are characterized by unstable requirements and artifacts which must be formally specified and verified, and are constrained by deadlines that cannot be postponed.
-We propose a blend of agile and formal engineering practices, enclosed by an iterative and incremental development process: the ``Mole'' development process.
+In this paper we focus on a specific set of development challenges
+with which neither AMs nor FMs are completely confortable.  The
+challenges are characterized by unstable requirements combined with
+artifacts that must be formally specified and verified all while the
+development team is constrained by deadlines that cannot be postponed.
+We propose a blend of agile and formal engineering practices, enclosed
+by an iterative and incremental development process: the \mole
+development process.
 
 % Mole
-The verification practices proposed recreate the fast feedback development environment we find in Test Driven Development, used with formal and informal artifacts.
-A total of four practices are presented, two of them are conservative, following closely the development practices of FMs and AMs, while the other pair are innovative, blending together elements from both worlds.
+The verification practices described in this paper recreate the fast
+feedback development environment we find in Test Driven Development in
+the presence of formal and informal artifacts.  A total of four
+practices are presented---two of them are conservative, following
+closely the development practices of formal methodologies and agile
+methodologies, while the other pair are innovative, blending together
+elements from both worlds.
 
 % Mole at work
-The practices have been applied succesfully to develop a driver for a custom embedded sensor board, equipped with board simulators and protocol verifiers. 
-The protocol and the board specifications were informal and unstable. 
-Yet a working software product was needed as soon as possible, with incrementally added functionality, and keeping it linked with the last specifications of the protocol and the board.
-``Mole'' verification practices enabled us to keep in sync formal specifications (JML annotations), informal specifications (test suites) and source code, through very fast and short development cycles.
-We developed handmade test suites, generated test suites and JML annotations, wich succesfully supported the verification of the driver and the simulators, and the board as well, when it was delivered.
+The practices have been applied succesfully to develop a driver for a
+custom embedded sensor board equipped with board simulators and
+protocol verifiers.  And, while the protocol and the board
+specifications were informal and unstable, a working software product
+was needed as soon as possible, with incrementally added
+functionality, and ensuring that the software and hardware products
+were linked via the last specifications of the protocol and the board.
+\mole verification practices enabled us to keep in sync formal
+specifications (JML annotations), informal specifications (test
+suites), and source code through very rapid, short development cycles.
+We developed handmade test suites and generated test suites and JML
+annotations, all of which successfully supported the verification of
+the software driver and the simulators, as well as the hardware board
+when it was delivered.  In fact, the resulting system had no flaws, as
+it worked correctly the very first time it was plugged in to the
+prototype hardware.
 
 % extending Mole
-The practices are very general, and are applicable to different development languages and FMs.
-One important constraint is on the FM, that must be able to cope with partial specifications; another constraint is on the tools available, that must be able to verify the system properties at runtime and partially automate the verification process of the formalized properties.
+The practices we describe are general and are applicable to different
+development languages and formal methodologies.  One important
+constraint is on the formal methodology is that it must cope with
+partial specifications.  Another constraint is on the tools available,
+which must verify system properties at runtime and partially automate
+the verification process of the formalized properties.
+
+TODO POP!
 
@@
 \end{document}
+
+%=====================================================================
+% hi Joe
+
+% what about having the chat next week?
+% on thursday, maybe?
+
+% here are my comments on the reviewers:
+
+% ---------------------------- REVIEW 1 --------------------------
+% OVERALL RATING: 3 (strong accept)
+
+% > This paper successfully bridges the gab between both worlds. I would be i=
+% nteresting to read at bit more about the gained experience with the techniq=
+% ues based on the presented case study (what was easy what difficult).
+
+% agree. to be included
+
+% > Furthermore I recommend that the authors work out a bit more detailed the=
+%  prerequisites to apply a mixture of formal and agile techniques (e.g. by m=
+% eans of a list of "thesis"). This would stimulate the discussion!
+
+% agree. to be included. it's already there, but not clear and polished enoug=
+% h
+
+% ---------------------------- REVIEW 2 --------------------------
+% OVERALL RATING: 2 (accept)
+
+% > The authors use the (somewhat formal) JML technology and introduce their =
+% TDF-methodology as a variant of more classic development cycles.
+
+% why "somewhat"?
+
+% > Content-wise it must be said that, given the amount of material in the te=
+% xt, the conclusions drawn and lessons learnt are a bit 'thin'. E.g. Does TD=
+% F really work? what is your experience?
+
+% partial agree. to be expanded, in the line of the previous comments
+
+% > It seems that the format of the printed area (text height and width) is n=
+% ot in accordance with the other conference contributions. When reformatted,=
+%  the already long paper will probably get longer. The (too large) wordiness=
+%  of the paper could perhaps be reduced somewhat to compensate.
+
+% uh
+% I used thir template... I'm "almost" sure
+% I'll double check this
+% maybe some following format options
+
+% > page 1, abstract: 'not always see eye to eye' seems an inappropriate way =
+% of stating 'do not meet face to face'.
+% > page 1, abstract end: highly-spices -> highly-spiced
+% > page 2, line 2: citation problem
+% > page 2, line 12: 'ther'?
+% > page 3, line 8 and 10: appliable in -> applicable to?
+% > page 3, line 8: yeilding -> yielding
+% > page 3, middle: .... the handmade test suites are automatically generated=
+%  .... ?
+% > page 3, the same section: the last two sentences seem crippled; split the=
+% m up!
+% > page 6, line -5: Identify -> Identifying
+% > page 9, line 1: confidant -> confident
+% > page 15, line 6: inoovative -> innovative
+% > page 15, line 12: in sink -> in sync
+% > page 17, last line: litt ref in disorder
+
+% ok
+
+% > Summing up the editorial part: Please use a spelling checker and perhaps =
+% some review by a native speaker
+% could help sharpening (and shortening) some sentences.
+
+% well... I've used a spell checker!
+% but we have to reread it once the modifications are in place
+
+% ---------------------------- REVIEW 3 --------------------------
+% OVERALL RATING: 2 (accept)
+
+% > they do not mention Liu's work on the SOFL foraml engineering method for =
+% software development that have been xtensively studies over the last 15 yea=
+% rs with the aim to solve the similar problem. In particular, the recent pap=
+% er on the application of SOFL to achieve agile process has provided a very =
+% different approach from the mentioned approaches
+% > Shaoying Liu, ``An Approach to Applying SOFL for Agile Process and Its Ap=
+% plication in Developing a Test Support Tool'', Innovations in Systems and S=
+% oftware Engineering, Springer London, 22 December, DOI 10.1007/s11334-0114-=
+% 3.
+
+% ok
+% to be read and analyzed
+
+% ---------------------------- REVIEW 4 --------------------------
+% OVERALL RATING: 1 (weak accept)
+
+% > I did find it disturbing that the paper never clearly distinguishes betwe=
+% en structural and functional testing . This is an
+% important distinction as it potentially evidences different types of proble=
+% ms.
+
+% I'll rewrite some parts to make the distinction clear
+
+% > Moreover, there is a distinct difference between automatically generated =
+% tests and manually produced tests as to how well they apply to one or the o=
+% ther.
+
+% the distinction is clear in the paper... I don't understand this
+
+% > The overall description of how AM and FM are merged is not that persuasiv=
+% e, as a result. In particular, Figure 2 does not
+% convey much nor profound value in its current incarnation and description.
+
+% mmhh
+% not much that can be done here... I think
+% how can we be more "persuasive"?
+
+% > It would help if the authors would more specifically address what type of=
+%  tests the JML specification helps generate. In the
+% case study, it is a structural coverage type test.
+
+% mmhh
+% I've explained this
+% it would take me pages to go into specific details... and I've cited
+% works where this is explained
+
+% > All that said, it is a bit unclear as to where the presented work is actu=
+% ally pushing the boundaries of the current state of
+% the field, other than in actually presenting the approach as an FM+AM hybri=
+% d.
+
+% we are actually talking about FM+AM hybrids... that IS the filed,
+% so... what's this question about?
+
+% > As a suggestion, then, it could be good if the verbosity and obtusity of =
+% the article is reduced
+
+% ok...
+
+% > and a more formal and incisive treatise is presented.
+
+% no idea how to do this
+% nor I'm sure I really want to do this
+
+% --=20
+% Vieri del Bianco

### 2010-08-24 papers/AgileFormality/paper.tex
reviewed bibliographies still need to be reviewed reread all

@@
   \usepackage{epstopdf}
-  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true,
-  colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue,
-  pdfauthor="Vieri del Bianco and Joseph R. Kiniry and Dragan
-  Stosic"]{hyperref} 
+  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Vieri del Bianco and Joseph R.
+Kiniry and Dragan Stosic"]{hyperref}
 \else
@@
 \newcommand{\STUs}{\ST \emph{Units}\xspace}
-\newcommand{\mole}{``\emph{Mole}''\xspace}
 
@@
 \newcommand{\sensorfarm}{The \STSBs\xspace}
+\newcommand{\mole}{``\emph{Mole}''\xspace}
 
@@
 
-\title{Agile Formality: A \mole of\\Software Engineering Practices}
+\title{Agile Formality:\\ A \mole of Software Engineering Practices}
 
@@
 
-  Members of the agile programming and formal methods communities do
-  not always see eye-to-eye.  These two communities often do not talk
-  to or learn from each other.  Only recently, as highlighted by the
-  September 2009 issue of IEEE Software, the IFIP workshop on
-  balancing agility and formalism in software engineering, and the
-  first edition of the international workshop for formal methods and
-  agile methods, ideas from the two communities begun synthesize.
-  While the problem-solving approaches and psychological attitudes of
-  members of the two communities differ widely, we exploit this clash
-  of viewpoints, creating a new development processes that actually
-  blends, rather than mashes together, best practices from the two
-  worlds.  This paper summarizes our process and a supporting complex
-  case study, showing that it is not only possible, but tasty, to
-  combine the ``chili pepper'' of formal methods and the ``chocolate''
-  of agile programming, thus producing a tasty \mole (as in the
-  highly-spiced Mexican sauce) of software engineering practices.
+Members of the agile programming and formal methods communities do not always see eye-to-eye.  
+These two communities often do not talk to or learn from each other.  
+Only recently, as highlighted by the September 2009 issue of IEEE Software, the IFIP workshop on balancing agility and formalism in software engineering, and the first edition of the international workshop for formal methods and agile methods, ideas from the two communities begun synthesize.
+While the problem-solving approaches and psychological attitudes of members of the two communities differ widely, we exploit this clash of viewpoints, creating a new development processes that actually blends, rather than mashes together, best practices from the two worlds.
+This paper summarizes our process and a supporting complex case study, showing that it is not only possible, but tasty, to combine the ``chili pepper'' of formal methods and the ``chocolate'' of agile programming, thus producing a tasty \mole (as in the highly-spiced Mexican sauce) of software engineering practices.
 
@@
 % agile and formal difficult to blend: mind attitudes
-Agile and formal development methodologies usually do not blend
-together well.  This is because of several reasons, the most important
-of which is often characterized as a radical difference in
-psychological attitudes about software development.
+Agile and formal development methodologies usually do not blend together well.  
+This is because of several reasons, the most important of which is often characterized as a radical difference in psychological attitudes about software development.
 
 % FM
-Formal methodologists often favour an in-depth, think-first approach,
-where the problem is understood, formalized, and solved; usually, but
-not always, adopting a waterfall-style of development.  Once a
-formalization is developed, development and verification proceed in an
-automated and interactive fashion.  Consequently, projects that use
-formal methodologies (FMs) typically focus on critical systems with
-fixed requirements and somewhat flexible deadlines (``We'll ship it
-when it is right.'').
+Formal methodologists often favour an in-depth, think-first approach, where the problem is understood, formalized, and solved; usually, but not always, adopting a waterfall-style of development.  
+Once a formalization is developed, development and verification proceed in an automated and interactive fashion.  
+Consequently, projects that use formal methodologies (FMs) typically focus on critical systems with fixed requirements and somewhat flexible deadlines (``We will ship it when it is right.'').
 
 % AM
-Agile methodologists favour an highly incremental and iterative
-approach, specifically tailored to cope with changing requirements and
-precise deadlines.  In projects that use agile methodologies (AMs), it
-is often the case that the problem is only partially understood, and
-there is a focus only on the aspects that must be implemented in the
-current development iteration.  The solution initially developed is
-usually not optimal, but as it is refined through continuous code
-refactorings, its quality and validity improve.  Test suites are used
-for several purposes, the most important of which are system
-verification, system documentation, and requirement specification, and
-they are strictly handmade.  AMs are typically used in development
-settings where changing requirements and rapid delivery of the product
-are paramount.
+Agile methodologists favour an highly incremental and iterative approach, specifically tailored to cope with changing requirements and precise deadlines.  
+In projects that use agile methodologies (AMs), it is often the case that the problem is only partially understood, and there is a focus only on the aspects that must be implemented in the current development iteration.
+The solution initially developed is usually not optimal, but as it is refined through continuous code refactorings, its quality and validity improve.  
+Test suites are used for several purposes, the most important are system verification, system documentation, and requirement specification, and they are strictly handmade.  
+AMs are typically used in development settings where changing requirements and rapid delivery of the product are paramount.
 
 % problems would benefit form agile and formal mix, characterize the problem
-The two worlds seem irreconcilable.  Nevertheless, problems exist that
-would greatly benefit from both of them~\cite{Black2009}.  The problem
-class in which we have particular interest have \emph{unstable
-  requirements}, are \emph{constrained by deadlines that cannot be
-  postponed}, but also have \emph{subsystems that must be formally
-  specified and verified}.
+The two worlds seem irreconcilable.  
+Nevertheless, problems exist that would greatly benefit from both of them~\cite{Black2009}.  
+The problem class in which we have particular interest have \emph{unstable requirements}, are \emph{constrained by deadlines that cannot be postponed}, but also have \emph{subsystems that must be formally specified and verified}.
 
-\subsection{Terminology}
-\label{sec:terminology}
+\paragraph*{Terminology}
+%\label{sec:terminology}
 
 % IEEE terminology
-In this paper the IEEE 610~\cite{Jay1990} standard terminology for
-validation and verification is used.  Paraphrasing the standard,
-\emph{verification} is the process of evaluating software to determine
-whether the products of a given development phase satisfy the
-conditions imposed, according to the specifications (``you build the
-system in the right way''); \emph{validation} is the process of
-evaluating software to determine whether it satisfies specified
-requirements ensuring it meets the user's needs (``you build the right
-system'').
+In this paper the IEEE 610~\cite{Jay1990} standard terminology for validation and verification is used.  
+Paraphrasing the standard, \emph{verification} is the process of evaluating software to determine whether the products of a given development phase satisfy the conditions imposed, according to the specifications (``you build the system in the right way''); \emph{validation} is the process of evaluating software to determine whether it satisfies specified requirements ensuring it meets the user's needs (``you build the right system'').
 
 % FM terminology
-Within the FMs community, verification and validation have different
-meanings than in the IEEE 610 standard.  Loosely \emph{verification}
-(in the paper \emph{formal verification}) means the use of formal
-methods to formally, statically verify that a system conforms to its
-specification; \emph{validation} (in the paper \emph{formal
-  validation}) means the use of traditional software engineering
-practices to check, statically or dynamically, that a system conforms
-to its specification (whether in the form of tests, requirements,
-etc.) by hand or via execution under some execution scenarios.  In the
-present paper both perspectives are respected and acknowledged.
+Within the FMs community, verification and validation have different meanings than in the IEEE 610 standard.
+Loosely \emph{verification} (in the paper \emph{formal verification}) means the use of formal methods to formally, statically verify that a system conforms to its specification; \emph{validation} (in the paper \emph{formal validation}) means the use of traditional software engineering practices to check, statically or dynamically, that a system conforms to its specification (whether in the form of tests, requirements, etc.) by hand or via execution under some execution scenarios.
+In the present paper both perspectives are respected and acknowledged.
 
 % tests terminology
-Tests are usually classified by dichotomies: unit vs. functional,
-black-box vs. white-box, testing vs. design, tester vs. coder.  Yet
-testing does not always fit into these dichotomies, especially in the
-case of agile testing approaches~\cite{Beck2007}.  The black-box
-vs. white-box dichotomy is particularly problematic, since box
-boundaries change one's perspective.  In this paper, functional tests
-are always black-box tests, coverage tests are always white-box tests,
-generated tests are always black-box tests (the box is the module
-specified); other strict rules do not hold.
+Tests are usually classified by dichotomies: unit vs. functional, black-box vs. white-box, testing vs. design, tester vs. coder. 
+Yet testing does not always fit into these dichotomies, especially in the case of agile testing approaches~\cite{Beck2007}.
+The black-box vs. white-box dichotomy is particularly problematic, since box boundaries change one's perspective.
+In this paper, functional tests are always black-box tests, coverage tests are always white-box tests, generated tests are always black-box tests (the box is the module specified); other strict rules do not hold.
 
 % development terminology
-Additionally, ``...development of a system,'' more precisely means
-``...analysis, design, development, verification, formal validation of
-a system.''
+Additionally, ``...development of a system,'' more precisely means ``...analysis, design, development, verification, formal validation of a system.''
 
@@
 
-Proposing new practices with no supporting evidence is an
-uninteresting proposition.  A complex case study is mandatory to
-test-bench the proposed practices.  Our case study focuses on the
-development of a device driver to control and communicate with an
-embedded custom circuit board equipped with more than a dozen sensors.
-The communication protocol with the board is asynchronous and
-packet-based and the board is novel and custom-designed.  The protocol
-is focused on controlling the board and reporting sensor measurements
-and board status.
+Proposing new practices with no supporting evidence is an uninteresting proposition.  
+A complex case study is mandatory to test-bench the proposed practices.  
+Our case study focuses on the development of a device driver to control and communicate with an embedded custom circuit board equipped with more than a dozen sensors.
+The communication protocol with the board is asynchronous and packet-based and the board is novel and custom-designed.  
+The protocol is focused on controlling the board and reporting sensor measurements and board status.
 
-This software driver has been developed using our process while
-satisfying seemingly contradictory requirements: (1)~the protocol
-implementation must be thoroughly verified; (2)~parts of the protocol
-must be formally specified; (3)~a board simulator must be designed,
-implemented, and verified to match the specifications of, and actual
-behaviour of, the physical circuit board; (4)~if board components
-change during development, the associated formal specifications and
-the simulator whose behavior reflects such changes must be updated
-accordingly; (5)~there are strict software delivery deadlines.
-Consequently, these process requirements constitute an environment
-that benefits from both agile and formal approaches to software
-development.
+This software driver has been developed using our process while satisfying seemingly contradictory requirements: (1)~the protocol implementation must be thoroughly verified; (2)~parts of the protocol must be formally specified; (3)~a board simulator must be designed, implemented, and verified to match the specifications of, and actual behaviour of, the physical circuit board; (4)~if board components change during development, the associated formal specifications and the simulator whose behavior reflects such changes must be updated accordingly; (5)~there are strict software delivery deadlines.
+Consequently, these process requirements constitute an environment that benefits from both agile and formal approaches to software development.
 
-A novel development process has been drafted to cope with the
-contrasting requirements of the case study, blending agile and formal
-practices.  In the following, the development process and the practices
-are summarized, and their successful application to the case study is
-reported, yielding several promising results.
+A novel development process has been drafted to cope with the contrasting requirements of the case study, blending agile and formal practices.  
+In the following, the development process and the practices are summarized, and their successful application to the case study is reported, yielding several promising results.
 
@@
 % what is needed, high-level
-To enable the integration of the two worlds, the highly iterative and
-incremental approach found in most of the AMs must be maintained, and
-formal validation methodologies, traditionally used in a waterfall
-development process (formally specify the system, implement the
-system, validate the implementation against the specification) must
-be adapted to a highly iterative one.
+To enable the integration of the two worlds, the highly iterative and incremental approach found in most of the AMs must be maintained, and formal validation methodologies, traditionally used in a waterfall development process (formally specify the system, implement the system, validate the implementation against the specification) must be adapted to a highly iterative one.
 
 % what must be considered: TDD
-The most common software verification practices in AMs must be
-considered when trying to integrate FMs and AMs since they are a
-fundamental part of the development process.  The most significant
-verification related practice is Test Driven Development
-(TDD)~\cite{Beck2003}.  TDD is a software development technique,
-originally defined in the Extreme Programming (XP)~\cite{Beck2004}
-methodology and is characterized by a process that focuses on writing
-unit tests for a module before writing the code of said module.
+The most common software verification practices in AMs must be considered when trying to integrate FMs and AMs since they are a fundamental part of the development process. 
+The most significant verification related practice is Test Driven Development (TDD)~\cite{Beck2003}. 
+TDD is a software development technique, originally defined in the Extreme Programming (XP)~\cite{Beck2004} methodology and is characterized by a process that focus on writing unit tests of a module before writing the code of said module.
 
 % test driven kinds
-The test driven approach relies on writing tests before
-implementations.  It is applicable to a scall and large-scale cycles.
-In the former case, very small scale TDD yields activity cycles as
-short as a few minutes.  In the latter, requirements are immediately
-translated into functional tests resulting in activity cycles as long
-as a full delivery iteration, which is usually no longer than a couple
-of months.  The test driven approach is a cornerstone of XP, but it is
-so popular that has been adopted in many other AMs as well.  It is
-also considered a good software development technique when used on its
-own, as it pre-dates XP and AMs considerably, regardless of the
-enclosing development process.
+The test driven approach relies on writing tests before implementations. 
+It is applicable to a very small scale and large scale cycles
+In the former case, TDD yields to activity cycles as short as a few minutes.
+In the latter, requirements are immediately translated into functional tests resultin in activity cycles as long as a full delivery iteration, which is usually no longer than a couple of months.
+The test driven approach is a cornerstone of XP, but it is so popular that has been adopted in many other AMs as well.
+It is also considered a good software development technique when used on its own, regardless of the enclosing development process.
 
 % previous related work
-Some tentative attempts to reconcile FMs and AMs have been developed
-by others.  The reoccuring theme of these attempts is that idea that
-handmade test suites are no longer used, but instead are replaced by
-different kinds of automated formal validation based on a system-level
-formal specification.  Automated validation can be as simple as
-enabling runtime checking of assertions or automatically generating a
-test suite based on the system properties, or as complex as proving
-the whole system's behaviour through formal verification.
+Some tentative attempts to reconcile FMs and AMs have been developed by others.
+The reoccuring theme of these attempts is that idea that handmade test suites are no longer used, but instead are replaced by different kinds of automated formal validation based on a system level formal specification.
+Automated validation can be as simple as enabling runtime checking of assertions or automatically generating a test suite based on the system properties, or as complex as proving the whole system's behaviour through formal verification.
 
 % a quick analysis of the related works
-Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an
-iterative and incremental process to generate X-Machines models, an
-extension of Finite State Machines.  The development process impose a
-complete specification of the system, and its verification is done
-exclusively using the tools provided by the X-Machines FM.  The
-development process proposed by Herranz and
-Moreno-Navarro~\cite{Herranz2003b} is quite similar.  They propose an
-iterative process to model a system using SLAM-SL, an object-oriented
-specification language supported by tools.  Some XP practices are
-fully adopted or considered compatible, like pair-programming,
-iterative and incremental development, and system refactoring; while
-tests are completely replaced by the verification tools provided by
-the SLAM suite.  A different FM, but similar approach, is found also
-in the work of Suhaib~\cite{Suhaib2005} where the XFM methodology is
-used.
+Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an iterative and incremental process to generate X-Machines models, an extension of Finite State Machines. 
+The development process impose a complete specification of the system, and its verification is done exclusively using the tools provided by the X-Machines FM. 
+The development process proposed by Herranz and Moreno-Navarro~\cite{Herranz2003b} is quite similar.
+They propose an iterative process to model a system using SLAM-SL, an object-oriented specification language supported by tools. 
+Some XP practices are fully adopted or considered compatible, like pair-programming, iterative and incremental development and system refactoring; while tests are completely replaced by the verification tools provided by the SLAM suite.
+A different FM, but similar approach, is found also in the work of Suhaib~\cite{Suhaib2005} where XFM methodology is introduced.
 
 % previous work: a common approach
-As already mentioned, in all of these attempts, test suites are
-automatically generated using the system formal specification, or are
-completely replaced by formal verification, typically through the use
-of theorem provers and/or static checkers.  The major precondition on
-the use of these approaches is, at the minimum, a specification of all
-the relevant aspects of the system under development.  Such a
-specification enables simple verification techniques such as runtime
-checking (via assertions and design by contract~\cite{Meyer1997} using
-preconditions, postconditions and invariants) and the automatic
-generation of test suites~\cite{Cheon2002,Cheon2004,Cheon2005}.  Via a
-complete and sound system specification, supported by an appropriate
-formal language (to specify the system model and its properties)
-coupled with a (possibly restricted) programming language, one can
-statically prove properties about the system using a variety of tools
-and
-techniques~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
+As already mentioned, in all these attempts, the test suites are automatically generated using system formal specification, or are completely replaced by formal verification, typically through the use of theorem provers and/or static checkers.
+The major precondition on the use of these approaches is, at the minimum, a specification of all the relevant aspects of the system under development. 
+Such a specification enables simple verification techniques such as runtime checking (via assertions and design by contract~\cite{Meyer1997} using preconditions, postconditions and invariants) and the automatic generation of test suites~\cite{Cheon2002,Cheon2004,Cheon2005}.
+Via a complete and sound system specification, supported by an appropriate formal language (to specify the system model and its properties) coupled with a (possibly restricted) programming language, one can statically prove properties about the system using a variety of tools and techniques~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.
 
 % previous work: why is not satisfying
-The underlying idea of this approach is simple: what was once the
-purpose of handwritten tests (verification, documentation, and
-requirement specification) are now the responsibility of the formal
-specifications, and a complete formal specification of the system
-under development is needed to apply these methodologies.  The problem
-is that this requirement is very difficult to fulfill since the effort
-required to write a complete formal specification of the system in
-most real-world complex cases is usually greater than writing a suite
-of tests~\cite{Glass2002}.
-
-If a complete formal specification is not available, all of the
-previous methodologies share a common problem: the parts informally
-specified are not verified at all.  These parts of the system are
-consequently verified with traditional methods, but the connections
-between development artifacts and related activities (code, tests,
-design, refactoring) are neither detailed nor enforced.  This is
-problematic, as the loops and relations between the different
-development artifacts are the inner engine of many AMs, as they impose
-an iterative and incremental pace to the development process.
+The underlying idea of this approach is simple: what was once the purpose of handwritten tests (verification, documentation and requirement specification) are now the respnsability of the formal specifications, and a complete formal specification of the system under development is needed to apply these methodologies.
+The problem is that this requirement is very difficult to fulfill since the effort required to write a complete formal specification of the system in most real world complex cases is usually greater than writing a suite of tests~\cite{Glass2002}.
+If a complete formal specification is not available, all of the previous methodologies share a common problem: the parts informally specified are not verified at all.
+These parts os the system are consequently verified with traditional methods, but the connections between development artifacts and related activities (code, tests, design, refactoring) are neither detailed nor enforced.
+This is problematic as the loops and relations between the different development artifacts are the inner engine of many AMs, as they impose an iterative and incremental pace to the development process.
 
 % outsider
-Liu takes a different approach wherein the SOFL
-methodology~\cite{Liu2004} is merged into agile
-processes~\cite{Liu2010}.  The SOFL methodology is based on a three
-step specification approach: (1) informal specification, (2)
-semi-formal specification, (3) formal specification.  The agile
-adaptation consists mainly in introducing shorter loops between
-activities and reducing the size of the set of formal specifications.
-The formal specifications are used only to help understand ambiguous
-statements in the semi-formal specification and are not maintained
-while the system evolves.  Testing and inspections are based on the
-semi-formal specification.  The formal specification are partial and
-are only used for documentation purposes and then discarded---they are
-not used at all in the verification of the system.
+Liu takes a different approach wherein the SOFL methodology~\cite{Liu2004} is merged into agile processes~\cite{Liu2010}. 
+The SOFL methodology is based on a three step specification approach: (1) informal specification, (2) semi-formal specification, (3) formal specification. 
+The agile adaptation consists mainly in introducing shorter loops between activities and reducing the size of the set of formal specifications.
+The formal specifications are used only to help understand ambiguous statements in the semi-formal specification and are not maintained while the system evolves.
+Testing and inspections are based on the semi-formal specification.
+The formal specification are partial and are only used for documentation purposes and than discarded---they are not used at all in the verification of the system.
 
 % what we want to achieve
-In contrast, our objective is to formally specify only those parts of
-the system that must be verified (in order to cope with constrained
-resources and unstable requirements) and to develop and refine a
-highly iterative and incremental development process that blends
-formal and agile practices.
+I contrast our objective is to formally specify only parts of the system (in order to cope with constrained resources and unstable requirements) and to develop and refine a highly iterative and incremental development process that blends formal and agile practices.
 
 % how we want to achive it
-The main open problem unaddressed in the literature is how to connect
-development artifacts of the two worlds of FMs and AMs together.  To
-answer this challenge we define activity cycles, similar to what is
-found in AMs (especially in XP), to solve this problem in an highly
-iterative development process.  We show how tests drive the formal
-specification, how the formal specifications drive test and code
-development, how handmade and automated tests coexist and support each
-other, how the unspecified parts of the system are incrementally
-specified.  That is, we show how to blend, and not just tack together,
-formal validation and agile verification practices: a \mole of
-verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly
-  spiced Mexican sauce made chiefly from chili peppers (agile) and
-  chocolate (formal), served with meat (working system).}.
+The main open problem unaddressed in the literature is how to connect development artifacts of the two worlds of AMs and FMs together.
+To answer this challenge we define activity cycles, similar to what is found in AMs (especially in XP), to solve this problem in an highly iterative development process.
+We show how tests drive the formal specification, how the formal specifications drive tests and code development, how handmade and automated tests coexist and support each other, how the unspecified parts of the system are incrementally specified.
+That is, we show how to blend, and not just tack together, formal validation and agile verification practices: a \mole of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers (agile) and chocolate (formal), served with meat (working system).}.
 
@@
 
-\subsection{Related Methodologies}
-\label{sec:relat-meth}
+\paragraph*{Related Methodologies}
+%\label{sec:relat-meth}
 
@@
 
-% ======================================================================
-\section{Blending Formal And Agile Development}
+\section{Blending Formal And Agile Development: the \mole Practices}
 \label{sec:blending_formal_and_agile_development}
@@
 
-\subsection{\mole Practices}
+%\subsection{\mole practices}
 
@@
 \begin{lstlisting}[
-  caption={A specification with JML annotation: a complex object invariant which constraints 
-    the number of samples for each medium data rate stream channel.},
+  caption={A complex specification with JML annotation: invariant constraining the number of samples for medium data rate streams.},
   label=lst:jml_example_invariant,
@@
   \centering
-  \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__Main}
+  \includegraphics[scale=0.6]{UML_model/Class_Diagram__Structure__Main}
   \caption{Main packages class diagram.}
@@
   \centering
-  \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__Test}
+  \includegraphics[scale=0.35]{UML_model/Class_Diagram__Structure__Test}
   \caption{Test packages class diagram.}
@@
   \centering
-  \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__StreamDriverT}
+  \includegraphics[scale=0.35]{UML_model/Class_Diagram__Structure__StreamDriverT}
   \caption{StreamDriver test packages class diagram: use of the stream
@@
 % preconditions on projects
-We believe that \mole practices are relatively easy to introduce and
-use.  Many kinds of projects could benefit from these practices, not
-only those that have unstable requirements but also demands for a
-certain level of formality.  Even projects with no requirements for
-formal specification at all benefit from these practices, since some
-simple or obvious system properties are easy to specify and, though
-their introduction, there is a corresponding reduction in the test
-suite development effort.
+We believe that \mole practices are relatively easy to introduce and use.
+Many kinds of projects could benefit from these practices, not only those that have unstable requirements but also demands for a certain level of formality.
+Even projects with no requirements for formal specification at all benefit from these practices, since some system properties are easy to specify and, through their introduction, there is a corresponding reduction in the test suite development effort.
 
 % precondition on formal methodology and associated formal suite
-The cornerstone of the practices is the formal methodology and its
-supporting tools, and the precondition on the tools' utility is not
-always easily fulfilled.  Methodologies that require a complete
-formalization are nearly impossibile to apply.  In addition, it is not
-clear if a tool suite that does not provide a runtime verification,
-either via runtime checking or simulation, can be broadly used
-succesfully, nor it is clear whether formal methods that are not based
-on the contract paradigm are useful in this context.  These are open
-questions that will need to be addressed in the future to determine
-the broad, concrete applicability of the \mole practices.
+The cornerstone of the practices is the formal methodology and its supporting tools, and the precondition on the tools' utility is not always easily fulfilled.
+Methodologies that require a complete formalization are nearly impossibile to apply.
+In addition, it is not clear if a tool suite that does not provide a runtime verification, either via runtime checking or simulation, can be broadly used succesfully, nor it is clear whether formal methods that are not based on the contract paradigm are useful in this context.
+These are open questions that will need to be addressed in the future to determine the broad applicability of the \mole practices.
 
 % when to use what
-There are no strict rules to determine which of the \mole practices to
-use when facing a specific problem.  We have found that experience
-trumps cookbooks, so we can only give some advice.  We suggesting
-using TDD there is no need for any specifications at all, FDD with
-generated tests if the specification are not too difficult and system
-scenarios is simple enough to be covered by generated tests, FDD with
-hand made tests if the scenarios are complex, and TDF when the
-specification is complex.  We feel that the experience we have
-gathered is not yet enough to distill richer idioms or patterns---such
-is future work.
+There are no strict rules to determine which of the \mole practices to use when facing a specific problem.
+Since experience trumps cookbooks, we can only give some advice.
+We suggest using TDD if there is no need for any specifications at all, FDD with generated tests if the specification are not too difficult and execution scenarios are simple enough to be covered by generated tests, FDD with hand made tests if the scenarios are complex, and TDF when the specification is complex.
+We feel that the experience we have gathered is not yet enogh to distill richer idioms or patterns---such is future work.
 
 % what is working and what is not
-Upon reflect, we find that the end result obtained via combining FDD
-with hand made tests is remarkable, as it is as natural to work with
-as TDD.  FDD with generated tests is sometimes annoying because of the
-time needed to run the full test suite, thus often only part of the
-test suite is executed so that the development loops do not take too
-long.  TDF is great for complex preconditions, but really difficult
-with postconditions, especially if the specification is written for
-existing code.  When writing after code, the corresponding test is not
-expected to fail, hence the feedback is limited and the contract is
-often not as strong as needed.  In this specific case, TDF
-effectiveness is limited.
+Upon reflect, we find that the end result obtained via combining FDD with hand made tests is remarkable, as it is as natural to work with as TDD. 
+FDD with generated tests is sometimes annoying because of the time needed to run the full test suite, thus often only part of the test suite is executed so that the development loops do not take too long.
+TDF is great for complex preconditions, but really difficult with postconditions, especially if the specification is written for existing code. 
+When writing after code, the corresponding test is not expected to fail, hence the feedback is limited and the contract is often not as strong as needed. 
+In this specific case, TDF effectiveness is limited.
 
@@
 % the problem
-In this paper we focus on a specific set of development challenges
-with which neither AMs nor FMs are completely confortable.  The
-challenges are characterized by unstable requirements combined with
-artifacts that must be formally specified and verified all while the
-development team is constrained by deadlines that cannot be postponed.
-We propose a blend of agile and formal engineering practices, enclosed
-by an iterative and incremental development process: the \mole
-development process.
+In this paper we focus on a specific set of development challenges with which neither AMs nor FMs are completely confortable.
+The challenges are characterized by unstable requirements combined with artifacts that must be formally specified and verified all while the developmemnt team is constrained by deadlines that cannot be postponed.
+We propose a blend of agile and formal engineering practices, enclosed by an iterative and incremental development process: the \mole development process.
 
 % Mole
-The verification practices described in this paper recreate the fast
-feedback development environment we find in Test Driven Development in
-the presence of formal and informal artifacts.  A total of four
-practices are presented---two of them are conservative, following
-closely the development practices of formal methodologies and agile
-methodologies, while the other pair are innovative, blending together
-elements from both worlds.
+The verification practices described in this paper recreate the fast feedback development environment we find in Test Driven Development in the presence of both formal and informal artifacts.
+A total of four practices are presented---two of them are conservative, following closely the development practices of FMs and AMs, while the other pair are innovative, blending together elements from both worlds.
 
 % Mole at work
-The practices have been applied succesfully to develop a driver for a
-custom embedded sensor board equipped with board simulators and
-protocol verifiers.  And, while the protocol and the board
-specifications were informal and unstable, a working software product
-was needed as soon as possible, with incrementally added
-functionality, and ensuring that the software and hardware products
-were linked via the last specifications of the protocol and the board.
-\mole verification practices enabled us to keep in sync formal
-specifications (JML annotations), informal specifications (test
-suites), and source code through very rapid, short development cycles.
-We developed handmade test suites and generated test suites and JML
-annotations, all of which successfully supported the verification of
-the software driver and the simulators, as well as the hardware board
-when it was delivered.  In fact, the resulting system had no flaws, as
-it worked correctly the very first time it was plugged in to the
-prototype hardware.
+The practices have been applied succesfully to develop a driver for a custom embedded sensor board equipped with board simulators and protocol verifiers. 
+And while the protocol and the board specifications were informal and unstable, a working software product was needed as soon as possible, with incrementally added functionality, and ensuring that the software and hardware products were linked via the last specifications of the protocol and the board.
+\mole verification practices enabled us to keep in sync formal specifications (JML annotations), informal specifications (test suites) and source code through very fast and short development cycles.
+We developed handmade test suites and JML annotations, and generated test suites, all of wich succesfully supported the verification of the software driver and the simulators as well as the hardware board when it was delivered.
+In fact, the resulting system had no flaws, as it worked correctly the very first time it was plugged in to the prototype hardware.
 
 % extending Mole
-The practices we describe are general and are applicable to different
-development languages and formal methodologies.  One important
-constraint is on the formal methodology is that it must cope with
-partial specifications.  Another constraint is on the tools available,
-which must verify system properties at runtime and partially automate
-the verification process of the formalized properties.
-
-TODO POP!
+The practices we describe are general, and are applicable to different development languages and FMs.
+One important constraint is on the FM, since it must cope with partial specifications.
+Another constraint is on the tools available, which must verify the system properties at runtime and partially automate the verification process of the formalized properties.
 
@@
 \end{document}
-
-%=====================================================================
-% hi Joe
-
-% what about having the chat next week?
-% on thursday, maybe?
-
-% here are my comments on the reviewers:
-
-% ---------------------------- REVIEW 1 --------------------------
-% OVERALL RATING: 3 (strong accept)
-
-% > This paper successfully bridges the gab between both worlds. I would be i=
-% nteresting to read at bit more about the gained experience with the techniq=
-% ues based on the presented case study (what was easy what difficult).
-
-% agree. to be included
-
-% > Furthermore I recommend that the authors work out a bit more detailed the=
-%  prerequisites to apply a mixture of formal and agile techniques (e.g. by m=
-% eans of a list of "thesis"). This would stimulate the discussion!
-
-% agree. to be included. it's already there, but not clear and polished enoug=
-% h
-
-% ---------------------------- REVIEW 2 --------------------------
-% OVERALL RATING: 2 (accept)
-
-% > The authors use the (somewhat formal) JML technology and introduce their =
-% TDF-methodology as a variant of more classic development cycles.
-
-% why "somewhat"?
-
-% > Content-wise it must be said that, given the amount of material in the te=
-% xt, the conclusions drawn and lessons learnt are a bit 'thin'. E.g. Does TD=
-% F really work? what is your experience?
-
-% partial agree. to be expanded, in the line of the previous comments
-
-% > It seems that the format of the printed area (text height and width) is n=
-% ot in accordance with the other conference contributions. When reformatted,=
-%  the already long paper will probably get longer. The (too large) wordiness=
-%  of the paper could perhaps be reduced somewhat to compensate.
-
-% uh
-% I used thir template... I'm "almost" sure
-% I'll double check this
-% maybe some following format options
-
-% > page 1, abstract: 'not always see eye to eye' seems an inappropriate way =
-% of stating 'do not meet face to face'.
-% > page 1, abstract end: highly-spices -> highly-spiced
-% > page 2, line 2: citation problem
-% > page 2, line 12: 'ther'?
-% > page 3, line 8 and 10: appliable in -> applicable to?
-% > page 3, line 8: yeilding -> yielding
-% > page 3, middle: .... the handmade test suites are automatically generated=
-%  .... ?
-% > page 3, the same section: the last two sentences seem crippled; split the=
-% m up!
-% > page 6, line -5: Identify -> Identifying
-% > page 9, line 1: confidant -> confident
-% > page 15, line 6: inoovative -> innovative
-% > page 15, line 12: in sink -> in sync
-% > page 17, last line: litt ref in disorder
-
-% ok
-
-% > Summing up the editorial part: Please use a spelling checker and perhaps =
-% some review by a native speaker
-% could help sharpening (and shortening) some sentences.
-
-% well... I've used a spell checker!
-% but we have to reread it once the modifications are in place
-
-% ---------------------------- REVIEW 3 --------------------------
-% OVERALL RATING: 2 (accept)
-
-% > they do not mention Liu's work on the SOFL foraml engineering method for =
-% software development that have been xtensively studies over the last 15 yea=
-% rs with the aim to solve the similar problem. In particular, the recent pap=
-% er on the application of SOFL to achieve agile process has provided a very =
-% different approach from the mentioned approaches
-% > Shaoying Liu, ``An Approach to Applying SOFL for Agile Process and Its Ap=
-% plication in Developing a Test Support Tool'', Innovations in Systems and S=
-% oftware Engineering, Springer London, 22 December, DOI 10.1007/s11334-0114-=
-% 3.
-
-% ok
-% to be read and analyzed
-
-% ---------------------------- REVIEW 4 --------------------------
-% OVERALL RATING: 1 (weak accept)
-
-% > I did find it disturbing that the paper never clearly distinguishes betwe=
-% en structural and functional testing . This is an
-% important distinction as it potentially evidences different types of proble=
-% ms.
-
-% I'll rewrite some parts to make the distinction clear
-
-% > Moreover, there is a distinct difference between automatically generated =
-% tests and manually produced tests as to how well they apply to one or the o=
-% ther.
-
-% the distinction is clear in the paper... I don't understand this
-
-% > The overall description of how AM and FM are merged is not that persuasiv=
-% e, as a result. In particular, Figure 2 does not
-% convey much nor profound value in its current incarnation and description.
-
-% mmhh
-% not much that can be done here... I think
-% how can we be more "persuasive"?
-
-% > It would help if the authors would more specifically address what type of=
-%  tests the JML specification helps generate. In the
-% case study, it is a structural coverage type test.
-
-% mmhh
-% I've explained this
-% it would take me pages to go into specific details... and I've cited
-% works where this is explained
-
-% > All that said, it is a bit unclear as to where the presented work is actu=
-% ally pushing the boundaries of the current state of
-% the field, other than in actually presenting the approach as an FM+AM hybri=
-% d.
-
-% we are actually talking about FM+AM hybrids... that IS the filed,
-% so... what's this question about?
-
-% > As a suggestion, then, it could be good if the verbosity and obtusity of =
-% the article is reduced
-
-% ok...
-
-% > and a more formal and incisive treatise is presented.
-
-% no idea how to do this
-% nor I'm sure I really want to do this
-
-% --=20
-% Vieri del Bianco
